---
title: "batch_correction_sample_reports"
author: "Amy Danson, Lea Wölbert"
date: '2022-12-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Summary report for data integration.

#### Load libraries, source code 

```{r load,  message = FALSE}

library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(cowplot, quietly = TRUE) 

```

```{r}

sce_07 <- snakemake@input[["sce_07"]]
sce_08 <- snakemake@input[["sce_08"]]
sce_09mnn <- snakemake@input[["sce_09mnn"]]

nr_hvgs <- snakemake@params[["nr_hvgs"]]
species_build <- snakemake@params[["species_build"]]
batches <- snakemake@params[["batches"]]
color_tables <- snakemake@params[["color_tables"]]
mnn_fast <- snakemake@params[["mnn_fast"]]
species_curr <- sce_07$Species_ID[1]
species_curr <- sce_07$Species_ID[1]

```

```{r}

sce_08 <- reduce_dims(sce_08, nr_hvgs = snakemake@params[["nr_hvgs"]])

if(mnn_fast){
  
  gene_var <- modelGeneVar(sce_09mnn, assay.type = "reconstructed")
  hvgs <- getTopHVGs(hvgs, n=gene_var)
  
  sce_09mnn <- runUMAP(sce_09mnn, dimred = "corrected",
                     external_neighbors=TRUE, subset_row = hvgs)

}else{
  
  gene_var <- modelGeneVar(sce_09mnn, assay.type = )
  hvgs <- getTopHVGs(hvgs, n=gene_var)
  
  sce_09mnn <- runPCA(sce_09mnn, ncomponents=25, subset_row = hvgs,
                      assay.type = )
  
  sce_09mnn <- runUMAP(sce_09mnn, dimred = "PCA",
                     external_neighbors=TRUE, subset_row = hvgs)
}
```


```{r}
source(file = "../source/colors.R")
source(file = "../source/plotting.R")
source(file = "../source/sce_functions.R")
```

```{r}
col_cts_all <- col_cts[names(col_cts) %in% sce_07$Identity_ref_all]
col_cts_fraction1 <- col_cts[names(col_cts) %in% sce_07$Identity_ref_fraction1]
col_cts_fraction2 <- col_cts[names(col_cts) %in% sce_07$Identity_ref_fraction2]
col_batch_exp_day <- col_alp[names(col_alp) %in% sce_07$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% sce_07$Batch_sequencing]
col_ann_age <- col_ann[names(col_ann) %in% sce_07$Age_ID]
col_ann_fraction <- col_ann[names(col_ann) %in% sce_07$Fraction_ID]
col_species <- col_spc[names(col_spc) %in% sce_07$Species_ID]
```

# Reference cell types 

Make one plotting wrapper for all objects.

```{r}

identity_plots <- function(sce, title, identifier){
  
  id_1 <- colData(sce)[identifier][1]
  
  plot_1 <- umap_base(sce, color_by = "Identity_ref_all")+ # own function
    ggtitle(paste0(id_1, title))+
    scale_color_manual("Identity1", values = col_cts_all)
  
  plot_1l <- umap_legend(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", values = col_cts_all)
  legend_1 <- get_legend(plot_1l)

    
  plot_2 <- umap_base(sce, color_by = "Identity_ref_fraction1")+
    ggtitle(paste0(id_1, title))+
    scale_color_manual("Identity2", values = col_cts_fraction1)
  
  plot_2l <- umap_legend(sce, color_by = "Identity_ref_fraction1")+
    scale_color_manual("Identity2", values = col_cts_fraction1)
  legend_2 <- get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "Identity_ref_fraction2")+
    ggtitle(paste0(id_1, title))+
    scale_color_manual("Identity3", values = col_cts_fraction2)
  
  plot_3l <- umap_legend(sce, color_by = "Identity_ref_fraction2")+
    scale_color_manual("Identity3", values = col_cts_fraction2)
  legend_3 <- get_legend(plot_3l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)
  
  return(list(PLOT_1, PLOT_2, PLOT_3))
}

```

## Species level

### After merging

```{r}

identity_species_plots_07 <- identity_plots(sce_07, 
                                            title = ", after merging",
                                            identifier = "Species_ID")

print(sce_07$Species_ID[1])
print(table(sce_07$Identity_ref_all))
print(table(sce_07$Identity_ref_fraction))

```

```{r fig.width=10, fig.height=5}
identity_species_plots_07
```

### After renormalization

```{r}

identity_species_plots_08 <- identity_plots(sce_08, 
                                            title = ", after renormalization",
                                            identifier = "Species_ID")

```

```{r fig.width=10, fig.height=5}
identity_species_plots_08
```

### After batch correction

```{r}

identity_species_plots_09 <- identity_plots(sce_09mnn, 
                                            title = ", after batch correction",
                                            identifier = "Species_ID")

```

```{r fig.width=10, fig.height=5}
identity_species_plots_09
```

## Contamination and mature cell types

Use only Baccin et al dataset because it's the only one containing both
stromal and hematopoetic cell types in a single dataset.

```{r}

# find contamination and mature cells in sce objects
find_contamination_mature <- function(sce, title){
  
  # CONTAMINATION
  sce_str <- sce[,sce$Fraction_ID == "str"]
  sce_hsc <- sce[,sce$Fraction_ID == "hsc"]

  sce_str <- find_contamination_ref_all(sce_str, input = "stromal") 
  sce_hsc <- find_contamination_ref_all(sce_hsc, input = "hsc") # own function

  if(grepl("TRUE", sce_str$right_fraction)){
    plot_4 <- umap_base(sce_str, color_by = "right_fraction")+
      scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                        "FALSE" = "orange"))
    plot_4l <- umap_legend(sce_str, color_by = "right_fraction")+
      scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                        "FALSE" = "orange"))
    legend_4 <- get_legend(plot_4l)

    PLOT_4 <- ggarrange(plot_4 + ggtitle(paste0(species_curr, 
                                                title, "Stromal")), legend_4)
  }

  if(grepl("TRUE", sce_hsc$right_fraction)){
    plot_5 <- umap_base(sce_hsc, color_by = "right_fraction")+
      scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                        "FALSE" = "orange"))
    plot_5l <- umap_legend(sce_hsc, color_by = "right_fraction")+
      scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                        "FALSE" = "orange"))
    legend_5 <- get_legend(plot_5l)
  
    PLOT_5 <- ggarrange(plot_5 + ggtitle(paste0(species_curr, 
                                                title, "HSPCs")), legend_5)
  }
  
  sce <- get_mature_cts_all(sce) # own function 

  plot_6 <- umap_base(sce, color_by = "mature_cells")+
  scale_color_manual("Mature cell types", values = c("FALSE" = "black", 
                                                    "TRUE" = "orange"))
  plot_6l <- umap_legend(sce, color_by = "mature_cells")+
  scale_color_manual("Mature cell types", values = c("FALSE" = "black", 
                                                    "TRUE" = "orange"))
  legend_6 <- get_legend(plot_6l)

  PLOT_6 <- ggarrange(plot_6 + ggtitle(paste0(species_curr, 
                                              title, "Mature HSCPs")), legend_6)
  
  return(list(PLOT_4, PLOT_5, PLOT_6))
}  

cont_plots_species_07 <- find_contamination(sce_07, 
                                            title = ", after merging")
cont_plots_species_08 <- find_contamination(sce_08, 
                                            title = ", after renormalization")
cont_plots_species_09 <- find_contamination(sce_09mnn, 
                                            title = ", after batch correction")

```

### After merging

```{r fig.width=10, fig.height=5}
cont_plots_species_07
```

### After renormalization

```{r fig.width=10, fig.height=5}
cont_plots_species_08
```

### After batch correction

```{r fig.width=10, fig.height=5}
cont_plots_species_09
```

# All QC metrics

```{r}

qc_plots <- function(sce, title){
  
  plot_1 <- umap_base(sce, color_by = "Object_ID")+ # own function
    ggtitle(paste0(species_curr, title))
  plot_1l <- umap_legend(sce, color_by = "Object_ID")
  legend_1 <- get_legend(plot_1l)
  
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", values = col_batch_seq)+
    ggtitle(paste0(species_curr, title))
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", values = col_batch_seq)
  legend_2 <- get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+ 
    scale_color_manual("Batch_experimental_day",values =  col_batch_exp_day)+
    ggtitle(paste0(species_curr, title))
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)
  legend_3 <- get_legend(plot_3l)
  
  plot_4 <- umap_base(sce, color_by = "Date_collected")+ 
    ggtitle(paste0(species_curr, title))
  plot_4l <- umap_legend(sce, color_by = "Date_collected")+
  legend_4 <- get_legend(plot_4l)
  
  # check everything else
  plot_5 <- umap_base(sce, color_by = "Species_ID")+ 
    scale_color_manual("Species", values = col_species)+
    ggtitle(paste0(species_curr, title))
  plot_5l <- umap_legend(sce, color_by = "Species_ID")+
    scale_color_manual("Species", values = col_species)+
  legend_5 <- get_legend(plot_5l)
  
  plot_6 <- umap_base(sce, color_by = "Age")+ 
    scale_color_manual("Age", values = col_ann_age)+
    ggtitle(paste0(species_curr, title))
  plot_6l <- umap_legend(sce, color_by = "Age")+
    scale_color_manual("Age", values = col_ann_age)+
  legend_6 <- get_legend(plot_6l) 
  
  plot_7 <- umap_base(sce, color_by = "Fraction")+ 
    scale_color_manual("Fraction", values = col_ann_fraction)+
    ggtitle(paste0(species_curr, title))
  plot_7l <- umap_legend(sce, color_by = "Fraction")+
    scale_color_manual("Fraction", values = col_ann_fraction)+
  legend_7 <- get_legend(plot_7l) 
 
  plot_8 <- umap_base(sce, color_by = "Antibody_combination")+ 
    ggtitle(paste0(species_curr, title))
  plot_8l <- umap_legend(sce, color_by = "Antibody_combination")+
  legend_8 <- get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "doublet_score")+ 
    ggtitle(paste0(species_curr, title))
  plot_9l <- umap_legend(sce, color_by = "doublet_score")+
  legend_9 <- get_legend(plot_9l)
  
  plot_10 <- umap_base(sce, color_by = "Keep_sample")+ 
    ggtitle(paste0(species_curr, title))
  plot_10l <- umap_legend(sce, color_by = "Keep_sample")+
  legend_10 <- get_legend(plot_10l)
  
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)
  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_5 <- ggarrange(plot_5, legend_5)
  PLOT_6 <- ggarrange(plot_6, legend_6)
  PLOT_7 <- ggarrange(plot_7, legend_7)
  PLOT_8 <- ggarrange(plot_8, legend_8)
  PLOT_9 <- ggarrange(plot_9, legend_9)
  PLOT_10 <- ggarrange(plot_10, legend_10)

  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10)
  return(plot_list)
  
}

qc_plots_species_07 <- qc_plots(sce_07, title = ", after merging")
qc_plots_species_08 <- qc_plots(sce_08, title = ", after renormalization")
qc_plots_species_09 <- qc_plots(sce_09mnn, title = ", after batch correction")

```

### After merging

```{r fig.width=10, fig.height=5}
qc_plots_species_07
```

### After renormalization

```{r fig.width=10, fig.height=5}
qc_plots_species_08
```

### After batch correction

```{r fig.width=10, fig.height=5}
qc_plots_species_09
```

## Facet grid UMAPs

To merge across ages and fractions using MNNcorrect, every age and fraction 
should have at least one overlapping BATCH.
Per definition, this is not the case for Object_ID as possible batch.

```{r}

facet_plots <- function(sce, title){

  plot_1 <- umap_base(sce, color_by = "Object_ID")+ 
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_1l <- umap_legend(sce, color_by = "Object_ID")
  legend_1 <- get_legend(plot_1l)
  
  col_batch_seq <- col_num[names(col_num) %in% sce$Batch_sequencing]
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)
  legend_2 <- get_legend(plot_2l)
  
  col_batch_exp_day <- col_alp[names(col_alp) %in% sce$Batch_exp_day]
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_3 <- get_legend(plot_3l)

  # check age
  plot_4 <- umap_base(sce, color_by = "Object_ID")+ 
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  plot_4l <- umap_legend(sce, color_by = "Object_ID")
  legend_4 <- get_legend(plot_4l)
  
  plot_5 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  plot_5l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)
  legend_5 <- get_legend(plot_5l)
  
  plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_6 <- get_legend(plot_6l)
  
  plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_6 <- get_legend(plot_6l)
  
  # Identities
  col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
  col_cts_fraction <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction]
 
  plot_7 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_7l <- umap_legend(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)
  legend_idrefall <- get_legend(plot_7l)
  
  plot_8 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_8l <- umap_legend(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)
  legend_idreffrac <- get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  
  plot_10 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Age))
  
  plot_11 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_12 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_13 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Batch_exp_day))
  
  plot_14 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species_curr, title))+
    facet_grid(cols = vars(sce$Batch_exp_day))
 
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)
  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_5 <- ggarrange(plot_5, legend_5)
  PLOT_6 <- ggarrange(plot_6, legend_6)
  PLOT_7 <- ggarrange(plot_7, legend_idrefall)
  PLOT_8 <- ggarrange(plot_8, legend_idreffrac)
  PLOT_9 <- ggarrange(plot_9, legend_idrefall)
  PLOT_10 <- ggarrange(plot_10, legend_idreffrac)
  PLOT_11 <- ggarrange(plot_11, legend_idrefall)
  PLOT_12 <- ggarrange(plot_12, legend_idreffrac)
  PLOT_13 <- ggarrange(plot_13, legend_idrefall)
  PLOT_14 <- ggarrange(plot_14, legend_idreffrac)
  
  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10, PLOT_11, PLOT_12, PLOT_13, PLOT_14)
  
  return(plot_list)
}

facet_plots_species_07 <- facet_plots(sce_07, title = ", after merging",)
facet_plots_species_08 <- facet_plots(sce_08, title = ", after renormalization")
facet_plots_species_09 <- facet_plots(sce_09mnn,
                                      title = ", after batch correction",)

```

### After merging

```{r fig.width=10, fig.height=5}
facet_plots_species_07
```

### After renormalization

```{r fig.width=10, fig.height=5}
facet_plots_species_08
```

### After batch correction

```{r fig.width=10, fig.height=5}
facet_plots_species_09
```


# Cell type proportions

Using only one object is enough for this

```{r}
sce <- sce_09mnn
```

## Per Sample

```{r}

ggdf <- data.frame(colData(sce))

plot1 <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_all,
                          fill=Identity_ref_all, color = Identity_ref_all))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot_all_leg <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_all,
                                 fill=Identity_ref_all))+
  geom_col()+
  scale_fill_manual("Identity1", values = col_cts_all)
legend_all <- get_legend(plot_all_leg)

plot2 <- ggplot(ggdf, aes(x = Object_ID, fill=Identity_ref_all, 
                          color = Identity_ref_all))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot3 <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_fraction1,
                          fill=Identity_ref_fraction1,
                          color = Identity_ref_fraction1))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot_frc1_leg <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_fraction1,
                                  fill=Identity_ref_fraction1))+
  geom_col()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)
legend_frc1 <- get_legend(plot_frc1_leg)
                    
plot4 <- ggplot(ggdf, aes(x = Object_ID, fill=Identity_ref_fraction1, 
                          color = Identity_ref_fraction1))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot5 <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_fraction2,
                          fill=Identity_ref_fraction2,
                          color = Identity_ref_fraction2))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot_frc2_leg <- ggplot(ggdf, aes(x = Object_ID, y=Identity_ref_fraction2,
                                  fill=Identity_ref_fraction2))+
  geom_col()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)
legend_frc2 <- get_legend(plot_frc2_leg)
   
plot6 <- ggplot(ggdf, aes(x = Object_ID, fill=Identity_ref_fraction2, 
                          color = Identity_ref_fraction2))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

PLOT1 <- ggarrange(plot1, legend_all)
PLOT2 <- ggarrange(plot2, legend_all)
PLOT3 <- ggarrange(plot3, legend_frc1, ncol = 1)
PLOT4 <- ggarrange(plot4, legend_frc1, ncol = 1)
PLOT5 <- ggarrange(plot5, legend_frc2, ncol = 1)
PLOT6 <- ggarrange(plot6, legend_frc2, ncol = 1)

PLOT1
PLOT2

```

```{r fig.width=10, fig.height = 11}

PLOT3
PLOT4
PLOT5
PLOT6

```

## Per Batch

```{r}

plot1 <- ggplot(ggdf, aes(x = Batch_exp_day, y=Identity_ref_all,
                          fill=Identity_ref_all, color = Identity_ref_all))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot2 <- ggplot(ggdf, aes(x = Batch_exp_day, fill=Identity_ref_all, 
                          color = Identity_ref_all))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot3 <- ggplot(ggdf, aes(x = Batch_exp_day, y=Identity_ref_fraction1,
                          fill=Identity_ref_fraction1,
                          color = Identity_ref_fraction1))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)
                    
plot4 <- ggplot(ggdf, aes(x = Batch_exp_day, fill=Identity_ref_fraction1, 
                          color = Identity_ref_fraction1))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot5 <- ggplot(ggdf, aes(x = Batch_exp_day, y=Identity_ref_fraction2,
                          fill=Identity_ref_fraction2,
                          color = Identity_ref_fraction2))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot6 <- ggplot(ggdf, aes(x = Batch_exp_day, fill=Identity_ref_fraction2, 
                          color = Identity_ref_fraction2))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

PLOT1 <- ggarrange(plot1, legend_all)
PLOT2 <- ggarrange(plot2, legend_all)
PLOT3 <- ggarrange(plot3, legend_frc1, ncol = 1)
PLOT4 <- ggarrange(plot4, legend_frc1, ncol = 1)
PLOT5 <- ggarrange(plot5, legend_frc2, ncol = 1)
PLOT6 <- ggarrange(plot6, legend_frc2, ncol = 1)

PLOT1
PLOT2
```

```{r fig.width=10, fig.height=11}

PLOT3
PLOT4
PLOT5
PLOT6

```

```{r}

ggdf_str <- data.frame(colData(sce)[sce$Fraction_ID == "str",])

plot1 <- ggplot(ggdf_str, aes(x = Batch_exp_day, y=Identity_ref_all,
                          fill=Identity_ref_all, color = Identity_ref_all))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot2 <- ggplot(ggdf_str, aes(x = Batch_exp_day, fill=Identity_ref_all,
                              color = Identity_ref_all))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot3 <- ggplot(ggdf_str, aes(x = Batch_exp_day, y=Identity_ref_fraction1,
                          fill=Identity_ref_fraction1,
                          color = Identity_ref_fraction1))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)
                    
plot4 <- ggplot(ggdf_str, aes(x = Batch_exp_day, fill=Identity_ref_fraction1, 
                          color = Identity_ref_fraction1))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot5 <- ggplot(ggdf_str, aes(x = Batch_exp_day, y=Identity_ref_fraction2,
                          fill=Identity_ref_fraction2,
                          color = Identity_ref_fraction2))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot6 <- ggplot(ggdf_str, aes(x = Batch_exp_day, fill=Identity_ref_fraction2, 
                          color = Identity_ref_fraction2))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

PLOT1 <- ggarrange(plot1, legend_all)
PLOT2 <- ggarrange(plot2, legend_all)
PLOT3 <- ggarrange(plot3, legend_frc1, ncol = 1)
PLOT4 <- ggarrange(plot4, legend_frc1, ncol = 1)
PLOT5 <- ggarrange(plot5, legend_frc2, ncol = 1)
PLOT6 <- ggarrange(plot6, legend_frc2, ncol = 1)

PLOT1
PLOT2
```

```{r fig.width=10, fig.height=11}

PLOT3
PLOT4
PLOT5
PLOT6

```

```{r}

ggdf_hsc <- data.frame(colData(sce)[sce$Fraction_ID == "hsc",])

plot1 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, y=Identity_ref_all,
                          fill=Identity_ref_all, color = Identity_ref_all))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot2 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, fill=Identity_ref_all,
                              color = Identity_ref_all))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot3 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, y=Identity_ref_fraction1,
                          fill=Identity_ref_fraction1,
                          color = Identity_ref_fraction1))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)
                    
plot4 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, fill=Identity_ref_fraction1, 
                          color = Identity_ref_fraction1))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot5 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, y=Identity_ref_fraction2,
                          fill=Identity_ref_fraction2,
                          color = Identity_ref_fraction2))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot6 <- ggplot(ggdf_hsc, aes(x = Batch_exp_day, fill=Identity_ref_fraction2, 
                          color = Identity_ref_fraction2))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

PLOT1 <- ggarrange(plot1, legend_all)
PLOT2 <- ggarrange(plot2, legend_all)
PLOT3 <- ggarrange(plot3, legend_frc1, ncol = 1)
PLOT4 <- ggarrange(plot4, legend_frc1, ncol = 1)
PLOT5 <- ggarrange(plot5, legend_frc2, ncol = 1)
PLOT6 <- ggarrange(plot6, legend_frc2, ncol = 1)

PLOT1
PLOT2
```

```{r fig.width=10, fig.height=11}

PLOT3
PLOT4
PLOT5
PLOT6

```

## Per species

```{r}

plot1 <- ggplot(ggdf, aes(x = Fraction_ID, y=Identity_ref_all,
                          fill=Identity_ref_all, color = Identity_ref_all))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity1", values = col_cts_all)+
  scale_color_manual(values = col_cts_all)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot3 <- ggplot(ggdf, aes(x = Fraction_ID, y=Identity_ref_fraction1,
                          fill=Identity_ref_fraction1,
                          color = Identity_ref_fraction1))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity2", values = col_cts_fraction1)+
  scale_color_manual(values = col_cts_fraction1)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

plot5 <- ggplot(ggdf, aes(x = Fraction_ID, y=Identity_ref_fraction2,
                          fill=Identity_ref_fraction2,
                          color = Identity_ref_fraction2))+
  geom_col()+
  theme_classic()+
  scale_fill_manual("Identity3", values = col_cts_fraction2)+
  scale_color_manual(values = col_cts_fraction2)+
  theme(legend.position = "none", axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(species_curr)

PLOT1 <- ggarrange(plot1, legend_all)
PLOT3 <- ggarrange(plot3, legend_frc1, ncol = 1)
PLOT5 <- ggarrange(plot5, legend_frc2, ncol = 1)

PLOT1

```

```{r fig.width=10, fig.height=11}

PLOT3
PLOT5

```

```{r}
sessionInfo()
```
