---
title: "bulk_quality_report"
author: "Lea Wölbert"
date: '2023-11-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

This data is NOT batch corrected.
DESeq objects are batch corrected by removing hidden sources of variations using
the sva package.
It is no surprising that aggregated samples would by separated by batch 
(Batch_experimental_day).
It was not possible to include Batch_experimental_day in the design because
in the case of BL6, Age_ID and Batch_experimental_day cannot be separated 
(all young BL6 are in one batch, all old BL6 in another batch).

#### Load objects

```{r seed,  message = FALSE}
set.seed(37)
```

```{r load,  message = FALSE}
library(pcaExplorer)
library(pheatmap)
library(RColorBrewer)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["functions"]])
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}

rlog_list = readRDS(snakemake@input[["rlog"]])
rlog_list <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/main_analysis/sce_objects/03_sce_analysis/01_DESeq2_crossage/04_dres/res_mcar_hsc_celltype")

print(names(rld))
```

```{r colors, meassage = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])
```

# QC

Distribution of expression values, should be uniform.

```{r}
plot1 <- distro_expr(rld, plot_type = "boxplot")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90))+
  theme(legend.position = "none")

plot1
```

```{r}
rld_pca <- prcomp(t(assay(rld)))

rld_cor <- correlatePCs(rld_pca, colData(rld))

plotPCcorrs(rld_cor)

```

```{r}

ggdf <- as.data.frame(rld_pca$x)
ggdf$Antibody_combination <- colData(rld)$Antibody_combination
ggdf$age <- colData(rld)$age
ggdf$condition <- colData(rld)$condition
ggdf$batch <- colData(rld)$batch
ggdf$Batch_sequencing <- colData(rld)$Batch_sequencing
ggdf$sample <- colData(rld)$sample
ggdf$ncells <- colData(rld)$ncells

head(ggdf)
#knitr::knit_exit()
```

# PCAs

```{r}
plot1 <- ggplot(ggdf, aes(x = PC1, y = PC2, color = condition))+
  geom_point()+
  theme_classic()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())+
  scale_color_manual("Species", values = col_spc)
plot1

plot2 <- ggplot(ggdf, aes(x = PC1, y = PC2, color = age))+
  geom_point()+
  theme_classic()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())+
  scale_color_manual("Age", values = col_age)
plot2

plot3 <- ggplot(ggdf, aes(x = PC1, y = PC2, color = batch))+
  geom_point()+
  theme_classic()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
plot3

plot4 <- ggplot(ggdf, aes(x = PC1, y = PC2, color = Antibody_combination))+
  geom_point()+
  theme_classic()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
plot4

plot5 <- ggplot(ggdf, aes(x = PC1, y = PC2, color = ncells))+
  geom_point()+
  theme_classic()+
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())
plot5

knitr::knit_exit()

```

```{r}
rld_pca <- prcomp(t(assay(rld)))
plot3 <- pcascree(rld_pca, type="pev")
print(plot3)
```

```{r}
hi_loadings(rld_pca, topN = 20, whichpc = 1)
```

```{r}
hi_loadings(rld_pca, topN = 20, whichpc = 2)
```

```{r}
sampleDists <- dist(t(assay(rld)))
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- rld$condition
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette(rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
```

```{r}
sessionInfo()
```