---
title: "ref_anno_sample_report"
author: "Lea Wölbert"
date: '2022-11-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on annotation with three reference datasets, annotated using
scmapcell and scmapcluster at sample level (Object_ID).

#### Load libraries, source code 

```{r load,  message = FALSE}

library(SingleCellExperiment, quietly = TRUE) 
library(DropletUtils, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(pheatmap, quietly = TRUE) 
set.seed(37)
```

```{r load_object, message = FALSE}

print(snakemake@input[[1]])
sce <- readRDS(file = snakemake@input[[1]])
color_tables <- snakemake@params[["color_tables"]]

# set name for plots
name_curr <- colData(sce)$Object_ID[1] # choose name to display in plots

```


```{r load_code, message = FALSE}

source(file = "../../source/colors.R")
source(file = "../../source/plotting.R")
source(file = "../../source/sce_functions.R")

sce <- factor_reference_cts(sce)# own function

col_cts_baccin_cell <- col_cts[match(levels(sce$baccin_celltype_scmapcell), 
                                     names(col_cts))]
col_cts_baccin_clst <- col_cts[match(levels(sce$baccin_celltype_scmapclust),
                                     names(col_cts))]
col_cts_dahlin_cell <- col_cts[match(levels(sce$dahlin_celltype_scmapcell), 
                                     names(col_cts))]
col_cts_dahlin_clst <- col_cts[match(levels(sce$dahlin_celltype_scmapclust), 
                                     names(col_cts))]
col_cts_dolgalev_cell <- col_cts[match(levels(sce$dolgalev_celltype_scmapcell),
                                       names(col_cts))]
col_cts_dolgalev_clst <- col_cts[match(levels(sce$dolgalev_celltype_scmapclust),
                                       names(col_cts))]

print(levels(sce$dolgalev_celltype_scmapclust))
print(levels(sce$baccin_celltype_scmapclust))
print(levels(sce$dahlin_celltype_scmapclust))

print(col_cts_baccin_cell)
print(col_cts_dahlin_clst)
print(col_cts_dolgalev_clst)
```

## Cell types

### All Identities

#### Baccin

```{r fig.width=8}

plot_1 <- umap_base(sce, color_by = "baccin_celltype_scmapcell")+ # own function
  ggtitle(name_curr)+
  scale_color_manual("baccin_celltype_scmapcell", values = col_cts_baccin_cell)
plot_1l <- umap_legend(sce, color_by = "baccin_celltype_scmapcell")+
  scale_color_manual("baccin_celltype_scmapcell", values = col_cts_baccin_cell)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+ 
  ggtitle(name_curr)+
  scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin_clst)
plot_2l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust")+
  scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin_clst)
legend_2 <- get_legend(plot_2l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)

PLOT_1
PLOT_2
```

```{r fig.width=8}

plot_3 <- umap_base(sce, color_by = "baccin_celltype_scmapcell_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_3l <- umap_legend(sce, color_by = "baccin_celltype_scmapcell_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "baccin_celltype_scmapcell_sim")
legend_3 <- get_legend(plot_3l)  

plot_4 <- umap_base(sce, color_by = "baccin_celltype_scmapclust_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_4l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "baccin_celltype_scmapclust_sim")
legend_4 <- get_legend(plot_4l)  

PLOT_3 <- ggarrange(plot_3, legend_3)
PLOT_4 <- ggarrange(plot_4, legend_4)

PLOT_3
PLOT_4
```

#### Dahlin

```{r fig.width=8}

plot_1 <- umap_base(sce, color_by = "dahlin_celltype_scmapcell")+ 
  ggtitle(name_curr)+
  scale_color_manual("dahlin_celltype_scmapcell", values = col_cts_dahlin_cell)
plot_1l <- umap_legend(sce, color_by = "dahlin_celltype_scmapcell")+
  scale_color_manual("dahlin_celltype_scmapcell", values = col_cts_dahlin_cell)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+ 
  ggtitle(name_curr)+
  scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin_clst)
plot_2l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust")+
  scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin_clst)
legend_2 <- get_legend(plot_2l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)

PLOT_1
PLOT_2
```

```{r fig.width=8}

plot_3 <- umap_base(sce, color_by = "dahlin_celltype_scmapcell_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_3l <- umap_legend(sce, color_by = "dahlin_celltype_scmapcell_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "dahlin_celltype_scmapcell_sim")
legend_3 <- get_legend(plot_3l)  

plot_4 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_4l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "dahlin_celltype_scmapclust_sim")
legend_4 <- get_legend(plot_4l)  

PLOT_3 <- ggarrange(plot_3, legend_3)
PLOT_4 <- ggarrange(plot_4, legend_4)

PLOT_3
PLOT_4
```

#### Dolgalev

```{r fig.width=8}

plot_1 <- umap_base(sce, color_by = "dolgalev_celltype_scmapcell")+ 
  ggtitle(name_curr)+
  scale_color_manual("dolgalev_celltype_scmapcell",
                     values = col_cts_dolgalev_cell)
plot_1l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapcell")+
  scale_color_manual("dolgalev_celltype_scmapcell", 
                     values = col_cts_dolgalev_cell)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+ 
  ggtitle(name_curr)+
  scale_color_manual("dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev_clst)
plot_2l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
  scale_color_manual("dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev_clst)
legend_2 <- get_legend(plot_2l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)

PLOT_1
PLOT_2
```

```{r fig.width=8}

plot_3 <- umap_base(sce, color_by = "dolgalev_celltype_scmapcell_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_3l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapcell_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "dolgalev_celltype_scmapcell_sim")
legend_3 <- get_legend(plot_3l)  

plot_4 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust_sim")+
  ggtitle(name_curr)+
  scale_colour_gradient(low = "black", high = "gold")
plot_4l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust_sim")+
  scale_colour_gradient(low = "black", high = "gold",
                        name = "dolgalev_celltype_scmapclust_sim")
legend_4 <- get_legend(plot_4l)  

PLOT_3 <- ggarrange(plot_3, legend_3)
PLOT_4 <- ggarrange(plot_4, legend_4)

PLOT_3
PLOT_4
```

## Contamination vs. correct cell types vs mature cell types

### Contamination

Use only Baccin et al dataset because it's the only one containing both
stromal and hematopoetic cell types in a single dataset.

```{r}

# find contamination in sce objects (scmapcluster)

if(grepl("str", name_curr)){
  sce <- find_contamination_ref_all(sce, input = "stromal") # own function
}else if(grepl("hsc", name_curr)){
  sce <- find_contamination_ref_all(sce, input = "hsc") 
}

```

```{r fig.width=8}

if(length(grepl("TRUE", sce$right_fraction)) >0){
  plot_4 <- umap_base(sce, color_by = "right_fraction")+
    ggtitle(name_curr)+
    scale_color_manual("Correct fraction", values = c("TRUE" = "grey80", 
                                                      "FALSE" = "orange"))
  plot_4l <- umap_legend(sce, color_by = "right_fraction")+
    scale_color_manual("Correct fraction", values = c("TRUE" = "grey80", 
                                                      "FALSE" = "orange"))
  legend_4 <- get_legend(plot_4l)

  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_4
}
```

```{r fig.width=8}

qcdf <- perCellQCMetrics(sce)
sce$sum <- qcdf$sum
sce$detected <- qcdf$detected

plot_5 <- umap_base(sce, color_by = "detected")+
  ggtitle(name_curr)+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))+
  ggtitle("With contaminants")
plot_5l <- umap_legend(sce, color_by = "detected")+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))
legend_5 <- get_legend(plot_5l)

PLOT_5 <- ggarrange(plot_5, legend_5)
PLOT_5

plot_6 <- umap_base(sce, color_by = "sum")+
  ggtitle(name_curr)+
  scale_color_gradientn("Library size", 
                         limits = c(0, 100000),
                         colors = c("black", "darkorange4", 
                                    "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))+
  ggtitle("With contaminants")
plot_6l <- umap_legend(sce, color_by = "sum")+
  scale_color_gradientn("Library size", 
                         limits = c(0, 1000000),
                          colors = c("black", "darkorange4", 
                                     "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))
legend_6 <- get_legend(plot_6l)

PLOT_6 <- ggarrange(plot_6, legend_6)
PLOT_6
```

```{r fig.width=8}

if(length(grepl("TRUE", sce$right_fraction))>0){
  sce_wo <- sce[,sce$right_fraction == TRUE]

  qcdf_wo <- perCellQCMetrics(sce_wo)
  sce_wo$sum <- qcdf_wo$sum
  sce_wo$detected <- qcdf_wo$detected

  plot_7 <- umap_base(sce_wo, color_by = "detected")+
    ggtitle(name_curr)+
    scale_color_gradientn("Number of detected genes", 
                           limits = c(0, 10000),
                           colors = c("black", "darkorange3", 
                                      "orange", "lightgoldenrod"))+
  ggtitle("Without contaminants")
  plot_7l <- umap_legend(sce_wo, color_by = "detected")+
    scale_color_gradientn("Number of detected genes", 
                           limits = c(0, 10000),
                           colors = c("black", "darkorange3",
                                      "orange", "lightgoldenrod"))
  legend_7 <- get_legend(plot_7l)

  PLOT_7 <- ggarrange(plot_7, legend_7)
  PLOT_7
  
  plot_8 <- umap_base(sce_wo, color_by = "sum")+
    ggtitle(name_curr)+
    scale_color_gradientn("Library size", 
                           limits = c(0, 100000),
                           colors = c("black", "darkorange4",
                                      "darkorange3", "darkorange1",
                                      "orange", "orange", "gold1", "gold",
                                      "lightgoldenrod", "lightgoldenrod", 
                                      "lightgoldenrod", "lightgoldenrod", 
                                      "lightgoldenrod", "lightgoldenrod"))+
    ggtitle("Without contaminants")
  plot_8l <- umap_legend(sce_wo, color_by = "sum")+
    scale_color_gradientn("Library size", 
                           limits = c(0, 1000000),
                            colors = c("black", "darkorange4",
                                       "darkorange3", "darkorange1",
                                       "orange", "orange", "gold1", "gold",
                                       "lightgoldenrod", "lightgoldenrod", 
                                       "lightgoldenrod", "lightgoldenrod", 
                                       "lightgoldenrod", "lightgoldenrod"))
  legend_8 <- get_legend(plot_8l)

  PLOT_8 <- ggarrange(plot_8, legend_8)
  PLOT_8
}
```

### Mature cells

```{r}
sce <- get_mature_cts_all(sce) # own function 
```

```{r fig.width=8}

plot_4 <- umap_base(sce, color_by = "mature_cells")+
  ggtitle(name_curr)+
  scale_color_manual("Mature cells", values = c("FALSE" = "grey80", 
                                                    "TRUE" = "orange"))
plot_4l <- umap_legend(sce, color_by = "mature_cells")+
  ggtitle(name_curr)+
  scale_color_manual("Mature cells", values = c("FALSE" = "grey80", 
                                                    "TRUE" = "orange"))
legend_4 <- get_legend(plot_4l)

PLOT_4 <- ggarrange(plot_4, legend_4)
PLOT_4
```

```{r}

qcdf <- perCellQCMetrics(sce)
sce$sum <- qcdf$sum
sce$detected <- qcdf$detected

plot_5 <- umap_base(sce, color_by = "detected")+
  ggtitle(name_curr)+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))+
  ggtitle("With mature cell types")
plot_5l <- umap_legend(sce, color_by = "detected")+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))
legend_5 <- get_legend(plot_5l)

PLOT_5 <- ggarrange(plot_5, legend_5)
PLOT_5

plot_6 <- umap_base(sce, color_by = "sum")+
  ggtitle(name_curr)+
  scale_color_gradientn("Library size", 
                         limits = c(0, 100000),
                         colors = c("black", "darkorange4", 
                                    "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))+
  ggtitle("With mature cell types")
plot_6l <- umap_legend(sce, color_by = "sum")+
  scale_color_gradientn("Library size", 
                         limits = c(0, 1000000),
                          colors = c("black", "darkorange4", "darkorange3",
                                     "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))
legend_6 <- get_legend(plot_6l)

PLOT_6 <- ggarrange(plot_6, legend_6)
PLOT_6
```

```{r}

sce_wo <- sce[,sce$mature_cells == FALSE]

qcdf_wo <- perCellQCMetrics(sce_wo)
sce_wo$sum <- qcdf_wo$sum
sce_wo$detected <- qcdf_wo$detected

plot_7 <- umap_base(sce_wo, color_by = "detected")+
  ggtitle(name_curr)+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))+
  ggtitle("Without contaminants")
plot_7l <- umap_legend(sce_wo, color_by = "detected")+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", 
                                    "orange", "lightgoldenrod"))
legend_7 <- get_legend(plot_7l)

PLOT_7 <- ggarrange(plot_7, legend_7)
PLOT_7

plot_8 <- umap_base(sce_wo, color_by = "sum")+
  ggtitle(name_curr)+
  scale_color_gradientn("Library size", 
                         limits = c(0, 100000),
                         colors = c("black", "darkorange4", 
                                    "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))+
  ggtitle("Without contaminants")
plot_8l <- umap_legend(sce_wo, color_by = "sum")+
  scale_color_gradientn("Library size", 
                         limits = c(0, 1000000),
                          colors = c("black", "darkorange4", 
                                     "darkorange3", "darkorange1",
                                     "orange", "orange", "gold1", "gold",
                                     "lightgoldenrod", "lightgoldenrod", 
                                     "lightgoldenrod", "lightgoldenrod", 
                                     "lightgoldenrod", "lightgoldenrod"))
legend_8 <- get_legend(plot_8l)

PLOT_8 <- ggarrange(plot_8, legend_8)
PLOT_8
```

### Cell type overlap scmapcell vs scmapcluster

The percentage of cells per Haas cell type that is assigned another
identity with another dataset.

#### In Baccin

```{r}

mat_comp <- as.matrix(table(unfactor(sce$baccin_celltype_scmapcell),
                            unfactor(sce$baccin_celltype_scmapclust)))
if(! 1 %in% dim(mat_comp)){

  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

  rownames(mat_comp) <- paste0(rownames(mat_comp), "_cell")
  colnames(mat_comp) <- paste0(colnames(mat_comp), "_clust")

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }

  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " Baccin overlap"))

  pheatmap(mat_comp_temp2, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " Baccin overlap"))
}else{
  print("not enough dimensions")
}
```

#### In Dahlin

```{r}

mat_comp <- as.matrix(table(unfactor(sce$dahlin_celltype_scmapcell),
                            unfactor(sce$dahlin_celltype_scmapclust)))
if(! 1 %in% dim(mat_comp)){
  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]
  
  rownames(mat_comp) <- paste0(rownames(mat_comp), "_cell")
  colnames(mat_comp) <- paste0(colnames(mat_comp), "_clust")

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }

  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " dahlin overlap"))

  pheatmap(mat_comp_temp2, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " dahlin overlap"))

}else{
  print("not enough dimensions")
}
```

#### In Dolgalev

```{r}

mat_comp <- as.matrix(table(unfactor(sce$dolgalev_celltype_scmapcell),
                            unfactor(sce$dolgalev_celltype_scmapclust)))
if(! 1 %in% dim(mat_comp)){
  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]
  
  rownames(mat_comp) <- paste0(rownames(mat_comp), "_cell")
  colnames(mat_comp) <- paste0(colnames(mat_comp), "_clust")

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }

  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " dolgalev overlap"))
  
  pheatmap(mat_comp_temp2, cluster_cols = FALSE, cluster_rows = FALSE, 
           main = paste0(name_curr, " dolgalev overlap"))
}else{
  print("not enough dimensions")
}
```

```{r}
sessionInfo()
```