---
title: "man_annotation_mmus_HSC"
author: "Lea Wölbert, Adrien Jolly"
date: '2022-11-29'
output: html_document
---

# MMUS_HSC ANNOTATION

Manually annotate cell types by using NMF, reference data sets and various
other data.
General plots are found in reports (.../data/sce_objects/reports/03_annotation).
This cannot be automated and must be done manually for each seperate data set.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r load,  message = FALSE}

library(SingleCellExperiment, quietly = TRUE) 
library(DropletUtils, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
#library(scater, quietly = TRUE) 
#library(scuttle, quietly = TRUE) 
#library(scran, quietly = TRUE) 
library(pheatmap)
library("config")

```

```{r}
sce <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/sce_objects/08_anmf/sce_mmus-08")
sce <- sce[,sce$Fraction_ID == "hsc"]
```

```{r}

config <- config::get(file = "../../config/config.yaml")
color_tables <- config$color_tables

source(file = "../../source/colors.R")
source(file = "../../source/plotting.R")
source(file = "../../source/sce_functions.R")

col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
col_cts_fraction1 <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction1]
col_cts_fraction2 <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction2]
col_batch_exp_day <- col_alp[names(col_alp) %in% sce$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% sce$Batch_sequencing]
col_ann_age <- col_ann[names(col_ann) %in% sce$Age_ID]
col_ann_fraction <- col_ann[names(col_ann) %in% sce$Fraction_ID]

```

# hsc_program 1

```{r}

print(table(sce$hsc_program_1 > 0))
print(table(sce$hsc_program_1 > 1))
print(table(sce$hsc_program_1 > 2))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Batch_exp_day, 
                                     color = sce$Batch_exp_day))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot( outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Experimental day", values = col_batch_exp_day)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Batch_sequencing, 
                                     color = sce$Batch_sequencing))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$Age_ID, 
                                     color = sce$Age_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Age", values = col_ann_age)

```

Looks like possible TC/NK cell or mature cell program

```{r}

program_1_genes <- data.frame(
  genes = rownames(metadata(sce)$Genescores_NMF_hsc),
  scores = metadata(sce)$Genescores_NMF_hsc$hsc_program_1)

program_1_genes <- program_1_genes[order(program_1_genes$scores, 
                                         decreasing = TRUE),]
program_1_genes[1:50,]

```

Trbc1, Tcf7, Tcrg.C1, Tox ... associated with TCs.
Bcl11b associated with B cell lymphoma.
Gata3 associated with erythropoesis.

Mostly TC associated but I guess generally differentiation related.
Would say contamination/mature cell related.
Could also be apoptosis related (Bcl2, Pdcd1)

```{r}

sce <- get_mature_cts_all(sce) # own function 

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$mature_cells, 
                                     color = sce$mature_cells))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Mature cells", values = c("FALSE" = "black",
                                                "TRUE" ="orange"))

sce <- find_contamination_ref_all(sce, input = "hsc") # own function

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_1, 
                                     x = sce$right_fraction, 
                                     color = sce$right_fraction))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Mature cells", values = c("TRUE" = "black",
                                                "FALSE" ="orange"))

```

Is indeed enriched in mature cells, but still over half of mature cells
have a usage of 0.

Check if its low quality cells associated with apoptosis.

```{r}

sce$is_p1_quick <- rep("FALSE", nrow(colData(sce)))
sce$is_p1_quick[sce$hsc_program_1 >1] <- "TRUE"

qcdf <- perCellQCMetrics(sce)

sce$sum <- qcdf$sum
sce$detected <- qcdf$detected

ggplot(data.frame(colData(sce)), aes(y = sce$sum, 
                                     x = sce$is_p1_quick, 
                                     color = sce$is_p1_quick))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Mature cells", values = c("TRUE" = "black",
                                                "FALSE" ="orange"))

ggplot(data.frame(colData(sce)), aes(y = sce$detected, 
                                     x = sce$is_p1_quick, 
                                     color = sce$is_p1_quick))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Mature cells", values = c("TRUE" = "black",
                                                "FALSE" ="orange"))

ggplot(data.frame(colData(sce)), aes(y = sce$detected, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  facet_grid(cols = vars(sce$is_p1_quick))


```

Quality is not significantly different I would say, or probably even a little 
better in cells with program_1 usage >1.

# Program 2

```{r}

print(table(sce$hsc_program_2 > 0))
print(table(sce$hsc_program_2 > 1))
print(table(sce$hsc_program_2 > 2))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)

```

These seems to be an "immature" signature, in contrast to program 1.
The "low-prog2" cells that still count as immature may be mostly Ery progs.

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$mature_cells, 
                                     color = sce$mature_cells))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Mature cells", values = c("FALSE" = "black",
                                                "TRUE" ="orange"))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2,
                                     x = sce$hsc_program_1))+
  geom_bin2d()+
  theme_classic()

gg_df <- data.frame(colData(sce)) 
gg_df <- gg_df[-which(gg_df$hsc_program_1 == 0 | gg_df$hsc_program_2 == 0),]

ggplot(gg_df, aes(y = hsc_program_2, x = hsc_program_1))+
  geom_bin2d()+
  theme_classic()

```

It doesn't appear like Program 1 and 2 are necessarily strictly exclusive
of each other. 
There are high-program 1 cells that are high, and some that are low in 
program 2 usage.

```{r}

program_2_genes <- data.frame(
  genes = rownames(metadata(sce)$Genescores_NMF_hsc),
  scores = metadata(sce)$Genescores_NMF_hsc$hsc_program_2)

program_2_genes <- program_2_genes[order(program_2_genes$scores, 
                                         decreasing = TRUE),]
program_2_genes[1:50,]

```

Ribosomal and mitochondrial gene expression suggest that it's a proliferation
or activity signature.
Makes sense that it's not found in mature cells.
But these are also the genes where batch effect would be found.

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Batch_exp_day, 
                                     color = sce$Batch_exp_day))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot( outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Experimental day", values = col_batch_exp_day)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Batch_sequencing, 
                                     color = sce$Batch_sequencing))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Age_ID, 
                                     color = sce$Age_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Age", values = col_ann_age)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Object_ID, 
                                     color = sce$Object_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")

```

this program is clearly enriched in old cells, but of the old objects,
it is especially enriched in two objects.

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)+
  facet_grid(cols = vars(sce$Age_ID))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  facet_grid(cols = vars(sce$Age_ID))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  facet_grid(cols = vars(sce$Age_ID))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Object_ID, 
                                     color = sce$Object_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  facet_grid(cols = vars(sce$Batch_exp_day))

```

```{r fig.height = 10}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  facet_grid(rows = vars(sce$Batch_exp_day))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_all)+
  facet_grid(rows = vars(sce$Object_ID))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction1)+
  facet_grid(rows = vars(sce$Object_ID))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_2, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  facet_grid(rows = vars(sce$Object_ID))

```

It is not enriched in old because one celltype is more abundant in old.
It is enriched in almost every old cell type compared to the same young cell 
type.
It is especially more enriched in old Ery progenitors in the two outlier batches.

After looking at the species report I am inclined to believe that this is
possibly a batch effect or possibly related to age.
Maybe an old ery signature? But it is also enriched in othe cells.

# Program 3

```{r}

print(table(sce$hsc_program_3 > 0))
print(table(sce$hsc_program_3 > 1))
print(table(sce$hsc_program_3 > 2))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)

```

Looks like a mature myeloid cell signature.
But also includes some other mature lymphoid cells.

```{r}

program_3_genes <- data.frame(
  genes = rownames(metadata(sce)$Genescores_NMF_hsc),
  scores = metadata(sce)$Genescores_NMF_hsc$hsc_program_3)

program_3_genes <- program_3_genes[order(program_3_genes$scores, 
                                         decreasing = TRUE),]
program_3_genes[1:50,]

```

Ngp, Lcn2, neutrphil related
Camp, Chil3, S100a8 Immune system related

Could be inflammatory/immune system program/mature immune cell program?

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Batch_exp_day, 
                                     color = sce$Batch_exp_day))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot( outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Experimental day", values = col_batch_exp_day)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Batch_sequencing, 
                                     color = sce$Batch_sequencing))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Age_ID, 
                                     color = sce$Age_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Age", values = col_ann_age)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3, 
                                     x = sce$Object_ID, 
                                     color = sce$Object_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")

```

Not specifically enriched in any batch or age.

# Program 4

```{r}

print(table(sce$hsc_program_4 > 0))
print(table(sce$hsc_program_4 > 1))
print(table(sce$hsc_program_4 > 2))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)

```

This seems monocyte/dendritic skewed but generally also lymphoid.

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_3,
                                     x = sce$hsc_program_4))+
  geom_bin2d()+
  theme_classic()

gg_df <- data.frame(colData(sce)) 
gg_df <- gg_df[-which(gg_df$hsc_program_3 == 0 | gg_df$hsc_program_4 == 0),]

ggplot(gg_df, aes(y = hsc_program_3, x = hsc_program_4))+
  geom_bin2d()+
  theme_classic()

gg_df <- data.frame(colData(sce)) 
gg_df <- gg_df[-which(gg_df$hsc_program_3 <1 | gg_df$hsc_program_4 <1 ),]

ggplot(gg_df, aes(y = hsc_program_3, x = hsc_program_4))+
  geom_bin2d()+
  theme_classic()

```

Seems like one clearly separated population of program 3 high cells that
does not overlap with program 4 high cells (and vice versa).

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Batch_exp_day, 
                                     color = sce$Batch_exp_day))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot( outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Experimental day", values = col_batch_exp_day)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Batch_sequencing, 
                                     color = sce$Batch_sequencing))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Age_ID, 
                                     color = sce$Age_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Age", values = col_ann_age)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Object_ID, 
                                     color = sce$Object_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")

```

Clearly enriched in Batch d

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Object_ID, 
                                     color = sce$Object_ID))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  facet_grid(cols = vars(sce$Batch_exp_day))

```

```{r fig.height = 12}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)+
  facet_grid(rows = vars(sce$Batch_exp_day))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  facet_grid(rows = vars(sce$Batch_exp_day))

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_4, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  facet_grid(rows = vars(sce$Batch_exp_day))

```

The enrichment in Batch d is not cell-type specific, this is therefore most
likely a batch effect program specifically for batch d.

```{r}

program_4_genes <- data.frame(
  genes = rownames(metadata(sce)$Genescores_NMF_hsc),
  scores = metadata(sce)$Genescores_NMF_hsc$hsc_program_4)

program_4_genes <- program_4_genes[order(program_4_genes$scores, 
                                         decreasing = TRUE),]
program_4_genes[1:50,]

```



# Program 5

```{r}

print(table(sce$hsc_program_5 > 0))
print(table(sce$hsc_program_5 > 1))
print(table(sce$hsc_program_5 > 2))

```

```{r}

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_5, 
                                     x = sce$Identity_ref_all, 
                                     color = sce$Identity_ref_all))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity1", values = col_cts_all)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_5, 
                                     x = sce$Identity_ref_fraction1, 
                                     color = sce$Identity_ref_fraction1))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity2", values = col_cts_fraction1)

ggplot(data.frame(colData(sce)), aes(y = sce$hsc_program_5, 
                                     x = sce$Identity_ref_fraction2, 
                                     color = sce$Identity_ref_fraction2))+
  ggbeeswarm::geom_quasirandom(size = 0.7)+
  geom_boxplot(outlier.alpha = 0, color = "black", alpha = 0)+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "none")+
  scale_color_manual("Identity3", values = col_cts_fraction2)

```

Probably a lymphoid cell program.

```{r}

program_5_genes <- data.frame(
  genes = rownames(metadata(sce)$Genescores_NMF_hsc),
  scores = metadata(sce)$Genescores_NMF_hsc$hsc_program_5)

program_5_genes <- program_5_genes[order(program_5_genes$scores, 
                                         decreasing = TRUE),]
program_5_genes[1:50,]

```


