---
title: "Tabula sapiens bone marrow batch-corrected reclustering"
author: "Lea Wölbert"
date: '2025-12-11'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

Similarly to before this script will be used to recluster and annotate 
one of the published datasets, in this case tabula sapiens bone marrow.

This time, I will perform the standard clustering pipeline, but 
- FIRST subset to signature gene orthologs only,
- THEN perform batch correction (or I'd need to calculate variable features),
- THEN cluster with the same resolution used for the non-batch-corrected
signature-based reclustering, 
- and THEN do a blind annotation.

This is to show the effect of batch correction as would be done in a 
normal preprocessing pipeline.

Again, I will save the resulting object from within this script 
for simplicity reasons.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)

# see which seurat version 
utils::sessionInfo()
```

```{r load_objects}

# the un-clustered original object
seu_ts_bone_marrow <- readRDS(snakemake@input[["seu_input"]])
#seu_ts_bone_marrow <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/ts_bone_marrow")

# this contains the ensembl IDs for orthologs of ONLY the signature genes
# used for subsetting later
ensembl_sign <- base::readRDS(snakemake@input[["ensembl_sign"]])
#ensembl_sign <- readRDS("/home/l012t/data/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/02_endf/ensembl_sign_hsc")
head(ensembl_sign)

# this contains info on the selected resolution for each dataset
# for clustering with each gene set
resolution_df_path <- snakemake@input[["resolution_df_path"]]
# resolution_df_path <- "/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt"

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")
```

```{r output_path}
# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/06_include_batch_correction/ts_bone_marrow.Rds")
print(output_path)
```

# Batch correction and clustering

```{r resolution}
# get the resolution used to recluster the same object with signature genes
# but without batch-correction
resolution <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "conserved_signature") %>%
  dplyr::pull()
resolution
```

Similar to the seurat pipeline roughly taken from sce_functions_reclustering.R
The seurat object should be pretty basic.

Adding a step for batch correction. 
Still keeping it very basic here.
Sticking to this vignette: https://satijalab.org/seurat/archive/v4.3/integration_introduction
because the original clustering was done in Seurat v4.3.0

```{r subset}

# subset right from the beginnging
# this is because if variable or integration features are calculated, 
# they will not include enough/all signature genes
# seems a little silly but this also seems to be the best option right now

ensembl_column_use <- seu_ts_bone_marrow@misc$ensembl_column_use
ensembl_column_use

length(unique(ensembl_sign$MMUS_SYMBOL)) # there are 229 signature genes

sign_ids <- unique(ensembl_sign[,ensembl_column_use])
length(sign_ids) # and 226 int features that are orthologs of those
table(sign_ids %in% rownames(seu_ts_bone_marrow)) # of which 225 are in the objects rownames

# subset seurat_object 
seu_ts_bone_marrow_sub <- seu_ts_bone_marrow[
  rownames(seu_ts_bone_marrow) %in% sign_ids,]
```

```{r seurat_clustering_pipeline_split}

# start from raw data
seu_ts_bone_marrow_sub@misc$data_use 

seu_ts_bone_marrow_list <- Seurat::SplitObject(
  seu_ts_bone_marrow_sub, 
  split.by = "donor_id")
# donor_id should cover both assay and sex which have had
# the strongest effects in the other clusterings and annotations 

# normalize and identify variable features for each dataset independently
seu_ts_bone_marrow_list <- lapply(
  X = seu_ts_bone_marrow_list, 
  FUN = function(seu){
    
    seu <- Seurat::NormalizeData(
      seu, 
      assay = "RNA", 
      verbose = FALSE)

    seu <- Seurat::FindVariableFeatures(
      seu, 
      selection.method = "vst", 
      nfeatures = nrow(seu)) # just use all genes

    return(seu)
})
seu_ts_bone_marrow_list
```

```{r seurat_clustering_pipeline_integration}
# start integration

int_features <- Seurat::SelectIntegrationFeatures(
  object.list = seu_ts_bone_marrow_list)
length(int_features) # it's the same signature genes

anchors_ts_bone_marrow <- Seurat::FindIntegrationAnchors(
  object.list = seu_ts_bone_marrow_list,
  anchor.features = int_features,
  verbose = FALSE)

combined_ts_bone_marrow <- Seurat::IntegrateData(
  anchorset = anchors_ts_bone_marrow,
  verbose = FALSE)

dim(combined_ts_bone_marrow)
combined_ts_bone_marrow
```

```{r seurat_clustering_pipeline_downstream}
# run the downstream analysis on the integrated data
DefaultAssay(combined_ts_bone_marrow) <- "integrated"

combined_ts_bone_marrow <- Seurat::ScaleData(
  combined_ts_bone_marrow, 
  verbose = FALSE)

# uses scaled data
combined_ts_bone_marrow <- Seurat::RunPCA(
  combined_ts_bone_marrow,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(combined_ts_bone_marrow@reductions$pca)
print(nr_pca)

# requires PC coordinates
combined_ts_bone_marrow <- Seurat::FindNeighbors(
  combined_ts_bone_marrow, 
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
combined_ts_bone_marrow <- Seurat::FindClusters(
  combined_ts_bone_marrow, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
combined_ts_bone_marrow <-  Seurat::RunUMAP(
  combined_ts_bone_marrow,
  dims = 1:10)
combined_ts_bone_marrow
```

# Evaluate

```{r table}
table(combined_ts_bone_marrow@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible remaining batch effects.

```{r umaps_base, fig.height = 6, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap")
``` 

```{r umaps_base_donid, fig.height = 6, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap", group.by = "donor_id")
``` 

```{r umaps_dev, fig.height = 6, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(combined_ts_bone_marrow@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(combined_ts_bone_marrow@meta.data$sex), ncol = 2)
``` 

```{r umaps_sex_assay, fig.height = 3, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap", group.by = "assay")+
  ggplot2::facet_wrap(
    facets = vars(combined_ts_bone_marrow@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay_donid, fig.height = 3, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap", group.by = "assay")+
  ggplot2::facet_wrap(
    facets = vars(combined_ts_bone_marrow@meta.data$donor_id), ncol = 2)
``` 

```{r umaps_sex_donid, fig.height = 3, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap", group.by = "sex")+
  ggplot2::facet_wrap(
      facets = vars(combined_ts_bone_marrow@meta.data$donor_id), ncol = 2)
``` 

```{r umaps_assay, fig.height = 6, fig.width = 6}
Seurat::DimPlot(combined_ts_bone_marrow, reduction = "umap")+
  ggplot2::facet_wrap(facets = vars(combined_ts_bone_marrow@meta.data$assay), ncol = 2)
``` 

It looks well-integrated, especially in comparison to the other instances
of reclustering and annotation.

Now continue with the blind annotation.

# Blind annotation

## Calculate markers

```{r marker_list}
clust_list <- as.list(levels(combined_ts_bone_marrow$seurat_clusters))

ct_markers_ts_bone_marrow <- lapply(clust_list, function(x){
  markers <- Seurat::FindMarkers(combined_ts_bone_marrow,
                                 test.use = "wilcox",
                                 ident.1 = x)

  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]

  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  # need to add only ensembl gene symbols as I will not be using others
  markers$symbol <- markers$genes
  markers$symbol <- ensembl_sign$HSAPIENS_SYMBOL[
    base::match(markers$genes, ensembl_sign$ENSG_ID)]

  return(markers)
})
names(ct_markers_ts_bone_marrow) <- levels(combined_ts_bone_marrow$seurat_clusters)

combined_ts_bone_marrow$batchc_sign_labels <- vector(length = ncol(combined_ts_bone_marrow))
```

Will go in the order of 1, 7, 0, 8 (interspersed with 0), 3, 6, 2, 5, 4.

### Show markers

```{r cluster1}
nr <- "1"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# HLF, CD34, MSI2, MECOM, MPL among the top genes are major HSC genes

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "HSC"
```

```{r cluster7}
nr <- "7"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# LY86, ANXA2, SPI1, CLEC12A, CTSZ, MEF2C seem granu/mono to me
# possibly more mono depending on the other clusters

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Granu/Mono prog"
```

```{r cluster0}
nr <- "0"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# PRTN3, CTSG, MPO, CLEC12A, CEBPE, LY86, ANXA2, SPI1, CALR, CD63, CALR
# so this is early myeloid with both neutro and mono potential

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Early Myeloid"
```

```{r cluster8}
nr <- "8"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# very short list of markers but those identified are granu-like baso/
# mast/eo (LMO4, CD63, GATA2)

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Baso/Eo/Mast/Granu"
```

```{r cluster3}
nr <- "3"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# definitely some cycling, but also some myeloid signature with
# ANXA2, CTSZ, LY86, SPI1

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Cycling myeloid"
```

```{r cluster6}
nr <- "6"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# TSPO2, HMBS, ERMAP, AQP1, KLF1, EPOR in that order look 
# already like fairly mature erythroid cells, let's see what the others
# look like

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Late Ery"
```

```{r cluster2, fig.height = 7, fig.width = 5}
nr <- "2"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# higher KLF1, GATA1 EPOR in this population make it seem earlier to me

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Early Ery"
```

```{r cluster5}
nr <- "5"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# ITGA2B, RAMP1, GATA2, TGFB1, MEF2C, ZFPM1 give it a more mk than ery
# profile, KLF1, etc is missing

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Mk"
```

```{r cluster4}
nr <- "4"
markers <- ct_markers_ts_bone_marrow[[nr]]$genes[
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) &
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05]

ct_markers_ts_bone_marrow[[nr]][
  !is.na(ct_markers_ts_bone_marrow[[nr]]$symbol) & 
    ct_markers_ts_bone_marrow[[nr]]$avg_log2FC > 0 &
    ct_markers_ts_bone_marrow[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

Seurat::FeaturePlot(combined_ts_bone_marrow, markers[1:6])
Seurat::FeaturePlot(combined_ts_bone_marrow, markers[7:12])
# also expression of ITGA2B, ZFPM1, GATA2, but KLF1, EPOR are also
# detected, so some Ery profile as well

combined_ts_bone_marrow$batchc_sign_labels[
  combined_ts_bone_marrow$seurat_clusters == nr] <- "Mk/Ery"

head(combined_ts_bone_marrow$batchc_sign_labels)
sort(unique(combined_ts_bone_marrow$batchc_sign_labels))

knitr::knit_exit()
```

Factorize for nicer plots.

```{r others2_level}
combined_ts_bone_marrow$batchc_sign_labels <- factor(
  combined_ts_bone_marrow$batchc_sign_labels,
  levels = c(  
    "HSC",
    "Early Myeloid",
    "Granu/Mono prog",
    "Cycling Myeloid",
    "Baso/Eo/Mast/Granu",
    "Mk",
    "Mk/Ery",
    "Early Ery",
    "Late Ery"
   ))

stopifnot(!is.na(combined_ts_bone_marrow$batchc_sign_labels))
```

## Save

```{r save}
saveRDS(combined_ts_bone_marrow, output_path)
```
