---
title: "HSC annotation"
author: "Lea Wölbert"
date: '2023-01-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r load,  message = FALSE}
library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(Seurat, quietly = TRUE) 
library(GOfuncR)
set.seed(37)
```

```{r}
color_tables <- "/omics/odcf/analysis/OE0538_projects/DO-0008/metadata/color_tables"
source(file = "../source/colors.R")
source(file = "../source/plotting.R")
source(file = "../source/sce_functions.R")

sce_hsc <- readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/sce_objects/10_lcls_fractions/sce_hsc-10")
fraction_curr <- sce_hsc$Fraction_ID[1]

```

```{r}

sce_hsc <- factor_reference_cts(sce_hsc)

col_cts_baccin <- col_cts[match(levels(sce_hsc$baccin_celltype_scmapclust), 
                                names(col_cts))]
col_cts_dahlin <- col_cts[match(levels(sce_hsc$dahlin_celltype_scmapclust), 
                                names(col_cts))]
col_cts_dolgalev <- col_cts[match(levels(sce_hsc$dolgalev_celltype_scmapclust),
                                  names(col_cts))]

col_batch_exp_day <- col_alp[names(col_alp) %in% sce_hsc$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% sce_hsc$Batch_sequencing]
col_ann_age <- col_ann[names(col_ann) %in% sce_hsc$Age_ID]
col_ann_fraction <- col_ann[names(col_ann) %in% sce_hsc$Fraction_ID]
col_species <- col_spc[names(col_spc) %in% sce_hsc$Species_ID]

col_clust_l <- col_num[names(col_num) %in% sce_hsc$cluster_louvain]

```

# Reference cell types 

Cells were annotated with scmap and three reference datasets.

```{r}

identity_plots <- function(sce){
  
  plot_1 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+ 
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin)
  plot_1l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust")+
    scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin)
  legend_1 <- get_legend(plot_1l)

  plot_2 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin)
  plot_2l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust")+
    scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin)
  legend_2 <- get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+ 
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("dolgalev_celltype_scmapclust",
                       values = col_cts_dolgalev)
  plot_3l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
    scale_color_manual("dolgalev_celltype_scmapclust", 
                       values = col_cts_dolgalev)
  legend_3 <- get_legend(plot_3l)
  
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)

  return(list(PLOT_1, PLOT_2, PLOT_3))
}

identity_plots <- identity_plots(sce_hsc)

```

```{r fig.width=8, fig.height=4}
identity_plots
```

# Clusters

```{r fig.width=8, fig.height=4}

plot_1 <- umap_base(sce_hsc, color_by = "cluster_louvain")+ 
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_1l <- umap_legend(sce_hsc, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_1 <- get_legend(plot_1l)

plot_2 <- pca_base2(sce_hsc, color_by = "cluster_louvain")+ 
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_2l <- pca_legend(sce_hsc, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_2 <- get_legend(plot_2l)

plot_3 <- pca_base3(sce_hsc, color_by = "cluster_louvain")+ 
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_3l <- pca_legend(sce_hsc, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_3 <- get_legend(plot_3l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
  
PLOT_1
PLOT_2
PLOT_3

```

# Literature marker gene expression

Using marker genes from literature to cell types.

### Functions

```{r}

gene_exp_umaps <- function(gene){
 
  plot_1 <- umap_gene(sce_hsc, gene)
  return(plot_1)

}

gene_exp_bees <- function(gene, ggdf){
  
  plot <- ggplot(ggdf, aes(x = cluster_louvain,
                           y = ggdf[,colnames(ggdf) == gene], 
                           color = cluster_louvain))+
    ggbeeswarm::geom_quasirandom(size = 0.2)+
    theme_classic()+
    scale_color_manual(values = col_clust_l)+
    geom_boxplot(color = "black", alpha = 0)+
    ylab(paste0(gene, " expression"))
  
  return(plot)
}

```


## HSCs

Not necessarily quiescent or LT.

```{r}

hsc_marker_genes <- list(
  "Procr", # LTHSCs
  "Cish", # LTHSCs
  "Esam", # LTHSCs
  "Fgd5", # LTHSCs
  "Tie1", # LTHSCs
  "Tinagl1", # LTHSCs
  "Trpc6", # LTHSCs
  "Hoxb5", # LTHSCs
  "Mmrn1", # LTHSCs
  "Ctnnal1", # LTHSCs 

  "Vwf", # HSCs, Mk cells, stroma cells

  "Hoxa9", # differentiation HSC TF
  "Junb", # HSC quiescence TF
  "Bclaf1", # HSC quiescence TF
  "Kdm5b", # non-quiescent HSCs TF
  "Klf1", # quiescent HSCs and erythroid priming TF
  "Tal1", # quiescent HSCs TF, cell cycle
  "Zbtb7a", # TF in self-renewing HSCs
  "Egr1", # "HSC quiescence TF
  "Foxp1", # non-quiescent HSC TF
  "Cebpe", # HSC quiescence TF
  "Gata2" # undifferentiated HSC TF, ery and Mk priming
)

```

For sources see Marker_list.xlsx

```{r}

plotlist_hsc <- lapply(hsc_marker_genes, gene_exp_umaps)
plotlist_hsc

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in hsc_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plot_list_hsc2 <- lapply(hsc_marker_genes, ggdf = ggdf, gene_exp_bees)
plot_list_hsc2

```

## Ery/Mk priming

```{r}

ermk_marker_genes <- list(
  "Gata1", # ery and Mk priming, also TF
  "Gypa", # MEPs
  "Hba-a1", # MEPs
  "Smim1", # MEPs
  "Snca", # MEPs
  "Col5a1", # MEPs
  "Vwf", # MEPs, HSCs, and stromal
  "Junb",  # ery and Mk priming TF
  "Zfpm1" # ery and Mk priming TF
)

```

```{r}

plotlist_erymk <- lapply(ermk_marker_genes, gene_exp_umaps)
plotlist_erymk

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in ermk_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plotlist_erymk2 <- lapply(ermk_marker_genes, ggdf = ggdf, gene_exp_bees)
plotlist_erymk2

```

# Ery priming

```{r}

ery_marker_genes <- list(
  "Gata1", # Erys, Mks, and TF for both priming
  "Klf1" # erythroid and quiescent HSCs
)

"Klf1" %in% rownames(sce_hsc)

```

```{r}

plotlist_ery <- lapply(ery_marker_genes, gene_exp_umaps)
plotlist_ery

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in ery_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plotlist_ery2 <- lapply(ery_marker_genes, ggdf = ggdf, gene_exp_bees)
plotlist_ery2

```

## MK priming

```{r}

mk_marker_genes <- list(
  "Fli1", # erythroid and quiescent HSCs
  "Gata1" # Erys, Mks, and TF for both priming
)

```

```{r}

plotlist_mk <- lapply(mk_marker_genes, gene_exp_umaps)
plotlist_mk

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in mk_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plotlist_mk2 <- lapply(mk_marker_genes, ggdf = ggdf, gene_exp_bees)
plotlist_mk2

```

# Lymphoid priming

```{r}

lym_marker_genes <- list(
  "Zbtb7a", # lymhpoid priming TF
  "Ikzf1" # lymhpoid priming TF
)

```

```{r}

plotlist_lym <- lapply(lym_marker_genes, gene_exp_umaps)
plotlist_lym

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in lym_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plotlist_lym2 <- lapply(lym_marker_genes, ggdf = ggdf, gene_exp_bees)
plotlist_lym2

```

## Myeloid priming

```{r}

mye_marker_genes <- list(
  "Cebpa", # Neu/Mast priming TF
  "Gata2", # mast cell priming TF
  "Spi1" # Neu/Mast priming TF
)

"Spi1" %in% rownames(sce_hsc)

```

```{r}

plotlist_mye <- lapply(mye_marker_genes, gene_exp_umaps)
plotlist_mye

```

```{r}

ggdf <- as.data.frame(colData(sce_hsc))
logc <- assays(sce_hsc)[["logcounts"]]

for(i in mye_marker_genes){
  ggdf$new <- logc[which(rownames(logc) == i),]
  colnames(ggdf)[colnames(ggdf) == "new"] <- i
}

plotlist_mye2 <- lapply(mye_marker_genes, ggdf = ggdf, gene_exp_bees)
plotlist_mye2

```

# Find marker genes

Use seurats wilcoxon test-based method
(see https://www.biorxiv.org/content/10.1101/2022.05.09.490241v1)

```{r}

seurat <- as.Seurat(sce_hsc, counts = "counts", data = "logcounts",) 
Idents(seurat) <- sce_hsc$cluster_louvain
tail(match(names(Idents(seurat)), rownames(colData(sce_hsc))))

hvgs <- modelGeneVar(sce_hsc)
hvgs <- getTopHVGs(hvgs, n=2000)

clustlist <- as.list(1:length(unique(sce_hsc$cluster_louvain)))
cluster_markers <- lapply(clustlist, function(x){
  markers <- FindMarkers(seurat, test.use = "wilcox", ident.1 = x, 
                         features = hvgs)
  markers <- markers[order(abs(markers$avg_log2FC), decreasing=TRUE),]
  return(markers)
})

cluster_markers

```

## Visualise

```{r}

c <- 0
marker_plotlist <- lapply(cluster_markers, function(x){
  
  c <- c + 1
  if(length(which(x$avg_log2FC >0))>6){
    markers_pos <- as.list(rownames(x[x$avg_log2FC >0,])[1:6])
  }else{
    markers_pos <- as.list(rownames(x[x$avg_log2FC >0,])) 
  }
  if(length(which(x$avg_log2FC <0))>6){
    markers_neg <- as.list(rownames(x[x$avg_log2FC <0,])[1:6])
  }else{
    markers_neg <- as.list(rownames(x[x$avg_log2FC <0,]))
  }

  umap_list_pos <- lapply(markers_pos, function(y){
    plot <- gene_exp_umaps(y)
    plot <- plot + ggtitle(paste0("cluster ", c, " positive marker"))
    return(plot)
  })
  umap_list_neg <- lapply(markers_neg, function(y){
    plot <- gene_exp_umaps(y)
    plot <- plot + ggtitle(paste0("cluster ", c, " negative marker"))
    return(plot)
  })
  
  ggdf <- as.data.frame(colData(sce_hsc))
  logc <- assays(sce_hsc)[["logcounts"]]

  for(i in markers_pos){
    ggdf$new <- logc[which(rownames(logc) == i),]
    colnames(ggdf)[colnames(ggdf) == "new"] <- i
  }
  bee_list_pos <- lapply(markers_pos, function(y){
   plot <- gene_exp_bees(y, ggdf = ggdf) 
   plot <- plot + ggtitle(paste0("cluster ", c, " positive marker"))
   return(plot)
  })
  
  for(i in markers_neg){
    ggdf$new <- logc[which(rownames(logc) == i),]
    colnames(ggdf)[colnames(ggdf) == "new"] <- i
  }
  bee_list_neg <- lapply(markers_neg, function(y){
   plot <- gene_exp_bees(y, ggdf = ggdf) 
   plot <- plot + ggtitle(paste0("cluster ", c, " negative marker"))
   return(plot)
  })
  
  return(list(umap_list_pos, bee_list_pos, umap_list_neg, bee_list_neg))

})

marker_plotlist

```

## GO analysis

```{r, message = FALSE}

get_go_results <- function(markers){
  
  go_input_df <- data.frame(
    gene_ids = rownames(markers),
    gene_scores =  markers$avg_log2FC
  )
                       
  go_result <- go_enrich(go_input_df, organismDb='Mus.musculus',
                         test='wilcoxon', n_randsets=100)
  return(go_result)  
}

go_results <- lapply(cluster_markers, get_go_results)

go_results_short <- lapply(go_results, function(x){
  results <- x$results[x$results$ontology == "biological_process",]
  return(results)
})

```

```{r}
go_results
go_results_short
```
