---
title: "clustering_report_full"
author: "Lea Wölbert"
date: '2022-12-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report for clustering as part of manual annotation at fraction level.

#### Load libraries, source code 

```{r load,  message = FALSE}
library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(dendextend, quietly = TRUE) 
#library(factoextra, quietly = TRUE) 
library(pheatmap, quietly = TRUE) 
library(bluster, quietly = TRUE)
library(RColorBrewer, quietly = TRUE)
set.seed(37)
```

```{r}
sce_inp <- readRDS(file = snakemake@input[["sce_input"]])
sce_h <- readRDS(file = snakemake@input[["sce_h"]])
sce_s <- readRDS(file = snakemake@input[["sce_s"]])
sce_l <- readRDS(file = snakemake@input[["sce_l"]])
fraction_curr <- sce_inp$Fraction_ID[1]

number_k <- snakemake@params[["number_k"]]
number_k_curr <- number_k[[fraction_curr]]

color_tables <- snakemake@params[["color_tables"]]
size_subs_hscs <- snakemake@params[["size_subs_hscs"]]

sce_l$cluster_seurat <- sce_s$cluster_seurat
sce_l$cluster_seurat_resolution <- sce_s$cluster_seurat_resolution

# check if cluster labels can just be transferred
print(ncol(sce_l))
print(tail(match(colnames(sce_l), colnames(sce_s))))

```

```{r}
source(file = "../../source/colors.R")
source(file = "../../source/plotting.R")
source(file = "../../source/sce_functions.R")
```

```{r}

sce_inp <- factor_reference_cts(sce_inp)
sce_h <- factor_reference_cts(sce_h)
sce_l <- factor_reference_cts(sce_l)

if(fraction_curr == "hsc"){
  col_cts["unassigned"] <- "honeydew2"
  col_num <- c(col_num, "DOWNSAMPLED" = "grey90")
}else if(fraction_curr == "str"){
  col_cts["unassigned"] <- "mistyrose3"
}

col_cts_baccin <- col_cts[match(levels(sce_l$baccin_celltype_scmapclust), 
                                names(col_cts))]
col_cts_dahlin <- col_cts[match(levels(sce_l$dahlin_celltype_scmapclust), 
                                names(col_cts))]
col_cts_dolgalev <- col_cts[match(levels(sce_l$dolgalev_celltype_scmapclust),
                                  names(col_cts))]

col_batch_exp_day <- col_alp[names(col_alp) %in% sce_l$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% sce_l$Batch_sequencing]
col_ann_age <- col_ann[names(col_ann) %in% sce_l$Age_ID]
col_ann_fraction <- col_ann[names(col_ann) %in% sce_l$Fraction_ID]
col_species <- col_spc[names(col_spc) %in% sce_l$Species_ID]

col_clust_h <- col_num[names(col_num) %in% sce_h$cluster_hierarchical]
col_clust_s <- col_num[names(col_num) %in% sce_l$cluster_seurat]
col_clust_l <- col_num[names(col_num) %in% sce_l$cluster_louvain]

```

Annotations for reference.

```{r}

identity_plots <- function(sce){
  
  plot_1 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+ 
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin)
  plot_1l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust")+
    scale_color_manual("baccin_celltype_scmapclust", values = col_cts_baccin)
  legend_1 <- get_legend(plot_1l)

  plot_2 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin)
  plot_2l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust")+
    scale_color_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin)
  legend_2 <- get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+ 
    ggtitle(paste0(fraction_curr))+
    scale_color_manual("dolgalev_celltype_scmapclust",
                       values = col_cts_dolgalev)
  plot_3l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
    scale_color_manual("dolgalev_celltype_scmapclust", 
                       values = col_cts_dolgalev)
  legend_3 <- get_legend(plot_3l)
  
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)

  return(list(PLOT_1, PLOT_2, PLOT_3))
}

identity_plots <- identity_plots(sce_h)

```

```{r fig.width=8, fig.height=4}
identity_plots
```

# Louvain

```{r}
print(table(sce_l$cluster_louvain))
```

```{r}
print(table(sce_l$cluster_louvain, sce_l$Species_ID))
```

```{r}
print(table(sce_l$cluster_louvain, sce_l$Age_ID))
```

## Clusters

```{r fig.width=8, fig.height=4}

plot_1 <- umap_base(sce_l, color_by = "cluster_louvain")+ 
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_1l <- umap_legend(sce_l, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_1 <- get_legend(plot_1l)

plot_2 <- pca_base2(sce_l, color_by = "cluster_louvain")+ 
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_2l <- pca_legend(sce_l, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_2 <- get_legend(plot_2l)

plot_3 <- pca_base3(sce_l, color_by = "cluster_louvain")+ 
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
plot_3l <- pca_legend(sce_l, color_by = "cluster_louvain")+
  scale_color_manual("cluster (louvain)", values = col_clust_l)
legend_3 <- get_legend(plot_3l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
  
PLOT_1
PLOT_2
PLOT_3
```

## QC

```{r fig.width=8, fig.height=4}

ggdf <- data.frame(colData(sce_l))

# legends
plot_baccin_leg <- ggplot(ggdf, aes(x = cluster_louvain, 
                                    fill=baccin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  scale_fill_manual("baccin_celltype_scmapclust", values = col_cts_baccin)

legend_baccin <- get_legend(plot_baccin_leg)

plot_dahlin_leg <- ggplot(ggdf, aes(x = cluster_louvain, 
                                    fill=dahlin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  scale_fill_manual("dahlin_celltype_scmapclust", values = col_cts_dahlin)

legend_dahlin <- get_legend(plot_dahlin_leg)

plot_dolgalev_leg <- ggplot(ggdf, aes(x = cluster_louvain, 
                                    fill=dolgalev_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  scale_fill_manual("dolgalev_celltype_scmapclust", values = col_cts_dolgalev)
legend_dolgalev <- get_legend(plot_dolgalev_leg)
# also use legends for later

# plots
plot1a <- ggplot(ggdf, aes(x = cluster_louvain, 
                          fill=Batch_exp_day, color = Batch_exp_day))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))

plot1b <- ggplot(ggdf, aes(x = cluster_louvain, fill=Batch_exp_day, 
                          color = Batch_exp_day))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))

plot2a <- ggplot(ggdf, aes(x = Batch_exp_day, 
                          fill=cluster_louvain, color = cluster_louvain))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (louvain)", values = col_clust_l)+
  scale_color_manual(name = "Cluster (louvain)", values = col_clust_l)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  facet_grid(cols = vars(ggdf$Fraction))

plot2b <- ggplot(ggdf, aes(x = Batch_exp_day, fill=cluster_louvain, 
                          color = cluster_louvain))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (louvain)",  values = col_clust_l)+
  scale_color_manual(name = "Cluster (louvain)", values = col_clust_l)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  facet_grid(cols = vars(ggdf$Fraction))

plot3a <- ggplot(ggdf, aes(x = cluster_louvain,
                          fill=Age_ID, color = Age_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))

plot3b <- ggplot(ggdf, aes(x = cluster_louvain, fill=Age_ID, 
                          color = Age_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))

plot4a <- ggplot(ggdf, aes(x = Age_ID, 
                          fill=cluster_louvain, color = cluster_louvain))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (louvain)", values = col_clust_l)+
  scale_color_manual(name = "Cluster (louvain)", values = col_clust_l)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))

plot4b <- ggplot(ggdf, aes(x = Age_ID,
                          fill=cluster_louvain, 
                          color = cluster_louvain))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (louvain)", values = col_clust_l)+
  scale_color_manual(name = "Cluster (louvain)", values = col_clust_l)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))


# Species
plot5a <- ggplot(ggdf, aes(x = cluster_louvain, 
                          fill=Species_ID, color = Species_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot5b <- ggplot(ggdf, aes(x = cluster_louvain, fill=Species_ID, 
                          color = Species_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot6a <- ggplot(ggdf, aes(x = Species_ID, 
                          fill=cluster_louvain, 
                          color = cluster_louvain))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_l)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_l)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot6b <- ggplot(ggdf, aes(x = Species_ID,
                          fill=cluster_louvain, 
                          color = cluster_louvain))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_l)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_l)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())


# Cell types
plot7a <- ggplot(ggdf, aes(x = cluster_louvain, 
                          fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust",
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust",
                     values = col_cts_baccin)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot7b <- ggplot(ggdf, aes(x = cluster_louvain, fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust",
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust",
                     values = col_cts_baccin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())


plot8a <- ggplot(ggdf, aes(x = cluster_louvain,
                          fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust", 
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust",
                     values = col_cts_dahlin)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot8b <- ggplot(ggdf, aes(x = cluster_louvain, fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust", 
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust",
                     values = col_cts_dahlin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())


plot9a <- ggplot(ggdf, aes(x = cluster_louvain,
                          fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust", 
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot9b <- ggplot(ggdf, aes(x = cluster_louvain,
                           fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust", 
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust",
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  theme(axis.title.y = element_blank())

plot1a
plot1b
plot2a
plot2b
plot3a
plot3b
plot4a
plot4b
plot5a
plot5b
plot6a
plot6b
```

```{r fig.width=8, fig.height = 11}

PLOT7a <- ggarrange(plot7a, legend_baccin, ncol = 1)
PLOT7b <- ggarrange(plot7b, legend_baccin, ncol = 1)
PLOT8a <- ggarrange(plot8a, legend_dahlin, ncol = 1)
PLOT8b <- ggarrange(plot8b, legend_dahlin, ncol = 1)
PLOT9a <- ggarrange(plot9a, legend_dolgalev, ncol = 1)
PLOT9b <- ggarrange(plot9b, legend_dolgalev, ncol = 1)

PLOT7a
PLOT7b
PLOT8a
PLOT8b
PLOT9a
PLOT9b
```

```{r fig.width=12, fig.height=4}

plot_x <- umap_base(sce_l, color_by = "cluster_louvain")+
  scale_color_manual("cluster_louvain", values = col_clust_l)+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  facet_grid(cols = vars(sce_l$Species_ID))
plot_xl <- umap_legend(sce_l, color_by = "cluster_louvain")+
  scale_color_manual("cluster_louvain", values = col_clust_l)
legend_x <- get_legend(plot_xl)

PLOT_X <-  ggarrange(plot_x, legend_x, widths = c(3,1))
PLOT_X
```

## Silhouette and Purity

```{r}

knitr::knit_exit()

# Silhouette
sil.approx <- approxSilhouette(reducedDim(sce_l, "PCA"), 
                               clusters = sce_l$cluster_louvain)
sil.data <- as.data.frame(sil.approx)
sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_l$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_l$cluster_louvain

ggplot(sil.data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

ggplot(sil.data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  facet_grid(cols = vars(sce_l$Species_ID))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

# Purity
pure_sce <- neighborPurity(reducedDim(sce_l, "PCA"), sce_l$cluster_louvain)
pure_data <- as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_l$cluster_louvain

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  facet_grid(cols = vars(sce_l$Species_ID))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, sce_l$k_louvain[1]))+
  scale_color_manual("cluster (louvain)", values = col_clust_l)+
  facet_grid(cols = vars(sce_l$Species_ID))+
  theme(legend.position = "none")+
  xlab("cluster (louvain)")+
  ylim(0, 1)+
  ylab("Cluster purity")
```

# Hierarchical clustering

```{r}
print(table(sce_h$cluster_hierarchical))
```

## Clusters

```{r fig.width=8, fig.height=4}

plot_1 <- umap_base(sce_h, color_by = "cluster_hierarchical")+ 
  ggtitle(paste0(fraction_curr, " ", 
                 length(which(sce_h$cluster_hierarchical != "DOWNSAMPLED")), 
                 " cells"))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
plot_1l <- umap_legend(sce_h, color_by = "cluster_hierarchical")+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
legend_1 <- get_legend(plot_1l)

plot_2 <- pca_base2(sce_h, color_by = "cluster_hierarchical")+ 
 ggtitle(paste0(fraction_curr, " ", 
                 length(which(sce_h$cluster_hierarchical != "DOWNSAMPLED")), 
                 " cells"))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
plot_2l <- pca_legend(sce_h, color_by = "cluster_hierarchical")+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
legend_2 <- get_legend(plot_2l)

plot_3 <- pca_base3(sce_h, color_by = "cluster_hierarchical")+ 
 ggtitle(paste0(fraction_curr, " ", 
                 length(which(sce_h$cluster_hierarchical != "DOWNSAMPLED")), 
                 " cells"))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
plot_3l <- pca_legend(sce_h, color_by = "cluster_hierarchical")+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)
legend_3 <- get_legend(plot_3l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
  
PLOT_1
PLOT_2
PLOT_3
```

## QC

```{r fig.width=8, fig.height=4}

ggdf <- data.frame(colData(sce_h))

# Experimental day
plot1a <- ggplot(ggdf, aes(x = cluster_hierarchical, 
                          fill=Batch_exp_day, color = Batch_exp_day))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
 ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot1b <- ggplot(ggdf, aes(x = cluster_hierarchical, fill=Batch_exp_day, 
                          color = Batch_exp_day))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot2a <- ggplot(ggdf, aes(x = Batch_exp_day, 
                          fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(ggdf$Species))+
  theme(axis.title.y = element_blank())

plot2b <- ggplot(ggdf, aes(x = Batch_exp_day, fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(ggdf$Species))+
  theme(axis.title.y = element_blank())

# Age
plot3a <- ggplot(ggdf, aes(x = cluster_hierarchical, 
                          fill=Age_ID, color = Age_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot3b <- ggplot(ggdf, aes(x = cluster_hierarchical, fill=Age_ID, 
                          color = Age_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot4a <- ggplot(ggdf, aes(x = Age_ID, 
                          fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot4b <- ggplot(ggdf, aes(x = Age_ID,
                          fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

# Species
plot5a <- ggplot(ggdf, aes(x = cluster_hierarchical,
                          fill=Species_ID, color = Species_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot5b <- ggplot(ggdf, aes(x = cluster_hierarchical, fill=Species_ID, 
                          color = Species_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot6a <- ggplot(ggdf, aes(x = Species_ID, 
                          fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot6b <- ggplot(ggdf, aes(x = Species_ID,
                          fill=cluster_hierarchical, 
                          color = cluster_hierarchical))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  scale_color_manual(name = "Cluster (hierarchical)", values = col_clust_h)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())


# Cell types
plot7a <- ggplot(ggdf, aes(x = cluster_hierarchical, 
                          fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust", 
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust",
                     values = col_cts_baccin)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot7b <- ggplot(ggdf, aes(x = cluster_hierarchical, 
                           fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust",
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust",
                     values = col_cts_baccin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())


plot8a <- ggplot(ggdf, aes(x = cluster_hierarchical,
                          fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust",
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust", 
                     values = col_cts_dahlin)+
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot8b <- ggplot(ggdf, aes(x = cluster_hierarchical,
                           fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust",
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust",
                     values = col_cts_dahlin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())


plot9a <- ggplot(ggdf, aes(x = cluster_hierarchical,
                          fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust",
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust",
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())

plot9b <- ggplot(ggdf, aes(x = cluster_hierarchical, 
                           fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust", 
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme(axis.title.y = element_blank())
 
plot1a
plot1b
plot2a
plot2b
plot3a
plot3b
plot4a
plot4b
plot5a
plot5b
plot6a
plot6b

```

```{r fig.width=8, fig.height = 11}

PLOT7a <- ggarrange(plot7a, legend_baccin, ncol = 1)
PLOT7b <- ggarrange(plot7b, legend_baccin, ncol = 1)
PLOT8a <- ggarrange(plot8a, legend_dahlin, ncol = 1)
PLOT8b <- ggarrange(plot8b, legend_dahlin, ncol = 1)
PLOT9a <- ggarrange(plot9a, legend_dolgalev, ncol = 1)
PLOT9b <- ggarrange(plot9b, legend_dolgalev, ncol = 1)

PLOT7a
PLOT7b
PLOT8a
PLOT8b
PLOT9a
PLOT9b
```

```{r fig.width=12, fig.height=4}

plot_x <- umap_base(sce_h, color_by = "cluster_hierarchical")+
  scale_color_manual("cluster_hierarchical", values = col_clust_h)+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(sce_h$Species_ID))
plot_xl <- umap_legend(sce_h, color_by = "cluster_hierarchical")+
  scale_color_manual("cluster_hierarchical", values = col_clust_h)
legend_x <- get_legend(plot_xl)

PLOT_X <-  ggarrange(plot_x, legend_x, widths = c(3,1))
PLOT_X
```

## Silhouette and Purity

From http://bioconductor.org/books/3.16/OSCA.advanced/clustering-redux.html

```{r}

# Silhouette 
sil_approx <- approxSilhouette(reducedDim(sce_h, "PCA"), 
                               clusters = sce_h$cluster_hierarchical)

sil_data <- as.data.frame(sil_approx)
sil_data$closest <- factor(ifelse(sil_data$width > 0, 
                                  sce_h$cluster_hierarchical, 
                                  sil_data$other))
sil_data$cluster <- sce_h$cluster_hierarchical

ggplot(sil_data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

ggplot(sil_data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(sce_h$Species_ID))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

# Purity
pure_sce <- neighborPurity(reducedDim(sce_h, "PCA"), 
                           sce_h$cluster_hierarchical)
pure_data <- as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_h$cluster_hierarchical

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(sce_h$Species_ID))+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  theme_classic()+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  ggtitle(paste0(fraction_curr, " ", ncol(sce_h), " cells"))+
  facet_grid(cols = vars(sce_h$Species_ID))+
  theme_classic()+
  scale_color_manual("cluster (hierarchical)", values = col_clust_h)+
  theme(legend.position = "none")+
  xlab("cluster (hierarchical)")+
  ylim(0, 1)+
  ylab("Cluster purity")
```

### BOTH

```{r}

mat <- reducedDim(sce_h, "PCA") # PCA is batch corrected
mat <- scale(mat)
dist <- dist(mat, method = 'euclidean')
ward <- hclust(dist, method = "ward.D2")
dend <- as.dendrogram(ward)

col_dend <- color_branches(dend, k = (number_k_curr-1), 
                               groupLabels = FALSE, col = col_clust_h)
plot(col_dend, main = paste0(fraction_curr, 
                                 " , ward.D2, k=", (number_k_curr-1)))

col_dend <- color_branches(dend, k = number_k_curr, 
                               groupLabels = FALSE, col = col_clust_h)
plot(col_dend, main = paste0(fraction_curr, 
                                 " , ward.D2, used k=", number_k_curr))

col_dend <- color_branches(dend, k = (number_k_curr+1), 
                               groupLabels = FALSE, col = col_clust_h)
plot(col_dend, main = paste0(fraction_curr, 
                                 " , ward.D2, k=", (number_k_curr+1)))
```

# Seurat clustering

```{r}
print(table(sce_l$cluster_seurat))
```

## Clusters

```{r fig.width=8, fig.height=4}

plot_1 <- umap_base(sce_l, color_by = "cluster_seurat")+ # own function
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
plot_1l <- umap_legend(sce_l, color_by = "cluster_seurat")+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
legend_1 <- get_legend(plot_1l)

plot_2 <- pca_base2(sce_l, color_by = "cluster_seurat")+ 
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
plot_2l <- pca_legend(sce_l, color_by = "cluster_seurat")+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
legend_2 <- get_legend(plot_2l)

plot_3 <- pca_base3(sce_l, color_by = "cluster_seurat")+ 
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
plot_3l <- pca_legend(sce_l, color_by = "cluster_seurat")+
  scale_color_manual("cluster (seurat)", values = col_clust_s)
legend_3 <- get_legend(plot_3l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
  
PLOT_1
PLOT_2
PLOT_3
```

## QC

```{r fig.width=8, fig.height=4}

ggdf <- data.frame(colData(sce_l))

plot1a <- ggplot(ggdf, aes(x = cluster_seurat, 
                          fill=Batch_exp_day, color = Batch_exp_day))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot1b <- ggplot(ggdf, aes(x = cluster_seurat, fill=Batch_exp_day, 
                          color = Batch_exp_day))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Experimental day", values = col_batch_exp_day)+
  scale_color_manual(name = "Experimental day", values = col_batch_exp_day)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot2a <- ggplot(ggdf, aes(x = Batch_exp_day,
                          fill=cluster_seurat, color = cluster_seurat))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(name = "Cluster (seurat)", values = col_clust_s)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  facet_grid(cols = vars(ggdf$Fraction))+
  theme(axis.title.y = element_blank())

plot2b <- ggplot(ggdf, aes(x = Batch_exp_day, fill=cluster_seurat, 
                          color = cluster_seurat))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(name = "Cluster (seurat)", values = col_clust_s)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  facet_grid(cols = vars(ggdf$Fraction))+
  theme(axis.title.y = element_blank())

plot3a <- ggplot(ggdf, aes(x = cluster_seurat, 
                          fill=Age_ID, color = Age_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot3b <- ggplot(ggdf, aes(x = cluster_seurat, fill=Age_ID, 
                          color = Age_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_ann_age)+
  scale_color_manual(name = "Age", values = col_ann_age)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot4a <- ggplot(ggdf, aes(x = Age_ID,
                          fill=cluster_seurat, color = cluster_seurat))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual("Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(values = col_clust_s)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot4b <- ggplot(ggdf, aes(x = Age_ID,
                          fill=cluster_seurat, 
                          color = cluster_seurat))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(name = "Cluster (seurat)", values = col_clust_s)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", ggdf$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

# Species
plot5a <- ggplot(ggdf, aes(x = cluster_seurat, 
                          fill=Species_ID, color = Species_ID))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot5b <- ggplot(ggdf, aes(x = cluster_seurat, fill=Species_ID, 
                          color = Species_ID))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Age", values = col_species)+
  scale_color_manual(name = "Age", values = col_species)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot6a <- ggplot(ggdf, aes(x = Species_ID, 
                          fill=cluster_seurat, 
                          color = cluster_seurat))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(name = "Cluster (seurat)", values = col_clust_s)+
  theme(axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot6b <- ggplot(ggdf, aes(x = Species_ID,
                          fill=cluster_seurat, 
                          color = cluster_seurat))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "Cluster (seurat)", values = col_clust_s)+
  scale_color_manual(name = "Cluster (seurat)", values = col_clust_s)+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())


# Cell types
plot7a <- ggplot(ggdf, aes(x = cluster_seurat,
                          fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust", 
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust", 
                     values = col_cts_baccin)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot7b <- ggplot(ggdf, aes(x = cluster_seurat, fill=baccin_celltype_scmapclust, 
                          color = baccin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "baccin_celltype_scmapclust", 
                    values = col_cts_baccin)+
  scale_color_manual(name = "baccin_celltype_scmapclust", 
                     values = col_cts_baccin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())


plot8a <- ggplot(ggdf, aes(x = cluster_seurat, 
                          fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust", 
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust", 
                     values = col_cts_dahlin)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot8b <- ggplot(ggdf, aes(x = cluster_seurat, fill=dahlin_celltype_scmapclust, 
                          color = dahlin_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dahlin_celltype_scmapclust", 
                    values = col_cts_dahlin)+
  scale_color_manual(name = "dahlin_celltype_scmapclust", 
                     values = col_cts_dahlin)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())


plot9a <- ggplot(ggdf, aes(x = cluster_seurat, 
                          fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "stack") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust",
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot9b <- ggplot(ggdf, aes(x = cluster_seurat, 
                           fill=dolgalev_celltype_scmapclust, 
                          color = dolgalev_celltype_scmapclust))+
  geom_bar(position = "fill") + 
  theme_classic()+
  scale_fill_manual(name = "dolgalev_celltype_scmapclust", 
                    values = col_cts_dolgalev)+
  scale_color_manual(name = "dolgalev_celltype_scmapclust", 
                     values = col_cts_dolgalev)+
  theme(legend.position = "none", axis.text.x = element_text(angle = 90))+
  ggtitle(paste0(fraction_curr, " ", sce_l$cluster_seurat_resolution[1]))+
  theme(axis.title.y = element_blank())

plot1a
plot1b
plot2a
plot2b
plot3a
plot3b
plot4a
plot4b
plot5a
plot5b
plot6a
plot6b
```

```{r fig.width=8, fig.height = 11}

PLOT7a <- ggarrange(plot7a, legend_baccin, ncol = 1)
PLOT7b <- ggarrange(plot7b, legend_baccin, ncol = 1)
PLOT8a <- ggarrange(plot8a, legend_dahlin, ncol = 1)
PLOT8b <- ggarrange(plot8b, legend_dahlin, ncol = 1)
PLOT9a <- ggarrange(plot9a, legend_dolgalev, ncol = 1)
PLOT9b <- ggarrange(plot9b, legend_dolgalev, ncol = 1)

PLOT7a
PLOT7b
PLOT8a
PLOT8b
PLOT9a
PLOT9b
```

```{r fig.width=12, fig.height=4}

plot_x <- umap_base(sce_l, color_by = "cluster_seurat")+
  scale_color_manual("cluster_seurat", values = col_clust_s)+
  ggtitle(fraction_curr)+
  facet_grid(cols = vars(sce_l$Species_ID))
plot_xl <- umap_legend(sce_l, color_by = "cluster_seurat")+
  scale_color_manual("cluster_seurat", values = col_clust_s)
legend_x <- get_legend(plot_xl)

PLOT_X <-  ggarrange(plot_x, legend_x, widths = c(3,1))
PLOT_X
```

## Silhouette and Purity

```{r}

# Silhouette
sil.approx <- approxSilhouette(reducedDim(sce_l, "PCA"), 
                               clusters = sce_l$cluster_seurat)
sil.data <- as.data.frame(sil.approx)
sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_l$cluster_seurat, 
                                  sil.data$other))
sil.data$cluster <- sce_l$cluster_seurat

ggplot(sil.data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

ggplot(sil.data, aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  facet_grid(cols = vars(sce_l$Species_ID))+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(-1, 1)+
  ylab("Silhouette width")

# Purity
pure_sce <- neighborPurity(reducedDim(sce_l, "PCA"), sce_l$cluster_seurat)
pure_data <- as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_l$cluster_seurat

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  facet_grid(cols = vars(sce_l$Species_ID))+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(0, 1)+
  ylab("Cluster purity")

ggplot(pure_data, aes(x=cluster, y=purity, color=cluster))+
  geom_boxplot(outlier.size = 0.1)+
  theme_classic()+
  ggtitle(fraction_curr)+
  scale_color_manual("cluster (seurat)", values = col_clust_s)+
  facet_grid(cols = vars(sce_l$Species_ID))+
  theme(legend.position = "none")+
  xlab("cluster (seurat)")+
  ylim(0, 1)+
  ylab("Cluster purity")
```

# Comparison of labels

## Clusters

### Seurat and Hierarchical

```{r fig.height = 10}

if(fraction_curr == "str"){
  sce_l$cluster_hierarchical <- sce_h$cluster_hierarchical
  sce_temp <- sce_l
  sce_temp$cluster_hierarchical <- paste0(sce_temp$cluster_hierarchical,
                                          "_hierarchical")
  sce_temp$cluster_seurat <- paste0(sce_temp$cluster_seurat, "_seurat")
  
  mat_comp <- as.matrix(table(sce_temp$cluster_hierarchical,
                              sce_temp$cluster_seurat))
  
  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }

  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, 
           main = paste0(fraction_curr, " Seurat/Hierarchical overlap"))
  
  pheatmap(mat_comp_temp2, 
           main = paste0(fraction_curr, " Seurat/Hierarchical overlap"))
}else{
  print("no comparison for hierarchical clustering for HSCs")
}
```

### Hierarchical and Louvain

```{r fig.height = 10}

if(fraction_curr == "str"){

  sce_temp <- sce_l
  sce_temp$cluster_hierarchical <- paste0(sce_temp$cluster_hierarchical,
                                          "_hierarchical")
  sce_temp$cluster_louvain <- paste0(sce_temp$cluster_louvain, "_louvain")
  
  mat_comp <- as.matrix(table(sce_temp$cluster_hierarchical,
                              sce_temp$cluster_louvain))

  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                         which(colMeans(mat_comp) > 0)]

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }

  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, 
           main = paste0(fraction_curr, " Louvain/Hierarchical overlap"))
  
  pheatmap(mat_comp_temp2, 
           main = paste0(fraction_curr, " Louvain/Hierarchical overlap"))
}else{
  print("no comparison for hierarchical clustering for HSCs")
}
```

### Seurat and Louvain

```{r fig.height = 10}

sce_temp <- sce_l
sce_temp$cluster_seurat <- paste0(sce_temp$cluster_seurat, "_seurat")
sce_temp$cluster_louvain <- paste0(sce_temp$cluster_louvain, "_louvain")

mat_comp <- as.matrix(table(sce_temp$cluster_seurat,
                            sce_temp$cluster_louvain))

mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

mat_comp_temp1 <- mat_comp
for(i in 1:nrow(mat_comp_temp1)){
  mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
}

mat_comp_temp2 <- t(mat_comp)
for(i in 1:nrow(mat_comp_temp2)){
  mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
}

pheatmap(mat_comp_temp1, 
         main = paste0(fraction_curr, " Louvain/Seurat overlap"))

pheatmap(mat_comp_temp2, 
         main = paste0(fraction_curr, " Louvain/Seurat overlap"))
```

## Reference dataset

Use only baccin et al. dataset.

### Hierarchical

```{r fig.height = 10}

if(fraction_curr == "str"){

  sce_temp <- sce_l
  sce_temp$cluster_hierarchical <- paste0(sce_temp$cluster_hierarchical,
                                          "_hierarchical")

  mat_comp <- as.matrix(table(sce_temp$baccin_celltype_scmapclust,
                              sce_temp$cluster_hierarchical))
  
  mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

  mat_comp_temp1 <- mat_comp
  for(i in 1:nrow(mat_comp_temp1)){
    mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
  }
  
  mat_comp_temp2 <- t(mat_comp)
  for(i in 1:nrow(mat_comp_temp2)){
    mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
  }

  pheatmap(mat_comp_temp1, 
           main = paste0(fraction_curr, " Cell type/Hierarchical overlap"))
  
  pheatmap(mat_comp_temp2, 
           main = paste0(fraction_curr, " Cell type/Hierarchical overlap"))

}else{
  print("no comparison for hierarchical clustering for HSCs")
}
```

### Seurat

```{r fig.height = 10}

sce_temp <- sce_l
sce_temp$cluster_seurat <- paste0(sce_temp$cluster_seurat, "_seurat")

mat_comp <- as.matrix(table(sce_temp$baccin_celltype_scmapclust,
                            sce_temp$cluster_seurat))

mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

mat_comp_temp1 <- mat_comp
for(i in 1:nrow(mat_comp_temp1)){
  mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
}

mat_comp_temp2 <- t(mat_comp)
for(i in 1:nrow(mat_comp_temp2)){
  mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
}

pheatmap(mat_comp_temp1, 
         main = paste0(fraction_curr, " Cell type/Seurat overlap"))

pheatmap(mat_comp_temp2, 
         main = paste0(fraction_curr, " Cell type/Seurat overlap"))
```

### Louvain 

```{r fig.height = 10}

sce_temp <- sce_l
sce_temp$cluster_louvain <- paste0(sce_temp$cluster_louvain, "_louvain")

mat_comp <- as.matrix(table(sce_temp$baccin_celltype_scmapclust,
                            sce_temp$cluster_louvain))

mat_comp <- mat_comp[which(rowMeans(mat_comp) > 0), 
                       which(colMeans(mat_comp) > 0)]

mat_comp_temp1 <- mat_comp
for(i in 1:nrow(mat_comp_temp1)){
  mat_comp_temp1[i,] <- mat_comp_temp1[i,]/sum(mat_comp_temp1[i,])
}

mat_comp_temp2 <- t(mat_comp)
for(i in 1:nrow(mat_comp_temp2)){
  mat_comp_temp2[i,] <- mat_comp_temp2[i,]/sum(mat_comp_temp2[i,])
}

pheatmap(mat_comp_temp1, 
         main = paste0(fraction_curr, " Cell type/Louvain overlap"))

pheatmap(mat_comp_temp2, 
         main = paste0(fraction_curr, " Cell type/Louvain overlap"))
```

```{r}
sessionInfo()
```