---
title: "Supplementary Figure 2"
date: '2026-02-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is going to be a very big figure containing many subplots.
Trying to make it still relatively understandable.
This figure's code was split into multiple .Rmds because it was too
much data.
This part is concerned with making the schematics and the heatmaps.

## Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(scuttle, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@params[["base_path"]]
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")
```

### SCE objects

#### Objects with merged clustering, containing all cells

Because some cells were removed between this step and the final annotation.
So that we can show all cells, even contamination, these objects will 
serve as a basis and other annotations or reductions can be added. 
For instance, for the alluvial plots and the pre-removal umaps.

```{r load_all_cells_hsc}
sce_hsc_all_cells <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"
))
```

```{r load_all_cells_str}
sce_str_all_cells <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"
))
```

#### Finally annotated objects

These are needed for the heatmaps and for the final annotation names.

```{r load_anno_hsc}
sce_hsc_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"
))
```

```{r load_anno_str}
sce_str_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"
))
```

#### Species-specific custom annotations

These objects were annotated by species and fraction, so there are
eight objects in total.
These will be depicted individually.

```{r load_species_specific_hsc}

sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))
```

```{r load_species_specific_str}
sce_mmus_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds"))
sce_mcas_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds"))
sce_mspr_str <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds"))
sce_mcar_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds"))
```

## Params

```{r params}
# global params for figures specifically
# this is a supp figure so text size can be smaller (6-8 pts)
source("../figures/determine_params.R")

# local params
#legend_text_size_allu <- 13
umap_point_size <- 0.4
heatmap_text_size <- 14
colors_z_score_ct <- c("steelblue3", "white", "red3")


# cell types to be depicted in the heatmaps
# keeping the stem compartment for comparisons
cts_keep_hsc <- c(
  "HSC",
  "Early MPP",
  "Activated MPP",
  "Early MPP cells 1",
  "Early MPP cells 2",
  "HSC/Early Mk",
  "MPP cells 1",
  "MPP cells 2",
  "Activated MPP cells 1",
  "Activated MPP cells 2",
  "Late MPP",
  "MPP cells"
)
# it looks like a lot but each species has different names

cts_keep_hsc[!cts_keep_hsc %in% factors_custom_hsc]
cts_keep_hsc[!cts_keep_hsc %in% names(col_cts_hsc)]

# keeping possible CAR and osteo populations
cts_keep_str <- c(
  "Adipo/CAR",
  "Adipo CAR", 
  "Osteo",
  "Adipo/Osteo CAR",
  "Balanced mesenchymal",
  "Early chondro",
  "Late chondro and Osteo",
  "Late chondro", 
  "Articular/synovial-like",
  "Fibro/Chondro (more Fibro)",
  "Fibro",
  "Chondrocyte",
  "Fibro/Chondro",
  "Fibroblast"
)

cts_keep_str[!cts_keep_str %in% factors_custom_str]
cts_keep_str[!cts_keep_str %in% names(col_cts_str)]
```

####  Output paths

```{r pdf_output}
# define output paths for pdf export

output_pdf_a_umap_1 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_a_umap_1.pdf")
output_pdf_a_umap_2 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_a_umap_2.pdf")
output_pdf_a_umap_3 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_a_umap_3.pdf")

output_pdf_b_umap_1 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_b_umap_1.pdf")
output_pdf_b_umap_2 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_b_umap_2.pdf")
output_pdf_b_umap_3 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_b_umap_3.pdf")
output_pdf_b_umap_4 <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_b_umap_4.pdf")

output_pdf_b_umap_3_legend<- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_b_umap_3_legend.pdf")

output_pdf_c_combined_hsc <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_c_combined_hsc.pdf")
output_pdf_c_combined_hsc_legend <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_c_combined_hsc_legend.pdf")

output_pdf_d_combined_str <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_d_combined_str.pdf")
output_pdf_d_combined_str_legend <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_d_combined_str_legend.pdf")
```

## Prep 

### Randomize barcodes

#### HSC objects

```{r rand_barc_hsc}
# randomize the barcodes for nicer visualisation
dim(sce_hsc_all_cells)
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]
dim(sce_hsc_all_cells)
```

Also for the BL6 custom object.

```{r rand_barc_hsc_anno}
# randomize the barcodes for nicer visualisation
dim(sce_mmus_hsc)
sce_mmus_hsc <- sce_mmus_hsc[,base::sample(
  x = c(1:ncol(sce_mmus_hsc)), 
  size = ncol(sce_mmus_hsc),
  replace = FALSE)]
dim(sce_mmus_hsc)
```

Also for the finally annotated object.

```{r rand_barc_mmus_hsc}
# randomize the barcodes for nicer visualisation
dim(sce_hsc_anno)
sce_hsc_anno <- sce_hsc_anno[,base::sample(
  x = c(1:ncol(sce_hsc_anno)), 
  size = ncol(sce_hsc_anno),
  replace = FALSE)]
dim(sce_hsc_anno)
```

#### STR objects

```{r rand_barc_str}
# randomize the barcodes for nicer visualisation
dim(sce_str_all_cells)
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]
dim(sce_str_all_cells)
```

### Final anno

Add the labels used for analysis and shown in main Figure 1.

```{r add_final_anno_hsc}

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_hsc_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_hsc_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"
```

```{r add_final_anno_str}

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"
```

### Custom labels

These are the labels from the species-specific custom clustering.

```{r add_custom_hsc}

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), 
  colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), 
  colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), 
  colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), 
  colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label
```

```{r add_custom_str}

sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label <- unfactor(sce_mmus_str$custom_label)
sce_mcas_str$custom_label <- unfactor(sce_mcas_str$custom_label)
sce_mspr_str$custom_label <- unfactor(sce_mspr_str$custom_label)
sce_mcar_str$custom_label <- unfactor(sce_mcar_str$custom_label)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), 
  colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), 
  colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), 
  colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), 
  colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label
```

### Factorize

#### HSC objects

```{r factors_hsc}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[
  sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[
  sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[
  sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[
  sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

Also need to do this for the example BL6 HSC fraction object.
For the schematic A.

```{r factors_mmus_hsc}
# change species names to currently used abbreviations
sce_mmus_hsc$species_pub <- vector(length = ncol(sce_mmus_hsc))
sce_mmus_hsc$species_pub[
  sce_mmus_hsc$Species_ID == "mmus"] <- "BL6"
sce_mmus_hsc$species_pub[
  sce_mmus_hsc$Species_ID == "mcas"] <- "CAST"
sce_mmus_hsc$species_pub[
  sce_mmus_hsc$Species_ID == "mspr"] <- "SPRET"
sce_mmus_hsc$species_pub[
  sce_mmus_hsc$Species_ID == "mcar"] <- "CAROLI"
sce_mmus_hsc$species_pub <- factor(
  sce_mmus_hsc$species_pub, 
  levels = names(col_spc_pub))
```

Also need to do this for the final merged anno HSC object.
For the schematic A.

```{r factors_hsc_anno}
# change species names to currently used abbreviations
sce_hsc_anno$species_pub <- vector(length = ncol(sce_hsc_anno))
sce_hsc_anno$species_pub[
  sce_hsc_anno$Species_ID == "mmus"] <- "BL6"
sce_hsc_anno$species_pub[
  sce_hsc_anno$Species_ID == "mcas"] <- "CAST"
sce_hsc_anno$species_pub[
  sce_hsc_anno$Species_ID == "mspr"] <- "SPRET"
sce_hsc_anno$species_pub[
  sce_hsc_anno$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_anno$species_pub <- factor(
  sce_hsc_anno$species_pub, 
  levels = names(col_spc_pub))
```

#### STR objects

```{r factors_str}
# change species names to currently used abbreviations
sce_str_all_cells$species_pub <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mmus"] <- "BL6"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mcas"] <- "CAST"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_str_all_cells$species_pub <- factor(
  sce_str_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label,
  levels = factors_custom_str
)
```

# Plots

## Schematic A

This will be three different UMAPs, using the species-specific custom BL6 
HSPC fraction object as example.

### Prep DF

```{r a_umap_prep_dfs}

umap_df_mmus_hsc <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_mmus_hsc, type = "UMAP"))

# first show the species
umap_df_mmus_hsc$species_pub <- sce_mmus_hsc$species_pub 
# then the louvain clusterings
umap_df_mmus_hsc$cluster_louvain <- sce_mmus_hsc$cluster_louvain 
# then the annotated popuations
umap_df_mmus_hsc$custom_label <- sce_mmus_hsc$custom_label 

colnames(umap_df_mmus_hsc)[1] <- "UMAP1"
colnames(umap_df_mmus_hsc)[2] <- "UMAP2"
```

### Plot 

```{r a_umap_1, fig.width = 2, fig.height = 2}
a_umap_1 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = species_pub
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_spc_pub)+
  theme_void()+
  theme(legend.position = "none")

a_umap_1
```

```{r a_umap_2, fig.width = 2, fig.height = 2}
a_umap_2 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = as.character(cluster_louvain)
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_num)+
  theme_void()+
  theme(legend.position = "none")

a_umap_2
```

```{r a_umap_3, fig.width = 2, fig.height = 2}
a_umap_3 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = custom_label
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_custom_hsc)+
  theme_void()+
  theme(legend.position = "none")

a_umap_3
```

#### EXPORT

```{r export_a_umap_1}
pdf(output_pdf_a_umap_1, width = 1.5, height = 1.5)
a_umap_1
dev.off()
```

```{r export_a_umap_2}
pdf(output_pdf_a_umap_2, width = 1.5, height = 1.5)
a_umap_2
dev.off()
```

```{r export_a_umap_3}
pdf(output_pdf_a_umap_3, width = 1.5, height = 1.5)
a_umap_3
dev.off()
```

## Schematic B

This is a schematic for the merged annotation including sub-clustering
based on several objects.
It will contain several umaps, as well.

### Prep DFs

```{r b_prep_mixed_umap}

# all cells, colored by species
# and also colored by louvain cluster
umap_df_sce_hsc_all_cells <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_all_cells, type = "UMAP"))

umap_df_sce_hsc_all_cells$species_pub <- sce_hsc_all_cells$species_pub 
umap_df_sce_hsc_all_cells$cluster_louvain <- sce_hsc_all_cells$cluster_louvain 

colnames(umap_df_sce_hsc_all_cells)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_all_cells)[2] <- "UMAP2"

# also only the HSC - MPP portion, colored by Mecom expression
sce_hsc_stem <- sce_hsc_all_cells[,sce_hsc_all_cells$cluster_louvain %in% c(2, 4)]
dim(sce_hsc_stem)

umap_df_sce_hsc_stem <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_stem, type = "UMAP"))

# logcounts have been batch-normalized
logcounts_stem <- SingleCellExperiment::logcounts(sce_hsc_stem)
umap_df_sce_hsc_stem$expression <- logcounts_stem["Mecom",]

colnames(umap_df_sce_hsc_stem)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_stem)[2] <- "UMAP2"
```

```{r b_prep_hsc_anno}

# final annotation, last umap of the plot
# but still containing the contaminating cells

umap_df_sce_hsc_anno <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_all_cells, type = "UMAP"))
umap_df_sce_hsc_anno$celltypes <- sce_hsc_all_cells$celltypes 
umap_df_sce_hsc_anno$species_pub <- sce_hsc_all_cells$species_pub 

colnames(umap_df_sce_hsc_anno)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_anno)[2] <- "UMAP2"
```

### Plots

```{r b_umap_1, fig.width = 2, fig.height = 2}
b_umap_1 <- ggplot2::ggplot(
  umap_df_sce_hsc_all_cells,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = species_pub
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_spc_pub)+
  theme_void()+
  theme(legend.position = "none")

b_umap_1
```

```{r b_umap_2, fig.width = 2, fig.height = 2}
b_umap_2 <- ggplot2::ggplot(
  umap_df_sce_hsc_all_cells,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = as.character(cluster_louvain)
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_num)+
  theme_void()+
  theme(legend.position = "none")

b_umap_2
```

```{r b_umap_3, fig.width = 2, fig.height = 2}
b_umap_3 <- ggplot2::ggplot(
  umap_df_sce_hsc_stem,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = expression
  ))+
  geom_point(size = 0.2)+
  scale_color_gradientn(colours = mycolors_to_max)+
  theme_void()+
  theme(legend.position = "none")

b_umap_3
```

```{r b_umap_4, fig.width = 2, fig.height = 2}
b_umap_4 <- ggplot2::ggplot(
  umap_df_sce_hsc_anno,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = celltypes
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_cts_hsc)+
  theme_void()+
  theme(legend.position = "none")

b_umap_4
```

#### Mecom legend

##### Legend STR

```{r b_legend_mecom, fig.width = 4, fig.height = 3}
legend_plot_mecom <- ggplot2::ggplot(
  umap_df_sce_hsc_stem,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = expression
  ))+
  ggplot2::geom_point()+
  ggplot2::scale_color_gradientn(colours = mycolors_to_max)+
  theme_all+
  ggplot2::theme(
    legend.text = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.title = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.position = "bottom"
  )

legend_mecom <- ggpubr::get_legend(legend_plot_mecom)
b_umap_3_legend <- ggpubr::ggarrange(legend_mecom)

b_umap_3_legend
```

#### EXPORT

```{r export_b_umap_1}
pdf(output_pdf_b_umap_1, width = 1.5, height = 1.5)
b_umap_1
dev.off()
```

```{r export_b_umap_2}
pdf(output_pdf_b_umap_2, width = 1.5, height = 1.5)
b_umap_2
dev.off()
```

```{r export_b_umap_3}
pdf(output_pdf_b_umap_3, width = 2.15, height = 2.78)
b_umap_3
dev.off()
```

```{r export_b_umap_3_legend}
pdf(output_pdf_b_umap_3_legend, width = 4, height = 2)
b_umap_3_legend
dev.off()
```

```{r export_b_umap_4}
pdf(output_pdf_b_umap_4, width = 1.5, height = 1.5)
b_umap_4
dev.off()
```

## Gene expression heatmaps 

### Prep data

Make appropriate dataframes.

The log-expression levels will be averaged per cell type and scaled in 
some way.
How exactly to scale each of the five datasets (merged + four species
datasets) was difficult to decide.
However, for simplicity, both different annotations (merged + custom)
will be treated the same way: averaging gene expression per cell type 
as is, then scaling across all cells ( = all cell types and all species).

I think this is the most simple and best options for the comparison we want
to make if we very clearly explain and state it.

I am using the logcounts based on the merged batch-normalisation for both the
orig clustering and the custom clusterings by using sce_hsc_all_cells as 
basis.
I think that is better comparable.
So by using the the merged batch-normalized logcounts, then even if a cell has 
two different labels in both, it has still the same expression levels.
If I used the species-specific values, the expression level would only be
normalized within each species, but not across, so we wouldn't really
be able to compare across species.
This is simply a way to compare the clustering partitions, so using the same
expression values for each cells is how I decided to go about this. 

#### HSCs

##### Aggregate by type of clustering

All cells from all cell types.

```{r aggregate_orig}

sce_hsc_all_cells_orig <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_orig$cluster_species <- base::paste0(
  sce_hsc_all_cells_orig$celltypes,
  "_",
  sce_hsc_all_cells_orig$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_orig <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_orig,
  id=colData(sce_hsc_all_cells_orig)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_orig_df <- logcounts(sce_agg_orig) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_orig$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_orig_df$clustering <- "original"

# separate cell type and species again
agg_orig_df <- agg_orig_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)

# factor species
agg_orig_df$species <- factor(
  agg_orig_df$species,
  levels = names(col_spc_pub))
```

```{r aggregate_cust}

sce_hsc_all_cells_cust <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_cust$cluster_species <- base::paste0(
  sce_hsc_all_cells_cust$custom_label,
  "_",
  sce_hsc_all_cells_cust$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_cust <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_cust,
  id=colData(sce_hsc_all_cells_cust)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_cust_df <- logcounts(sce_agg_cust) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_cust$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_cust_df$clustering <- "custom"

# separate cell type and species again
agg_cust_df <- agg_cust_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)
```

##### Subset to required genes and scale

```{r scale_function}
scale_expression <- function(expression_df){

  for(gene in unique(expression_df$genes)){
    expr_vector_g <- expression_df[
      expression_df$genes == gene,]$avg_logcounts

    # calculate z score per gene
    z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
      stats::sd(expr_vector_g)
    expression_df$zscore_gene[expression_df$genes == gene] <- z_score_g
  }
  return(expression_df)
}
```

```{r determine_genes}
# stem and MPP genes from the utils source .R script
marker_genes <- unlist(marker_genes_list_hsc_figs2)
```

```{r scale_orig}

# only keep genes for visualisation
agg_orig_df_genes <- agg_orig_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_orig_df_genes$genes <- factor(
  agg_orig_df_genes$genes,
  levels = rev(marker_genes))

agg_orig_df_genes$cluster <- factor(
  agg_orig_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$celltypes))

# scale expression per gene using z-scores
agg_orig_df_genes$zscore_gene <- vector(
  length = nrow(agg_orig_df_genes), 
  mode = "numeric")

agg_orig_df_vis <- scale_expression(agg_orig_df_genes)
```

```{r scale_cust}

# only keep genes for visualisation
agg_cust_df_genes <- agg_cust_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_cust_df_genes$genes <- factor(
  agg_cust_df_genes$genes,
  levels = rev(marker_genes))

agg_cust_df_genes$cluster <- factor(
  agg_cust_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$custom_label))

# scale expression per gene using z-scores
agg_cust_df_genes$zscore_gene <- vector(
  length = nrow(agg_cust_df_genes), 
  mode = "numeric")
```

```{r scale_cust_across}

agg_cust_df_genes <- scale_expression(agg_cust_df_genes)

agg_cust_df_mmus <- agg_cust_df_genes %>% 
  dplyr::filter(species == "BL6")
agg_cust_df_mcas <- agg_cust_df_genes %>% 
  dplyr::filter(species == "CAST")
agg_cust_df_mspr <- agg_cust_df_genes %>% 
  dplyr::filter(species == "SPRET")
agg_cust_df_mcar <- agg_cust_df_genes %>% 
  dplyr::filter(species == "CAROLI")
```

##### Subset to the required cell types

After scaling.

```{r subset_ct_keep}
agg_cust_df_mmus <- agg_cust_df_mmus[agg_cust_df_mmus$cluster %in% cts_keep_hsc,]
agg_cust_df_mcas <- agg_cust_df_mcas[agg_cust_df_mcas$cluster %in% cts_keep_hsc,]
agg_cust_df_mspr <- agg_cust_df_mspr[agg_cust_df_mspr$cluster %in% cts_keep_hsc,]
agg_cust_df_mcar <- agg_cust_df_mcar[agg_cust_df_mcar$cluster %in% cts_keep_hsc,]

agg_orig_df_vis <- agg_orig_df_vis[agg_orig_df_vis$cluster %in% cts_keep_hsc,]
```

#### Plot

##### Species-specific custom

Now four separate heatmaps per species-specific custom clustering.

```{r c_base_heatmap_cust_mmus, fig.height = 6, fig.width = 3}

c_base_heatmap_cust_mmus <- ggplot2::ggplot(
  agg_cust_df_mmus,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-7, -3.5, 0, 3.5, 7), 
    limits = c(-7, 7))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
c_base_heatmap_cust_mmus
```

```{r c_theme_heatmap_cust_mmus, fig.height = 6, fig.width = 2}
c_theme_heatmap_cust_mmus <- c_base_heatmap_cust_mmus+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    strip.background = element_blank(),
    axis.line = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
c_theme_heatmap_cust_mmus
```

```{r c_base_heatmap_cust_mcas, fig.height = 6, fig.width = 3}

c_base_heatmap_cust_mcas <- ggplot2::ggplot(
  agg_cust_df_mcas,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-7, -3.5, 0, 3.5, 7), 
    limits = c(-7, 7))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
c_base_heatmap_cust_mcas
```

```{r c_theme_heatmap_cust_mcas, fig.height = 6, fig.width = 2}
c_theme_heatmap_cust_mcas <- c_base_heatmap_cust_mcas+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
c_theme_heatmap_cust_mcas
```

```{r c_base_heatmap_cust_mspr, fig.height = 6, fig.width = 3}

c_base_heatmap_cust_mspr <- ggplot2::ggplot(
  agg_cust_df_mspr,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-7, -3.5, 0, 3.5, 7), 
    limits = c(-7, 7))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
c_base_heatmap_cust_mspr
```

```{r c_theme_heatmap_cust_mspr, fig.height = 6, fig.width = 2}
c_theme_heatmap_cust_mspr <- c_base_heatmap_cust_mspr+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
c_theme_heatmap_cust_mspr
```

```{r c_base_heatmap_cust_mcar, fig.height = 6, fig.width = 3}

c_base_heatmap_cust_mcar <- ggplot2::ggplot(
  agg_cust_df_mcar,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-7, -3.5, 0, 3.5, 7), 
    limits = c(-7, 7))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
c_base_heatmap_cust_mcar
```

```{r c_theme_heatmap_cust_mcar, fig.height = 6, fig.width = 2}
c_theme_heatmap_cust_mcar <- c_base_heatmap_cust_mcar+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
c_theme_heatmap_cust_mcar
```

Build the species-specific heatmaps together

```{r c_custom_hsc_heatmap, fig.height = 6, fig.width = 3.4}

c_custom_hsc_heatmap <- ggpubr::ggarrange(
  c_theme_heatmap_cust_mmus+ggplot2::theme(plot.margin = margin(0.1,0.1,0.4,0.1)),
  c_theme_heatmap_cust_mcas+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  c_theme_heatmap_cust_mspr+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  c_theme_heatmap_cust_mcar+ggplot2::theme(plot.margin = margin(0,0.1,0.1,0.1)),
  nrow = 1,
  widths = c(8, 6, 3, 2),
  align = c("h")
)

c_custom_hsc_heatmap
```

##### Original

Shared orig expression heatmap separated by species.

```{r c_base_heatmap_orig_all, fig.height = 6, fig.width = 5}

c_base_heatmap_orig_all <- ggplot2::ggplot(
  agg_orig_df_vis,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-7, -3.5, 0, 3.5, 7), 
    limits = c(-7, 7))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::facet_grid(cols = vars(agg_orig_df_vis$species))
c_base_heatmap_orig_all
```

```{r c_orig_hsc_heatmap, fig.height = 6, fig.width = 2.4}
c_orig_hsc_heatmap <- c_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.line = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
c_orig_hsc_heatmap
```

##### Combine

```{r c_combine_hsc, fig.height = 6.1, fig.width = 6.5}

c_combine_hsc <- ggpubr::ggarrange(
  c_custom_hsc_heatmap,
  c_orig_hsc_heatmap,
  nrow = 1,
  widths = c(3,2.1),
  align = c("h")
)

c_combine_hsc
```

##### Legend HSC

```{r c_combine_hsc_legend, fig.width = 3, fig.height = 3}
legend_plot_hsc <- c_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    legend.text = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.title = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.position = "bottom"
  )

legend_hsc <- ggpubr::get_legend(legend_plot_hsc)
c_combine_hsc_legend <- ggpubr::ggarrange(legend_hsc)

c_combine_hsc_legend
```

##### Export

```{r export_c_combine_hsc}
pdf(output_pdf_c_combined_hsc, width = 6.9, height = 6.1)
c_combine_hsc
dev.off()
```

```{r export_c_combine_hsc_legend}
pdf(output_pdf_c_combined_hsc_legend, width = 4, height = 1.2)
c_combine_hsc_legend
dev.off()
```

#### STR

##### Aggregate by type of clustering

```{r aggregate_orig_str}

sce_str_all_cells_orig <- sce_str_all_cells

# put info to be separated into one column
sce_str_all_cells_orig$cluster_species <- base::paste0(
  sce_str_all_cells_orig$celltypes,
  "_",
  sce_str_all_cells_orig$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_orig <- scuttle::aggregateAcrossCells(
  sce_str_all_cells_orig,
  id=colData(sce_str_all_cells_orig)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_orig_df <- logcounts(sce_agg_orig) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_orig$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_orig_df$clustering <- "original"

# separate cell type and species again
agg_orig_df <- agg_orig_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)

# factor species
agg_orig_df$species <- factor(
  agg_orig_df$species,
  levels = names(col_spc_pub))

head(agg_orig_df)
```

```{r aggregate_cust_str}

sce_str_all_cells_cust <- sce_str_all_cells

# put info to be separated into one column
sce_str_all_cells_cust$cluster_species <- base::paste0(
  sce_str_all_cells_cust$custom_label,
  "_",
  sce_str_all_cells_cust$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_cust <- scuttle::aggregateAcrossCells(
  sce_str_all_cells_cust,
  id=colData(sce_str_all_cells_cust)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_cust_df <- logcounts(sce_agg_cust) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_cust$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_cust_df$clustering <- "custom"

# separate cell type and species again
agg_cust_df <- agg_cust_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)

head(agg_cust_df)
```

##### Subset to required genes and scale

```{r scale_function_str}
scale_expression <- function(expression_df){

  for(gene in unique(expression_df$genes)){
    expr_vector_g <- expression_df[
      expression_df$genes == gene,]$avg_logcounts

    # calculate z score per gene
    z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
      stats::sd(expr_vector_g)
    expression_df$zscore_gene[expression_df$genes == gene] <- z_score_g
  }
  return(expression_df)
}
```

```{r determine_genes_str}
# stem and MPP genes from the utils source .R script
marker_genes <- unlist(marker_genes_list_str_figs2)
```

```{r scale_orig_str}

# only keep genes for visualisation
agg_orig_df_genes <- agg_orig_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_orig_df_genes$genes <- factor(
  agg_orig_df_genes$genes,
  levels = rev(marker_genes))

agg_orig_df_genes$cluster <- factor(
  agg_orig_df_genes$cluster,
  levels = levels(sce_str_all_cells$celltypes))

# scale expression per gene using z-scores
agg_orig_df_genes$zscore_gene <- vector(
  length = nrow(agg_orig_df_genes), 
  mode = "numeric")

agg_orig_df_vis <- scale_expression(agg_orig_df_genes)
```

```{r scale_cust_str}

# only keep genes for visualisation
agg_cust_df_genes <- agg_cust_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_cust_df_genes$genes <- factor(
  agg_cust_df_genes$genes,
  levels = rev(marker_genes))

agg_cust_df_genes$cluster <- factor(
  agg_cust_df_genes$cluster,
  levels = levels(sce_str_all_cells$custom_label))

# scale expression per gene using z-scores
agg_cust_df_genes$zscore_gene <- vector(
  length = nrow(agg_cust_df_genes), 
  mode = "numeric")
```

```{r scale_cust_across_str}

# expression is scaled ACROSS ALL SPECIES TOGETHER
agg_cust_df_genes <- scale_expression(agg_cust_df_genes)

# then that object is split by species
agg_cust_df_mmus <- agg_cust_df_genes %>% 
  dplyr::filter(species == "BL6")
agg_cust_df_mcas <- agg_cust_df_genes %>% 
  dplyr::filter(species == "CAST")
agg_cust_df_mspr <- agg_cust_df_genes %>% 
  dplyr::filter(species == "SPRET")
agg_cust_df_mcar <- agg_cust_df_genes %>% 
  dplyr::filter(species == "CAROLI")
```

##### Subset to the required cell types

After scaling.

```{r subset_ct_keep_str}
agg_cust_df_mmus <- agg_cust_df_mmus[agg_cust_df_mmus$cluster %in% cts_keep_str,]
agg_cust_df_mcas <- agg_cust_df_mcas[agg_cust_df_mcas$cluster %in% cts_keep_str,]
agg_cust_df_mspr <- agg_cust_df_mspr[agg_cust_df_mspr$cluster %in% cts_keep_str,]
agg_cust_df_mcar <- agg_cust_df_mcar[agg_cust_df_mcar$cluster %in% cts_keep_str,]

agg_orig_df_vis <- agg_orig_df_vis[agg_orig_df_vis$cluster %in% cts_keep_str,]
```

### Plot

#### Custom species-specific

Four separate heatmaps per species-specific custom clustering.

```{r d_base_heatmap_cust_mmus, fig.height = 6, fig.width = 3}

d_base_heatmap_cust_mmus <- ggplot2::ggplot(
  agg_cust_df_mmus,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-9, -4.5, 0, 4.5, 9), 
    limits = c(-9, 9))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
d_base_heatmap_cust_mmus
```

```{r d_theme_heatmap_cust_mmus, fig.height = 6, fig.width = 2}
d_theme_heatmap_cust_mmus <- d_base_heatmap_cust_mmus+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    strip.background = element_blank(),
    axis.line = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
d_theme_heatmap_cust_mmus
```
```{r d_base_heatmap_cust_mcas, fig.height = 6, fig.width = 3}

d_base_heatmap_cust_mcas <- ggplot2::ggplot(
  agg_cust_df_mcas,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-9, -4.5, 0, 4.5, 9), 
    limits = c(-9, 9))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
d_base_heatmap_cust_mcas
```

```{r d_theme_heatmap_cust_mcas, fig.height = 6, fig.width = 2}
d_theme_heatmap_cust_mcas <- d_base_heatmap_cust_mcas+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
d_theme_heatmap_cust_mcas
```

```{r d_base_heatmap_cust_mspr, fig.height = 6, fig.width = 3}

d_base_heatmap_cust_mspr <- ggplot2::ggplot(
  agg_cust_df_mspr,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-9, -4.5, 0, 4.5, 9), 
    limits = c(-9, 9))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
d_base_heatmap_cust_mspr
```

```{r d_theme_heatmap_cust_mspr, fig.height = 10, fig.width = 3}
d_theme_heatmap_cust_mspr <- d_base_heatmap_cust_mspr+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
d_theme_heatmap_cust_mspr
```

```{r d_base_heatmap_cust_mcar, fig.height = 6, fig.width = 3}

d_base_heatmap_cust_mcar <- ggplot2::ggplot(
  agg_cust_df_mcar,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-9, -4.5, 0, 4.5, 9), 
    limits = c(-9, 9))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
d_base_heatmap_cust_mcar
```

```{r d_theme_heatmap_cust_mcar, fig.height = 6, fig.width = 2}
d_theme_heatmap_cust_mcar <- d_base_heatmap_cust_mcar+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
d_theme_heatmap_cust_mcar
```

Build the species-specific heatmaps together

```{r d_custom_heatmap, fig.height = 6.8, fig.width = 4}

d_custom_heatmap <- ggpubr::ggarrange(
  d_theme_heatmap_cust_mmus+ggplot2::theme(
    plot.margin = margin(0.1,0.1,0.3,0.1)),
  d_theme_heatmap_cust_mcas+ggplot2::theme(
    plot.margin = margin(0,0.1,0.3,0.1)),
  d_theme_heatmap_cust_mspr+ggplot2::theme(
    plot.margin = margin(0,0.1,0.3,0.1)),
  d_theme_heatmap_cust_mcar+ggplot2::theme(
    plot.margin = margin(0,0.1,0.1,0.1)),
  nrow = 1,
  widths = c(9, 5, 6, 5),
  align = c("h")
)

d_custom_heatmap

```

#### Original merged

A shared orig expression heatmap separated by species.

```{r d_base_heatmap_orig_all, fig.height = 6, fig.width = 5}

d_base_heatmap_orig_all <- ggplot2::ggplot(
  agg_orig_df_vis,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradientn(
    "z-score per gene\nacross all cell types", 
    colours = colors_z_score_ct,
    breaks = c(-9, -4.5, 0, 4.5, 9), 
    limits = c(-9, 9))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::facet_grid(cols = vars(agg_orig_df_vis$species))
d_base_heatmap_orig_all
```

```{r d_heatmap_orig, fig.height = 6.8, fig.width = 3.1}
d_heatmap_orig <- d_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 270,
      size = heatmap_text_size,
      hjust = 0,
      vjust = 0.5
    ),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.line = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
d_heatmap_orig
```

```{r d_combine_str, fig.height = 6.9, fig.width = 8}

d_combine_str <- ggpubr::ggarrange(
  d_custom_heatmap,
  d_heatmap_orig,
  nrow = 1,
  widths = c(3,2.3),
  align = c("h")
)

d_combine_str
```

##### Legend STR

```{r d_combine_str_legend, fig.width = 3, fig.height = 3}
legend_plot_str <- d_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    legend.text = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.title = element_text(
      size = heatmap_text_size,
      color = legend_text_color
    ),
    legend.position = "bottom"
  )

legend_str <- ggpubr::get_legend(legend_plot_str)
d_combine_str_legend <- ggpubr::ggarrange(legend_str)

d_combine_str_legend
```

##### Export

```{r output_pdf_d_combined_str}
pdf(output_pdf_d_combined_str, width = 8.2, height = 6.8)
d_combine_str
dev.off()
```

```{r output_pdf_d_combined_str_legend}
pdf(output_pdf_d_combined_str_legend, width = 1.2, height = 4)
d_combine_str_legend
dev.off()
```

```{r session_info}
utils::sessionInfo()
```
