---
title: "Supplementary Figure 3"
date: '2024-11-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggalluvial, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")

col_cts_hsc <- c(col_cts_hsc, "removed" = "grey70")
col_cts_str <- c(col_cts_str, "removed" = "grey70")
```

## Load and prepare objects

### Load objects with all cells

After the original louvain clustering of the merged dataset.

```{r load_sce_hsc_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_hsc_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_hsc_all_cells) <- list(
  counts = assays(sce_hsc_all_cells)$counts,
  logcounts = assays(sce_hsc_all_cells)$logcounts
)
reducedDims(sce_hsc_all_cells) <- list(
  UMAP = reducedDims(sce_hsc_all_cells)$UMAP
)

gc()
```

```{r load_sce_str_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_str_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_str_all_cells) <- list(
  counts = assays(sce_str_all_cells)$counts,
  logcounts = assays(sce_str_all_cells)$logcounts
)
reducedDims(sce_str_all_cells) <- list(
  UMAP = reducedDims(sce_str_all_cells)$UMAP
)

gc()
```

### Load the objects with the custom labels

After species-specific clustering.

```{r load_custom_cluster_labels_hsc}
# add all labels to the object that contains all cells

# load objects
sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label

rm(
  sce_mmus_hsc,
  sce_mcas_hsc,
  sce_mspr_hsc,
  sce_mcar_hsc)
gc()
```

```{r load_custom_cluster_labels_str}
# add all labels to the object that contains all cells

# load objects
sce_mmus_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds"))
sce_mcas_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds"))
sce_mspr_str <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds"))
sce_mcar_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds"))

sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label <- unfactor(sce_mmus_str$custom_label)
sce_mcas_str$custom_label <- unfactor(sce_mcas_str$custom_label)
sce_mspr_str$custom_label <- unfactor(sce_mspr_str$custom_label)
sce_mcar_str$custom_label <- unfactor(sce_mcar_str$custom_label)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label

gc()
```

### Load annotated objects

Final annotation after sub-clustering.

```{r load_anno_hsc}
sce_hsc_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_str.rds")

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

rm(sce_hsc_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]

print(sort(unique(sce_hsc_all_cells$custom_label)))
```

```{r load_anno_str}
sce_str_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_str.rds")

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

rm(sce_str_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]

print(sort(unique(sce_str_all_cells$custom_label)))
```

### Add factors

```{r prep_hsc}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

```{r prep_str}
# change species names to currently used abbreviations
sce_str_all_cells$species_pub <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mmus"] <- "BL6"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcas"] <- "CAST"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcar"] <- "CAROLI"

sce_str_all_cells$species_pub <- factor(
  sce_str_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label,
  levels = factors_custom_str
)
```

## Other prep

```{r pdf_output}
# define output paths for pdf export


```

```{r params}
# global params
source("../figures/determine_params.R")
```

# Build 

Containing all cells, including contamination.
Both annotations (original merged and custom species-specific).

## Make DFs for UMAPs

```{r prep_hsc_df1}
hsc_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_all_cells, type = "UMAP"))

hsc_umap$celltypes <- sce_hsc_all_cells$celltypes 
hsc_umap$custom_label <- sce_hsc_all_cells$custom_label 
hsc_umap$species <- sce_hsc_all_cells$species_pub 
```

```{r prep_str_df1}
str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str_all_cells, type = "UMAP"))

str_umap$celltypes <- sce_str_all_cells$celltypes 
str_umap$custom_label <- sce_str_all_cells$custom_label 
str_umap$species <- sce_str_all_cells$species_pub 

# add the remaining UMAP coordinates from each species-specific PCA/BC
str_umap$UMAP_spec1 <- vector(length = nrow(str_umap))
str_umap$UMAP_spec2 <- vector(length = nrow(str_umap))

str_umap$UMAP_spec1[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,2]
```

## Make DFs for Alluvial

### MMUS

```{r mat_mmus_str}
mat_mmus_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

mmus_df_str <- mat_mmus_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mmus_df_str$Var2 <- factor(
  mmus_df_str$Var2, 
  levels = rev(factors_custom_str))

mmus_df_str <- mmus_df_str %>%
  dplyr::mutate(
    Var2_c = fct_inorder(paste0(" ", Var2)))
```

## Make UMAPs

```{r a_base_plot_str_celltypes_mmus, fig.height = 3, fig.width = 3}
a_base_plot_str_celltypes_mmus <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r b_base_plot_str_custom_mmus, fig.height = 3, fig.width = 3}
b_base_plot_str_custom_mmus <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
b_base_plot_str_custom_mmus
```

```{r c_base_plot_str_custom_mmus, fig.height = 3, fig.width = 3}
c_base_plot_str_custom2_mmus <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(
    base::min(str_umap$UMAP_spec1)-1, 
    base::max(str_umap$UMAP_spec1)+1))+
  ggplot2::ylim(c(
    base::min(str_umap$UMAP_spec2)-1, 
    base::max(str_umap$UMAP_spec2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
c_base_plot_str_custom2_mmus
```

## Make Alluvials

```{r c_base_plot_str_celltypes_mmus, fig.height = 3, fig.width = 5}
c_base_plot_str_celltypes_mmus <- ggplot2::ggplot(
  data = mmus_df_str,
  aes(
    axis1 = Var1,
    axis2 = Var2_c, 
    y = Freq)
    )+
  ggalluvial::geom_alluvium(
    aes(fill = Var1),
    width = 0.05
    )+
  ggalluvial::geom_stratum(
    fill = "black",
    alpha = 0.2,
    color = "black",
    size = 0.2,
    width = 0.05
    )+
  ggrepel::geom_text_repel(
    stat = "stratum",
    aes(
      x = after_stat(x),
      label = after_stat(stratum),
      hjust = 0),
    nudge_x = -0.5,      
    direction = "y",         
    box.padding = 0.1,
    alpha = c( # make the y axis text "disappear" due to 0 alpha
      rep(1, length(levels(mmus_df_str$Var1))),
      rep(0, length(unique(unfactor(
        mmus_df_str[mmus_df_str$Freq != 0,]$Var2_c)))))     
    )+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_cts_str)
c_base_plot_str_celltypes_mmus
```












```{r a_base_plot_hsc_1}

a_base_plot_hsc_1 <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1)+1))+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_clust)
```

```{r a_base_plot_hsc_1_mecom}
names(mycolors_to_max) <- as.character(
  as.numeric(names(mycolors_to_max))*max(hsc_umap_mecom$expression))

a_base_plot_hsc_1_mecom <- ggplot2::ggplot(
  hsc_umap_mecom[hsc_umap_mecom$color_by %in% c("Stem/MPP1", "Stem/MPP2"),], 
  aes(x = V1,
      y = V2,
      color = expression))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1)+1))+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2)+1))+
  ggplot2::scale_color_gradientn("Mecom expression", colours = mycolors_to_max)
a_base_plot_hsc_1_mecom

legend_mecom <- cowplot::get_legend(a_base_plot_hsc_1_mecom+theme_all_supp)
legend_mecom <- ggpubr::ggarrange(legend_mecom)
legend_mecom
```

```{r a_base_plot_hsc_1_subcl}
a_base_plot_hsc_1_subcl <- ggplot2::ggplot(
  hsc_umap_subcl[hsc_umap_subcl$color_by %in% c("Early MPP", "HSC"),], 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1)+1))+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2)+1))+
  ggplot2::scale_color_manual(
    values = col_cts_hsc[names(col_cts_hsc) %in%  c("HSC", "Early MPP")], 
    na.value = "white")
```

```{r a_base_plot_hsc_1_final}
a_base_plot_hsc_1_final <- ggplot2::ggplot(
  hsc_umap_fin, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap_fin$X1)-1, base::max(hsc_umap_fin$X1)+1))+
  ggplot2::ylim(c(base::min(hsc_umap_fin$X2)-1, base::max(hsc_umap_fin$X2)+1))+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r b_base_plot_str_1}
b_base_plot_str_1 <- ggplot2::ggplot(
  str_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))
```

```{r b_base_plot_str_2}
b_base_plot_str_2 <- ggplot2::ggplot(
  str_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = c(col_cts_str, "removed" = "grey50"))
```

```{r a_theme_plot_hsc_spc, fig.width = 3, fig.height = 3}
a_theme_plot_hsc_spc <- a_base_plot_hsc_spc+
  cowplot::theme_nothing()
a_theme_plot_hsc_spc
```

```{r a_theme_plot_hsc_1, fig.width = 3, fig.height = 3}
a_theme_plot_hsc_1 <- a_base_plot_hsc_1+
  cowplot::theme_nothing()
a_theme_plot_hsc_1
```

```{r a_theme_plot_hsc_1_mecom, fig.width = 3, fig.height = 3}
a_theme_plot_hsc_1_mecom <- a_base_plot_hsc_1_mecom+
  cowplot::theme_nothing()
a_theme_plot_hsc_1_mecom
```

```{r a_theme_plot_hsc_1_subcl, fig.width = 3, fig.height = 3}
a_theme_plot_hsc_1_subcl <- a_base_plot_hsc_1_subcl+
  cowplot::theme_nothing()
a_theme_plot_hsc_1_subcl
```

```{r a_theme_plot_hsc_1_final, fig.width = 3, fig.height = 3}
a_theme_plot_hsc_1_final <- a_base_plot_hsc_1_final+
  cowplot::theme_nothing()
a_theme_plot_hsc_1_final
```

```{r b_theme_plot_str_1, fig.width = 3, fig.height = 3}
b_theme_plot_str_1 <- b_base_plot_str_1+
  cowplot::theme_nothing()
b_theme_plot_str_1
```

```{r b_theme_plot_str_2, fig.width = 3, fig.height = 3}
b_theme_plot_str_2 <- b_base_plot_str_2+
  cowplot::theme_nothing()
b_theme_plot_str_2
```

## Legends

```{r hsc_legends}

col_clust <- col_num[1:length(levels(hsc_umap$color_by))]
names(col_clust) <- levels(hsc_umap$color_by)

umap_legend_plot_hsc_1 <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = col_clust)

umap_legend_theme_plot_hsc_1 <- umap_legend_plot_hsc_1+
  theme_all_supp+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
umap_legend_hsc_1 <- ggpubr::get_legend(umap_legend_theme_plot_hsc_1)
umap_legend_hsc_1 <- ggpubr::ggarrange(umap_legend_hsc_1)

umap_legend_hsc_1



umap_legend_plot_hsc_2 <- ggplot2::ggplot(
  hsc_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = c(col_cts_hsc, "removed" = "grey50"))

umap_legend_theme_plot_hsc_2 <- umap_legend_plot_hsc_2+
  theme_all_supp+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
umap_legend_hsc_2 <- ggpubr::get_legend(umap_legend_theme_plot_hsc_2)
umap_legend_hsc_2 <- ggpubr::ggarrange(umap_legend_hsc_2)

umap_legend_hsc_2

```

```{r str_legends}

umap_legend_plot_str_1 <- ggplot2::ggplot(
  str_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  theme_all_supp+
  ggplot2::scale_color_discrete("")
  # ggplot2::scale_color_discrete(
  #   "",
  #   labels = c(levels(str_umap$color_by)[1:9],
  #              levels(str_umap$color_by)[10:21]))

umap_legend_theme_plot_str_1 <- umap_legend_plot_str_1+
  theme_all_supp+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
umap_legend_str_1 <- ggpubr::get_legend(umap_legend_theme_plot_str_1)
umap_legend_str_1 <- ggpubr::ggarrange(umap_legend_str_1)

umap_legend_str_1

umap_legend_plot_str_2 <- ggplot2::ggplot(
  str_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = c(col_cts_str, "removed" = "grey50"))

umap_legend_theme_plot_str_2 <- umap_legend_plot_str_2+
  theme_all_supp+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
umap_legend_str_2 <- ggpubr::get_legend(umap_legend_theme_plot_str_2)
umap_legend_str_2 <- ggpubr::ggarrange(umap_legend_str_2)

umap_legend_str_2
```

```{r assemble_a, fig.width = 16, fig.height = 4}

# export b together
b_plot <- ggpubr::ggarrange(
  b_theme_plot_str_1,
  umap_legend_str_1,
  b_theme_plot_str_2,
  umap_legend_str_2,
  ncol = 4,
  widths = c(1,1.8,1,1)
)

b_plot
```

```{r export_b}
pdf(output_pdf_b, width = 16, height = 4)
b_plot
dev.off()
```

```{r output_pdf_a_1}
pdf(output_pdf_a_1, width = 4, height = 4)
a_theme_plot_hsc_spc
dev.off()
```

```{r output_pdf_a_2}
pdf(output_pdf_a_2, width = 4, height = 4)
a_theme_plot_hsc_1
dev.off()
```

```{r output_pdf_a_3}
pdf(output_pdf_a_3, width = 4, height = 4)
a_theme_plot_hsc_1_mecom
dev.off()
```

```{r output_pdf_a_4}
pdf(output_pdf_a_4, width = 3, height = 3)
legend_mecom
dev.off()
```

```{r output_pdf_a_5}
pdf(output_pdf_a_5, width = 4, height = 4)
a_theme_plot_hsc_1_subcl
dev.off()
```

```{r output_pdf_a_6}
pdf(output_pdf_a_6, width = 4, height = 4)
a_theme_plot_hsc_1_final
dev.off()
```

```{r output_pdf_a_7}
pdf(output_pdf_a_7, width = 6, height = 5)
umap_legend_hsc_1
dev.off()
```

```{r output_pdf_a_8}
pdf(output_pdf_a_8, width = 3, height = 4)
umap_legend_hsc_2
dev.off()
```

# Sankey plots

## Prep HSC

```{r load_spsp_hsc}

reducedDim(sce_hsc) <- NULL
reducedDim(sce_str) <- NULL

sce_hsc_spsp_mmus <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mmus_hsc-02"))
reducedDim(sce_hsc_spsp_mmus) <- NULL
assays(sce_hsc_spsp_mmus) <- list()

sce_hsc_spsp_mcas <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcas_hsc-02"))
reducedDim(sce_hsc_spsp_mcas) <- NULL
assays(sce_hsc_spsp_mcas) <- list()

sce_hsc_spsp_mspr <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mspr_hsc-02"))
reducedDim(sce_hsc_spsp_mspr) <- NULL
assays(sce_hsc_spsp_mspr) <- list()

sce_hsc_spsp_mcar <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcar_hsc-02"))
reducedDim(sce_hsc_spsp_mcar) <- NULL
assays(sce_hsc_spsp_mcar) <- list()
```

```{r prepare_spsp}
sce_hsc_spsp_mmus$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mmus), colnames(sce_hsc_ann))])
sce_hsc_spsp_mcas$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mcas), colnames(sce_hsc_ann))])
sce_hsc_spsp_mspr$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mspr), colnames(sce_hsc_ann))])
sce_hsc_spsp_mcar$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mcar), colnames(sce_hsc_ann))])
```

```{r prepare_spsp_mmus}
hsc_spsp_mmus_df <- base::as.data.frame(
  table(sce_hsc_spsp_mmus$annotation_all,
       unfactor(sce_hsc_spsp_mmus$cluster_louvain)))

hsc_spsp_mmus_df$Var1 <- factor(
  hsc_spsp_mmus_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mmus_df$Var2 <- factor(
  hsc_spsp_mmus_df$Var2,
  levels = c("5", "2", "6", "8", "9", "4", "3", "7", "1"))
stopifnot(!is.na(hsc_spsp_mmus_df$Var2))
```

```{r prepare_spsp_mcas}
hsc_spsp_mcas_df <- base::as.data.frame(
  table(sce_hsc_spsp_mcas$annotation_all,
       unfactor(sce_hsc_spsp_mcas$cluster_louvain)))

hsc_spsp_mcas_df$Var1 <- factor(
  hsc_spsp_mcas_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mcas_df$Var2 <- factor(
  hsc_spsp_mcas_df$Var2,
  levels = c("1", "7", "8", "3", "2", "9", "6", "4", "10", "5"))
stopifnot(!is.na(hsc_spsp_mcas_df$Var2))
```

```{r prepare_spsp_mspr}
hsc_spsp_mspr_df <- base::as.data.frame(
  table(sce_hsc_spsp_mspr$annotation_all,
       unfactor(sce_hsc_spsp_mspr$cluster_louvain)))

hsc_spsp_mspr_df$Var1 <- factor(
  hsc_spsp_mspr_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mspr_df$Var2 <- factor(
  hsc_spsp_mspr_df$Var2,
  levels = c("3", "2", "5", "10", "11", "6", "7", "1", "8", "4", "9"))
stopifnot(!is.na(hsc_spsp_mspr_df$Var2))
```

```{r prepare_spsp_mcar}
hsc_spsp_mcar_df <- base::as.data.frame(
  table(sce_hsc_spsp_mcar$annotation_all,
       unfactor(sce_hsc_spsp_mcar$cluster_louvain)))

hsc_spsp_mcar_df$Var1 <- factor(
  hsc_spsp_mcar_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mcar_df$Var2 <- factor(
  hsc_spsp_mcar_df$Var2,
  levels = c("2", "1", "6", "8", "4", "3", "5", "9", "7", "10", "11"))
stopifnot(!is.na(hsc_spsp_mcar_df$Var2))
```

```{r mmus_hsc_base_plot, fig.height = 5, fig.width = 4}
mmus_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mmus_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mcas_hsc_base_plot, fig.height = 5, fig.width = 4}
mcas_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mcas_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mspr_hsc_base_plot, fig.height = 5, fig.width = 4}
mspr_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mspr_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mcar_hsc_base_plot, fig.height = 5, fig.width = 4}
mcar_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mcar_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)
```

```{r assemble_c_hsc, fig.width = 13, fig.height = 4}

c_hsc_plot <- ggpubr::ggarrange(
  mmus_hsc_base_plot,
  mcas_hsc_base_plot,
  mspr_hsc_base_plot,
  mcar_hsc_base_plot,
  nrow = 1
)
c_hsc_plot
```

## Prep Str

```{r load_spsp_str}

sce_str_spsp_mmus <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mmus_str-02"))
reducedDim(sce_str_spsp_mmus) <- NULL
assays(sce_str_spsp_mmus) <- list()

sce_str_spsp_mcas <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcas_str-02"))
reducedDim(sce_str_spsp_mcas) <- NULL
assays(sce_str_spsp_mcas) <- list()

sce_str_spsp_mspr <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mspr_str-02"))
reducedDim(sce_str_spsp_mspr) <- NULL
assays(sce_str_spsp_mspr) <- list()

sce_str_spsp_mcar <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcar_str-02"))
reducedDim(sce_str_spsp_mcar) <- NULL
assays(sce_str_spsp_mcar) <- list()
```


```{r prepare_spsp_str}
sce_str_spsp_mmus$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mmus), colnames(sce_str_ann))])
sce_str_spsp_mcas$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mcas), colnames(sce_str_ann))])
sce_str_spsp_mspr$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mspr), colnames(sce_str_ann))])
sce_str_spsp_mcar$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mcar), colnames(sce_str_ann))])
```

```{r prepare_spsp_mmus_str}
str_spsp_mmus_df <- base::as.data.frame(
  table(sce_str_spsp_mmus$annotation_all,
       unfactor(sce_str_spsp_mmus$cluster_louvain)))
base::table(sce_str_spsp_mmus$annotation_all,
            sce_str_spsp_mmus$cluster_louvain)

str_spsp_mmus_df$Var1 <- factor(
  str_spsp_mmus_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mmus_df$Var2 <- factor(
  str_spsp_mmus_df$Var2,
  levels = c("1", "6", "8", "11", "4", "3", "2", "5", "10", "9", "7", "12"))
stopifnot(!is.na(str_spsp_mmus_df$Var2))
# 7 and 12 are empty
```

```{r prepare_spsp_mcas_str}
str_spsp_mcas_df <- base::as.data.frame(
  table(sce_str_spsp_mcas$annotation_all,
       unfactor(sce_str_spsp_mcas$cluster_louvain)))

base::table(sce_str_spsp_mcas$annotation_all,
            sce_str_spsp_mcas$cluster_louvain)

str_spsp_mcas_df$Var1 <- factor(
  str_spsp_mcas_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mcas_df$Var2 <- factor(
 str_spsp_mcas_df$Var2,
 levels = c("1", "5", "6", "16", "9"))
stopifnot(!is.na(str_spsp_mmus_df$Var2))
# "11", "12", "14", "15", "7", "8", "9", "2", "3", "4" are empty
```

```{r prepare_spsp_mspr_str}
str_spsp_mspr_df <- base::as.data.frame(
  table(sce_str_spsp_mspr$annotation_all,
       unfactor(sce_str_spsp_mspr$cluster_louvain)))

base::table(sce_str_spsp_mspr$annotation_all,
            sce_str_spsp_mspr$cluster_louvain)

str_spsp_mspr_df$Var1 <- factor(
  str_spsp_mspr_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mspr_df$Var2 <- factor(
  str_spsp_mspr_df$Var2,
  levels = c("3", "4", "9", "10", "6", "2", "5", "7", "1", "8", "11", "13", "12", "14"))
stopifnot(!is.na(str_spsp_mspr_df$Var2))
```

```{r prepare_spsp_mcar_str}
str_spsp_mcar_df <- base::as.data.frame(
  table(sce_str_spsp_mcar$annotation_all,
       unfactor(sce_str_spsp_mcar$cluster_louvain)))

base::table(sce_str_spsp_mcar$annotation_all,
            sce_str_spsp_mcar$cluster_louvain)

str_spsp_mcar_df$Var1 <- factor(
  str_spsp_mcar_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mcar_df$Var2 <- factor(
  str_spsp_mcar_df$Var2,
  levels = c("4", "7", "8", "11", "9", "10", "12", "1", "2", "3", "5", "6"))
stopifnot(!is.na(str_spsp_mcar_df$Var2))
# 9, 10, 12, 1, 2, 3, 5, 6 are empty
```

```{r mmus_str_base_plot, fig.height = 5, fig.width = 4}
mmus_str_base_plot <- ggplot2::ggplot(data = str_spsp_mmus_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mcas_str_base_plot, fig.height = 5, fig.width = 4}
mcas_str_base_plot <- ggplot2::ggplot(data = str_spsp_mcas_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mspr_str_base_plot, fig.height = 5, fig.width = 4}
mspr_str_base_plot <- ggplot2::ggplot(data = str_spsp_mspr_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mcar_str_base_plot, fig.height = 5, fig.width = 4}
mcar_str_base_plot <- ggplot2::ggplot(data = str_spsp_mcar_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r assemble_c_str, fig.width = 13, fig.height = 4}

c_str_plot <- ggpubr::ggarrange(
  mmus_str_base_plot,
  mcas_str_base_plot,
  mspr_str_base_plot,
  mcar_str_base_plot,
  nrow = 1
)
c_str_plot
```

```{r,  fig.width = 13, fig.height = 8}
c_plot <- ggpubr::ggarrange(
  c_hsc_plot,
  c_str_plot,
  nrow = 2
)
c_plot
```

```{r}
pdf(output_pdf_c, width = 16, height = 4)
c_plot
dev.off()
```

# Species proportions

## Plot

```{r d_hsc_base_plot}

df_hsc <- as.data.frame(colData(sce_hsc_ann)[
  colnames(colData(sce_hsc_ann)) %in% c("Species_ID", "Age_ID", "celltypes")])
table_hsc <- table(sce_hsc_ann$celltypes, sce_hsc_ann$Species_ID)
table_hsc

df_hsc$Species_ID <- factor(df_hsc$Species_ID, levels = names(col_spc))
df_hsc$Age_ID <- factor(df_hsc$Age_ID, levels = names(col_age))

d_hsc_base_plot <- ggplot2::ggplot(
  df_hsc, 
  aes(x = Age_ID,
      fill = Species_ID))+
  ggplot2::geom_bar()+
  facet_grid(cols = vars(df_hsc$celltypes))
```

Except for mcar lymph, there are at least 97 cells per specues and cell types.

```{r d_hsc_theme_plot}
d_hsc_theme_plot <- d_hsc_base_plot+
  theme_all_supp+
  ggplot2::xlab("Age and cell types")+
  ggplot2::ylab("# cells")+
  ggplot2::scale_fill_manual("Species", values = col_spc)+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.text.x = element_text(
      size = axis_text_size,
      angle = 90,
      vjust = 0.5,
      hjust = 0
    ),
    strip.background = element_blank())
d_hsc_theme_plot
```

```{r d_str_base_plot}

df_str <- as.data.frame(colData(sce_str_ann)[
  colnames(colData(sce_str_ann)) %in% c("Species_ID", "Age_ID", "celltypes")])
table_str <- table(sce_str_ann$celltypes, sce_str_ann$Species_ID)
table_str

df_str$Species_ID <- factor(df_str$Species_ID, levels = names(col_spc))
df_str$Age_ID <- factor(df_str$Age_ID, levels = names(col_age))

d_str_base_plot <- ggplot2::ggplot(
  df_str, 
  aes(x = Age_ID,
      fill = Species_ID))+
  ggplot2::geom_bar()+
  facet_grid(cols = vars(df_str$celltypes))
```

```{r d_str_theme_plot}
d_str_theme_plot <- d_str_base_plot+
  theme_all_supp+
  ggplot2::xlab("Age and cell types")+
  ggplot2::ylab("# cells")+
  ggplot2::scale_fill_manual("Species", values = col_spc)+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.text.x = element_text(
      size = axis_text_size,
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.background = element_blank())
d_str_theme_plot
```

```{r assemble_d, fig.width = 17, fig.height = 6}
d_plot <- ggpubr::ggarrange(
  d_hsc_theme_plot+ggplot2::theme(legend.position = "none"),
  align = "hv",
  d_str_theme_plot,
  widths = c(1, 1.2)
)
d_plot
```

```{r export_g}
pdf(output_pdf_d, width = 17, height = 6)
d_plot
dev.off()
```

```{r sessioninfo}
sessionInfo()
```
