---
title: "Supplementary Figure X Reclusterings Overview"
date: '2025-12-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(Seurat, quietly = TRUE)

```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../figures/determine_params.R")
```

## Load and prepare objects

### Original

```{r load_orig}
# custom clustered datasets
seu_ts_bone_marrow_orig <- readRDS(base::paste0(
  base_path,
  "/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/ts_bone_marrow"))
```

### After manual custom clustering

```{r load_cust}
seu_ts_bone_marrow_custom <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_bone_marrow.Rds"))
```

### Reclustered with different signature genes

```{r load_recl}
seu_ts_bone_marrow_recl_list <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_ts_bone_marrow_list"))
```

### Blind annotation after signature reclustering

```{r load_blind}
seu_ts_bone_marrow_blind <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/ts_bone_marrow_both.Rds"))
```

### Tresholds

```{r load_threshold}
seu_ts_bone_marrow_by_p <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/05_thresholds/ts_bone_marrow_recl_by_p.rds"))

seu_ts_bone_marrow_by_t <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/05_thresholds/ts_bone_marrow_recl_by_t.rds"))
```

### After batch correction signature reclustered

```{r load_bc}
seu_ts_bone_marrow_bc <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/06_include_batch_correction/ts_bone_marrow.Rds"))
```


## Prep

get the selected object by using the resolution df with the recorded
chosen optimal resolution

```{r resolution_df}
resolution_df_path <- base::paste0(
  base_path,
  "/metadata/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(
  file = resolution_df_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")
```

```{r get_res}
res_ts_bone_marrow_sign <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "conserved_signature") %>%
  dplyr::pull()

res_ts_bone_marrow_mark <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "conserved_markers") %>%
  dplyr::pull()

res_ts_bone_marrow_mmms <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "mmusall_markers") %>%
  dplyr::pull()

res_ts_bone_marrow_rand <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "random_features") %>%
  dplyr::pull()
```

```{r get_selected}
seu_ts_bone_marrow_sign <- seu_ts_bone_marrow_recl_list$seu_sign[[res_ts_bone_marrow_sign]]
seu_ts_bone_marrow_mark <- seu_ts_bone_marrow_recl_list$seu_mark[[res_ts_bone_marrow_mark]]
seu_ts_bone_marrow_mmms <- seu_ts_bone_marrow_recl_list$seu_mmms[[res_ts_bone_marrow_mmms]]
seu_ts_bone_marrow_rand <- seu_ts_bone_marrow_recl_list$seu_rand[[res_ts_bone_marrow_rand]]
```

```{r add_clusters}
# add all the different cluster labels

seu_ts_bone_marrow <- seu_ts_bone_marrow_orig

# custom labels after custom expertise-based reclustering
seu_ts_bone_marrow$custom_label <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$custom_label <- seu_ts_bone_marrow_custom$custom_label[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_custom))]

# after reclustering with random genes (as many as there are signature genes)
seu_ts_bone_marrow$seurat_rand <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$seurat_rand <- seu_ts_bone_marrow_rand$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_rand))]

# after reclustering with BL6-only markers
seu_ts_bone_marrow$seurat_mmms <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$seurat_mmms <- seu_ts_bone_marrow_mmms$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_mmms))]

# after reclustering with conserved markers
seu_ts_bone_marrow$seurat_mark <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$seurat_mark <- seu_ts_bone_marrow_mark$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_mark))]

# after reclustering with signature genes
seu_ts_bone_marrow$seurat_sign <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$seurat_sign <- seu_ts_bone_marrow_sign$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_sign))]

# after also adding a blind annotation of these same signature-based clusters
seu_ts_bone_marrow$sign_blind_labels <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$sign_blind_labels <- seu_ts_bone_marrow_blind$signature_label[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_blind))]

# after batch correction for signature clustering
seu_ts_bone_marrow$seurat_sign_bc <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$seurat_sign_bc <- seu_ts_bone_marrow_bc$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_bc))]

# after batch correction for signature clustering AND blind annotation
seu_ts_bone_marrow$sign_bc_labels <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$sign_bc_labels <- seu_ts_bone_marrow_bc$batchc_sign_labels[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_bc))]

for (treshold_p in names(seu_ts_bone_marrow_by_p)){
  
  seu_obj <- seu_ts_bone_marrow_by_p[[treshold_p]]

  colname <- base::paste0("by_p_", treshold_p)

  seu_ts_bone_marrow$new_col <- vector(length = ncol(seu_ts_bone_marrow))
  seu_ts_bone_marrow$new_col <- seu_obj$seurat_clusters[
    match(
      colnames(seu_ts_bone_marrow),
      colnames(seu_obj))]

  colnames(seu_ts_bone_marrow@meta.data)[
    colnames(seu_ts_bone_marrow@meta.data) == "new_col"] <- colname
}

for (treshold_t in names(seu_ts_bone_marrow_by_t)){
  
  seu_obj <- seu_ts_bone_marrow_by_t[[treshold_t]]

  colname <- base::paste0("by_t_", treshold_t)

  seu_ts_bone_marrow$new_col <- vector(length = ncol(seu_ts_bone_marrow))
  seu_ts_bone_marrow$new_col <- seu_obj$seurat_clusters[
    match(
      colnames(seu_ts_bone_marrow),
      colnames(seu_obj))]

  colnames(seu_ts_bone_marrow@meta.data)[
    colnames(seu_ts_bone_marrow@meta.data) == "new_col"] <- colname
}

colnames(seu_ts_bone_marrow@meta.data)

```

# Heatmaps

Between original cell types and custom clustering.

