---
title: "Supplementary Figure 3 HSPC heatmaps"
date: '2025-12-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

HSPC species-specific re-annotation.
Do the heatmaps here as they need different libraries.

Using the unified annotatiosn here.

#### Load objects

```{r set_export}
export <- FALSE
```

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggrepel, quietly = TRUE)
library(scuttle, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@params[["base_path"]]
#base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")
```

## Load and prepare objects

### Load objects with all cells

After the original louvain clustering of the merged dataset.

```{r load_sce_hsc_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_hsc_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_hsc_all_cells) <- list(
  counts = assays(sce_hsc_all_cells)$counts,
  logcounts = assays(sce_hsc_all_cells)$logcounts)
reducedDims(sce_hsc_all_cells) <- list(
  UMAP = reducedDims(sce_hsc_all_cells)$UMAP)

gc()
```

### Load the objects with the custom labels

After species-specific clustering.

```{r load_custom_cluster_labels_hsc}
# add all labels to the object that contains all cells

# load objects
sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label

gc()
```

### Load annotated objects

Final annotation after sub-clustering.

```{r load_anno_hsc}
sce_hsc_anno <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"))

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_hsc_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

rm(sce_hsc_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_hsc_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]

print(sort(unique(sce_hsc_all_cells$custom_label)))
```

### Add factors

```{r prep_hsc}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"

sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

```{r control_cols}
# make sure the color vector fits the label names 
# the factors must be right because they have been controlled in 02/01/custom
col_custom_hsc[!names(col_custom_hsc) %in% c(sce_hsc_all_cells$custom_label)]
sce_hsc_all_cells$custom_label[!sce_hsc_all_cells$custom_label %in% names(col_custom_hsc)]

stopifnot(names(col_custom_hsc) %in% c(sce_hsc_all_cells$custom_label))
stopifnot(c(sce_hsc_all_cells$custom_label) %in% names(col_custom_hsc))
```

## Other prep

```{r pdf_output}
# define output paths for pdf export
if(export){

}
```

```{r params}
# global params
source("../figures/determine_params.R")

heatmap_text_size <- 12

# keeping the stem compartment for comparisons
cts_keep_hsc <- c(
  "HSC",
  "Early MPP",
  "Activated MPP",
  "Early MPP cells 1",
  "Early MPP cells 2",
  "HSC/Early Mk",
  "MPP cells 1",
  "MPP cells 2",
  "Activated MPP cells 1",
  "Activated MPP cells 2",
  "Late MPP",
  "MPP cells"
)
# it looks like a lot but each species has different names

cts_keep_hsc[!cts_keep_hsc %in% factors_custom_hsc]
cts_keep_hsc[!cts_keep_hsc %in% names(col_cts_hsc)]

# keeping possible CAR and osteo populations
cts_keep_str <- c(
  "Adipo/CAR",
  "Osteo",
  "Adipo/Osteo CAR",
  "Early MPP cells",
  "HSC/Early Mk",
  "Activated MPP cells",
  "Late MPP",
  "MPP cells"
)

cts_keep_str[!cts_keep_str %in% factors_custom_str]
cts_keep_str[!cts_keep_str %in% names(col_cts_str)]
```

# Gene expression heatmaps

Absolute minimum number of genes required to show cell types.

## Prep

Make an appropriate dataframe.
I want to have gene expression scaled per cell type-average. 
For both the original merged and the custom species-specific clusterings.
Even though only some cell types are shown, I initially think I want to 
scale across all cell types - open for discussion. 
I think it's OK if we very clearly state it.

I also think it wouldn't be fair to compare otherwise, since the cells that are
part of the stem compartment in one annotation may not be exactly the same
as the cells in the stem compartment of the other annotation, so we'Re not 
comparing equal datasets.

I am using the logcounts based on the merged batch-normalisation for both the
orig clustering and the custom clusterings by using sce_hsc_all_cells as 
basis.
I think that is better comparable.
So by using the the merged batch-normalized logcounts, then even if a cell has 
two different labels in both, it has still the same expression levels.
That seems fair.
Although the logcounts are averaged and then scaled anyways.

### Aggregate by type of clustering

All cells from all cell types

```{r aggregate_orig}

sce_hsc_all_cells_orig <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_orig$cluster_species <- base::paste0(
  sce_hsc_all_cells_orig$celltypes,
  "_",
  sce_hsc_all_cells_orig$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_orig <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_orig,
  id=colData(sce_hsc_all_cells_orig)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_orig_df <- logcounts(sce_agg_orig) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_orig$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_orig_df$clustering <- "original"

# separate cell type and species again
agg_orig_df <- agg_orig_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)

# factor species
agg_orig_df$species <- factor(
  agg_orig_df$species,
  levels = names(col_spc_pub))
```

```{r aggregate_cust}

sce_hsc_all_cells_cust <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_cust$cluster_species <- base::paste0(
  sce_hsc_all_cells_cust$custom_label,
  "_",
  sce_hsc_all_cells_cust$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_cust <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_cust,
  id=colData(sce_hsc_all_cells_cust)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_cust_df <- logcounts(sce_agg_cust) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_cust$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_cust_df$clustering <- "custom"

# separate cell type and species again
agg_cust_df <- agg_cust_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)
```

### Subset to required genes and scale

Scale within each clustering.

```{r scale_function}
scale_expression <- function(expression_df){

  for(gene in unique(expression_df$genes)){
    expr_vector_g <- expression_df[
      expression_df$genes == gene,]$avg_logcounts

    # calculate z score per gene
    z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
      stats::sd(expr_vector_g)
    expression_df$zscore_gene[expression_df$genes == gene] <- z_score_g
  }
  return(expression_df)
}
```

```{r determine_genes}
# stem and MPP genes from the utils source .R script
marker_genes <- unlist(marker_genes_list_hsc[
  names(marker_genes_list_hsc) %in% c(
  "HSC",
  "MPP",
  "CC entry/activation"
  )])
```

```{r scale_orig}

# only keep genes for visualisation
agg_orig_df_genes <- agg_orig_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_orig_df_genes$genes <- factor(
  agg_orig_df_genes$genes,
  levels = rev(marker_genes))

agg_orig_df_genes$cluster <- factor(
  agg_orig_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$celltypes))

# scale expression per gene using z-scores
agg_orig_df_genes$zscore_gene <- vector(
  length = nrow(agg_orig_df_genes), 
  mode = "numeric")

agg_orig_df_vis <- scale_expression(agg_orig_df_genes)
```

```{r scale_cust}

# only keep genes for visualisation
agg_cust_df_genes <- agg_cust_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_cust_df_genes$genes <- factor(
  agg_cust_df_genes$genes,
  levels = rev(marker_genes))

agg_cust_df_genes$cluster <- factor(
  agg_cust_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$custom_label))

# scale expression per gene using z-scores
agg_cust_df_genes$zscore_gene <- vector(
  length = nrow(agg_cust_df_genes), 
  mode = "numeric")

# but scale WITHIN each species!
agg_cust_df_mmus <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "BL6"))
agg_cust_df_mcas <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "CAST"))
agg_cust_df_mspr <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "SPRET"))
agg_cust_df_mcar <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "CAROLI"))

# will not put back together because they contain different cell types
```

### Subset to the required cell types

After scaling.

```{r subset_ct_keep}
agg_cust_df_mmus <- agg_cust_df_mmus[agg_cust_df_mmus$cluster %in% cts_keep_hsc,]
agg_cust_df_mcas <- agg_cust_df_mcas[agg_cust_df_mcas$cluster %in% cts_keep_hsc,]
agg_cust_df_mspr <- agg_cust_df_mspr[agg_cust_df_mspr$cluster %in% cts_keep_hsc,]
agg_cust_df_mcar <- agg_cust_df_mcar[agg_cust_df_mcar$cluster %in% cts_keep_hsc,]

agg_orig_df_vis <- agg_orig_df_vis[agg_orig_df_vis$cluster %in% cts_keep_hsc,]
```

## Plot

First a shared orig expression heatmap separated by species.

```{r e_base_heatmap_orig_all, fig.height = 6, fig.width = 5}

e_base_heatmap_orig_all <- ggplot2::ggplot(
  agg_orig_df_vis,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::facet_grid(cols = vars(agg_orig_df_vis$species))
e_base_heatmap_orig_all
```

```{r e_theme_heatmap_orig_all, fig.height = 6, fig.width = 5}
e_theme_heatmap_orig_all <- e_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5,
      face = "italic"
    ),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.line = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
e_theme_heatmap_orig_all
```

Now four separate heatmaps per species-specific custom clustering.

```{r f_base_heatmap_cust_mmus, fig.height = 6, fig.width = 3}

f_base_heatmap_cust_mmus <- ggplot2::ggplot(
  agg_cust_df_mmus,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
f_base_heatmap_cust_mmus
```

```{r f_theme_heatmap_cust_mmus, fig.height = 6, fig.width = 2}
f_theme_heatmap_cust_mmus <- f_base_heatmap_cust_mmus+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5,
      face = "italic"
    ),
    strip.background = element_blank(),
    axis.line = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
f_theme_heatmap_cust_mmus
```

```{r g_base_heatmap_cust_mcas, fig.height = 6, fig.width = 3}

g_base_heatmap_cust_mcas <- ggplot2::ggplot(
  agg_cust_df_mcas,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
g_base_heatmap_cust_mcas
```

```{r g_theme_heatmap_cust_mcas, fig.height = 6, fig.width = 2}
g_theme_heatmap_cust_mcas <- g_base_heatmap_cust_mcas+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
g_theme_heatmap_cust_mcas
```

```{r h_base_heatmap_cust_mspr, fig.height = 6, fig.width = 3}

h_base_heatmap_cust_mspr <- ggplot2::ggplot(
  agg_cust_df_mspr,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
h_base_heatmap_cust_mspr
```

```{r h_theme_heatmap_cust_mspr, fig.height = 6, fig.width = 2}
h_theme_heatmap_cust_mspr <- h_base_heatmap_cust_mspr+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
h_theme_heatmap_cust_mspr
```

```{r i_base_heatmap_cust_mcar, fig.height = 6, fig.width = 3}

i_base_heatmap_cust_mcar <- ggplot2::ggplot(
  agg_cust_df_mcar,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
i_base_heatmap_cust_mcar
```

```{r i_theme_heatmap_cust_mcar, fig.height = 6, fig.width = 2}
i_theme_heatmap_cust_mcar <- i_base_heatmap_cust_mcar+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
i_theme_heatmap_cust_mcar
```

# Build

Build the species-specific heatmaps together

```{r fghi_heatmap, fig.height = 6, fig.width = 5}

fghi_heatmap <- ggpubr::ggarrange(
  f_theme_heatmap_cust_mmus+ggplot2::theme(plot.margin = margin(0.1,0.1,0.4,0.1)),
  g_theme_heatmap_cust_mcas+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  h_theme_heatmap_cust_mspr+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  i_theme_heatmap_cust_mcar+ggplot2::theme(plot.margin = margin(0,0.1,0.1,0.1)),
  nrow = 1,
  widths = c(6, 7, 3, 2),
  align = c("h")
)

fghi_heatmap

```

# Legends