---
title: "Figure S3 known age-associated gene heatmaps"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd generates the Figure 2f-style stacked heatmaps for known age-associated genes (**Vwf**, **Clu**, **Selp**), showing mean expression across **species x age** conditions and **cell types**, with tiles outlined when the gene is a conserved signature gene for that cell type.

## Load packages

```{r load_packages, message=FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(tidyr, quietly = TRUE)
library(ggplot2, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(scales, quietly = TRUE)
```

## Paths and object loading

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r load_objects}
sce_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"))

geneset_list_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_hsc"))

colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")

source("../source/colors.R")

source("../figures/determine_params.R")

colors_z_score_ct <- c("steelblue3", "white", "red3")
```

## Stacked gene heatmaps across species/age and cell types

### Parameters

```{r params}
genes_to_plot <- c("Vwf", "Clu", "Selp")

order_condition <- c(
  "mmus yng","mmus old",
  "mcas yng","mcas old",
  "mspr yng","mspr old",
  "mcar yng","mcar old"
)
```

### Prepare data

```{r prep_data}
# Expression matrix for selected genes (cells x genes) from SCE logcounts
expr_mat <- t(as.matrix(SingleCellExperiment::logcounts(sce_hsc)[genes_to_plot, , drop = FALSE]))
expr_df  <- as.data.frame(expr_mat) %>%
  tibble::rownames_to_column("cell_id")

# Minimal metadata needed
meta_df <- as.data.frame(SummarizedExperiment::colData(sce_hsc)) %>%
  tibble::rownames_to_column("cell_id") %>%
  dplyr::select(cell_id, Species_ID, Age_ID, celltypes)

# Long format: one row per cell x gene
df_long <- expr_df %>%
  dplyr::left_join(meta_df, by = "cell_id") %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(genes_to_plot),
    names_to = "gene",
    values_to = "expr"
  )

# Aggregate: mean expression per gene x species x age x celltype
df_agg <- df_long %>%
  dplyr::group_by(gene, Species_ID, Age_ID, celltypes) %>%
  dplyr::summarize(mean_expr = mean(expr), .groups = "drop") %>%
  dplyr::mutate(
    signature = mapply(
      function(ct, g) g %in% geneset_list_hsc[[ct]][["conserved_signature"]],
      ct = celltypes, g = gene
    ),
    condition = base::paste(Species_ID, Age_ID),
    conditionrev = factor(condition, levels = base::rev(order_condition)),
    celltypes = factor(celltypes, levels = names(geneset_list_hsc))
  ) %>%
  dplyr::group_by(gene) %>%
  dplyr::mutate(z_score = as.numeric(base::scale(mean_expr))) %>%  # z-score per gene
  dplyr::ungroup()
```

### Plot helpers

```{r plot_helper}
one_gene_heatmap <- function(gene_name, show_y = FALSE, show_x = FALSE) {
  p <- ggplot2::ggplot(
    df_agg %>% dplyr::filter(gene == gene_name),
    ggplot2::aes(
      x = celltypes,
      y = conditionrev,
      fill = z_score,
      color = signature,
      size = signature
    )
  ) +
    ggplot2::geom_tile(width = 0.95, height = 0.95) +
    ggplot2::scale_color_manual(
      guide = "none",
      values = c(`FALSE` = scales::alpha("black", 0), `TRUE` = "grey50")
    ) +
    ggplot2::scale_size_manual(
      guide = "none",
      values = c(`FALSE` = 0, `TRUE` = 0.4)
    ) +
    ggplot2::scale_fill_gradientn(
      "z-score per gene",
      colours = colors_z_score_ct,
      breaks = c(-5, -2.5, 0, 2.5, 5),
      limits = c(-5, 5),
      na.value = NA
    ) +
    ggplot2::labs(x = NULL, y = NULL, title = gene_name) +
    ggplot2::theme_classic()

  p <- p + theme_all

  # y-axis behavior
  p <- p +
    ggplot2::theme(
      ggplot2::element_text(hjust = 1),
      axis.title.y = ggplot2::element_blank()
    )

  # x-axis behavior
  if (show_x) {
    p <- p +
      ggplot2::theme(
        axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = ggplot2::element_blank(),
        axis.ticks.x = ggplot2::element_line()
      )
  } else {
    p <- p +
      ggplot2::theme(
        axis.text.x = ggplot2::element_blank(),
        axis.title.x = ggplot2::element_blank(),
        axis.ticks.x = ggplot2::element_blank()
      )
  }

  p +
    ggplot2::scale_y_discrete(expand = c(0, 0)) +
    ggplot2::scale_x_discrete(expand = c(0, 0))
}
```

### Generate stacked plot

```{r plot_stacked, fig.width=6, fig.height=10}
p1 <- one_gene_heatmap(genes_to_plot[1], show_x = FALSE) +
  ggplot2::theme(legend.position = "none")
p2 <- one_gene_heatmap(genes_to_plot[2], show_x = FALSE) +
  ggplot2::theme(legend.position = "none")
p3 <- one_gene_heatmap(genes_to_plot[3], show_x = TRUE) +
  ggplot2::theme(legend.position = "none")


heatmap_conds_stacked_plot <- ggpubr::ggarrange(
  p1,
  p2,
  p3,
  ncol = 1,
  heights = c(1, 1, 1.7)
)

heatmap_conds_stacked_plot
```

### Legend

```{r plot_legend, fig.width=3, fig.height=1}
legend_cond_plot <- one_gene_heatmap(genes_to_plot[1], show_y = TRUE, show_x = TRUE) +
  ggplot2::theme(legend.position = "bottom")

legend_cond <- ggpubr::get_legend(legend_cond_plot)
legend_cond_theme <- ggpubr::ggarrange(legend_cond)

legend_cond_theme
```

### Export to PDF

```{r export_paths}
output_pdf_heatmap_conds_stacked_plot <- base::paste0(
  base_path, "/manuscript1_rev/pdfs/figure_s3_known_age.pdf")
output_pdf_heatmap_conds_stacked_legend <- base::paste0(
  base_path, "/manuscript1_rev/pdfs/figure_s3_known_age_legend.pdf")
```

```{r export_pdf}
pdf(output_pdf_heatmap_conds_stacked_plot, width = 6, height = 10)
heatmap_conds_stacked_plot
dev.off()

pdf(output_pdf_heatmap_conds_stacked_legend, width = 3, height = 1)
legend_cond_theme
dev.off()
```
