---
title: "Supplemental Figure 5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate Supplemental Figure 5, corresponding for Figure 5.

This is a big supplementary figure supporting the blidn re-annotation of
human HSPC datasets and also contains info on the batch correction. 
Need to think carefully how to structure the text here so let's not do 
the figure toooo perfectly at first.

Instead of ordering by ABC I will order by type of plot as ABC order
may change later.

The corr_pval plot has instead been added to figure S4. 

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load_packages, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@params[["base_path"]]
#base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
#setwd("repositories/Interspecies_BM_phd/code/figures_rev")
```

```{r load_for_umaps_batches_o1}

dataset_o1 <- "ts_bone_marrow"

# batch corrected
seu_batchcorrected_o1 <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/06_include_batch_correction/ts_bone_marrow.Rds"))

# also load the object containing both the custom and the blind annotation
seu_both_o1 <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/ts_bone_marrow_both.Rds"))
```

```{r load_for_umaps_batches_o2}

dataset_o2 <- "ts_hscs_progenitors"

# batch corrected
seu_batchcorrected_o2 <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/06_include_batch_correction/ts_hscs_progenitors.Rds"))

# also load the object containing both the custom and the blind annotation
seu_blind_anno_o2 <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/ts_hscs_progenitors.Rds"))
```

```{r source}
colors_path <- base::paste0(
  base_path,
  "/metadata/colors/colors.txt")

source("../source/colors.R")
source("../source/plotting.R")
source("../figures/determine_params.R")

supplementary_text_size = 14
```

```{r export_path}

output_pdf_batch_umaps_o1 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_o1.pdf")
output_pdf_batch_umaps_legend1_o1 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_legend1_o1.pdf")
output_pdf_batch_umaps_legend2_o1 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_legend2_o1.pdf")

output_pdf_batch_umaps_o2 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_o2.pdf")
output_pdf_batch_umaps_legend1_o2 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_legend1_o2.pdf")
output_pdf_batch_umaps_legend2_o2 <- base::paste0(
  base_path, "/manuscript1_rev/figures5/batch_umaps_legend2_o2.pdf")
```

# Batch UMAPs

Find a good way to visualise the batches using UMAPs in relatively small space.

I will make more plots than I eventually need so I know which ones are most useful
for the figure.
Not all of them will fit.

```{r theme_umaps_supp}
theme_umaps_supp <- ggplot2::theme(
    legend.text = element_text(
      size = supplementary_text_size,
      color = axis_text_color,
      face = axis_text_face
    ),
    axis.title = element_text(
      size = supplementary_text_size,
      color = axis_text_color,
      face = axis_text_face
    ),
    strip.text = element_text(
      size = supplementary_text_size,
      color = axis_text_color,
      face = axis_text_face
    ),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    strip.background = element_blank(),
    plot.title = element_blank(),
    legend.key.size = unit(3, "mm")
  )
```

## Dataset 1

The UMAP coordinates will be taken from the object that was reclustered with 
signature genes and then re-annotated, because that is the one showing batch effect best.

```{r prep_umap_o1}
# add all annotations together in one object
# the one containing the signature-based umap coordinates
seu_all_o1 <- seu_both_o1

# this is the same as was originally called "annotated_reclustered" in Fig5
# but for clarity here, will call it signature label
# reclustered and annotated using signature genes
head(seu_all_o1$signature_label)
head(seu_all_o1$annotated_reclustered)
identical(seu_all_o1$annotated_reclustered, seu_all_o1$signature_label)

# these are the labels obtained after batch correction, then reclustering
# and annotation, all using only signature genes
seu_all_o1$batchc_sign_labels <- vector(length = ncol(seu_all_o1))
seu_all_o1$batchc_sign_labels <- seu_batchcorrected_o1$batchc_sign_labels[
  match(
    colnames(seu_all_o1),
    colnames(seu_batchcorrected_o1))]
head(seu_all_o1$batchc_sign_labels)

# also the original annotation
head(seu_all_o1$cell_type)
```

```{r factor_umap_o1}

seu_all_o1$assay_factored  <- unfactor(seu_all_o1$assay)

seu_all_o1$assay_factored[which(seu_all_o1$assay_factored == "10x 5' transcription profiling")] <- "10x 5'"

seu_all_o1$assay_factored <- factor(
  seu_all_o1$assay_factored,
  levels = c(
    "10x 3' v3",
    "10x 5'",
    "Smart-seq2"))
```

### Original cell type

```{r umap_batch_o1_cell_type_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o1_cell_type_assay <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "cell_type")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$assay_factored))+
  theme_umaps_supp
umap_batch_o1_cell_type_assay
```

```{r umap_batch_o1_cell_type_sex, fig.width = 6, fig.height = 2.5}
umap_batch_o1_cell_type_sex <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "cell_type")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$sex))+
  theme_umaps_supp
umap_batch_o1_cell_type_sex
```

### Custom clustering + anno

```{r umap_batch_o1_custom_label_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o1_custom_label_assay <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "custom_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$assay_factored))+
  theme_umaps_supp
umap_batch_o1_custom_label_assay
```

```{r a_umap_batch_o1_custom_label_sex, fig.width = 6, fig.height = 2.5}
a_umap_batch_o1_custom_label_sex <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "custom_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$sex))+
  theme_umaps_supp
a_umap_batch_o1_custom_label_sex
```

### Reclustering + anno with signature genes

```{r umap_batch_o1_signature_label_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o1_signature_label_assay <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "signature_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$assay_factored))+
  theme_umaps_supp
umap_batch_o1_signature_label_assay
```

```{r a_umap_batch_o1_signature_label_sex, fig.width = 6, fig.height = 2.5}
a_umap_batch_o1_signature_label_sex <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "signature_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$sex))+
  theme_umaps_supp
a_umap_batch_o1_signature_label_sex
```

### Batch correction + reclustering + anno with signature genes

```{r umap_batch_o1_batchc_sign_labels_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o1_batchc_sign_labels_assay <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$assay_factored))+
  theme_umaps_supp
umap_batch_o1_batchc_sign_labels_assay
```

```{r a_umap_batch_o1_batchc_sign_labels_sex, fig.width = 6, fig.height = 2.5}
a_umap_batch_o1_batchc_sign_labels_sex <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o1$sex))+
  theme_umaps_supp
a_umap_batch_o1_batchc_sign_labels_sex
```

```{r a_umap_all_together_o1, fig.width = 3, fig.height = 3}
a_umap_all_together_o1 <- Seurat::DimPlot(
  seu_all_o1, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  theme_umaps_supp+ 
  ggplot2::theme(legend.position = "none")
a_umap_all_together_o1
```

### Combine the selected plots

```{r, chosen_batch_umap_o1, fig.width = 6, fig.height = 6}

# col_anno_ts_bone_marrow from determine_params.R

chosen_umap_o1_a <- umap_batch_o1_signature_label_assay+
  ggplot2::scale_color_manual(
    "Reclustered and annotated\nwith signature genes",
    values = col_anno_ts_bone_marrow[names(col_anno_ts_bone_marrow) %in% 
      seu_all_o1$signature_label])
chosen_umap_o1_b <- umap_batch_o1_batchc_sign_labels_assay+
  ggplot2::scale_color_manual(
    "Batch corrected,\nreclustered and annotated\nwith signature genes",
    values = col_anno_ts_bone_marrow[names(col_anno_ts_bone_marrow) %in% 
      seu_all_o1$batchc_sign_labels])  

chosen_batch_umap_o1 <- ggpubr::ggarrange(
  chosen_umap_o1_a + ggplot2::theme(legend.position = "none"),
  chosen_umap_o1_b + ggplot2::theme(
    legend.position = "none",
    strip.text.x = element_blank(),
    axis.title = element_blank()),
  ncol = 1,
  align = "hv"
)
chosen_batch_umap_o1
```

```{r chosen_umap_o1_a_legend, fig.width = 3, fig.height = 3}

chosen_umap_o1_a_legend_theme <- ggpubr::get_legend(chosen_umap_o1_a)
chosen_umap_o1_a_legend <- ggpubr::ggarrange(chosen_umap_o1_a_legend_theme)
chosen_umap_o1_a_legend
```

```{r chosen_umap_o1_b_legend, fig.width = 3, fig.height = 3}

chosen_umap_o1_b_legend_theme <- ggpubr::get_legend(chosen_umap_o1_b)
chosen_umap_o1_b_legend <- ggpubr::ggarrange(chosen_umap_o1_b_legend_theme)
chosen_umap_o1_b_legend
```

#### Export

```{r output_pdf_batch_umaps_o1}
pdf(output_pdf_batch_umaps_o1, width = 6, height = 6)
chosen_batch_umap_o1
dev.off()
```

```{r output_pdf_batch_umaps_legend1_o1}
pdf(output_pdf_batch_umaps_legend1_o1, width = 4, height = 4)
chosen_umap_o1_a_legend
dev.off()
```

```{r output_pdf_batch_umaps_legend2_o1}
pdf(output_pdf_batch_umaps_legend2_o1, width = 4, height = 4)
chosen_umap_o1_b_legend
dev.off()
```

## Dataset 2

The UMAP coordinates will be taken from the object that was reclustered with 
signature genes and then re-annotated, because that is the one showing batch effect best.

```{r prep_umap_o2}
# add all annotations together in one object
# the one containing the signature-based umap coordinates
seu_all_o2 <- seu_blind_anno_o2

# this is the same as was originally called "annotated_reclustered" in Fig5
# but for clarity here, will call it signature label
# reclustered and annotated using signature genes
head(seu_all_o2$annotated_reclustered)
seu_all_o2$signature_label <- seu_all_o2$annotated_reclustered
identical(seu_all_o2$annotated_reclustered, seu_all_o2$signature_label)

# these are the labels obtained after batch correction, then reclustering
# and annotation, all using only signature genes
seu_all_o2$batchc_sign_labels <- vector(length = ncol(seu_all_o2))
seu_all_o2$batchc_sign_labels <- seu_batchcorrected_o2$batchc_sign_labels[
  match(
    colnames(seu_all_o2),
    colnames(seu_batchcorrected_o2))]
head(seu_all_o2$batchc_sign_labels)

# also the original annotation
head(seu_all_o2$cell_type)
```

```{r factor_umap_o2}

seu_all_o2$assay_factored  <- unfactor(seu_all_o2$assay)

seu_all_o2$assay_factored <- factor(
  seu_all_o2$assay_factored,
  levels = c(
    "10x 3' v2",
    "10x 5' v1"
    ))
```

### Original cell type

```{r umap_batch_o2_cell_type_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o2_cell_type_assay <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "cell_type")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$assay_factored))+
  theme_umaps_supp
umap_batch_o2_cell_type_assay
```

```{r umap_batch_o2_cell_type_sex, fig.width = 8, fig.height = 2.5}
umap_batch_o2_cell_type_sex <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "cell_type")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$sex))+
  theme_umaps_supp
umap_batch_o2_cell_type_sex
```

### Reclustering + anno with signature genes

```{r umap_batch_o2_signature_label_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o2_signature_label_assay <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "signature_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$assay_factored))+
  theme_umaps_supp
umap_batch_o2_signature_label_assay
```

```{r a_umap_batch_o2_signature_label_sex, fig.width = 6, fig.height = 2.5}
a_umap_batch_o2_signature_label_sex <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "signature_label")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$sex))+
  theme_umaps_supp
a_umap_batch_o2_signature_label_sex
```

### Batch correction + reclustering + anno with signature genes

```{r umap_batch_o2_batchc_sign_labels_assay, fig.width = 8.25, fig.height = 2.5}
umap_batch_o2_batchc_sign_labels_assay <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$assay_factored))+
  theme_umaps_supp
umap_batch_o2_batchc_sign_labels_assay
```

```{r a_umap_batch_o2_batchc_sign_labels_sex, fig.width = 6, fig.height = 2.5}
a_umap_batch_o2_batchc_sign_labels_sex <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  ggplot2::facet_wrap(
    nrow = 1,
    facets = vars(seu_all_o2$sex))+
  theme_umaps_supp
a_umap_batch_o2_batchc_sign_labels_sex
```

```{r a_umap_all_together_o2, fig.width = 3, fig.height = 3}
a_umap_all_together_o2 <- Seurat::DimPlot(
  seu_all_o2, 
  dims = c(1,2), 
  group.by = "batchc_sign_labels")+
  theme_umaps_supp+ 
  ggplot2::theme(legend.position = "none")
a_umap_all_together_o2
```

#### Selected

```{r, chosen_batch_umap_o2, fig.width = 5, fig.height = 6}

# col_anno_ts_bone_marrow from determine_params.R

chosen_umap_o2_a <- umap_batch_o2_signature_label_assay+
  ggplot2::scale_color_manual(
    "Reclustered and annotated\nwith signature genes",
    values = col_anno_ts_bone_marrow[names(col_anno_ts_bone_marrow) %in% 
      seu_all_o2$signature_label])
chosen_umap_o2_b <- umap_batch_o2_batchc_sign_labels_assay+
  ggplot2::scale_color_manual(
    "Batch corrected,\nreclustered and annotated\nwith signature genes",
    values = col_anno_ts_bone_marrow[names(col_anno_ts_bone_marrow) %in% 
      seu_all_o2$batchc_sign_labels])  

chosen_batch_umap_o2 <- ggpubr::ggarrange(
  chosen_umap_o2_a + ggplot2::theme(legend.position = "none"),
  chosen_umap_o2_b + ggplot2::theme(
    legend.position = "none",
    strip.text.x = element_blank(),
    axis.title = element_blank()),
  ncol = 1,
  align = "hv"
)
chosen_batch_umap_o2
```

```{r chosen_umap_o2_a_legend, fig.width = 3, fig.height = 5}

chosen_umap_o2_a_legend_theme <- ggpubr::get_legend(chosen_umap_o2_a)
chosen_umap_o2_a_legend <- ggpubr::ggarrange(chosen_umap_o2_a_legend_theme)
chosen_umap_o2_a_legend
```

```{r chosen_umap_o2_b_legend, fig.width = 3, fig.height = 5}

chosen_umap_o2_b_legend_theme <- ggpubr::get_legend(chosen_umap_o2_b)
chosen_umap_o2_b_legend <- ggpubr::ggarrange(chosen_umap_o2_b_legend_theme)
chosen_umap_o2_b_legend
```

```{r output_pdf_batch_umaps_o2}
pdf(output_pdf_batch_umaps_o2, width = 6, height = 6)
chosen_batch_umap_o2
dev.off()
```

```{r output_pdf_batch_umaps_legend1_o2}
pdf(output_pdf_batch_umaps_legend1_o2, width = 4, height = 4)
chosen_umap_o2_a_legend
dev.off()
```

```{r output_pdf_batch_umaps_legend2_o2}
pdf(output_pdf_batch_umaps_legend2_o2, width = 4, height = 4)
chosen_umap_o2_b_legend
dev.off()
```

###