---
title: "Supplementary Figure 3"
date: '2025-12-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

HSPC cell species-specific re-annotation.
Do the heatmaps here.

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggrepel, quietly = TRUE)
library(scuttle, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")

col_cts_hsc <- c(col_cts_hsc, "removed" = "grey70")
col_cts_hsc
```

## Load and prepare objects

### Load objects with all cells

After the original louvain clustering of the merged dataset.

```{r load_sce_hsc_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_hsc_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_hsc_all_cells) <- list(
  counts = assays(sce_hsc_all_cells)$counts,
  logcounts = assays(sce_hsc_all_cells)$logcounts
)
reducedDims(sce_hsc_all_cells) <- list(
  UMAP = reducedDims(sce_hsc_all_cells)$UMAP
)

gc()
```

### Load the objects with the custom labels

After species-specific clustering.

```{r load_custom_cluster_labels_hsc}
# add all labels to the object that contains all cells

# load objects
sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label

gc()
```

### Load annotated objects

Final annotation after sub-clustering.

```{r load_anno_hsc}
sce_hsc_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_hsc.rds")

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_hsc_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

rm(sce_hsc_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_hsc_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]

print(sort(unique(sce_hsc_all_cells$custom_label)))
```

### Add factors

```{r prep_hsc}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"

sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

```{r control_cols}
# make sure the color vector fits the label names 
# the factors must be right because they have been controlled in 02/01/custom
col_custom_hsc[!names(col_custom_hsc) %in% c(sce_hsc_all_cells$custom_label)]
sce_hsc_all_cells$custom_label[!sce_hsc_all_cells$custom_label %in% names(col_custom_hsc)]

stopifnot(names(col_custom_hsc) %in% c(sce_hsc_all_cells$custom_label))
stopifnot(c(sce_hsc_all_cells$custom_label) %in% names(col_custom_hsc))
```

## Other prep

```{r pdf_output}
# define output paths for pdf export


```

```{r params}
# global params
source("../figures/determine_params.R")

heatmap_text_size <- 12
```

# Gene expression heatmaps

Absolute minimum number of genes required to show cell types.

## Prep

Make an appropriate dataframe.
I want to have gene expression scaled per cell type-average. 
For both the original merged and the custom species-specific clusterings to
show that each annotated cluster actually contains cells expressing the
well-known markers genes.

I am using the logcounts based on the merged batch-normalisation for both the
orig clustering and the custom clusterings by using sce_hsc_all_cells as 
basis.
I think that is better comparable.
So even if a cell has two different labels in both, it has still the same 
expression levels no matter what.

### aggregate by type of clustering

```{r aggregate_orig}

sce_hsc_all_cells_orig <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_orig$cluster_species <- base::paste0(
  sce_hsc_all_cells_orig$celltypes,
  "_",
  sce_hsc_all_cells_orig$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_orig <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_orig,
  id=colData(sce_hsc_all_cells_orig)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_orig_df <- logcounts(sce_agg_orig) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_orig$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_orig_df$clustering <- "original"

# separate cell type and species again
agg_orig_df <- agg_orig_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)

# factor species
agg_orig_df$species <- factor(
  agg_orig_df$species,
  levels = names(col_spc_pub))
```

```{r aggregate_cust}

sce_hsc_all_cells_cust <- sce_hsc_all_cells

# put info to be separated into one column
sce_hsc_all_cells_cust$cluster_species <- base::paste0(
  sce_hsc_all_cells_cust$custom_label,
  "_",
  sce_hsc_all_cells_cust$species_pub  
)

# aggregate to get mean logcounts per species and cell type
sce_agg_cust <- scuttle::aggregateAcrossCells(
  sce_hsc_all_cells_cust,
  id=colData(sce_hsc_all_cells_cust)[,c("cluster_species")], 
  statistics = "mean", 
  use.assay.type = "logcounts") # batch-normalized (not corrected)

# turn into df formatted for visualisation
agg_cust_df <- logcounts(sce_agg_cust) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(sce_agg_cust$cluster_species),
    names_to = "cluster_species",
    values_to = "avg_logcounts")

# add info on which clustering these values are from
agg_cust_df$clustering <- "custom"

# separate cell type and species again
agg_cust_df <- agg_cust_df %>%
  tidyr::separate(
    col = cluster_species, 
    into = c("cluster", "species"),
    sep = "_",
    remove = FALSE)
```

### subset to required genes and scale

Scale within each clustering.

```{r scale_function}
scale_expression <- function(expression_df){

  for(gene in unique(expression_df$genes)){
    expr_vector_g <- expression_df[
      expression_df$genes == gene,]$avg_logcounts

    # calculate z score per gene
    z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
      stats::sd(expr_vector_g)
    expression_df$zscore_gene[expression_df$genes == gene] <- z_score_g
  }
  return(expression_df)
}
```

```{r determine_genes}
marker_genes <- best_marker_genes_hsc
```

```{r scale_orig}

# only keep genes for visualisation
agg_orig_df_genes <- agg_orig_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_orig_df_genes$genes <- factor(
  agg_orig_df_genes$genes,
  levels = rev(marker_genes))

agg_orig_df_genes$cluster <- factor(
  agg_orig_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$celltypes))

# scale expression per gene using z-scores
agg_orig_df_genes$zscore_gene <- vector(
  length = nrow(agg_orig_df_genes), 
  mode = "numeric")

agg_orig_df_vis <- scale_expression(agg_orig_df_genes)
```

```{r scale_cust}

# only keep genes for visualisation
agg_cust_df_genes <- agg_cust_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_cust_df_genes$genes <- factor(
  agg_cust_df_genes$genes,
  levels = rev(marker_genes))

agg_cust_df_genes$cluster <- factor(
  agg_cust_df_genes$cluster,
  levels = levels(sce_hsc_all_cells$custom_label))

# scale expression per gene using z-scores
agg_cust_df_genes$zscore_gene <- vector(
  length = nrow(agg_cust_df_genes), 
  mode = "numeric")

# but scale WITHIN each species!
agg_cust_df_mmus <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "BL6"))
agg_cust_df_mcas <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "CAST"))
agg_cust_df_mspr <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "SPRET"))
agg_cust_df_mcar <- scale_expression(agg_cust_df_genes %>% 
  dplyr::filter(species == "CAROLI"))

# will not put back together because they contain different cell types
```

## Plot

First a shared orig expression heatmap separated by species.

```{r e_base_heatmap_orig_all, fig.height = 10, fig.width = 16}

e_base_heatmap_orig_all <- ggplot2::ggplot(
  agg_orig_df_vis,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::facet_grid(cols = vars(agg_orig_df_vis$species))
e_base_heatmap_orig_all
```

```{r e_theme_heatmap_orig_all, fig.height = 11, fig.width = 16}
e_theme_heatmap_orig_all <- e_base_heatmap_orig_all+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5,
      face = "italic"
    ),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.line = element_blank(),
    axis.title = element_blank(),
    legend.position = "none")
e_theme_heatmap_orig_all
```

Now four separate heatmaps per species-specific custom clustering.

```{r f_base_heatmap_cust_mmus, fig.height = 10, fig.width = 4}

f_base_heatmap_cust_mmus <- ggplot2::ggplot(
  agg_cust_df_mmus,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
f_base_heatmap_cust_mmus
```

```{r f_theme_heatmap_cust_mmus, fig.height = 10, fig.width = 4}
f_theme_heatmap_cust_mmus <- f_base_heatmap_cust_mmus+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.text.y = element_text(
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5,
      face = "italic"
    ),
    strip.background = element_blank(),
    axis.line = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
f_theme_heatmap_cust_mmus
```
```{r g_base_heatmap_cust_mcas, fig.height = 10, fig.width = 5}

g_base_heatmap_cust_mcas <- ggplot2::ggplot(
  agg_cust_df_mcas,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
g_base_heatmap_cust_mcas
```

```{r g_theme_heatmap_cust_mcas, fig.height = 10, fig.width = 4}
g_theme_heatmap_cust_mcas <- g_base_heatmap_cust_mcas+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
g_theme_heatmap_cust_mcas
```

```{r h_base_heatmap_cust_mspr, fig.height = 10, fig.width = 4}

h_base_heatmap_cust_mspr <- ggplot2::ggplot(
  agg_cust_df_mspr,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
h_base_heatmap_cust_mspr
```

```{r h_theme_heatmap_cust_mspr, fig.height = 10, fig.width = 3}
h_theme_heatmap_cust_mspr <- h_base_heatmap_cust_mspr+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
h_theme_heatmap_cust_mspr
```

```{r i_base_heatmap_cust_mcar, fig.height = 10, fig.width = 5}

i_base_heatmap_cust_mcar <- ggplot2::ggplot(
  agg_cust_df_mcar,
  aes(
    x = cluster,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
i_base_heatmap_cust_mcar
```

```{r i_theme_heatmap_cust_mcar, fig.height = 10, fig.width = 5}
i_theme_heatmap_cust_mcar <- i_base_heatmap_cust_mcar+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      size = heatmap_text_size,
      hjust = 1,
      vjust = 0.5
    ),
    axis.ticks.y = element_blank(),
    axis.line = element_blank(),
    axis.text.y = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(
      size = heatmap_text_size,
      face = "bold"
    ),
    axis.title = element_blank(),
    legend.position = "none")
i_theme_heatmap_cust_mcar
```

# Build

Build the species-specific heatmaps together

```{r fghi_heatmap, fig.height = 11, fig.width = 16}

fghi_heatmap <- ggpubr::ggarrange(
  f_theme_heatmap_cust_mmus+ggplot2::theme(plot.margin = margin(0.1,0.1,0.4,0.1)),
  g_theme_heatmap_cust_mcas+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  h_theme_heatmap_cust_mspr+ggplot2::theme(plot.margin = margin(0,0.1,0.4,0.1)),
  i_theme_heatmap_cust_mcar+ggplot2::theme(plot.margin = margin(0,0.1,0.1,0.1)),
  nrow = 1,
  widths = c(3.8, 3.1, 3.5, 3.5),
  align = c("h")
)

fghi_heatmap

```

# Legends