---
title: "Supplementary Figure 2"
date: '2026-02-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is going to be a very big figure containing many subplots.
Trying to make it still relatively understandable.
This figure will contain data too big to run locally (for me).

## Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggalluvial, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")
```

### SCE objects

#### Objects with merged clustering, containing all cells

Because some cells were removed between this step and the final annotation.
So that we can show all cells, even contamination, these objects will 
serve as a basis and other annotations or reductions can be added. 
For instance, for the alluvial plots and the pre-removal umaps.

```{r load_all_cells_hsc}
sce_hsc_all_cells <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"
))
```

```{r load_all_cells_str}
sce_str_all_cells <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"
))
```

#### Finally annotated objects

These are needed for the heatmaps and for the final annotation names.

```{r load_anno_hsc}
sce_hsc_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"
))
```

```{r load_anno_str}
sce_str_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"
))
```

#### Species-specific custom annotations

These objects were annotated by species and fraction, so there are
eight objects in total.
These will be depicted individually.

```{r load_species_specific_hsc}

sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))
```

```{r load_species_specific_str}
sce_mmus_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds"))
sce_mcas_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds"))
sce_mspr_str <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds"))
sce_mcar_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds"))
```

## Params

```{r params}
# global params for figures specifically
# this is a supp figure so text size can be smaller (6-8 pts)
source("../figures/determine_params.R")

# local params
legend_text_size_allu <- 14
umap_point_size <- 0.4
```

## Prep 

### Randomize barcodes

#### HSC objects

```{r rand_barc_hsc}
# randomize the barcodes for nicer visualisation
dim(sce_hsc_all_cells)
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]
dim(sce_hsc_all_cells)
```

Also for the BL6 custom object

```{r rand_barc_hsc_anno}
# randomize the barcodes for nicer visualisation
dim(sce_mmus_hsc)
sce_mmus_hsc <- sce_mmus_hsc[,base::sample(
  x = c(1:ncol(sce_mmus_hsc)), 
  size = ncol(sce_mmus_hsc),
  replace = FALSE)]
dim(sce_mmus_hsc)
```

Also for the finally annotated object.

```{r rand_barc_mmus_hsc}
# randomize the barcodes for nicer visualisation
dim(sce_hsc_anno)
sce_hsc_anno <- sce_hsc_anno[,base::sample(
  x = c(1:ncol(sce_hsc_anno)), 
  size = ncol(sce_hsc_anno),
  replace = FALSE)]
dim(sce_hsc_anno)
```

#### STR objects

```{r rand_barc_str}
# randomize the barcodes for nicer visualisation
dim(sce_str_all_cells)
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]
dim(sce_str_all_cells)
```

### Final anno

Add the labels used for analysis and shown in main Figure 1.

```{r add_final_anno_hsc}

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_hsc_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_hsc_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"
```

```{r add_final_anno_str}

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"
```

### Custom labels

These are the labels from the species-specific custom clustering.

```{r add_custom_hsc}

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label
```

```{r add_custom_str}

sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label <- unfactor(sce_mmus_str$custom_label)
sce_mcas_str$custom_label <- unfactor(sce_mcas_str$custom_label)
sce_mspr_str$custom_label <- unfactor(sce_mspr_str$custom_label)
sce_mcar_str$custom_label <- unfactor(sce_mcar_str$custom_label)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label
```

### Factorize

#### HSC objects

```{r factors_hsc}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

Also need to do this for the example BL6 HSC fraction object.
For the schematic A.

```{r factors_mmus_hsc}
# change species names to currently used abbreviations
sce_mmus_hsc$species_pub <- vector(length = ncol(sce_mmus_hsc))
sce_mmus_hsc$species_pub[sce_mmus_hsc$Species_ID == "mmus"] <- "BL6"
sce_mmus_hsc$species_pub[sce_mmus_hsc$Species_ID == "mcas"] <- "CAST"
sce_mmus_hsc$species_pub[sce_mmus_hsc$Species_ID == "mspr"] <- "SPRET"
sce_mmus_hsc$species_pub[sce_mmus_hsc$Species_ID == "mcar"] <- "CAROLI"
sce_mmus_hsc$species_pub <- factor(
  sce_mmus_hsc$species_pub, 
  levels = names(col_spc_pub))
```

Also need to do this for the final merged anno HSC object.
For the schematic A.

```{r factors_hsc_anno}
# change species names to currently used abbreviations
sce_hsc_anno$species_pub <- vector(length = ncol(sce_hsc_anno))
sce_hsc_anno$species_pub[sce_hsc_anno$Species_ID == "mmus"] <- "BL6"
sce_hsc_anno$species_pub[sce_hsc_anno$Species_ID == "mcas"] <- "CAST"
sce_hsc_anno$species_pub[sce_hsc_anno$Species_ID == "mspr"] <- "SPRET"
sce_hsc_anno$species_pub[sce_hsc_anno$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_anno$species_pub <- factor(
  sce_hsc_anno$species_pub, 
  levels = names(col_spc_pub))
```

#### STR objects

```{r factors_str}
# change species names to currently used abbreviations
sce_str_all_cells$species_pub <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mmus"] <- "BL6"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcas"] <- "CAST"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_str_all_cells$species_pub <- factor(
  sce_str_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label,
  levels = factors_custom_str
)
```

# Plots

## Schematic A

This will be three different UMAPs, using the species-specific custom BL6 
HSPC fraction object as example.

### Prep DF

```{r a_umap_prep_dfs}

umap_df_mmus_hsc <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_mmus_hsc, type = "UMAP"))

# first show the species
umap_df_mmus_hsc$species_pub <- sce_mmus_hsc$species_pub 
# then the louvain clusterings
umap_df_mmus_hsc$cluster_louvain <- sce_mmus_hsc$cluster_louvain 
# then the annotated popuations
umap_df_mmus_hsc$custom_label <- sce_mmus_hsc$custom_label 

colnames(umap_df_mmus_hsc)[1] <- "UMAP1"
colnames(umap_df_mmus_hsc)[2] <- "UMAP2"
```

### Plot 

```{r a_umap_1, fig.width = 2, fig.height = 2}
a_umap_1 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = species_pub
  ))+
  geom_point(size = umap_point_size.8)+
  scale_color_manual(values = col_spc_pub)+
  theme_void()+
  theme(legend.position = "none")

a_umap_1
```

```{r a_umap_2, fig.width = 2, fig.height = 2}
a_umap_2 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = as.character(cluster_louvain)
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_num)+
  theme_void()+
  theme(legend.position = "none")

a_umap_2
```

```{r a_umap_3, fig.width = 2, fig.height = 2}
a_umap_3 <- ggplot2::ggplot(
  umap_df_mmus_hsc,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = custom_label
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_custom_hsc)+
  theme_void()+
  theme(legend.position = "none")

a_umap_3
```

## Schematic B

This is a schematic for the merged annotation including sub-clustering
based on several objects.
It will contain several umaps, as well.

### Prep DFs

```{r b_prep_mixed_umap}

# all cells, colored by species
# and also colored by louvain cluster
umap_df_sce_hsc_all_cells <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_all_cells, type = "UMAP"))

umap_df_sce_hsc_all_cells$species_pub <- sce_hsc_all_cells$species_pub 
umap_df_sce_hsc_all_cells$cluster_louvain <- sce_hsc_all_cells$cluster_louvain 

colnames(umap_df_sce_hsc_all_cells)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_all_cells)[2] <- "UMAP2"

# also only the HSC - MPP portion, colored by Mecom expression
sce_hsc_stem <- sce_hsc_all_cells[,sce_hsc_all_cells$cluster_louvain %in% c(2, 4)]
dim(sce_hsc_stem)

umap_df_sce_hsc_stem <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_stem, type = "UMAP"))

# logcounts have been batch-normalized
logcounts_stem <- SingleCellExperiment::logcounts(sce_hsc_stem)
umap_df_sce_hsc_stem$expression <- logcounts_stem["Mecom",]

colnames(umap_df_sce_hsc_stem)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_stem)[2] <- "UMAP2"
```

```{r b_prep_hsc_anno}

# final annotation, last umap of the plot
# contaminating cells have been removed
umap_df_sce_hsc_anno <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_anno, type = "UMAP"))
umap_df_sce_hsc_anno$celltypes <- sce_hsc_anno$celltypes 
umap_df_sce_hsc_anno$species_pub <- sce_hsc_anno$species_pub 

colnames(umap_df_sce_hsc_anno)[1] <- "UMAP1"
colnames(umap_df_sce_hsc_anno)[2] <- "UMAP2"
```

### Plots

```{r b_umap_1, fig.width = 2, fig.height = 2}
b_umap_1 <- ggplot2::ggplot(
  umap_df_sce_hsc_all_cells,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = species_pub
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_spc_pub)+
  theme_void()+
  theme(legend.position = "none")

b_umap_1
```

```{r b_umap_2, fig.width = 2, fig.height = 2}
b_umap_2 <- ggplot2::ggplot(
  umap_df_sce_hsc_all_cells,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = as.character(cluster_louvain)
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_num)+
  theme_void()+
  theme(legend.position = "none")

b_umap_2
```

```{r b_umap_3, fig.width = 2, fig.height = 2}
b_umap_3 <- ggplot2::ggplot(
  umap_df_sce_hsc_stem,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = expression
  ))+
  geom_point(size = 0.8)+
  scale_color_gradientn(colours = mycolors_to_max)+
  theme_void()+
  theme(legend.position = "none")

b_umap_3
```

```{r b_umap_4, fig.width = 2, fig.height = 2}
b_umap_4 <- ggplot2::ggplot(
  umap_df_sce_hsc_anno,
  aes(
    x = UMAP1,
    y = UMAP2,
    color = celltypes
  ))+
  geom_point(size = umap_point_size)+
  scale_color_manual(values = col_cts_hsc)+
  theme_void()+
  theme(legend.position = "none")

b_umap_4
```

