---
title: "Figure 5"
date: '2024-09-23'
output: html_document
---

Niche cell reclustering (ours, mouse, human).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate Figure 5: niche reclustering

#### Load objects

```{r seed, message = FALSE}
base::RNGkind("L'Ecuyer-CMRG")
set.seed(37)
```

```{r load_packages, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(Seurat, quietly = TRUE)
```

```{r define_basepath}
base_path <- snakemake@params[["base_path"]]
```

```{r load}
# list of gene sets for niche (conserved signature, conserved markers, etc.)
geneset_list_str <-base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_str"))

colors_path <- base::paste0(
  base_path,
  "/metadata/colors/colors.txt")

source("../source/colors.R")
source("../source/plotting.R")
```

```{r load_own}
#### OWN DATASET (Niche)

# annotated SCE
sce_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"))

# SCE with reclustering vectors for each gene set
sce_str_recl <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/03_recl/sce_str"))
```

```{r load_other2}
#### OTHER MOUSE DATASET 2 (Niche)
#### MUS_TIK_STROMAL
# number o2 is remnant from previous installment

dataset_o2 <- "mus_tik_stromal"

mus_tik_str_blind_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/mus_tik_stromal.Rds"
))
```

```{r load_other3}
#### OTHER HUMAN DATASET 3 (Niche)
#### li_all_stromal
# number o3 is remnant from previous installments

dataset_o3 <- "li_all_stromal"

# list of seurat objects with reclustering vectors for each gene set and resolution
li_all_str_blind_anno <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/li_all_stromal.Rds"))
```

```{r params}
# global params
source("../figures/determine_params.R")

colors_z_score_ct <- c("steelblue3", "white", "red3")
cts_exclude <- c("Lymphoid EC", "Chondrocyte")
print(cts_exclude)
```

```{r export_paths}
output_pdf_heatmap_expr <- base::paste0(
  base_path, "/manuscript1_rev/figure5/figure5_heatmap_expr.pdf")
output_pdf_heatmap_expr_legend <- base::paste0(
  base_path, "/manuscript1_rev/figure5/figure5_heatmap_expr_legend.pdf")
output_pdf_b_plot <- base::paste0(
  base_path, "/manuscript1_rev/figure5/figure5_b.pdf")
output_pdf_cd_plot <- base::paste0(
  base_path, "/manuscript1_rev/figure5/figure5_cd.pdf")
```

## Obtain data

```{r obtain_data}
# percentage of BL6-only markers that are not signature genes
BL6_genes <- vector()
signature_genes <- vector()
for(ct in names(geneset_list_str)){
  BL6_genes <- base::unique(
    c(BL6_genes,
      geneset_list_str[[ct]]$conserved_df$gene[
        !is.na(geneset_list_str[[ct]]$conserved_df$mmus)]))
  signature_genes <- base::unique(
    c(signature_genes, 
      geneset_list_str[[ct]]$conserved_signature))
}

perc <- (length(BL6_genes)-length(signature_genes[
  signature_genes %in% BL6_genes]))/length(BL6_genes)
print(perc)

# number of transcribed genes in total
print(length(which(rowSums(assays(sce_str)$counts_before_BC) > 0)))

# get all signature genes
conserved_signature_list <- lapply(geneset_list_str, function(geneset){
  conserved_signature <- geneset$conserved_signature
  return(conserved_signature)
})
signt <- base::unique(unlist(conserved_signature_list))
print(length(signt))

# get all conserved markers
conserved_markers_list <- lapply(geneset_list_str, function(geneset){
  conserved_markers <- geneset$conserved_markers
  return(conserved_markers)
})
consm <- base::unique(unlist(conserved_markers_list))
print(length(consm))

# get all BL6-only markers
mmus_marker_list <- lapply(geneset_list_str, function(geneset){
  mmus_markers <- geneset$conserved_df$gene[
    which(!is.na(geneset$conserved_df$mmus))]
  return(mmus_markers)
})
mmusm <- base::unique(unlist(mmus_marker_list))
print(length(mmusm))

# the number of each gene set without subclustering genes
sce_str_recl$cluster_signt_genes_used[1]
sce_str_recl$cluster_consm_genes_used[1]
sce_str_recl$cluster_mmusm_genes_used[1]
```

Find the genes in each cell type that are BL6 specific, for interest.
Nothing really jumps out though, at least to me. 

```{r print_ct_bl6specific}
str_cons_df_ct <- geneset_list_str$Osteo$conserved_df
str_mmus_specific <- str_cons_df_ct$gene[which(
  !is.na(str_cons_df_ct$mmus) &
    is.na(str_cons_df_ct$mcas) &
    is.na(str_cons_df_ct$mspr) &
    is.na(str_cons_df_ct$mcar))]
str_mmus_specific

str_cons_df_ct <- geneset_list_str$`Adipo/CAR`$conserved_df
str_mmus_specific <- str_cons_df_ct$gene[which(
  !is.na(str_cons_df_ct$mmus) &
    is.na(str_cons_df_ct$mcas) &
    is.na(str_cons_df_ct$mspr) &
    is.na(str_cons_df_ct$mcar))]
str_mmus_specific

str_cons_df_ct <- geneset_list_str$`Fibro/Chondro`$conserved_df
str_mmus_specific <- str_cons_df_ct$gene[which(
  !is.na(str_cons_df_ct$mmus) &
    is.na(str_cons_df_ct$mcas) &
    is.na(str_cons_df_ct$mspr) &
    is.na(str_cons_df_ct$mcar))]
str_mmus_specific

str_cons_df_ct <- geneset_list_str$Fibroblast$conserved_df
str_mmus_specific <- str_cons_df_ct$gene[which(
  !is.na(str_cons_df_ct$mmus) &
    is.na(str_cons_df_ct$mcas) &
    is.na(str_cons_df_ct$mspr) &
    is.na(str_cons_df_ct$mcar))]
str_mmus_specific

str_cons_df_ct <- geneset_list_str$`Peri/SMC`$conserved_df
str_mmus_specific <- str_cons_df_ct$gene[which(
  !is.na(str_cons_df_ct$mmus) &
    is.na(str_cons_df_ct$mcas) &
    is.na(str_cons_df_ct$mspr) &
    is.na(str_cons_df_ct$mcar))]
str_mmus_specific
```

## Expression heat map

Expression of conserved signature genes in all mouse species across HSPC
cell types (scaled by z-score per gene).
For Figure 5b.

### Prep

```{r a_prep1}
# get all unique signature genes
all_sign_genes <- vector()
for(i in 1:length(geneset_list_str)){
  sign_genes <- geneset_list_str[[i]]$conserved_signature
  all_sign_genes <- c(all_sign_genes, sign_genes)
}
all_sign_genes_str <- base::unique(all_sign_genes)
```

```{r a_prep2}
# aggregate per cell type and species
cts_str <- as.list(names(geneset_list_str))

sce_str$species_pub <- vector(length = ncol(sce_str))
sce_str$species_pub[sce_str$Species_ID == "mmus"] <- "BL6"
sce_str$species_pub[sce_str$Species_ID == "mcas"] <- "CAST"
sce_str$species_pub[sce_str$Species_ID == "mspr"] <- "SPRET"
sce_str$species_pub[sce_str$Species_ID == "mcar"] <- "CAROLI"
sce_str$species_pub <- factor(sce_str$species_pub, 
                              levels = names(col_spc_pub))
# remove cell types to be removed
cts_str <- cts_str[!cts_str %in% cts_exclude]
cts_list <- as.list(cts_str)

sce_str <- sce_str[,!sce_str$celltypes %in% cts_exclude]
sce_str$celltypes <- factor(sce_str$celltypes, levels(sce_str$celltypes)[
  !levels(sce_str$celltypes) %in% cts_exclude])

# separate SCE objects by cell type, then aggregate the counts to obtain
# the average counts per species for each cell type separately
# use the multi-batchnorm-corrected logcounts 

agg_str_list <- lapply(cts_str, function(ct){
  
  sce_ct <- sce_str[,sce_str$celltypes == ct]
  agg_str <- scuttle::aggregateAcrossCells(
    sce_ct, 
    id=colData(sce_ct)[,c("species_pub")], 
    statistics = "mean",
    use.assay.type = "logcounts") #multibatchnorm
  
  return(agg_str)
})
names(agg_str_list) <- names(geneset_list_str)
```

```{r a_prep3}
expr_df_list_str <- lapply(agg_str_list, function(agg_str){
  
  # get current cell type and relevant signature genes
  ct_curr <- agg_str$celltypes[1]
  
  sign_genes_ct <- geneset_list_str[[
    which(names(geneset_list_str) == ct_curr)]]$conserved_signature
  
  # get a df from average logcounts per ct and species
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str)$logcounts))
  
  # subset to all signature genes of this tissue
  vis_df <- vis_df[,colnames(vis_df) %in% all_sign_genes_str]
  
  # change df format 
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")
  
  # add ct info
  vis_df$cell_type <- base::rep(ct_curr)

  # add info on genes that are signature genes of this ct
  vis_df$is_sign_gene <- vector(length = nrow(vis_df))
  vis_df$is_sign_gene[vis_df$gene %in% sign_genes_ct] <- TRUE

  # scale expression of all genes per cell type 
  # for ordering of genes ONLY
  vis_df$scaled_expression_per_ct <- base::scale(vis_df$expression)
  vis_df <- vis_df[base::order(vis_df$scaled_expression_per_ct, 
                               decreasing = TRUE),]

  return(vis_df)
})

# bind dfs 
agg_str_df <- dplyr::bind_rows(expr_df_list_str)
```

```{r a_prep4}
# scale expression per gene using z-scores
agg_str_df$zscore_gene <- vector(length = nrow(agg_str_df), mode = "numeric")
for(g in base::unique(agg_str_df$gene)){
  
  expr_vector_g <- agg_str_df[agg_str_df$gene == g,]$expression

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_str_df$zscore_gene[agg_str_df$gene == g] <- z_score_g
}
```

```{r a_prep5}
# get a well ordered list of unique genes for factorization (after order above)
all_sign_genes_for_factor <- base::unique(agg_str_df$gene[
  agg_str_df$is_sign_gene == TRUE])

# factorize
agg_str_df$gene <- factor(agg_str_df$gene,
                          levels = all_sign_genes_for_factor)
agg_str_df$species <- factor(agg_str_df$species,
                             levels = base::rev(names(col_spc_pub)))

#print(levels(agg_str_df$gene))
```

```{r a_prep6}
interest_genes <- c("Cxcl12",
                    "Marcks",
                    "Kitl",
                    "Cd34",
                    "Efnb2",
                    "Mmp13",
                    "Tie1",
                    "Actn4",
                    "Lrg1")

agg_str_df$is_gene <- vector(length = nrow(agg_str_df))
agg_str_df$is_gene[agg_str_df$gene %in% interest_genes] <- "TRUE"
agg_str_df$is_gene <- as.logical(agg_str_df$is_gene)

print(levels(agg_str_df$gene)[levels(agg_str_df$gene) %in% interest_genes])
```

### Plot

```{r heatmap_expr_base_plot}
heatmap_expr_base_plot <- ggplot2::ggplot(
  agg_str_df,
  aes(y = species, x = gene, fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::facet_grid(rows = vars(agg_str_df$cell_type))
```

```{r heatmap_expr_theme_plot}
heatmap_expr_theme_plot <- heatmap_expr_base_plot+
  theme_all+
  ggplot2::theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    plot.title = element_text(
      vjust = 0.5,
      hjust = 0.5,
      color = col_cons_long["conserved identity signature"]),
    axis.title.y = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_text(
      angle = 0,
      vjust = 0.5,
      hjust = 1),
    strip.text.y = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color,
      angle = 0,
      vjust = 0.5,
      hjust = 0),
    strip.background.y = element_blank())+
  ggplot2::ggtitle("Signature genes")+
  ggplot2::scale_fill_gradientn(
    "z-score per gene", 
    colours = colors_z_score_ct,
    breaks = c(-5, -2.5,  0, 2.5, 5), 
    limits = c(-5,5))+
  ggplot2::scale_x_discrete(
    position =  "bottom",
    breaks = agg_str_df$gene[agg_str_df$is_gene])

heatmap_expr_legend <- ggpubr::get_legend(heatmap_expr_theme_plot)
heatmap_expr_legend <- ggpubr::ggarrange(heatmap_expr_legend)
```

### Assemble

```{r a_assemble, fig.width = 9, fig.height = 9}
heatmap_expr_plot <- ggpubr::ggarrange(
  heatmap_expr_theme_plot + theme(legend.position = "none",
                                  plot.title = element_blank())
)
heatmap_expr_plot
```

### Export

```{r a_plot_export}
pdf(output_pdf_heatmap_expr, width = 6.5, height = 4)
heatmap_expr_plot
dev.off()
```

```{r a_legend_export}
pdf(output_pdf_heatmap_expr_legend, width = 3, height = 3)
heatmap_expr_legend
dev.off()
```

## Re-clustering Heatmaps 

Figure 5c, Figure 5d, Figure 5e.

Heatmaps of own or published niche data, reclusterted using signature genes.

For other datasets, this includes blind re-annotation, see 03_sce_analysis/04_signatures/04_blind_annotation.

### Plots

#### Own

Figure 5c.

```{r mat_sign}
mat_sign <- base::table(
  sce_str_recl$celltypes, 
  sce_str_recl$cluster_signt)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign)){
  mat_sign[,i] <- mat_sign[,i]/base::sum(mat_sign[,i])
}
sign_df <- base::as.data.frame(mat_sign)

#manually factorise clusters for nice order
sign_df$Var2 <- factor(
  sign_df$Var2,
  levels = base::rev(c(1, 8, 2, 4, 5, 3, 6, 7)))
```

```{r sign_base_plot}
sign_base_plot <- ggplot2::ggplot(
  sign_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r sign_theme_plot, fig.width = 5, fig.height = 7}
sign_theme_plot <- sign_base_plot+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Reclustered")
sign_theme_plot
```

### Other 2

Figure 5d. 
Tikhonova et al mouse stromal cells. 

```{r mat_sign_o2}
mat_sign_o2 <- base::table(mus_tik_str_blind_anno$cell_type, 
                           mus_tik_str_blind_anno$annotated_reclustered)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign_o2)){
  mat_sign_o2[,i] <- mat_sign_o2[,i]/base::sum(mat_sign_o2[,i])
}
sign_df_o2 <- base::as.data.frame(mat_sign_o2)

sign_df_o2$Var2 <- unfactor(sign_df_o2$Var2)

sign_df_o2$Var2 <- factor(
  sign_df_o2$Var2,
  levels = base::rev(c(
    "Adipo/CAR1, pbs",
    "Adipo/CAR2, pbs",
    "Adipo/CAR3, non-pbs",
    "Osteo/CAR",
    "Osteo-Chondro",
    "Fibro-Chondro",
    "Chondro",
    "Osteo1",
    "Osteo2",
    "Osteo3",
    "Art/Cap-EC",
    "Ven EC1, non-pbs",
    "Ven EC2, pbs",
    "Ven EC3, non-pbs"
)))

sign_df_o2$Var1 <- unfactor(sign_df_o2$Var1)
sign_df_o2$Var1[sign_df_o2$Var1 == "P1"] <- "P1 adipogenesis1"
sign_df_o2$Var1[sign_df_o2$Var1 == "P2"] <- "P2 adipogenesis2"
sign_df_o2$Var1[sign_df_o2$Var1 == "P3"] <- "P3 osteo early"
sign_df_o2$Var1[sign_df_o2$Var1 == "P4"] <- "P4 osteo later"

sign_df_o2$Var1[sign_df_o2$Var1 == "O1"] <- "O1 osteo, Col16a"
sign_df_o2$Var1[sign_df_o2$Var1 == "O2"] <- "O2 osteo-chondro"
sign_df_o2$Var1[sign_df_o2$Var1 == "O3"] <- "O3 osteoblasts"

sign_df_o2$Var1[sign_df_o2$Var1 == "V1"] <- "V1 arterial"
sign_df_o2$Var1[sign_df_o2$Var1 == "V2"] <- "V2 sinusoidal" 

sign_df_o2$Var1 <- factor(
  sign_df_o2$Var1,
  levels = c(
    "P1 adipogenesis1",
    "P2 adipogenesis2",
    "P3 osteo early",
    "P4 osteo later",
    "O1 osteo, Col16a",
    "O2 osteo-chondro",
    "O3 osteoblasts",
    "V1 arterial",
    "V2 sinusoidal" 
))
```

```{r sign_base_plot_o2}
sign_base_plot_o2 <- ggplot2::ggplot(
  sign_df_o2, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r sign_theme_plot_o2, fig.height = 7, fig.width = 7}
sign_theme_plot_o2 <- sign_base_plot_o2+
  theme_all+
  ggplot2::scale_y_discrete(position = "right")+
  ggplot2::theme(
    legend.position = "none",
    axis.text.y = element_text(
      hjust = 0),
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::xlab("Original cell types")+
  ggplot2::ylab("Reclustered & Re-annotated")
sign_theme_plot_o2
```

### Other 3

Figure 5c.

```{r mat_sign_o3}
mat_sign_o3 <- base::table(li_all_str_blind_anno$cell_type, 
                           li_all_str_blind_anno$annotated_reclustered)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign_o3)){
  mat_sign_o3[,i] <- mat_sign_o3[,i]/base::sum(mat_sign_o3[,i])
}
sign_df_o3 <- base::as.data.frame(mat_sign_o3)

sign_df_o3$Var2 <- factor(
  sign_df_o3$Var2,
  levels = base::rev(c(
    "Fibro-like CAR1",
    "Fibro-like CAR2",
    "CAR",
    "Adipo-like CAR",
    "Osteo-Chondro",
    "EC1",
    "EC2",
    "EC3"
    )))

sign_df_o3$Var1 <- unfactor(sign_df_o3$Var1)
sign_df_o3$Var1[
  sign_df_o3$Var1 == "multipotent stromal stem cells"] <- "stromal stem cell"
sign_df_o3$Var1[
  sign_df_o3$Var1 == "balanced prog."] <- "balanced prog"
sign_df_o3$Var1[
  sign_df_o3$Var1 == "highly adipocytic gene-expressing progenitors"] <- "adipo-prog"
sign_df_o3$Var1[
  sign_df_o3$Var1 == "osteochondrogenic progenitors 1"] <- "osteo-chondro prog 1"
sign_df_o3$Var1[
  sign_df_o3$Var1 == "osteochondrogenic progenitors 2"] <- "osteo-chondro prog 2"

sign_df_o3$Var1 <- factor(
  sign_df_o3$Var1,
  levels = c(
    "stromal stem cell",
    "balanced prog",
    "adipo-prog",
    "pre-osteoblast",
    "osteo-chondro prog 1",
    "osteo-chondro prog 2",
    "pre-fibroblast 1",
    "pre-fibroblast 2",
    "pre-fibroblast 3",
    "endothelial"
    ))
```

```{r sign_base_plot_o3}
sign_base_plot_o3 <- ggplot2::ggplot(
  sign_df_o3, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r sign_theme_plot_o3, fig.width = 6, fig.height = 6}
sign_theme_plot_o3 <- sign_base_plot_o3+
  theme_all+
  ggplot2::scale_y_discrete(position = "right")+
  ggplot2::theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text.y = element_text(
      hjust = 1),
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5),
    axis.ticks.y = element_blank())+
   ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
sign_theme_plot_o3
```

### Assemble

```{r assemble_cd, fig.width = 15, fig.height =5.5}

cd_plots <- ggpubr::ggarrange(
  ncol = 2,
  align = "h",
  sign_theme_plot_o2,
  sign_theme_plot_o3,
  widths = c(1.3, 1.4))

cd_plots
```

### Export

```{r b_plot_export}
pdf(output_pdf_b_plot, width = 3, height = 5.2)
sign_theme_plot
dev.off()
```

```{r cd_plot_export}
pdf(output_pdf_cd_plot, width = 13, height = 7)
cd_plots
dev.off()
```

```{r session_info}
utils::sessionInfo()
```
