---
title: "Supplementary Figure X"
date: '2025-12-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Compare the custom_clustering of other datasets to their original 
cell type annotation.

These are ts_bone_marrow, ts_all_stromal, and mus_weinreb, and they were
selected for custom_clustering because they had relatively lower 
cell type resolution or clustering from the original publication because
many cell types had to be removed.

Also check how many signature genes (or orthologs) are detected in each
cluster's or cell types marker gene list to see if this is comparable.

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(Seurat, quietly = TRUE)

```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../figures/determine_params.R")
```

## Load and prepare objects

```{r load_objects}
# custom clustered datasets

seu_ts_bone_marrow <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_bone_marrow.Rds"))

seu_ts_all_stromal <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_all_stromal.Rds"))

seu_mus_weinreb_hspc <- readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/mus_weinreb_hspc.Rds"))
```

```{r load_recl}

seu_ts_bone_marrow_recl_list <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_ts_bone_marrow_list"))

seu_ts_all_stromal_recl_list <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_ts_all_stromal_list"))

seu_mus_weinreb_hspc_recl_list <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_mus_weinreb_hspc_list"))
```

## Prep

get the selected object by using the resolution df with the recorded
chosen optimal resolution

```{r resolution_df}
resolution_df_path <- base::paste0(
  base_path,
  "/metadata/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(
  file = resolution_df_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")
```

```{r get_res}
res_ts_bone_marrow <- resolution_df %>%
  dplyr::filter(dataset == "ts_bone_marrow") %>%
  dplyr::filter(conservation_level == "conserved_signature") %>%
  dplyr::pull()

res_ts_all_stromal <- resolution_df %>%
  dplyr::filter(dataset == "ts_all_stromal") %>%
  dplyr::filter(conservation_level == "conserved_signature") %>%
  dplyr::pull()

res_mus_weinreb_hspc <- resolution_df %>%
  dplyr::filter(dataset == "mus_weinreb_hspc") %>%
  dplyr::filter(conservation_level == "conserved_signature") %>%
  dplyr::pull()
```

```{r get_selected}
seu_ts_bone_marrow_sign <- seu_ts_bone_marrow_recl_list$seu_sign[[res_ts_bone_marrow]]
seu_ts_all_stromal_sign <- seu_ts_all_stromal_recl_list$seu_sign[[res_ts_all_stromal]]
seu_mus_weinreb_hspc_sign <- seu_mus_weinreb_hspc_recl_list$seu_sign[[res_mus_weinreb_hspc]]
```

```{r add_clusters}
# add the custom-labelled clusters

seu_ts_bone_marrow$signature_clust <- vector(length = ncol(seu_ts_bone_marrow))
seu_ts_bone_marrow$signature_clust <- seu_ts_bone_marrow_sign$seurat_clusters[
  match(
    colnames(seu_ts_bone_marrow),
    colnames(seu_ts_bone_marrow_sign))]

seu_ts_all_stromal$signature_clust <- vector(length = ncol(seu_ts_all_stromal))
seu_ts_all_stromal$signature_clust <- seu_ts_all_stromal_sign$seurat_clusters[
  match(
    colnames(seu_ts_all_stromal),
    colnames(seu_ts_all_stromal_sign))]

seu_mus_weinreb_hspc$signature_clust <- vector(length = ncol(seu_mus_weinreb_hspc))
seu_mus_weinreb_hspc$signature_clust <- seu_mus_weinreb_hspc_sign$seurat_clusters[
  match(
    colnames(seu_mus_weinreb_hspc),
    colnames(seu_mus_weinreb_hspc_sign))]
```

# Heatmaps

Between original cell types and custom clustering.

## ts_bone_marrow

```{r mat_ts_bone_marrow}

mat_ts_bone_marrow <- base::table(
  seu_ts_bone_marrow$cell_type, 
  seu_ts_bone_marrow$custom_label)

for(i in 1:ncol(mat_ts_bone_marrow)){
  mat_ts_bone_marrow[,i] <- mat_ts_bone_marrow[,i]/base::sum(mat_ts_bone_marrow[,i])
}
ts_bone_marrow_df <- base::as.data.frame(mat_ts_bone_marrow)

ts_bone_marrow_df$Var2 <- factor(
  ts_bone_marrow_df$Var2,
  levels = rev(levels(seu_ts_bone_marrow$custom_label)))
```

```{r base_heatmap_ts_bone_marrow}
base_heatmap_ts_bone_marrow <- ggplot2::ggplot(
  ts_bone_marrow_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_ts_bone_marrow
```

```{r theme_heatmap_ts_bone_marrow}
theme_heatmap_ts_bone_marrow <- base_heatmap_ts_bone_marrow+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Custom clustering and annotation")+
  ggplot2::xlab("Original annotation")
  
theme_heatmap_ts_bone_marrow
```

## ts_all_stromal

```{r mat_ts_all_stromal}

mat_ts_all_stromal <- base::table(
  seu_ts_all_stromal$cell_type, 
  seu_ts_all_stromal$custom_label)

for(i in 1:ncol(mat_ts_all_stromal)){
  mat_ts_all_stromal[,i] <- mat_ts_all_stromal[,i]/base::sum(mat_ts_all_stromal[,i])
}
ts_all_stromal_df <- base::as.data.frame(mat_ts_all_stromal)

ts_all_stromal_df$Var2 <- factor(
  ts_all_stromal_df$Var2,
  levels = rev(levels(seu_ts_all_stromal$custom_label)))
```

```{r base_heatmap_ts_all_stromal}
base_heatmap_ts_all_stromal <- ggplot2::ggplot(
  ts_all_stromal_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_ts_all_stromal
```

```{r theme_heatmap_ts_all_stromal}
theme_heatmap_ts_all_stromal <- base_heatmap_ts_all_stromal+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Custom clustering and annotation")+
  ggplot2::xlab("Original annotation")
  
theme_heatmap_ts_all_stromal
```

## mus_weinreb_hspc

```{r mat_ts_all_stromal}

mat_mus_weinreb_hspc <- base::table(
  seu_mus_weinreb_hspc$cell_type, 
  seu_mus_weinreb_hspc$custom_label)

for(i in 1:ncol(mat_mus_weinreb_hspc)){
  mat_mus_weinreb_hspc[,i] <- mat_mus_weinreb_hspc[,i]/base::sum(mat_mus_weinreb_hspc[,i])
}
mus_weinreb_hspc_df <- base::as.data.frame(mat_mus_weinreb_hspc)

mus_weinreb_hspc_df$Var2 <- factor(
  mus_weinreb_hspc_df$Var2,
  levels = rev(levels(seu_mus_weinreb_hspc$custom_label)))
```

```{r base_heatmap_weinreb_hspc_df}
base_heatmap_weinreb_hspc_df <- ggplot2::ggplot(
  mus_weinreb_hspc_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_weinreb_hspc_df
```

```{r theme_heatmap_weinreb_hspc_df}
theme_heatmap_weinreb_hspc_df <- base_heatmap_weinreb_hspc_df+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Custom clustering and annotation")+
  ggplot2::xlab("Original annotation")
  
theme_heatmap_weinreb_hspc_df
```

# Heatmaps vs signature clusters

## ts_bone_marrow_sign

```{r mat_ts_bone_marrow}

mat_ts_bone_marrow_sign <- base::table(
  seu_ts_bone_marrow$custom_label, 
  seu_ts_bone_marrow$signature_clust)

for(i in 1:ncol(mat_ts_bone_marrow_sign)){
  mat_ts_bone_marrow_sign[,i] <- mat_ts_bone_marrow_sign[,i]/base::sum(mat_ts_bone_marrow_sign[,i])
}
ts_bone_marrow_sign_df <- base::as.data.frame(mat_ts_bone_marrow_sign)

ts_bone_marrow_sign_df$Var1 <- factor(
  ts_bone_marrow_sign_df$Var1,
  levels = levels(seu_ts_bone_marrow$custom_label))

ts_bone_marrow_sign_df$Var2 <- factor(
  ts_bone_marrow_sign_df$Var2,
  levels = rev(c(
    "0",
    "4",
    "3",
    "2",
    "7",
    "6",
    "5",
    "1",
    "8"
  )))

```

```{r base_heatmap_ts_bone_marrow_sign}
base_heatmap_ts_bone_marrow_sign <- ggplot2::ggplot(
  ts_bone_marrow_sign_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_ts_bone_marrow_sign
```

```{r theme_heatmap_ts_bone_marrow_sign}
theme_heatmap_ts_bone_marrow_sign <- base_heatmap_ts_bone_marrow_sign+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Re-clustered")+
  ggplot2::xlab("Custom clustering and annotation")
  
theme_heatmap_ts_bone_marrow_sign
```

## ts_all_stromal

```{r mat_ts_all_stromal}

mat_ts_all_stromal_sign <- base::table(
  seu_ts_all_stromal$cell_type, 
  seu_ts_all_stromal$custom_label)

for(i in 1:ncol(mat_ts_all_stromal)){
  mat_ts_all_stromal[,i] <- mat_ts_all_stromal[,i]/base::sum(mat_ts_all_stromal[,i])
}
ts_all_stromal_df <- base::as.data.frame(mat_ts_all_stromal)

ts_all_stromal_df$Var2 <- factor(
  ts_all_stromal_df$Var2,
  levels = rev(levels(seu_ts_all_stromal$custom_label)))
```

```{r base_heatmap_ts_all_stromal}
base_heatmap_ts_all_stromal <- ggplot2::ggplot(
  ts_all_stromal_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_ts_all_stromal
```

```{r theme_heatmap_ts_all_stromal}
theme_heatmap_ts_all_stromal <- base_heatmap_ts_all_stromal+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Custom clustering and annotation")+
  ggplot2::xlab("Original annotation")
  
theme_heatmap_ts_all_stromal
```

## mus_weinreb_hspc

```{r mat_ts_all_stromal}

mat_mus_weinreb_hspc <- base::table(
  seu_mus_weinreb_hspc$cell_type, 
  seu_mus_weinreb_hspc$custom_label)

for(i in 1:ncol(mat_mus_weinreb_hspc)){
  mat_mus_weinreb_hspc[,i] <- mat_mus_weinreb_hspc[,i]/base::sum(mat_mus_weinreb_hspc[,i])
}
mus_weinreb_hspc_df <- base::as.data.frame(mat_mus_weinreb_hspc)

mus_weinreb_hspc_df$Var2 <- factor(
  mus_weinreb_hspc_df$Var2,
  levels = rev(levels(seu_mus_weinreb_hspc$custom_label)))
```

```{r base_heatmap_weinreb_hspc_df}
base_heatmap_weinreb_hspc_df <- ggplot2::ggplot(
  mus_weinreb_hspc_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
base_heatmap_weinreb_hspc_df
```

```{r theme_heatmap_weinreb_hspc_df}
theme_heatmap_weinreb_hspc_df <- base_heatmap_weinreb_hspc_df+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    #axis.title = element_blank(),
    #axis.text = element_blank(),
    axis.text.x = element_text(
      angle = 90
    ),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ylab("Custom clustering and annotation")+
  ggplot2::xlab("Original annotation")
  
theme_heatmap_weinreb_hspc_df
```


