---
title: "Supplementary Figure 2 Alluvial Portion stromal fraction"
date: '2025-12-02'
output: html_document
---

Making alluvial plots and umaps comparing the species-specific,
custom clustering with the merged annotation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggalluvial, quietly = TRUE)
library(ggrepel, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@params[["base_path"]]
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")

col_cts_str <- c(col_cts_str, "removed" = "grey70")
col_cts_str
```

## Load and prepare objects

### Load objects with all cells

After the original louvain clustering of the merged dataset.

```{r load_sce_str_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_str_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_str_all_cells) <- list(
  counts = assays(sce_str_all_cells)$counts,
  logcounts = assays(sce_str_all_cells)$logcounts
)
reducedDims(sce_str_all_cells) <- list(
  UMAP = reducedDims(sce_str_all_cells)$UMAP
)

gc()
```

### Load the objects with the custom labels

After species-specific clustering.

```{r load_custom_cluster_labels_str}
# add all labels to the object that contains all cells

# load objects
sce_mmus_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds"))
sce_mcas_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds"))
sce_mspr_str <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds"))
sce_mcar_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds"))

# using the unified labels now to save space but still name the slot
# custom label for downstream compatibility
sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label_unif <- unfactor(sce_mmus_str$custom_label_unif)
sce_mcas_str$custom_label_unif <- unfactor(sce_mcas_str$custom_label_unif)
sce_mspr_str$custom_label_unif <- unfactor(sce_mspr_str$custom_label_unif)
sce_mcar_str$custom_label_unif <- unfactor(sce_mcar_str$custom_label_unif)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), 
  colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label_unif
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), 
  colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label_unif
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), 
  colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label_unif
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), 
  colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label_unif

gc()
```

### Load annotated objects

Final annotation after sub-clustering.

```{r load_anno_str}
sce_str_anno <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"))

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

rm(sce_str_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are now called "removed"
levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]

print(sort(unique(sce_str_all_cells$custom_label)))
```

### Add factors

```{r prep_str}
# change species names to currently used abbreviations
sce_str_all_cells$species_pub <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mmus"] <- "BL6"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mcas"] <- "CAST"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_str_all_cells$species_pub[
  sce_str_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_str_all_cells$species_pub <- factor(
  sce_str_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label,
  levels = factors_custom_str_unif
)
```

```{r control_cols}
# make sure the color vector fits the label names 
# the factors must be right because they have been controlled in 02/01/custom
col_custom_str_unif[
  !names(col_custom_str_unif) %in% c(sce_str_all_cells$custom_label)]
unique(sce_str_all_cells$custom_label[!sce_str_all_cells$custom_label %in% 
  names(col_custom_str_unif)])

table(is.na(sce_str_all_cells$custom_label))

stopifnot(names(col_custom_str_unif) %in% c(sce_str_all_cells$custom_label))
stopifnot(c(sce_str_all_cells$custom_label) %in% names(col_custom_str_unif))
```

## Other prep

```{r pdf_output}
# define output paths for pdf export

output_pdf_allu_str_mmus <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_allu_str_mmus.pdf")
output_pdf_allu_str_mcas <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_allu_str_mcas.pdf")
output_pdf_allu_str_mspr <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_allu_str_mspr.pdf")
output_pdf_allu_str_mcar <- base::paste0(
  base_path,
  "/manuscript1_rev/figures2/figures2_allu_str_mcar.pdf")
```

```{r params}
# global params
source("../figures/determine_params.R")

# local params
legend_text_size_allu <- 14 # seems closest to 6pt font size at the end
geom_repel_size <- 4.9 # seems closest to 6pt font size at the end

# also adjust the width of the geoms for the alluvial plot to make it slim
gap <-  0.11 
w_alluvial <- 0.025
w_stratum <- 0.025
nudge <- -0.5
# decided to use the same params for all plots so that they align hopefully
```

# Base Plots 

Containing all cells, including contamination.
Both annotations (original merged and custom species-specific).

## Make DFs for UMAPs

```{r prep_str_df1}
str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str_all_cells, type = "UMAP"))

str_umap$celltypes <- sce_str_all_cells$celltypes 
str_umap$custom_label <- sce_str_all_cells$custom_label 
str_umap$species <- sce_str_all_cells$species_pub 

# add the remaining UMAP coordinates from each species-specific PCA/BC
str_umap$UMAP_spec1 <- vector(length = nrow(str_umap))
str_umap$UMAP_spec2 <- vector(length = nrow(str_umap))

str_umap$UMAP_spec1[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,2]
```

## Make DFs for Alluvial

### MMUS

```{r mat_mmus_str}
mat_mmus_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

mmus_df_str <- mat_mmus_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq)) %>%
  dplyr::rename(original_anno = Var1) %>%
  dplyr::rename(custom_anno = Var2)

mmus_df_str$custom_anno <- factor(
  mmus_df_str$custom_anno, 
  levels = factors_custom_str_unif)

mmus_df_str <- mmus_df_str %>%
  dplyr::mutate(
    original_anno_space = forcats::fct_inorder(paste0(" ", original_anno)))

# the  " " is added to avoid confusing ggalluvium with "Chondro" in one
# alluvium side and "Chondro" also in the other alluvium side
# by adding the space, the names and factors are not identical anymore
``` 

### MCAS

```{r mat_mcas_str}
mat_mcas_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcas"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcas"])

mcas_df_str <- mat_mcas_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))%>%
  dplyr::rename(original_anno = Var1) %>%
  dplyr::rename(custom_anno = Var2)

mcas_df_str$custom_anno <- factor(
  mmus_df_str$custom_anno, 
  levels = factors_custom_str_unif)

mcas_df_str <- mcas_df_str %>%
  dplyr::mutate(
    original_anno_space = forcats::fct_inorder(paste0(" ", original_anno)))
```

### MSPR

```{r mat_mspr_str}
mat_mspr_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mspr"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mspr"])

mspr_df_str <- mat_mspr_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))%>%
  dplyr::rename(original_anno = Var1) %>%
  dplyr::rename(custom_anno = Var2)

mspr_df_str$custom_anno <- factor(
  mmus_df_str$custom_anno, 
  levels = factors_custom_str_unif)

mspr_df_str <- mspr_df_str %>%
  dplyr::mutate(
    original_anno_space = forcats::fct_inorder(paste0(" ", original_anno)))
```

#### MCAR

```{r mat_mcar_str}
mat_mcar_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcar"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcar"])

mcar_df_str <- mat_mcar_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))%>%
  dplyr::rename(original_anno = Var1) %>%
  dplyr::rename(custom_anno = Var2)

mcar_df_str$custom_anno <- factor(
  mmus_df_str$custom_anno, 
  levels = factors_custom_str_unif)

mcar_df_str <- mcar_df_str %>%
  dplyr::mutate(
    original_anno_space = forcats::fct_inorder(paste0(" ", original_anno)))
```

## Make UMAPs

### One plot for the cell types for all later

But can use mmus

```{r base_plot_str_celltypes, fig.width = 4, fig.height = 3}
# this type of plot is not needed after all!
base_plot_str_celltypes <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
base_plot_str_celltypes
```

### MMUS

```{r base_plot_str_celltypes_mmus_2, fig.width = 4, fig.height = 3}
base_plot_str_celltypes_mmus_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mmus_2
```

```{r base_plot_str_celltypes_mmus_3, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mmus_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mmus_3
```

### MCAS

```{r base_plot_str_celltypes_mcas_2, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mcas_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAST",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mcas_2
```

```{r base_plot_str_celltypes_mcas_3, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mcas_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAST",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mcas_3
```

### MSPR

```{r base_plot_str_celltypes_mspr_2, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mspr_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "SPRET",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mspr_2
```

```{r base_plot_str_celltypes_mspr_3, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mspr_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "SPRET",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mspr_3
```

### MCAR

```{r base_plot_str_celltypes_mcar_2, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mcar_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAROLI",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mcar_2
```

```{r base_plot_str_celltypes_mcar_3, fig.height = 3, fig.width = 4}
base_plot_str_celltypes_mcar_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAROLI",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str_unif)
base_plot_str_celltypes_mcar_3
```

## Make Alluvials

The colors of the boxes are actually later manually adjusted;
here, the boxes are empty.
A 100% opaque filling of the cell type color will be added 
manually in Affinity. This was, for me, the easiest option. 

### MMUS

```{r pre_plot_prep_mmus}
mmus_df_str2 <- mmus_df_str %>%
  dplyr::mutate(flow_id = row_number())

# there are no overlapping names between custom anno and original_anno_space
# specifically because I added the space
table(is.na(match(
  mmus_df_str2$original_anno_space, 
  mmus_df_str2$custom_anno)))
table(is.na(match(
  mmus_df_str2$custom_anno, 
  mmus_df_str2$original_anno_space)))

table(is.na(mmus_df_str2$custom_anno))
unique(mmus_df_str2$custom_anno)

mmus_df_str2$fill_by <- mmus_df_str2$custom_anno
unique(mmus_df_str2$fill_by)

# convert wide to lodes form, this gives me more control
mmus_lodes <- ggalluvial::to_lodes_form(
  mmus_df_str2,
  axes = c("custom_anno", "original_anno_space"),
  id   = "flow_id") %>% 
  dplyr::select(!original_anno) 

# x = key variable, indicating to which axis a value corresponds,
# so each type of annotation
# stratum, = value taking by the axis variable indicated by x, 
# so each cell type
# alluvium = indexing scheme that links rows, in this case the flow ID

ggalluvial::is_lodes_form(mmus_lodes, key = x, value = stratum, id = flow_id)

# manually set numeric x positions to bring the axes close together
mmus_lodes <- mmus_lodes %>%
  dplyr::mutate(x = if_else(x == "custom_anno", 1, 1 + gap))
```

```{r base_allu_str_mmus_new, fig.width = 4.7, fig.height = 2.5}

base_allu_str_mmus <- ggplot(
  mmus_lodes,
  aes(
    x = x,
    stratum  = stratum,
    alluvium = flow_id,
    y = Freq
    )
  )+
  ggalluvial::geom_alluvium(
    aes(fill = fill_by),
    width = w_alluvial
  )+
  ggalluvial::geom_stratum(
    color = "black",
    size  = 0.2,
    width = w_stratum
  )+
  ggrepel::geom_text_repel(
    data = mmus_lodes %>% filter(x == "1"),
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = geom_repel_size,
    color = axis_text_color,
    nudge_x = nudge,
    direction = "y",
    box.padding = 0.1,
    hjust = 1
  )+
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::scale_fill_manual(values = col_custom_str_unif) +
  ggplot2::scale_x_continuous(
    breaks = c(1, 1 + gap),
    labels = c("", ""),
    expand = c(0, 0)
  )+
  ggplot2::coord_cartesian(clip = "off")
base_allu_str_mmus
```

### MCAS

```{r pre_plot_prep_mcas}
mcas_df_str2 <- mcas_df_str %>%
  dplyr::mutate(flow_id = row_number())

# there are no overlapping names between custom anno and original_anno_space
# specifically because I added the space
table(is.na(match(
  mcas_df_str2$original_anno_space, 
  mcas_df_str2$custom_anno)))
table(is.na(match(
  mcas_df_str2$custom_anno, 
  mcas_df_str2$original_anno_space)))

mcas_df_str2$fill_by <- mcas_df_str2$custom_anno

# convert wide to lodes form, this gives me more control
mcas_lodes <- ggalluvial::to_lodes_form(
  mcas_df_str2,
  axes = c("custom_anno", "original_anno_space"),
  id   = "flow_id") %>% 
  dplyr::select(!original_anno) 

# x = key variable, indicating to which axis a value corresponds,
# so each type of annotation
# stratum, = value taking by the axis variable indicated by x, 
# so each cell type
# alluvium = indexing scheme that links rows, in this case the flow ID

ggalluvial::is_lodes_form(mcas_lodes, key = x, value = stratum, id = flow_id)

# manually set numeric x positions to bring the axes close together
mcas_lodes <- mcas_lodes %>%
  dplyr::mutate(x = if_else(x == "custom_anno", 1, 1 + gap))
```

```{r base_allu_str_mcas, fig.width = 4.7, fig.height = 2.5}

base_allu_str_mcas <- ggplot(
  mcas_lodes,
  aes(
    x = x,
    stratum  = stratum,
    alluvium = flow_id,
    y = Freq
    )
  )+
  ggalluvial::geom_alluvium(
    aes(fill = fill_by),
    width = w_alluvial
  )+
  ggalluvial::geom_stratum(
    color = "black",
    size  = 0.2,
    width = w_stratum
  )+
  ggrepel::geom_text_repel(
    data = mcas_lodes %>% filter(x == "1"),
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = geom_repel_size,
    color = axis_text_color,
    nudge_x = nudge,
    direction = "y",
    box.padding = 0.1,
    hjust = 1
  )+
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::scale_fill_manual(values = col_custom_str_unif) +
  ggplot2::scale_x_continuous(
    breaks = c(1, 1 + gap),
    labels = c("", ""),
    expand = c(0, 0)
  )+
  ggplot2::coord_cartesian(clip = "off")
base_allu_str_mcas
```

### MSPR

```{r pre_plot_prep_mspr}
mspr_df_str2 <- mspr_df_str %>%
  dplyr::mutate(flow_id = row_number())

# there are no overlapping names between custom anno and original_anno_space
# specifically because I added the space
table(is.na(match(
  mspr_df_str2$original_anno_space,
  mspr_df_str2$custom_anno)))
table(is.na(match(
  mspr_df_str2$custom_anno, 
  mspr_df_str2$original_anno_space)))

mspr_df_str2$fill_by <- mspr_df_str2$custom_anno

# convert wide to lodes form, this gives me more control
mspr_lodes <- ggalluvial::to_lodes_form(
  mspr_df_str2,
  axes = c("custom_anno", "original_anno_space"),
  id   = "flow_id") %>% 
  dplyr::select(!original_anno) 

# x = key variable, indicating to which axis a value corresponds,
# so each type of annotation
# stratum, = value taking by the axis variable indicated by x, 
# so each cell type
# alluvium = indexing scheme that links rows, in this case the flow ID

ggalluvial::is_lodes_form(mspr_lodes, key = x, value = stratum, id = flow_id)

# manually set numeric x positions to bring the axes close together
mspr_lodes <- mspr_lodes %>%
  dplyr::mutate(x = if_else(x == "custom_anno", 1, 1 + gap))
```

```{r base_allu_str_mspr, fig.width = 4.7, fig.height = 2.5}

base_allu_str_mspr <- ggplot(
  mspr_lodes,
  aes(
    x = x,
    stratum  = stratum,
    alluvium = flow_id,
    y = Freq
    )
  )+
  ggalluvial::geom_alluvium(
    aes(fill = fill_by),
    width = w_alluvial
  )+
  ggalluvial::geom_stratum( 
    color = "black",
    size  = 0.2,
    width = w_stratum
  )+
  ggrepel::geom_text_repel(
    data = mspr_lodes %>% filter(x == "1"),
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = geom_repel_size,
    color = axis_text_color,
    nudge_x = nudge,
    direction = "y",
    box.padding = 0.1,
    hjust = 1
  )+
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::scale_fill_manual(values = col_custom_str_unif) +
  ggplot2::scale_x_continuous(
    breaks = c(1, 1 + gap),
    labels = c("", ""),
    expand = c(0, 0)
  )+
  ggplot2::coord_cartesian(clip = "off")
base_allu_str_mspr
```

### MCAR

```{r pre_plot_prep_mcar}
mcar_df_str2 <- mcar_df_str %>%
  dplyr::mutate(flow_id = row_number())

# there are no overlapping names between custom anno and original_anno_space
# specifically because I added the space
table(is.na(match(
  mcar_df_str2$original_anno_space,
  mcar_df_str2$custom_anno)))
table(is.na(match(
  mcar_df_str2$custom_anno, 
  mcar_df_str2$original_anno_space)))

mcar_df_str2$fill_by <- mcar_df_str2$custom_anno

# convert wide to lodes form, this gives me more control
mcar_lodes <- ggalluvial::to_lodes_form(
  mcar_df_str2,
  axes = c("custom_anno", "original_anno_space"),
  id   = "flow_id") %>% 
  dplyr::select(!original_anno) 

# x = key variable, indicating to which axis a value corresponds,
# so each type of annotation
# stratum, = value taking by the axis variable indicated by x, 
# so each cell type
# alluvium = indexing scheme that links rows, in this case the flow ID

ggalluvial::is_lodes_form(mcar_lodes, key = x, value = stratum, id = flow_id)

# manually set numeric x positions to bring the axes close together
mcar_lodes <- mcar_lodes %>%
  dplyr::mutate(x = if_else(x == "custom_anno", 1, 1 + gap))
```

```{r base_allu_str_mcar, fig.width = 4.7, fig.height = 2.5}

base_allu_str_mcar <- ggplot(
  mcar_lodes,
  aes(
    x = x,
    stratum  = stratum,
    alluvium = flow_id,
    y = Freq
    )
  )+
  ggalluvial::geom_alluvium(
    aes(fill = fill_by),
    width = w_alluvial
  )+
  ggalluvial::geom_stratum(  
    color = "black",
    size  = 0.2,
    width = w_stratum
  )+
  ggrepel::geom_text_repel(
    data = mcar_lodes %>% filter(x == "1"),
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = geom_repel_size,
    color = axis_text_color,
    nudge_x = nudge,
    direction = "y",
    box.padding = 0.1,
    hjust = 1
  )+
  ggplot2::theme_void() +
  ggplot2::theme(legend.position = "none") +
  ggplot2::scale_fill_manual(values = col_custom_str_unif) +
  ggplot2::scale_x_continuous(
    breaks = c(1, 1 + gap),
    labels = c("", ""),
    expand = c(0, 0)
  )+
  ggplot2::coord_cartesian(clip = "off")
base_allu_str_mcar
```

# Build

## List of cell types for all plots

```{r celltypes_legend, fig.width = 1, fig.height = 3} 

# this legend can be used for all species
celltypes_legend_plot <- base_plot_str_celltypes + 
  theme_all+
  ggplot2::geom_point(size = 3)+
  ggplot2::theme(
    legend.title = element_blank(),
    legend.text = element_text(
      face = legend_text_face,
      color = legend_text_color,
      size = legend_text_size_allu
    ),
    legend.spacing.y = unit(0.08, 'cm'),
    legend.key.size = unit(0.42, "cm"))

celltypes_legend_theme <- ggpubr::get_legend(celltypes_legend_plot)
celltypes_legend <- ggpubr::ggarrange(celltypes_legend_theme)

celltypes_legend
```

## A plot MMUS

UMAP in species-specific space
 
```{r a_mmus_plot_4, fig.width = 3, fig.height = 3} 

a_mmus_plot_4 <- base_plot_str_celltypes_mmus_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

a_mmus_plot_4
```

UMAP in merged space

```{r a_mmus_plot_3, fig.width = 3, fig.height = 3} 

a_mmus_plot_3 <- base_plot_str_celltypes_mmus_2+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

a_mmus_plot_3
```

alluvial plot

```{r a_mmus_plot_2, fig.width = 4, fig.height = 3} 

a_mmus_plot_2 <- base_allu_str_mmus+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

a_mmus_plot_2
```

```{r arrange_a, fig.width = 12.2, fig.height = 2.2}

plot_a <- ggpubr::ggarrange(
  a_mmus_plot_4+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  a_mmus_plot_3+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  a_mmus_plot_2,
  celltypes_legend,
  nrow = 1,
  widths = c(1.75, 1.75, 7.1, 1.6)
)

plot_a
```

```{r export_allu_str_mmus}
pdf(output_pdf_allu_str_mmus, width = 12.2, height = 2.2)
plot_a
dev.off()
```

## B plot MCAS

```{r b_mcas_plot_3, fig.width = 3, fig.height = 3} 

b_mcas_plot_3 <- base_plot_str_celltypes_mcas_2+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

b_mcas_plot_3
```

```{r b_mcas_plot_4, fig.width = 3, fig.height = 3} 

b_mcas_plot_4 <- base_plot_str_celltypes_mcas_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

b_mcas_plot_4
```

```{r b_mcas_plot_2, fig.width = 3, fig.height = 3} 

b_mcas_plot_2 <- base_allu_str_mcas+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

b_mcas_plot_2
```

```{r arrange_b, fig.width = 12.2, fig.height = 2.2}

plot_b <- ggpubr::ggarrange(
  b_mcas_plot_4+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  b_mcas_plot_3+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  b_mcas_plot_2,
  celltypes_legend,
  nrow = 1,
  widths = c(1.75, 1.75, 7.1, 1.6)
)

plot_b
```

```{r export_allu_str_mcas}
pdf(output_pdf_allu_str_mcas, width = 12.2, height = 2.2)
plot_b
dev.off()
```

## C plot MSPR

```{r c_mspr_plot_3, fig.width = 3, fig.height = 3} 

c_mspr_plot_3 <- base_plot_str_celltypes_mspr_2+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

c_mspr_plot_3
```

```{r c_mspr_plot_4, fig.width = 3, fig.height = 3} 

c_mspr_plot_4 <- base_plot_str_celltypes_mspr_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

c_mspr_plot_4
```

```{r c_mspr_plot_2, fig.width = 4, fig.height = 3} 

c_mspr_plot_2 <- base_allu_str_mspr+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

c_mspr_plot_2
```

```{r arrange_c, fig.width = 12.2, fig.height = 2.2}

plot_c <- ggpubr::ggarrange(
  c_mspr_plot_4+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  c_mspr_plot_3+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  c_mspr_plot_2,
  celltypes_legend,
  nrow = 1,
  widths = c(1.75, 1.75, 7.1, 1.6)
)

plot_c
```

```{r export_allu_str_mspr}
pdf(output_pdf_allu_str_mspr, width = 12.2, height = 2.2)
plot_c
dev.off()
```

## D plot MCAR

```{r d_mcar_plot_2, fig.width = 4, fig.height = 3} 

d_mcar_plot_2 <- base_allu_str_mcar+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

d_mcar_plot_2
```

```{r d_mcar_plot_3, fig.width = 3, fig.height = 3} 

d_mcar_plot_3 <- base_plot_str_celltypes_mcar_2+
  ggplot2::theme_void()+
  ggplot2::theme(
    legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

d_mcar_plot_3
```

```{r d_mcar_plot_4, fig.width = 3, fig.height = 3} 

d_mcar_plot_4 <- base_plot_str_celltypes_mcar_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none",
    panel.border = element_rect(
      fill = NA,
      color = "grey30",
      linewidth = 1,
      linetype = 1
    ))

d_mcar_plot_4
```

```{r arrange_d, fig.width = 12.2, fig.height = 2.2}

plot_d <- ggpubr::ggarrange(
  d_mcar_plot_4+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  d_mcar_plot_3+ggplot2::theme(
    plot.margin = margin(t = 0, r = 0, b = 20, l = 0, unit = "pt")),
  d_mcar_plot_2,
  celltypes_legend,
  nrow = 1,
  widths = c(1.75, 1.75, 7.1, 1.6)
)

plot_d
```

```{r export_allu_str_mcar}
pdf(output_pdf_allu_str_mcar, width = 12.2, height = 2.2)
plot_d
dev.off()
```

```{r session_info}
utils::sessionInfo()
```
