---
title: "Figure 1"
date: '2024-08-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate Figure 1 

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(scuttle, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@params[["base_path"]]
#base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
#setwd("repositories/Interspecies_BM_phd/code/figures_rev")
```

```{r load_sce}
# fully annotated SCE objects
sce_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"))
sce_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"))
```

```{r load_short}
# list containing genes in the correct order for the dotplot
order_list_dotplot <- base::paste0(
  base_path, 
  "/metadata/scRNAseq/02_sce_anno/dotplot_order.txt")

dotplot_order_df <- utils::read.csv(order_list_dotplot, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

# list of genes used for subclustering
# only for printing info not for making plots
gene_list_subclustering <- base::paste0(
  base_path, 
  "/metadata/scRNAseq/02_sce_anno/gene_list_subclustering.txt")

subclustering_df <- utils::read.csv(gene_list_subclustering, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

# color table
colors_path <- base::paste0(
  base_path,
  "/metadata/colors/colors.txt")
print(colors_path)
source("../source/colors.R")
print(col_spc_pub)
```

```{r prep_sces}

# randomize the barcodes for nicer visualisation
sce_hsc <- sce_hsc[,base::sample(
  x = c(1:ncol(sce_hsc)), 
  size = ncol(sce_hsc),
  replace = FALSE)]

sce_str <- sce_str[,base::sample(
  x = c(1:ncol(sce_str)), 
  size = ncol(sce_str),
  replace = FALSE)]

# change species names to currently used abbreviations
sce_hsc$species_pub <- vector(length = ncol(sce_hsc))
sce_hsc$species_pub[sce_hsc$Species_ID == "mmus"] <- "BL6"
sce_hsc$species_pub[sce_hsc$Species_ID == "mcas"] <- "CAST"
sce_hsc$species_pub[sce_hsc$Species_ID == "mspr"] <- "SPRET"
sce_hsc$species_pub[sce_hsc$Species_ID == "mcar"] <- "CAROLI"
sce_hsc$species_pub <- factor(sce_hsc$species_pub, 
                              levels = names(col_spc_pub))
print(col_spc_pub)

sce_str$species_pub <- vector(length = ncol(sce_str))
sce_str$species_pub[sce_str$Species_ID == "mmus"] <- "BL6"
sce_str$species_pub[sce_str$Species_ID == "mcas"] <- "CAST"
sce_str$species_pub[sce_str$Species_ID == "mspr"] <- "SPRET"
sce_str$species_pub[sce_str$Species_ID == "mcar"] <- "CAROLI"
sce_str$species_pub <- factor(sce_str$species_pub, 
                              levels = names(col_spc_pub))
```

```{r pdf_output}
# define output paths for pdf export

  output_pdf_umap_species_hsc <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_species_hsc.pdf")
  output_pdf_umap_species_str <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_species_str.pdf")
  output_pdf_umap_species_legend <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_species_legend.pdf")

  output_pdf_umap_spec_sep_hsc <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_specsep_hsc.pdf") 
  output_pdf_umap_spec_sep_str <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_specsep_str.pdf") 
  output_pdf_umap_spec_sep_legend <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_umap_specsep_legend.pdf") 

  output_pdf_dotplot <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_dotplot.pdf") 
  output_pdf_dotplot_legend <- base::paste0(
    base_path,
    "/manuscript1_rev/figure1/figure1_dotplot_legend.pdf") 

```

```{r params}
# global params
source("../figures/determine_params.R")

# specific params
ct_hsc <- "HSC"
ct_str <- "Adipo/CAR"
```

## Obtain data

Obtain data needed only for manuscript text, e.g. cell numbers.

```{r obtain_data}
# nr of cells
ncol(sce_hsc)
ncol(sce_str)

# median nr of UMIs per cell
# remove lots of unneccessary stuff to avoid overloading the environment
sce_hsc_met <- sce_hsc
sce_str_met <- sce_str
rowData(sce_hsc_met) <- rowData(sce_hsc_met)[,c(2, 3, 4)]
rowData(sce_str_met) <- rowData(sce_str_met)[,c(2, 3, 4)]

reducedDim(sce_hsc_met) <- NULL
reducedDim(sce_str_met) <- NULL

assays(sce_hsc_met) <- list(assays(sce_hsc_met)[["counts_before_BC"]])
names(assays(sce_hsc_met)) <- "counts_before_BC"
assays(sce_str_met) <- list(assays(sce_str_met)[["counts_before_BC"]])
names(assays(sce_str_met)) <- "counts_before_BC"

qc_df <- scuttle::perCellQCMetrics(cbind(sce_hsc_met, sce_str_met),
                                   assay.type = "counts_before_BC")
median(qc_df$sum)

rm(sce_hsc_met)
rm(sce_str_met)
gc()

# nr of subclustering genes, and which genes are subclustering genes
subclustering_df[subclustering_df$purpose == "subclustering",]

data.frame(table(sce_hsc$Object_ID, sce_hsc$PID)) %>% subset(Freq >0)
data.frame(table(sce_str$Object_ID, sce_str$PID)) %>% subset(Freq >0)
table(sce_hsc$Object_ID)
table(sce_str$Object_ID)

table(sce_hsc$Age_weeks)
table(sce_str$Age_weeks)
```

## UMAPs Species

HSPCs (left) and Niche cells (right) clustered in UMAP space, 
colored by species.

#### Prep

```{r species_prep}
hsc_umap_species <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc, type = "UMAP"))
hsc_umap_species$color_by <- sce_hsc$species_pub
hsc_umap_species <- hsc_umap_species[base::sample(
  1:nrow(hsc_umap_species),
  nrow(hsc_umap_species),
  replace = FALSE),]

str_umap_species <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str, type = "UMAP"))
str_umap_species$color_by <- sce_str$species_pub
str_umap_species <- str_umap_species[base::sample(
  1:nrow(str_umap_species),
  nrow(str_umap_species),
  replace = FALSE),]
```

#### Plot

```{r species_base_plot_left}
stopifnot(unique(hsc_umap_species$color_by) %in% names(col_spc_pub))
species_base_plot_left <- ggplot2::ggplot(
  hsc_umap_species, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c((base::min(hsc_umap_species$X1)-1), 
                  (base::max(hsc_umap_species$X1))+1))+
  ggplot2::ylim(c((base::min(hsc_umap_species$X2)-1),
                  (base::max(hsc_umap_species$X2))+1))+
  ggplot2::scale_color_manual("Species", values = col_spc_pub)
species_base_plot_left
```

```{r species_theme_plot_left, fig.width = 4, fig.height = 3}
species_theme_plot_left <- species_base_plot_left+
  cowplot::theme_nothing()
species_theme_plot_left
```

```{r species_base_plot_right}
species_base_plot_right <- ggplot2::ggplot(
  str_umap_species, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c((base::min(str_umap_species$X1)-1),
                  (base::max(str_umap_species$X1))+1))+
  ggplot2::ylim(c((base::min(str_umap_species$X2)-1), 
                  (base::max(str_umap_species$X2))+1))+
  ggplot2::scale_color_manual("Species", values = col_spc_pub)
```

```{r species_theme_plot_right, fig.width = 4, fig.height = 3}
species_theme_plot_right <- species_base_plot_right+
  cowplot::theme_nothing()
species_theme_plot_right
```

#### Legend

Only one required

```{r species_legend_plot}
species_legend_plot <- ggplot2::ggplot(
  hsc_umap_species, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("Mouse species", values = col_spc_pub)
```

```{r species_legend_theme_plot, fig.width = 2, fig.height = 2}
species_legend_theme_plot <- species_legend_plot+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
species_theme_legend <- ggpubr::get_legend(species_legend_theme_plot)
species_legend <- ggpubr::ggarrange(species_theme_legend)

species_legend
```

## UMAPs celltype separated species 

HSPCs (left) and Niche cells (right) clustered in UMAP space, 
separated by mouse species.

#### PREP

```{r d_prep2}
hsc_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc, type = "UMAP"))
hsc_umap$color_by <- sce_hsc$celltypes 
hsc_umap$species <- sce_hsc$species_pub 
hsc_umap <- hsc_umap[base::sample(
  1:nrow(hsc_umap),
  nrow(hsc_umap),
  replace = FALSE),]
hsc_umap$color_by <- factor(
  hsc_umap$color_by,
  levels = names(col_cts_hsc)
)

str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str, type = "UMAP"))
str_umap$color_by <- sce_str$celltypes
str_umap$species <- sce_str$species_pub 
str_umap <- str_umap[base::sample(
  1:nrow(str_umap),
  nrow(str_umap),
  replace = FALSE),]
str_umap$color_by <- factor(
  str_umap$color_by,
  levels = names(col_cts_str)
)
```

#### PLOT

```{r ctspc_base_plot_left}
ctspc_base_plot_left <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$X1)-1), (base::max(hsc_umap$X1)+1))+
  ggplot2::ylim(c(base::min(hsc_umap$X2)-1), (base::max(hsc_umap$X2)+1))+
  ggplot2::scale_color_manual(values = col_cts_hsc)+
  facet_grid(rows = vars(hsc_umap$species))
```

```{r ctspc_theme_plot_left, fig.height = 9, fig.width = 3}
ctspc_theme_plot_left <- ctspc_base_plot_left+
  cowplot::theme_nothing()
ctspc_theme_plot_left
```

```{r ctspc_base_plot_right}
ctspc_base_plot_right <- ggplot2::ggplot(
  str_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$X1)-1), (base::max(str_umap$X1)+1))+
  ggplot2::ylim(c(base::min(str_umap$X2)-1), (base::max(str_umap$X2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)+
  facet_grid(rows = vars(str_umap$species))
```

```{r d_theme_plot_right, fig.height = 10, fig.width = 4}
ctspc_theme_plot_right <- ctspc_base_plot_right+
  cowplot::theme_nothing()
ctspc_theme_plot_right
```

#### Legend

```{r ct_legend_plot_left}
ct_legend_plot_left <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("Cell type", values = col_cts_hsc)
```

```{r ct_legend_plot_right}
ct_legend_plot_right <- ggplot2::ggplot(
  str_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("Cell type", values = col_cts_str)
```

```{r ct_theme_legend_plot_left, fig.width = 3, fig.height = 4}
ct_theme_legend_plot_left <- ct_legend_plot_left+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
ct_legend_left_theme <- ggpubr::get_legend(ct_theme_legend_plot_left)

ct_legend_left <- ggpubr::ggarrange(ct_legend_left_theme) 
ct_legend_left
```


```{r ct_theme_legend_plot_right, fig.width = 3, fig.height = 3}
ct_theme_legend_plot_right <- ct_legend_plot_right+
  theme_all+
  ggplot2::theme(
    legend.title = element_blank(),
    legend.key.height = unit(0.3, "lines")
  )
ct_legend_right_theme <- ggpubr::get_legend(ct_theme_legend_plot_right)

ct_legend_right <- ggpubr::ggarrange(ct_legend_right_theme) 
ct_legend_right
```

```{r e_theme_legend, fig.height = 7, fig.width = 4}
ct_legend <- ggpubr::ggarrange(
  ncol = 1,
  align = "v",
  ct_legend_left,
  ct_legend_right)

ct_legend
```

## Assemble UMAPs

```{r umap_construct1, fig.width = 8, fig.height = 3}
species_plot <- ggpubr::ggarrange(
  align = "hv",
  species_theme_plot_left,
  species_theme_plot_right,
  ncol = 2
)
species_plot
```

```{r umap_construct2, fig.width = 8, fig.height = 10}
species_sep_plot <- ggpubr::ggarrange(
  align = "hv",
  ctspc_theme_plot_left,
  ctspc_theme_plot_right,  
  ncol = 2
)
species_sep_plot
```

#### EXPORT

```{r export_species_plot_left_pdf}
pdf(output_pdf_umap_species_hsc, width = 4, height = 3)
species_theme_plot_left
dev.off()
```

```{r export_species_plot_right_pdf}
pdf(output_pdf_umap_species_str, width = 4, height = 3)
species_theme_plot_right
dev.off()
```

```{r export_species_legend_pdf}
pdf(output_pdf_umap_species_legend, width = 2, height = 2)
species_legend
dev.off()
```

```{r export_ctspc_left_pdf}
pdf(output_pdf_umap_spec_sep_hsc, width = 3, height = 9)
ctspc_theme_plot_left
dev.off()
```

```{r export_ctspc_right_pdf}
pdf(output_pdf_umap_spec_sep_str, width = 3, height = 9)
ctspc_theme_plot_right
dev.off()
```

```{r export_ctspc_legend_pdf}
pdf(output_pdf_umap_spec_sep_legend, width = 4, height = 7)
ct_legend
dev.off()
```

## DOTPLOT

Average gene expression of well-known marker genes that were not used for
cell type sub-clustering per species for HSCs (top left) and Adipo/CARs 
(top right) or per cell type for HSPC fraction (bottom left) or Niche fraction 
(bottom right).

#### PREP

This is for all cell types together.

```{r dp_prepare}
# get genes
gene_df_hsc <- dotplot_order_df[dotplot_order_df$fraction == "hsc",]
gene_df_hsc$gene <- factor(gene_df_hsc$gene, levels = gene_df_hsc$gene)
genes_hsc <- gene_df_hsc$gene
genes_hsc

gene_df_str <- dotplot_order_df[dotplot_order_df$fraction == "str",]
gene_df_str$gene <- factor(gene_df_str$gene, levels = gene_df_str$gene)
genes_str <- gene_df_str$gene
genes_str

# aggregate across cell types, get average expression, use normalized counts
agg_hsc <- scuttle::aggregateAcrossCells(
  sce_hsc, 
  id=colData(sce_hsc)[,c("celltypes")], 
  statistics = "mean",
  use.assay.type = "logcounts")

agg_str <- scuttle::aggregateAcrossCells(
  sce_str, 
  id=colData(sce_str)[,c("celltypes")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# make data frames for visualisation
vis_df_hsc <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg_hsc)$logcounts[
    rownames(SummarizedExperiment::assays(agg_hsc)$logcounts) %in% genes_hsc,]))

vis_df_str <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg_str)$logcounts[
    rownames(SummarizedExperiment::assays(agg_str)$logcounts) %in% genes_str,]))

# format data frames
vis_df_hsc <- tibble::rownames_to_column(vis_df_hsc, "celltypes")
vis_df_hsc <- tidyr::pivot_longer(
  vis_df_hsc, 
  cols = c(2:ncol(vis_df_hsc)), 
  values_to = "expression", 
  names_to = "gene")

vis_df_str <- tibble::rownames_to_column(vis_df_str, "celltypes")
vis_df_str <- tidyr::pivot_longer(
  vis_df_str, 
  cols = c(2:ncol(vis_df_str)), 
  values_to = "expression", 
  names_to = "gene")

# factor cell types for correct order
vis_df_hsc$celltypes <- factor(
  vis_df_hsc$celltypes, 
  levels = base::rev(names(col_cts_hsc)))

vis_df_str$celltypes <- factor(
  vis_df_str$celltypes, 
  levels = base::rev(names(col_cts_str)))

# factor genes for correct order
vis_df_hsc$gene <- factor(
  vis_df_hsc$gene, 
  levels = levels(gene_df_hsc$gene))

vis_df_str$gene <- factor(
  vis_df_str$gene, 
  levels = levels(gene_df_str$gene))
```

```{r dp_function}
get_perc_expressed <- function(vis_df, sce){
  
  vis_df <- vis_df
  sce <- sce
  
  vis_df$percent_expressed <- vector(length = nrow(vis_df))
  
  cts <- base::unique(vis_df$celltypes)
  gs <- base::unique(vis_df$gene)
  
  for(ct in cts){
    
    # normalized (MultiBatchNorm)
    lc_temp <- SummarizedExperiment::assays(sce)$logcounts[,sce$celltypes == ct]
    
    for(g in gs){
      
      # the number of cells expressing g > 0 times / absolute cells * 100 per ct
      perc_expr <- length(which(lc_temp[
        rownames(lc_temp) == g,] > 0))/ncol(lc_temp)*100
      
      vis_df$percent_expressed[
        which(vis_df$celltypes == ct & vis_df$gene == g)] <- perc_expr
    }
  }
  return(vis_df)
}

vis_df_hsc <- get_perc_expressed(vis_df_hsc, sce_hsc)
vis_df_str <- get_perc_expressed(vis_df_str, sce_str)
```

This is for just the selected cell types.

```{r dp_prepare_ct}
# aggregate
sce_hsc_ct <- sce_hsc[,sce_hsc$celltypes == ct_hsc]
agg_ct_hsc <- scuttle::aggregateAcrossCells(
  sce_hsc_ct,
  id=colData(sce_hsc_ct)[,c("species_pub")], 
  statistics = "mean",
  use.assay.type = "logcounts")

sce_str_ct <- sce_str[,sce_str$celltypes == ct_str]
agg_ct_str <- scuttle::aggregateAcrossCells(
  sce_str_ct,
  id=colData(sce_str_ct)[,c("species_pub")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# make vis_dfs
vis_df_ct_hsc <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg_ct_hsc)$logcounts[
    rownames(SummarizedExperiment::assays(agg_ct_hsc)$logcounts) %in% 
      genes_hsc,]))

vis_df_ct_str <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg_ct_str)$logcounts[
    rownames(SummarizedExperiment::assays(agg_ct_str)$logcounts) %in% 
      genes_str,]))

# format
vis_df_ct_hsc <- tibble::rownames_to_column(vis_df_ct_hsc, "species")
vis_df_ct_hsc <- tidyr::pivot_longer(
  vis_df_ct_hsc,
  cols = c(2:ncol(vis_df_ct_hsc)), 
  values_to = "expression", 
  names_to = "gene")

vis_df_ct_str <- tibble::rownames_to_column(vis_df_ct_str, "species")
vis_df_ct_str <- tidyr::pivot_longer(
  vis_df_ct_str,
  cols = c(2:ncol(vis_df_ct_str)), 
  values_to = "expression", 
  names_to = "gene")

vis_df_ct_hsc$celltypes <- base::rep(ct_hsc, nrow(vis_df_ct_hsc))
vis_df_ct_str$celltypes <- base::rep(ct_str, nrow(vis_df_ct_str))

# factor
vis_df_ct_hsc$species <- factor(vis_df_ct_hsc$species, 
                                levels = base::rev(names(col_spc_pub)))
vis_df_ct_hsc$gene <- factor(vis_df_ct_hsc$gene,
                             levels = levels(gene_df_hsc$gene))

vis_df_ct_str$species <- factor(vis_df_ct_str$species, 
                              levels = base::rev(names(col_spc_pub)))
vis_df_ct_str$gene <- factor(vis_df_ct_str$gene,
                             levels = levels(gene_df_str$gene))

# perc expressed
vis_df_ct_hsc <- get_perc_expressed(vis_df_ct_hsc, sce_hsc)
vis_df_ct_str <- get_perc_expressed(vis_df_ct_str, sce_str)
```

```{r background_hsc}
# make background grid
background_hsc <- data.frame(
  xmin = c( # start of each segment 
    "Mllt3", # HSC
    "Cd34", # MPP
    "Cebpa", # Myeloid
    "Irf8", # Mono
    "Cebpe", # Neutro
    "Satb1", # Lymph
    "Gata2", # BM
    "Itga2b", # Mk
    "Gata1", # Mk ery
    "Klf1", # Ery
    "Mcm6", # Cell cycle entry
    "Mki67" # Cycling
    ), 
  xmax = c( # end of each segment 
    "Hlf", # HSC
    "Cd34", # MPP
    "Cebpa", # Myeloid
    "Ccr2", # Mono
    "S100a8", # Neutro 
    "Il7r", # Lymph
    "Lmo4", # BM
    "Zfpm1", # Mk
    "Gata1", # Mk ery
    "Hba-a1", # Ery
    "Lig1", # Cell cycle entry
    "Cenpa" # Cycling
  ), 
  ymin = -Inf,              # covers full height
  ymax = Inf,
  fill = c(
    "grey80",
    rep(c("white", "grey80"), 5),
    "white")  # alternating colors
)

background_hsc$xmin[!background_hsc$xmin %in% vis_df_ct_hsc$gene]
background_hsc$xmax[!background_hsc$xmax %in% vis_df_ct_hsc$gene]

background_hsc$xmin[!background_hsc$xmin %in% vis_df_hsc$gene]
background_hsc$xmax[!background_hsc$xmax %in% vis_df_hsc$gene]

background_hsc$xmin <- base::match(background_hsc$xmin, 
                                   levels(vis_df_ct_hsc$gene))
background_hsc$xmax <- base::match(background_hsc$xmax, 
                                   levels(vis_df_ct_hsc$gene))

background_hsc$xmin
background_hsc$xmax

background_hsc
```

```{r background_str}
# make background grid
background_str <- data.frame(
  xmin = c(
    "Cxcl12", # Niche
    "Lepr", # Adipo
    "Spp1", # Osteo
    "Sox9", # Chondro
    "Comp", # Fibro/Chondro
    "Clec3b", # Fibro
    "Cdh5", # Pan EC
    "Sox17", # Art
    "Car4", # Cap
    "Ackr1", # Ven
    "Tfpi", # Sinusoid
    "Lyve1", # Lymphatic
    "Acta2" # Peri SMC
    ),  
  xmax = c(
    "Kitl", # Niche
    "Lepr", # Adipo
    "Bglap", # Osteo
    "Fmod", # Chondro
    "Apod", # Fibro/Chondro
    "Cd34", #  Fibro
    "Pecam1", # PAN EC
    "Efnb2", # Art
    "Rgcc", # Cap
    "Lrg1", # Ven
    "Stab2", # Sinusoid
    "Ccl21a", # Lymphatic
    "Rgs5" # Peri SMC
  ), 
  ymin = -Inf,             
  ymax = Inf,
  fill = c(
    "grey80",
    rep(c("white", "grey80"),  6)) 
)

background_str$xmin[!background_str$xmin %in% vis_df_ct_str$gene]
background_str$xmax[!background_str$xmax %in% vis_df_ct_str$gene]

background_str$xmin[!background_str$xmin %in% vis_df_str$gene]
background_str$xmax[!background_str$xmax %in% vis_df_str$gene]


background_str$xmin <- base::match(background_str$xmin,
                                   levels(vis_df_ct_str$gene))
background_str$xmax <- base::match(background_str$xmax,
                                   levels(vis_df_ct_str$gene))

background_str$xmin
background_str$xmax

background_str
```

#### PLOT

```{r dp_base_plot_top_left}
dp_base_plot_top_left <- ggplot2::ggplot(
  vis_df_ct_hsc, 
  aes(x = gene,
      y = species,
      color = percent_expressed,
      size = expression))+
  ggplot2::geom_rect(
    data = background_hsc, 
    aes(xmin = xmin - 0.5,
        xmax = xmax + 0.5, 
        ymin = ymin,
        ymax = ymax),
    fill = background_hsc$fill,
    inherit.aes = FALSE, 
    alpha = 0.2)+
  ggplot2::geom_point()
# printing the base plots from snkmk_isbm doesn't work for
# unclear reasons!!
```

```{r dp_theme_plot_top_left}
dp_theme_plot_top_left <- dp_base_plot_top_left+
  theme_all+
  ggplot2::scale_x_discrete(position = "top")+
  ggplot2::theme(
    axis.title = element_blank(),
    axis.text.x.top = element_text(
      size = axis_text_size_small,
      hjust = 0,
      vjust = 0.5,
      angle = 90,
      color = axis_text_color,
      face = axis_text_face),
    axis.text.y = element_text(
      hjust = 1),
    legend.position = "none")+
  ggplot2::ylab("Species")+
  ggplot2::xlab("Gene")+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    limits = c(0, base::max(vis_df_hsc$expression)))+
  ggplot2::scale_colour_gradientn(
     "% expressed",
     colors = mycolors_to100,
     limits = c(0, 100))
dp_theme_plot_top_left
```

```{r dp_base_plot_top_right}
dp_base_plot_top_right <- ggplot2::ggplot(
  vis_df_ct_str, 
  aes(x = gene,
      y = species,
      color = percent_expressed,
      size = expression))+
  ggplot2::geom_rect(
    data = background_str, 
    aes(xmin = xmin - 0.5,
        xmax = xmax + 0.5, 
        ymin = ymin,
        ymax = ymax),
    fill = background_str$fill,
    inherit.aes = FALSE, 
    alpha = 0.2)+
  ggplot2::scale_x_discrete(position = "top")+
  ggplot2::geom_point()
```

```{r dp_theme_plot_top_right}
dp_theme_plot_top_right <- dp_base_plot_top_right+
  theme_all+
  ggplot2::theme(
    axis.title = element_blank(),
    axis.text.x.top = element_text(
      size = axis_text_size_small,
      hjust = 0,
      vjust = 0.5,
      angle = 90,
      color = axis_text_color,
      face = axis_text_face),
    axis.text.y = element_blank(),
    legend.position = "none")+
  ggplot2::xlab("Gene")+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    limits = c(0, base::max(vis_df_ct_str$expression)))+
  ggplot2::scale_colour_gradientn(
     "% expressed",
     colors = mycolors_to100,
     limits = c(0, 100)
  )
dp_theme_plot_top_right
```

```{r dp_base_plot_bottom_left}
dp_base_plot_bottom_left <- ggplot2::ggplot(
  vis_df_hsc, 
  aes(x = gene,
      y = celltypes,
      color = percent_expressed,
      size = expression))+
  ggplot2::geom_point() 
```

```{r dp_theme_plot_bottom_left, fig.width = 25, fig.height = 6}
dp_theme_plot_bottom_left <- dp_base_plot_bottom_left+
  theme_all+
  ggplot2::theme(
    axis.title = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(
      hjust = 1,
      size = axis_text_size,
      color = axis_text_color,
      face = axis_text_face),
    legend.position = "none")+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    limits = c(0, base::max(vis_df_hsc$expression)))+
  ggplot2::scale_colour_gradientn(
     "% expressed",
     colors = mycolors_to100,
     limits = c(0, 100)
  )
dp_theme_plot_bottom_left
```

```{r dp_base_plot_bottom_right}
dp_base_plot_bottom_right <- ggplot2::ggplot(
  vis_df_str, 
  aes(x = gene,
      y = celltypes,
      color = percent_expressed,
      size = expression))+
  ggplot2::geom_point()
```

```{r dp_theme_plot_bottom_right, fig.width = 20,fig.height = 5}
dp_theme_plot_bottom_right <- dp_base_plot_bottom_right+
  theme_all+
  ggplot2::theme(
    axis.title = element_blank(),
    axis.text.y = element_text(
      hjust = 1,
      size = axis_text_size,
      color = axis_text_color,
      face = axis_text_face),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none")+
  ggplot2::ylab("Cell type")+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    limits = c(0, base::max(vis_df_ct_str$expression)))+
  ggplot2::scale_colour_gradientn(
     "% expressed",
     colors = mycolors_to100,
     limits = c(0, 100)
  )
dp_theme_plot_bottom_right
```

#### LEGEND

Since logcounts have different upper limits in HSPCs vs Niche, need two.

```{r dp_legend_base_plot_left}
dp_legend_base_plot_left <- ggplot2::ggplot(
  vis_df_hsc, 
  aes(x = gene,
      y = celltypes,
      size = expression))+
  ggplot2::geom_point()
```

```{r dp_legend_theme_plot_left}
dp_legend_theme_plot_left <- dp_legend_base_plot_left+
  theme_all+
  ggplot2::theme(
    legend.box.just = "left",
    legend.position = "bottom",
    legend.box = "vertical", # Stack legends vertically
    legend.spacing.y = unit(0.5, 'cm'))+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    breaks = c(0, 1, 2),
    limits = c(0, base::max(vis_df_hsc$expression)))
dp_legend_left <- ggpubr::get_legend(dp_legend_theme_plot_left)
```

```{r dp_legend_base_plot_middle}
dp_legend_base_plot_middle <- ggplot2::ggplot(
  vis_df_str, 
  aes(x = gene,
      y = celltypes,
      color = percent_expressed))+
  ggplot2::geom_point()
```

```{r dp_legend_theme_plot_middle}
dp_legend_theme_plot_middle <- dp_legend_base_plot_middle+
  theme_all+
  ggplot2::theme(
    legend.box.just = "left",
    legend.position = "bottom",
    legend.box = "vertical", # Stack legends vertically
    legend.spacing.y = unit(0.5, 'cm'))+
  ggplot2::scale_colour_gradientn(
    "% expressed",
    breaks = c(0, 50, 100),
    colors = mycolors_to100,
    limits = c(0, 100)
  )
dp_legend_middle <- ggpubr::get_legend(dp_legend_theme_plot_middle)
```

```{r dp_legend_base_plot_right}
dp_legend_base_plot_right <- ggplot2::ggplot(
  vis_df_str, 
  aes(x = gene,
      y = celltypes,
      size = expression))+
  ggplot2::geom_point()
```

```{r dp_legend_theme_plot_right}
dp_legend_theme_plot_right <- dp_legend_base_plot_right+
  theme_all+
  ggplot2::theme(
    legend.box.just = "left",
    legend.position = "bottom",
    legend.box = "vertical", # Stack legends vertically
    legend.spacing.y = unit(0.5, 'cm'))+
  ggplot2::scale_size_area(
    "logcounts",
    max_size = max_size_dotplots,
    limits = c(0, base::max(vis_df_ct_str$expression)))
dp_legend_right <- ggpubr::get_legend(dp_legend_theme_plot_right)
```

### Species proportions tiny

Add little barplots showing the nr of cells per cell type per species.

```{r x_hsc_base_plot}

df_hsc <- as.data.frame(colData(sce_hsc)[
  colnames(colData(sce_hsc)) %in% c("species_pub", "celltypes")])
table_hsc <- table(sce_hsc$celltypes, sce_hsc$species_pub)
table_hsc

df_hsc$species_pub <- factor(df_hsc$species_pub, levels = names(col_spc_pub))
df_hsc$celltypes <- factor(df_hsc$celltypes, levels = rev(names(col_cts_hsc)))

x_hsc_base_plot <- ggplot2::ggplot(
  df_hsc, 
  aes(y = celltypes,
      fill = species_pub))+
  ggplot2::geom_bar()
```

Except for mcar lymph, there are at least 97 cells per species and cell types.

```{r x_hsc_theme_plot}
x_hsc_theme_plot <- x_hsc_base_plot+
  theme_all_supp+
  ggplot2::scale_fill_manual("Species", values = col_spc_pub)+
  ggplot2::scale_x_continuous(
    limits = c(0, 25000),
    breaks = c(0, 10000, 20000),
    labels = c(0, "", 20000))+
  ggplot2::theme(
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_text(
      angle = 0,
      vjust = 0.5,
      hjust = 0.5,
      size = axis_text_size,
      color = axis_text_color,
      face = axis_text_face),
    axis.title.x = element_text(
      size = axis_title_size,
    ),
    legend.position = "none")+
  ggplot2::xlab("# cells")
x_hsc_theme_plot
```

```{r x_str_base_plot}

df_str <- as.data.frame(colData(sce_str)[
  colnames(colData(sce_str)) %in% c("species_pub", "celltypes")])
table_str <- table(sce_str$celltypes, sce_str$species_pub)
table_str

df_str$species_pub <- factor(df_str$species_pub, levels = names(col_spc_pub))
df_str$celltypes <- factor(df_str$celltypes, levels = rev(names(col_cts_str)))

x_str_base_plot <- ggplot2::ggplot(
  df_str, 
  aes(y = celltypes,
      fill = species_pub))+
  ggplot2::geom_bar()
```

```{r x_str_theme_plot}
x_str_theme_plot <- x_str_base_plot+
  theme_all_supp+
  ggplot2::xlab("# cells")+
  ggplot2::scale_fill_manual("Species", values = col_spc_pub)+
  ggplot2::scale_x_continuous(
    limits = c(0, 10000),
    breaks = c(0, 4000, 8000),
    labels = c(0, "", 8000))+
  ggplot2::theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(
      angle = 0,
      vjust = 0.5,
      hjust = 0.5,
      size = axis_text_size,
      color = axis_text_color,
      face = axis_text_face),
    axis.title.x = element_text(
      size = axis_title_size
    ),
    legend.position = "none")
x_str_theme_plot
```

```{r x_legend, fig.width = 2, fig.height = 2}
x_legend_base_plot <- ggplot2::ggplot(
  df_str, 
  aes(y = celltypes,
      fill = species_pub))+
  ggplot2::geom_bar()+
  ggplot2::scale_fill_manual("Species", values = col_spc_pub)+
  ggplot2::theme(
    legend.title = element_text(
      size = legend_title_size,
      face = legend_title_face,
      color = legend_title_color),
    legend.text = element_text(
      size = legend_text_size,
      face = legend_text_face,
      color = legend_text_color
    ))

x_theme_legend <- ggpubr::get_legend(x_legend_base_plot)
x_legend <- ggpubr::ggarrange(x_theme_legend)

x_legend
```

#### ASSEMBLE

```{r dp_theme_plot2, fig.width = 16.5, fig.height = 6}
dp_theme_plot2 <- ggpubr::ggarrange(
  dp_theme_plot_top_left + theme(
    plot.margin = margin(t = 10, r = 0, b = 20, l = 62, unit = "pt")), 
  NULL,
  dp_theme_plot_top_right + theme(
    plot.margin = margin(t = 14, r = 0, b = 20, l = 157, unit = "pt")),
  NULL,
  dp_theme_plot_bottom_left + theme(
    plot.margin = margin(t = 5, r = 0, b = 47, l = 5, unit = "pt")),
  x_hsc_theme_plot+ theme(
    plot.margin = margin(t = 3, r = 3, b = 4, l = 2, unit = "pt")),
  dp_theme_plot_bottom_right + theme(
    plot.margin = margin(t = 5, r = 0, b = 47, l = 5, unit = "pt")),
  x_str_theme_plot+ theme(
    plot.margin = margin(t = 3, r = 3, b = 4, l = 2, unit = "pt")),
  heights = c(1.1,1.8),
  widths = c(1, 0.16, 1.09, 0.16),
  nrow = 2,
  ncol = 4
  )

dp_theme_plot2
```

```{r dp_theme_legends, fig.width = 6, fig.height = 7}
dp_theme_legends <- ggpubr::ggarrange(
  nrow = 4,
  dp_legend_left,
  dp_legend_middle,
  dp_legend_right,
  x_legend
  )

dp_theme_legends
```

#### EXPORT

```{r export_dp}
pdf(output_pdf_dotplot, width = 16.5, height = 6)
dp_theme_plot2
dev.off()
```

```{r export_dp_legend}
pdf(output_pdf_dotplot_legend, width = 6, height = 6) 
dp_theme_legends
dev.off()
```

# UTILS

```{r sessioninfo}
utils::sessionInfo()
```
