---
title: "Supplementary Figure 3"
date: '2024-11-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(ggalluvial, quietly = TRUE)
library(ggrepel, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
source("../source/custom_clustering_utils.R")

col_cts_hsc <- c(col_cts_hsc, "removed" = "grey70")
col_cts_str <- c(col_cts_str, "removed" = "grey70")

col_cts_hsc
col_cts_str
```

## Load and prepare objects

### Load objects with all cells

After the original louvain clustering of the merged dataset.

```{r load_sce_hsc_all_cells, eval = FALSE, include = FALSE}
# load data for manual use here

# SCE objects with all cells
sce_hsc_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_hsc_all_cells) <- list(
  counts = assays(sce_hsc_all_cells)$counts,
  logcounts = assays(sce_hsc_all_cells)$logcounts
)
reducedDims(sce_hsc_all_cells) <- list(
  UMAP = reducedDims(sce_hsc_all_cells)$UMAP
)

gc()
```

```{r load_sce_str_all_cells}
# load data for manual use here

# SCE objects with all cells
sce_str_all_cells <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"))

# remove unnecessary assays because the objects become too large to load
assays(sce_str_all_cells) <- list(
  counts = assays(sce_str_all_cells)$counts,
  logcounts = assays(sce_str_all_cells)$logcounts
)
reducedDims(sce_str_all_cells) <- list(
  UMAP = reducedDims(sce_str_all_cells)$UMAP
)

gc()
```

### Load the objects with the custom labels

After species-specific clustering.

```{r load_custom_cluster_labels_hsc, eval = FALSE, include = FALSE}
# add all labels to the object that contains all cells

# load objects
sce_mmus_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_hsc_custom.rds"))
sce_mcas_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_hsc_custom.rds"))
sce_mspr_hsc <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_hsc_custom.rds"))
sce_mcar_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_hsc_custom.rds"))

sce_hsc_all_cells$custom_label <- vector(length = ncol(sce_hsc_all_cells))

sce_mmus_hsc$custom_label <- unfactor(sce_mmus_hsc$custom_label)
sce_mcas_hsc$custom_label <- unfactor(sce_mcas_hsc$custom_label)
sce_mspr_hsc$custom_label <- unfactor(sce_mspr_hsc$custom_label)
sce_mcar_hsc$custom_label <- unfactor(sce_mcar_hsc$custom_label)

sce_hsc_all_cells$custom_label[match(
  colnames(sce_mmus_hsc), colnames(sce_hsc_all_cells))] <- sce_mmus_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcas_hsc), colnames(sce_hsc_all_cells))] <- sce_mcas_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mspr_hsc), colnames(sce_hsc_all_cells))] <- sce_mspr_hsc$custom_label
sce_hsc_all_cells$custom_label[match(
  colnames(sce_mcar_hsc), colnames(sce_hsc_all_cells))] <- sce_mcar_hsc$custom_label

rm(
  sce_mmus_hsc,
  sce_mcas_hsc,
  sce_mspr_hsc,
  sce_mcar_hsc)
gc()
```

```{r load_custom_cluster_labels_str}
# add all labels to the object that contains all cells

# load objects
sce_mmus_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds"))
sce_mcas_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds"))
sce_mspr_str <-base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds"))
sce_mcar_str <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds"))

sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label <- unfactor(sce_mmus_str$custom_label)
sce_mcas_str$custom_label <- unfactor(sce_mcas_str$custom_label)
sce_mspr_str$custom_label <- unfactor(sce_mspr_str$custom_label)
sce_mcar_str$custom_label <- unfactor(sce_mcar_str$custom_label)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label

gc()
```

### Load annotated objects

Final annotation after sub-clustering.

```{r load_anno_hsc, eval = FALSE, include = FALSE}
sce_hsc_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_str.rds")

sce_hsc_all_cells$celltypes <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_hsc_all_cells), colnames(sce_hsc_anno))]

rm(sce_hsc_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_hsc_all_cells$celltypes) <- c(
  levels(sce_hsc_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_hsc_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_hsc_all_cells <- sce_hsc_all_cells[,base::sample(
  x = c(1:ncol(sce_hsc_all_cells)), 
  size = ncol(sce_hsc_all_cells),
  replace = FALSE)]

print(sort(unique(sce_hsc_all_cells$custom_label)))
```

```{r load_anno_str}
sce_str_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_str.rds")

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

rm(sce_str_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]

print(sort(unique(sce_str_all_cells$custom_label)))
```

### Add factors

```{r prep_hsc, eval = FALSE, include = FALSE}
# change species names to currently used abbreviations
sce_hsc_all_cells$species_pub <- vector(length = ncol(sce_hsc_all_cells))
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mmus"] <- "BL6"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcas"] <- "CAST"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_hsc_all_cells$species_pub[sce_hsc_all_cells$Species_ID == "mcar"] <- "CAROLI"
sce_hsc_all_cells$species_pub <- factor(
  sce_hsc_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_hsc_all_cells$custom_label <- factor(
  sce_hsc_all_cells$custom_label,
  levels = factors_custom_hsc
)
```

```{r prep_str}
# change species names to currently used abbreviations
sce_str_all_cells$species_pub <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mmus"] <- "BL6"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcas"] <- "CAST"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mspr"] <- "SPRET"
sce_str_all_cells$species_pub[sce_str_all_cells$Species_ID == "mcar"] <- "CAROLI"

sce_str_all_cells$species_pub <- factor(
  sce_str_all_cells$species_pub, 
  levels = names(col_spc_pub))

sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label,
  levels = factors_custom_str
)
```

## Other prep

```{r pdf_output}
# define output paths for pdf export


```

```{r params}
# global params
source("../figures/determine_params.R")
```

# Base Plots 

Containing all cells, including contamination.
Both annotations (original merged and custom species-specific).

## Make DFs for UMAPs

```{r prep_str_df1}
str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str_all_cells, type = "UMAP"))

str_umap$celltypes <- sce_str_all_cells$celltypes 
str_umap$custom_label <- sce_str_all_cells$custom_label 
str_umap$species <- sce_str_all_cells$species_pub 

# add the remaining UMAP coordinates from each species-specific PCA/BC
str_umap$UMAP_spec1 <- vector(length = nrow(str_umap))
str_umap$UMAP_spec2 <- vector(length = nrow(str_umap))

str_umap$UMAP_spec1[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mmus_str), 
  rownames(str_umap))] <- reducedDims(sce_mmus_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcas_str), 
  rownames(str_umap))] <- reducedDims(sce_mcas_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mspr_str), 
  rownames(str_umap))] <- reducedDims(sce_mspr_str)$UMAP[,2]

str_umap$UMAP_spec1[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,1]
str_umap$UMAP_spec2[match(
  colnames(sce_mcar_str), 
  rownames(str_umap))] <- reducedDims(sce_mcar_str)$UMAP[,2]
```

## Make DFs for Alluvial

### MMUS

```{r mat_mmus_str}
mat_mmus_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

mmus_df_str <- mat_mmus_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mmus_df_str$Var2 <- factor(
  mmus_df_str$Var2, 
  levels = rev(factors_custom_str))

mmus_df_str <- mmus_df_str %>%
  dplyr::mutate(
    Var2_c = fct_inorder(paste0(" ", Var2)))
```

### MCAS

```{r mat_mcas_str}
mat_mcas_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcas"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcas"])

mcas_df_str <- mat_mcas_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcas_df_str$Var2 <- factor(
  mmus_df_str$Var2, 
  levels = rev(factors_custom_str))

mcas_df_str <- mcas_df_str %>%
  dplyr::mutate(
    Var2_c = fct_inorder(paste0(" ", Var2)))
```

### MSPR

```{r mat_mspr_str}
mat_mspr_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mspr"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mspr"])

mspr_df_str <- mat_mspr_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mspr_df_str$Var2 <- factor(
  mmus_df_str$Var2, 
  levels = rev(factors_custom_str))

mspr_df_str <- mspr_df_str %>%
  dplyr::mutate(
    Var2_c = fct_inorder(paste0(" ", Var2)))
```

```{r mat_mcar_str}
mat_mcar_str <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcar"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcar"])

mcar_df_str <- mat_mcar_str %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcar_df_str$Var2 <- factor(
  mmus_df_str$Var2, 
  levels = rev(factors_custom_str))

mcar_df_str <- mcar_df_str %>%
  dplyr::mutate(
    Var2_c = fct_inorder(paste0(" ", Var2)))
```

## Make UMAPs

### MMUS

```{r base_plot_str_celltypes_mmus_1}
base_plot_str_celltypes_mmus_1 <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r base_plot_str_celltypes_mmus_2}
base_plot_str_celltypes_mmus_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
```

```{r base_plot_str_celltypes_mmus_3, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mmus_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "BL6",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str)
```

### MCAS

```{r base_plot_str_celltypes_mcas_1, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcas_1 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAST",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r base_plot_str_celltypes_mcas_2, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcas_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAST",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
```

```{r base_plot_str_celltypes_mcas_3, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcas_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAST",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str)
```

### MSPR

```{r base_plot_str_celltypes_mspr_1, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mspr_1 <- ggplot2::ggplot(
  str_umap[str_umap$species == "SPRET",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r base_plot_str_celltypes_mspr_2, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mspr_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "SPRET",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
```

```{r base_plot_str_celltypes_mspr_3, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mspr_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "SPRET",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str)
```

### MCAR

```{r base_plot_str_celltypes_mcar_1, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcar_1 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAROLI",], 
  aes(x = V1,
      y = V2,
      color = celltypes))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r base_plot_str_celltypes_mcar_2, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcar_2 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAROLI",], 
  aes(x = V1,
      y = V2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1)+1))+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2)+1))+
  ggplot2::scale_color_manual(values = col_custom_str)
```

```{r base_plot_str_celltypes_mcar_3, fig.height = 3, fig.width = 3}
base_plot_str_celltypes_mcar_3 <- ggplot2::ggplot(
  str_umap[str_umap$species == "CAROLI",], 
  aes(x = UMAP_spec1,
      y = UMAP_spec2,
      color = custom_label))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::scale_color_manual(values = col_custom_str)
```

## Make Alluvials

### MMUS

```{r base_allu_str_mmus, fig.height = 3, fig.width = 5}
base_allu_str_mmus <- ggplot2::ggplot(
  data = mmus_df_str,
  aes(
    axis1 = Var1,
    axis2 = Var2_c, 
    y = Freq)
    )+
  ggalluvial::geom_alluvium(
    aes(fill = Var1),
    width = 0.05
    )+
  ggalluvial::geom_stratum(
    fill = "black",
    alpha = 0.2,
    color = "black",
    size = 0.2,
    width = 0.05
    )+
  ggrepel::geom_text_repel(
    stat = "stratum",
    aes(
      x = after_stat(x),
      label = after_stat(stratum),
      hjust = 0),
    nudge_x = -0.7,      
    direction = "y",         
    box.padding = 0.1,
    alpha = c( # make the y axis text "disappear" due to 0 alpha
      rep(1, length(levels(mmus_df_str$Var1))),
      rep(0, length(unique(unfactor(
        mmus_df_str[mmus_df_str$Freq != 0,]$Var2_c)))))     
    )+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_cts_str)
base_allu_str_mmus
```

### MCAS

```{r base_allu_str_mcas, fig.height = 3, fig.width = 5}
base_allu_str_mcas <- ggplot2::ggplot(
  data = mcas_df_str,
  aes(
    axis1 = Var1,
    axis2 = Var2_c, 
    y = Freq)
    )+
  ggalluvial::geom_alluvium(
    aes(fill = Var1),
    width = 0.05
    )+
  ggalluvial::geom_stratum(
    fill = "black",
    alpha = 0.2,
    color = "black",
    size = 0.2,
    width = 0.05
    )+
  ggrepel::geom_text_repel(
    stat = "stratum",
    aes(
      x = after_stat(x),
      label = after_stat(stratum),
      hjust = 0),
    nudge_x = -0.7,      
    direction = "y",         
    box.padding = 0.1,
    alpha = c( # make the y axis text "disappear" due to 0 alpha
      rep(1, length(levels(mcas_df_str$Var1))),
      rep(0, length(unique(unfactor(
        mmus_df_str[mcas_df_str$Freq != 0,]$Var2_c)))))     
    )+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_cts_str)
base_allu_str_mcas
```

### MSPR

```{r base_allu_str_mspr, fig.height = 3, fig.width = 5}
base_allu_str_mspr <- ggplot2::ggplot(
  data = mspr_df_str,
  aes(
    axis1 = Var1,
    axis2 = Var2_c, 
    y = Freq)
    )+
  ggalluvial::geom_alluvium(
    aes(fill = Var1),
    width = 0.05
    )+
  ggalluvial::geom_stratum(
    fill = "black",
    alpha = 0.2,
    color = "black",
    size = 0.2,
    width = 0.05
    )+
  ggrepel::geom_text_repel(
    stat = "stratum",
    aes(
      x = after_stat(x),
      label = after_stat(stratum),
      hjust = 0),
    nudge_x = -0.7,      
    direction = "y",         
    box.padding = 0.1,
    alpha = c( # make the y axis text "disappear" due to 0 alpha
      rep(1, length(levels(mspr_df_str$Var1))),
      rep(0, length(unique(unfactor(
        mmus_df_str[mspr_df_str$Freq != 0,]$Var2_c)))))     
    )+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_cts_str)
base_allu_str_mspr
```

### MCAR

```{r base_allu_str_mcar, fig.height = 3, fig.width = 5}
base_allu_str_mcar <- ggplot2::ggplot(
  data = mcar_df_str,
  aes(
    axis1 = Var1,
    axis2 = Var2_c, 
    y = Freq)
    )+
  ggalluvial::geom_alluvium(
    aes(fill = Var1),
    width = 0.05
    )+
  ggalluvial::geom_stratum(
    fill = "black",
    alpha = 0.2,
    color = "black",
    size = 0.2,
    width = 0.05
    )+
  ggrepel::geom_text_repel(
    stat = "stratum",
    aes(
      x = after_stat(x),
      label = after_stat(stratum),
      hjust = 0),
    nudge_x = -0.7,      
    direction = "y",         
    box.padding = 0.1,
    alpha = c( # make the y axis text "disappear" due to 0 alpha
      rep(1, length(levels(mcar_df_str$Var1))),
      rep(0, length(unique(unfactor(
        mmus_df_str[mcar_df_str$Freq != 0,]$Var2_c)))))     
    )+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_cts_str)
base_allu_str_mcar
```

# Build

## A plot MMUS

```{r a_mmus_plot_1, fig.width = 3, fig.height = 3} 

a_mmus_plot_1 <- base_plot_str_celltypes_mmus_1+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

a_mmus_plot_1
```

```{r a_mmus_legend_1, fig.width = 1, fig.height = 3} 

# this legend can be used for all species
a_mmus_legend_1_plot <- base_plot_str_celltypes_mmus_1 + 
  theme_all+
  ggplot2::geom_point(size = 4)+
  ggplot2::theme(
    legend.text = element_blank(),
    legend.title = element_blank())

a_mmus_legend_1 <- ggpubr::get_legend(a_mmus_legend_1_plot)
a_mmus_legend_1 <- ggpubr::ggarrange(a_mmus_legend_1)

a_mmus_legend_1
```

```{r a_mmus_plot_2, fig.width = 4, fig.height = 3} 

a_mmus_plot_2 <- base_allu_str_mmus+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

a_mmus_plot_2
```

```{r a_mmus_legend_2, fig.width = 5, fig.height = 3} 

# this legend can be used for all species
a_mmus_legend_2_plot <- base_plot_str_celltypes_mmus_2 + 
  theme_all+
  ggplot2::geom_point(size = 3)+
  ggplot2::theme(
    legend.text = element_text(size = 12),
    legend.spacing.y = unit(0.1, 'cm'),
    legend.key.size = unit(0.45, "cm"),
    legend.title = element_blank())

a_mmus_legend_2_temp <- ggpubr::get_legend(a_mmus_legend_2_plot)
a_mmus_legend_2 <- ggpubr::ggarrange(a_mmus_legend_2_temp)

a_mmus_legend_2
```

```{r a_mmus_plot_3, fig.width = 3, fig.height = 3} 

a_mmus_plot_3 <- base_plot_str_celltypes_mmus_2+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

a_mmus_plot_3
```

```{r a_mmus_plot_4, fig.width = 3, fig.height = 3} 

a_mmus_plot_4 <- base_plot_str_celltypes_mmus_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

a_mmus_plot_4
```

```{r arrange_a, fig.width = 16, fig.height = 3}

plot_a <- ggpubr::ggarrange(
  a_mmus_plot_1,
  a_mmus_legend_1,
  a_mmus_plot_2,
  a_mmus_legend_2,
  a_mmus_plot_3,
  a_mmus_plot_4,
  nrow = 1,
  widths = c(2.5, 0.3, 4, 4.2, 2.5, 2.5)
)

plot_a
```

## B plot MCAS

```{r b_mcas_plot_1, fig.width = 3, fig.height = 3} 

b_mcas_plot_1 <- base_plot_str_celltypes_mcas_1+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

b_mcas_plot_1
```

```{r b_mcas_plot_2, fig.width = 4, fig.height = 3} 

b_mcas_plot_2 <- base_allu_str_mcas+
  ggplot2::theme(
    legend.position = "none",
    plot.background = element_blank())

b_mcas_plot_2
```

```{r b_mcas_legend_2, fig.width = 4, fig.height = 2} 

# this legend can be used for all species
b_mcas_legend_2_plot <- base_plot_str_celltypes_mcas_2 + 
  theme_all+
  ggplot2::geom_point(size = 4)+
  ggplot2::theme(
    legend.text = element_text(size = 12),
    legend.spacing.y = unit(0.1, 'cm'),
    legend.key.size = unit(0.45, "cm"),
    legend.title = element_blank())

b_mcas_legend_2 <- ggpubr::get_legend(b_mcas_legend_2_plot)
b_mcas_legend_2 <- ggpubr::ggarrange(b_mcas_legend_2)

b_mcas_legend_2
```

```{r b_mcas_plot_3, fig.width = 3, fig.height = 3} 

b_mcas_plot_3 <- base_plot_str_celltypes_mcas_2+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

b_mcas_plot_3
```

```{r b_mcas_plot_4, fig.width = 3, fig.height = 3} 

b_mcas_plot_4 <- base_plot_str_celltypes_mcas_3+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")

b_mcas_plot_4
```

```{r arrange_b, fig.width = 16, fig.height = 3}

plot_b <- ggpubr::ggarrange(
  b_mcas_plot_1,
  a_mmus_legend_1,
  b_mcas_plot_2,
  b_mcas_legend_2,
  b_mcas_plot_3,
  b_mcas_plot_4,
  nrow = 1,
  widths = c(2.5, 0.3, 4, 4.2, 2.5, 2.5)
)

plot_b
```


