---
title: "Identifying reclaimable signature genes in other datasets' markers"
date: '2025-12-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  message    = FALSE,
  warning    = FALSE,
  fig.align  = "left"
)
```

```{r seed}
RNGkind("L'Ecuyer-CMRG")
set.seed(37)
```

```{r libraries}
library(tidyverse)
library(Seurat)
```

```{r paths}
base_path    <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
#project_path <- "/omics/groups/OE0433/internal/veronica/github/Interspecies_BM_phd"
```

```{r source}
colors_path <- base::paste0(
  base_path,
  "metadata/colors/colors.txt"
)

source("../source/colors.R")
source("../source/plotting.R")
source("../figures/determine_params.R")

#source(file.path(project_path, "code/source/colors.R"))
#source(file.path(project_path, "code/source/plotting.R"))
#source(file.path(project_path, "code/figures/determine_params.R"))
```

---

# Load input objects

```{r load_objects}
hsc_sig <- read.delim(
  file.path(
    base_path,
    "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/09_tabl/hsc_signature_table.csv"
  ),
  sep = ";"
)

str_sig <- read.delim(
  file.path(
    base_path,
    "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/09_tabl/str_signature_table.csv"
  ),
  sep = ";"
)

ensembl_hum <- readRDS(
  file.path(
    base_path, 
    "metadata/scRNAseq/03_sce_analysis/ensembl/ensembl_hum"
))

ensembl_mus <- readRDS(
  file.path(
    base_path, 
    "metadata/scRNAseq/03_sce_analysis/ensembl/ensembl_mus"
))

tm_hspc <- readRDS(file.path(
  base_path,
  "metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/mus_tm_bonemarrow"
))

weinreb_hspc <- readRDS(file.path(
  base_path,
  "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/mus_weinreb_hspc.Rds"
))

tabula_sapiens_hspc <- readRDS(file.path(
  base_path,
  "metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/ts_hscs_progenitors"
))

tabula_sapiens_bm <- readRDS(file.path(
  base_path,
  "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_bone_marrow.Rds"
))

tabula_sapiens_stromal <- readRDS(file.path(
  base_path,
  "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_all_stromal.Rds"
))
```

---

# Create ortholog table (mouse symbols -> human Ensembl IDs)

```{r build_orthologs}
# Build mouse symbol -> human Ensembl ortholog table
# using: ensembl_hum (has mmusculus_homolog_ensembl_gene) and ensembl_mus

ortholog_mouse2human <- ensembl_hum %>%
  dplyr::select(
    human_ensembl = ensembl_gene_id,              # human Ensembl
    mm_ensembl    = mmusculus_homolog_ensembl_gene,
    human_gene_name = external_gene_name          # optional, for debugging
  ) %>%
  dplyr::filter(!is.na(mm_ensembl)) %>%
  dplyr::left_join(
    ensembl_mus %>%
      dplyr::select(
        mm_ensembl = ensembl_gene_id,
        sig_gene   = external_gene_name          # mouse gene symbol
      ),
    by = "mm_ensembl"
  ) %>%
  dplyr::filter(!is.na(sig_gene)) %>%
  dplyr::transmute(
    sig_gene,                   # mouse signature symbol (matches hsc_sig$str_sig$signature_gene)
    seurat_gene    = human_ensembl,
    human_gene_name
  ) %>%
  dplyr::distinct()

```

---

# Helper function to calculate markers with optional ortholog mapping

```{r calculate_sig_marker_conservation}
analyze_seurat_vs_signatures <- function(seurat_obj,
                                          signatures,
                                          ortholog_df     = NULL,
                                          logfc_threshold = 0.25,
                                          min_pct         = 0.1) {
  # signatures: data.frame with at least:
  #   signature_gene (mouse symbol)
  #   celltype
  #   subclustering_gene_for_ct (logical) – we ignore it here unless you want a filter

  signatures_df <- signatures %>%
    dplyr::filter(
      !is.na(signature_gene),
      !is.na(celltype)
      # if you want to drop subclustering genes:
      # , !subclustering_gene_for_ct
    )

  # list: celltype -> vector of raw signature genes (mouse symbols)
  sig_list_raw <- signatures_df %>%
    dplyr::group_by(celltype) %>%
    dplyr::summarise(
      genes = list(unique(signature_gene)),
      .groups = "drop"
    )

  sig_cts <- sig_list_raw$celltype
  signature_genes_raw <- rlang::set_names(sig_list_raw$genes, sig_cts)

  gene_universe <- rownames(seurat_obj)

  # Map signature genes into the ID space of the Seurat object
  if (is.null(ortholog_df)) {
    # same-ID-space case (mouse signatures vs mouse Seurat object)
    signature_genes_mapped <- lapply(signature_genes_raw, function(gs) {
      intersect(gs, gene_universe)
    })
  } else {
    # cross-ID case: sig_gene (mouse symbol) -> seurat_gene (human Ensembl)
    ortholog_df_filt <- ortholog_df %>%
      dplyr::filter(seurat_gene %in% gene_universe)

    signature_genes_mapped <- lapply(sig_cts, function(ct) {
      gs <- signature_genes_raw[[ct]]
      ortholog_df_filt %>%
        dplyr::filter(sig_gene %in% gs) %>%
        dplyr::pull(seurat_gene) %>%
        unique()
    })
    names(signature_genes_mapped) <- sig_cts
  }

  # union of all signature genes (now in seurat_obj ID space)
  signatures_all <- unique(unlist(signature_genes_mapped))

  # use custom_label_unif as identities
  if (is.null(seurat_obj$custom_label_unif)) {
    stop("seurat_obj has no meta.data column 'custom_label_unif'")
  }
  Idents(seurat_obj) <- seurat_obj$custom_label_unif
  custom_cts <- levels(Idents(seurat_obj))

  custom_sig_by_ct     <- vector("list", length(custom_cts))
  names(custom_sig_by_ct) <- custom_cts
  custom_markers_by_ct <- vector("list", length(custom_cts))
  names(custom_markers_by_ct) <- custom_cts
  per_custom_ct_rows <- vector("list", length(custom_cts))
  names(per_custom_ct_rows) <- custom_cts

  for (ct in custom_cts) {
    mk <- Seurat::FindMarkers(
      object          = seurat_obj,
      ident.1         = ct,
      only.pos        = TRUE,
      logfc.threshold = logfc_threshold,
      min.pct         = min_pct
    )

    genes_ct <- rownames(mk)
    custom_markers_by_ct[[ct]] <- genes_ct

    overlap_any <- intersect(genes_ct, signatures_all)
    custom_sig_by_ct[[ct]] <- overlap_any

    per_custom_ct_rows[[ct]] <- tibble::tibble(
      custom_label_unif = ct,
      n_custom_markers  = length(genes_ct),
      n_overlap_any_set = length(overlap_any)
    )
  }

  all_custom_markers <- unique(unlist(custom_markers_by_ct))

  per_signature_rows <- purrr::map_dfr(sig_cts, function(hct) {
    raw_genes    <- unique(signature_genes_raw[[hct]])      # mouse symbols
    mapped_genes <- unique(signature_genes_mapped[[hct]])   # in seurat ID space

    n_sig_total <- length(raw_genes)

    sig_detectable   <- intersect(mapped_genes, gene_universe)
    n_sig_detectable <- length(sig_detectable)

    overlap <- intersect(mapped_genes, all_custom_markers)

    tibble::tibble(
      celltype                = hct,
      n_sig_total             = n_sig_total,
      n_sig_detectable        = n_sig_detectable,
      n_overlap_any_custom_ct = length(overlap),
      prop_overlap_any_custom =
        if (n_sig_detectable > 0) length(overlap) / n_sig_detectable else NA_real_
    )
  })

  list(
    per_custom_ct          = dplyr::bind_rows(per_custom_ct_rows),
    custom_sig_by_ct       = custom_sig_by_ct,
    custom_markers_by_ct   = custom_markers_by_ct,
    per_signature          = per_signature_rows,
    signature_genes_raw    = signature_genes_raw,    # list: celltype -> mouse symbols
    signature_genes_mapped = signature_genes_mapped  # list: celltype -> IDs in seurat_obj
  )
}
```

---

# Run the analysis and save outputs

```{r run_weinreb}
# Mouse tabula muris dataset
# not reclustered but function doesn't know that
tm_hspc$custom_label_unif <- tm_hspc$cell_type

# same ID space as signatures, so no ortholog mapping needed
tm_hspc_markers <- analyze_seurat_vs_signatures(
  seurat_obj = tm_hspc,
  signatures = hsc_sig,
  ortholog_df = NULL
)

saveRDS(
  tm_hspc_markers,
  file.path(
    base_path,
    "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/tm_hspc_markers.rds"
)
```

```{r run_weinreb}
# Mouse dataset; same ID space as signatures, so no ortholog mapping needed
weinreb_markers <- analyze_seurat_vs_signatures(
  seurat_obj  = weinreb_hspc,
  signatures  = hsc_sig,
  ortholog_df = NULL
)

saveRDS(
  weinreb_markers,
  file.path(base_path, "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/mus_weinreb_markers.rds")
)
```

```{r run_sapiens_bm}
# Human hspc dataset
# not reclustered but function doesn't know that
tabula_sapiens_hspc$custom_label_unif <- tabula_sapiens_hspc$cell_type

# use precomputed ortholog mapping from mouse symbols to human Ensembl IDs
tabula_sapiens_hspc_markers <- analyze_seurat_vs_signatures(
  seurat_obj = tabula_sapiens_hspc,
  signatures = hsc_sig,
  ortholog_df = ortholog_mouse2human
)

saveRDS(
  tabula_sapiens_hspc_markers,
  file.path(
    base_path,
    "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_hspc_markers.rds")
)
```

```{r run_sapiens_bm}
# Human bone marrow dataset
# use precomputed ortholog mapping from mouse symbols to human Ensembl IDs
tabula_sapiens_bm_markers <- analyze_seurat_vs_signatures(
  seurat_obj  = tabula_sapiens_bm,
  signatures  = hsc_sig,
  ortholog_df = ortholog_mouse2human
)

saveRDS(
  tabula_sapiens_bm_markers,
  file.path(base_path, "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_bm_markers.rds")
)
```

```{r run_sapiens_str}
# Human stromal dataset; same mapping
tabula_sapiens_str_markers <- analyze_seurat_vs_signatures(
  seurat_obj  = tabula_sapiens_stromal,
  signatures  = str_sig,
  ortholog_df = ortholog_mouse2human
)

saveRDS(
  tabula_sapiens_str_markers,
  file.path(base_path, "scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_str_markers.rds")
)
```

---

# Plot: mouse tabula muris HSPC (same-species, symbol-based)

```{r plot_tm_hspc}
custom_markers_by_ct <- tm_hspc_markers$custom_markers_by_ct
per_signature <- tm_hspc_markers$per_signature
signature_genes_mapped <- tm_hspc_markers$signature_genes_mapped

custom_order <- names(custom_markers_by_ct)
# use whatever signature order you want; example assumes col_cts_hsc exists
signature_order <- rev(names(col_cts_hsc))

pairwise_overlap_tm <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif,
      celltype,
      ~ length(intersect(
        custom_markers_by_ct[[.x]],
        signature_genes_mapped[[.y]]
      ))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable = dplyr::if_else(
      n_sig_detectable > 0,
      n_overlap / n_sig_detectable,
      NA_real_
    ),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype = factor(celltype, levels = signature_order)
  )

tm_hspc_dot <- pairwise_overlap_tm %>%
  ggplot(aes(x = custom_label_unif, y = celltype)) +
  geom_point(aes(size = n_overlap, color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes", range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature genes",
    low = "white",
    high = "blue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Mouse Tabula Muris HSPC"
  ) +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

tm_hspc_dot
```

# Plot: mouse Weinreb HSPC (same-species, symbol-based)

```{r plot_weinreb}
custom_markers_by_ct   <- weinreb_markers$custom_markers_by_ct
per_signature          <- weinreb_markers$per_signature
signature_genes_mapped <- weinreb_markers$signature_genes_mapped

custom_order    <- names(custom_markers_by_ct)
# use whatever signature order you want; example assumes col_cts_hsc exists
signature_order <- rev(names(col_cts_hsc))

pairwise_overlap_weinreb <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype          = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif, celltype,
      ~ length(intersect(custom_markers_by_ct[[.x]],
                         signature_genes_mapped[[.y]]))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable =
      dplyr::if_else(n_sig_detectable > 0, n_overlap / n_sig_detectable, NA_real_),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype          = factor(celltype,          levels = signature_order)
  )

weinreb_dot <- pairwise_overlap_weinreb %>%
  ggplot(aes(x = custom_label_unif, y = celltype)) +
  geom_point(aes(size  = n_overlap,
                 color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes",
                        range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature genes",
    low  = "white",
    high = "blue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Mouse Weinreb HSPC"
  )  +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

weinreb_dot
```

---

# Plot: human Tabula Sapiens bone marrow (ortholog-based)

```{r plot_ts_bm}
custom_markers_by_ct   <- tabula_sapiens_bm_markers$custom_markers_by_ct
per_signature          <- tabula_sapiens_bm_markers$per_signature
signature_genes_mapped <- tabula_sapiens_bm_markers$signature_genes_mapped

custom_order    <- names(custom_markers_by_ct)
signature_order <- rev(names(col_cts_hsc))

pairwise_overlap_ts_bm <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype          = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif, celltype,
      ~ length(intersect(custom_markers_by_ct[[.x]],
                         signature_genes_mapped[[.y]]))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable =
      dplyr::if_else(n_sig_detectable > 0, n_overlap / n_sig_detectable, NA_real_),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype          = factor(celltype,          levels = signature_order)
  )

ts_bm_dot <-  ggplot(pairwise_overlap_ts_bm,
  aes(x = custom_label_unif, 
      y = celltype)) +
  geom_point(aes(size  = n_overlap,
                 color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes",
                        range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature orthologs",
    low  = "white",
    high = "blue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Tabula Sapiens bone marrow"
  )  +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ts_bm_dot
```

# Plot: human Tabula Sapiens HSPC (ortholog-based)

```{r plot_ts_hspc}
custom_markers_by_ct <- tabula_sapiens_hspc_markers$custom_markers_by_ct
per_signature <- tabula_sapiens_hspc_markers$per_signature
signature_genes_mapped <- tabula_sapiens_hspc_markers$signature_genes_mapped

custom_order <- names(custom_markers_by_ct)
signature_order <- rev(names(col_cts_hsc))

pairwise_overlap_ts_hspc <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif,
      celltype,
      ~ length(intersect(
        custom_markers_by_ct[[.x]],
        signature_genes_mapped[[.y]]
      ))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable = dplyr::if_else(
      n_sig_detectable > 0,
      n_overlap / n_sig_detectable,
      NA_real_
    ),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype = factor(celltype, levels = signature_order)
  )

ts_hspc_dot <- ggplot(
  pairwise_overlap_ts_hspc,
  aes(x = custom_label_unif, y = celltype)
) +
  geom_point(aes(size = n_overlap, color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes", range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature orthologs",
    low = "white",
    high = "blue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Tabula Sapiens HSPC"
  ) +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ts_hspc_dot
```

---

# Plot: human Tabula Sapiens stromal (ortholog-based)

```{r plot_ts_stromal}
custom_markers_by_ct   <- tabula_sapiens_str_markers$custom_markers_by_ct
per_signature          <- tabula_sapiens_str_markers$per_signature
signature_genes_mapped <- tabula_sapiens_str_markers$signature_genes_mapped

custom_order    <- names(custom_markers_by_ct)
# define an appropriate order for stromal signatures; here we use names as-is
signature_order <- rev(names(col_cts_str))
signature_order <- signature_order[signature_order %in% names(signature_genes_mapped)] # some celltypes omitted


pairwise_overlap_ts_str <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype          = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif, celltype,
      ~ length(intersect(custom_markers_by_ct[[.x]],
                         signature_genes_mapped[[.y]]))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable =
      dplyr::if_else(n_sig_detectable > 0, n_overlap / n_sig_detectable, NA_real_),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype          = factor(celltype,          levels = signature_order)
  )

ts_str_dot <- ggplot(pairwise_overlap_ts_str,
  aes(x = custom_label_unif, 
      y = celltype)) +
  geom_point(aes(size  = n_overlap,
                 color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes",
                        range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature orthologs",
    low  = "white",
    high = "blue"
  )  +
    labs(
    x = NULL,
    y = NULL,
    title = "Tabula Sapiens Stromal"
  )  +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ts_str_dot
```

# save pdfs

```{r export_pdf}
pdf(
  file = file.path(
    base_path,
    "manuscript1_rev/pdfs/reclaiming_sigs_tm_hspc.pdf"
  ),
  height = 8,
  width = 8
)
print(tm_hspc_dot)
dev.off()

pdf(
  file = file.path(
    base_path,
    "manuscript1_rev/pdfs/reclaiming_sigs_weinreb.pdf"
  ),
  height = 8,
  width = 8
)
print(weinreb_dot)
dev.off()

pdf(
  file = file.path(base_path, "manuscript1_rev/pdfs/reclaiming_sigs_ts_bm.pdf"),
  height = 8,
  width = 8
)
print(ts_bm_dot)
dev.off()

pdf(
  file = file.path(
    base_path,
    "manuscript1_rev/pdfs/reclaiming_sigs_ts_hspc.pdf"
  ),
  height = 10,
  width = 8
)
print(ts_hspc_dot)
dev.off()

pdf(
  file = file.path(
    base_path,
    "manuscript1_rev/pdfs/reclaiming_sigs_ts_str.pdf"
  ),
  height = 10,
  width = 8
)
print(ts_str_dot)
dev.off()
```

# Additional analysis: Weinreb with original annotation

```{r}
weinreb_hspc$custom_label_unif <- weinreb_hspc$cell_type

weinreb_markers_2 <- analyze_seurat_vs_signatures(
  seurat_obj = weinreb_hspc,
  signatures = hsc_sig,
  ortholog_df = NULL
)

custom_markers_by_ct <- weinreb_markers_2$custom_markers_by_ct
per_signature <- weinreb_markers_2$per_signature
signature_genes_mapped <- weinreb_markers_2$signature_genes_mapped

custom_order <- names(custom_markers_by_ct)
# use whatever signature order you want; example assumes col_cts_hsc exists
signature_order <- rev(names(col_cts_hsc))

pairwise_overlap_weinreb_2 <- tidyr::crossing(
  custom_label_unif = custom_order,
  celltype = signature_order
) %>%
  dplyr::mutate(
    n_overlap = purrr::map2_int(
      custom_label_unif,
      celltype,
      ~ length(intersect(
        custom_markers_by_ct[[.x]],
        signature_genes_mapped[[.y]]
      ))
    )
  ) %>%
  dplyr::left_join(
    per_signature %>% dplyr::select(celltype, n_sig_detectable),
    by = "celltype"
  ) %>%
  dplyr::mutate(
    prop_of_sig_detectable = dplyr::if_else(
      n_sig_detectable > 0,
      n_overlap / n_sig_detectable,
      NA_real_
    ),
    custom_label_unif = factor(custom_label_unif, levels = custom_order),
    celltype = factor(celltype, levels = signature_order)
  )

weinreb_dot_2 <- pairwise_overlap_weinreb_2 %>%
  ggplot(aes(x = custom_label_unif, y = celltype)) +
  geom_point(aes(size = n_overlap, color = prop_of_sig_detectable)) +
  scale_size_continuous(name = "# overlapping genes", range = c(5, 9)) +
  scale_color_gradient(
    name = "Frac. recovered\nsignature genes",
    low = "white",
    high = "blue"
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "Mouse Weinreb HSPC"
  ) +
  theme_all +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

pdf(
  file = file.path(
    base_path,
    "manuscript1_rev/pdfs/reclaiming_sigs_weinreb_2.pdf"
  ),
  height = 8,
  width = 8
)
print(weinreb_dot_2)
dev.off()
```