---
title: "Genesets summary"
author: "Lea Wölbert"
date: '2023-11-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on gene sets, especially conserved signatures.

#### Load objects

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(scater, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(ComplexUpset, quietly = TRUE)
```

```{r source, message = FALSE}
source(file = snakemake@params[["plotting"]])
source(file = snakemake@params[["functions_reclustering"]])
```

```{r manual, eval = FALSE, include = FALSE}

geneset_list_hsc <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_hsc")
geneset_list_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_str")

sce_hsc <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10")
sce_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10")

colors_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/colors/colors.txt"
source("../../source/colors.R")
source("../../source/sce_functions_reclustering.R")
```

```{r load_objects}
sce_hsc <- base::readRDS(snakemake@input[["sce_hsc"]])
sce_str <- base::readRDS(snakemake@input[["sce_str"]])

geneset_list_hsc <- base::readRDS(snakemake@input[["geneset_list_hsc"]])
geneset_list_str <- base::readRDS(snakemake@input[["geneset_list_str"]])

cts_exclude <- snakemake@params[["cts_exclude"]]

# subset to required cell types (sce_str only)
sce_str <- sce_str[,which(!sce_str$celltypes %in% cts_exclude)]

# remove cell type levels that are not required anymore
sce_str$celltypes <- factor(
  sce_str$celltypes,
  levels = levels(sce_str$celltypes)[
    which(levels(sce_str$celltypes) %in% base::unique(sce_str$celltypes))])
```

```{r colors, meassage = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])
```

# Preparation 

Put all dfs containing info on conservation level per cell type together.

```{r}
cts_str <- as.list(names(geneset_list_str))
cts_hsc <- as.list(names(geneset_list_hsc))

conserved_df_list_str <- lapply(cts_str, function(c){
  
  geneset_list_str[[c]]$conserved_df$cell_type <- c
  geneset_list_str[[c]]$conserved_df$fraction <- "str"
  return(geneset_list_str[[c]]$conserved_df)
})

conserved_df_list_hsc <- lapply(cts_hsc, function(c){
  
  geneset_list_hsc[[c]]$conserved_df$cell_type <- c
  geneset_list_hsc[[c]]$conserved_df$fraction <- "hsc"
  return(geneset_list_hsc[[c]]$conserved_df)
})

conserved_df_str <- dplyr::bind_rows(conserved_df_list_str)
conserved_df_hsc <- dplyr::bind_rows(conserved_df_list_hsc)
cons_large_df <- BiocGenerics::rbind(conserved_df_str, conserved_df_hsc)
print(head(cons_large_df))
```

# Upset plots

Make one df specifically for upset plot

```{r prepare_upset}
cons_large_df_up <- cons_large_df
cons_large_df_up$mmus[which(!is.na(cons_large_df_up$mmus))] <- TRUE
cons_large_df_up$mcas[which(!is.na(cons_large_df_up$mcas))] <- TRUE
cons_large_df_up$mspr[which(!is.na(cons_large_df_up$mspr))] <- TRUE
cons_large_df_up$mcar[which(!is.na(cons_large_df_up$mcar))] <- TRUE
cons_large_df_up$mmus[which(is.na(cons_large_df_up$mmus))] <- FALSE
cons_large_df_up$mcas[which(is.na(cons_large_df_up$mcas))] <- FALSE
cons_large_df_up$mspr[which(is.na(cons_large_df_up$mspr))] <- FALSE
cons_large_df_up$mcar[which(is.na(cons_large_df_up$mcar))] <- FALSE

head(cons_large_df_up)
base::table(cons_large_df_up$conserved_signature)

base::table(base::duplicated(cons_large_df_up[
  cons_large_df_up$ndge == TRUE,]))
base::table(base::duplicated(cons_large_df_up[
  cons_large_df_up$conserved_marker == TRUE,]))
base::table(base::duplicated(cons_large_df_up[
  cons_large_df_up$conserved_signature == TRUE,]))
```

```{r upset}
ComplexUpset::upset(cons_large_df_up[which(
  cons_large_df_up$ndge == TRUE & cons_large_df_up$fraction == "str"),], 
  intersect = c("mmus", "mcas", "mspr",  "mcar"))+
  xlab("")+
  ggplot2::ggtitle("all marker genes, Stromal cells")

ComplexUpset::upset(cons_large_df_up[which(
  cons_large_df_up$ndge == TRUE & cons_large_df_up$fraction == "hsc"),], 
  intersect = c("mmus", "mcas", "mspr",  "mcar"))+
  xlab("")+
  ggplot2::ggtitle("all marker genes, HSPCs")
```

# Check Proportion of cells

That express certain genes.

### HSCs

```{r perc_cells_hsc}

hsc_conserved_signature <- base::unique(conserved_df_hsc$gene[
  conserved_df_hsc$conserved_signature == TRUE])
hsc_conserved_markers <- base::unique(conserved_df_hsc$gene[
  conserved_df_hsc$conserved_marker == TRUE])
hsc_ndges <- base::unique(conserved_df_hsc$gene[
  conserved_df_hsc$ndge == TRUE])

# use own function to get the proportion of cells a gene is expressed in
prop_expr_df_hsc <- prop_expressed_total_sce(
  sce = sce_hsc,
  geneset = c(base::unique(c(hsc_conserved_signature,
                             hsc_conserved_markers,
                             hsc_ndges))))

prop_expr_df_hsc_sign <- prop_expr_df_hsc[prop_expr_df_hsc$gene %in%
                                            hsc_conserved_signature,]
prop_expr_df_hsc_mark <- prop_expr_df_hsc[prop_expr_df_hsc$gene %in%
                                            hsc_conserved_markers,] 
prop_expr_df_hsc_ndge <- prop_expr_df_hsc[prop_expr_df_hsc$gene %in%
                                            hsc_ndges,]

prop_expr_df_hsc_sign$type <- "conserved_signature"
prop_expr_df_hsc_mark$type <- "conserved_markers"
prop_expr_df_hsc_ndge$type <- "ndges"

prop_expr_df_hsc <- base::rbind(prop_expr_df_hsc_sign,
                                prop_expr_df_hsc_mark,
                                prop_expr_df_hsc_ndge)

plot <- ggplot2::ggplot(prop_expr_df_hsc,
                        aes(x = type, y = prop_cells))+
  ggplot2::geom_jitter(color = "grey70")+
  ggplot2::geom_boxplot(alpha = 0, color = "black")+
  ggplot2::theme_classic()
plot

print(base::min(prop_expr_df_hsc$prop_cells))
print(base::min(prop_expr_df_hsc_sign$prop_cells))
print(base::min(prop_expr_df_hsc_mark$prop_cells))
print(base::min(prop_expr_df_hsc_ndge$prop_cells))
```

### Stromal

```{r perc_cells_str}

str_conserved_signature <- base::unique(conserved_df_str$gene[
  conserved_df_str$conserved_signature == TRUE])
str_conserved_markers <- base::unique(conserved_df_str$gene[
  conserved_df_str$conserved_marker == TRUE])
str_ndges <- base::unique(conserved_df_str$gene[
  conserved_df_str$ndge == TRUE])

prop_expr_df_str <- prop_expressed_total_sce(
  sce = sce_str,
  geneset = c(base::unique(c(str_conserved_signature,
                             str_conserved_markers,
                             str_ndges))))

prop_expr_df_str_sign <- prop_expr_df_str[prop_expr_df_str$gene %in%
                                            str_conserved_signature,]
prop_expr_df_str_mark <- prop_expr_df_str[prop_expr_df_str$gene %in%
                                            str_conserved_markers,] 
prop_expr_df_str_ndge <- prop_expr_df_str[prop_expr_df_str$gene %in%
                                            str_ndges,]

prop_expr_df_str_sign$type <- "conserved_signature"
prop_expr_df_str_mark$type <- "conserved_markers"
prop_expr_df_str_ndge$type <- "ndges"

prop_expr_df_str <- base::rbind(prop_expr_df_str_sign,
                                prop_expr_df_str_mark,
                                prop_expr_df_str_ndge)

plot <- ggplot2::ggplot(prop_expr_df_str,
                        aes(x = type, y = prop_cells))+
  ggplot2::geom_jitter(color = "grey70")+
  ggplot2::geom_boxplot(alpha = 0, color = "black")+
  ggplot2::theme_classic()
plot

print(base::min(prop_expr_df_str$prop_cells))
print(base::min(prop_expr_df_str_sign$prop_cells))
print(base::min(prop_expr_df_str_mark$prop_cells))
print(base::min(prop_expr_df_str_ndge$prop_cells))
```

# Categorize

## Total markers

Make a df with categories of conservation for visualisation.

```{r prepare_ggdf_category}
ggdf_category <- cons_large_df

ggdf_category$marker_only <- base::rep(FALSE, nrow(ggdf_category))
ggdf_category$conserved_marker_only <- base::rep(FALSE, nrow(ggdf_category))
ggdf_category$conserved_signature_only <- base::rep(FALSE, nrow(ggdf_category))

# genes that are only non-conserved markers
ggdf_category$marker_only[
  which(ggdf_category$conserved_marker == FALSE &
          ggdf_category$conserved_signature == FALSE)] <- TRUE
base::table(ggdf_category$marker_only)

# genes that are conserved markers, but not part of a conserved signature
# because they were not nDGE
ggdf_category$conserved_marker_only[
  which(ggdf_category$conserved_marker == TRUE &
          ggdf_category$conserved_signature == FALSE)] <- TRUE
base::table(ggdf_category$conserved_marker_only)

# genes that are part of a conserved signature
ggdf_category$conserved_signature_only[
  which(ggdf_category$conserved_marker_only == FALSE &
          ggdf_category$conserved_signature == TRUE)] <- TRUE
base::table(ggdf_category$conserved_signature_only)

# pivot longer by category
ggdf_category_long <- base::as.data.frame(tidyr::pivot_longer(
  ggdf_category,
  cols = c("marker_only",
           "conserved_marker_only",
           "conserved_signature_only"), 
  names_to = "conservation_level", 
  values_to = "is_true"))

# remove unneccessary rows that have been generated by pivot longer
ggdf_category_long <- ggdf_category_long[ggdf_category_long$is_true == TRUE,]

# rename and factor
ggdf_category_long$conservation_level <- base::gsub(
  "_only", "", ggdf_category_long$conservation_level)
ggdf_category_long$conservation_level <- factor(
  ggdf_category_long$conservation_level,
  levels = c("marker",
             "conserved_marker",
             "conserved_signature"))

ggdf_category_long$cell_type_rev <- factor(
  ggdf_category_long$cell_type,
  levels = base::rev(c(names(col_cts_hsc), names(col_cts_str))))
ggdf_category_long$cell_type <- factor(
  ggdf_category_long$cell_type,
  levels = c(names(col_cts_hsc), names(col_cts_str)))
```

```{r plot_category1}
plot_cat_1 <- ggplot2::ggplot(ggdf_category_long,
                              aes(fill = conservation_level, 
                                  y = cell_type_rev))+
  ggplot2::geom_bar(position = "stack")+
  ggplot2::theme_classic()+
  ggplot2::xlab("Nr of detected marker genes")+
  ggplot2::scale_fill_manual("", values = col_mode)+
  ggplot2::theme(strip.text.x = element_text(angle = 90),
                 axis.text.x = element_text(angle = 90),
                 legend.position = "bottom")+
  ggplot2::ylab("Cell type")
plot_cat_1
```

## Per species

```{r prepare_ggdf_category_species}
# pivot longer by species
ggdf_category_long_species <- base::as.data.frame(tidyr::pivot_longer(
  ggdf_category_long,
  cols = c("mmus",
           "mcas",
           "mspr",
           "mcar"), 
  names_to = "species", 
  values_to = "enrichment"))

# remove  unneccessary rows that have been generated by pivot longer
ggdf_category_long_species <- ggdf_category_long_species[
  !is.na(ggdf_category_long_species$enrichment),]

# factorize species
ggdf_category_long_species$species <- factor(
  ggdf_category_long_species$species,
  levels = names(col_spc))
# factorize species
ggdf_category_long_species$species_rev <- factor(
  ggdf_category_long_species$species,
  levels = base::rev(names(col_spc)))
```

```{r plot_category2, fig.width = 5.5, fig.height = 7}
plot_cat2 <- ggplot2::ggplot(ggdf_category_long_species[
  ggdf_category_long_species$cell_type %in% names(col_cts_hsc),],
  aes(fill = conservation_level, y = species_rev))+
  ggplot2::geom_bar(position = "stack")+
  ggplot2::theme_classic()+
  ggplot2::xlab("Nr of detected marker genes")+
  ggplot2::xlim(c(0,450))+
  ggplot2::scale_fill_manual(values = col_mode)+
  ggplot2::facet_grid(rows = vars(cell_type), switch = "y")+
  ggplot2::scale_y_discrete(position = "right")+
  ggplot2::theme(
    legend.position = "top",
    legend.title = element_blank(),
    strip.text.y.left = element_text(angle = 0),
    axis.text.y = element_text(angle = 0),
    axis.title.y = element_blank(),
    strip.text = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    strip.background = element_rect(fill = "grey97", linetype = 0))

plot_cat2
```

```{r, fig.width = 5.5, fig.height = 7}
plot_cat3 <- plot_cat2+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.text.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "none") 
plot_cat3
```

```{r plot_category4, fig.width = 5.6, fig.height = 4.5}
plot_cat4 <- ggplot2::ggplot(ggdf_category_long_species[
  ggdf_category_long_species$cell_type %in% names(col_cts_str),],
  aes(fill = conservation_level, y = species_rev))+
  ggplot2::geom_bar(position = "stack")+
  ggplot2::theme_classic()+
  ggplot2::xlab("Nr of detected marker genes")+
  ggplot2::scale_fill_manual(values = col_mode)+
  ggplot2::facet_grid(rows = vars(cell_type), switch = "y")+
  ggplot2::xlim(c(0,450))+
  ggplot2::scale_y_discrete(position = "right")+
  ggplot2::theme(
    legend.position = "none",
    strip.text.y.left = element_text(angle = 0),
    axis.text.y = element_text(angle = 0),
    axis.title.y = element_blank(),
    strip.text = element_text(size = 14),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    strip.background = element_rect(fill = "grey97", linetype = 0))

plot_cat4
```

```{r plot_category5, fig.width = 5.6, fig.height = 4.5}
plot_cat5 <- plot_cat4+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.text.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "none") 
plot_cat5
```

## Is it signature or not?

```{r signature_true}
ggdf_issign <- ggdf_category_long_species

ggdf_issign$is_signature <- vector(length = nrow(ggdf_issign))
ggdf_issign$is_signature[
  ggdf_issign$conservation_level == "marker"] <- FALSE
ggdf_issign$is_signature[
  ggdf_issign$conservation_level == "conserved_marker"] <- FALSE
ggdf_issign$is_signature[
  ggdf_issign$conservation_level == "conserved_signature"] <- TRUE

ggdf_issign$is_signature <- factor(ggdf_issign$is_signature, 
                                   levels =c(FALSE, TRUE))
```

```{r plot_signature_true1, fig.width = 6, fig.height = 4}
ggplot2::ggplot(ggdf_issign[ggdf_issign$species == "mmus",],
       aes(fill = is_signature, x = cell_type))+
  ggplot2::geom_bar(position = "stack")+
  ggplot2::theme_classic()+
  ggplot2::ylab("Nr of detected marker genes")+
  ggplot2::scale_fill_manual("Conserved signature", values = col_mode_log)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle("BL6 marker genes")+
  ggplot2::xlab("Cell type")

ggplot2::ggplot(ggdf_issign[ggdf_issign$species == "mmus",],
       aes(fill = is_signature, x = cell_type))+
  ggplot2::geom_bar(position = "fill")+
  ggplot2::theme_classic()+
  ggplot2::ylab("Proportion of conserved signature genes")+
  ggplot2::scale_fill_manual("Conserved signature", values = col_mode_log)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle("BL6 marker genes")+
  ggplot2::xlab("Cell type")
```

# Signature gene expression

## All together

```{r prepare_sces}
cts_hsc <- as.list(names(geneset_list_hsc))
cts_str <- as.list(names(geneset_list_str))

# this may not be the most elegant solution and it takes quite long
# separate SCE objects by cell type, then aggregate the counts to obtain
# the average counts per species for each cell type separately
# use the multi-batchnorm-corrected logcounts (for now)
agg_hsc_list <- lapply(cts_hsc, function(ct){
  
  print(ct)
  sce_ct <- sce_hsc[,sce_hsc$celltypes == ct]
  
  agg_hsc <- scuttle::aggregateAcrossCells(
    sce_ct, 
    id=colData(sce_ct)[,c("Species_ID")], 
    statistics = "mean",
    use.assay.type = "logcounts") #multibatchnorm
  
  return(agg_hsc)
})
names(agg_hsc_list) <- names(geneset_list_hsc)

agg_str_list <- lapply(cts_str, function(ct){
  
  print(ct)
  sce_ct <- sce_str[,sce_str$celltypes == ct]
  
  agg_str <- scuttle::aggregateAcrossCells(
    sce_ct, 
    id=colData(sce_ct)[,c("Species_ID")], 
    statistics = "mean",
    use.assay.type = "logcounts") #multibatchnorm
  
  return(agg_str)
})
names(agg_str_list) <- names(geneset_list_str)

```

### HSCs

```{r calc_expr_hsc_all}

# get all unique signature genes
all_sign_genes <- vector()
for(i in 1:length(geneset_list_hsc)){
  sign_genes <- geneset_list_hsc[[i]]$conserved_signature
  all_sign_genes <- c(all_sign_genes, sign_genes)
}
all_sign_genes_hsc <- base::unique(all_sign_genes)

expr_df_list_hsc <- lapply(agg_hsc_list, function(agg_hsc){
  
  # get current cell type and relevant signature genes
  ct_curr <- agg_hsc$celltypes[1]
  sign_genes_ct <- geneset_list_hsc[[ct_curr]]$conserved_signature
  
  # get a df from average logcounts per ct and species
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_hsc)$logcounts))
  
  # subset to all signature genes of this tissue
  vis_df <- vis_df[,colnames(vis_df) %in% all_sign_genes_hsc]
  
  # change df format 
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")
  
  # add ct info
  vis_df$cell_type <- base::rep(ct_curr)
  
  # scale expression of all genes per cell type (which genes are highest 
  # or lowest per cell type)
  vis_df$scaled_expression <- base::scale(vis_df$expression)
  
  # add info on genes that are signature genes of this ct
  vis_df$is_sign_gene <- vector(length = nrow(vis_df))
  vis_df$is_sign_gene[vis_df$gene %in% sign_genes_ct] <- TRUE
  
  vis_df <- vis_df[base::order(vis_df$scaled_expression, decreasing = TRUE),]

  return(vis_df)
})

# bind dfs 
agg_hsc_df <- dplyr::bind_rows(expr_df_list_hsc)

# add "sample"
agg_hsc_df <- agg_hsc_df %>%
  mutate("sample" = base::paste0(cell_type, "_", species))
head(agg_hsc_df)

# get a well ordered list of unique genes for factorization
all_sign_genes_for_factor <- base::unique(agg_hsc_df$gene[
  agg_hsc_df$is_sign_gene == TRUE])

# factorize
agg_hsc_df$gene <- factor(agg_hsc_df$gene,
                          levels = base::rev(all_sign_genes_for_factor))
agg_hsc_df$species <- factor(agg_hsc_df$species,
                             levels = names(col_spc))

# scale expression per gene (in which cell type is a gene highes or lowest
# expressed)
agg_hsc_df$scaled_per_gene <- vector(length = nrow(agg_hsc_df), mode = "numeric")
agg_hsc_df$zscore_gene <- vector(length = nrow(agg_hsc_df), mode = "numeric")
for(g in base::unique(agg_hsc_df$gene)){
  
  expr_vector_g <- agg_hsc_df[agg_hsc_df$gene == g,]$expression
  
  # scale expression levels per gene
  agg_hsc_df$scaled_per_gene[agg_hsc_df$gene == g] <- base::scale(expr_vector_g)
  
  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/stats::sd(expr_vector_g)
  agg_hsc_df$zscore_gene[agg_hsc_df$gene == g] <- z_score_g
  
}
```

```{r hsc_hm_all1}
ggplot2::ggplot(agg_hsc_df,
                aes(x = species, y = gene, fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "red3",
                                 limits = c(0, max(agg_hsc_df$expression)))+
  ggplot2::facet_grid(cols = vars(agg_hsc_df$cell_type))

```

```{r hsc_hm_all2}
ggplot2::ggplot(agg_hsc_df,
                aes(x = species, y = gene, fill = scaled_expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("scaled per cell type", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-max(agg_hsc_df$scaled_expression),
                                            0,
                                            max(agg_hsc_df$scaled_expression)), 
                                 limits = c(-max(agg_hsc_df$scaled_expression),
                                            max(agg_hsc_df$scaled_expression)))+
  ggplot2::facet_grid(cols = vars(agg_hsc_df$cell_type))

```

```{r hsc_hm_all3}
ggplot2::ggplot(agg_hsc_df,
                aes(x = species, y = gene, fill = scaled_per_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("scaled per gene", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-5, -2.5,  0, 2.5, 5), 
                                 limits = c(-5,5))+
  ggplot2::facet_grid(cols = vars(agg_hsc_df$cell_type))

```

```{r hsc_hm_all4}
ggplot2::ggplot(agg_hsc_df,
                aes(x = species, y = gene, fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("z-scores per gene", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-5, -2.5,  0, 2.5, 5), 
                                 limits = c(-5,5))+
  ggplot2::facet_grid(cols = vars(agg_hsc_df$cell_type))

```

### Stromal cells

```{r calc_expr_str_all}

all_sign_genes <- vector()
for(i in 1:length(geneset_list_str)){
  sign_genes <- geneset_list_str[[i]]$conserved_signature
  all_sign_genes <- c(all_sign_genes, sign_genes)
}
all_sign_genes_str <- base::unique(all_sign_genes)


expr_df_list_str <- lapply(agg_str_list, function(agg_str){
  
  ct_curr <- agg_str$celltypes[1]
  ct_curr_pos <- which(names(geneset_list_str) == ct_curr)
  sign_genes_ct <- geneset_list_str[[ct_curr_pos]]$conserved_signature
  
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str)$logcounts))
  
  vis_df <- vis_df[,colnames(vis_df) %in% all_sign_genes_str]
  
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")
  
  vis_df$cell_type <- base::rep(ct_curr)
  vis_df$scaled_expression <- base::scale(vis_df$expression)
  
  vis_df$is_sign_gene <- vector(length = nrow(vis_df))
  vis_df$is_sign_gene[vis_df$gene %in% sign_genes_ct] <- TRUE
  
  vis_df <- vis_df[base::order(vis_df$scaled_expression, decreasing = TRUE),]

  return(vis_df)
})

agg_str_df <- dplyr::bind_rows(expr_df_list_str)

# add "sample"
agg_str_df <- agg_str_df %>%
  mutate("sample" = base::paste0(cell_type, "_", species))
head(agg_str_df)

# get a well ordered list of unique genes for factorization
all_sign_genes_for_factor <- base::unique(agg_str_df$gene[
  agg_str_df$is_sign_gene == TRUE])

agg_str_df$gene <- factor(agg_str_df$gene,
                          levels = base::rev(all_sign_genes_for_factor))
agg_str_df$species <- factor(agg_str_df$species,
                             levels = names(col_spc))

# scale expression per gene (in which cell type is a gene highes or lowest
# expressed)
agg_str_df$scaled_per_gene <- vector(length = nrow(agg_str_df), mode = "numeric")
agg_str_df$zscore_gene <- vector(length = nrow(agg_str_df), mode = "numeric")
for(g in base::unique(agg_str_df$gene)){
  
  expr_vector_g <- agg_str_df[agg_str_df$gene == g,]$expression
  
  # scale expression levels per gene
  agg_str_df$scaled_per_gene[agg_str_df$gene == g] <- base::scale(expr_vector_g)
  
  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/stats::sd(expr_vector_g)
  agg_str_df$zscore_gene[agg_str_df$gene == g] <- z_score_g
  
}
```

```{r str_hm_all1}
ggplot2::ggplot(agg_str_df,
                aes(x = species, y = gene, fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "red3",
                                 limits = c(0, max(agg_str_df$expression)))+
  ggplot2::facet_grid(cols = vars(agg_str_df$cell_type))

```

```{r str_hm_all2}
ggplot2::ggplot(agg_str_df,
                aes(x = species, y = gene, fill = scaled_expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("scaled per cell type", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-max(agg_str_df$scaled_expression),
                                            0,
                                            max(agg_str_df$scaled_expression)), 
                                 limits = c(-max(agg_str_df$scaled_expression),
                                            max(agg_str_df$scaled_expression)))+
  ggplot2::facet_grid(cols = vars(agg_str_df$cell_type))

```

```{r str_hm_all3}
ggplot2::ggplot(agg_str_df,
                aes(x = species, y = gene, fill = scaled_per_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("scaled per gene", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-4, -2,  0, 2, 4), 
                                 limits = c(-4.5,4.5)
                                )+
  ggplot2::facet_grid(cols = vars(agg_str_df$cell_type))

```

```{r str_hm_all4}
ggplot2::ggplot(agg_str_df,
                aes(x = species, y = gene, fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank(),
                 axis.text.x = element_text(angle = 90),
                 strip.text.x = element_text(angle = 90))+
  ggplot2::scale_fill_gradientn("z-scores per gene", 
                                 colours = c("blue3", "white", "red3"),
                                 breaks = c(-4, -2,  0, 2, 4), 
                                 limits = c(-4.5,4.5)
                                )+
  ggplot2::facet_grid(cols = vars(agg_str_df$cell_type))

```

## Per cell type

Generate plots for each cell type.

```{r plot_hsc_signatures1}

plotlist_hsc <- lapply(cts_hsc, function(ct){

  print(ct)
  hsc_genes <- geneset_list_hsc[[ct]]$conserved_signature
  
  hsc_genes_nosub <- hsc_genes[!hsc_genes %in%
                                 geneset_list_hsc[[ct]]$genes_subclustering]
  
  agg_hsc_ct <- agg_hsc_list[[ct]]
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_hsc_ct)$logcounts))
  vis_df <- vis_df[,colnames(vis_df) %in% hsc_genes]
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")

  vis_df$species <- factor(vis_df$species, levels = names(col_spc))
  
  # make plots
  plot1 <- ggplot2::ggplot(vis_df, 
                           aes(x = species, 
                               y = reorder(gene, expression), 
                               fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "blue2",
                                 limits = c(0, max(vis_df$expression)))+
  ggplot2::ggtitle(base::paste(ct, "markers"))+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "bottom",
                 axis.text.x = element_text(angle = 90, 
                                            size = 12,
                                            hjust = 0.45),
                 axis.text.y = element_text(size = 12))

  plotlist <- list(plot1)
  
  if(!identical(hsc_genes, hsc_genes_nosub)){
    vis_df <- vis_df[vis_df$gene %in% hsc_genes_nosub,]
    vis_df <- vis_df[order(vis_df$expression),]
    
    plot2 <- ggplot2::ggplot(vis_df, 
                             aes(x = species, 
                                 y = reorder(gene, expression), 
                                 fill = expression))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::scale_fill_continuous("average logcounts",
                                   low = "white", high = "blue2",
                                   limits = c(0, max(vis_df$expression)))+
    ggplot2::ggtitle(base::paste(ct, "markers"))+
    ggplot2::theme(axis.title.y = element_blank(),
                   axis.title.x = element_blank(),
                   legend.position = "bottom",
                   axis.text.x = element_text(angle = 90, 
                                              size = 12,
                                              hjust = 0.45),
                   axis.text.y = element_text(size = 12))
    plotlist <- list(plot1, plot2)
  }
  return(plotlist)
})
names(plotlist_hsc) <- names(geneset_list_hsc)
```

```{r plot_expr_hsc, fig.height = 6, fig.width = 4}
print(plotlist_hsc[["HSCs"]]) 
```

```{r plot_expr_earlympps, fig.height = 5.5, fig.width = 4}
print(plotlist_hsc[["Early MPPs"]])
```

```{r plot_expr_activatedympps, fig.height = 3, fig.width = 4}
print(plotlist_hsc[["Activated MPPs"]])
```

```{r plot_expr_earlymyeloid, fig.height = 5, fig.width = 4}
print(plotlist_hsc[["Early Myeloid"]])
```

```{r plot_expr_lymphoid, fig.height = 5, fig.width = 4}
print(plotlist_hsc[["Lymphoid"]])
```

```{r plot_expr_neutro, fig.height = 10, fig.width = 4}
print(plotlist_hsc[["Neutrophil progs."]])
```

```{r plot_expr_mono, fig.height = 10, fig.width = 4}
print(plotlist_hsc[["Monocyte progs."]])
```

```{r plot_expr_mk, fig.height = 4, fig.width = 4}
print(plotlist_hsc[["Mk progs."]])
```

```{r plot_expr_mkery, fig.height = 4, fig.width = 4}
print(plotlist_hsc[["Mk/Ery progs."]])
```

```{r plot_expr_bmprogs, fig.height = 4, fig.width = 4}
print(plotlist_hsc[["Baso/Mast progs."]])
```

```{r plot_expr_ery, fig.height = 15, fig.width = 4}
print(plotlist_hsc[["Erythroid progs."]])
```

```{r plot_expr_cycling, fig.height = 12, fig.width = 4}
print(plotlist_hsc[["Cycling"]])
```

```{r plot_str_signatures}
cts_str <- as.list(names(geneset_list_str))

plotlist_str <- lapply(cts_str, function(ct){
  
  print(ct)
  str_genes <- geneset_list_str[[ct]]$conserved_signature
  
  str_genes_nosub <- str_genes[!str_genes %in%
                                 geneset_list_str[[ct]]$genes_subclustering]
  
  agg_str_ct <- agg_str_list[[ct]]
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str_ct)$logcounts))
  vis_df <- vis_df[,colnames(vis_df) %in% str_genes]
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")

  vis_df$species <- factor(vis_df$species, levels = names(col_spc))
  
  # make plots
  plot1 <- ggplot2::ggplot(vis_df, 
                           aes(x = species, 
                               y = reorder(gene, expression), 
                               fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "blue2",
                                 limits = c(0, max(vis_df$expression)))+
  ggplot2::ggtitle(base::paste(ct, "markers"))+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "bottom",
                 axis.text.x = element_text(angle = 90, 
                                            size = 12,
                                            hjust = 0.45),
                 axis.text.y = element_text(size = 12))

  plotlist <- list(plot1)
  
  if(!identical(str_genes, str_genes_nosub)){
    vis_df <- vis_df[vis_df$gene %in% str_genes_nosub,]
    vis_df <- vis_df[order(vis_df$expression),]
    
    plot2 <- ggplot2::ggplot(vis_df, 
                             aes(x = species, 
                                 y = reorder(gene, expression), 
                                 fill = expression))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::scale_fill_continuous("average logcounts",
                                   low = "white", high = "blue2",
                                   limits = c(0, max(vis_df$expression)))+
    ggplot2::ggtitle(base::paste(ct, "markers"))+
    ggplot2::theme(axis.title.y = element_blank(),
                   axis.title.x = element_blank(),
                   legend.position = "bottom",
                   axis.text.x = element_text(angle = 90, 
                                              size = 12,
                                              hjust = 0.45),
                   axis.text.y = element_text(size = 12))
  plotlist <- list(plot1, plot2)
  }
  return(plotlist)
})
names(plotlist_str) <- names(geneset_list_str)
```

```{r plot_expr_adipocars, fig.height = 17, fig.width = 4}
print(plotlist_str[["Adipo/CARs"]])
```

```{r plot_expr_osteo, fig.height = 10, fig.width = 4}
print(plotlist_str[["Osteo"]])
```

```{r plot_expr_fibrochondro, fig.height = 10, fig.width = 4}
print(plotlist_str[["Fibro/Chondro"]])
```

```{r plot_expr_fibro, fig.height = 18, fig.width = 4}
print(plotlist_str[["Fibroblasts"]])
```

```{r plot_expr_aec, fig.height = 25, fig.width = 4}
print(plotlist_str[["Arteriolar/Capillary ECs"]])
```

```{r plot_expr_svec, fig.height = 15, fig.width = 4}
print(plotlist_str[["Sinusoidal/Venular ECs"]])
```

```{r plot_expr_sinusoid, fig.height = 15, fig.width = 4}
print(plotlist_str[["Sinusoidal ECs"]])
```

```{r plot_expr_perismooth, fig.height = 10, fig.width = 4}
print(plotlist_str[["Pericytes/Smooth Muscle"]])
```

# All BL6 marker gene expression

Generate plots for each cell type.
This will help us find interesting genes to validate using RNA-FISH (HCR).

```{r plot_hsc_bl6m1}

plotlist_hsc_bl6_markers <- lapply(cts_hsc, function(ct){

  print(ct)
  # get all BL6 markers per cell type
  mmms_genes <- geneset_list_hsc[[ct]]$conserved_df$gene[
    which(!is.na(geneset_list_hsc[[ct]]$conserved_df$mmus))]
 
  mmms_genes_nosub <- mmms_genes[!mmms_genes %in%
                                 geneset_list_hsc[[ct]]$genes_subclustering]
  
  agg_hsc_ct <- agg_hsc_list[[ct]]
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_hsc_ct)$logcounts))
  vis_df <- vis_df[,colnames(vis_df) %in% mmms_genes]
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")

  vis_df$species <- factor(vis_df$species, levels = names(col_spc))
  
  # make plots
  plot1 <- ggplot2::ggplot(vis_df, 
                           aes(x = species, 
                               y = reorder(gene, expression), 
                               fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "blue2",
                                 limits = c(0, max(vis_df$expression)))+
  ggplot2::ggtitle(base::paste(ct, "markers"))+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "bottom",
                 axis.text.x = element_text(angle = 90, 
                                            size = 12,
                                            hjust = 0.45),
                 axis.text.y = element_text(size = 12))

  plotlist <- list(plot1)
  
  if(!identical(mmms_genes, mmms_genes_nosub)){
    vis_df <- vis_df[vis_df$gene %in% mmms_genes_nosub,]
    vis_df <- vis_df[order(vis_df$expression),]
    
    plot2 <- ggplot2::ggplot(vis_df, 
                             aes(x = species, 
                                 y = reorder(gene, expression), 
                                 fill = expression))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::scale_fill_continuous("average logcounts",
                                   low = "white", high = "blue2",
                                   limits = c(0, max(vis_df$expression)))+
    ggplot2::ggtitle(base::paste(ct, "markers"))+
    ggplot2::theme(axis.title.y = element_blank(),
                   axis.title.x = element_blank(),
                   legend.position = "bottom",
                   axis.text.x = element_text(angle = 90, 
                                              size = 12,
                                              hjust = 0.45),
                   axis.text.y = element_text(size = 12))
    plotlist <- list(plot1, plot2)
  }
  return(plotlist)
})
names(plotlist_hsc_bl6_markers) <- names(geneset_list_hsc)
```

```{r plot_expr_hsc_bl6, fig.height = 13, fig.width = 4}
print(plotlist_hsc_bl6_markers[["HSCs"]]) 
```

```{r plot_expr_earlympps_bl6, fig.height = 6, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Early MPPs"]])
```

```{r plot_expr_activatedympps_bl6, fig.height = 7, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Activated MPPs"]])
```

```{r plot_expr_earlymyeloid_bl6, fig.height = 13, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Early Myeloid"]])
```

```{r plot_expr_lymphoid_bl6, fig.height = 17, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Lymphoid"]])
```

```{r plot_expr_neutro_bl6, fig.height = 35, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Neutrophil progs."]])
```

```{r plot_expr_mono_bl6, fig.height = 35, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Monocyte progs."]])
```

```{r plot_expr_mk_bl6, fig.height = 10, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Mk progs."]])
```

```{r plot_expr_mkery_bl6, fig.height = 14, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Mk/Ery progs."]])
```

```{r plot_expr_bmprogs_bl6, fig.height = 14, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Baso/Mast progs."]])
```

```{r plot_expr_ery_bl6, fig.height = 40, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Erythroid progs."]])
```

```{r plot_expr_cycling_bl6, fig.height = 20, fig.width = 4}
print(plotlist_hsc_bl6_markers[["Cycling"]])
```

```{r plot_str_bl6}
cts_str <- as.list(names(geneset_list_str))

plotlist_allbl6_markers_str <- lapply(cts_str, function(ct){
  
  print(ct)
   # get all BL6 markers per cell type
  mmms_genes <- geneset_list_str[[ct]]$conserved_df$gene[
    which(!is.na(geneset_list_str[[ct]]$conserved_df$mmus))]
 
  mmms_genes_nosub <- mmms_genes[!mmms_genes %in%
                                 geneset_list_str[[ct]]$genes_subclustering]
  
  agg_str_ct <- agg_str_list[[ct]]
  vis_df <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str_ct)$logcounts))
  vis_df <- vis_df[,colnames(vis_df) %in% mmms_genes]
  vis_df <- tibble::rownames_to_column(vis_df, "species")
  vis_df <- tidyr::pivot_longer(vis_df, 
                                cols = c(2:ncol(vis_df)),
                                values_to = "expression",
                                names_to = "gene")

  vis_df$species <- factor(vis_df$species, levels = names(col_spc))
  
  # make plots
  plot1 <- ggplot2::ggplot(vis_df, 
                           aes(x = species, 
                               y = reorder(gene, expression), 
                               fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_continuous("average logcounts",
                                 low = "white", high = "blue2",
                                 limits = c(0, max(vis_df$expression)))+
  ggplot2::ggtitle(base::paste(ct, "markers"))+
  ggplot2::theme(axis.title.y = element_blank(),
                 axis.title.x = element_blank(),
                 legend.position = "bottom",
                 axis.text.x = element_text(angle = 90, 
                                            size = 12,
                                            hjust = 0.45),
                 axis.text.y = element_text(size = 12))

  plotlist <- list(plot1)
  
  if(!identical(mmms_genes, mmms_genes_nosub)){
    vis_df <- vis_df[vis_df$gene %in% mmms_genes_nosub,]
    vis_df <- vis_df[order(vis_df$expression),]
    
    plot2 <- ggplot2::ggplot(vis_df, 
                             aes(x = species, 
                                 y = reorder(gene, expression), 
                                 fill = expression))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::scale_fill_continuous("average logcounts",
                                   low = "white", high = "blue2",
                                   limits = c(0, max(vis_df$expression)))+
    ggplot2::ggtitle(base::paste(ct, "markers"))+
    ggplot2::theme(axis.title.y = element_blank(),
                   axis.title.x = element_blank(),
                   legend.position = "bottom",
                   axis.text.x = element_text(angle = 90, 
                                              size = 12,
                                              hjust = 0.45),
                   axis.text.y = element_text(size = 12))
  plotlist <- list(plot1, plot2)
  }
  return(plotlist)
})
names(plotlist_allbl6_markers_str) <- names(geneset_list_str)
```

```{r plot_expr_adipocars_bl6, fig.height = 60, fig.width = 4}
print(plotlist_allbl6_markers_str[["Adipo/CARs"]])
```

```{r plot_expr_osteo_bl6, fig.height = 45, fig.width = 4}
print(plotlist_allbl6_markers_str[["Osteo"]])
```

```{r plot_expr_fibrochondro_bl6, fig.height = 50, fig.width = 4}
print(plotlist_allbl6_markers_str[["Fibro/Chondro"]])
```

```{r plot_expr_fibro_bl6, fig.height = 68, fig.width = 4}
print(plotlist_allbl6_markers_str[["Fibroblasts"]])
```

```{r plot_expr_aec_bl6, fig.height = 70, fig.width = 4}
print(plotlist_allbl6_markers_str[["Arteriolar/Capillary ECs"]])
```

```{r plot_expr_svec_bl6, fig.height = 50, fig.width = 4}
print(plotlist_allbl6_markers_str[["Sinusoidal/Venular ECs"]])
```

```{r plot_expr_sinusoid_bl6, fig.height = 35, fig.width = 4}
print(plotlist_allbl6_markers_str[["Sinusoidal ECs"]])
```

```{r plot_expr_perismooth_bl6, fig.height = 44, fig.width = 4}
print(plotlist_allbl6_markers_str[["Pericytes/Smooth Muscle"]])
```

### HSCs only, special plot

Making one example plot for HSCs using all BL6 markers

```{r hsc_vis_df}

hsc_cons_df <- geneset_list_hsc$HSCs$conserved_df

agg_hscs <- agg_hsc_list$HSCs

vis_df_hsc <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg_hscs)$logcounts)) # batchnorm
vis_df_hsc <- vis_df_hsc[,colnames(vis_df_hsc) %in% hsc_cons_df$gene]
vis_df_hsc <- tibble::rownames_to_column(vis_df_hsc, "species")
vis_df_hsc <- tidyr::pivot_longer(vis_df_hsc, 
                                  cols = c(2:ncol(vis_df_hsc)),
                                  values_to = "expression",
                                  names_to = "gene")

vis_df_hsc$species <- factor(vis_df_hsc$species, levels = names(col_spc))

# scale expression per gene (in which cell type is a gene highes or lowest
# expressed)
vis_df_hsc$scaled_per_gene <- vector(length = nrow(vis_df_hsc), mode = "numeric")
vis_df_hsc$zscore_gene <- vector(length = nrow(vis_df_hsc), mode = "numeric")
for(g in base::unique(vis_df_hsc$gene)){
  
  expr_vector_g <- vis_df_hsc[vis_df_hsc$gene == g,]$expression
  
  vis_df_hsc$scaled_per_gene[vis_df_hsc$gene == g] <- base::scale(expr_vector_g)
  
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/stats::sd(expr_vector_g)
  vis_df_hsc$zscore_gene[vis_df_hsc$gene == g] <- z_score_g
  
}
```

```{r order_hsc_genes}
# order genes semi-manually for nice visualisation

# get gene sets of interest
hsc_conserved_signature <- geneset_list_hsc$HSCs$conserved_signature
hsc_bl6_specific <- hsc_cons_df$gene[which(!is.na(hsc_cons_df$mmus) &
                                             is.na(hsc_cons_df$mcas) &
                                             is.na(hsc_cons_df$mspr) &
                                             is.na(hsc_cons_df$mcar))]


# gene set 1 = bl6-specific markers
vis_df_hsc_bl6spec <- vis_df_hsc[vis_df_hsc$gene %in% hsc_bl6_specific,]
vis_df_hsc_bl6spec <- vis_df_hsc_bl6spec[
  base::order(vis_df_hsc_bl6spec$expression, decreasing = TRUE),]

bl6_spec_genes <- base::unique(vis_df_hsc_bl6spec$gene)
print(length(bl6_spec_genes))

# gene set 2 = neither
vis_df_hsc_nei <- vis_df_hsc[!vis_df_hsc$gene %in% c(hsc_bl6_specific,
                                                     hsc_conserved_signature),]
vis_df_hsc_nei <- vis_df_hsc_nei[
  base::order(vis_df_hsc_nei$expression, decreasing = TRUE),]
nei_genes <- base::unique(vis_df_hsc_nei$gene)
print(length(nei_genes))

# gene set 3 = signature
vis_df_hsc_sig <- vis_df_hsc[vis_df_hsc$gene %in% hsc_conserved_signature,]
vis_df_hsc_sig <- vis_df_hsc_sig[
  base::order(vis_df_hsc_sig$expression, decreasing = TRUE),]

sig_genes <- base::unique(vis_df_hsc_sig$gene)
print(length(sig_genes))

# bind genes, then use to factor
genes_for_factor_hsc <- c(sig_genes, nei_genes, bl6_spec_genes)

vis_df_hsc$gene <- factor(vis_df_hsc$gene, 
                          levels = base::rev(genes_for_factor_hsc))
```

```{r hsc_markers1, fig.width = 4, fig.height = 4}
ggplot2::ggplot(vis_df_hsc,
                aes(x = species, y = gene, fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn("average logcounts",
                                 colours = c("white",  "orange", "red3"))
```

```{r hsc_markers2, fig.width = 4, fig.height = 6}
ggplot2::ggplot(vis_df_hsc[vis_df_hsc$gene %in% c(
  sig_genes, bl6_spec_genes),],
                aes(x = species, y = gene, fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradientn("average logcounts",
                                 colours = c("white", "orange", "red3"))
```

```{r hsc_markers3, fig.width = 4, fig.height = 3}
ggplot2::ggplot(vis_df_hsc[vis_df_hsc$gene %in% c(
  sig_genes, bl6_spec_genes),],
                aes(x = species, y = gene, fill = expression))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn("average logcounts",
                                 colours = c("white", "orange", "red3"))
```

```{r hsc_markers4, fig.width = 4, fig.height = 3}
ggplot2::ggplot(vis_df_hsc[vis_df_hsc$gene %in% c(
  sig_genes, bl6_spec_genes),],
                aes(x = species, y = gene, fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.y = element_blank(),
                 axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn("zscore per gene",
                                 colours = c("blue3", "white", "red3"),
                                limits = c(-5, 5))
```

```{r sessioninfo}
utils::sessionInfo()
```