---
title: "Reclustering permutation report - geneset comparison"
author: "Lea Wölbert"
date: '2024-08-28'
output: html_document
---

Plotting the reclustered values for overview/quality assurance.
Looking at permutation of:

- conserved marker genes
- all BL6 marker genes

to determine if they perform better than the same number of random genes
containing the conserved signature or conserved marker gene set.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(Seurat, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r load_objects1}
# current fraction
fraction_curr <- snakemake@wildcards[["fraction"]] 
print(fraction_curr)
```

```{r load_objects2}
# permutation dataframe, containing all scores per iteration
perm_score_df_mark <- base::readRDS(snakemake@input[["perm_score_df_mark"]])
perm_score_df_mmms <- base::readRDS(snakemake@input[["perm_score_df_mmms"]])

# score dataframe, containing the calculated scores per re-clustered dataset
score_df <- base::readRDS(snakemake@input[["score_df"]])
```

```{r load_objects3}
score_df_mark <- score_df[score_df$conservation_level == "conserved_markers",]
score_df_mmms <- score_df[score_df$conservation_level == "mmusall_markers",]
score_df_sign <- score_df[score_df$conservation_level == "conserved_signature",]
```

```{r load_objects4}
# load ensembl list to know how many genes of each set were used 
geneset_list <- base::readRDS(snakemake@input[["geneset_list"]])
```

```{r prepare_genesets}
# get subclustering genes for control
subclustering_gene_list <- lapply(geneset_list, function(geneset){
  subclustering_genes <- geneset$genes_subclustering
  return(subclustering_genes)
})

subclustering_genes <- base::unique(unlist(subclustering_gene_list))
# get all signature genes (excluding subclustering genes)
conserved_signature_list <- lapply(geneset_list, function(geneset){
  conserved_signature <- geneset$conserved_signature
  return(conserved_signature)
})
sign_IDs <- base::unique(unlist(conserved_signature_list))
sign_IDs <- sign_IDs[!sign_IDs %in% subclustering_genes]
print(length(sign_IDs))

# get all conserved markers
conserved_markers_list <- lapply(geneset_list, function(geneset){
  conserved_markers <- geneset$conserved_markers
  return(conserved_markers)
})
mark_IDs <- base::unique(unlist(conserved_markers_list))
mark_IDs <- mark_IDs[!mark_IDs %in% subclustering_genes]
print(length(mark_IDs))

# get all BL6 marker IDs
mmus_marker_list <- lapply(geneset_list, function(geneset){
  mmus_markers <- geneset$conserved_df$gene[
    which(!is.na(geneset$conserved_df$mmus))]
  return(mmus_markers)
})
mmms_IDs <- base::unique(unlist(mmus_marker_list))
mmms_IDs <- mmms_IDs[!mmms_IDs %in% subclustering_genes]
print(length(mmms_IDs))

# get all ndges
ndge_list <- lapply(geneset_list, function(geneset){
  ndges <- geneset$ndges
  return(ndges)
})
ndges <- base::unique(unlist(ndge_list))
ndges <- ndges[!ndges %in% subclustering_genes]
print(length(ndges))

max_genes <- length(mmms_IDs)
```

```{r colors}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])
```

```{r it}
iterations <- nrow(perm_score_df_mmms)
```

```{r print}
print(head(score_df_mark))
print(head(score_df_mmms))
print(head(perm_score_df_mark))
print(head(perm_score_df_mmms))
```

# Conserved Markers vs Signature + Random

## Histograms

```{r perm_plots_mark}

scores <- c(
  "adjusted_rand_index",
  "mean_cluster_purity",
  "mean_prop_cells_cluster")

scores_v <- c(
  "adjusted_rand_index",
  "mean_prop_cells_cluster",
  "variation_information") # reclustering scores

for(score in scores_v){
  
  print(score)
  
  score_df_temp <- score_df_mark[score_df_mark$type == score,]
  perm_scores <- perm_score_df_mark[,which(
    colnames(perm_score_df_mark) == score)]
  
  if(score != "variation_information"){
    max <- 1
    pval = length(which(perm_scores > score_df_temp$value))/length(perm_scores)
    print(pval)
  
  }else{
    max <- base::max(c(score_df_temp$value, perm_scores))
    # for variation_information score, the smaller, the better
    pval = length(which(perm_scores < score_df_temp$value))/length(perm_scores)
  }
  
  plot1 <- ggplot2::ggplot(as.data.frame(perm_scores), aes(x = perm_scores))+
    ggplot2::geom_histogram(binwidth = 0.02)+
    ggplot2::geom_vline(xintercept = score_df_temp$value, 
                        color = col_cons["conserved_markers"])+
    ggplot2::theme_classic()+
    ggplot2::xlim(c(0, max))+
    ggplot2::ggtitle(base::paste(score, 
                                 fraction_curr, 
                                 iterations,
                                 "iterations,",
                                 perm_score_df_mark$cut_off_prop[1],
                                 "% of cells,",
                                 perm_score_df_mark$nr_genes_used,
                                 "genes used"))+
    annotate("text", 
             x = 0.8,
             y = 0.8, 
             label = base::paste0("pval = ", as.character(pval)))
  print(plot1)
}
```

## In two plots

All scores between 0 and 1

```{r all1, fig.width = 3.5, fig.height = 4}

score_df_temp <- score_df_mark[score_df_mark$type %in% scores,]
score_df_temp <- score_df_temp[
  score_df_temp$conservation_level == "conserved_markers",]
score_df_temp_sign <- score_df_sign[score_df_sign$type %in% scores,]
score_df_temp <- base::rbind(score_df_temp, score_df_temp_sign)

perm_score_df_temp <- pivot_longer(perm_score_df_mark,
                                   c = 1:ncol(perm_score_df_mark),
                                   values_to = "value",
                                   names_to = "type")
perm_score_df_temp <- perm_score_df_temp[perm_score_df_temp$type %in% scores,]

plot1 <- ggplot2::ggplot(score_df_temp, 
                aes(x = type, y = value))+
  ggplot2::geom_violin(data = perm_score_df_temp,
                       scale = "width",
                       position = position_nudge(x = 0.2, y = 0))+
  ggplot2::geom_point(size = 2,
                      alpha = 0.7,
                      aes(color = conservation_level),
                      position = position_nudge(x = -0.2, y = 0))+
  ggplot2::scale_color_manual("conservation level", values = col_cons)+
  ggplot2::ylim(c(0,1))+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(base::paste(fraction_curr, 
                               iterations,
                               "iterations,",
                               "\n",
                               perm_score_df_mark$cut_off_prop[1],
                               "% of cells,",
                               "\n",
                               perm_score_df_mark$nr_genes_used,
                               "genes used"))
```

```{r all2, fig.width = 3, fig.height = 3}

score_df_temp <- score_df_mark[
  score_df_mark$type %in% c("variation_information"),]
score_df_temp <- score_df_temp[
  score_df_temp$conservation_level == "conserved_markers",]
score_df_temp_sign <- score_df_sign[
  score_df_sign$type %in% c("variation_information"),]
score_df_temp <- rbind(score_df_temp, score_df_temp_sign)

perm_score_df_temp <- pivot_longer(perm_score_df_mark, 
                                   c = 1:ncol(perm_score_df_mark),
                                   values_to = "value",
                                   names_to = "type")
perm_score_df_temp <- perm_score_df_temp[
  perm_score_df_temp$type %in% c("variation_information"),]

plot2 <- ggplot2::ggplot(score_df_temp, aes(x = type, y = value))+
  ggplot2::geom_violin(data = perm_score_df_temp,
                       scale = "width", 
                       position = position_nudge(x = 0.2, y = 0))+
  ggplot2::geom_point(size = 2,
                      alpha = 0.7,
                      aes(color = conservation_level),
                      position = position_nudge(x = -0.2, y = 0))+
  ggplot2::ylim(c(0, base::max(c(perm_score_df_temp$value, 
                                 score_df_temp$value))))+
  ggplot2::scale_color_manual("conservation level", values = col_cons)+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle("")
```

```{r numbers_mark_sign}
print("cell types:")
print(score_df_mark$nr_celltypes[1])
print(score_df_sign$nr_celltypes[1])
print(base::table(perm_score_df_mark$nr_celltypes))

print("clusters:")
print(score_df_mark$nr_clusters[1])
print(score_df_sign$nr_clusters[1])
print(base::table(perm_score_df_mark$nr_clusters))
```


```{r plot_mark_sign_x, fig.width = 5, fig.height = 5}
plot <- ggpubr::ggarrange(
  plot1 + ggplot2::theme(legend.position = "none"), 
  plot2 + ggplot2::theme(plot.title = element_blank(),
                          axis.title.y = element_blank()),
  align = "h",
  widths = c(1, 1.5))
plot
```

## Nr of genes per gene set

Prepare identically to 05_permutation_genesets.R

```{r prepare_mark_nrs, fig.width = 4, fig.height = 5}

# get the number for each
nr_sign <- length(sign_IDs) 
nr_mark <- length(mark_IDs[!mark_IDs %in% sign_IDs])
nr_random_mark <- length(mark_IDs) - nr_sign

vis_df_mark <- data.frame(type1 = c("conserved signature", 
                                    "conserved markers", 
                                    "conserved markers",
                                    "conserved marker perm",
                                    "conserved marker perm"),
                          type2 = c("conserved_signature",
                                    "conserved_signature",
                                    "conserved_markers",
                                    "conserved_signature",
                                    "random_features"),
                          number = c(nr_sign, 
                                     nr_sign,
                                     nr_mark, 
                                     nr_sign,
                                     nr_random_mark))

vis_df_mark$type1 <- factor(
  vis_df_mark$type1, levels = c("conserved signature", 
                                "conserved markers",
                                "conserved marker perm"))
vis_df_mark$type2 <- factor(
  vis_df_mark$type2, levels = c("conserved_markers",
                                "random_features",
                                "conserved_signature"))

ggplot2::ggplot(vis_df_mark, aes(x = type1, y = number, fill = type2))+
  ggplot2::geom_col(position = position_stack())+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::xlab("conservation_level")+
  ggplot2::ylab("nr of genes")+
  ggplot2::scale_fill_manual("gene set", 
                             values = col_cons)+
  ggplot2::ylim(c(0, max_genes))
  
```

# All BL6 Markers vs Signature + Random

## Histograms

```{r perm_plots_mmms}

for(score in scores_v){
  
  print(score)
  
  score_df_temp <- score_df_mmms[score_df_mmms$type == score,]
  perm_scores <- perm_score_df_mmms[,which(
    colnames(perm_score_df_mmms) == score)]
  
  if(score != "variation_information"){
    max <- 1
    pval = length(which(perm_scores > score_df_temp$value))/length(perm_scores)
    print(pval)
  
  }else{
    max <- base::max(c(score_df_temp$value, perm_scores))
    # for variation_information score, the smaller, the better
    pval = length(which(perm_scores < score_df_temp$value))/length(perm_scores)
  }
  
  plot1 <- ggplot2::ggplot(as.data.frame(perm_scores), aes(x = perm_scores))+
    ggplot2::geom_histogram(binwidth = 0.02)+
    ggplot2::geom_vline(xintercept = score_df_temp$value, 
                        color = col_cons["mmusall_markers"])+
    ggplot2::theme_classic()+
    ggplot2::xlim(c(0, max))+
    ggplot2::ggtitle(base::paste(score, 
                                 fraction_curr, 
                                 iterations,
                                 "iterations,",
                                 perm_score_df_mmms$cut_off_prop[1],
                                 "% of cells,",
                                 perm_score_df_mmms$nr_genes_used,
                                 "genes used"))+
    annotate("text", 
             x = 0.8,
             y = 0.8, 
             label = base::paste0("pval = ", as.character(pval)))
  print(plot1)
}
```

## In two plots

All scores between 0 and 1

```{r all1_mmms, fig.width = 3.5, fig.height = 4.4}

score_df_temp <- score_df_mmms[score_df_mmms$type %in% scores,]
score_df_temp <- score_df_temp[
  score_df_temp$conservation_level == "mmusall_markers",]
score_df_temp_sign <- score_df_sign[score_df_sign$type %in% scores,]
score_df_temp <- rbind(score_df_temp, score_df_temp_sign)

perm_score_df_temp <- pivot_longer(perm_score_df_mmms,
                                   c = 1:ncol(perm_score_df_mmms),
                                   values_to = "value",
                                   names_to = "type")
perm_score_df_temp <- perm_score_df_temp[perm_score_df_temp$type %in% scores,]

plot1 <- ggplot2::ggplot(score_df_temp, aes(x = type, y = value))+
  ggplot2::geom_violin(data = perm_score_df_temp,
                       scale = "width",
                       position = position_nudge(x = 0.2, y = 0))+
  ggplot2::geom_point(alpha = 0.7,
                      aes(color = conservation_level),
                      size = 2,
                      position = position_nudge(x = -0.2, y = 0))+
  ggplot2::scale_color_manual("conservation_level", values = col_cons)+
  ggplot2::ylim(c(0,1))+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(base::paste(fraction_curr, 
                               iterations,
                               "iterations,",
                               "\n",
                               perm_score_df_mmms$cut_off_prop[1],
                               "% of cells,",
                               "\n",
                               perm_score_df_mmms$nr_genes_used,
                               "genes"))
```

```{r all2_mmms, fig.width = 3, fig.height = 3}

score_df_temp <- score_df_mmms[
  score_df_mmms$type %in% c("variation_information"),]
score_df_temp <- score_df_temp[
  score_df_temp$conservation_level == "mmusall_markers",]
score_df_temp_sign <- score_df_sign[
  score_df_sign$type %in% c("variation_information"),]
score_df_temp <- rbind(score_df_temp, score_df_temp_sign)

perm_score_df_temp <- pivot_longer(perm_score_df_mmms, 
                                   c = 1:ncol(perm_score_df_mmms),
                                   values_to = "value",
                                   names_to = "type")
perm_score_df_temp <- perm_score_df_temp[
  perm_score_df_temp$type %in% c("variation_information"),]

plot2 <- ggplot2::ggplot(score_df_temp, aes(x = type, y = value))+
  ggplot2::geom_violin(data = perm_score_df_temp,
                       scale = "width", 
                       position = position_nudge(x = 0.2, y = 0))+
  ggplot2::geom_point(aes(color = conservation_level), 
                      alpha = 0.7,
                      size = 2,
                      position = position_nudge(x = -0.2, y = 0))+
  ggplot2::scale_color_manual("conservation level", values = col_cons)+
  ggplot2::ylim(c(0, base::max(c(perm_score_df_temp$value,
                                 score_df_temp$value))))+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle("")
```

```{r numbers_mmms_sign}
print("cell types:")
print(score_df_mmms$nr_celltypes[1])
print(score_df_sign$nr_celltypes[1])
print(base::table(perm_score_df_mmms$nr_celltypes))

print("clusters:")
print(score_df_mmms$nr_clusters[1])
print(score_df_sign$nr_clusters[1])
print(base::table(perm_score_df_mmms$nr_clusters))
```

```{r plot_mmms_sign_x, fig.width = 5, fig.height = 5}
plot <- ggpubr::ggarrange(
  plot1 + ggplot2::theme(legend.position = "none"), 
  plot2 + ggplot2::theme(plot.title = element_blank(),
                          axis.title.y = element_blank()),
  align = "h",
  widths = c(1, 1.5))
plot
```

## Nr of genes per gene set

Prepare identically to 05_permutation_genesets.R

```{r prepare_mark_nrs_mmms, fig.width = 4, fig.height = 5}

# get the number for each
nr_sign <- length(sign_IDs) 
nr_mmms <- length(mmms_IDs[!mmms_IDs %in% sign_IDs])
nr_random_mmms <- length(mmms_IDs) - nr_sign

vis_df_mark <- data.frame(type1 = c("conserved signature", 
                                    "all BL6 markers", 
                                    "all BL6 markers",
                                    "all BL6 marker perm",
                                    "all BL6 marker perm"),
                          type2 = c("conserved_signature",
                                    "conserved_signature",
                                    "mmusall_markers",
                                    "conserved_signature",
                                    "random_features"),
                          number = c(nr_sign, 
                                     nr_sign,
                                     nr_mmms, 
                                     nr_sign,
                                     nr_random_mmms))

vis_df_mark$type1 <- factor(
  vis_df_mark$type1, levels = c("conserved signature", 
                                "all BL6 markers",
                                "all BL6 marker perm"))
vis_df_mark$type2 <- factor(
  vis_df_mark$type2, levels = c("mmusall_markers",
                                "random_features",
                                "conserved_signature"))

ggplot2::ggplot(vis_df_mark, aes(x = type1, y = number, fill = type2))+
  ggplot2::geom_col(position = position_stack())+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::xlab("conservation_level")+
  ggplot2::ylab("nr of genes")+
  ggplot2::scale_fill_manual("gene set", 
                             values = col_cons)+
  ggplot2::ylim(c(0, max_genes))
  
```

```{r session_info}
sessionInfo()
```
