---
title: "Reclustering of datasets from other species"
author: "Lea Wölbert"
date: '2024-04-15'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on datasets from other species that were re-clustered using:

- conserved signature genes
- conserved marker genes
- all BL6 marker genes
- random genes

This will also help picking a resolution. 

1. Choose a frame of resolution where the number of cell types per cluster
or clusters per cell types doesn't change much. 
This ensures that changes in the scores are not mainly caused by large 
differences in the number of clusters, since some scores are biased towards
a smaller or a larger number of clusters.
2. Pick a resolution with the best combined scores within the frame.
3. If there are multiple resolutions with very similar scores, choose
a higher resolution to get a higher number of potentially interesting clusters.
4. Seurat objects re-clustered with random genes get the same resolution as 
Seurat objects re-clustered with signature genes.
This is because when getting multiple random sets of genes, the "optimal" 
resolution may change each iteration (for permutation tests downstream). 
choosing the same resolution as is therefore the easiest option for direct
comparison.

#### Load objects

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(Seurat, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r load_objects}
seu_list_all <- base::readRDS(snakemake@input[["seu_list"]])
score_df_list_all <- base::readRDS(snakemake@input[["score_df_list"]])

dataset_curr <- snakemake@wildcards[["dataset"]]
```

```{r colors}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

cols_per <- col_cl_per_ct
col_vec1 <- col_score1
col_vec2 <-  col_score2
```

## Functions

### Plotting from seurat object

#### Wide

Cell types on y axis.
Proportions of cells of a cluster per cell type.

```{r wide_function}

vis_perc_wide_gg <- function(seu, title_add= NULL){
  
  title <- base::paste(seu@misc$used_genes, seu@misc$nr_genes_used, title_add)
  
  mat <- base::table(seu$cell_type, seu$seurat_clusters)
  mat1 <- mat/rowSums(mat)*100
  
  df_vis <- base::as.data.frame(mat1)
  df_vis$Var1 <- factor(df_vis$Var1, levels = base::rev(levels(seu$cell_type)))
  
  plot_wide <- ggplot2::ggplot(df_vis, aes(y = Var1, x = Var2, fill= Freq))+ 
    ggplot2::geom_tile()+
    ggplot2::scale_fill_continuous("% cells/cell type", 
                                   limits = c(0, 100), 
                                   breaks = seq(0, 100,by = 20),
                                   low = "white",
                                   high = "blue")+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle=0),
                   axis.ticks = element_blank(),
                   legend.position = "none")+
    ggplot2::xlab("new clusters")+
    ggplot2::ylab("original cell types")+
    ggplot2::ggtitle(base::paste("resolution", 
                                 seu@misc$resolution[1], 
                                 title))
  
  return(plot_wide)
}
```

#### Long

Cell types on x axis.
Proportion of cells of a cell type per cluster.

```{r long_function}
vis_perc_long_gg <- function(seu, title_add= NULL){
  
  title <- base::paste(seu@misc$used_genes, seu@misc$nr_genes_used, title_add)
    
  mat <- t(base::table(seu$cell_type, seu$seurat_clusters))
  mat2 <- mat/rowSums(mat)*100
  
  df_vis <- base::as.data.frame(mat2)
  df_vis$Var2 <- factor(df_vis$Var2, levels = levels(seu$cell_type))

  plot_long <- ggplot2::ggplot(df_vis, aes(y = Var1, x = Var2, fill= Freq))+ 
    ggplot2::geom_tile()+
    ggplot2::scale_fill_continuous("% cells/cluster", 
                                   limits = c(0, 100),
                                   breaks = seq(0, 100, by = 20),
                                   low = "white", 
                                   high = "blue")+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.ticks = element_blank(),
                   legend.position = "none",
                   axis.text.x = element_text(angle = 60,
                                              vjust = 1, 
                                              hjust = 1))+
    ggplot2::ylab("new clusters")+
    ggplot2::xlab("original cell types")+
    ggplot2::ggtitle(base::paste("resolution", seu@misc$resolution[1], title))

  return(plot_long)
}
```

### Nr of clusters or cell types 

Get the mean and meadian nr of cell types per cluster and clusters per ct.
This will help with picking a good resolution where these numbers are not 
too extreme/relatively stable.

```{r nr_clusters_function}

vis_nr_cl_nr_cell <- function(seu_list, title){
  
  vis_df_list <- lapply(seu_list, function(seu){
    
    mat <- base::table(seu$cell_type, seu$seurat_clusters)
    
    # get the nr of cell types per cluster for each cluster
    nr_ct_per_cluster_all <- vector()
    for(cluster in colnames(mat)){
      nr_ct_per_cluster <- length(which(mat[,cluster] !=0))
      nr_ct_per_cluster_all <- c(nr_ct_per_cluster_all, nr_ct_per_cluster)
    }
  
    # same for nr of clusters per cell type for each cell type
    nr_clusters_per_ct_all <- vector()
    for(ct in rownames(mat)){
      nr_clusters_per_ct <- length(which(mat[ct,] !=0))
      nr_clusters_per_ct_all <- c(nr_clusters_per_ct_all, nr_clusters_per_ct)
    }
  
    vis_df <- base::data.frame(
        "mean_ct_per_cluster" = base::mean(nr_ct_per_cluster_all),
        "median_ct_per_cluster" = stats::median(nr_ct_per_cluster_all),
        "mean_cluster_per_ct" = base::mean(nr_clusters_per_ct_all),
        "median_cluster_per_ct" = stats::median(nr_clusters_per_ct_all),
        "nr_clusters" = ncol(mat),
        "resolution" = seu@misc$resolution,
        "conservation_level" = seu@misc$used_genes
    )
    return(vis_df)
  })
  
  vis_df_all <- dplyr::bind_rows(vis_df_list)
  
  # the nr of original cell types is always the same
  vis_df_all$nr_celltypes <- base::rep(
    length(base::unique(seu_list[[1]]$cell_type)), nrow(vis_df_all))
  
  # pivot_longer
  vis_df_all <- pivot_longer(vis_df_all, 
                             names_to = "type", 
                             values_to = "value",
                             cols = c("mean_ct_per_cluster",
                                      "median_ct_per_cluster",
                                      "mean_cluster_per_ct",
                                      "median_cluster_per_ct",
                                      "nr_clusters"))
  
  title <- base::paste(
    title, 
    seu_list[[1]]@misc$nr_genes_used, 
    vis_df_all$conservation_level[1])
  
  plot1 <- ggplot2::ggplot(vis_df_all, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point(alpha = 0.7)+
    ggplot2::scale_color_manual("", values = cols_per)+
    ggplot2::geom_hline(yintercept = vis_df_all$nr_celltypes[1], 
                        color = "grey60", 
                        linetype = "dashed")+
    ggplot2::theme_classic()+
    ggplot2::ylim(c(0, base::max(vis_df_all$nr_celltypes[1], 
                                 vis_df_all$value)))+
    ggplot2::ggtitle(title)
  
  return(plot1)
}

# run the function but visualize later
plot_list_nr <- lapply(seu_list_all, vis_nr_cl_nr_cell, title = dataset_curr)
```

### Scores

```{r score_vis_function}
vis_clustering_scores <- function(score_df_list, title){
  
  score_df <- dplyr::bind_rows(score_df_list)
  title <- base::paste(
    title, 
    score_df$nr_genes_used[1], 
    score_df$conservation_level[1])

  vis_df_1 <- score_df[score_df$type %in% c("mean_prop_cells_cluster",
                                            "mean_cluster_purity"),]
  
  plot1 <- ggplot2::ggplot(vis_df_1, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point()+
    ggplot2::theme_classic()+
    ggplot2::geom_hline(yintercept = 0.5,
                        color = "grey20",
                        linetype = "dashed")+
    ggplot2::ylim(c(0, 1))+
    ggplot2::ggtitle(title)+
    ggplot2::scale_color_manual("score", values = col_vec1)
  
  vis_df_2 <- score_df[score_df$type %in% c("adjusted_rand_index",
                                            "variation_information"),]

  plot2 <- ggplot2::ggplot(vis_df_2, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point()+
    ggplot2::theme_classic()+
    ggplot2::geom_hline(yintercept = 0.5,
                        color = "grey20",
                        linetype = "dashed")+
    ggplot2::ylim(c(0, base::max(vis_df_2$value)))+
    ggplot2::ggtitle(title)+
    ggplot2::scale_color_manual("score", values = col_vec2)
  
  return(list(plot1, plot2))
  
}
# run the function, visualize later
plotlist_scores <- lapply(score_df_list_all, 
                          vis_clustering_scores, 
                          title = dataset_curr)
```

# Plot

## Conserved Signature Genes

```{r sign1}
seu_sign_list <- seu_list_all$seu_sign
```

```{r print_nr_plot_sign, fig.width = 7, fig.height = 3}
plot_list_nr$seu_sign
```

```{r print_scores_sign, fig.width = 7, fig.height = 3}
plotlist_scores$seu_sign
```

```{r sign2, fig.width = 15, fig.height=30}
plot_listw <- lapply(seu_sign_list,
                     vis_perc_wide_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listw, ncol = 2)
```

```{r sign3, fig.height = 40, fig.width=12}
plot_listl <- lapply(seu_sign_list, 
                     vis_perc_long_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listl, ncol = 3)
```

# Conserved Markers

```{r markers1}
seu_marker_list <- seu_list_all$seu_mark
```

```{r print_nr_plot_mark, fig.width = 7, fig.height = 3}
plot_list_nr$seu_mark
```

```{r print_scores_mark, fig.width = 7, fig.height = 3}
plotlist_scores$seu_mark
```

```{r markers2, fig.width = 14, fig.height=30}
plot_listw <- lapply(seu_marker_list, 
                     vis_perc_wide_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listw, ncol = 2)
```

```{r markers3, fig.height = 40, fig.width=12}
plot_listl <- lapply(seu_marker_list,
                     vis_perc_long_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listl, ncol = 3)
```

# All BL6 Markers

```{r bl6markers1}
seu_mmusm_list <- seu_list_all$seu_mmms
```

```{r print_nr_plot_mmms, fig.width = 7, fig.height = 3}
plot_list_nr$seu_mmms
```

```{r print_scores_mmms, fig.width = 7, fig.height = 3}
plotlist_scores$seu_mmms
```

```{r bl6markers2, fig.width = 14, fig.height=30}
plot_listw <- lapply(seu_mmusm_list, 
                     vis_perc_wide_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listw, ncol = 2)
```

```{r bl6markers3, fig.height = 40, fig.width=12}
plot_listl <- lapply(seu_mmusm_list,
                     vis_perc_long_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listl, ncol = 3)
```

## Random

```{r random1}
seu_random_list <- seu_list_all$seu_rand
```

```{r print_nr_plot_rand, fig.width = 7, fig.height = 3}
plot_list_nr$seu_rand
```

```{r print_scores_rand, fig.width = 7, fig.height = 3}
plotlist_scores$seu_rand
```

```{r random2, fig.width = 14, fig.height=30}
plot_listw <- lapply(seu_random_list, 
                     vis_perc_wide_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listw, ncol = 2)
```

```{r random3, fig.height = 40, fig.width=12}
plot_listl <- lapply(seu_random_list, 
                     vis_perc_long_gg, 
                     title_add = dataset_curr)
cowplot::plot_grid(plotlist = plot_listl, ncol = 3)
```

```{r sessioninfo}
utils::sessionInfo()
```
