---
title: "Reclustering other datasets, selected resolution"
author: "Lea Wölbert"
date: '2024-04-19'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on datasets from other species that were re-clustered using:

- conserved signature genes
- conserved marker genes
- all BL6 marker genes
- random genes

Using only the selected resolution with the best combination of markers.

#### Load objects

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(Seurat, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r load_objects}
seu_list_all <- base::readRDS(snakemake@input[["seu_list"]])
score_df_list_all <- base::readRDS(snakemake@input[["score_df_list"]])

dataset_curr <- snakemake@wildcards[["dataset"]]

resolution_df_path <- snakemake@input[["resolution_df"]]
resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")

frame_df_path <- snakemake@input[["frame_df"]]
frame_df <- utils::read.csv(
  file = frame_df_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

resolution_df <- resolution_df[resolution_df$dataset == dataset_curr,]
frame_df <- frame_df[frame_df$dataset == dataset_curr,]
```

```{r subset_selected}
# subset to selected resolution from resolution_df
selected_seu_list <- lapply(seu_list_all, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df[resolution_df$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[which(names(seu_list) == res_curr)]]
  return(seu_return)
  
})
selected_seu_list

dataset_list <- names(score_df_list_all)
selected_score_df_list <- lapply(dataset_list, function(dataset){
  
  seu_list_curr <- seu_list_all[[dataset]]
  score_df_list <- score_df_list_all[[dataset]]

  conslev_curr <- seu_list_curr[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df[resolution_df$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  score_df_return <- score_df_list[[res_curr]]
  return(score_df_return)
  
})

selected_seu_list
selected_score_df_list
```

```{r colors}

colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

col_vec_cons <- col_cons

col_vec1 <- col_score1
col_vec2 <- col_score2
cols_per <- col_cl_per_ct
```

## Functions

### Plotting Mat

#### Wide

Cell types on y axis.
Proportions of cells of a cluster per cell type.

```{r wide_function}

vis_perc_wide_gg <- function(seu, correction_vec_list = NULL, title_add= NULL){
  
  nr_genes <- nrow(seu)

  title <- base::paste(seu@misc$used_genes, title_add, "\n" , nr_genes, "genes")
  print(seu@misc$used_genes)
  
  mat <- base::table(seu$cell_type, seu$seurat_clusters)
  mat1 <- mat/rowSums(mat)*100
  
  # for nicer look, choose manually below
  if(!is.null(correction_vec_list)){
    correction_vec <- correction_vec_list[[seu@misc$used_genes]]
    mat1 <- mat1[,correction_vec]
  }
  
  df_vis <- base::as.data.frame(mat1)
  df_vis$Var1 <- factor(df_vis$Var1, levels = base::rev(levels(seu$cell_type)))
  
  plot_wide <- ggplot2::ggplot(df_vis, aes(y = Var1, x = Var2, fill= Freq))+ 
    ggplot2::geom_tile()+
    ggplot2::scale_fill_continuous("% cells/cell type", 
                                   limits = c(0, 100), 
                                   breaks = seq(0, 100,by = 20),
                                   low = "white",
                                   high = "blue")+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle=0),
                   axis.ticks = element_blank(),
                   legend.position = "none")+
    ggplot2::xlab("new clusters")+
    ggplot2::ylab("original cell types")+
    ggplot2::ggtitle(base::paste("resolution", 
                                 seu@misc$resolution[1], 
                                 title))
  
  return(plot_wide)
}
```

#### Long

Cell types on x axis.
Proportion of cells of a cell type per cluster.

```{r long_function}
vis_perc_long_gg <- function(seu, correction_vec_list = NULL, title_add= NULL){
  
  nr_genes <- nrow(seu)

  title <- base::paste(seu@misc$used_genes, title_add, "\n" , nr_genes, "genes")
  print(seu@misc$used_genes)
    
  mat <- t(base::table(seu$cell_type, seu$seurat_clusters))
  mat2 <- mat/rowSums(mat)*100
  
  if(!is.null(correction_vec_list)){
    correction_vec <- correction_vec_list[[seu@misc$used_genes]]
    mat2 <- mat2[rev(correction_vec),]
  }
  
  df_vis <- base::as.data.frame(mat2)
  df_vis$Var2 <- factor(df_vis$Var2, levels = levels(seu$cell_type))

  plot_long <- ggplot2::ggplot(df_vis, aes(y = Var1, x = Var2, fill= Freq))+ 
    ggplot2::geom_tile()+
    ggplot2::scale_fill_continuous("% cells/cluster", 
                                   limits = c(0, 100),
                                   breaks = seq(0, 100, by = 20),
                                   low = "white", 
                                   high = "blue")+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.ticks = element_blank(),
                   legend.position = "none",
                   axis.text.x = element_text(angle = 60,
                                              vjust = 1, 
                                              hjust = 1))+
    ggplot2::ylab("new clusters")+
    ggplot2::xlab("original cell types")+
    ggplot2::ggtitle(base::paste("resolution", seu@misc$resolution[1], title))

  return(plot_long)
}
```

```{r nr_clusters_function}

vis_nr_cl_nr_cell <- function(seu_list, title){
  
  vis_df_list <- lapply(seu_list, function(seu){
    
    mat <- base::table(seu$cell_type, seu$seurat_clusters)
    
    # get the nr of cell types per cluster for each cluster
    nr_ct_per_cluster_all <- vector()
    for(cluster in colnames(mat)){
      nr_ct_per_cluster <- length(which(mat[,cluster] !=0))
      nr_ct_per_cluster_all <- c(nr_ct_per_cluster_all, nr_ct_per_cluster)
    }
  
    # same for nr of clusters per cell type for each cell type
    nr_clusters_per_ct_all <- vector()
    for(ct in rownames(mat)){
      nr_clusters_per_ct <- length(which(mat[ct,] !=0))
      nr_clusters_per_ct_all <- c(nr_clusters_per_ct_all, nr_clusters_per_ct)
    }
  
    vis_df <- base::data.frame(
        "mean_ct_per_cluster" = base::mean(nr_ct_per_cluster_all),
        "median_ct_per_cluster" = stats::median(nr_ct_per_cluster_all),
        "mean_cluster_per_ct" = base::mean(nr_clusters_per_ct_all),
        "median_cluster_per_ct" = stats::median(nr_clusters_per_ct_all),
        "nr_clusters" = ncol(mat),
        "resolution" = seu@misc$resolution,
        "conservation_level" = seu@misc$used_genes
    )
    return(vis_df)
  })
  
  vis_df_all <- dplyr::bind_rows(vis_df_list)
  
  frame_df_temp <- frame_df[
    frame_df$conservation_level == seu_list[[1]]@misc$used_genes,]
  frame_low <- as.numeric(frame_df_temp$frame_low)
  frame_high <- as.numeric(frame_df_temp$frame_high)
  
  resolution_chosen <- as.numeric(
    resolution_df$resolution[
      resolution_df$conservation_level == seu_list[[1]]@misc$used_genes])
  
  # the nr of original cell types is always the same
  vis_df_all$nr_celltypes <- base::rep(
    length(base::unique(seu_list[[1]]$cell_type)), nrow(vis_df_all))
  
  # pivot_longer
  vis_df_all <- pivot_longer(vis_df_all, 
                             names_to = "type", 
                             values_to = "value",
                             cols = c("mean_ct_per_cluster",
                                      "median_ct_per_cluster",
                                      "mean_cluster_per_ct",
                                      "median_cluster_per_ct",
                                      "nr_clusters"))
  
  title <- base::paste(title, vis_df_all$conservation_level[1])
  
  plot1 <- ggplot2::ggplot(vis_df_all, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point(alpha = 0.7)+
    ggplot2::scale_color_manual("", values = cols_per)+
    ggplot2::geom_hline(yintercept = vis_df_all$nr_celltypes[1], 
                        color = "grey60", 
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_low, 
                        color = "grey60", 
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_high, 
                        color = "grey60", 
                        linetype = "dashed")+ 
    ggplot2::geom_vline(xintercept = resolution_chosen, 
                        color = "orange2", 
                        linetype = "dashed")+ 
    ggplot2::theme_classic()+
    ggplot2::ylim(c(0, base::max(vis_df_all$nr_celltypes[1], 
                                 vis_df_all$value)))+
    ggplot2::ggtitle(title)
  
  return(plot1)
}

# run the function but visualize later
plot_list_nr <- lapply(seu_list_all, vis_nr_cl_nr_cell, title = dataset_curr)
```

### Scores

```{r score_vis_function}
vis_clustering_scores <- function(score_df_list, title){
  
  score_df <- dplyr::bind_rows(score_df_list)
  
  cons_level <- score_df$conservation_level[1]
  
  title <- base::paste(title, score_df$conservation_level[1])

  frame_df_temp <- frame_df[
    frame_df$conservation_level == cons_level,]
  frame_low <- as.numeric(frame_df_temp$frame_low)
  frame_high <- as.numeric(frame_df_temp$frame_high)
  
  resolution_chosen <- as.numeric(
    resolution_df$resolution[
      resolution_df$conservation_level == cons_level])
  
  vis_df_1 <- score_df[score_df$type %in% c("mean_prop_cells_cluster",
                                            "mean_cluster_purity"),]
  
  plot1 <- ggplot2::ggplot(vis_df_1, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point()+
    ggplot2::theme_classic()+
    ggplot2::geom_hline(yintercept = 0.5,
                        color = "grey20",
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_low, 
                        color = "grey60", 
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_high, 
                        color = "grey60", 
                        linetype = "dashed")+ 
    ggplot2::geom_vline(xintercept = resolution_chosen, 
                        color = "orange2", 
                        linetype = "dashed")+ 
    ggplot2::ylim(c(0, 1))+
    ggplot2::ggtitle(title)+
    ggplot2::scale_color_manual("score", values = col_vec1)
  
  vis_df_2 <- score_df[score_df$type %in% c("adjusted_rand_index",
                                            "variation_information"),]

  plot2 <- ggplot2::ggplot(vis_df_2, 
                           aes(x = resolution, y = value, color = type))+
    ggplot2::geom_point()+
    ggplot2::theme_classic()+
    ggplot2::geom_hline(yintercept = 0.5,
                        color = "grey20",
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_low, 
                        color = "grey60", 
                        linetype = "dashed")+
    ggplot2::geom_vline(xintercept = frame_high, 
                        color = "grey60", 
                        linetype = "dashed")+ 
    ggplot2::geom_vline(xintercept = resolution_chosen, 
                        color = "orange2", 
                        linetype = "dashed")+ 
    ggplot2::ylim(c(0, base::max(vis_df_2$value)))+
    ggplot2::ggtitle(title)+
    ggplot2::scale_color_manual("score", values = col_vec2)
  
  return(list(plot1, plot2))
  
}
# run the function, visualize later
plotlist_scores <- lapply(score_df_list_all, 
                          vis_clustering_scores, 
                          title = dataset_curr)
```

# Plot

## Scores

```{r score_df}
score_df <- dplyr::bind_rows(selected_score_df_list)
```

```{r all_0_to_1, fig.width = 4, fig.height = 3.5}

title <- base::paste(dataset_curr)

ggplot2::ggplot(
  score_df[!score_df$type %in% c("variation_information", 
                                 "nr_celltypes",
                                 "nr_clusters"),], 
  aes(y = value, x = type, color = conservation_level))+
  ggplot2::geom_point(alpha = 0.7, position=position_dodge(width=0.5))+
  ggplot2::theme_classic()+
  ggplot2::ylim(c(0, 1))+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::scale_color_manual("conservation level", values = col_vec_cons)+
  ggplot2::ggtitle(title)
```

```{r vi, fig.width = 3, fig.height = 3.5}
ggplot2::ggplot(score_df[score_df$type == "variation_information",], 
                aes(y = value, x = type, color = conservation_level))+
  ggplot2::geom_point(alpha = 0.7, position=position_dodge(width=0.5))+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::scale_color_manual("conservation level", 
                              values = col_vec_cons)+
  ggplot2::ylim(c(0, max(score_df$value[
    score_df$type == "variation_information"])))+
  ggplot2::ggtitle(title)
```

```{r print_nr_plot, fig.width = 7, fig.height = 3}
plot_list_nr
```

```{r print_scores, fig.width = 7, fig.height = 3}
plotlist_scores
```

## Heatmaps

```{r wide1, fig.width = 8, fig.height = 4}
# include = FALSE, eval = FALSE
# only for manually choosing the correction vector
plot_list_wide <- lapply(selected_seu_list, 
                         vis_perc_wide_gg,
                         title_add = dataset_curr)
plot_list_wide
```

```{r long1, fig.width = 6, fig.height = 5}
# only for manually choosing the correction vector
plot_list_long <- lapply(selected_seu_list, 
                         vis_perc_long_gg,
                         title_add = dataset_curr)
plot_list_long
```

```{r sessioninfo}
utils::sessionInfo()
```
