---
title: "Preprocessing reports at sample level"
author: "Amy Danson, Lea Wölbert"
date: '2022-09-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

QC report for all samples separately (Object_ID).
Some operations from the scripts must be repeated to generate relevant plots,
but any new objects cannot be saved from an .Rmd script using snakemake.
Repeat various steps for visualisation.

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r load_packages, message = FALSE}
library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
```

```{r source, message = FALSE}
source(file = snakemake@params[["plotting"]])
```

```{r load_objects, message = FALSE}

sce_input <- base::readRDS(file = snakemake@input[["sce_input"]])
sce_drop <- base::readRDS(file = snakemake@input[["sce_drop"]])
sce_outl <- base::readRDS(file = snakemake@input[["sce_outl"]])
sce_norm <- base::readRDS(file = snakemake@input[["sce_norm"]])
sce_dimr <- base::readRDS(file = snakemake@input[["sce_dimr"]])

# set name for plots
name_curr <- colData(sce_drop)$Object_ID[1] 
```

```{r declare_cutoffs, message = FALSE}
# call metadata
cutoff_umis <- snakemake@params[["cutoff_umis"]]
cutoff_doublets <- snakemake@params[["cutoff_doublets"]] 
cutoff_sum <- snakemake@params[["cutoff_sum"]] 
cutoff_detected <- snakemake@params[["cutoff_detected"]]
cutoff_mitos <- snakemake@params[["cutoff_mitos"]]
```

```{r colors, message = FALSE}
source(file = snakemake@params[["colors"]])
```

## Step 1: 

### Remove empty droplets

Check that the barcode rank and p value distributions look appropriate.
Also check if any droplets with low FDR (that are likely cells) were dropped

```{r remove_droplets}

# show barcoded ranks 
# before droplet removal
bcrank <- DropletUtils::barcodeRanks(SingleCellExperiment::counts(sce_input)) 
uniq_pos <- !base::duplicated(bcrank$rank)

base::plot(bcrank$rank[uniq_pos],
           bcrank$total[uniq_pos],
           log="xy", 
           xlab="Rank", 
           ylab="Total UMI count", 
           cex.lab=1.2, 
           main = name_curr)
graphics::abline(h = metadata(bcrank)$inflection,
                 col = marker1)
graphics::abline(h = metadata(bcrank)$knee, 
                 col = marker2)
graphics::legend("bottomleft",
                 legend=c("Inflection", "Knee"), 
                 col=c(marker1, marker2), 
                 cex=1.2)

# show pval distribution    
cutoff_umis <- cutoff_umis
out <- DropletUtils::emptyDrops(SingleCellExperiment::counts(sce_input), 
                                lower = cutoff_umis, 
                                test.ambient=TRUE)  

graphics::hist(out$PValue[out$Total <= cutoff_umis & out$Total > 0],
               xlab="P-value",
               main = paste(name_curr, "cutoff", cutoff_umis), 
               col="grey80")  

# show FDR of droplets that were kept or removed
check <- base::table(Sig=out$FDR <= 0.001, Limited=out$Limited)
print(check)
```

### Remove doublets

Check if cells with high doublet scores form conspicuous clusters

```{r plot_doublets}
scater::plotUMAP(sce_dimr, colour_by = "doublet_score")+ # requires reduced dims
  ggplot2::ggtitle(name_curr)+
  ggplot2::theme(legend.key.size = unit(0.3, "cm"), 
                 legend.key.width = unit(0.3, "cm"),
                 legend.spacing = unit(0.06, "cm"), 
                 legend.text = element_text(size = 7))
```

## Step 2: Remove outliers

Check the quality of the single cells and if cells were removed appropriately

```{r remove_outliers}

# get mito count just to check
mito_genes <- grep("mt-", rownames(sce_drop))
qcdf_outl <- scuttle::perCellQCMetrics(sce_drop, subsets=list(Mito=mito_genes)) 
# before outlier removal

# set QC cutoffs below which cells are removed
sum_out <- qcdf_outl$sum < cutoff_sum 
det_out <- qcdf_outl$detected < cutoff_detected 
mito_out <- qcdf_outl$subsets_Mito_percent > cutoff_mitos

# remove all cells to discard and add info to qc
remove_pos <- sum_out | det_out | mito_out
qcdf_outl$removed <- remove_pos

qcdf_gg <- base::as.data.frame(qcdf_outl[,c(1,2,5,7)])

qcdf_gg <- tidyr::pivot_longer(qcdf_gg, 
                               names_to = "metric", 
                               cols = c(1, 2, 3))

ggplot2::ggplot(qcdf_gg[qcdf_gg$metric == "sum",], 
                aes(y = value, x = metric,  color = removed))+
  ggplot2::theme_classic()+
  ggplot2::geom_jitter(size = 0.1, alpha = 50)+
  ggplot2::geom_violin(color = "black", alpha = 0, draw_quantiles = 0.5)+  
  ggplot2::scale_color_manual("Removed", 
                              values = c("FALSE" = "grey80", 
                                         "TRUE" = "orange"))+
  ggplot2::ylab("Library size")+
  ggplot2::ggtitle(name_curr)+
  ggplot2::theme(axis.title.x = element_blank(), 
                 axis.text.x = element_blank())+
  ggplot2::geom_abline(intercept = cutoff_sum, 
                       color = marker1,  
                       slope = 0) 

ggplot2::ggplot(qcdf_gg[qcdf_gg$metric == "detected",], 
                aes(y = value, x = metric, color = removed))+
  ggplot2::theme_classic()+
  ggplot2::geom_jitter(size = 0.1, alpha = 50)+
  ggplot2::geom_violin(color = "black", alpha = 0, draw_quantiles = 0.5)+ 
  ggplot2::scale_color_manual("Removed", 
                              values = c("FALSE" = "grey80", 
                                         "TRUE" = "orange"))+
  ggplot2::ylab("Detected genes")+
  ggplot2::ggtitle(name_curr)+
  ggplot2::theme(axis.title.x = element_blank(), 
                 axis.text.x = element_blank())+
  ggplot2::geom_abline(intercept = cutoff_detected, color = marker1, slope = 0) 

ggplot2::ggplot(qcdf_gg[qcdf_gg$metric == "subsets_Mito_percent",], 
                aes(y = value, x = metric, color = removed))+
  ggplot2::theme_classic()+
  ggplot2::geom_jitter(size = 0.1, alpha = 50)+
  ggplot2::geom_violin(color = "black", alpha = 0, draw_quantiles = 0.5)+  
  ggplot2::scale_color_manual("Removed",
                              values = c("FALSE" = "grey80", 
                                         "TRUE" = "orange"))+
  ggplot2::ylab("Mitochondrial gene percent")+
  ggplot2::ggtitle(name_curr)+
  ggplot2::theme(axis.title.x = element_blank(), 
                 axis.text.x = element_blank())+
  ggplot2::geom_abline(intercept = cutoff_mitos, color = marker1, slope = 0) 
```

## Step 3: Normalization

```{r normalization}
# calculate two different size factors for direct comparison
quick_clust <- scran::quickCluster(sce_outl) # before normalization
sizef_deconv <- scran::calculateSumFactors(sce_outl, cluster = quick_clust)
sizef_lib <- scuttle::librarySizeFactors(sce_outl)

base::plot(sizef_lib, 
           sizef_deconv,
           xlab = "Library size factor",
           ylab = "Deconv. size factor",
           log = 'xy',
           pch=16, 
           main = paste("Library/Deconv", name_curr))
graphics::abline(a=0,
                 b=1,
                 col="red")
    
# distribution of size factors should be close to normal dist
graphics::hist(log10(sizef_deconv),
               xlab="Log10[Deconv factor]", 
               col='grey80', 
               main = name_curr)
    
# directly compare deconv factors to library size
qcdf_03 <- scuttle::perCellQCMetrics(sce_outl) # after outlier removal
base::plot(sizef_deconv, 
           qcdf_03$sum, 
           log = "xy",
           ylab = "Library size",
           xlab = "Deconv. Size factor", 
           main = paste("Deconv/libsize", name_curr))
```

```{r print_objects}
# print final object
print(name_curr)
print(sce_norm) # after normalization
print(colData(sce_norm))
print(ncol(sce_norm))
```

## Remaining quality

```{r remaining_quality}

qcdf_dimr <- scuttle::perCellQCMetrics(sce_dimr)

sce_dimr$sum <- qcdf_dimr$sum
sce_dimr$detected <- qcdf_dimr$detected

plot_1 <- umap_base(sce_dimr, color_by = "sum")+ # own function
  ggplot2::scale_color_gradientn("Library size", 
                                 limits = c(0, 100000),
                                 colors = color_vector_sum)+
  ggplot2::ggtitle(name_curr)

plot_1l <- umap_legend(sce_dimr, color_by = "sum")+
  ggplot2::scale_color_gradientn("Library size", 
                                 limits = c(0, 100000),
                                 colors = color_vector_sum)
legend_1 <- ggpubr::get_legend(plot_1l)

PLOT_1 <- ggpubr::ggarrange(plot_1, legend_1)
PLOT_1

plot_2 <- umap_base(sce_dimr, color_by = "detected")+ 
  ggplot2::scale_color_gradientn("Number of detected genes", 
                                 limits = c(0, 10000),
                                 colors = color_vector_det)+
  ggplot2::ggtitle(name_curr)

plot_2l <- umap_legend(sce_dimr, color_by = "detected")+
  ggplot2::scale_color_gradientn("Number of detected genes", 
                                 limits = c(0, 10000),
                                 colors = color_vector_det)
legend_2 <- ggpubr::get_legend(plot_2l)

PLOT_2 <- ggpubr::ggarrange(plot_2, legend_2)
PLOT_2
```

```{r session_info}
utils::sessionInfo()
```