---
title: "integration_sample_reports"
author: "Amy Danson, Lea Wölbert"
date: '2022-10-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Summary report for data integration.

#### Load libraries, source code 

```{r load,  message = FALSE}

library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(cowplot, quietly = TRUE) 

```

```{r}
source(file = "../source/colors.R")
source(file = "../source/plotting.R")
source(file = "../source/sce_functions.R")
```

```{r}

sce_07 <- snakemake@input[["sce_07"]]
sce_08 <- snakemake@input[["sce_08"]]
sce_09mnn <- snakemake@input[["sce_09mnn"]]

nr_hvgs <- snakemake@params[["nr_hvgs"]]
species_build <- snakemake@params[["species_build"]]
batches <- snakemake@params[["batches"]]

```

# Preliminary Cell types 

## Sample level

UMAPS of the preliminary celltypes to get an overview/idea of the data.

```{r}

individual_sce_list <-list()
for(i in colData(sce_07)$Object_ID){
  individual_sce_list[[i]] <- sce_07[,which(colData(sce_07)$Object_ID == i)]
}

# generic dimensionality reduction
individual_sce_list <- lapply(individual_sce_list, reduce_dims) # own function

```

```{r}

identity_plots <- function(sce, title, identifier){
  
  id_1 <- colData(sce)[identifier][1]
  
  col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
  col_cts_fraction <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction]
  
  plot_1 <- umap_base(sce, color_by = "Identity_ref_all")+ # own function
    ggtitle(paste0(id_1, title))+
    scale_color_manual("Identity1", values = col_cts_all)
  
  plot_1l <- umap_legend(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", values = col_cts_all)
  legend_1 <- get_legend(plot_1l)

    
  plot_2 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    ggtitle(paste0(id_1, title))+
    scale_color_manual("Identity2", col_cts_fraction)
  
  plot_2l <- umap_legend(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)
  legend_2 <- get_legend(plot_2l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  
  return(list(PLOT_1, PLOT_2))
}

```

```{r}

identity_individual_plots <- lapply(individual_sce_list, 
                                    identity_plots, 
                                    identifier = "Object_ID",
                                    title = "")

lapply(individual_sce_list, function(sce){
  
  print(sce$Object_ID[1])
  print(table(sce$Identity_ref_all))
  print(table(sce$Identity_ref_fraction))
})

```

```{r fig.width=10, fig.height=5}
identity_individual_plots
```

## Species level

### After merging

```{r}

# generic dimensionality reduction
sce_07 <- reduce_dims(sce_07)

identity_species_plots_07 <- identity_plots(sce_07, 
                                            title = ", after merging",
                                            identifier = "Species_ID")

print(sce_07$Species_ID[1])
print(table(sce_07$Identity_ref_all))
print(table(sce_07$Identity_ref_fraction))

```

```{r fig.width=10, fig.height=5}
identity_species_plots_07
```

### After renormalization

```{r}

# generic dimensionality reduction
sce_08 <- reduce_dims(sce_08)

identity_species_plots_08 <- identity_plots(sce_08, 
                                            title = ", after renormalization",
                                            identifier = "Species_ID")

```

```{r fig.width=10, fig.height=5}
identity_species_plots_08
```

### After batch correction

```{r}

# does not need PCA calculation 
sce_09mnn <- reduce_dims(sce_09mnn)

hvgs <- modelGeneVar(sce_09mnn)
hvgs <- getTopHVGs(hvgs, n=nr_hvgs)

sce_09mnn <- runUMAP(sce_09mnn, dimred = "corrected",
                     external_neighbors=TRUE, subset_row = hvgs)

identity_species_plots_09 <- identity_plots(sce_09mnn, 
                                            title = ", after batch correction",
                                            identifier = "Species_ID")

```

```{r fig.width=10, fig.height=5}
identity_species_plots_09
```

# All QC metrics

```{r}

qc_plots <- function(sce, title){
  
  species <- sce$Species_ID[1]

  plot_1 <- umap_base(sce, color_by = "Object_ID")+ # own function
    ggtitle(paste0(species, title))
  plot_1l <- umap_legend(sce, color_by = "Object_ID")
  legend_1 <- get_legend(plot_1l)
  
  col_batch_seq <- col_num[names(col_num) %in% sce$Batch_sequencing]
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)+
    ggtitle(paste0(species, title))
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)
  legend_2 <- get_legend(plot_2l)
  
  col_batch_exp_day <- col_alp[names(col_alp) %in% sce$Batch_exp_day]
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+ 
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species, title))
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_3 <- get_legend(plot_3l)
  
  plot_4 <- umap_base(sce, color_by = "Date_collected")+ 
    ggtitle(paste0(species, title))
  plot_4l <- umap_legend(sce, color_by = "Date_collected")+
  legend_4 <- get_legend(plot_4l)
  
  # check everything else
  col_species <- col_spc[names(col_spc) %in% sce$Species_ID]
  plot_5 <- umap_base(sce, color_by = "Species_ID")+ 
    scale_color_manual("Species", col_species)+
    ggtitle(paste0(species, title))
  plot_5l <- umap_legend(sce, color_by = "Species_ID")+
    scale_color_manual("Species", col_species)+
  legend_5 <- get_legend(plot_5l)
  
  col_ages <- col_ann[names(col_ann) %in% sce$Age]
  plot_6 <- umap_base(sce, color_by = "Age")+ 
    scale_color_manual("Age", col_ages)+
    ggtitle(paste0(species, title))
  plot_6l <- umap_legend(sce, color_by = "Age")+
    scale_color_manual("Age", col_ages)+
  legend_6 <- get_legend(plot_6l) 
  
  col_fractions <- col_ann[names(col_ann) %in% sce$Fraction]
  plot_7 <- umap_base(sce, color_by = "Fraction")+ 
    scale_color_manual("Fraction", col_fractions)+
    ggtitle(paste0(species, title))
  plot_7l <- umap_legend(sce, color_by = "Fraction")+
    scale_color_manual("Fraction", col_fractions)+
  legend_7 <- get_legend(plot_7l) 
 
  plot_8 <- umap_base(sce, color_by = "Antibody_combination")+ 
    ggtitle(paste0(species, title))
  plot_8l <- umap_legend(sce, color_by = "Antibody_combination")+
  legend_8 <- get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "doublet_score")+ 
    ggtitle(paste0(species, title))
  plot_9l <- umap_legend(sce, color_by = "doublet_score")+
  legend_9 <- get_legend(plot_9l)
  
  plot_10 <- umap_base(sce, color_by = "Keep_sample")+ 
    ggtitle(paste0(species, title))
  plot_10l <- umap_legend(sce, color_by = "Keep_sample")+
  legend_10 <- get_legend(plot_10l)
  
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)
  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_5 <- ggarrange(plot_5, legend_5)
  PLOT_6 <- ggarrange(plot_6, legend_6)
  PLOT_7 <- ggarrange(plot_7, legend_7)
  PLOT_8 <- ggarrange(plot_8, legend_8)
  PLOT_9 <- ggarrange(plot_9, legend_9)
  PLOT_10 <- ggarrange(plot_10, legend_10)

  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10)
  return(plot_list)
  
}

qc_plots_species_07 <- qc_plots(sce_07, title = ", after merging",)
qc_plots_species_08 <- qc_plots(sce_08, title = ", after renormalization")
qc_plots_species_09 <- qc_plots(sce_09mnn, title = ", after batch correction",)

```

### After merging

```{r fig.width=10, fig.height=5}
qc_plots_species_07
```

### After renormalization

```{r fig.width=10, fig.height=5}
qc_plots_species_08
```

### After batch correction

```{r fig.width=10, fig.height=5}
qc_plots_species_09
```

## Facet grid UMAPs

To merge across ages and fractions using MNNcorrect, every age and fraction 
should have at least one overlapping BATCH.
Per definition, this is not the case for Object_ID as possible batch.

```{r}

facet_plots <- function(sce, title){
  
  # Check Fraction
  species <- sce$Species_ID[1]
  
  plot_1 <- umap_base(sce, color_by = "Object_ID")+ 
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_1l <- umap_legend(sce, color_by = "Object_ID")
  legend_1 <- get_legend(plot_1l)
  
  col_batch_seq <- col_num[names(col_num) %in% sce$Batch_sequencing]
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)
  legend_2 <- get_legend(plot_2l)
  
  col_batch_exp_day <- col_alp[names(col_alp) %in% sce$Batch_exp_day]
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_3 <- get_legend(plot_3l)

  # check age
  plot_4 <- umap_base(sce, color_by = "Object_ID")+ 
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  plot_4l <- umap_legend(sce, color_by = "Object_ID")
  legend_4 <- get_legend(plot_4l)
  
  plot_5 <- umap_base(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  plot_5l <- umap_legend(sce, color_by = "Batch_sequencing")+
    scale_color_manual("Batch_sequencing", col_batch_seq)
  legend_5 <- get_legend(plot_5l)
  
  plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_6 <- get_legend(plot_6l)
  
  plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
    scale_color_manual("Batch_experimental_day", col_batch_exp_day)
  legend_6 <- get_legend(plot_6l)
  
  # Identities
  col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
  col_cts_fraction <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction]
 
  plot_7 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_7l <- umap_legend(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)
  legend_idrefall <- get_legend(plot_7l)
  
  plot_8 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Fraction))
  plot_8l <- umap_legend(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)
  legend_idreffrac <- get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  
  plot_10 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Age))
  
  plot_11 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_12 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_13 <- umap_base(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", col_cts_all)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Batch_exp_day))
  
  plot_14 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)+
    ggtitle(paste0(species, title))+
    facet_grid(cols = vars(sce$Batch_exp_day))
 
  PLOT_1 <- ggarrange(plot_1, legend_1)
  PLOT_2 <- ggarrange(plot_2, legend_2)
  PLOT_3 <- ggarrange(plot_3, legend_3)
  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_5 <- ggarrange(plot_5, legend_5)
  PLOT_6 <- ggarrange(plot_6, legend_6)
  PLOT_7 <- ggarrange(plot_7, legend_idrefall)
  PLOT_8 <- ggarrange(plot_8, legend_idreffrac)
  PLOT_9 <- ggarrange(plot_9, legend_idrefall)
  PLOT_10 <- ggarrange(plot_10, legend_idreffrac)
  PLOT_11 <- ggarrange(plot_11, legend_idrefall)
  PLOT_12 <- ggarrange(plot_12, legend_idreffrac)
  PLOT_13 <- ggarrange(plot_13, legend_idrefall)
  PLOT_14 <- ggarrange(plot_14, legend_idreffrac)
  
  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10, PLOT_11, PLOT_12, PLOT_13, PLOT_14)
  
  return(plot_list)
}

facet_plots_species_07 <- facet_plots(sce_07, title = ", after merging",)
facet_plots_species_08 <- facet_plots(sce_08, title = ", after renormalization")
facet_plots_species_09 <- facet_plots(sce_09mnn,
                                      title = ", after batch correction",)

```

### After merging

```{r fig.width=10, fig.height=5}
facet_plots_species_07
```

### After renormalization

```{r fig.width=10, fig.height=5}
facet_plots_species_08
```

### After batch correction

```{r fig.width=10, fig.height=5}
facet_plots_species_09
```

```{r}
sessionInfo()
```
