---
title: "integration_sample_reports"
author: "Amy Danson, Lea Wölbert"
date: '2022-10-26'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Summary report for data integration.

#### Load libraries, source code 

```{r load,  message = FALSE}

library(DropletUtils, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(cowplot, quietly = TRUE) 

```

```{r}
source(file = "../source/colors.R")
source(file = "../source/imaging.R")
source(file = "../source/sce_functions.R")
```


```{r}

sce_07 <- snakemake@input[["sce_07"]]
sce_08 <- snakemake@input[["sce_08"]]
sce_09mnn <- snakemake@input[["sce_09mnn"]]

```

```{r}

nr_hvgs <- snakemake@params[["nr_hvgs"]]
species_all <- snakemake@params[["species_all"]]
batches <- snakemake@params[["batches"]]

```

# Preliminary Cell types 

### Sample level

UMAPS of the preliminary celltypes to get an overview/idea of the data.

```{r}

individual_sce_list <-list()
for(i in colData(sce_07)$Object_ID){
  individual_sce_list[[i]] <- sce_07[,which(colData(sce_07)$Object_ID == i)]
}

# generic dimensionality reduction
individual_sce_list <- lapply(individual_sce_list, reduce_dims) # own function

```


```{r}

identity_plots <- function(sce, title, identifier){
  
  id_1 <- colData(sce)[identifier][1]
  
  col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
  col_cts_fraction <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction]
  
  plot_1 <- umap_base(sce, color_by = "Identity_ref_all")+ # own function
    ggtitle(paste0(Object_ID, title))+
    scale_color_manual("Identity1", values = col_cts_all)
  
  plot_1l <- umap_legend(sce, color_by = "Identity_ref_all")+
    scale_color_manual("Identity1", values = col_cts_all)
  legend_1 <- get_legend(plot_1l)

    
  plot_2 <- umap_base(sce, color_by = "Identity_ref_fraction")+
    ggtitle(paste0(Object_ID, title))+
    scale_color_manual("Identity2", col_cts_fraction)
  
  plot_2l <- umap_legend(sce, color_by = "Identity_ref_fraction")+
    scale_color_manual("Identity2", col_cts_fraction)
  legend_2 <- get_legend(plot_2l)

  PLOT_1 <- ggarrange(
    plot_1,
    legend_1
  )
  PLOT_2 <- ggarrange(
    plot_2a,
    legend_2
  )
  
  return(list(PLOT_1, PLOT_2))
}

```

```{r fig.width=10, fig.height=5}

identity_individual_plots <- lapply(individual_sce_list, 
                                    identity_plots, 
                                    identifier = "Object_ID",
                                    title = "")

identity_individual_plots

```

More Info on preliminary cell types.

```{r}

lapply(individual_sce_list, function(sce){
  
  print(sce$Object_ID[1])
  print(table(sce$Identity_ref_all))
  print(table(sce$Identity_ref_fraction))
  
})

```

### Species level

```{r}

# generic dimensionality reduction
sce_07 <- reduce_dims(sce_07)

identity_species_plots <- identity_plots(sce_07, 
                                         title = ", merged only",
                                         identifier = "Species_ID")

identity_species_plots

print(sce_07$Species_ID[1])
print(table(sce_07$Identity_ref_all))
print(table(sce_07$Identity_ref_fraction))

```

# All metrics

```{r}

make_single_plots <- function(sce){
  
  species <- sce$Species_ID[1]
  
  plot_1 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Object_ID))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
  
   plot_2 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Batch_sequencing))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
 
   plot_3 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Batch_exp_day))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
  
  plot_4 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Date_collected))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))

  # check everything else
  plot_5 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Species_ID))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))+
    scale_color_manual("Species", values = col_spc)
  
  plot_6 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Age))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))+
    scale_color_manual("Age", values = col_age)
 
  plot_7 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Fraction))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))+
    scale_color_manual("Fraction", values = col_frc)
  
  plot_8 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Antibody_combination))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
 
  plot_9 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$doublet_score))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
  
   plot_10 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Keep_sample))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))
  
  plot_11 <- ggplot(data.frame(reducedDims(sce)[["UMAP"]]),
                    aes(x = X1, y = X2, color = sce$Identity))+
    geom_point(size = 0.01)+
    theme_classic()+
    ylab("UMAP 2")+
    xlab("UMAP 1")+
    theme(axis.text = element_blank())+
    theme(legend.key.size = unit(0.3, "cm"), 
          legend.key.width = unit(0.3, "cm"),
          legend.spacing = unit(0.06, "cm"), 
          legend.text = element_text(size = 7))+ 
    ggtitle(paste0(species, ", merged only"))+
    scale_color_manual("REF Identity", values = col_cts_all[names(col_cts_all) %in% mspr2$Identity])

  plot_list <- list(plot_1, plot_2, plot_3, plot_4, plot_5, plot_6, plot_7,
                    plot_8, plot_9, plot_10, plot_11)
  
  return(plot_list)
  
}

single_plot_list_007 <- make_single_plots(sce_07)
test <- make_single_plots(mspr2)

```

```{r fig.width = 10}
single_plot_list_007
```

### Facet grid UMAPs

To merge across ages and fractions using MNNcorrect, every age and fraction 
should have at least one overlapping BATCH.
Per definition, this is not the case for Object_ID as possible batch.

Check Fraction:

```{r}

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Batch_exp_day))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Fraction_ID))


ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Batch_sequencing))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Fraction_ID))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Object_ID))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_wrap(vars(mspr2$Fraction_ID))

```

Check Age:

```{r}

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Batch_exp_day))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Age_ID))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Batch_sequencing))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Age_ID))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Object_ID))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_wrap( vars(mspr2$Age_ID))

```

Check preliminary cell types

```{r}

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Identity_ref))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Fraction))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Identity_ref))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Age))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Identity_ref))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, " merged"))+
  facet_grid(cols = vars(mspr2$Batch_exp_day))

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Identity_ref))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_grid(cols = vars(mspr2$Batch_sequencing))

```


```{r fig.width = 10, fig.height= 10}

ggplot(data.frame(reducedDims(mspr2)[["UMAP"]]),
                 aes(x = X1, y = X2, color = mspr2$Identity))+
  geom_point(size = 0.01)+
  theme_classic()+
  ylab("UMAP 2")+
  xlab("UMAP 1")+
  theme(axis.text = element_blank())+
  theme(legend.key.size = unit(0.3, "cm"), 
        legend.key.width = unit(0.3, "cm"),
        legend.spacing = unit(0.06, "cm"), 
        legend.text = element_text(size = 7))+ 
  ggtitle(paste0(species, ", merged only"))+
  facet_wrap(vars(mspr2$Object_ID))+
    scale_color_manual("REF Identity", values = col_cts_all[names(col_cts_all) %in% mspr2$Identity])

names(col_cts_all)
unique(mspr2$Identity_ref)
```

