---
title: "Figure ideas Trkb"
author: "Lea Wölbert"
date: '2026-05-11'
output: html_document
---

This .Rmd file generates a lot of plots to get a deeper understanding
of possibly identities of TrkB+ cells and also to test which type of
plot can nicely show that.

```{r seed}
set.seed(37)
```

```{r load,  message = FALSE}
library(SingleCellExperiment)
library(tidyverse)
library(ggpubr)
library(DESeq2)
```

```{r load_objects}
# bulk RNAseq output
vsd_trkb <- readRDS(snakemake@input[["vsd_trkb"]])
results_df_trkb <- readRDS(snakemake@input[["results_df_trkb"]])

# corresponding single cell RNAseq data
sce_str <- readRDS(snakemake@input[["sce_str"]])

#stromal conserved signature gene list
sign_genes <- readRDS(snakemake@input[["sign_genes"]])
```

```{r}
pval_cutoff <- 0.05
lfc_cutoff <- 0.25

col_mice <- c(
      "05" = "red4",
      "06" = "orange3",
      "07" = "goldenrod2",
      "08" = "yellow3"
    )

#source("../../figures/determine_params.R")

# color table
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])
source(snakemake@params[["plotting"]])
```

## Normalised counts

```{r counts_table_trkb}

# get the counts from vsd, transformed only for visualisation
counts_trkb <- assays(vsd_trkb)[[1]]

counts_long_trkb <- counts_trkb %>% 
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_name") %>%
  tidyr::pivot_longer(
    names_to = "Object_ID", 
    values_to = "counts", 
    cols = c(1:ncol(counts_trkb)+1))

 counts_long_trkb <- counts_long_trkb  %>% 
    tidyr::separate(
      Object_ID,
      into = c(
        "Species_ID", 
        "Fraction_ID", 
        "Target_ID", 
        "Condition_ID", 
        "Mouse_ID"),
      sep = "_",
      remove = FALSE
  ) 
```

#### Signature genes

```{r trkb_signature_genes}

# get the significant results
results_df_trkb_significant <- results_df_trkb %>%
    filter(padj < pval_cutoff)

# get the Fibro signature genes
Fibro_sign <- sign_genes$Fibroblast$conserved_signature
Fibro_sign_005 <- Fibro_sign[which(Fibro_sign %in% 
  rownames(results_df_trkb_significant))]
sort(Fibro_sign_005)
sort(Fibro_sign[!Fibro_sign %in% 
  rownames(results_df_trkb_significant)])
table(Fibro_sign %in% rownames(results_df_trkb_significant))

# check how many signature genes were captured by the trkb+ population
for(i in names(sign_genes)){
    print(i)
    print(table(
        sign_genes[[i]]$conserved_signature %in% 
            rownames(results_df_trkb_significant)))
}  
# Fibro, fibro chondro, and fibro clusters are the only ones where 
# the majority of signature genes is detected as significant
```

```{r signature_gene_df}

signature_gene_df <- data.frame()

# make a df with all signature genes per cell type
for(i in names(sign_genes)){
    print(i)

    sign_df <- data.frame(
      "celltype" = rep(
        i, 
        length = length(sign_genes[[i]]$conserved_signature)),
      "genes" = sign_genes[[i]]$conserved_signature)
    
    signature_gene_df <- rbind(
      signature_gene_df,
      sign_df
      )
}  

signature_gene_df <- signature_gene_df %>%
  dplyr::mutate(
    is_significant = genes %in% rownames(results_df_trkb_significant))

res_df <- signature_gene_df %>%
  dplyr::group_by(celltype) %>%
  dplyr::summarise(
    n_true  = sum(is_significant == TRUE),
    n_false = sum(is_significant == FALSE),
    total_genes = n()
  ) %>%
  dplyr::mutate(percentage_significant = n_true/total_genes) 

res_df$celltype <- factor(
  res_df$celltype,
  levels = names(col_cts_str)
)

res_df
```

```{r plot_percentage, fig.width = 4, fig.height = 4}
plot_percentage <- ggplot(
  res_df, 
  aes(
    x = celltype,
    y = percentage_significant))+
  ggplot2::geom_col()+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90))+
  ggplot2::ylab("proportion signature genes, DE significant")+
  ggplot2::ylim(c(0,1))
plot_percentage
```

```{r test_related_celltypes}
osteo_sign <- sign_genes$Osteo$conserved_signature
osteo_sign_005 <- osteo_sign[which(osteo_sign %in% 
  rownames(results_df_trkb_significant))]
sort(osteo_sign_005)
sort(osteo_sign[!osteo_sign %in% rownames(results_df_trkb_significant)])

fibrochondro_sign <- sign_genes$'Fibro/Chondro'$conserved_signature
fibrochondro_sign_005 <- fibrochondro_sign[which(fibrochondro_sign %in% 
  rownames(results_df_trkb_significant))]
sort(fibrochondro_sign_005)
sort(fibrochondro_sign[!fibrochondro_sign %in% 
  rownames(results_df_trkb_significant)])

# the overlap of the detected signature genes is not super high
table(is.na(match(Fibro_sign_005, osteo_sign_005)))
table(is.na(match(fibrochondro_sign_005, osteo_sign_005)))
table(is.na(match(fibrochondro_sign_005, Fibro_sign_005)))
```

#### Gene lists

Compiled a lot of lists of genes from DESeq2 results lists,
from literature, from PCA loadings, or otherwise 
possibly interesting genes.

Roughly sorted into categories or ordered, but with any
marker gene, 100 % exact cell type specificity does not exist.  

I'm trying to get a rough understanding of the Neo1+ cells
by looking at a lot of genes, not just a handfull.

```{r trkb_genes}

# interesting genes in the DE list from trkb samples
# and genes I was interested in from literature
gene_vis_trkb <- unique(c(
  "Ntrk2",
  "Cd34",
  "Col5a3",
  "Abi3bp",
  "Abca8a",
  "Abca8b",
  "Lama2",
  "Lamb2",
  "Egfr",
  "Fgfr1",
  "Rarres2",
  "Ecm1",
  "S100a10",
  "S100a6",
  "Ackr3",
  "Dpt",
  "Nes",
  "Cspg4",
  "Pdgfrb",
  "Prg4",
  "Tnn",
  "Sox9",
  "Ptprz1",
  "Col1a1",
  "Col1a2",
  "Alpl",
  "Vcam1",
  "Flnc",
  "Tln2",
  "Ntrk3",

  "Neo1",
  "Runx2",
  "Postn",
  "Col1a1",
  "Col1a2",
  "Col12a1",
  "Col5a1",
  "Col5a2",

  "Mxra8",
  "Sparc",
  "Tnn",
  "Dlx5",

  "Pdpn",
  "Prelp",
  "Tnxb",
  "Fbn1",
  "Vcam1",
  "Pxdc1",
  "Bmp4",
  "Pdgfra",
  "Pdgfrb",
  "Pdgfc",
  "Pappa",
  "Nrp1",
  "F13a1",
  "Lama5",
  "Itgb8",
  "Itga8",
  "Adamts2",
  "Nr2f2",
  "Tagln",
  "Mcam",
  "Myh11",
  "Carmn",
  "Enpep",
  "Ddr2"
  ))


# significant, from Fibro list, and also somehow known
# nicer looking fibro genes
# MUST BE DE except for Ntrk2
gene_vis_TOP <- unique(c(
  "Ntrk2", # not significant
  "Abca8a",
  "Abca8b",
  "Egfr",
  "Anxa5",
  "Ank2",
  "Sod3",
  "Fgfr1",
  "Lama2",
  "Lama5",
  "Lamb2",
  "Ackr3",
  "Nes",
  "Dpt",
  "Nr2f2",
  "Fgfr1",
  "Fbn1",
  "Col1a1",
  "Cola12",
  "Col5a1",
  "Col5a2",
  "Col5a3",
  "Ahnak",
  "Cd34",
  "Dpt",
  "Ackr3",
  "Zeb2",

  "Pdpn",
  "Prelp",
  "Nr2f2",

  "Tagln",
  "Myh11",
  "Rgs5"
))

# list of genes I want to show,
# can include MSCs, fibro, osteo, other
final_pick <- unique(c(

  # MSC/Fibro identity, DE significant
  "Pdgfrb",
  "Cspg4", 
  "Dpt",
  "Fbn1",
  "Zeb2",

  # fibro/ECM
  "Col5a1",
  "Col6a2",
  "Ddr2",
  "Cd34",
  "Prelp",

  # from osteo signature list (shared)
  "Slit2",
  "Runx2",
  "Alpl",

  # Peri
  "Lama2",
  "Lama5",
  "Nr2f2",
  "Tagln",
  "Rgs5"

))

# add back Ntrk2 and also add negative controls
final_pick <- final_pick[final_pick %in% 
  rownames(results_df_trkb_significant)]

final_pick <- c("Ntrk2", final_pick, c(
  "Kitl", # Niche
  "Hoxa10", # Osteo
  "Pecam1", # EC
  "Procr" # HSC
))

trkb_trkb_compare_overlap <- unique(c(
  
  # general fibro/MSC markers
  "Pdgfrb", # DE
  "Nes",
  "Cspg4",

  "Col3a1", # not DE
  "Col6a3",
  "Sfrp2",
  "Ccdc80",
  "Nid2",
  "Mfap2",
  "Eln",
  "Fbln7",
  "Timp1",
  "Timp3",

  "Abca8a",
  "Abca8b",
  "Egfr",
  "Anxa5",
  "Ahnak",

  # trkb+ only

  "Cd34", # from fibro signature list
  "Rarres2",
  "Crip1",
  "Ugdh", # moderately fibro
  "Lmna",
  "Sema3c",
 
  "Pdgfc", # not from signature list but somewhow fibro related
  "Pappa",
  "Nrp1",
  "Rarres2",
  "F13a1",
  "Enpep",
  "Lama5",
  "Lama2",
  "Lamb2",
  "Itgb8",
  "Itga8",
  "Adamts2",
  "Adamts7",
  "Adamts8",
  "Nr2f2",
  "Fn1",
  "Itga11",

  # with SMC flavor
  "Tagln",
  "Myh11",
  "Carmn",
  "Mcam",
  "Myod1",

  # with Adipo flavor
  "Lpl",

  # with chondro flavor
  "Sox5",
  "Prg4",
  
  # with neuro flavor
  "Ntrk3",
  "Ngfr",
  "Plp",
  "Mapt",

  # from osteo signature list but only DE in trkb
  "Gpc1",
  "Enah",
  "Col5a1",
  "Zeb2",

  # trkb+ and trkb+ fibro shared

  "Cdh2",
  "Col5a1",
  "Col5a2",
  "Emilin1",
  "Mmp13",
  "Runx2",
  "Pcolce",
  "Ptprd",
  "Mrxa8",

  # not list
  "Postn",
  "Ibsp",
  # trkb+ and trkb+ osteo shared
  "Ptprz1",
  "Col1a1",
  "Sparc",
  "Pdpn",

  # osteo only but also not sign in trkb+
  "Ptgis",
  "Dlx5",
  "Dcn",
  "Mef2c",

  # trkb+ only (DE significant)
  "Sp7",
  "Spp1",
  "Satb2",
  "Blap",
  "Bglap2",
  "Pth1r",
  "Wnt16",
  "Fzd1",
  "Fzd8",
  "Lrp6",
  "Gli1",
  "Gli3",
  "Hoxa10",
  "Clec11a",
  "Gja1",
  "Mmp14",
  "Mmp2",
  "Thbs1",
  "Thbs3",
  "Smad3",
  "Smad7",
  "Smad9",
  "Tgfbr1",
  "Cdh11",
  "Col24a1",
  "Itgbl1",
  "Nt5e",
  "Serpinf1",
  "Lox",
  "Loxl3",
  "P4ha1",
  "P4ha2",
  "P3h1",
  "P3h3",

  # with chondro flavor
  "Cilp2",
  
  # other osteo-ish
  "Col6a2",
  "Pgx3",
  "Osr2",

  # chondro flavor
  "Chad",
  "Acan",
  "Sox9",
  "Matn3",
  "Fmod"
))

trkb_neo_compare_final <- unique(c(
  # unique trkb fibro genes
  "Cd34",
  "Rarres2",
  "Lama5",
  "Tagln",
  "Nr2f2",
  "Adamts8",
  "Myh11",

  # shared genes
  "Ackr3",
  "Dpt",
  "Sod3",
  "Col1a1",
  "Col5a1",
  "Col5a2",

  # shared osteo genes that look good in both
  "Pdpn",
  "Sdc4",
  "Slit2",
  "Mxra8",

  # not so good looking shared osteo genes
  "Lrp4",
  "Alpl",
  "Runx2",
  "Mmp13",

  # osteo only osteo genes
  "Satb2",
  "Mmp14",
  "Pth1r",
  "Sp7",

  # osteo only literature osteo genes
  # DE
  "Wnt16", 
  "Nt5e",
  "Fzd8", 
  "Spp1",
  "Smad9",
  "Hoxa10",
  "Dact1",
  "Cdh11"
))

negative_controls <- unique(c(
  # Adipo/CAR/Niche
  "Kitl",
  "Cxcl14",
  "Cxcl12",
  "Lepr",
  "Adipoq",
  "Plin1",
  "Lpl",
  "Pappa",

  # fibro
  "Pdgfra",
  "Lum",
  "Dcn",
  "Fn1",
  "Scara5",
  "Sema3c",

  # Peri/SMC or muscle
  "Acta2",
  "Acta1",
  "Tnnc2",
  "Tagln",
  "Rgs5",
  "Rgs4",
  "Myh11",
  "Steap4",

  # chondro
  "Acan",
  "Chad",
  "Creb5",
  "Comp",
  "Fmod",
  "Mgp",
  "Sox9",

  # ECs
  "Pecam1",
  "Cdh5",
  "Plvap",
  "Lyve1",
  "Ccl21a",
  "Efnb2",
  "Emcn",
  "Flt4",
  "Prox1",
  "Rgcc",
  "Sox17",
  "Stab2",

  # hematopoietic
  "Mecom",
  "Procr",
  "Meis1",
  "Mllt3",
  "Mpl",
  "Spi1",
  "Pbx1",
  "Hoxa9",
  "Flt3",
  "Mpo",
  "Irf8",
  "Ctsg",
  "Elane",
  "S100a8",
  "S100a9",
  "Ms4a2",
  "Ms4a6c",
  "Dntt",
  "Cd79b",
  "Il7r",
  "Pf4",
  "Zfpm1",
  "Klf1",
  "Epor",
  "Hba-a1"

))

negative_controls <- negative_controls[!negative_controls %in% 
  rownames(results_df_trkb_significant)]

gene_list_trkb <- list(
    gene_vis_trkb,
    Fibro_sign_005,
    Fibro_sign,
    gene_vis_TOP,
    final_pick,
    trkb_trkb_compare_overlap,
    trkb_neo_compare_final,
    negative_controls
)

gene_list_trkb_rev <- list(
    rev(gene_vis_trkb),
    rev(Fibro_sign_005),
    rev(Fibro_sign),
    rev(gene_vis_TOP),
    rev(final_pick),
    rev(trkb_trkb_compare_overlap),
    rev(trkb_neo_compare_final),
   rev( negative_controls)
)
```

#### process for visualisation

```{r counts_long_trkb_genes_list}
counts_long_trkb_genes_list <- lapply(
  gene_list_trkb, 
  function(gene_list){
  
  counts_long_trkb_genes <- counts_long_trkb %>%
    dplyr::filter(gene_name %in% gene_list)

  counts_long_trkb_genes$gene_name <- factor(
    counts_long_trkb_genes$gene_name,
    levels = gene_list)

  counts_long_trkb_genes <- counts_long_trkb_genes  %>% 
    tidyr::separate(
      Object_ID,
      into = c(
        "Species_ID", 
        "Fraction_ID", 
        "Target_ID", 
        "Condition_ID", 
        "Mouse_ID"),
      sep = "_",
      remove = FALSE
  ) 

  counts_long_trkb_genes <- counts_long_trkb_genes %>%
    dplyr::group_by(gene_name) %>%
    dplyr::mutate(scaled = as.numeric(scale(counts))) %>%
    dplyr::ungroup()

  return(counts_long_trkb_genes)
})
```

```{r counts_long_trkb_genes_list_rev}
counts_long_trkb_genes_list_rev <- lapply(
  gene_list_trkb_rev, 
  function(gene_list){
  
  counts_long_trkb_genes <- counts_long_trkb %>%
    dplyr::filter(gene_name %in% gene_list)

  counts_long_trkb_genes$gene_name <- factor(
    counts_long_trkb_genes$gene_name,
    levels = gene_list)

  counts_long_trkb_genes <- counts_long_trkb_genes  %>% 
    tidyr::separate(
      Object_ID,
      into = c(
        "Species_ID", 
        "Fraction_ID", 
        "Target_ID", 
        "Condition_ID", 
        "Mouse_ID"),
      sep = "_",
      remove = FALSE
  ) 

  counts_long_trkb_genes <- counts_long_trkb_genes %>%
    dplyr::group_by(gene_name) %>%
    dplyr::mutate(scaled = as.numeric(scale(counts))) %>%
    dplyr::ungroup()

  return(counts_long_trkb_genes)
})
```

#### bulk heatmaps

```{r plot1a, fig.width = 3, fig.height = 8}

plot1a <- ggplot2::ggplot(
    counts_long_trkb_genes_list_rev[[1]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("hand-picked")
plot1a
```

```{r plot1b, fig.width = 3, fig.height = 11}

plot1b <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[2]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Fibro signature genes, DE significant")
plot1b
```

```{r plot1c, fig.width = 3, fig.height = 14}

plot1c <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[3]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("all Fibro signature genes")
plot1c
```

```{r plot1_top, fig.width = 3, fig.height = 5}

plot1_top <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[4]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top Fibro genes")
plot1_top
```

```{r plot1_final, fig.width = 3, fig.height = 5}

plot1_final <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[5]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top picks")
plot1_final
```

```{r plot1_final_overlap, fig.width = 3, fig.height = 15}

plot1_final_overlap <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[6]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Osteo overlap")
plot1_final_overlap
```

```{r plot1_compare, fig.width = 3, fig.height = 6}

plot1_compare <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[7]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("trkb compare")
plot1_compare
```

```{r plot1_negative, fig.width = 3, fig.height = 5}

plot1_negative <- ggplot2::ggplot(
  counts_long_trkb_genes_list_rev[[8]],
    aes(  
      x = Object_ID,
      y = gene_name,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue",
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Negative")
plot1_negative
```

#### paired dotplot, expression

```{r plot1_paired_a, fig.width = 14, fig.height = 3}
plot1_paired_a <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[1]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("hand-picked from literature + DESeq2 list")
plot1_paired_a
```

```{r plot1_paired_b, fig.width = 18, fig.height = 3}
plot1_paired_b <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[2]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("Fibro signature genes, DE significant")
plot1_paired_b
```

```{r plot1_paired_c, fig.width = 22, fig.height = 3}
plot1_paired_c <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[3]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("all Fibro signature genes")
plot1_paired_c
```

```{r plot1_paired_top, fig.width = 8, fig.height = 3}
plot1_paired_top <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[4]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("Top Fibro genes")
plot1_paired_top
```

```{r plot1_paired_final, fig.width = 7, fig.height = 3.5}
plot1_paired_final <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[5]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("Top picks")
plot1_paired_final
```

```{r plot1_paired_overlap, fig.width = 25, fig.height = 3.5}
plot1_paired_overlap <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[6]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("Osteo overlap")
plot1_paired_overlap
```

```{r plot1_paired_compare, fig.width = 7, fig.height = 3.5}
plot1_paired_compare <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[7]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("trkb-Trkb compare")
plot1_paired_compare
```

```{r plot1_paired_negative, fig.width = 10, fig.height = 3.5}
plot1_paired_negative <- ggplot2::ggplot(
    counts_long_trkb_genes_list[[8]],
    aes(  
      x = Condition_ID,
      y = counts,
      color = Mouse_ID,
      group = Mouse_ID))+
    ggplot2::geom_point()+  
    ggplot2::geom_line(alpha = 0.6) +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~gene_name)+
    ggplot2::ggtitle("Negative")
plot1_paired_negative
```

#### selected dotplot, all cell types

```{r ct_trkb_df_long_scld}

# make a df that contains scaled counts of all signature genes of each cell
# type, for each sample

# prepare df for dotplot visualitation
ct_trkb_df <- signature_gene_df

# prepare counts
counts_trkb_prep <- counts_trkb %>%
  as.data.frame() %>%
  tibble::rownames_to_column("genes")

# add counts to each signature genes
ct_trkb_df <- left_join(ct_trkb_df, counts_trkb_prep, by = "genes")

ct_trkb_df_long <- ct_trkb_df %>% 
  tidyr::pivot_longer(
    names_to = "Object_ID", 
    values_to = "counts", 
    cols = c(4:10))

ct_trkb_df_long_scld <- ct_trkb_df_long %>%
  dplyr::group_by(genes) %>%
  #dplyr::mutate(averaged_counts = mean(as.numeric(counts), na.rm = TRUE)) %>%
  dplyr::mutate(scaled = as.numeric(scale(counts))) %>%
  dplyr::ungroup() %>%
  as.data.frame()

ct_trkb_df_long_scld <- ct_trkb_df_long_scld %>%
  dplyr::group_by(celltype, Object_ID) %>%
  dplyr::mutate(median_scaled_counts = median(as.numeric(scaled), na.rm = TRUE)) %>%
  dplyr::mutate(mean_scaled_counts = mean(as.numeric(scaled), na.rm = TRUE)) %>%
  dplyr::ungroup() %>%
  as.data.frame()

ct_trkb_df_long_scld <- ct_trkb_df_long_scld  %>% 
    tidyr::separate(
      Object_ID,
      into = c(
        "Species_ID", 
        "Fraction_ID", 
        "Target_ID", 
        "Condition_ID", 
        "Mouse_ID"),
      sep = "_",
      remove = FALSE
  ) 

ct_trkb_df_long_scld$celltype <- factor(
  ct_trkb_df_long_scld$celltype,
  levels = c(
    "Adipo/CAR",
    "Osteo",
    "Fibro/Chondro",
    "Fibroblast",
    "Art/Cap EC",
    "Sinusoid/Ven EC",
    "Sinusoid EC",
    "Peri/SMC"
  ))
```

```{r ct_trkb_df_long_scld_np}

ct_trkb_df_long <- ct_trkb_df_long  %>% 
    tidyr::separate(
      Object_ID,
      into = c(
        "Species_ID", 
        "Fraction_ID", 
        "Target_ID", 
        "Condition_ID", 
        "Mouse_ID"),
      sep = "_",
      remove = FALSE
  ) 

ct_trkb_df_long_scld_np <- ct_trkb_df_long %>%
  dplyr::group_by(genes) %>%
  dplyr::mutate(scaled = as.numeric(scale(counts))) %>%
  dplyr::ungroup() %>%
  as.data.frame()

ct_trkb_df_long_scld_np <- ct_trkb_df_long_scld_np %>%
  dplyr::group_by(celltype, Condition_ID) %>%
  dplyr::mutate(median_scaled_counts = median(as.numeric(scaled), na.rm = TRUE)) %>%
  dplyr::mutate(mean_scaled_counts = mean(as.numeric(scaled), na.rm = TRUE)) %>%
  dplyr::ungroup() %>%
  as.data.frame()

ct_trkb_df_long_scld_np$celltype <- factor(
  ct_trkb_df_long_scld_np$celltype,
  levels = c(
    "Adipo/CAR",
    "Osteo",
    "Fibro/Chondro",
    "Fibroblast",
    "Art/Cap EC",
    "Sinusoid/Ven EC",
    "Sinusoid EC",
    "Peri/SMC"
  ))

```

```{r plot_celltype_scaled_mean, fig.width = 9, fig.height = 5}
plot_celltype_scaled_mean <- ggplot2::ggplot(
    ct_trkb_df_long_scld,
    aes(  
      x = Object_ID,
      y = scaled,
      color = Mouse_ID))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld,
      aes(y = mean_scaled_counts, 
          color = Mouse_ID),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_mean
```

```{r plot_celltype_scaled_mean_color, fig.width = 9, fig.height = 5}
plot_celltype_scaled_mean_color <- ggplot2::ggplot(
    ct_trkb_df_long_scld,
    aes(  
      x = Object_ID,
      y = scaled,
      color = mean_scaled_counts))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld,
      aes(y = mean_scaled_counts, 
          color = mean_scaled_counts),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red",
      midpoint = 0)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_mean_color
```

```{r plot_celltype_scaled_median, fig.width = 9, fig.height = 5}
plot_celltype_scaled_median <- ggplot2::ggplot(
    ct_trkb_df_long_scld,
    aes(  
      x = Object_ID,
      y = scaled,
      color = Mouse_ID))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld,
      aes(y = median_scaled_counts, 
          color = Mouse_ID),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual(values = col_mice)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_median
```

```{r plot_celltype_scaled_median_color, fig.width = 9, fig.height = 5}
plot_celltype_scaled_median_color <- ggplot2::ggplot(
    ct_trkb_df_long_scld,
    aes(  
      x = Object_ID,
      y = scaled,
      color = median_scaled_counts))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld,
      aes(y = median_scaled_counts, 
          color = median_scaled_counts),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_median_color
```

now by condition (np = negative positive)

```{r plot_celltype_scaled_mean_color_np, fig.width = 6, fig.height = 5}
plot_celltype_scaled_mean_color_np <- ggplot2::ggplot(
    ct_trkb_df_long_scld_np,
    aes(  
      x = Condition_ID,
      y = scaled,
      color = mean_scaled_counts))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld_np,
      aes(y = mean_scaled_counts, 
          color = mean_scaled_counts),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_mean_color_np
```

```{r plot_celltype_scaled_median_color_np, fig.width = 6, fig.height = 5}
plot_celltype_scaled_median_color_np <- ggplot2::ggplot(
    ct_trkb_df_long_scld_np,
    aes(  
      x = Condition_ID,
      y = scaled,
      color = median_scaled_counts))+
    ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggplot2::geom_point(
      data = ct_trkb_df_long_scld_np,
      aes(y = median_scaled_counts, 
          color = median_scaled_counts),
      size = 4,
      position = position_dodge(width = 0.3))+
    ggplot2::geom_hline(yintercept = 0, color = "blue3") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::facet_grid(~celltype)

plot_celltype_scaled_median_color_np
```

#### try logfc instead of counts

```{r logfc_trkb_df}

# make a df that contains scaled counts of all signature genes of each cell
# type, for each sample

# prepare df for dotplot visualitation
logfc_trkb_df <- signature_gene_df

# add 200 random genes
random_df <- data.frame(
  "celltype" = rep("Random", 150),
  "genes" = rownames(results_df_trkb)[sample(1:nrow(results_df_trkb), 150)]
)

random_df <- random_df %>%
  mutate(is_significant = genes %in% rownames(results_df_trkb_significant))

logfc_trkb_df <- rbind(logfc_trkb_df, random_df)

# prepare results_df
results_df_trkb_prep <- results_df_trkb %>%
  rownames_to_column("genes")

# add counts to each signature genes
logfc_trkb_df <- dplyr::left_join(logfc_trkb_df, results_df_trkb_prep, by = "genes")

logfc_trkb_df <- logfc_trkb_df %>% 
  dplyr::mutate(
    padj_plot = ifelse(is.na(padj), 2, -log10(padj)),
    padj_missing = is.na(padj),
    is_significant_plot = ifelse(is.na(padj), "pval adj. = NA", is_significant))
  

logfc_trkb_df$celltype <- factor(
  logfc_trkb_df$celltype,
  levels = c(
    "Adipo/CAR",
    "Osteo",
    "Fibro/Chondro",
    "Fibroblast",
    "Art/Cap EC",
    "Sinusoid/Ven EC",
    "Sinusoid EC",
    "Peri/SMC",
    "Random"
  ))
```

```{r plot_logfc, fig.width = 7, fig.height = 5}
plot_logfc <- ggplot2::ggplot(
    logfc_trkb_df,
    aes(  
      x = is_significant,
      y = log2FoldChange,
      color = is_significant_plot,
      size = padj_plot))+
    #ggplot2::geom_violin(fill = "grey90", color = "grey90")+  
    ggbeeswarm::geom_quasirandom(alpha = 0.3)+
    ggplot2::geom_hline(yintercept = 0, color = "grey50") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_color_manual("", values = c(
      "FALSE" = "grey50",
      "pval adj. = NA" = "grey80",
      "TRUE" = "blue"))+
    ggplot2::facet_grid(~celltype)

plot_logfc

print(length(unique(logfc_trkb_df$genes[
  logfc_trkb_df$celltype != "Random" & 
  !is.na(logfc_trkb_df$log2FoldChange)])))
print(length(unique(signature_gene_df$genes)))
```

NOTE!! several genes were excluded from this analysis due to lack of detection 
in the bulkRNAseq dataset. 

```{r barplot_logfc, fog.width = 5, fig.height = 2.5}
barplot_logfc <- ggplot2::ggplot(
    logfc_trkb_df,
    aes(  
      x = celltype,
      fill = is_significant))+
    ggplot2::geom_bar(position = "dodge", alpha = 0.85)+  
    #ggplot2::geom_hline(yintercept = 0, color = "grey50") +
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.background = element_blank(),
      axis.text.x = element_text(angle = 90),
      strip.text.x = element_text(angle = 90, hjust = 0))+
    ggplot2::scale_fill_manual("", values = c(
      "FALSE" = "grey50",
      #"pval adj. = NA" = "grey50",
      "TRUE" = "blue"))
barplot_logfc
```

### SCE object

```{r prep_sce_str_trkb}

sce_str_msc <- sce_str[,sce_str$celltypes %in% c(
    "Adipo/CAR",
    "Osteo",
    "Fibroblast",
    #"Chondrocyte", # removed
    "Fibro/Chondro",
    "Peri/SMC"
)]

sce_str_msc$is_Fibro <- FALSE
sce_str_msc$is_Fibro[sce_str_msc$celltypes == "Fibroblast"] <- TRUE
```

##### Fibro cells compared to all others

```{r prep_sce_str_trkb_is_Fibro}

vis_df_sce_msc_genes_list_is_Fibro <- lapply(gene_list_trkb, function(gene_list){

  sce_str_msc_genes <- sce_str_msc[rownames(sce_str_msc) %in% gene_list,]

  agg_str_msc_genes <- scuttle::aggregateAcrossCells(
    sce_str_msc_genes, 
    id=colData(sce_str_msc_genes)[,c("is_Fibro")], 
    statistics = "mean",
    use.assay.type = "logcounts")

  vis_df_sce_msc_genes <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str_msc_genes)$logcounts))

  vis_df_sce_msc_genes <- tibble::rownames_to_column(
    vis_df_sce_msc_genes, 
    "is_Fibro")
  vis_df_sce_msc_genes <- tidyr::pivot_longer(
    vis_df_sce_msc_genes, 
    cols = c(2:ncol(vis_df_sce_msc_genes)), 
    values_to = "expression", 
    names_to = "gene")

  vis_df_sce_msc_genes <- vis_df_sce_msc_genes %>%
      dplyr::group_by(gene) %>%
      dplyr::mutate(scaled = as.numeric(scale(expression))) %>%
      dplyr::ungroup()

  vis_df_sce_msc_genes$gene <- factor(
    vis_df_sce_msc_genes$gene,
    levels = gene_list)

  return(vis_df_sce_msc_genes)

})
```

```{r plot2a, fig.width = 3, fig.height = 6}
plot2a <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[1]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("hand-picked from literature + DESeq2 list")
plot2a
```

```{r plot2b, fig.width = 3, fig.height = 8.8}
plot2b <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[2]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Fibro signature genes, DE significant")
plot2b
```

```{r plot2c, fig.width = 3, fig.height = 13}
plot2c <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[3]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("all Fibro signature genes")
plot2c
```

```{r plot2_top, fig.width = 3, fig.height = 4}
plot2_top <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[4]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top Fibro genes")
plot2_top
```

```{r plot2_final, fig.width = 3, fig.height = 4.5}
plot2_final <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[5]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top picks")
plot2_final
```

```{r plot2_overlap, fig.width = 3, fig.height = 10}
plot2_overlap <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[6]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Osteo overlap")
plot2_overlap
```

```{r plot2_compare, fig.width = 3, fig.height = 6}
plot2_compare <-  ggplot2::ggplot(
    vis_df_sce_msc_genes_list_is_Fibro[[7]],
    aes(  
      x = is_Fibro,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Osteo compare")
plot2_compare
```

##### Per cell type

```{r vis_df_sce_msc_genes_list_celltypes}

vis_df_sce_msc_genes_list_celltypes <- lapply(gene_list_trkb_rev, function(gene_list){

  sce_str_msc_genes <- sce_str_msc[rownames(sce_str_msc) %in% gene_list,]

  agg_str_msc_genes <- scuttle::aggregateAcrossCells(
    sce_str_msc_genes, 
    id=colData(sce_str_msc_genes)[,c("celltypes")], 
    statistics = "mean",
    use.assay.type = "logcounts")

  vis_df_sce_msc_genes <- base::as.data.frame(
    t(SummarizedExperiment::assays(agg_str_msc_genes)$logcounts))

  vis_df_sce_msc_genes <- tibble::rownames_to_column(vis_df_sce_msc_genes, "celltypes")
  vis_df_sce_msc_genes <- tidyr::pivot_longer(
    vis_df_sce_msc_genes, 
    cols = c(2:ncol(vis_df_sce_msc_genes)), 
    values_to = "expression", 
    names_to = "gene")

  vis_df_sce_msc_genes <- vis_df_sce_msc_genes %>%
      dplyr::group_by(gene) %>%
      dplyr::mutate(scaled = as.numeric(scale(expression))) %>%
      ungroup()

  vis_df_sce_msc_genes$gene <- factor(vis_df_sce_msc_genes$gene,
  levels = gene_list)
  vis_df_sce_msc_genes$celltypes <- factor(vis_df_sce_msc_genes$celltypes,
  levels = c(
      "Adipo/CAR",
      "Osteo",
      #"Chondrocyte", # removed
      "Fibro/Chondro",
      "Fibroblast",
      "Peri/SMC"))

  return(vis_df_sce_msc_genes)
})

```

```{r plot3a, fig.width = 3, fig.height = 7}
plot3a <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[1]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("hand-picked from literature + DESeq2 list")
plot3a
```

```{r plot3b, fig.width = 3, fig.height = 10}
plot3b <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[2]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Fibro signature genes, DE significant")
plot3b
```

```{r plot3c, fig.width = 3, fig.height = 14}
plot3c <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[3]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("all Fibro signature genes")
plot3c
```

```{r plot3_top, fig.width = 3, fig.height = 4.5}
plot3_top <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[4]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top Fibro genes")
plot3_top
```

```{r plot3_final, fig.width = 3, fig.height = 4}
plot3_final <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[5]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Top picks")
plot3_final
```

```{r plot3_overlap, fig.width = 3, fig.height = 12}
plot3_overlap <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[6]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Osteo Overlap")
plot3_overlap
```

```{r plot3_compare, fig.width = 3, fig.height = 5}
plot3_compare <- ggplot2::ggplot(
    vis_df_sce_msc_genes_list_celltypes[[7]],
    aes(  
      x = celltypes,
      y = gene,
      fill = scaled))+
    ggplot2::geom_tile()+
    ggplot2::theme_classic()+
    ggplot2::theme(axis.text.x = element_text(angle = 90))+
    ggplot2::scale_fill_gradient2(
      low = "blue", 
      mid = "white", 
      high = "red", 
      midpoint = 0)+
    ggplot2::ggtitle("Trkb-trkb Compare")
plot3_compare
```

## UMAPs

```{r umap, fig.width = 4, fig.height = 3}

umap1_gene_list <- gene_list_trkb[[5]] 
umap1_gene_list <- umap1_gene_list[umap1_gene_list %in% rownames(sce_str)]
for(gene in umap1_gene_list){
  print(umap_gene(sce_str, gene)+ # own function
    ggplot2::ggtitle(gene))
}

umap_gene(sce_str, "Ntrk2")
```

```{r umap2, fig.width = 4, fig.height = 3}

umap2_gene_list <- gene_list_trkb[[7]] 
umap2_gene_list <- umap2_gene_list[umap2_gene_list %in% rownames(sce_str)]

for(gene in umap2_gene_list){
  print(umap_gene(sce_str, gene)+
    ggplot2::ggtitle(gene))
}
```

```{r umap_trkb, fig.width = 4, fig.height = 3}

df_trkb <- assays(sce_str)[["logcounts"]][
  rownames(SummarizedExperiment::assays(sce_str)[["logcounts"]]) == "Ntrk2",] %>%
    as.matrix() %>%
    as.data.frame()

nrow(df_trkb)

df_trkb$UMAP1 <- SingleCellExperiment::reducedDims(sce_str)[["UMAP"]][,1]
df_trkb$UMAP2 <- SingleCellExperiment::reducedDims(sce_str)[["UMAP"]][,2]
nrow(df_trkb)

df_trkb <- df_trkb %>%
  arrange(V1)
nrow(df_trkb)

plot <- ggplot2::ggplot(
    df_trkb,
    aes(x = UMAP1, 
        y = UMAP2, 
        color = V1))+
    ggplot2::scale_color_gradientn(
      "TrkB/Ntrk2", 
      colors = c("grey30", "orange", "gold"))+
    ggplot2::geom_point(size = 0.01)+
    ggplot2::theme_classic()+
    ggplot2::theme(
      axis.text = element_blank(),
      axis.ticks = element_blank())+
    ggplot2::ylab("UMAP 2")+
    ggplot2::xlab("UMAP 1")
plot
```

## TrkB expression profile

```{r prep_trkb_violin}

sce_str_removed <- sce_str[,!sce_str$celltypes %in% c(
  "Chondrocyte",
  "Lymphatic EC"
)]
sce_str_trkb <- sce_str_removed[rownames(sce_str_removed) == "Ntrk2",]

logcounts_trkb <- logcounts(sce_str_trkb) %>%
  as.matrix() %>%
  t() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "barcode") %>%
  tibble::add_column(celltypes = colData(sce_str_trkb)$celltypes)
```

```{r trkb_violin, fig.width = 4, fig.height = 3}
trkb_violin <- ggplot2::ggplot(
  logcounts_trkb, 
  aes(
    x = celltypes,
    y = Ntrk2,
    fill = celltypes
  ))+
  ggplot2::geom_violin(
    draw_quantiles = c(0.25, 0.5, 0.75), 
    scale = "width" )+
  ggplot2::theme_classic()+
  ggplot2::theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  ggplot2::scale_fill_manual(values = col_cts_str)
trkb_violin
```

```{r session_info}
sessionInfo()
```
