---
title: "Figure 6"
output: html_document
date: "2024-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate Figure 6

Also includes one plot for Supplementary Figure 6
Also includes Supplementary Table for Zebrafish Pvalues

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load_packages, message = FALSE}
library(tidyverse, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(UpSetR, quietly = TRUE)
library(ggpubr, quietly = TRUE)
```

```{r data_paths}

# define data paths
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
brain_path <- "/scRNAseq/main_analysis/sce_objects/08_sce_brain"

colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")

source("../source/colors.R")
```

```{r params}
# global params
source("determine_params.R")
```

```{r export_paths}
output_pdf_b_plot <- base::paste0(
  base_path, "/manuscript1/figure6/figure6_b.pdf")
output_pdf_b_legend <- base::paste0(
  base_path, "/manuscript1/figure6/figure6_b_legend.pdf")
output_pdf_c_plot <- base::paste0(
  base_path, "/manuscript1/figure6/figure6_c.pdf")
output_pdf_d_plot <- base::paste0(
  base_path, "/manuscript1/figure6/figure6_d.pdf")
output_pdf_e_plot <- base::paste0(
  base_path, "/manuscript1/figure6/figure6_e.pdf")

output_pdf_s_a_plot <- base::paste0(
  base_path, "/manuscript1/supp_figure8/supp_figure8_allhs_heatmap.pdf")

output_table_zebrafish <- base::paste0(
  base_path, "/manuscript1/supplementary_table_zebrafish.csv")
```

# Print data

```{r print_integrated}
# nr of cells per dataset

integrated <- readRDS(
  paste0(base_path, "/metadata/scRNAseq/08_sce_brain/sample.combined_exc_4_species_integration.RDS"))

table(integrated@meta.data$orig.ident == "human")
table(integrated@meta.data$orig.ident == "macaque")
table(integrated@meta.data$orig.ident == "marmoset")
table(integrated@meta.data$orig.ident == "mouse")

integrated_mouse_metadata <- integrated@meta.data[integrated@meta.data$orig.ident == "mouse",]
nrow(integrated_mouse_metadata)

table(integrated_mouse_metadata$subclass_label != "L6 IT Car3")
```


```{r print_info}

# conserved marker info
cons_markers <- readRDS(
  paste0(base_path, brain_path, "/02_marker_conserved_primates.rds"))
markers_conservation <- readRDS(
  paste0(base_path, brain_path, "/02_marker_expression_primates.rds"))

# nDEG info
nDEGs <- readRDS(paste0(base_path, brain_path, "/01_list_nDEGs_all.rds"))

# get the signature
conserved_signature <- lapply(names(cons_markers), function(x){
  markers <- cons_markers[[x]]
  ndeg <- nDEGs[[x]]
  return(markers[markers %in% ndeg])
})
names(conserved_signature) <- names(cons_markers)

seu_sign <- unique(unlist(conserved_signature))
seu_sign
length(seu_sign)

# human-only markers info
human_markers <- do.call(rbind, markers_conservation) %>%
  as.data.frame() %>%
  filter(!is.na(human)) %>% 
  rownames() %>%
  sub("\\..*", "", .) %>%
  unique()

length(human_markers)
```

## Primate "own" reclustering heatmaps

Primate cell type identification using only primate-derived conserved identity
signature or conserved marker genes.
For figure 6b.

#### Prep

```{r b_prep1}
# establish order of clusters for plotting based on 04_recluster.R
# sig = conserved identity signature genes
factorize <- list(human = rev(c(9, 6, 4, 15, 2, 12, 10, 0, 5, 13, 8, 16, 1, 14, 3, 11, 7)),
                  marmoset = rev(c(0, 3, 12, 1, 5, 13, 4, 15, 16, 7, 9, 10, 8, 2, 6, 14)),
                  macaque = rev(c(3, 0, 2, 7, 4, 10, 5, 6, 8, 1, 9)))
```

```{r b_prep2}
# loop through all three primate species sig genes sets
# to create normalized matrices for plotting
plot_gg <- list()
for(sp in c("human", "marmoset", "macaque")){
  
  data <- readRDS(paste0(base_path, # reclustered primate datasets
                         brain_path,
                         "/04_recluster/", sp, "_reclust_sig.rds"))

  plot <- data@meta.data[,c("subclass_label", "seurat_clusters")] %>%
    table() %>% as.matrix()

  plot_gg[[sp]] <- plot %>%
    as.data.frame() %>%
    group_by(seurat_clusters) %>%
    mutate(clust_prop = Freq/sum(Freq)) %>%
    ungroup %>%
    group_by(subclass_label) %>%
    mutate(subclass_prop = Freq/sum(Freq),
           order = match(seurat_clusters, factorize[[sp]]))
}

plot_gg_all <- bind_rows(plot_gg, .id = 'species')
```

#### Plot Components

```{r b_base_plot}
b_plot <- ggplot(plot_gg_all, 
       aes(x = factor(subclass_label,
                      levels = c("L2/3 IT", "L5 IT", "L6 IT", "L6 IT Car3", 
                                 "L5 ET", "L5/6 NP", "L6 CT", "L6b")), 
           y = order,
           fill = clust_prop)) + 
  facet_wrap(~species, 
             ncol = 1,
             strip.position = "left") +
  geom_tile()
```

```{r b_set_theme, fig.width = 5, fig.height = 9}
b_plot_theme <- b_plot + 
      ggplot2::theme_classic() +
      theme_all+
      ggplot2::theme(
        strip.background = element_blank(),
        strip.text.y.left = element_blank(),
        legend.position = "none",
        #axis.title.x = element_blank(),
        axis.text = element_blank(),
        axis.text.x = element_text(
                          size = axis_text_size,
                          face = axis_text_face,
                          color = axis_text_color,
                          angle = 90,
                          hjust = 1,
                          vjust = 0.5),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank(),
        plot.title = element_text(
          size = axis_title_size,
          face = axis_title_face,
          hjust = 0.5,
          color = col_cons_long["conserved identity signature"]))  +
       ggplot2::ggtitle("Signature genes") +
       ggplot2::xlab("Reference annotation") +
       ggplot2::ylab("Reclustered with") +
       ggplot2::scale_fill_gradientn(
            colors = mycolors_to1,
            limits = c(0, 1),
            breaks = c(0, 0.5, 1))
b_plot_theme
```

#### Assemble and Export

```{r b_plot_export}
pdf(output_pdf_b_plot, width = 5, height = 9)
b_plot_theme
dev.off()
```

## Mouse "other" reclustering

Mouse cell type identification using only primate marker genes.
Figure 6c.

#### Prep

```{r c_prep1}
# establish order of clusters for plotting based on 04_recluster.R
# hs = human only markers; sig = conserved identity signature genes
factorize_mouse <- list(hs = c(11, 12, 9, 5, 8, 1, 3, 4, 2, 14, 6, 10, 7, 15, 17, 16,0),
                        sig = c(0, 9, 6, 3, 7, 10, 8, 2, 4, 5, 11, 1))
```

```{r c_prep2}
# loop through all three primate species and across core and sig genes sets
# to create normalized matrices for plotting
plot_gg_mouse <- list()
for(genes in c("hs", "sig")){
  data <- readRDS(paste0(base_path, # reclustered mouse datasets, depending on geneset used for reclustering
                         brain_path, 
                         "/04_recluster/mouse_reclust_", genes, ".rds"))
  
  plot <- data@meta.data[,c("subclass_label", "seurat_clusters")] %>%
    table() %>% as.matrix()
  
  plot_gg_mouse[[genes]] <- plot %>%
    as.data.frame() %>%
    group_by(seurat_clusters) %>%
    mutate(clust_prop = Freq/sum(Freq)) %>%
    ungroup %>%
    group_by(subclass_label) %>%
    mutate(subclass_prop = Freq/sum(Freq),
           order = match(seurat_clusters, factorize_mouse[[genes]])) %>%
    filter(subclass_label != "L6 IT Car3")
}
```

#### Plot Components

```{r c_base_plot}
plotted_gg_mouse <- list()
for(genes in c("hs", "sig")){
  plotted_gg_mouse[[genes]] <- ggplot(plot_gg_mouse[[genes]], 
         aes(x = factor(subclass_label, 
                              levels = rev(c("L2/3 IT", "L5 IT", "L6 IT", 
                                         "L5 ET", "L5/6 NP", "L6 CT", "L6b"))), 
             y = order,
             fill = clust_prop)) + 
    geom_tile()
}
plotted_gg_mouse
```

```{r c_set_theme_hs}
# this will be a supplementary plot
plotted_gg_mouse[["hs"]] <- plotted_gg_mouse[["hs"]] + 
    ggplot2::theme_classic() +
    theme_all+
    ggplot2::theme(
      legend.position = "none",
      axis.text = element_blank(),
      axis.text.x = element_text(
                    size = axis_text_size,
                    face = axis_text_face,
                    color = axis_text_color,
                    angle = 90,
                    hjust = 1,
                    vjust = 0.5),
      axis.ticks.y = element_blank(),
      axis.line.y = element_blank(),
      axis.title.y = element_text(
        size = axis_title_size,
        face = axis_title_face,
        color = axis_title_color),
      plot.title = element_text(
        size = axis_title_size,
        face = axis_title_face,
        hjust = 0.5,
        color = col_cons_long["BL6 markers"])) +
    ggplot2::ggtitle("Total human markers") +
    ggplot2::scale_fill_gradientn(
      colors = mycolors_to1,
      limits = c(0, 1),
      breaks = c(0, 0.5, 1))+
    ggplot2::ylab("Reclustered with") +
    ggplot2::xlab("Reference annotation")
plotted_gg_mouse[["hs"]]
```


```{r c_set_theme_sig}
plotted_gg_mouse[["sig"]] <- plotted_gg_mouse[["sig"]] + 
    theme_all+
    ggplot2::theme(
      legend.position = "none",
      axis.title.x = element_blank(),
      axis.text.y = element_blank(),
      axis.text.x = element_text(
        angle = 90,
        hjust = 0.5,
        vjust = 0.5),
      axis.ticks.y = element_blank(),
      axis.line.y = element_blank(),
      axis.title.y = element_text(
        size = axis_title_size,
        face = axis_title_face,
        color = axis_title_color),
      plot.title = element_text(
        size = axis_title_size,
        face = axis_title_face,
        hjust = 0.5,
        color = col_cons_long["conserved identity signature"]))  +
     ggplot2::ggtitle("Signature genes") +
     ggplot2::scale_fill_gradientn(
          colors = mycolors_to1,
          limits = c(0, 1),
          breaks = c(0, 0.5, 1)) +
     ggplot2::ylab("Reclustered with")
plotted_gg_mouse[["sig"]]
```

#### Export

```{r c_plot_export}
pdf(output_pdf_c_plot, width = 3.5, height = 4)
plotted_gg_mouse[["sig"]]
dev.off()
```

#### Supplementary

Export Supplementary Figure 6 A

```{r s_a_plot_export}
pdf(output_pdf_s_a_plot, width = 5, height = 4)
plotted_gg_mouse[["hs"]]
dev.off()
```

## Barplot of accurately assigned cells

Number of mouse cortex cells that are able to be accurately re-annotated using 
clustering based on human marker genes or primate identity signature genes.
Figure 6d.

#### Prep

```{r d_prep1}
mouse_hs <- readRDS(paste0(base_path, brain_path, 
                           "/04_recluster/mouse_reclust_hs.rds")) # mouse cells reclustered with human marker genes
mouse_sig <- readRDS(paste0(base_path, brain_path, 
                            "/04_recluster/mouse_reclust_sig.rds")) # mouse cells reclustered with signature genes
```

```{r d_prep2}
plot_hs <- mouse_hs@meta.data[,c("subclass_label", "seurat_clusters")] %>%
  table() %>% 
  as.matrix() %>%
  as.data.frame() %>%
  group_by(seurat_clusters) %>%
  mutate(clust_prop = Freq/sum(Freq),
         clust_assigned = (max(clust_prop)==clust_prop)) %>%
  filter(subclass_label != "L6 IT Car3") %>%
  group_by(clust_assigned, subclass_label) %>%
  mutate(Count = sum(Freq)) %>%
  select(-seurat_clusters, -clust_prop, -Freq) %>%
  unique()

plot_sig <- mouse_sig@meta.data[,c("subclass_label", "seurat_clusters")] %>%
  table() %>% 
  as.matrix() %>%
  as.data.frame() %>%
  group_by(seurat_clusters) %>%
  mutate(clust_prop = Freq/sum(Freq),
         clust_assigned = (max(clust_prop)==clust_prop)) %>%
  filter(subclass_label != "L6 IT Car3") %>%
  group_by(clust_assigned, subclass_label) %>%
  mutate(Count = sum(Freq)) %>%
  select(-seurat_clusters, -clust_prop, -Freq) %>%
  unique()

all_data <- rbind(
  mutate(plot_hs, genes = "human markers"),
  mutate(plot_sig, genes = "identity signature")
)
```

#### Assemble

```{r d_construct, fig.width = 4, fig.height = 4}
d_plot <- ggplot(all_data,
       aes(y = factor(genes, 
                      levels = c("identity signature", "human markers")),
           x = Count,
           fill = clust_assigned)) +
  facet_wrap(~factor(subclass_label, 
                     levels = c("L2/3 IT", "L5 IT", "L6 IT", 
                                "L5 ET", "L5/6 NP", "L6 CT", "L6b")), 
             ncol = 1,
             strip.position = "right") +
  geom_col(position = "stack") +
  theme_classic() +
  theme(strip.text.y = element_text(angle = 0),
    strip.background = element_rect(color = "white"),
    axis.text.y = element_text(color = c(
      col_cons_long["conserved identity signature"], 
      col_cons_long["total human markers"]))) +
  scale_x_continuous(breaks = c(0, 1000, 2000)) +
  labs(y = NULL, 
       x = "Mouse cells", 
       fill = "Accurately\nAssigned") +
  scale_fill_manual(values = c(max(mycolors_to1_grey), 
                               min(mycolors_to1_grey)))
head(all_data)
```

```{r d_construct2, fig.width = 5, fig.height = 5}
d_plot <- ggplot(
  all_data,
  aes(x = factor(genes, levels = c("identity signature", "human markers")),
      y = Count,
      fill = clust_assigned)) +
  geom_col(position = "stack") +
  facet_wrap(~factor(subclass_label, 
                     levels = c("L2/3 IT", "L5 IT", "L6 IT", 
                                "L5 ET", "L5/6 NP", "L6 CT", "L6b")), 
             ncol = 7,
             strip.position = "top") +
  scale_x_discrete(position = "top") +
  scale_y_reverse(breaks = c(0, 1000, 2000)) +
  theme_all+
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_text(
      angle = 90,
      hjust = 0,
      color = c(col_cons_long["conserved identity signature"], 
                col_cons_long["BL6 markers"])),
    strip.text.x.top = element_blank(),
    strip.background = element_rect(color = "white")) +
  labs(x = NULL, 
       y = "Mouse cells", 
       fill = "Accurately\nAssigned") +
  scale_fill_manual(values = c("grey80", 
                               "grey30"))
d_plot
```

#### Export

```{r d_plot_export}
pdf(output_pdf_d_plot, width = 3.5, height = 5)
d_plot
dev.off()
```

## Upset plot of accurately assigned cells.

Summarized consensus annotation information across clustering methods
Figure 6e.

#### Prep

```{r e_prep}
plot_hs_clust <- mouse_hs@meta.data[,c("subclass_label", "seurat_clusters")] %>%
  table() %>% 
  as.matrix() %>%
  as.data.frame() %>%
  group_by(seurat_clusters) %>%
  mutate(clust_assigned = (max(Freq)==Freq)) %>%
  filter(clust_assigned == T) %>%
  select(-clust_assigned, -Freq)

plot_sig_clust <- mouse_sig@meta.data[,c("subclass_label", "seurat_clusters")] %>%
  table() %>% 
  as.matrix() %>%
  as.data.frame() %>%
  group_by(seurat_clusters) %>%
  mutate(clust_assigned = (max(Freq)==Freq)) %>%
  filter(clust_assigned == T) %>%
  select(-clust_assigned, -Freq)

test_hs <- lapply(mouse_hs@meta.data$seurat_clusters, 
               function(x){
                 plot_hs_clust[which(plot_hs_clust$seurat_clusters == x), "subclass_label"]}) %>% 
  unlist()
mouse_hs@meta.data$correct <- mouse_hs@meta.data$subclass_label == test_hs

test_sig <- lapply(mouse_sig@meta.data$seurat_clusters, 
               function(x){
                 plot_sig_clust[which(plot_sig_clust$seurat_clusters == x), "subclass_label"]}) %>% 
  unlist()
mouse_sig@meta.data$correct <- mouse_sig@meta.data$subclass_label == test_sig

rownames <- rownames(mouse_sig@meta.data)
mouse_sig@meta.data <- merge(mouse_sig@meta.data, # correct.x is signature markers
                             mouse_hs@meta.data[,c("nCount_RNA", "nFeature_RNA", "sample_id", "correct")], # correct.y is human markers
                             by = c("nCount_RNA", "nFeature_RNA", "sample_id")) %>%
  mutate(correct = paste(correct.x, correct.y))
rownames(mouse_sig@meta.data) <- rownames

# table(mouse_sig@meta.data$correct)
# 
# FALSE FALSE  FALSE TRUE  TRUE FALSE   TRUE TRUE 
# 388         306         398        7150 
```

#### Assemble

```{r e_construct, fig.height = 4, fig.width = 5}
input <- c(
  'Bakken&signature&human' = as.integer(table(mouse_sig@meta.data$correct)[4]),
  'Bakken&signature' = as.integer(table(mouse_sig@meta.data$correct)[3]),
  'Bakken&human' = as.integer(table(mouse_sig@meta.data$correct)[2]),
  'Bakken' = as.integer(table(mouse_sig@meta.data$correct)[1])
)

ups <- upset(fromExpression(input), 
             order.by = "degree",
             sets = c("human", "signature", "Bakken"),
             keep.order = T,
             mainbar.y.label = 'Mouse cells',
      number.angles = 0, 
      text.scale = 2, 
      point.size = 3, 
      line.size = 1
      )

skip_set_size_plot <- function(ups) {
  main <- ups$Main_bar
  ## find panel grob
  panel_idx <- grep("panel", main$layout$name, fixed = TRUE)
  ## find text grob
  text_idx <- which(
    vapply(main$grobs[[panel_idx]]$children, 
           \(x) inherits(x, "text"), 
           logical(1)))
  tG <- main$grobs[[panel_idx]]$children[[text_idx]]
  tG$label <- scales::label_percent(0.1)(as.numeric(tG$label) / 
                                                  sum(as.numeric(tG$label)))
  main$grobs[[panel_idx]]$children[[text_idx]] <- tG
  grid::grid.newpage()
  grid::grid.draw(gridExtra::arrangeGrob(main, 
                                         ups$Matrix, 
                                         heights = ups$mb.ratio))
}
print(skip_set_size_plot(ups))
```

#### Export

```{r e_plot_export}
pdf(output_pdf_e_plot, width = 3.5, height = 4)
skip_set_size_plot(ups)
dev.off()
```

# Zebrafish

```{r zebrafish}

pval_df_zeb_sign_vs_rand <-  readRDS(
  paste0(base_path, "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/08_expp/sign-vs-rand_zeb_all_hspc"))
pval_df_zeb_mark_vs_rand <-  readRDS(
  paste0(base_path, "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/08_expp/mark-vs-rand_zeb_all_hspc"))
pval_df_zeb_mmms_vs_rand <-  readRDS(
  paste0(base_path, "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/08_expp/mmms-vs-rand_zeb_all_hspc"))

pval_df_zeb_mark_vs_signrand <-  readRDS(
  paste0(base_path, "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/08_expp/mark-vs-signrand_zeb_all_hspc"))
pval_df_zeb_mmms_vs_signrand <-  readRDS(
  paste0(base_path, "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/08_expp/mmms-vs-signrand_zeb_all_hspc"))

zeb_df <- rbind(
  pval_df_zeb_sign_vs_rand,
  pval_df_zeb_mark_vs_rand,
  pval_df_zeb_mmms_vs_rand,
  pval_df_zeb_mark_vs_signrand,
  pval_df_zeb_mmms_vs_signrand
)

zeb_df_narrow <- zeb_df[,c(1, 7, 13)]
zeb_df_narrow$pval
zeb_df_narrow$pval_corrected <- stats::p.adjust(zeb_df_narrow$pval, method = "BH")

zeb_df_narrow$type[zeb_df_narrow$type == "mean_prop_cells_cluster"] <- "Mean Proportion"
zeb_df_narrow$type[zeb_df_narrow$type == "adjusted_rand_index"] <- "Adjusted Rand Index"
zeb_df_narrow$type[zeb_df_narrow$type == "variation_information"] <- "Variation of Information"

zeb_df_narrow$comparison[zeb_df_narrow$comparison == "sign-vs-rand"] <- "Signature vs. random"
zeb_df_narrow$comparison[zeb_df_narrow$comparison == "mark-vs-rand"] <- "Conserved markers vs. random"
zeb_df_narrow$comparison[zeb_df_narrow$comparison == "mmms-vs-rand"] <- "BL6-only markers vs. random"
zeb_df_narrow$comparison[zeb_df_narrow$comparison == "mark-vs-signrand"] <- "Conserved markers vs signature + random"
zeb_df_narrow$comparison[zeb_df_narrow$comparison == "mmms-vs-signrand"] <- "BL6-only markers vs signature + random"

zeb_df_narrow <- zeb_df_narrow[,c(1, 2, 4)]
print(zeb_df_narrow)

readr::write_delim(zeb_df_narrow, output_table_zebrafish, delim = ";", append = FALSE)
```

# Utils

```{r session_info}
utils::sessionInfo()
```
