---
title: "Supplementary Figure 5"
date: '2024-11-28'
output: html_document
---

For Figure 3.
Figure S4 = reclustering score selection and optimization
Figure S5 = reclustering permuation tests

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
library(ggpubr, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r load_manual}
# list of reclustered HSPC SCE objects
sce_list_hsc <-  base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/03_recl/reso_reclustered_list_hsc"))
```

```{r load_manual_other2}
#-------------------------------------------------------------------------------
#### OTHER MOUSE DATASET 2 (HSPC)
#### MUS_WEINREB_HSPC

dataset_o2 <- "mus_weinreb_hspc"

# list of seurat objects with reclustering vectors for each gene set and resolution
seu_hsc_recl_list_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_mus_weinreb_hspc_list"))

# list of dataframes of re-clustering scores for each gene set and resolution
score_df_hsc_list_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis//04_signatures/02_reclustering_other/04_rcls/score_df_mus_weinreb_hspc_list"))

### PERMUTATION ###

# perm score DF from background permutation (SIGN)
psdf_hsc_sign_rand_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/05_psig/perm_score_df_mus_weinreb_hspc"))

# perm score DF from background permutation (MARK)

psdf_hsc_mark_rand_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/06_pmrk/perm_score_df_mus_weinreb_hspc"))

# perm score DF from background permutation (MMMS)

psdf_hsc_mmms_rand_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/07_pmms/perm_score_df_mus_weinreb_hspc"))


# perm score DF from permutation comparing to SIGN (MMMS-SIGN)

psdf_hsc_mmms_signrand_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/07_perg/perm_score_df_mmms_mus_weinreb_hspc"))

```

Load the rest later to save some space

```{r load_rest}

# df containing the resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  base_path,
  "/metadata/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")

source("../source/colors.R")
source("../source/plotting.R")
```

```{r pdf_output}
# define output paths for pdf export
output_own_heatmap <- base::paste0(
  base_path,
  "/manuscript1/supp_figure5/sfigure5_a.pdf")

output_weinreb_heatmap <- base::paste0(
  base_path,
  "/manuscript1/supp_figure5/sfigure5_b.pdf")

output_weinreb_perm <- base::paste0(
  base_path,
  "/manuscript1/supp_figure5/sfigure5_c.pdf")

output_mark_signrand_perm <- base::paste0(
  base_path,
  "/manuscript1/supp_figure5/sfigure5_d.pdf")
```

```{r params}
# global params
source("determine_params.R")


## CHECK SCORE MAX!!
score2 <- "variation_information"
score1 <- "adjusted_rand_index"
score3 <- "mean_prop_cells_cluster"

score2_long <- "Variation\nof information"
score1_long <- "Adjusted\nRand index"
score3_long <- "Mean prop.\ncells/cluster"

nr_genes_title <- "# re-clustering\ngenes"
max_genes <- 1000

point_plot_size <- 7
```

Functions

```{r density_plot_function}
make_density_plot <- function(
  score_df,
  perm_score_df,
  color_cons,
  ylabel
  ){

  # make temp plot to extract maximum y value
  p_temp <- ggplot2::ggplot(
    perm_score_df,
    aes(x = value))+
    geom_density()
  
  # calculate the maximum y value (density at its peak)
  density_data <- ggplot2::ggplot_build(p_temp)$data[[1]]
  max_y <- max(density_data$y)
  
  # define the y value for the point as 125% of the maximum y value
  y_value <- 1.25 * max_y
  
  # define the y limits for a nicer plot + y value to place the break/label
  y_lim_low <- -0.25 * max_y
  y_lim_high <- 1.5 * max_y
  y_lim_break <- (y_lim_low + y_lim_high)/2 
  
  # add to score DF
  score_df$y_value <- y_value
  
  # plot
  p_return <- ggplot2::ggplot(
    perm_score_df,
    aes(x = value))+
    geom_density(
      fill = col_cons_long[["random genes"]],
      color = col_cons_long[["random genes"]])+
    geom_vline(
      color = "grey60",
      xintercept = score_df$value,
      linetype = "dashed"
    )+
    geom_point(
      inherit.aes = FALSE, 
      shape = 23,
      fill = color_cons,
      data = score_df,
      aes(x = value, y = y_value),
      size = point_plot_size,
      color = color_cons)+
    scale_y_continuous(
      limits = c(y_lim_low, y_lim_high),
      breaks = y_lim_break, 
      labels = ylabel)
  
  return(p_return)
}
```

# Prep

Weinreb getting the object reclustered with the preferred resolution

```{r subset_o2}
resolution_df_o2 <- resolution_df[resolution_df$dataset == dataset_o2,]

selected_seu_list_o2 <- lapply(seu_hsc_recl_list_o2, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  resdf_temp <- resolution_df_o2[
    resolution_df_o2$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
})
```

```{r rm_02}
rm(seu_hsc_recl_list_o2)
gc()
```

# Reclustering Heatmaps

### HSC Signature

```{r heatmap_own_sign_mat}

sce_sign <- sce_list_hsc$`0.65`$sce_signt

mat_sign <- base::table(
  sce_sign$celltypes, 
  sce_sign$reclustered)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign)){
  mat_sign[,i] <- mat_sign[,i]/base::sum(mat_sign[,i])
}
sign_df <- base::as.data.frame(mat_sign)

# manually factorise clusters for nice order
sign_df$Var2 <- factor(
  sign_df$Var2,
  levels = base::rev(c(5, 4, 3, 2, 6, 8, 7, 1)))
```

```{r heatmap_own_sign_base_plot}
heatmap_own_sign_base_plot <- ggplot2::ggplot(
  sign_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r heatmap_own_sign_theme_plot}
heatmap_own_sign_theme_plot <- heatmap_own_sign_base_plot+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    plot.title = element_blank(),   
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),)+
  ggplot2::xlab("Reference cell types")+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
heatmap_own_sign_theme_plot
```

### HSC Conserved Markers

```{r mat_mark}

sce_mark <- sce_list_hsc$`0.45`$sce_consm

mat_mark <- base::table(
  sce_mark$celltypes, 
  sce_mark$reclustered)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_mark)){
  mat_mark[,i] <- mat_mark[,i]/base::sum(mat_mark[,i])
}
mark_df <- base::as.data.frame(mat_mark)

# manually factorise clusters for nice order
mark_df$Var2 <- factor(
  mark_df$Var2,
  levels = base::rev(c(4, 6, 3, 2, 5, 8, 7, 1)))
```

```{r heatmap_own_mark_base_plot}
heatmap_own_mark_base_plot <- ggplot2::ggplot(
  mark_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r heatmap_own_mark_theme_plot}
heatmap_own_mark_theme_plot <- heatmap_own_mark_base_plot+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
heatmap_own_mark_theme_plot
```

### HSC Conserved Markers

```{r mat_mmms}

sce_mmms <- sce_list_hsc$`0.65`$sce_mmusm

mat_mmms <- base::table(
  sce_mmms$celltypes, 
  sce_mmms$reclustered)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_mmms)){
  mat_mmms[,i] <- mat_mmms[,i]/base::sum(mat_mmms[,i])
}
mmms_df <- base::as.data.frame(mat_mmms)

# manually factorise clusters for nice order
mmms_df$Var2 <- factor(
  mmms_df$Var2,
  levels = base::rev(c(12, 6, 8, 5, 4, 3, 2, 11, 10, 7, 13, 9, 1)))
```

```{r heatmap_own_mmms_base_plot}
heatmap_own_mmms_base_plot <- ggplot2::ggplot(
  mmms_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r heatmap_own_mmms_theme_plot}
heatmap_own_mmms_theme_plot <- heatmap_own_mmms_base_plot+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank(),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
heatmap_own_mmms_theme_plot
```

```{r heatmap_own_assemble, fig.height = 7.8, fig.width = 3.5}

heatmap_own_theme_plot <- ggpubr::ggarrange(
  ncol = 1,
  heatmap_own_mmms_theme_plot,
  heatmap_own_mark_theme_plot,
  heatmap_own_sign_theme_plot,
  align = "v",
  heights = c(1, 1, 2.2)
)

heatmap_own_theme_plot
```

```{r export_heatmap}
pdf(output_own_heatmap, width = 3.5, height = 7.8)
heatmap_own_theme_plot
dev.off()
```

```{r rm_own}
rm(sce_list_hsc)
gc()
```

### Weinreb Heatmap

#### Signature

```{r mat_sign_o2}
mat_sign_o2 <- base::table(selected_seu_list_o2$seu_sign$cell_type, 
                           selected_seu_list_o2$seu_sign$seurat_clusters)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign_o2)){
  mat_sign_o2[,i] <- mat_sign_o2[,i]/base::sum(mat_sign_o2[,i])
}
sign_df_o2 <- base::as.data.frame(mat_sign_o2)

# manually factorise clusters
sign_df_o2$Var2 <- factor(
  sign_df_o2$Var2,
  levels = base::rev(c(3, 4, 6, 8, 9, 7, 5, 2, 11, 10, 1, 0)))
```

```{r sign_base_plot_o2}
sign_base_plot_o2 <- ggplot2::ggplot(
  sign_df_o2, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r sign_theme_plot}
sign_theme_plot_o2 <- sign_base_plot_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::xlab("Reference cell types")
sign_theme_plot_o2
```

#### Conserved markers

```{r mat_mark_o2}
mat_mark_o2 <- base::table(selected_seu_list_o2$seu_mark$cell_type, 
                           selected_seu_list_o2$seu_mark$seurat_clusters)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_mark_o2)){
  mat_mark_o2[,i] <- mat_mark_o2[,i]/base::sum(mat_mark_o2[,i])
}
mark_df_o2 <- base::as.data.frame(mat_mark_o2)

# manually factorise clusters
mark_df_o2$Var2 <- factor(
  mark_df_o2$Var2,
  levels = base::rev(c(2, 3, 7, 8, 9, 4, 5, 6, 10, 12, 11, 1, 0)))
```

```{r mark_base_plot_o2}
mark_base_plot_o2 <- ggplot2::ggplot(
  mark_df_o2, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r mark_theme_plot_o2}
mark_theme_plot_o2 <- mark_base_plot_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
mark_theme_plot_o2
```


#### BL6-only

```{r mat_mmms_o2}
mat_mmms_o2 <- base::table(selected_seu_list_o2$seu_mmms$cell_type, 
                           selected_seu_list_o2$seu_mmms$seurat_clusters)

for(i in 1:ncol(mat_mmms_o2)){
  mat_mmms_o2[,i] <- mat_mmms_o2[,i]/base::sum(mat_mmms_o2[,i])
}
mmms_df_o2 <- base::as.data.frame(mat_mmms_o2)

mmms_df_o2$Var2 <- factor(
  mmms_df_o2$Var2,
  levels = base::rev(
    c(16, 2, 15, 3, 12, 8, 9, 10, 5, 4, 6, 11, 14, 13, 7, 1, 0)))
```

```{r mmms_base_plot_o2}
mmms_base_plot_o2 <- ggplot2::ggplot(
  mmms_df_o2, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r mmms_theme_plot_o2}
mmms_theme_plot_o2 <- mmms_base_plot_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks.y = element_blank())+
  ggplot2::scale_fill_gradientn(
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))
mmms_theme_plot_o2
```

### Assemble

```{r weinreb_assemble, fig.width = 3.5, fig.height = 7.8}

weinreb_heatmap_theme_plot <- ggpubr::ggarrange(
  ncol = 1,
  mmms_theme_plot_o2,
  mark_theme_plot_o2,
  sign_theme_plot_o2,
  align = "v",
  heights = c(1, 1, 2.3)
)

weinreb_heatmap_theme_plot
```

```{r export_heatmap_weinreb}
pdf(output_weinreb_heatmap, width = 3.5, height = 7.8)
weinreb_heatmap_theme_plot
dev.off()
```

# Permutation plots

## Prep

### Weinreb 

```{r o2_scoredf_1}

selected_score_df_list_o2 <- lapply(score_df_hsc_list_o2, function(score_df_list){
  
  conslev_curr <- score_df_list[[1]]$conservation_level[1]
  resdf_temp <- resolution_df_o2[
    resolution_df_o2$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  score_df_return <- score_df_list[[res_curr]]
  return(score_df_return)
})
score_df_hsc_o2 <- dplyr::bind_rows(selected_score_df_list_o2)

# flip VI
score_df_hsc_o2$value[
  score_df_hsc_o2$type == "variation_information"] <- 0-score_df_hsc_o2$value[
    score_df_hsc_o2$type == "variation_information"]
```

Get the actual scores

Signature

```{r o2_scoredf_2}
# signature actual scores
score_df_sign_o2 <- score_df_hsc_o2[
  score_df_hsc_o2$conservation_level == "conserved_signature",]
score_df_sign_o2$conservation_level_long <- "conserved identity signature"

score_df_sign_score_1_o2 <- score_df_sign_o2[
  score_df_sign_o2$type == score1,]
score_df_sign_score_2_o2 <- score_df_sign_o2[
  score_df_sign_o2$type == score2,]
score_df_sign_score_3_o2 <- score_df_sign_o2[
  score_df_sign_o2$type == score3,]
```

Mark

```{r o2_scoredf_3}
# conserved markers actual scores
score_df_mark_o2 <- score_df_hsc_o2[
  score_df_hsc_o2$conservation_level == "conserved_markers",]
score_df_mark_o2$conservation_level_long <- "conserved markers"

score_df_mark_score_1_o2 <- score_df_mark_o2[
  score_df_mark_o2$type == score1,]
score_df_mark_score_2_o2 <- score_df_mark_o2[
  score_df_mark_o2$type == score2,]
score_df_mark_score_3_o2 <- score_df_mark_o2[
  score_df_mark_o2$type == score3,]
```

BL6-only

```{r o2_scoredf_4}
# bl6 markers actual scores
score_df_mmms_o2  <- score_df_hsc_o2[
  score_df_hsc_o2$conservation_level == "mmusall_markers",]
score_df_mmms_o2$conservation_level_long <- "BL6 markers"

score_df_mmms_score_1_o2 <- score_df_mmms_o2[
  score_df_mmms_o2$type == score1,]
score_df_mmms_score_2_o2 <- score_df_mmms_o2[
  score_df_mmms_o2$type == score2,]
score_df_mmms_score_3_o2 <- score_df_mmms_o2[
  score_df_mmms_o2$type == score3,]
```

get the permuted scores (vs random) in that same order

```{r o2_scoredf_5}
# permuted signature and bl6 vs random scores
head(psdf_hsc_sign_rand_o2)
head(psdf_hsc_mark_rand_o2)
head(psdf_hsc_mmms_rand_o2)
psdf_hsc_sign_rand_o2$variation_information <- 0-
  psdf_hsc_sign_rand_o2$variation_information
psdf_hsc_mark_rand_o2$variation_information <- 0-
  psdf_hsc_mark_rand_o2$variation_information
psdf_hsc_mmms_rand_o2$variation_information <- 0-
  psdf_hsc_mmms_rand_o2$variation_information

# permuted bl6 vs signrand scores
head(psdf_hsc_mmms_signrand_o2)
psdf_hsc_mmms_signrand_o2$variation_information <- 0-
  psdf_hsc_mmms_signrand_o2$variation_information

```

```{r o2_scoredf_6}
# sign vs rand
psdf_sign_rand_long_o2 <- tidyr::pivot_longer(
  psdf_hsc_sign_rand_o2, 
  c = c("mean_prop_cells_cluster", "adjusted_rand_index", 
        "variation_information"),
  values_to = "value",
  names_to = "type")

psdf_sign_rand_long_o2$conservation_level_long <- base::rep(
  "conserved identity signature", nrow(psdf_sign_rand_long_o2))

psdf_sign_rand_score_1_o2 <- psdf_sign_rand_long_o2[
  psdf_sign_rand_long_o2$type == score1,]
psdf_sign_rand_score_2_o2 <- psdf_sign_rand_long_o2[
  psdf_sign_rand_long_o2$type == score2,]
psdf_sign_rand_score_3_o2 <- psdf_sign_rand_long_o2[
  psdf_sign_rand_long_o2$type == score3,]
```

```{r o2_scoredf_7}
# mark vs rand
psdf_mark_rand_long_o2 <- tidyr::pivot_longer(
  psdf_hsc_mark_rand_o2, 
  c = c("mean_prop_cells_cluster", "adjusted_rand_index", 
        "variation_information"),
  values_to = "value",
  names_to = "type")

psdf_mark_rand_long_o2$conservation_level_long <- base::rep(
  "conserved markers", nrow(psdf_mark_rand_long_o2))

psdf_mark_rand_score_1_o2 <- psdf_mark_rand_long_o2[
  psdf_mark_rand_long_o2$type == score1,]
psdf_mark_rand_score_2_o2 <- psdf_mark_rand_long_o2[
  psdf_mark_rand_long_o2$type == score2,]
psdf_mark_rand_score_3_o2 <- psdf_mark_rand_long_o2[
  psdf_mark_rand_long_o2$type == score3,]
```

```{r o2_scoredf_8}
# mmms vs rand
psdf_mmms_rand_long_o2 <- tidyr::pivot_longer(
  psdf_hsc_mmms_rand_o2, 
  c = c("mean_prop_cells_cluster", 
        "adjusted_rand_index",
        "variation_information"),
  values_to = "value",
  names_to = "type")

psdf_mmms_rand_long_o2$conservation_level_long <- base::rep(
  "conserved identity signature", nrow(psdf_mmms_rand_long_o2))

psdf_mmms_rand_score_1_o2 <- psdf_mmms_rand_long_o2[
  psdf_mmms_rand_long_o2$type == score1,]
psdf_mmms_rand_score_2_o2 <- psdf_mmms_rand_long_o2[
  psdf_mmms_rand_long_o2$type == score2,]
psdf_mmms_rand_score_3_o2 <- psdf_mmms_rand_long_o2[
  psdf_mmms_rand_long_o2$type == score3,]
```

```{r o2_scoredf_9}
# mmms vs signrand
psdf_mmms_signrand_long_o2 <- tidyr::pivot_longer(
  psdf_hsc_mmms_signrand_o2,
  c = 1:ncol(psdf_hsc_mmms_signrand_o2),
  values_to = "value",
  names_to = "type")
psdf_mmms_signrand_long_o2$conservation_level_long = base::rep(
  "BL6 markers", nrow(psdf_mmms_signrand_long_o2))

psdf_mmms_signrand_score_1_o2 <- psdf_mmms_signrand_long_o2[
  psdf_mmms_signrand_long_o2$type == score1,]
psdf_mmms_signrand_score_2_o2 <- psdf_mmms_signrand_long_o2[
  psdf_mmms_signrand_long_o2$type == score2,]
psdf_mmms_signrand_score_3_o2 <- psdf_mmms_signrand_long_o2[
  psdf_mmms_signrand_long_o2$type  == score3,]
```

Boxplot prep

```{r o2_numberdf_1}

# get the number of genes used for each gene set
nr_sign_used_o2 <- selected_score_df_list_o2$seu_sign$nr_genes_used[1]
nr_sign_rand_o2 <- psdf_hsc_sign_rand_o2$nr_genes_used[1]

nr_mark_used_o2 <- selected_score_df_list_o2$seu_mark$nr_genes_used[1]
nr_mark_used_o2 <- psdf_hsc_mark_rand_o2$nr_genes_used[1]
nr_mark_left_o2 <- nr_mark_used_o2 - nr_sign_used_o2

nr_mmms_used_o2 <- selected_score_df_list_o2$seu_mmms$nr_genes_used[1]
nr_mmms_perm_o2 <- psdf_hsc_mmms_rand_o2$nr_genes_used[1]

nr_mmms_left_o2 <- nr_mmms_used_o2 - nr_sign_used_o2
nr_mmms_rand_o2 <- nr_mmms_left_o2

print(nr_sign_used_o2)
print(nr_mmms_used_o2)
```

```{r o2_numberdf_2}
vis_df_sign_rand_o2 <- base::data.frame(
  type_y = c(
    "conserved identity signature",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used_o2,
    nr_sign_rand_o2))

vis_df_sign_rand_o2$type_y <- factor(
  vis_df_sign_rand_o2$type_y, levels = base::rev(c(
  "conserved identity signature",
  "permutation gene set")))
vis_df_sign_rand_o2$type_fill <- factor(
  vis_df_sign_rand_o2$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "random genes")))
```

```{r o2_numberdf_3}

vis_df_mark_rand_o2 <- base::data.frame(
  type_y = c(
    "conserved markers",
    "conserved markers",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "conserved markers",
    "random genes"),
  number = c(
    nr_sign_used_o2,
    nr_mark_left_o2,
    nr_mark_used_o2))

vis_df_mark_rand_o2$type_fill <- factor(
  vis_df_mark_rand_o2$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "conserved markers",
  "random genes")))
vis_df_mark_rand_o2$type_y <- factor(
  vis_df_mark_rand_o2$type_y, levels = base::rev(c(
  "conserved markers",
  "permutation gene set")))
```

```{r o2_numberdf_4}

vis_df_mmms_rand_o2 <- base::data.frame(
  type_y = c(
    "BL6 markers",
    "BL6 markers",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "BL6 markers",
    "random genes"),
  number = c(
    nr_sign_used_o2,
    nr_mmms_left_o2,
    nr_mmms_perm_o2))

vis_df_mmms_rand_o2$type_fill <- factor(
  vis_df_mmms_rand_o2$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "BL6 markers",
  "random genes")))
vis_df_mmms_rand_o2$type_y <- factor(
  vis_df_mmms_rand_o2$type_y, levels = base::rev(c(
  "BL6 markers",
  "permutation gene set")))
```

```{r o2_numberdf_5}

vis_df_mmms_signrand_o2 <- base::data.frame(
  type_y = c(
    "BL6 markers",
    "BL6 markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "BL6 markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used_o2,
    nr_mmms_left_o2,
    nr_sign_used_o2,
    nr_mmms_rand_o2))

vis_df_mmms_signrand_o2$type_fill <- factor(
  vis_df_mmms_signrand_o2$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "BL6 markers",
  "random genes")))
vis_df_mmms_signrand_o2$type_y <- factor(
  vis_df_mmms_signrand_o2$type_y, levels = base::rev(c(
  "BL6 markers",
  "permutation gene set")))

```

### Markers vs. Signature: Own, O1 and O2

#### Load

```{r load_mark_signrand}

#### COMPARING MARK-SIGNRAND

####### OWN

# SCE with reclustering vectors for each gene set
sce_hsc_recl <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/03_recl/sce_hsc"))

# HSC actual scores
score_df_hsc <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/04_rcls/score_df_hsc"))

# perm score DF from permutation comparing to SIGN (MARK-SIGN)

psdf_hsc_mark_signrand <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/06_perg/perm_score_df_mark_hsc"))

###### OTHER 1

dataset_o1 <- "mus_tm_bonemarrow"

# list of seurat objects with reclustering vectors for each gene set and resolution
seu_hsc_recl_list_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_mus_tm_bonemarrow_list"))

# list of dataframes of re-clustering scores for each gene set and resolution
score_df_hsc_list_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/04_rcls/score_df_mus_tm_bonemarrow_list"))

# perm score DF from permutation comparing to SIGN (MARK-SIGN)

psdf_hsc_mark_signrand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/07_perg/perm_score_df_mark_mus_tm_bonemarrow"))


###### OTHER 2

# perm score DF from permutation comparing to SIGN (MARK-SIGN)

psdf_hsc_mark_signrand_o2 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/07_perg/perm_score_df_mark_mus_weinreb_hspc"))
```

#### OWN

Actual MARK scores

```{r own_scoredf}
score_df_hsc$value[score_df_hsc$type == "variation_information"] <- 0-
  score_df_hsc$value[score_df_hsc$type == "variation_information"] 

score_df_mark <- score_df_hsc[
  score_df_hsc$conservation_level == "conserved_markers",]
score_df_mark$conservation_level_long <- "conserved markers"

score_df_mark_score_1 <- score_df_mark[score_df_mark$type == score1,]
score_df_mark_score_2 <- score_df_mark[score_df_mark$type == score2,]
score_df_mark_score_3 <- score_df_mark[score_df_mark$type == score3,]
```

Permuted scores MARK-SIGN

```{r own_perm_score_df_sign1}
psdf_hsc_mark_signrand$variation_information <- 0-
  psdf_hsc_mark_signrand$variation_information
```

```{r own_perm_score_df_sign2}
psdf_mark_signrand_long <- tidyr::pivot_longer(
  psdf_hsc_mark_signrand,
  c = 1:ncol(psdf_hsc_mark_signrand),
  values_to = "value",
  names_to = "type")

psdf_mark_signrand_long$conservation_level_long = base::rep(
  "conserved markers", nrow(psdf_mark_signrand_long))

psdf_mark_signrand_score_1 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type == score1,]
psdf_mark_signrand_score_2 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type == score2,]
psdf_mark_signrand_score_3 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type == score3,]
```

Prepare boxplot

```{r own_numberdf_1}
# get the number of genes used for each gene set

# actual re-clustering
nr_sign_used_own <- sce_hsc_recl$cluster_signt_genes_used[1]
nr_mark_used_own <- sce_hsc_recl$cluster_consm_genes_used[1]

# for gene sets, get left-over random gene number and also gene number of
# gene sets that are not cons sign
nr_mark_left_own <- nr_mark_used_own - nr_sign_used_own

```

```{r own_numberdf_2}
vis_df_mark_signrand_own <- base::data.frame(
  type_y = c(
    "conserved markers",
    "conserved markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "conserved markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used_own,
    nr_mark_left_own,
    nr_sign_used_own,
    nr_mark_left_own))

vis_df_mark_signrand_own$type_fill <- factor(
  vis_df_mark_signrand_own$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "conserved markers",
  "random genes")))
vis_df_mark_signrand_own$type_y <- factor(
  vis_df_mark_signrand_own$type_y, levels = base::rev(c(
  "conserved markers",
  "permutation gene set")))
vis_df_mark_signrand_own
```

#### Other 1 

```{r subset_o1}

# extract only the info on the object re-clustered with the chosen resolution
# this is also required for the heatmaps

# seurat object
resolution_df_o1 <- resolution_df[resolution_df$dataset == dataset_o1,]

selected_seu_list_o1 <- lapply(seu_hsc_recl_list_o1, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  resdf_temp <- resolution_df_o1[
    resolution_df_o1$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
})

# original score DF
selected_score_df_list_o1 <- lapply(score_df_hsc_list_o1, function(score_df_list){
  
  conslev_curr <- score_df_list[[1]]$conservation_level[1]
  resdf_temp <- resolution_df_o1[
    resolution_df_o1$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  score_df_return <- score_df_list[[res_curr]]
  return(score_df_return)
  
})
score_df_hsc_o1 <- dplyr::bind_rows(selected_score_df_list_o1)

# flip VI
score_df_hsc_o1$value[
  score_df_hsc_o1$type == "variation_information"] <- 0-score_df_hsc_o1$value[
    score_df_hsc_o1$type == "variation_information"]
```

```{r remove_o1_2}
rm(seu_hsc_recl_list_o1)
gc()
```

Mark actual scores

```{r o1_scoredf_1}
score_df_mark_o1 <- score_df_hsc_o1[
  score_df_hsc_o1$conservation_level == "conserved_markers",]
score_df_mark_o1$conservation_level_long <- "conserved markers"

score_df_mark_score_1_o1 <- score_df_mark_o1[
  score_df_mark_o1$type == score1,]
score_df_mark_score_2_o1 <- score_df_mark_o1[
  score_df_mark_o1$type == score2,]
score_df_mark_score_3_o1 <- score_df_mark_o1[
  score_df_mark_o1$type == score3,]

```

Mark vs signature + random permuted scores

```{r o1_scoredf_2}

# flip VI 
psdf_hsc_mark_signrand_o1$variation_information <- 0-
  psdf_hsc_mark_signrand_o1$variation_information

psdf_mark_signrand_long_o1 <- tidyr::pivot_longer(
  psdf_hsc_mark_signrand_o1,
  c = 1:ncol(psdf_hsc_mark_signrand_o1),
  values_to = "value",
  names_to = "type")
psdf_mark_signrand_long_o1$conservation_level_long = base::rep(
  "conserved markers", nrow(psdf_mark_signrand_long_o1))

psdf_mark_signrand_score_1_o1 <- psdf_mark_signrand_long_o1[
  psdf_mark_signrand_long_o1$type == score1,]
psdf_mark_signrand_score_2_o1 <- psdf_mark_signrand_long_o1[
  psdf_mark_signrand_long_o1$type == score2,]
psdf_mark_signrand_score_3_o1 <- psdf_mark_signrand_long_o1[
  psdf_mark_signrand_long_o1$type == score3,]

```

```{r o1_numberdf_1}

# get the number of genes used for each gene set
nr_sign_used_o1 <- selected_score_df_list_o1$seu_sign$nr_genes_used[1]

nr_mark_used_o1 <- selected_score_df_list_o1$seu_mark$nr_genes_used[1]
nr_mark_perm_o1 <- psdf_hsc_mark_signrand_o1$nr_genes_used[1]

nr_mark_left_o1 <- nr_mark_perm_o1 - nr_sign_used_o1
nr_mark_rand_o1 <- nr_mark_left_o1
```

```{r o1_numberdf_2}
vis_df_mark_signrand_o1 <- base::data.frame(
  type_y = c(
    "conserved markers",
    "conserved markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "conserved markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used_o1,
    nr_mark_left_o1,
    nr_sign_used_o1,
    nr_mark_rand_o1))

vis_df_mark_signrand_o1$type_fill <- factor(
  vis_df_mark_signrand_o1$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "conserved markers",
  "random genes")))
vis_df_mark_signrand_o1$type_y <- factor(
  vis_df_mark_signrand_o1$type_y, levels = base::rev(c(
  "conserved markers",
  "permutation gene set")))
vis_df_mark_signrand_o1
```

### Other 2

Actual MARK scores

```{r o2_scoredf_0}
# score_df_mark_o2 etc. already generated for the Weinreb plots
```

Permuted MARK-SIGN scores

```{r o2_scoredf_10}

# flip VI 
psdf_hsc_mark_signrand_o2$variation_information <- 0-
  psdf_hsc_mark_signrand_o2$variation_information

psdf_mark_signrand_long_o2 <- pivot_longer(
  psdf_hsc_mark_signrand_o2,
  c = 1:ncol(psdf_hsc_mark_signrand_o2),
  values_to = "value",
  names_to = "type")
psdf_mark_signrand_long_o2$conservation_level_long = base::rep(
  "conserved markers", nrow(psdf_mark_signrand_long_o2))

psdf_mark_signrand_score_1_o2 <- psdf_mark_signrand_long_o2[
  psdf_mark_signrand_long_o2$type == score1,]
psdf_mark_signrand_score_2_o2 <- psdf_mark_signrand_long_o2[
  psdf_mark_signrand_long_o2$type == score2,]
psdf_mark_signrand_score_3_o2 <- psdf_mark_signrand_long_o2[
  psdf_mark_signrand_long_o2$type == score3,]
```

```{r o1_numberdf_6}

# get the number of genes used for each gene set
nr_mark_used_o2 <- selected_score_df_list_o2$seu_mark$nr_genes_used[1]
nr_mark_perm_o2 <- psdf_hsc_mark_signrand_o2$nr_genes_used[1]

nr_mark_left_o2 <- nr_mark_perm_o2 - nr_sign_used_o2
nr_mark_rand_o2 <- nr_mark_left_o2
```

```{r o2_numberdf_7}
vis_df_mark_signrand_o2 <- base::data.frame(
  type_y = c(
    "conserved markers",
    "conserved markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "conserved markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used_o2,
    nr_mark_left_o2,
    nr_sign_used_o2,
    nr_mark_rand_o2))

vis_df_mark_signrand_o2$type_fill <- factor(
  vis_df_mark_signrand_o2$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "conserved markers",
  "random genes")))
vis_df_mark_signrand_o2$type_y <- factor(
  vis_df_mark_signrand_o2$type_y, levels = base::rev(c(
  "conserved markers",
  "permutation gene set")))
vis_df_mark_signrand_o2
```

### Mins and Maxs all together

```{r minmax}

all_ari <- c(
  
 
  score_df_hsc$value[score_df_hsc$type == "adjusted_rand_index"],
  psdf_hsc_mark_signrand$adjusted_rand_index,

  # other dataset 1
  score_df_hsc_o1$value[score_df_hsc_o1$type == "adjusted_rand_index"],
  psdf_hsc_mark_signrand_o1$adjusted_rand_index,

  #other dataset 2
  score_df_hsc_o2$value[score_df_hsc_o2$type == "adjusted_rand_index"],

  psdf_hsc_sign_rand_o2$adjusted_rand_index,
  psdf_hsc_mark_rand_o2$adjusted_rand_index,
  psdf_hsc_mmms_rand_o2$adjusted_rand_index,
  psdf_hsc_mmms_signrand_o2$adjusted_rand_index,
  psdf_hsc_mark_signrand_o2$adjusted_rand_index
)

all_vi <- c(
  
  # actual mark scores OWN and permutated scores mark-signrand OWN
  score_df_hsc$value[score_df_hsc$type == "variation_information"],
  psdf_hsc_mark_signrand$variation_information,

  # other dataset 1
  score_df_hsc_o1$value[score_df_hsc_o1$type == "variation_information"],
  psdf_hsc_mark_signrand_o1$variation_information,

  # other dataset 2
  score_df_hsc_o2$value[score_df_hsc_o2$type == "variation_information"],

  psdf_hsc_sign_rand_o2$variation_information,
  psdf_hsc_mark_rand_o2$variation_information,
  psdf_hsc_mmms_rand_o2$variation_information,
  psdf_hsc_mmms_signrand_o2$variation_information,
  psdf_hsc_mark_signrand_o2$variation_information
)

all_meanprop <- c(
  
 
  score_df_hsc$value[score_df_hsc$type == "mean_prop_cells_cluster"],
  psdf_hsc_mark_signrand$mean_prop_cells_cluster,

  # other dataset 1
  score_df_hsc_o1$value[score_df_hsc_o1$type == "mean_prop_cells_cluster"],
  psdf_hsc_mark_signrand_o1$mean_prop_cells_cluster,

  #other dataset 2
  score_df_hsc_o2$value[score_df_hsc_o2$type == "mean_prop_cells_cluster"],

  psdf_hsc_sign_rand_o2$mean_prop_cells_cluster,
  psdf_hsc_mark_rand_o2$mean_prop_cells_cluster,
  psdf_hsc_mmms_rand_o2$mean_prop_cells_cluster,
  psdf_hsc_mmms_signrand_o2$mean_prop_cells_cluster,
  psdf_hsc_mark_signrand_o2$mean_prop_cells_cluster
)

min_ari <- 0
max_ari <- base::max(all_ari) + 0.1

stopifnot(all_vi <= 0)
min_vi <- base::min(all_vi- 0.5)
max_vi <- 0

min_mp <- 0
max_mp <- base::max(all_meanprop) + 0.1

max_genes <- 1000

print(score1)
print(score2)
print(score3)

score1_min <- min_ari
score1_max <- max_ari
score2_min <- min_vi
score2_max <- max_vi
score3_min <- min_mp
score3_max <- max_mp
```

## Plots Weinreb

### Boxplots

#### MMMS, MARK, and SIGN vs Rand

```{r base_plot_bp1_o2}
base_plot_bp1_o2 <- ggplot2::ggplot(
  vis_df_mmms_rand_o2,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp1_o2
```

```{r theme_plot_bp1_o2}
theme_plot_bp1_o2<- base_plot_bp1_o2+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size),
    axis.title = element_blank(),
    axis.text.y = element_blank())+
  ggplot2::ggtitle(nr_genes_title)
```

```{r base_plot_bp2_o2}
base_plot_bp2_o2 <- ggplot2::ggplot(
  vis_df_mark_rand_o2,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp2_o2
```

```{r theme_plot_bp2_o2}
theme_plot_bp2_o2<- base_plot_bp2_o2+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_blank())
```

```{r base_plot_bp3_o2}
base_plot_bp3_o2 <- ggplot2::ggplot(
  vis_df_sign_rand_o2,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      pattern = type_fill,
      color = type_fill,
      pattern_alpha = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp3_o2
```

```{r theme_plot_bp3_o2}
theme_plot_bp3_o2 <- base_plot_bp3_o2+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_blank())

```

#### MMMS vs SIGNRAND boxplots

```{r base_plot_bp1_o2_signrand}
base_plot_bp1_o2_signrand <- ggplot2::ggplot(
  vis_df_mmms_signrand_o2,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill,
      pattern = type_fill,
      pattern_alpha = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp1_o2_signrand
```

```{r theme_plot_bp1_o2_signrand}
theme_plot_bp1_o2_signrand <- base_plot_bp1_o2_signrand+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_blank())
```

### Adjusted Rand index (Score 1)

#### MMMS and SIGN vs Rand

```{r base_plot_ai1_o2}
base_plot_ai1_o2 <- make_density_plot(
  score_df = score_df_mmms_score_1_o2,
  perm_score_df = psdf_mmms_rand_score_1_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_ai1_o2
```

```{r theme_plot_ai1_o2}
theme_plot_ai1_o2 <- base_plot_ai1_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

```{r base_plot_ai2_o2}
base_plot_ai2_o2 <- make_density_plot(
  score_df = score_df_mark_score_1_o2,
  perm_score_df = psdf_mark_rand_score_1_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "Conserved markers"
)
base_plot_ai2_o2
```

```{r theme_plot_ai2_o2}
theme_plot_ai2_o2 <- base_plot_ai2_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

```{r base_plot_ai3_o2}
base_plot_ai3_o2 <- make_density_plot(
  score_df = score_df_sign_score_1_o2,
  perm_score_df = psdf_sign_rand_score_1_o2,
  color_cons = col_cons_long["conserved identity signature"],
  ylabel = "Conserved identity signature"
)
base_plot_ai3_o2
```

```{r theme_plot_ai3_o2}
theme_plot_ai3_o2 <- base_plot_ai3_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

#### MMMS vs signrand

```{r base_plot_ai3_o2_signrand}
base_plot_ai3_o2_signrand <- make_density_plot(
  score_df = score_df_mmms_score_1_o2,
  perm_score_df = psdf_mmms_signrand_score_1_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_ai3_o2_signrand
```

```{r theme_plot_ai1_o2_signrand}
theme_plot_ai1_o2_signrand <- base_plot_ai3_o2_signrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

### Variation information (Score 2)

#### MMMS and SIGN vs RAND

```{r base_plot_vi1_o2}
base_plot_vi1_o2 <- make_density_plot(
  score_df = score_df_mmms_score_2_o2,
  perm_score_df = psdf_mmms_rand_score_2_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_vi1_o2
```

```{r theme_plot_vi1_o2}
theme_plot_vi1_o2 <- base_plot_vi1_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(base::paste0("0-", score2_long))
```

```{r base_plot_vi2_o2}
base_plot_vi2_o2 <- make_density_plot(
  score_df = score_df_mark_score_2_o2,
  perm_score_df = psdf_mark_rand_score_2_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "Conserved markers"
)
base_plot_vi2_o2
```

```{r theme_plot_vi2_o2}
theme_plot_vi2_o2 <- base_plot_vi2_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

```{r base_plot_vi3_o2}
base_plot_vi3_o2 <- make_density_plot(
  score_df = score_df_sign_score_2_o2,
  perm_score_df = psdf_sign_rand_score_2_o2,
  color_cons = col_cons_long["conserved identity signature"],
  ylabel = "Conserved identity signature"
)
base_plot_vi3_o2
```

```{r theme_plot_vi3_o2}
theme_plot_vi3_o2 <- base_plot_vi3_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

#### MMMS vs SIGNRAND

```{r base_plot_vi1_o2_signrand}
base_plot_vi1_o2_signrand <- make_density_plot(
  score_df = score_df_mmms_score_2_o2,
  perm_score_df = psdf_mmms_signrand_score_2_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_vi1_o2_signrand
```

```{r theme_plot_vi1_o2_signrand}
theme_plot_vi1_o2_signrand <- base_plot_vi1_o2_signrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

### Mean prop (Score 3)

#### SIGN and MMMs vs RAND

```{r base_plot_mp1_o2}
base_plot_mp1_o2 <- make_density_plot(
  score_df = score_df_mmms_score_3_o2,
  perm_score_df = psdf_mmms_rand_score_3_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_mp1_o2
```

```{r theme_plot_mp1_o2}
theme_plot_mp1_o2 <- base_plot_mp1_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

```{r base_plot_mp2_o2}
base_plot_mp2_o2 <- make_density_plot(
  score_df = score_df_mark_score_3_o2,
  perm_score_df = psdf_mark_rand_score_3_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "Conserved markers"
)
base_plot_mp2_o2
```

```{r theme_plot_mp2_o2}
theme_plot_mp2_o2 <- base_plot_mp2_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

```{r base_plot_mp3_o2}
base_plot_mp3_o2 <- make_density_plot(
  score_df = score_df_sign_score_3_o2,
  perm_score_df = psdf_sign_rand_score_3_o2,
  color_cons = col_cons_long["conserved identity signature"],
  ylabel = "Conserved identity signature"
)
base_plot_mp3_o2
```

```{r theme_plot_mp3_o2}
theme_plot_mp3_o2 <- base_plot_mp3_o2+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

#### MMMS vs SIGNRAND

```{r base_plot_mp1_o2_signrand}
base_plot_mp1_o2_signrand <- make_density_plot(
  score_df = score_df_mmms_score_3_o2,
  perm_score_df = psdf_mmms_signrand_score_3_o2,
  color_cons = col_cons_long["BL6 markers"],
  ylabel = "BL6 markers"
)
base_plot_mp1_o2_signrand
```

```{r theme_plot_mp1_o2_signrand}
theme_plot_mp1_o2_signrand <- base_plot_mp1_o2_signrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

### Assemble

```{r plot_bp_o2}

plot_bp_o2 <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1, 1),
  align = "v",
  theme_plot_bp1_o2,
  theme_plot_bp2_o2,
  theme_plot_bp3_o2,
  theme_plot_bp1_o2_signrand
)

plot_ai_o2 <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1, 1),
  align = "v",
  theme_plot_ai1_o2,
  theme_plot_ai2_o2,
  theme_plot_ai3_o2,
  theme_plot_ai1_o2_signrand
)

plot_vi_o2 <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1, 1),
  align = "v",
  theme_plot_vi1_o2,
  theme_plot_vi2_o2,
  theme_plot_vi3_o2,
  theme_plot_vi1_o2_signrand
)

plot_mp_o2 <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1, 1),
  align = "v",
  theme_plot_mp1_o2,
  theme_plot_mp2_o2,
  theme_plot_mp3_o2,
  theme_plot_mp1_o2_signrand
)

```

```{r big_plot_o2, fig.width = 12, fig.height = 7}
big_plot_o2 <- ggpubr::ggarrange(
  ncol = 4,
  align = "hv",
  plot_bp_o2,
  plot_ai_o2,
  plot_vi_o2,
  plot_mp_o2, 
  widths = c(1, 1, 1, 1)
)
big_plot_o2
```

```{r output_weinreb_perm}
pdf(output_weinreb_perm, width = 12, height = 7)
big_plot_o2
dev.off()
```

## Plots MARK-SIGNRAND

### Boxplots

#### MMMS, MARK, and SIGN vs Rand

```{r base_plot_bp1_own_marksignrand}
base_plot_bp1_own_marksignrand <- ggplot2::ggplot(
  vis_df_mark_signrand_own,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill,
      pattern = type_fill,
      pattern_alpha = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp1_own_marksignrand
```

```{r theme_plot_bp1_own_marksignrand}
theme_plot_bp1_own_marksignrand <- base_plot_bp1_own_marksignrand+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size),
    axis.title = element_blank(),
    axis.text.y = element_blank())+
  ggplot2::ggtitle(nr_genes_title)
```

```{r base_plot_bp1_o1_marksignrand}
base_plot_bp1_o1_marksignrand <- ggplot2::ggplot(
  vis_df_mark_signrand_o1,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill,
      pattern = type_fill,
      pattern_alpha = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp1_o1_marksignrand
```

```{r theme_plot_bp1_o1_marksignrand}
theme_plot_bp1_o1_marksignrand <- base_plot_bp1_o1_marksignrand+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_blank())
```

```{r base_plot_bp1_o2_marksignrand}
base_plot_bp1_o2_marksignrand <- ggplot2::ggplot(
  vis_df_mark_signrand_o2,
  aes(y = type_y, 
      x =  number,
      fill = type_fill,
      color = type_fill,
      pattern = type_fill,
      pattern_alpha = type_fill))+
  geom_col()+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::scale_color_manual(values = col_cons_long)
base_plot_bp1_o2_marksignrand
```

```{r theme_plot_bp1_o2_marksignrand, fig.width = 4, fig.height = 2}
theme_plot_bp1_o2_marksignrand <- base_plot_bp1_o2_marksignrand+
  theme_all+
  ggplot2::scale_x_continuous(
    breaks=c(0,  400, 800),
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.y = element_blank())
```

### Adjusted Rand index (Score 1)

```{r base_plot_ai1_own_marksignrand}
base_plot_ai1_own_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_1,
  perm_score_df = psdf_mark_signrand_score_1,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_ai1_own_marksignrand
```

```{r theme_plot_ai1_own_marksignrand}
theme_plot_ai1_own_marksignrand <- base_plot_ai1_own_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

```{r base_plot_ai1_o1_marksignrand}
base_plot_ai1_o1_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_1_o1,
  perm_score_df = psdf_mark_signrand_score_1_o1,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_ai1_o1_marksignrand
```

```{r theme_plot_ai1_o1_marksignrand}
theme_plot_ai1_o1_marksignrand <- base_plot_ai1_o1_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

```{r base_plot_ai1_o2_marksignrand}
base_plot_ai1_o2_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_1_o2,
  perm_score_df = psdf_mark_signrand_score_1_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_ai1_o2_marksignrand
```

```{r theme_plot_ai1_o2_marksignrand}
theme_plot_ai1_o2_marksignrand <- base_plot_ai1_o2_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.5, 1),
    position = "top", 
    limits = c(score1_min, score1_max))+
  ggplot2::ggtitle(score1_long) 
```

### Variation of Information (Score 2)

```{r base_plot_vi1_own_marksignrand}
base_plot_vi1_own_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_2,
  perm_score_df = psdf_mark_signrand_score_2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_vi1_own_marksignrand
```

```{r theme_plot_vi1_own_marksignrand}
theme_plot_vi1_own_marksignrand <- base_plot_vi1_own_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

```{r base_plot_vi1_o1_marksignrand}
base_plot_vi1_o1_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_2_o1,
  perm_score_df = psdf_mark_signrand_score_2_o1,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_vi1_o1_marksignrand
```

```{r theme_plot_vi1_o1_marksignrand}
theme_plot_vi1_o1_marksignrand <- base_plot_vi1_o1_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

```{r base_plot_vi1_o2_marksignrand}
base_plot_vi1_o2_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_2_o2,
  perm_score_df = psdf_mark_signrand_score_2_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_vi1_o2_marksignrand
```

```{r theme_plot_vi1_o2_marksignrand}
theme_plot_vi1_o2_marksignrand <- base_plot_vi1_o2_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(-5, -2.5, 0),
    position = "top", 
    limits = c(score2_min, score2_max))+
  ggplot2::ggtitle(score2_long) 
```

### Mean Prop (Score 3)

```{r base_plot_mp1_own_marksignrand}
base_plot_mp1_own_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_3,
  perm_score_df = psdf_mark_signrand_score_3,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_mp1_own_marksignrand
```

```{r theme_plot_mp1_own_marksignrand}
theme_plot_mp1_own_marksignrand <- base_plot_mp1_own_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      hjust = 0.5,
      size = axis_title_size))+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

```{r base_plot_mp1_o1_marksignrand}
base_plot_mp1_o1_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_3_o1,
  perm_score_df = psdf_mark_signrand_score_3_o1,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_mp1_o1_marksignrand
```

```{r theme_plot_mp1_o1_marksignrand}
theme_plot_mp1_o1_marksignrand <- base_plot_mp1_o1_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

```{r base_plot_mp1_o2_marksignrand}
base_plot_mp1_o2_marksignrand <- make_density_plot(
  score_df = score_df_mark_score_3_o2,
  perm_score_df = psdf_mark_signrand_score_3_o2,
  color_cons = col_cons_long["conserved markers"],
  ylabel = "conserved markers"
)
base_plot_mp1_o2_marksignrand
```

```{r theme_plot_mp1_o2_marksignrand}
theme_plot_mp1_o2_marksignrand <- base_plot_mp1_o2_marksignrand+
  theme_all+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank(),
    plot.title = element_blank())+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.2, 0.4),
    position = "top", 
    limits = c(score3_min, score3_max))+
  ggplot2::ggtitle(score3_long) 
```

### Assemble

```{r plot_bp_mark}

plot_bp_mark <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1),
  align = "v",
  theme_plot_bp1_own_marksignrand,
  theme_plot_bp1_o1_marksignrand,
  theme_plot_bp1_o2_marksignrand
)

plot_ai_mark <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1),
  align = "v",
  theme_plot_ai1_own_marksignrand,
  theme_plot_ai1_o1_marksignrand,
  theme_plot_ai1_o2_marksignrand
)

plot_vi_mark <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1),
  align = "v",
  theme_plot_vi1_own_marksignrand,
  theme_plot_vi1_o1_marksignrand,
  theme_plot_vi1_o2_marksignrand
)

plot_mp_mark <- ggpubr::ggarrange(
  ncol = 1,
  heights = c(1.8, 1, 1),
  align = "v",
  theme_plot_mp1_own_marksignrand,
  theme_plot_mp1_o1_marksignrand,
  theme_plot_mp1_o2_marksignrand
)

```

```{r big_plot_mark, fig.width = 12, fig.height = 5.5}
big_plot_mark <- ggpubr::ggarrange(
  ncol = 4,
  align = "hv",
  plot_bp_mark,
  plot_ai_mark,
  plot_vi_mark,
  plot_mp_mark, 
  widths = c(1, 1, 1, 1)
)
big_plot_mark
```

```{r output_mark_signrand_perm}
pdf(output_mark_signrand_perm, width = 12, height = 5.5)
big_plot_mark
dev.off()
```

```{r}
utils::sessionInfo()
```

