---
title: "Figure 3"
date: '2024-09-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate Figure 3, current title:
"Mouse cell type signature genes alone are sufficient to identify cell types."

#### Load objects

```{r seed, message = FALSE}
base::RNGkind(kind = "Mersenne-Twister")
set.seed(37)
```

```{r load_packages, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(gghalves, quietly = TRUE) # not in conda env
library(ggpattern, quietly = TRUE) # not in conda env
```

```{r load_manual_basepath, eval = FALSE, include = FALSE}
base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data"
```

```{r load_manual, eval = FALSE, include = FALSE}
#-------------------------------------------------------------------------------
#### OWN DATASET (HSPCs)

# SCE with reclustering vectors for each gene set
sce_hsc_recl <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/03_recl/sce_hsc"))

# dataframe of re-clustering scores for each gene set
score_df_hsc <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/04_rcls/score_df_hsc"))

# dataframes of re-clustering scores after permutation comparing gene sets
psdf_hsc_mark_signrand <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/05_perg/perm_score_df_mark_hsc"))
psdf_hsc_mmms_signrand <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/05_perg/perm_score_df_mmms_hsc"))

# dataframes of re-clustering scores after background permutation
psdf_hsc_sign_rand <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/06_psig/perm_score_df_hsc"))
```

```{r load_manual_other1, eval = FALSE, include = FALSE}
#-------------------------------------------------------------------------------
#### OTHER MOUSE DATASET (HSPC)
# TODO: ADJUST TO DESIRED DATASET

# list of seurat objects with reclustering vectors for each gene set and resolution
seu_hsc_recl_list_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_mus_tm_bonemarrow_list"))

# list of dataframes of re-clustering scores for each gene set and resolution
score_df_hsc_list_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/04_rcls/score_df_mus_tm_bonemarrow_list"))

# dataframes of re-clustering scores after permutation comparing gene sets
psdf_hsc_mark_signrand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/05_perg/perm_score_df_mark_li_all_stromal"))
psdf_hsc_mmms_signrand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/05_perg/perm_score_df_mmms_li_all_stromal"))
psdf_hsc_mmms_markrand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/05_perg/perm_score_df_mmms_mark_li_all_stromal"))

# perm score DF from background permutation (SIGN)
psdf_hsc_sign_rand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/06_psig/perm_score_df_li_all_stromal"))

# perm score DF from background permutation (MARK)
psdf_hsc_mark_rand_o1 <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/07_pmrk/perm_score_df_li_all_stromal"))
```

```{r load_manual_remaining, eval = FALSE, include = FALSE}
#-------------------------------------------------------------------------------
#### REMAINING

# list of gene sets for HSPCs (conserved signature, conserved markers, etc.)
geneset_list_hsc <-base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_hsc"))

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  base_path,
  "/metadata/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")

colors_path <- base::paste0(
  base_path,
  "/metadata/colors/colors.txt")

source("../source/colors.R")
source("../source/plotting.R")
source("../source/sce_functions_reclustering.R")
```

```{r load_snakemake}

#-------------------------------------------------------------------------------
#### OWN DATASET (HSPCs)

# SCE with reclustering vectors for each gene set
sce_hsc_recl <- base::readRDS(snakemake@input[["sce_hsc_recl"]])
  
# dataframe of re-clustering scores for each gene set
score_df_hsc <- base::readRDS(snakemake@input[["score_df_hsc"]])

# dataframes of re-clustering scores after permutation comparing gene sets
psdf_hsc_mark_signrand <- base::readRDS(snakemake@input[["psdf_hsc_mark_signrand"]])
psdf_hsc_mmms_signrand <- base::readRDS(snakemake@input[["psdf_hsc_mmms_signrand"]])

# dataframe of re-clustering scores after background permutation
psdf_hsc_sign_rand <- base::readRDS(snakemake@input[["psdf_hsc_sign_rand"]])


#-------------------------------------------------------------------------------
#### OTHER (HSPC) MOUSE DATASET 1 

# list of seurat objects with reclustering vectors for each gene set
seu_hsc_recl_list_o1 <- base::readRDS(snakemake@input[["seu_hsc_recl_list_o1"]])

# list of dataframes of re-clustering scores for each gene set and resolution
score_df_hsc_list_o1 <- base::readRDS(snakemake@input[["score_df_hsc_list_o1"]])

# dataframes of re-clustering scores after permutation comparing gene sets
psdf_hsc_mark_signrand_o1 <- base::readRDS(snakemake@input[["psdf_hsc_mark_signrand_o1"]])
psdf_hsc_mmms_signrand_o1 <- base::readRDS(snakemake@input[["psdf_hsc_mmms_signrand_o1"]])
psdf_hsc_mmms_markrand_o1 <- base::readRDS(snakemake@input[["psdf_hsc_mmms_markrand_o1"]])

# dataframes of re-clustering scores after background permutation
psdf_hsc_sign_rand_o1 <- base::readRDS(snakemake@input[["psdf_hsc_sign_rand_o1"]])
psdf_hsc_mark_rand_o1 <- base::readRDS(snakemake@input[["psdf_hsc_mark_rand_o1"]])

#-------------------------------------------------------------------------------
#### REMAINING

# list of gene sets for HSPCs (conserved signature, conserved markers, etc.)
geneset_list_hsc <- base::readRDS(snakemake@input[["geneset_list_hsc"]])

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- snakemake@input[["resolution_df_path"]]

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")

# colors
colors_path <- snakemake@params[["colors_path"]]
colors_source <- snakemake@params[["colors"]]
source(colors_source)

```

```{r params}
# global params
source("determine_params.R")

col_cons <- c("conserved_signature" = "#db5b00ff",
              "conserved_markers" = "#eb9523ff",
              "mmusall_markers" = "#a7c5a3ff",
              "ndges" = "#ea91adff",
              "random_features" = "grey80")

col_cons_long <- c(
  "conserved identity signature" = "#db5b00ff",
  "conserved markers" = "#eb9523ff",
  "total BL6 markers" = "#a7c5a3ff",
  "ndges" = "#ea91adff",
  "random genes" = "grey80")

point_plot_size <- 4

pattern_fill_color <- "white"
pattern_line_colour <- "white"
pattern_density_val <- 0.4
pattern_spacing_val <- 0.1


color_hline_background <- "grey97"

```

```{r export_paths}
output_pdf_a_plot <- base::paste0(
  base_path, "/manuscript1/figures/figure3_a.pdf")
output_pdf_a_legend <- base::paste0(
  base_path, "/manuscript1/figures/figure3_a_legend.pdf")
output_pdf_b_plot <- base::paste0(
  base_path, "/manuscript1/figures/figure3_b.pdf")
output_pdf_b_legend <- base::paste0(
  base_path, "/manuscript1/figures/figure3_b_legend.pdf")
output_pdf_c_plot <- base::paste0(
  base_path, "/manuscript1/figures/figure3_c.pdf")
```

# A

Heatmaps of own re-clustered HSPC mouse data using signature genes/
conserved markers/BL6 markers/nDGEs.

#### Prep

```{r a_prep_conserved_signature}
mat_sign <- base::table(
  sce_hsc_recl$celltypes, 
  sce_hsc_recl$cluster_signt)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(mat_sign)){
  mat_sign[,i] <- mat_sign[,i]/base::sum(mat_sign[,i])
}
sign_df <- base::as.data.frame(mat_sign)

# manually factorise clusters for nice order
sign_df$Var2 <- factor(
  sign_df$Var2,
  levels = base::rev(c(1, 3, 2, 5, 8, 4, 6, 7)))
```

```{r a_prep_conserved_markers}
mat_mark <- base::table(
  sce_hsc_recl$celltypes, 
  sce_hsc_recl$cluster_consm)

for(i in 1:ncol(mat_mark)){
  mat_mark[,i] <- mat_mark[,i]/base::sum(mat_mark[,i])
}
mark_df <- base::as.data.frame(mat_mark)

mark_df$Var2 <- factor(
  mark_df$Var2,
  levels = base::rev(c(1, 3, 8, 10, 5, 2, 11, 4, 6, 7, 12, 9)))
```

```{r a_prep_all_bl6_markers}
mat_mmms <- base::table(
  sce_hsc_recl$celltypes, 
  sce_hsc_recl$cluster_mmusm)

for(i in 1:ncol(mat_mmms)){
  mat_mmms[,i] <- mat_mmms[,i]/base::sum(mat_mmms[,i])
}
mmms_df <- base::as.data.frame(mat_mmms)

mmms_df$Var2 <- factor(
  mmms_df$Var2,
  levels = base::rev(c(1, 3, 10, 11, 12, 5, 2, 9, 4, 6, 14, 13, 8, 7)))
```

#### Plot Components

```{r a_sign_base_plot}
a_sign_base_plot <- ggplot2::ggplot(
  sign_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r a_mark_base_plot}
a_mark_base_plot <- ggplot2::ggplot(
  mark_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r a_mmms_base_plot}
a_mmms_base_plot <- ggplot2::ggplot(
  mmms_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r a_sign_theme_plot}
a_sign_theme_plot <- a_sign_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      hjust = 0.5,
      color = col_cons_long["conserved identity signature"]))+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("Conserved identity signature")
```

```{r a_mark_theme_plot}
a_mark_theme_plot <- a_mark_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_text(
      size = axis_title_size,
      color = axis_title_color,
      face = axis_title_face),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      hjust = 0.5,
      color = col_cons_long["conserved markers"]))+
  ggplot2::ylab("New clusters")+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("Conserved markers")
```

```{r a_mmms_theme_plot}
a_mmms_theme_plot <- a_mmms_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_text(
      size = axis_title_size,
      color = axis_title_color,
      face = axis_title_face,
      margin = margin(t = 0.3, unit = "cm")),
    axis.text.x = element_text(
      size = axis_text_size,
      color = axis_text_color,
      face = axis_text_face,
      angle = 90,
      hjust = 0,
      vjust = 0.5),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      hjust = 0.5,
      color = col_cons_long["total BL6 markers"]))+
  ggplot2::xlab("Original cell types")+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("total BL6 markers")
```

```{r a_legend_theme_plot}
a_legend_plot <- a_mmms_base_plot+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::theme(
    legend.title = element_text(
      face = legend_title_face,
      size = legend_title_size,
      color = legend_title_color,
      margin = margin(r = 10),
      vjust = 0.85),
    legend.text = element_text(
      face = legend_text_face,
      size = legend_text_size,
      color = legend_text_color),
    legend.position = "bottom",
  )
a_legend_theme <- ggpubr::get_legend(a_legend_plot)
```

#### Construct

```{r a_construct, fig.height = 7, fig.width = 3.2}

a_theme_plot <- ggpubr::ggarrange(
  ncol = 1,
  a_sign_theme_plot,
  a_mark_theme_plot,
  a_mmms_theme_plot+
  ggplot2::theme(
    axis.text.x = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color, 
      angle = 90,
      hjust = 0,
      vjust = 0.5)),
  align = "v",
  heights = c(1, 1, 2.1)
)

a_theme_plot
```

```{r a_construct_legend, fig.width = 6, fig.height = 1}
a_legend <- ggpubr::ggarrange(a_legend_theme)
```

```{r a_plot_export}
pdf(output_pdf_a_plot, width = 3.2, height = 7)
a_theme_plot
dev.off()
```

```{r a_legend_export}
pdf(output_pdf_a_legend, width = 6, height = 1)
a_legend
dev.off()
```

## B

(Re-) clustering metrics to compare re-clustering of own dataset with signature
genes, conserved markers or all BL6 markers, permuted with different sets of
random genes as indicated (signature + random_conserved_markers,
signature + conserved markers + random_bl6_markers). 

#### Prep

##### perm plots

Separate score DFs per conservation level and type of score.

```{r b_prep1}
score_df_sign <- score_df_hsc[
  score_df_hsc$conservation_level == "conserved_signature",]
# add nicer looking conservation level annotation
score_df_sign$conservation_level_long <- "conserved identity signature"

score_df_sign_score_1 <- score_df_sign[
  score_df_sign$type == "adjusted_rand_index",]
score_df_sign_score_2 <- score_df_sign[
  score_df_sign$type == "variation_information",]
score_df_sign_score_3 <- score_df_sign[
  score_df_sign$type == "mean_prop_cells_cluster",]
```

```{r b_prep2}
score_df_mark <- score_df_hsc[
  score_df_hsc$conservation_level == "conserved_markers",]
score_df_mark$conservation_level_long <- "conserved markers"

score_df_mark_score_1 <- score_df_mark[
  score_df_mark$type == "adjusted_rand_index",]
score_df_mark_score_2 <- score_df_mark[
  score_df_mark$type == "variation_information",]
score_df_mark_score_3 <- score_df_mark[
  score_df_mark$type == "mean_prop_cells_cluster",]
```

```{r b_prep3}
score_df_mmms <- score_df_hsc[
  score_df_hsc$conservation_level == "mmusall_markers",]
score_df_mmms$conservation_level_long <- "total BL6 markers"

score_df_mmms_score_1 <- score_df_mmms[
  score_df_mmms$type == "adjusted_rand_index",]
score_df_mmms_score_2 <- score_df_mmms[
  score_df_mmms$type == "variation_information",]
score_df_mmms_score_3 <- score_df_mmms[
  score_df_mmms$type == "mean_prop_cells_cluster",]
```

Also separate the permutation score dfs by conservation level and type of score.

```{r b_prep4}
psdf_mark_signrand_long <- pivot_longer(
  psdf_hsc_mark_signrand, 
  c = 1:ncol(psdf_hsc_mark_signrand),
  values_to = "value",
  names_to = "type")

psdf_mark_signrand_long$conservation_level_long = base::rep(
  "conserved markers", nrow(psdf_mark_signrand_long))

psdf_mark_signrand_score_1 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type %in% c("adjusted_rand_index"),]
psdf_mark_signrand_score_2 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type %in% c("variation_information"),]
psdf_mark_signrand_score_3 <- psdf_mark_signrand_long[
  psdf_mark_signrand_long$type %in% c("mean_prop_cells_cluster"),]
```

```{r b_prep5}
psdf_mmms_signrand_long <- pivot_longer(
  psdf_hsc_mmms_signrand, 
  c = 1:ncol(psdf_hsc_mmms_signrand),
  values_to = "value",
  names_to = "type")

psdf_mmms_signrand_long$conservation_level_long = base::rep(
  "total BL6 markers", nrow(psdf_mmms_signrand_long))

psdf_mmms_signrand_score_1 <- psdf_mmms_signrand_long[
  psdf_mmms_signrand_long$type %in% c("adjusted_rand_index"),]
psdf_mmms_signrand_score_2 <- psdf_mmms_signrand_long[
  psdf_mmms_signrand_long$type %in% c("variation_information"),]
psdf_mmms_signrand_score_3 <- psdf_mmms_signrand_long[
  psdf_mmms_signrand_long$type %in% c("mean_prop_cells_cluster"),]
```

get the maximum number of variation of information scores for plot limits

```{r max_vi}
max_vi <- base::max(c(
  
  # actual scores
  score_df_hsc$value[score_df_hsc$type == "variation_information"],
  
  # permuted background scores for conserved signature
  psdf_hsc_sign_rand$variation_information,
  
  # pemruted gene set scores
  psdf_hsc_mmms_signrand$variation_information,
  psdf_hsc_mark_signrand$variation_information)) + 0.5
```

##### box plots

```{r b_prep6}
# get the number of genes used for each gene set
nr_sign_used <- sce_hsc_recl$cluster_signt_genes_used[1]
nr_mark_used <- sce_hsc_recl$cluster_consm_genes_used[1]
nr_mmms_used <- sce_hsc_recl$cluster_mmusm_genes_used[1]

nr_mark_left <- nr_mark_used - nr_sign_used
nr_mark_rand <- nr_mark_left # same number but keep for tracking purpose

nr_mmms_left <- nr_mmms_used - nr_sign_used
nr_mmms_rand <- nr_mmms_left

max_genes <- nr_mmms_used
```

make dfs for visualisation of gene numbers with boxplots.

```{r b_prep7}
vis_df_sign <- base::data.frame(
  type_y = "conserved identity signature",
  type_fill = "conserved identity signature",
  number = nr_sign_used
)
```

```{r b_prep8}
vis_df_mark <- base::data.frame(
  type_y = c(
    "conserved markers",
    "conserved markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "conserved markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used,
    nr_mark_left,
    nr_sign_used,
    nr_mark_rand)
)

# factor for plot
vis_df_mark$type_fill <- factor(vis_df_mark$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "conserved markers",
  "random genes"
)))
vis_df_mark$type_y <- factor(vis_df_mark$type_y, levels = base::rev(c(
  "conserved markers",
  "permutation gene set"
)))
```

```{r b_prep9}
vis_df_mmms <- base::data.frame(
  type_y = c(
    "total BL6 markers",
    "total BL6 markers",
    "permutation gene set",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "total BL6 markers",
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used,
    nr_mmms_left,
    nr_sign_used,
    nr_mmms_rand)
)

vis_df_mmms$type_fill <- factor(vis_df_mmms$type_fill, levels = base::rev(c(
  "conserved identity signature",
  "total BL6 markers",
  "random genes"
)))
vis_df_mmms$type_y <- factor(vis_df_mmms$type_y, levels = base::rev(c(
  "total BL6 markers",
  "permutation gene set"
)))
```

#### Plot Components

##### part left, score1: Adjusted Rand index

```{r b_base_plot_left1}
b_base_plot_left1 <- ggplot2::ggplot(
  score_df_sign_score_1,
  aes(y = conservation_level_long,
      x = value,
      color = conservation_level_long))+
  ggplot2::geom_vline(
    xintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_vline(
    xintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size)
```

```{r b_theme_plot_left1}
b_theme_plot_left1 <- b_base_plot_left1+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color, 
      angle = 0,
      hjust = 0.5,
      vjust = 0.5),
    axis.text.y = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color,
      vjust = 0.5),
    axis.title = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color, 
      hjust = 0.5))+
  ggplot2::scale_x_continuous(
    breaks=c(0, 0.25, 0.5, 0.75, 1),
    position = "top", 
    limits = c(0, 1))+
  ggplot2::ggtitle("Adjusted rand index")
```

```{r b_base_plot_left2}
b_base_plot_left2 <- ggplot2::ggplot(
  score_df_mark_score_1,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_1,
    color = col_cons_long["conserved markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mark_signrand_score_1,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_1,
    color = col_cons_long["conserved markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r b_theme_plot_left2}
b_theme_plot_left2 <- b_base_plot_left2+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color,
      vjust = 0.5),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_y_continuous( # because flipped
    limits = c(0, 1))+
  ggplot2::scale_x_discrete(expand = expansion(add = c(-0.8, -0.1)))
```

```{r b_base_plot_left3}
b_base_plot_left3 <- ggplot2::ggplot(
  score_df_mmms_score_1,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_1,
    color = col_cons_long["total BL6 markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mmms_signrand_score_1,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_1,
    color = col_cons_long["total BL6 markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r b_theme_plot_left3}
b_theme_plot_left3 <- b_base_plot_left3+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color,
      vjust = 0.5),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_y_continuous(
    limits = c(0, 1))+
  ggplot2::scale_x_discrete(expand = expansion(add = c(-0.8, -0.1)))
```

```{r b_construct_a, fig.height = 2.8, fig.width = 5}
b_part_left <- ggpubr::ggarrange(
  ncol = 1, 
  b_theme_plot_left1,
  b_theme_plot_left2,
  b_theme_plot_left3,
  heights = c(1.05, 1, 1),
  align = "v"
)
b_part_left
```

##### part b middle - left score2: Variation of Information

```{r b_base_plot_mleft1}
b_base_plot_mleft1 <- ggplot2::ggplot(
  score_df_sign_score_2,
  aes(y = conservation_level_long,
      x = value,
      color = conservation_level_long))+
  ggplot2::geom_vline(
    xintercept = 1.5,
      color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_vline(
    xintercept = 4.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size)
```

```{r b_theme_plot_mleft1}
b_theme_plot_mleft1 <- b_base_plot_mleft1+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color, 
      angle = 0,
      hjust = 0.5,
      vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color, 
      hjust = 0.5))+
  ggplot2::scale_x_continuous(
    position = "top",
    limit = c(0, max_vi),
    breaks=c(0, 1, 2, 3, 4, 5))+
  ggplot2::ggtitle("Variation of information")
```

```{r b_base_plot_mleft2}
b_base_plot_mleft2 <- ggplot2::ggplot(
  score_df_mark_score_2,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 1.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 4.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_2,
    color = col_cons_long["conserved markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mark_signrand_score_2,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_2,
    color = col_cons_long["conserved markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
   ggplot2::coord_flip()
```

```{r b_theme_plot_mleft2}
b_theme_plot_mleft2 <- b_base_plot_mleft2+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.8, -0.1)))+
  ggplot2::scale_y_continuous(
   limit = c(0, max_vi),
   breaks=c(0, 1, 2, 3, 4, 5))
```

```{r b_base_plot_mleft3}
b_base_plot_mleft3 <- ggplot2::ggplot(
  score_df_mmms_score_2,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 1.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 4.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_2,
    color = col_cons_long["total BL6 markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mmms_signrand_score_2,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_2,
    color = col_cons_long["total BL6 markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r b_theme_plot_mleft3}
b_theme_plot_mleft3 <- b_base_plot_mleft3+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_y_continuous(
    limits = c(0, max_vi))+
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.8, -0.1)))+
 ggplot2::scale_y_continuous(
   limit = c(0, max_vi),
   breaks=c(0, 1, 2, 3, 4, 5))
```

```{r b_construct_mleft, fig.height = 2.8, fig.width = 2.8}
b_part_mleft <- ggpubr::ggarrange(
  ncol = 1, 
  b_theme_plot_mleft1,
  b_theme_plot_mleft2,
  b_theme_plot_mleft3,
  heights = c(1.05, 1, 1),
  align = "v"
)
b_part_mleft
```

##### part c (right, score3) mean prop cells

```{r b_base_plot_mright1}
b_base_plot_mright1 <- ggplot2::ggplot(
  score_df_sign_score_3,
  aes(y = conservation_level_long,
      x = value,
      color = conservation_level_long))+
  ggplot2::geom_vline(
    xintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_vline(
    xintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size)
```

```{r b_theme_plot_mright1}
b_theme_plot_mright1 <- b_base_plot_mright1+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color, 
      angle = 0,
      hjust = 0.5,
      vjust = 0.5),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color, 
      hjust = 0.5))+
  ggplot2::scale_x_continuous(
    position = "top", 
    limits = c(0, 1))+
  ggplot2::ggtitle("Mean prop. cells/cluster")
```

```{r b_base_plot_mright2}
b_base_plot_mright2 <- ggplot2::ggplot(
  score_df_mark_score_3,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_3,
    color = col_cons_long["conserved markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mark_signrand_score_3,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mark_signrand_score_3,
    color = col_cons_long["conserved markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r b_theme_plot_mright2}
b_theme_plot_mright2 <- b_base_plot_mright2+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_y_continuous(
    limits = c(0, 1))+
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.8, -0.1)))
```

```{r b_base_plot_mright3}
b_base_plot_mright3 <- ggplot2::ggplot(
  score_df_mmms_score_3,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.25, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_3,
    color = col_cons_long["total BL6 markers"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_mmms_signrand_score_3,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_mmms_signrand_score_3,
    color = col_cons_long["total BL6 markers"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r b_theme_plot_mright3}
b_theme_plot_mright3 <- b_base_plot_mright3+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title = element_blank())+
  ggplot2::scale_y_continuous(
    limits = c(0, 1))+
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.8, -0.1)))
```

```{r b_construct_c, fig.height = 2.8, fig.width = 2.8}
b_part_mright <- ggpubr::ggarrange(
  ncol = 1, 
  b_theme_plot_mright1,
  b_theme_plot_mright2,
  b_theme_plot_mright3,
  heights = c(1.05, 1, 1),
  align = "v"
)
b_part_mright
```

##### part right: boxplots

```{r b_base_plot_right1}
b_base_plot_right1 <- ggplot2::ggplot(
  vis_df_sign,
  aes(y = type_y,
      x = number, 
      fill = type_fill))+
  ggplot2::geom_col(position = position_stack())
```

```{r b_theme_plot_right1}
b_theme_plot_right1 <- b_base_plot_right1+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(
      size = axis_text_size,
      face = axis_text_face,
      color = axis_text_color),
    axis.title.x = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color,
      hjust = 0.5))+
  ggplot2::scale_x_continuous(
    position = "top", 
    limits = c(0, max_genes))+
  ggplot2::ggtitle("# re-clustering genes")+
  ggplot2::scale_fill_manual(values = col_cons_long)
```

```{r b_base_plot_right2}
b_base_plot_right2 <- ggplot2::ggplot(
  vis_df_mark,
  aes(y = type_y, 
      x=  number,
      fill = type_fill))+
  ggplot2::geom_col(position = position_stack())+
  ggpattern::geom_bar_pattern(
    data = base::data.frame(
      "type_y" = base::rep("permutation gene set", nr_mark_used)),
    inherit.aes = FALSE,
    mapping = aes(y = type_y),
    position = position_dodge(preserve = "single"),
    pattern = "circle",
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)
```

```{r b_theme_plot_right2}
b_theme_plot_right2 <- b_base_plot_right2+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank())+
  ggplot2::scale_x_continuous(
    limits = c(0, max_genes))+
  ggplot2::xlab("Number of genes")+
  ggplot2::scale_fill_manual(values = col_cons_long)
```

```{r b_base_plot_right3}
b_base_plot_right3 <- ggplot2::ggplot(
  vis_df_mmms,
  aes(y = type_y,
      x = number, 
      fill = type_fill))+
  ggplot2::geom_col(position = position_stack())+
  ggpattern::geom_bar_pattern(
    data = base::data.frame(
      "type_y" = base::rep("permutation gene set", nr_mmms_used)),
    inherit.aes = FALSE,
    mapping = aes(y = type_y),
    pattern = "circle",
    position = position_dodge(preserve = "single"),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)
```

```{r b_theme_plot_right3}
b_theme_plot_right3 <- b_base_plot_right3+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank())+
  ggplot2::scale_x_continuous(
    limits = c(0, max_genes))+
  ggplot2::xlab("Number of genes")+
  ggplot2::scale_fill_manual(values = col_cons_long)
```

```{r b_legend_plot}

vis_df_legend <- base::rbind(vis_df_mmms, vis_df_mark)
vis_df_legend$type_fill <- factor(
  vis_df_legend$type_fill,
  levels = c("conserved identity signature",
             "conserved markers",
             "total BL6 markers",
             "random genes")
)

b_legend_right_plot <- ggplot2::ggplot(
  vis_df_legend,
    aes(y = type_y,
      x = number, 
      fill = type_fill))+
  ggplot2::geom_col(position = position_stack())+
  ggpattern::geom_bar_pattern(
    data = base::data.frame(
      "type_y" = base::rep("permutation gene set", nr_mmms_used)),
    inherit.aes = FALSE,
    mapping = aes(y = type_y),
    pattern = "circle",
    position = position_dodge(preserve = "single"),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  ggplot2::theme(
    legend.title = element_text(
      size = legend_title_size,
      face = legend_title_face,
      color = legend_title_color),
    legend.text = element_text(
      size = legend_text_size,
      face = legend_text_face,
      color = legend_text_color))+
  ggplot2::scale_fill_manual(
    "re-clustering gene sets",
    values = col_cons_long)

b_legend_right <- ggpubr::get_legend(b_legend_right_plot)
```

```{r b_legend, fig.width = 3, fig.height = 1.8}
b_legend <- ggpubr::ggarrange(b_legend_right)
b_legend
```

```{r b_construct_right, fig.height = 2.8, fig.width = 2.8}
b_part_right <- ggpubr::ggarrange(
  ncol = 1, 
  b_theme_plot_right1, # barplot width depends on plot size
  b_theme_plot_right2,
  b_theme_plot_right3,
  heights = c(1.05, 1, 1),
  align = "v"
)
b_part_right
```

```{r b_construct_all, fig.height = 3.5, fig.width = 13}

# the ratio between upper panel and lower two panels depends on the
# height of the entire plot

b_all_plot <- ggpubr::ggarrange(
  ncol = 4,
  b_part_left,
  b_part_mleft,
  b_part_mright,
  b_part_right,
  widths = c(3.8, 2, 2, 1.5),
  align = "v"
)
b_all_plot
```

```{r b_plot_export}
pdf(output_pdf_b_plot, width = 12.7, height = 3.7)
b_all_plot
dev.off()
```

```{r b_legend_export}
pdf(output_pdf_b_legend, width = 3, height = 1.8)
b_legend
dev.off()
```

## C

#### Prep

Can re-use conserved signature DFs generated for part B.
Make DFs for conserved signature permuted scores.

```{r c_prep1}
psdf_sign_rand_long <- pivot_longer(
  psdf_hsc_sign_rand, 
  c = 1:ncol(psdf_hsc_sign_rand),
  values_to = "value",
  names_to = "type")

psdf_sign_rand_long$conservation_level_long <- base::rep(
  "conserved identity signature", nrow(psdf_sign_rand_long))

psdf_sign_rand_score_1 <- psdf_sign_rand_long[psdf_sign_rand_long$type %in% 
                                               c("adjusted_rand_index"),]
psdf_sign_rand_score_2 <- psdf_sign_rand_long[psdf_sign_rand_long$type %in% 
                                               c("variation_information"),]
psdf_sign_rand_score_3 <- psdf_sign_rand_long[psdf_sign_rand_long$type %in% 
                                               c("mean_prop_cells_cluster"),]
```

Make DF with gene numbers also for conserved signature background permutation.

```{r c_prep2}
vis_df_sign_rand <- base::data.frame(
  type_y = c(
    "conserved identity signature",
    "permutation gene set"),
  type_fill = c(
    "conserved identity signature",
    "random genes"),
  number = c(
    nr_sign_used,
    nr_sign_used))

vis_df_sign_rand$type_fill <- factor(
  vis_df_sign_rand$type_fill,
  levels = base::rev(c(
    "conserved identity signature",
    "random genes")))

vis_df_sign_rand$type_y <- factor(
  vis_df_sign_rand$type_y, 
  levels = base::rev(c(
    "conserved identity signature",
    "permutation gene set")))
```

##### part left: adjusted rand index

```{r c_base_plot_left}
c_base_plot_left <- ggplot2::ggplot(
  score_df_sign_score_1,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.5, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_1,
    color = col_cons_long["conserved identity signature"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_sign_rand_score_1,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_1,
    color = col_cons_long["conserved identity signature"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r c_theme_plot_left}
c_theme_plot_left <- c_base_plot_left+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = element_text(
      face = axis_text_face,
      color = axis_text_color,
      size = axis_text_size),
    axis.text.y = element_text(
      face = axis_text_face,
      color = axis_text_color,
      size = axis_text_size),
    axis.title = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color,
      hjust = 0.5))+
  ggplot2::scale_y_continuous(
    position = "right", limits = c(0,1))+ # because flipped
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.9)))+
  ggplot2::ggtitle("Adjusted rand index")
```

##### part left middle: variation information

```{r c_base_plot_mleft}
c_base_plot_mleft <- ggplot2::ggplot(
  score_df_sign_score_2,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 1.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 4.5,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.5, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_2,
    color = col_cons_long["conserved identity signature"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggpattern::geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_sign_rand_score_2,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_2,
    color = col_cons_long["conserved identity signature"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r c_theme_plot_mleft}
c_theme_plot_mleft <- c_base_plot_mleft+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.text.x = element_text(
      face = axis_text_face,
      color = axis_text_color,
      size = axis_text_size),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color,
      hjust = 0.5))+
  ggplot2::scale_y_continuous(
    position = "right", 
    breaks=c(0, 1, 2, 3, 4, 5),
    limits = c(0, max_vi))+
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-0.9)))+
  ggplot2::ggtitle("Variation of information")
```

#### part middle right: prop cells/cluster

```{r c_base_plot_mright}
c_base_plot_mright <- ggplot2::ggplot(
  score_df_sign_score_3,
  aes(x = conservation_level_long,
      y = value,
      fill = conservation_level_long,
      color = conservation_level_long))+
  ggplot2::geom_hline(
    yintercept = 0.25,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_hline(
    yintercept = 0.75,
    color = color_hline_background,
    linewidth = 3)+
  ggplot2::geom_point(
    size = point_plot_size,
    position = position_nudge(x = 0.5, y = 0))+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_3,
    color = col_cons_long["conserved identity signature"],
    alpha = 1,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  geom_violin_pattern(
    pattern = "circle",
    inherit.aes = FALSE,
    scale = "width",
    trim = FALSE,
    position = position_nudge(x = -0.5, y = 0),
    data = psdf_sign_rand_score_3,
    mapping = aes(y = value, x = conservation_level_long),
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)+
  gghalves::geom_half_violin(
    data = psdf_sign_rand_score_3,
    color = col_cons_long["conserved identity signature"],
    alpha = 0,
    side = "r",
    scale = "width",
    position = position_nudge(x = -0.5, y = 0))+
  ggplot2::coord_flip()
```

```{r c_theme_plot_mright}
c_theme_plot_mright <- c_base_plot_mright+
  ggplot2::theme_classic()+
  ggplot2::scale_color_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::scale_fill_manual(
    "Score",
    values = col_cons_long)+
  ggplot2::theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.text.x = element_text(
      face = axis_text_face,
      color = axis_text_color,
      size = axis_text_size),
    axis.title = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color,
      hjust = 0.5))+
  ggplot2::scale_y_continuous(
    position = "right", 
    limits = c(0, 1))+ 
  ggplot2::scale_x_discrete(
    expand = expansion(add = c(-1)))+
  ggplot2::ggtitle("Mean prop. cells/cluster")
```

#### part right: barplots

```{r c_base_plot_right}
c_base_plot_right <- ggplot2::ggplot(
  vis_df_sign_rand,
  aes(y = type_y, 
        x = number, 
      fill = type_fill))+
  ggplot2::geom_col(position = position_stack())+
  ggpattern::geom_bar_pattern(
    data = base::data.frame(
      "type_y" = base::rep("permutation gene set", nr_sign_used)),
    inherit.aes = FALSE,
    mapping = aes(y = type_y),
    position = position_dodge(preserve = "single"),
    pattern = "circle",
    color = NA, 
    alpha = 0,
    pattern_fill = pattern_fill_color,
    pattern_colour = pattern_line_colour,
    pattern_density =pattern_density_val,
    pattern_spacing = pattern_spacing_val)
```

```{r c_theme_plot_right}
c_theme_plot_right <- c_base_plot_right+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    legend.position = "none",
    axis.text.x = element_text(
      size = axis_text_size,
      face = axis_text_face,
      color = axis_text_color,
      hjust = 0.5),
    axis.title.x = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = plot_title_color,
      hjust = 0.5))+
  ggplot2::scale_x_continuous(
    limits = c(0, max_genes),
    position = "top")+
  ggplot2::scale_fill_manual(values = col_cons_long)+
  ggplot2::ggtitle("# re-clustering genes")
```

#### Assemble

```{r c_assemble_plot, fig.width = 12.7, fig.height = 2}

c_theme_plot <- ggpubr::ggarrange(
  c_theme_plot_left,
  c_theme_plot_mleft,
  c_theme_plot_mright,
  c_theme_plot_right,
  nrow = 1,
  widths = c(3.8, 2, 2, 1.5)
)
c_theme_plot
```

## D 

#### Prep

```{r d_prep1}

selected_seu_list <- lapply(seu_hsc_recl_list, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df[resolution_df$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
  
})
selected_seu_list

```

```{r d_prep_conserved_signature}
d_mat_sign <- base::table(selected_seu_list$seu_sign$cell_type, 
                          selected_seu_list$seu_sign$seurat_clusters)

# normalise per column = new clusters and turn into DF
for(i in 1:ncol(d_mat_sign)){
  d_mat_sign[,i] <- d_mat_sign[,i]/base::sum(d_mat_sign[,i])
}
d_sign_df <- base::as.data.frame(d_mat_sign)

# manually factorise clusters
# d_sign_df$Var2 <- factor(
#   d_sign_df$Var2,
#   levels = base::rev(c(1, 3, 2, 5, 8, 4, 6, 7)))
```

```{r a_prep_conserved_markers}
d_mat_mark <- base::table(selected_seu_list$seu_mark$cell_type, 
                          selected_seu_list$seu_mark$seurat_clusters)

for(i in 1:ncol(d_mat_mark)){
  d_mat_mark[,i] <- d_mat_mark[,i]/base::sum(d_mat_mark[,i])
  }
d_mark_df <- base::as.data.frame(d_mat_mark)

# d_mark_df$Var2 <- factor(
#   d_mark_df$Var2,
#   levels = base::rev(c(1, 3, 8, 10, 5, 2, 11, 4, 6, 7, 12, 9)))
```

```{r a_prep_all_bl6_markers}
d_mat_mmms <- base::table(selected_seu_list$seu_mmms$cell_type, 
                          selected_seu_list$seu_mmms$seurat_clusters)

for(i in 1:ncol(d_mat_mmms)){
  d_mat_mmms[,i] <- d_mat_mmms[,i]/base::sum(d_mat_mmms[,i])
}
d_mmms_df <- base::as.data.frame(d_mat_mmms)

# mmms_df$Var2 <- factor(
#   mmms_df$Var2,
#   levels = base::rev(c(1, 3, 10, 11, 12, 5, 2, 9, 4, 6, 14, 13, 8, 7)))
```

#### Plot Components

```{r d_sign_base_plot}
d_sign_base_plot <- ggplot2::ggplot(
  d_sign_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r d_mark_base_plot}
d_mark_base_plot <- ggplot2::ggplot(
  d_mark_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r d_mmms_base_plot}
d_mmms_base_plot <- ggplot2::ggplot(
  d_mmms_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()
```

```{r d_sign_theme_plot}
d_sign_theme_plot <- d_sign_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = col_cons_long["conserved identity signature"]
    )
  )+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("Conserved identity signature")
```

```{r d_mark_theme_plot}
d_mark_theme_plot <- d_mark_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = col_cons_long["conserved markers"]
    )
  )+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("Conserved markers")
```

```{r d_mmms_theme_plot}
d_mmms_theme_plot <- d_mmms_base_plot+
  ggplot2::theme_classic()+
  ggplot2::theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    plot.title = element_text(
      size = plot_title_size,
      face = plot_title_face,
      color = col_cons_long["total BL6 markers"]))+
  ggplot2::scale_fill_gradientn(
    "Proportion of cells per cluster",
    colors = mycolors_to1,
    limits = c(0, 1),
    breaks = c(0, 0.5, 1))+
  ggplot2::ggtitle("total BL6 markers")
```

#### Construct

```{r d_construct, fig.height = 8, fig.width = 3}

d_theme_plot <- ggpubr::ggarrange(
  ncol = 1,
  d_sign_theme_plot,
  d_mark_theme_plot,
  d_mmms_theme_plot+
  ggplot2::theme(
    axis.text.x = element_text(
      face = axis_text_face,
      size = axis_text_size,
      color = axis_text_color, 
      angle = 90,
      hjust = 0,
      vjust = 0.5)),
  align = "v",
  heights = c(1, 1.5, 2.4)
)

d_theme_plot
```

## E

#### Prep

##### perm plots

```{r}
selected_score_df_list <- lapply(score_df_hsc_list_other1, function(score_df_list){
  
  conslev_curr <- score_df_list[[1]]$conservation_level[1]
  
  resdf_temp <- resolution_df[resolution_df$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  score_df_return <- score_df_list[[res_curr]]
  return(score_df_return)
  
})
selected_score_df_list
score_df_hsc_other1 <- bind_rows(selected_score_df_list)
```

```{r e_prep1}
score_df_sign_other1 <- score_df_hsc_other1[score_df_hsc_other1$conservation_level == "conserved_signature",]
score_df_sign_score_1_other1 <- score_df_sign_other1[score_df_sign_other1$type == "adjusted_rand_index",]
score_df_sign_score_2_other1 <- score_df_sign_other1[score_df_sign_other1$type == "variation_information",]
score_df_sign_score_3_other1 <- score_df_sign_other1[score_df_sign_other1$type == "mean_prop_cells_cluster",]

score_df_sign_score_1$conservation_level <- "conserved identity signature"
score_df_sign_score_2$conservation_level <- "conserved identity signature"
score_df_sign_score_3$conservation_level <- "conserved identity signature"
```

```{r e_prep2}
score_df_mark_other1 <- score_df_hsc_other1[score_df_hsc_other1$conservation_level == "conserved_markers",]
score_df_mark_score_1_other1 <- score_df_mark_other1[score_df_mark_other1$type == "adjusted_rand_index",]
score_df_mark_score_2_other1 <- score_df_mark_other1[score_df_mark_other1$type == "variation_information",]
score_df_mark_score_3_other1 <- score_df_mark_other1[score_df_mark_other1$type == "mean_prop_cells_cluster",]

score_df_mark_score_1_other1$conservation_level <- "conserved markers"
score_df_mark_score_2_other1$conservation_level <- "conserved markers"
score_df_mark_score_3_other1$conservation_level <- "conserved markers"
```

!# TODO: CONTINUE HERE

```{r e_prep3}
psdf_mark_signrand_long_other1 <- pivot_longer(
  psdf_hsc_mark_signrand_other1, 
  c = 1:ncol(psdf_hsc_mark_signrand_other1),
  values_to = "value",
  names_to = "type")

psdf_mark_signrand_score_1_other1 <- psdf_mark_signrand_long_other1[psdf_mark_signrand_long_other1$type %in% c("adjusted_rand_index"),]
psdf_mark_signrand_score_2_other1 <- psdf_mark_signrand_long_other1[psdf_mark_signrand_long_other1$type %in% c("variation_information"),]
psdf_mark_signrand_score_3_other1 <- psdf_mark_signrand_long_other1[psdf_mark_signrand_long_other1$type %in% c("mean_prop_cells_cluster"),]

psdf_mark_signrand_score_1_other1$conservation_level = base::rep("conserved markers", nrow(psdf_mark_signrand_score_1_other1))
psdf_mark_signrand_score_2_other1$conservation_level = base::rep("conserved markers", nrow(psdf_mark_signrand_score_2_other1))
psdf_mark_signrand_score_3_other1$conservation_level = base::rep("conserved markers", nrow(psdf_mark_signrand_score_3_other1))
```

```{r e_prep4}
score_df_mmms_other_1  <- score_df_hsc_other_1[score_df_hsc_other_1 $conservation_level == "mmusall_markers",]
score_df_mmms_score_1_other_1 <- score_df_mmms_other_1[score_df_mmms_other_1$type == "adjusted_rand_index",]
score_df_mmms_score_2_other_1 <- score_df_mmms_other_1[score_df_mmms_other_1$type == "variation_information",]
score_df_mmms_score_3_other_1 <- score_df_mmms_other_1[score_df_mmms_other_1$type == "mean_prop_cells_cluster",]

score_df_mmms_score_1_other_1$conservation_level <- "total BL6 markers"
score_df_mmms_score_2_other_1$conservation_level <- "total BL6 markers"
score_df_mmms_score_3_other_1$conservation_level <- "total BL6 markers"
```

```{r e_prep5}
psdf_mmms_signrand_long_other_1 <- pivot_longer(
  psdf_hsc_mmms_signrand_other1, 
  c = 1:ncol(psdf_hsc_mmms_signrand_other1),
  values_to = "value",
  names_to = "type")

psdf_mmms_signrand_score_1_other_1 <- psdf_mmms_signrand_long_other_1[psdf_mmms_signrand_long_other_1$type %in% c("adjusted_rand_index"),]
psdf_mmms_signrand_score_2_other_1 <- psdf_mmms_signrand_long_other_1[psdf_mmms_signrand_long_other_1$type %in% c("variation_information"),]
psdf_mmms_signrand_score_3_other_1 <- psdf_mmms_signrand_long_other_1[psdf_mmms_signrand_long_other_1$type %in% c("mean_prop_cells_cluster"),]

psdf_mmms_signrand_score_1_other_1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_signrand_score_1_other_1))
psdf_mmms_signrand_score_2_other_1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_signrand_score_2_other_1))
psdf_mmms_signrand_score_3_other_1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_signrand_score_3_other_1))
```

```{r e_prep7}
psdf_mmms_markrand_long_other_1 <- pivot_longer(
  psdf_hsc_mmms_markrand_other1, 
  c = 1:ncol(psdf_hsc_mmms_markrand_other1),
  values_to = "value",
  names_to = "type")

psdf_mmms_markrand_score_1_other1 <- psdf_mmms_markrand_long_other_1[psdf_mmms_markrand_long_other_1$type %in% c("adjusted_rand_index"),]
psdf_mmms_markrand_score_2_other1 <- psdf_mmms_markrand_long_other_1[psdf_mmms_markrand_long_other_1$type %in% c("variation_information"),]
psdf_mmms_markrand_score_3_other1 <- psdf_mmms_markrand_long_other_1[psdf_mmms_markrand_long_other_1$type %in% c("mean_prop_cells_cluster"),]

psdf_mmms_markrand_score_1_other1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_markrand_score_1_other1))
psdf_mmms_markrand_score_2_other1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_markrand_score_2_other1))
psdf_mmms_markrand_score_3_other1$conservation_level = base::rep("total BL6 markers", nrow(psdf_mmms_markrand_score_3_other1))
```

```{r max_vi}
max_vi <- base::max(c(
  score_df_hsc_other1$value[score_df_hsc_other1$type == "variation_information"],
  psdf_hsc_sign_rand_other1$variation_information,
  psdf_hsc_mmms_signrand_other1$variation_information,
  psdf_hsc_mark_signrand_other1$variation_information,
  psdf_hsc_mmms_markrand_other1$variation_information)) + 0.2
```


