---
title: "Supplementary Figure 1"
date: '2024-11-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load objects

```{r seed, message = FALSE}
base::RNGkind(kind = "Mersenne-Twister")
set.seed(37)
```

```{r load, message = FALSE}
library(tidyverse, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
```

```{r base_path}
base_path <- "/omics/odcf/analysis/OE0538_projects_temp/DO-0008/data_temp/reproducibility3_test_rerun/data"
```

```{r load_manual_hsc}
# load data for manual use here

# SCE objects with all cells
sce_hsc <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_hsc-02"))
sce_hsc_ann <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"))
sce_hsc_fin <- base::readRDS(base::paste0(
  base_path, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_hsc-10"))

# remove some unneccessary data to save space
reducedDim(sce_hsc_ann) <- NULL
assays(sce_hsc_ann) <- list()
logc <- logcounts(sce_hsc)
assays(sce_hsc) <- list()
assays(sce_hsc_fin) <- list()
```

```{r load_manual}
sce_str <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02"))
sce_str_ann <- base::readRDS(base::paste0(
  base_path,
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10"))

reducedDim(sce_str_ann) <- NULL
assays(sce_str_ann) <- list()
assays(sce_str) <- list()
```

```{r source}
colors_path <- base::paste0(
  base_path, 
  "/metadata/colors/colors.txt")
 
source("../source/colors.R")
source("../source/plotting.R")
```

```{r pdf_output}
# define output paths for pdf export
output_pdf_e <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_e.pdf")

output_pdf_f <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_f.pdf")

output_pdf_g <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_g.pdf")

output_pdf_x <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_e.pdf")

output_pdf_guide0 <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide0.pdf")
output_pdf_guide1 <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide1.pdf")
output_pdf_guide2 <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide2.pdf")
output_pdf_guide2_legend <-  base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide2_legend.pdf")
output_pdf_guide3 <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide3.pdf")
output_pdf_guide4 <- base::paste0(
  base_path,
  "/manuscript1/supp_figure1/sfigure1_guide4.pdf")
```

```{r params}
# global params
source("determine_params.R")
```

# UMAPs 

with mature hematopoietic cells and original louvain clusters

## Prepare objects

```{r transfer_labels_hsc}

sce_hsc$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc), colnames(sce_hsc_ann))])
sce_hsc$annotation_all_subcl <- sce_hsc_ann$annotation_subcluster[
  base::match(colnames(sce_hsc), colnames(sce_hsc_ann))]

# 11 12 13

sce_hsc$annotation_all[sce_hsc$cluster_louvain == "11"] <- "mature antigen-processing cells"
sce_hsc$annotation_all_subcl[sce_hsc$cluster_louvain == "11"] <- "mature antigen-processing cells"

sce_hsc$annotation_all[sce_hsc$cluster_louvain == "12"] <- "low-quality"
sce_hsc$annotation_all_subcl[sce_hsc$cluster_louvain == "12"] <- "low-quality"

sce_hsc$annotation_all[sce_hsc$cluster_louvain == "13"] <- "mature neutrophil"
sce_hsc$annotation_all_subcl[sce_hsc$cluster_louvain == "13"] <- "mature neutrophil"

sce_hsc$annotation_all <- factor(sce_hsc$annotation_all,
  levels = c(levels(sce_hsc_ann$annotation_cluster), 
             "mature antigen-processing cells",
             "low-quality",
             "mature neutrophil"))

sce_hsc$annotation_all_subcl <- factor(sce_hsc$annotation_all_subcl,
  levels = c(names(col_cts_hsc), 
             "mature antigen-processing cells",
             "low-quality",
             "mature neutrophil"))
```

```{r prep_hsc}
# change species names to currently used abbreviations
sce_hsc$species_pub <- vector(length = ncol(sce_hsc))
sce_hsc$species_pub[sce_hsc$Species_ID == "mmus"] <- "BL6"
sce_hsc$species_pub[sce_hsc$Species_ID == "mcas"] <- "CAST"
sce_hsc$species_pub[sce_hsc$Species_ID == "mspr"] <- "SPRET"
sce_hsc$species_pub[sce_hsc$Species_ID == "mcar"] <- "CAROLI"
sce_hsc$species_pub <- factor(sce_hsc$species_pub, 
                              levels = names(col_spc_pub))
```

```{r transfer_labels_str}
unique(sce_str$cluster_louvain)

sce_str_ann$annotation_cluster <- factor(
  sce_str_ann$annotation_cluster,
  levels = levels(sce_str_ann$annotation_cluster)[c(2, 1, 3:9)])

sce_str$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  match(colnames(sce_str), colnames(sce_str_ann))])
sce_str$annotation_all_subcl <- sce_str_ann$annotation_subcluster[
  match(colnames(sce_str), colnames(sce_str_ann))]

unique(sce_str$cluster_louvain[is.na(sce_str$annotation_all_subcl)])

# 1  2  3  5  8  11 14 15 16 17 19

sce_str$annotation_all[sce_str$cluster_louvain == "1"] <- "erythrocytic 1"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "1"] <- "erythrocytic 1"

sce_str$annotation_all[sce_str$cluster_louvain == "2"] <- "granulocytic 1"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "2"] <- "granulocytic 1"

sce_str$annotation_all[sce_str$cluster_louvain == "3"] <- "basophil lineage"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "3"] <- "basophil lineage"

sce_str$annotation_all[sce_str$cluster_louvain == "5"] <- "granulocytic 2"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "5"] <- "granulocytic 2"

sce_str$annotation_all[sce_str$cluster_louvain == "8"] <- "neutrophil"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "8"] <- "neutrophil"

sce_str$annotation_all[sce_str$cluster_louvain == "11"] <- "b lymphocyte 1"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "11"] <- "b lymphocyte 1"

sce_str$annotation_all[sce_str$cluster_louvain == "14"] <- "activated lymphocyte"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "14"] <- "activated lymphocyte"

sce_str$annotation_all[sce_str$cluster_louvain == "15"] <- "low quality"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "15"] <- "low quality"

sce_str$annotation_all[sce_str$cluster_louvain == "16"] <- "b lymphocyte 2"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "16"] <- "b lymphocyte 2"

sce_str$annotation_all[sce_str$cluster_louvain == "17"] <- "antigen presenting cell"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "17"] <- "antigen presenting cell"

sce_str$annotation_all[sce_str$cluster_louvain == "19"] <- "skeletal muscle"
sce_str$annotation_all_subcl[sce_str$cluster_louvain == "19"] <- "skeletal muscle"

sce_str$annotation_all <- factor(sce_str$annotation_all,
  levels = c(levels(sce_str_ann$annotation_cluster), 
             "erythrocytic 1",
             "granulocytic 1",
             "granulocytic 2",
             "basophil lineage",
             "neutrophil",
             "b lymphocyte 1",
             "b lymphocyte 2",
             "activated lymphocyte",
             "antigen presenting cell",
             "low quality",
             "skeletal muscle"))

sce_str$annotation_all_subcl <- factor(sce_str$annotation_all_subcl,
  levels = c(names(col_cts_str), 
             "erythrocytic 1",
             "granulocytic 1",
             "granulocytic 2",
             "basophil lineage",
             "neutrophil",
             "b lymphocyte 1",
             "b lymphocyte 2",
             "activated lymphocyte",
             "antigen presenting cell",
             "low quality",
             "skeletal muscle"))
```

```{r prep_str}
# change species names to currently used abbreviations
sce_str$species_pub <- vector(length = ncol(sce_str))
sce_str$species_pub[sce_str$Species_ID == "mmus"] <- "BL6"
sce_str$species_pub[sce_str$Species_ID == "mcas"] <- "CAST"
sce_str$species_pub[sce_str$Species_ID == "mspr"] <- "SPRET"
sce_str$species_pub[sce_str$Species_ID == "mcar"] <- "CAROLI"
sce_str$species_pub <- factor(sce_str$species_pub, 
                              levels = names(col_spc_pub))
```

## Make DFs

```{r prep_hsc2}
hsc_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc, type = "UMAP"))
hsc_umap$color_by <- sce_hsc$annotation_all 
hsc_umap$species <- sce_hsc$species_pub 

hsc_umap_fin <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc_fin, type = "UMAP"))
hsc_umap_fin$color_by <- sce_hsc_fin$celltypes

hsc_umap_mecom <- hsc_umap
hsc_umap_mecom$expression <- logc[rownames(logc) == "Mecom",]

hsc_umap <- hsc_umap[base::sample(
  1:nrow(hsc_umap),
  nrow(hsc_umap),
  replace = FALSE),]
hsc_umap$color_by <- factor(
  hsc_umap$color_by,
  levels = levels(sce_hsc$annotation_all)
)

hsc_umap_subcl <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc, type = "UMAP"))
hsc_umap_subcl$color_by <- unfactor(sce_hsc$annotation_all_subcl)
hsc_umap_subcl$species <- sce_hsc$species_pub 
hsc_umap_subcl <- hsc_umap_subcl[base::sample(
  1:nrow(hsc_umap_subcl),
  nrow(hsc_umap_subcl),
  replace = FALSE),]
hsc_umap_subcl$color_by[!hsc_umap_subcl$color_by %in% names(col_cts_hsc)] <- "removed"

hsc_umap_subcl$color_by <- factor(
  hsc_umap_subcl$color_by,
  levels = c(names(col_cts_hsc), "removed")
)

col_clust <- col_num[1:length(levels(hsc_umap$color_by))]
names(col_clust) <- levels(hsc_umap$color_by)

```

```{r prep_str2}
str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str, type = "UMAP"))
str_umap$color_by <- sce_str$annotation_all 
str_umap$species <- sce_str$species_pub 
str_umap <- str_umap[base::sample(
  1:nrow(str_umap),
  nrow(str_umap),
  replace = FALSE),]
str_umap$color_by <- factor(
  str_umap$color_by,
  levels = levels(sce_str$annotation_all)
)

str_umap_subcl <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str, type = "UMAP"))
str_umap_subcl$color_by <- unfactor(sce_str$annotation_all_subcl)
str_umap_subcl$species <- sce_str$species_pub 
str_umap_subcl <- str_umap_subcl[base::sample(
  1:nrow(str_umap_subcl),
  nrow(str_umap_subcl),
  replace = FALSE),]
str_umap_subcl$color_by[!str_umap_subcl$color_by %in% names(col_cts_str)] <- "removed"

str_umap_subcl$color_by <- factor(
  str_umap_subcl$color_by,
  levels = c(names(col_cts_str), "removed")
)
```

## Plots

```{r e_base_plot_hsc_spc}

e_base_plot_hsc_spc <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = V1,
      y = V2,
      color = species))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2))+1)+
  ggplot2::scale_color_manual(values = col_spc_pub)
```

```{r e_base_plot_hsc_1}

e_base_plot_hsc_1 <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2))+1)+
  ggplot2::scale_color_manual(values = col_clust)
```

```{r e_base_plot_hsc_1_mecom}
names(mycolors_to_max) <- as.character(
  as.numeric(names(mycolors_to_max))*max(hsc_umap_mecom$expression))

e_base_plot_hsc_1_mecom <- ggplot2::ggplot(
  hsc_umap_mecom[hsc_umap_mecom$color_by %in% c("Stem/MPP1", "Stem/MPP2"),], 
  aes(x = V1,
      y = V2,
      color = expression))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2))+1)+
  ggplot2::scale_color_gradientn("Mecom expression", colours = mycolors_to_max)

legend_mecom <- cowplot::get_legend(e_base_plot_hsc_1_mecom+theme_all)
legend_mecom <- ggpubr::ggarrange(legend_mecom)
```

```{r e_base_plot_hsc_1_subcl}
e_base_plot_hsc_1_subcl <- ggplot2::ggplot(
  hsc_umap_subcl[hsc_umap_subcl$color_by %in% c("Early MPP", "HSC"),], 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$V1)-1, base::max(hsc_umap$V1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap$V2)-1, base::max(hsc_umap$V2))+1)+
  ggplot2::scale_color_manual(values = c("HSC" = "grey10", "Early MPP" = "grey70"), na.value = "white")
```

```{r e_base_plot_hsc_1_final}
e_base_plot_hsc_1_final <- ggplot2::ggplot(
  hsc_umap_fin, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap_fin$X1)-1, base::max(hsc_umap_fin$X1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap_fin$X2)-1, base::max(hsc_umap_fin$X2))+1)+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r e_base_plot_hsc_2}
e_base_plot_hsc_2 <- ggplot2::ggplot(
  hsc_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap_subcl$V1)-1, base::max(hsc_umap_subcl$V2))+1)+
  ggplot2::ylim(c(base::min(hsc_umap_subcl$V2)-1, base::max(hsc_umap_subcl$V2))+1)+
  ggplot2::scale_color_manual(values = c(col_cts_hsc, "removed" = "grey50"))
```

```{r e_base_plot_str_1}
e_base_plot_str_1 <- ggplot2::ggplot(
  str_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$V1)-1, base::max(str_umap$V1))+1)+
  ggplot2::ylim(c(base::min(str_umap$V2)-1, base::max(str_umap$V2))+1)
```

```{r e_base_plot_str_2}
e_base_plot_str_2 <- ggplot2::ggplot(
  str_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap_subcl$V1)-1, base::max(str_umap_subcl$V2))+1)+
  ggplot2::ylim(c(base::min(str_umap_subcl$V2)-1, base::max(str_umap_subcl$V2))+1)+
  ggplot2::scale_color_manual(values = c(col_cts_str, "removed" = "grey50"))
```

```{r e_theme_plot_hsc_spc, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_spc <- e_base_plot_hsc_spc+
  cowplot::theme_nothing()
e_theme_plot_hsc_spc
```

```{r e_theme_plot_hsc_1, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_1 <- e_base_plot_hsc_1+
  cowplot::theme_nothing()
e_theme_plot_hsc_1
```

```{r e_theme_plot_hsc_1_mecom, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_1_mecom <- e_base_plot_hsc_1_mecom+
  cowplot::theme_nothing()
e_theme_plot_hsc_1_mecom
```

```{r e_theme_plot_hsc_1_subcl, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_1_subcl <- e_base_plot_hsc_1_subcl+
  cowplot::theme_nothing()
e_theme_plot_hsc_1_subcl
```

```{r e_theme_plot_hsc_1_final, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_1_final <- e_base_plot_hsc_1_final+
  cowplot::theme_nothing()
e_theme_plot_hsc_1_final
```

```{r e_theme_plot_hsc_2, fig.width = 3, fig.height = 3}
e_theme_plot_hsc_2 <- e_base_plot_hsc_2+
  cowplot::theme_nothing()
e_theme_plot_hsc_2
```


```{r e_theme_plot_str_1, fig.width = 3, fig.height = 3}
e_theme_plot_str_1 <- e_base_plot_str_1+
  cowplot::theme_nothing()
```

```{r e_theme_plot_str_2, fig.width = 3, fig.height = 3}
e_theme_plot_str_2 <- e_base_plot_str_2+
  cowplot::theme_nothing()
```

## Legends

```{r hsc_legends}

col_clust <- col_num[1:length(levels(hsc_umap$color_by))]
names(col_clust) <- levels(hsc_umap$color_by)

e_legend_plot_hsc_1 <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = col_clust)

e_legend_theme_plot_hsc_1 <- e_legend_plot_hsc_1+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
e_legend_hsc_1 <- ggpubr::get_legend(e_legend_theme_plot_hsc_1)
e_legend_hsc_1 <- ggpubr::ggarrange(e_legend_hsc_1)

e_legend_hsc_1



e_legend_plot_hsc_2 <- ggplot2::ggplot(
  hsc_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = c(col_cts_hsc, "removed" = "grey50"))

e_legend_theme_plot_hsc_2 <- e_legend_plot_hsc_2+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
e_legend_hsc_2 <- ggpubr::get_legend(e_legend_theme_plot_hsc_2)
e_legend_hsc_2 <- ggpubr::ggarrange(e_legend_hsc_2)

e_legend_hsc_2

```

```{r str_legends}

e_legend_plot_str_1 <- ggplot2::ggplot(
  str_umap, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_discrete("")

e_legend_theme_plot_str_1 <- e_legend_plot_str_1+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
e_legend_str_1 <- ggpubr::get_legend(e_legend_theme_plot_str_1)
e_legend_str_1 <- ggpubr::ggarrange(e_legend_str_1)

e_legend_str_1

e_legend_plot_str_2 <- ggplot2::ggplot(
  str_umap_subcl, 
  aes(x = V1,
      y = V2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_legend_point_size,
    alpha = umap_legend_point_alpha)+
  ggplot2::scale_color_manual("", values = c(col_cts_str, "removed" = "grey50"))

e_legend_theme_plot_str_2 <- e_legend_plot_str_2+
  theme_all+
  ggplot2::theme(legend.key.height = unit(0.3, "lines"))
e_legend_str_2 <- ggpubr::get_legend(e_legend_theme_plot_str_2)
e_legend_str_2 <- ggpubr::ggarrange(e_legend_str_2)

e_legend_str_2
```

```{r assemble_e, fig.width = 16, fig.height = 4}

e_plot <- ggpubr::ggarrange(
  e_theme_plot_hsc_1,
  e_theme_plot_hsc_2,
  e_theme_plot_str_1,
  e_theme_plot_str_2,
  e_legend_hsc_1,
  e_legend_hsc_2,
  e_legend_str_1,
  e_legend_str_2,
  ncol = 4,
  widths = c(1,1,1,1),
  heights = c(1,1)
)

e_plot

```

```{r export_e}
pdf(output_pdf_e, width = 16, height = 4)
e_plot
dev.off()
```

```{r output_pdf_guide0}
pdf(output_pdf_guide0, width = 4, height = 4)
e_theme_plot_hsc_spc
dev.off()
```

```{r output_pdf_guide1}
pdf(output_pdf_guide1, width = 4, height = 4)
e_theme_plot_hsc_1
dev.off()
```

```{r output_pdf_guide2}
pdf(output_pdf_guide2, width = 4, height = 4)
e_theme_plot_hsc_1_mecom
dev.off()
```

```{r output_pdf_guide2_legend}
pdf(output_pdf_guide2_legend, width = 3, height = 3)
legend_mecom
dev.off()
```

```{r output_pdf_guide3}
pdf(output_pdf_guide3, width = 4, height = 4)
e_theme_plot_hsc_1_subcl
dev.off()
```

```{r output_pdf_guide4}
pdf(output_pdf_guide4, width = 4, height = 4)
e_theme_plot_hsc_1_final
dev.off()
```

# Sankey plots

## Prep HSC

```{r load_spsp_hsc}

reducedDim(sce_hsc) <- NULL
reducedDim(sce_str) <- NULL

sce_hsc_spsp_mmus <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mmus_hsc-02"))
reducedDim(sce_hsc_spsp_mmus) <- NULL
assays(sce_hsc_spsp_mmus) <- list()

sce_hsc_spsp_mcas <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcas_hsc-02"))
reducedDim(sce_hsc_spsp_mcas) <- NULL
assays(sce_hsc_spsp_mcas) <- list()

sce_hsc_spsp_mspr <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mspr_hsc-02"))
reducedDim(sce_hsc_spsp_mspr) <- NULL
assays(sce_hsc_spsp_mspr) <- list()

sce_hsc_spsp_mcar <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcar_hsc-02"))
reducedDim(sce_hsc_spsp_mcar) <- NULL
assays(sce_hsc_spsp_mcar) <- list()
```

```{r prepare_spsp}
sce_hsc_spsp_mmus$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mmus), colnames(sce_hsc_ann))])
sce_hsc_spsp_mcas$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mcas), colnames(sce_hsc_ann))])
sce_hsc_spsp_mspr$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mspr), colnames(sce_hsc_ann))])
sce_hsc_spsp_mcar$annotation_all <- unfactor(sce_hsc_ann$annotation_cluster[
  base::match(colnames(sce_hsc_spsp_mcar), colnames(sce_hsc_ann))])
```

```{r prepare_spsp_mmus}
hsc_spsp_mmus_df <- base::as.data.frame(
  table(sce_hsc_spsp_mmus$annotation_all,
       unfactor(sce_hsc_spsp_mmus$cluster_louvain)))

hsc_spsp_mmus_df$Var1 <- factor(
  hsc_spsp_mmus_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mmus_df$Var2 <- factor(
  hsc_spsp_mmus_df$Var2,
  levels = c("5", "2", "6", "8", "9", "4", "3", "7", "1"))
stopifnot(!is.na(hsc_spsp_mmus_df$Var2))
```

```{r prepare_spsp_mcas}
hsc_spsp_mcas_df <- base::as.data.frame(
  table(sce_hsc_spsp_mcas$annotation_all,
       unfactor(sce_hsc_spsp_mcas$cluster_louvain)))

hsc_spsp_mcas_df$Var1 <- factor(
  hsc_spsp_mcas_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mcas_df$Var2 <- factor(
  hsc_spsp_mcas_df$Var2,
  levels = c("1", "7", "8", "3", "2", "9", "6", "4", "10", "5"))
stopifnot(!is.na(hsc_spsp_mcas_df$Var2))
```

```{r prepare_spsp_mspr}
hsc_spsp_mspr_df <- base::as.data.frame(
  table(sce_hsc_spsp_mspr$annotation_all,
       unfactor(sce_hsc_spsp_mspr$cluster_louvain)))

hsc_spsp_mspr_df$Var1 <- factor(
  hsc_spsp_mspr_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mspr_df$Var2 <- factor(
  hsc_spsp_mspr_df$Var2,
  levels = c("3", "2", "5", "10", "11", "6", "7", "1", "8", "4", "9"))
stopifnot(!is.na(hsc_spsp_mspr_df$Var2))
```

```{r prepare_spsp_mcar}
hsc_spsp_mcar_df <- base::as.data.frame(
  table(sce_hsc_spsp_mcar$annotation_all,
       unfactor(sce_hsc_spsp_mcar$cluster_louvain)))

hsc_spsp_mcar_df$Var1 <- factor(
  hsc_spsp_mcar_df$Var1,
  levels = levels(sce_hsc_ann$annotation_cluster))

hsc_spsp_mcar_df$Var2 <- factor(
  hsc_spsp_mcar_df$Var2,
  levels = c("2", "1", "6", "8", "4", "3", "5", "9", "7", "10", "11"))
stopifnot(!is.na(hsc_spsp_mcar_df$Var2))
```

```{r mmus_hsc_base_plot, fig.height = 5, fig.width = 4}
mmus_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mmus_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mcas_hsc_base_plot, fig.height = 5, fig.width = 4}
mcas_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mcas_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mspr_hsc_base_plot, fig.height = 5, fig.width = 4}
mspr_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mspr_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)

```

```{r mcar_hsc_base_plot, fig.height = 5, fig.width = 4}
mcar_hsc_base_plot <- ggplot2::ggplot(data = hsc_spsp_mcar_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_fill_manual(values = col_clust)
```

```{r assemble_g_hsc, fig.width = 13, fig.height = 4}

f_hsc_plot <- ggpubr::ggarrange(
  mmus_hsc_base_plot,
  mcas_hsc_base_plot,
  mspr_hsc_base_plot,
  mcar_hsc_base_plot,
  nrow = 1
)
f_hsc_plot
```

## Prep Str

```{r load_spsp_str}

sce_str_spsp_mmus <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mmus_str-02"))
reducedDim(sce_str_spsp_mmus) <- NULL
assays(sce_str_spsp_mmus) <- list()

sce_str_spsp_mcas <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcas_str-02"))
reducedDim(sce_str_spsp_mcas) <- NULL
assays(sce_str_spsp_mcas) <- list()

sce_str_spsp_mspr <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mspr_str-02"))
reducedDim(sce_str_spsp_mspr) <- NULL
assays(sce_str_spsp_mspr) <- list()

sce_str_spsp_mcar <- base::readRDS(base::paste0(
  base_path_input, 
  "/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/comparison/sce_mcar_str-02"))
reducedDim(sce_str_spsp_mcar) <- NULL
assays(sce_str_spsp_mcar) <- list()
```


```{r prepare_spsp_str}
sce_str_spsp_mmus$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mmus), colnames(sce_str_ann))])
sce_str_spsp_mcas$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mcas), colnames(sce_str_ann))])
sce_str_spsp_mspr$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mspr), colnames(sce_str_ann))])
sce_str_spsp_mcar$annotation_all <- unfactor(sce_str_ann$annotation_cluster[
  base::match(colnames(sce_str_spsp_mcar), colnames(sce_str_ann))])
```

```{r prepare_spsp_mmus_str}
str_spsp_mmus_df <- base::as.data.frame(
  table(sce_str_spsp_mmus$annotation_all,
       unfactor(sce_str_spsp_mmus$cluster_louvain)))
base::table(sce_str_spsp_mmus$annotation_all,
            sce_str_spsp_mmus$cluster_louvain)

str_spsp_mmus_df$Var1 <- factor(
  str_spsp_mmus_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mmus_df$Var2 <- factor(
  str_spsp_mmus_df$Var2,
  levels = c("1", "5", "7", "11", "3", "2", "4", "9", "10", "8", "6", "12"))
stopifnot(!is.na(str_spsp_mmus_df$Var2))
# 6 and 12 are empty
```

```{r prepare_spsp_mcas_str}
str_spsp_mcas_df <- base::as.data.frame(
  table(sce_str_spsp_mcas$annotation_all,
       unfactor(sce_str_spsp_mcas$cluster_louvain)))

base::table(sce_str_spsp_mcas$annotation_all,
            sce_str_spsp_mcas$cluster_louvain)

str_spsp_mcas_df$Var1 <- factor(
  str_spsp_mcas_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mcas_df$Var2 <- factor(
  str_spsp_mcas_df$Var2,
  levels = c("1", "13", "5", "6", "10", "11", "12", "14", "15", "16", "7", "8", "9", "2", "3", "4"))
stopifnot(!is.na(str_spsp_mmus_df$Var2))
# "11", "12", "14", "15", "16", "7", "8", "9", "2", "3", "4" are empty
```

```{r prepare_spsp_mspr_str}
str_spsp_mspr_df <- base::as.data.frame(
  table(sce_str_spsp_mspr$annotation_all,
       unfactor(sce_str_spsp_mspr$cluster_louvain)))

base::table(sce_str_spsp_mspr$annotation_all,
            sce_str_spsp_mspr$cluster_louvain)

str_spsp_mspr_df$Var1 <- factor(
  str_spsp_mspr_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mspr_df$Var2 <- factor(
  str_spsp_mspr_df$Var2,
  levels = c("4", "11", "13", "3", "5", "8", "9", "2", "1", "6", "7", "12",  "10"))
stopifnot(!is.na(str_spsp_mspr_df$Var2))
```

```{r prepare_spsp_mcar_str}
str_spsp_mcar_df <- base::as.data.frame(
  table(sce_str_spsp_mcar$annotation_all,
       unfactor(sce_str_spsp_mcar$cluster_louvain)))

base::table(sce_str_spsp_mcar$annotation_all,
            sce_str_spsp_mcar$cluster_louvain)

str_spsp_mcar_df$Var1 <- factor(
  str_spsp_mcar_df$Var1,
  levels = levels(sce_str_ann$annotation_cluster))

str_spsp_mcar_df$Var2 <- factor(
  str_spsp_mcar_df$Var2,
  levels = c("4", "7", "8", "11", "9", "10", "12", "1", "2", "3", "5", "6"))
stopifnot(!is.na(str_spsp_mcar_df$Var2))
# 9, 10, 12, 1, 2, 3, 5, 6 are empty
```

```{r mmus_str_base_plot, fig.height = 5, fig.width = 4}
mmus_str_base_plot <- ggplot2::ggplot(data = str_spsp_mmus_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mcas_str_base_plot, fig.height = 5, fig.width = 4}
mcas_str_base_plot <- ggplot2::ggplot(data = str_spsp_mcas_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mspr_str_base_plot, fig.height = 5, fig.width = 4}
mspr_str_base_plot <- ggplot2::ggplot(data = str_spsp_mspr_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r mcar_str_base_plot, fig.height = 5, fig.width = 4}
mcar_str_base_plot <- ggplot2::ggplot(data = str_spsp_mcar_df,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
```

```{r assemble_f_str, fig.width = 13, fig.height = 4}

f_str_plot <- ggpubr::ggarrange(
  mmus_str_base_plot,
  mcas_str_base_plot,
  mspr_str_base_plot,
  mcar_str_base_plot,
  nrow = 1
)
f_str_plot
```

```{r,  fig.width = 13, fig.height = 8}
f_plot <- ggpubr::ggarrange(
  f_hsc_plot,
  f_str_plot,
  nrow = 2
)
f_plot
```

```{r}
pdf(output_pdf_f, width = 16, height = 4)
f_plot
dev.off()
```



# Species proportions

## Plot

```{r g_hsc_base_plot}

df_hsc <- as.data.frame(colData(sce_hsc_ann)[colnames(colData(sce_hsc_ann)) %in% c("Species_ID", "Age_ID", "celltypes")])
table_hsc <- table(sce_hsc_ann$celltypes, sce_hsc_ann$Species_ID)
table_hsc

df_hsc$Species_ID <- factor(df_hsc$Species_ID, levels = names(col_spc))
df_hsc$Age_ID <- factor(df_hsc$Age_ID, levels = names(col_age))

g_hsc_base_plot <- ggplot2::ggplot(
  df_hsc, 
  aes(x = Age_ID,
      fill = Species_ID))+
  ggplot2::geom_bar()+
  facet_grid(cols = vars(df_hsc$celltypes))
g_hsc_base_plot
```

Except for mcar lymph, there are at least 97 cells per specues and cell types.

```{r g_hsc_theme_plot}
g_hsc_theme_plot <- g_hsc_base_plot+
  theme_all+
  ggplot2::xlab("Age and cell types")+
  ggplot2::ylab("# cells")+
  ggplot2::scale_fill_manual("Species", values = col_spc)+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.text.x = element_text(
      size = axis_text_size,
      angle = 90,
      vjust = 0.5,
      hjust = 0
    ),
    strip.background = element_blank())
g_hsc_theme_plot
```

```{r g_str_base_plot}

df_str <- as.data.frame(colData(sce_str_ann)[colnames(colData(sce_str_ann)) %in% c("Species_ID", "Age_ID", "celltypes")])
table_str <- table(sce_str_ann$celltypes, sce_str_ann$Species_ID)
table_str

df_str$Species_ID <- factor(df_str$Species_ID, levels = names(col_spc))
df_str$Age_ID <- factor(df_str$Age_ID, levels = names(col_age))

g_str_base_plot <- ggplot2::ggplot(
  df_str, 
  aes(x = Age_ID,
      fill = Species_ID))+
  ggplot2::geom_bar()+
  facet_grid(cols = vars(df_str$celltypes))
g_str_base_plot
```

```{r g_str_theme_plot}
g_str_theme_plot <- g_str_base_plot+
  theme_all+
  ggplot2::xlab("Age and cell types")+
  ggplot2::ylab("# cells")+
  ggplot2::scale_fill_manual("Species", values = col_spc)+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.text.x = element_text(
      size = axis_text_size,
      angle = 90,
      vjust = 0.5,
      hjust = 0),
    strip.background = element_blank())
g_str_theme_plot
```

```{r assemble_g, fig.width = 16, fig.height = 6}
g_plot <- ggpubr::ggarrange(
  g_hsc_theme_plot+ggplot2::theme(legend.position = "none"),
  align = "hv",
  g_str_theme_plot,
  widths = c(1, 1.2)
)
g_plot
```

```{r export_g}
pdf(output_pdf_g, width = 16, height = 6)
g_plot
dev.off()
```

## UMAPs Cell Types

HSPCs (left) and Niche cells (right) clustered in UMAP space, 
colored by cell type.

#### Plot Components

```{r ct_prep, eval = FALSE, include = FALSE}
hsc_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_hsc, type = "UMAP"))
hsc_umap$color_by <- sce_hsc$celltypes
hsc_umap <- hsc_umap[base::sample(
  1:nrow(hsc_umap),
  nrow(hsc_umap),
  replace = FALSE),]

str_umap <- base::as.data.frame(
  SingleCellExperiment::reducedDim(
    sce_str, type = "UMAP"))
str_umap$color_by <- sce_str$celltypes
str_umap <- str_umap[base::sample(
  1:nrow(str_umap),
  nrow(str_umap),
  replace = FALSE),]
```

```{r ct_base_plot_left, eval = FALSE, include = FALSE}
ct_base_plot_left <- ggplot2::ggplot(
  hsc_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(hsc_umap$X1)-1, base::max(hsc_umap$X1))+1)+
  ggplot2::ylim(c(base::min(hsc_umap$X2)-1, base::max(hsc_umap$X2))+1)+
  ggplot2::scale_color_manual("Cell type", values = col_cts_hsc)
```

```{r ct_theme_plot_left, eval = FALSE, include = FALSE}
ct_theme_plot_left <- ct_base_plot_left+
  cowplot::theme_nothing()
```

```{r ct_base_plot_right, eval = FALSE, include = FALSE}
ct_base_plot_right <- ggplot2::ggplot(
  str_umap, 
  aes(x = X1,
      y = X2,
      color = color_by))+
  ggplot2::geom_point(
    size = umap_point_size,
    alpha = umap_point_alpha)+
  ggplot2::xlim(c(base::min(str_umap$X1)-1, base::max(str_umap$X1))+1)+
  ggplot2::ylim(c(base::min(str_umap$X2)-1, base::max(str_umap$X2))+1)+
  ggplot2::scale_color_manual("Cell type", values = col_cts_str)
```

```{r ct_theme_plot_right, eval = FALSE, include = FALSE}
ct_theme_plot_right <- ct_base_plot_right+
  cowplot::theme_nothing()
```

```{r sessioninfo}
sessionInfo()
```
