---
title: "check mapping specific"
author: "Lea Wölbert"
date: '2023-04-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report QC for all samples separately (Object_ID).
Some operations from the scripts must be repeated to generate relevant plots,
but any new objects cannot be saved from an .Rmd script.

#### Load libraries, source code 

```{r load,  message = FALSE}

library(SingleCellExperiment, quietly = TRUE) 
library(DropletUtils, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
library(pheatmap, quietly = TRUE) 
library(biomaRt)
library(Seurat)

set.seed(37)
```

```{r load_objects, message = FALSE}

source(file = "../source/plotting.R")

sce_og <- readRDS(file = snakemake@input[["sce_07"]])
sce_fg_path <- snakemake@input[["sce_fg_path"]]

# set name for plots
name_curr <- colData(sce_og)$Object_ID[1] # name to display in plots
nr_hvgs <- snakemake@params[["nr_hvgs"]]

spec_curr <- snakemake@wildcards[["species"]]
fraction_curr <- snakemake@wildcards[["fraction"]]
individuals <- snakemake@params[["individuals"]]

#sce_og <- readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/main_analysis/sce_objects/07_mrge_species/sce_mspr_hsc-07")
#sce_fg_path <- list("/omics/odcf/analysis/OE0538_projects/DO-0008/data/fourgenomes/sce_objects/01_cellranger_output/mspr/sce_mspr_old_hsc_2_0-01", "/omics/odcf/analysis/OE0538_projects/DO-0008/data/fourgenomes/sce_objects/01_cellranger_output/mspr/sce_mspr_old_hsc_1_0-01")
#individuals <- c("mspr_old_hsc_1_0", "mspr_old_hsc_2_0")
#spec_curr <- "mspr"

sce_fg_list <- list()
for(i in individuals){
  if(grepl(spec_curr, i) & grepl(fraction_curr, i)){
    print(i)
    print(paste0(sce_fg_path, "/", spec_curr, "/sce_", i, "-01"))
    sce_fg_list[[i]] <- readRDS(file = paste0(sce_fg_path, "/", spec_curr, "/sce_", i, "-01"))
    colnames(sce_fg_list[[i]]) <- paste0(colnames(sce_fg_list[[i]]), "_", i)
  }
}

#print(sce_fg_list)

if(fraction_curr == "str"){
  knitr::knit_exit()
}

gene_list <- read.csv(snakemake@params[["gene_list_hsc_all"]], 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

class(gene_list)
print(gene_list)

genes <- unique(gene_list$gene)
print(genes)

```

# Prepare objects

```{r}

# subset barcodes
sce_fg_list_qc <- lapply(sce_fg_list, function(sce_fg){
  sce_fg <- sce_fg[,which(!is.na(match(colnames(sce_fg), colnames(sce_og))))]
  return(sce_fg)
})

```

MSPR objects don't have symbols

This part was taken from Perrine's script on how to download the relevant data 
from BioMart.

```{r}

if(grepl("mspr", spec_curr)){

  ensembl_mspr <- useMart(biomart="ENSEMBL_MART_ENSEMBL",
                          host="https://www.ensembl.org",
                          dataset="mspretus_gene_ensembl")
  ensembl_list_mspr=getBM(attributes=c("ensembl_gene_id",
                                       "mmusculus_homolog_ensembl_gene",
                                       "mmusculus_homolog_associated_gene_name"), 
                          mart=ensembl_mspr)

  sce_fg_list_qc <- lapply(sce_fg_list_qc, function(sce_fg){
    
    intersect_IDs <- intersect(rowData(sce_fg)$Symbol, ensembl_list_mspr$ensembl_gene_id)

    ensembl_list_mspr <- ensembl_list_mspr[ensembl_list_mspr$ensembl_gene_id %in% intersect_IDs,]
    sce_fg <- sce_fg[rowData(sce_fg)$Symbol %in% intersect_IDs,]

    ensembl_list_mspr <- ensembl_list_mspr[!duplicated(ensembl_list_mspr$ensembl_gene_id),]
    sce_fg <- sce_fg[!duplicated(rowData(sce_fg)$Symbol),]


    rowData(sce_fg)$Symbol[
      match(ensembl_list_mspr$ensembl_gene_id,
            rownames(sce_fg))] <- ensembl_list_mspr$mmusculus_homolog_associated_gene_name

    rowData(sce_fg)$Symbol[rowData(sce_fg)$Symbol == ""] <- rowData(sce_fg)$ID[rowData(sce_fg)$Symbol == ""]
    
    return(sce_fg)
  })
}

```

```{r}

sce_fg_list_shared <- lapply(sce_fg_list_qc, function(sce_fg){
  
  shared_genes <- intersect(rowData(sce_og)$Symbol, rowData(sce_fg)$Symbol)
  length(shared_genes)

  sce_fg_shared <- sce_fg[rowData(sce_fg)$Symbol %in% shared_genes,]

  sce_fg_shared <- sce_fg_shared[order(rowData(sce_fg_shared)$Symbol),]

  sce_fg_shared <- sce_fg_shared[!duplicated(rowData(sce_fg_shared)$Symbol),]
  return(sce_fg_shared)
  
})

shared_genes <- intersect(rowData(sce_og)$Symbol, rowData(sce_fg_list_shared[[1]])$Symbol)
sce_og_shared <- sce_og[rowData(sce_og)$Symbol %in% shared_genes,]
sce_og_shared <- sce_og_shared[order(rowData(sce_og_shared)$Symbol),]
sce_og_shared <- sce_og_shared[!duplicated(rowData(sce_og_shared)$Symbol),]

```

merge all fg objects

objects that are not in the species merges have ncol = 0
by using cbind they get filteres out automatically

```{r}

for(i in 1:length(sce_fg_list_shared)){
  if(i == 1){
      sce_fg_merged <- sce_fg_list_shared[[i]] # set first item of list
  }else{
      sce_fg_merged <- cbind(sce_fg_merged, sce_fg_list_shared[[i]]) 
  }
}

sce_fg_merged
sce_og_shared

```

```{r}

colnames(sce_og_shared) <- paste0(colnames(sce_og_shared), "_og")
sce_og_shared$Genome <- rep("OneGenome", ncol(sce_og_shared))

colnames(sce_fg_merged) <- paste0(colnames(sce_fg_merged), "_fg")
sce_fg_merged$Genome <- rep("FourGenomes", ncol(sce_fg_merged))

rownames(sce_og_shared) <- rowData(sce_og_shared)$Symbol 
rownames(sce_fg_merged) <- rowData(sce_fg_merged)$Symbol 

colData(sce_og_shared) <- colData(sce_og_shared)[,colnames(colData(sce_og_shared)) %in% colnames(colData(sce_fg_merged))]

colnames(rowData(sce_og_shared))[colnames(rowData(sce_og_shared)) == "ID"] <- "ID_og"
colnames(rowData(sce_fg_merged))[colnames(rowData(sce_fg_merged)) == "ID"] <- "ID_fg"

assays(sce_og_shared) <- list(counts = counts(sce_og_shared))
assays(sce_fg_merged)$counts[1:10, 1:10]
assays(sce_og_shared)$counts[1:10, 1:10]

reducedDim(sce_og_shared) <- NULL
reducedDim(sce_og_shared) <- NULL

# combine
sce_tog <- cbind(sce_og_shared, sce_fg_merged)

# normalize
set.seed(37)

quick_clust <- quickCluster(sce_tog) # quick clustering to accelerate calculation
sce_tog <- computeSumFactors(sce_tog, cluster = quick_clust)
sce_tog <- logNormCounts(sce_tog) # log-transformation

sce_tog

```

# Markergenes

```{r}

# convert to seurat
seurat <- as.Seurat(sce_tog, counts = "counts", data = "logcounts",) 
Idents(seurat) <- sce_tog$Genome

# get hvgs for marker genes
hvgs <- modelGeneVar(sce_tog)
hvgs <- getTopHVGs(hvgs, n=nr_hvgs)

# find marker genes for each cluster
genomes <- unique(sce_tog$Genome)
clustlist <- as.list(genomes)
cluster_markers <- lapply(clustlist, function(x){
  markers <- FindMarkers(seurat, test.use = "wilcox", ident.1 = x, 
                         features = hvgs)
  markers <- markers[order(abs(markers$avg_log2FC), decreasing=TRUE),]
  markers$which_cluster <- rep(x, nrow(markers))
  return(markers)
})

names(cluster_markers) <- genomes
markergenes_og <- cluster_markers[["OneGenome"]]

markergenes_og[,c(2, 3, 5)]

markergenes_og_temp <- markergenes_og
if(nrow(markergenes_og_temp) > 200){markergenes_og_temp <- markergenes_og_temp[1:200,]}

```

```{r, fig.height=20, fig.width=7}

sce_tog_temp <- sce_tog[which(rownames(sce_tog) %in% rownames(markergenes_og_temp)),]

mat <- as.matrix(
  assays(sce_tog_temp)$logcounts
)

anno_col <- data.frame(
  features = colnames(sce_tog_temp),
  Genome = colData(sce_tog_temp)$Genome)

anno_col <- column_to_rownames(anno_col, var = "features")

pheatmap(
  mat,
  cluster_cols = FALSE,
  annotation_col = anno_col,
  show_colnames = FALSE
)

```

# specific genes

```{r}

genes <- unique(gene_list[,1])
print(genes)
print(genes[which(genes %in% rownames(markergenes_og))])

```

```{r}

sce_tog_temp <- sce_tog[which(rownames(sce_tog) %in% genes),]

if(is.null(genes)){knitr::knit_exit()}

mat <- as.matrix(
  assays(sce_tog_temp)$logcounts
)

anno_col <- data.frame(
  features = colnames(sce_tog_temp),
  Genome = colData(sce_tog_temp)$Genome)

anno_col <- column_to_rownames(anno_col, var = "features")

print(mat[1:10,1:10])

pheatmap(
  mat,
  cluster_cols = FALSE,
  annotation_col = anno_col,
  show_colnames = FALSE,
  show_rownames = FALSE
)


```



