---
title: "man_annotation_report"
author: "Lea Wölbert, Adrien Jolly"
date: '2022-11-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r load,  message = FALSE}

library(SingleCellExperiment, quietly = TRUE) 
library(DropletUtils, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
#library(scater, quietly = TRUE) 
#library(scuttle, quietly = TRUE) 
#library(scran, quietly = TRUE) 
library(pheatmap)

```

```{r load_object, message = FALSE}

sce <- readRDS(file = snakemake@input[["sce_08"]])
color_tables <- snakemake@params[["color_tables"]]

# set name for plots
spec_curr <- colData(sce)$Species[1] # choose name to display in plots

```

```{r load_code, message = FALSE}

source(file = "../source/colors.R")
source(file = "../source/plotting.R")
source(file = "../source/sce_functions.R")

col_cts_all <- col_cts[names(col_cts) %in% sce$Identity_ref_all]
col_cts_fraction1 <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction1]
col_cts_fraction2 <- col_cts[names(col_cts) %in% sce$Identity_ref_fraction2]
col_batch_exp_day <- col_alp[names(col_alp) %in% sce$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% sce$Batch_sequencing]
col_ann_age <- col_ann[names(col_ann) %in% sce$Age_ID]
col_ann_fraction <- col_ann[names(col_ann) %in% sce$Fraction_ID]

```

# Merged QC

## Identities

```{r fig.width=11}

plot_1 <- umap_base(sce, color_by = "Identity_ref_all")+ # own function
  scale_color_manual("Identity1", values = col_cts_all)
plot_1l <- umap_legend(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "Identity_ref_fraction1")+ 
  scale_color_manual("Identity2", values = col_cts_fraction1)
plot_2l <- umap_legend(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)
legend_2 <- get_legend(plot_2l)

plot_3 <- umap_base(sce, color_by = "Identity_ref_fraction2")+ 
  scale_color_manual("Identity3", values = col_cts_fraction2)
plot_3l <- umap_legend(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)
legend_3 <- get_legend(plot_3l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)

PLOT_1+ggtitle(spec_curr)
PLOT_2+ggtitle(spec_curr)
PLOT_3+ggtitle(spec_curr)

```

### Contamination

Use only Baccin et al dataset because it's the only one containing both
stromal and hematopoetic cell types in a single dataset.

```{r}

# find contamination in sce objects
sce_str <- sce[,sce$Fraction_ID == "str"]
sce_hsc <- sce[,sce$Fraction_ID == "hsc"]

sce_str <- find_contamination_ref_all(sce_str, input = "stromal") # own function
sce_hsc <- find_contamination_ref_all(sce_hsc, input = "hsc") # own function

```

```{r fig.width=11}

if(grepl("TRUE", sce_str$right_fraction)){
  plot_4 <- umap_base(sce_str, color_by = "right_fraction")+
    scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                      "FALSE" = "orange"))
  plot_4l <- umap_legend(sce_str, color_by = "right_fraction")+
    scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                      "FALSE" = "orange"))
  legend_4 <- get_legend(plot_4l)

  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_4+ggtitle("Stromal")
}

if(grepl("TRUE", sce_hsc$right_fraction)){
  plot_4 <- umap_base(sce_hsc, color_by = "right_fraction")+
    scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                      "FALSE" = "orange"))
  plot_4l <- umap_legend(sce_hsc, color_by = "right_fraction")+
    scale_color_manual("Correct fraction", values = c("TRUE" = "black", 
                                                      "FALSE" = "orange"))
  legend_4 <- get_legend(plot_4l)

  PLOT_4 <- ggarrange(plot_4, legend_4)
  PLOT_4+ggtitle("HSPCs")
}

```

### Mature cell types

```{r}

sce <- get_mature_cts_all(sce) # own function 

```

```{r fig.width=8}

plot_4 <- umap_base(sce, color_by = "mature_cells")+
  scale_color_manual("Correct fraction", values = c("FALSE" = "black", 
                                                    "TRUE" = "orange"))
plot_4l <- umap_legend(sce, color_by = "mature_cells")+
  scale_color_manual("Correct fraction", values = c("FALSE" = "black", 
                                                    "TRUE" = "orange"))
legend_4 <- get_legend(plot_4l)

PLOT_4 <- ggarrange(plot_4, legend_4)
PLOT_4+ggtitle(spec_curr)

```

## QC Metrics

```{r fig.width=8}

plot_1 <- umap_base(sce, color_by = "Batch_exp_day")+ # own function
  scale_color_manual("Experimental day", values = col_batch_exp_day)
plot_1l <- umap_legend(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Experimental day", values = col_batch_exp_day)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Sequencing day", values = col_batch_seq)
plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Sequencing day", values = col_batch_seq)
legend_2 <- get_legend(plot_2l)

plot_3 <- umap_base(sce, color_by = "Object_ID")
plot_3l <- umap_legend(sce, color_by = "Object_ID")
legend_3 <- get_legend(plot_3l)

plot_4 <- umap_base(sce, color_by = "Age_ID")+
  scale_color_manual("Age", values = col_ann_age)
plot_4l <- umap_legend(sce, color_by = "Age_ID")+
  scale_color_manual("Age", values = col_ann_age)
legend_4 <- get_legend(plot_4l)

plot_5 <- umap_base(sce, color_by = "Fraction_ID")+
  scale_color_manual("Fraction", values = col_ann_fraction)
plot_5l <- umap_legend(sce, color_by = "Fraction_ID")+
  scale_color_manual("Fraction", values = col_ann_fraction)
legend_5 <- get_legend(plot_5l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
PLOT_4 <- ggarrange(plot_4, legend_4)
PLOT_5 <- ggarrange(plot_5, legend_5)

PLOT_1+ggtitle(spec_curr)
PLOT_2+ggtitle(spec_curr)
PLOT_3+ggtitle(spec_curr)
PLOT_4+ggtitle(spec_curr)
PLOT_5+ggtitle(spec_curr)

```

```{r fig.width=11}

plot_1 <- umap_base(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_1l <- umap_legend(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)
legend_1 <- get_legend(plot_1l)

plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)
legend_2 <- get_legend(plot_2l)

plot_3 <- umap_base(sce, color_by = "Object_ID")+ 
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_3l <- umap_legend(sce, color_by = "Object_ID")
legend_3 <- get_legend(plot_3l)

# check age
plot_4 <- umap_base(sce, color_by = "Object_ID")+ 
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))
plot_4l <- umap_legend(sce, color_by = "Object_ID")
legend_4 <- get_legend(plot_4l)

plot_5 <- umap_base(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))
plot_5l <- umap_legend(sce, color_by = "Batch_sequencing")+
  scale_color_manual("Batch_sequencing", values = col_batch_seq)
legend_5 <- get_legend(plot_5l)

plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))
plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)
legend_6 <- get_legend(plot_6l)

plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))
plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
  scale_color_manual("Batch_experimental_day", values = col_batch_exp_day)
legend_6 <- get_legend(plot_6l)

plot_7 <- umap_base(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_7l <- umap_legend(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)
legend_idrefall <- get_legend(plot_7l)

plot_8 <- umap_base(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_8l <- umap_legend(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)
legend_idreffrac1 <- get_legend(plot_8l)

plot_9 <- umap_base(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Fraction))
plot_9l <- umap_legend(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)
legend_idreffrac2 <- get_legend(plot_9l)

plot_10 <- umap_base(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))

plot_11 <- umap_base(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))

plot_12 <- umap_base(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Age))

plot_13 <- umap_base(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_sequencing))

plot_14 <- umap_base(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_sequencing))

plot_15 <- umap_base(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_sequencing))

plot_16 <- umap_base(sce, color_by = "Identity_ref_all")+
  scale_color_manual("Identity1", values = col_cts_all)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_exp_day))

plot_17 <- umap_base(sce, color_by = "Identity_ref_fraction1")+
  scale_color_manual("Identity2", values = col_cts_fraction1)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_exp_day))

plot_18 <- umap_base(sce, color_by = "Identity_ref_fraction2")+
  scale_color_manual("Identity3", values = col_cts_fraction2)+
  ggtitle(spec_curr)+
  facet_grid(cols = vars(sce$Batch_exp_day)) 

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_3 <- ggarrange(plot_3, legend_3)
PLOT_4 <- ggarrange(plot_4, legend_4)
PLOT_5 <- ggarrange(plot_5, legend_5)
PLOT_6 <- ggarrange(plot_6, legend_6)
PLOT_7 <- ggarrange(plot_7, legend_idrefall)
PLOT_8 <- ggarrange(plot_8, legend_idreffrac1)
PLOT_9 <- ggarrange(plot_9, legend_idreffrac2)
PLOT_10 <- ggarrange(plot_10, legend_idrefall)
PLOT_11 <- ggarrange(plot_11, legend_idreffrac1)
PLOT_12 <- ggarrange(plot_12, legend_idreffrac2)
PLOT_13 <- ggarrange(plot_13, legend_idrefall)
PLOT_14 <- ggarrange(plot_14, legend_idreffrac1)
PLOT_15 <- ggarrange(plot_14, legend_idreffrac2)
PLOT_16 <- ggarrange(plot_14, legend_idrefall)
PLOT_17 <- ggarrange(plot_14, legend_idreffrac1)
PLOT_18 <- ggarrange(plot_14, legend_idreffrac2)

PLOT_1
PLOT_2
PLOT_3
PLOT_4
PLOT_5
PLOT_6
PLOT_7
PLOT_8
PLOT_9
PLOT_10
PLOT_11
PLOT_12
PLOT_13
PLOT_14
PLOT_15
PLOT_16
PLOT_17
PLOT_18

```

# NMF programs

## Heatmaps

```{r}

colnames_int <- c(
  "Batch_exp_day",
  "Batch_sequencing",
  "Object_ID",
  "Age_ID",
  "Identity_ref_all",
  "Identity_ref_fraction1",
  "Identity_ref_fraction2"
)

```

### Stromal programs

```{r fig.height = 6.5}

nmf_str_usg <- colData(sce_str)[,grepl("str_program", 
                                       colnames(colData(sce_str)))]

for(c in colnames_int){
  variables <- unique(colData(sce_str)[,which(colnames(colData(sce)) == c)])
  mat_str <- matrix(nrow = length(variables), ncol = ncol(nmf_str_usg))

  #computes the mean factor usage per cell identity normalized 0-1
  for(i in 1:length(variables)){
    v <- variables[i]
    mat_str[i,] <- colMeans(as.matrix(nmf_str_usg[
      which(colData(sce_str)[,which(colnames(colData(sce)) == c)] == v),]))/max(
        colMeans(as.matrix(nmf_str_usg[
      which(colData(sce_str)[,which(colnames(colData(sce)) == c)] == v),])))
  }

  colnames(mat_str) <- colnames(nmf_str_usg)
  rownames(mat_str) <- variables

  if(c %in% c("Batch_exp_day","Batch_sequencing", "Object_ID", "Age_ID")){
    mat_str <- t(mat_str)
  }
  pheatmap(mat_str)
}

```

### HSC programs

```{r fig.height = 6.5}

nmf_hsc_usg <- colData(sce_hsc)[,grepl("hsc_program", 
                                       colnames(colData(sce_hsc)))]

for(c in colnames_int){
  variables <- unique(colData(sce_hsc)[,which(colnames(colData(sce)) == c)])
  mat_str <- matrix(nrow = length(variables), ncol = ncol(nmf_hsc_usg))

  #computes the mean factor usage per cell identity normalized 0-1
  for(i in 1:length(variables)){
    v <- variables[i]
    mat_str[i,] <- colMeans(as.matrix(nmf_hsc_usg[
      which(colData(sce_hsc)[,which(colnames(colData(sce)) == c)] == v),]))/max(
        colMeans(as.matrix(nmf_hsc_usg[
      which(colData(sce_hsc)[,which(colnames(colData(sce)) == c)] == v),])))
  }

  colnames(mat_str) <- colnames(nmf_hsc_usg)
  rownames(mat_str) <- variables

  if(c %in% c("Batch_exp_day","Batch_sequencing", "Object_ID", "Age_ID")){
    mat_str <- t(mat_str)
  }
  pheatmap(mat_str)
}

```

## UMAPs

Stromal cells

```{r fig.width=8}

for(i in colnames(nmf_str_usg)){
  plot_1 <- umap_base(sce, color_by = i)
  plot_1l <- umap_legend(sce, color_by = i)+
    scale_color_continuous(name = i)
  legend_1 <- get_legend(plot_1l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  print(PLOT_1+ggtitle(spec_curr))
  
}

for(i in colnames(nmf_str_usg)){
  plot_1 <- umap_base(sce_str, color_by = i)
  plot_1l <- umap_legend(sce_str, color_by = i)+
    scale_color_continuous(name = i)
  legend_1 <- get_legend(plot_1l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  print(PLOT_1+ggtitle(paste0(spec_curr, " stromal cells only")))
}

```

HSCs

```{r fig.width=8}

for(i in colnames(nmf_hsc_usg)){
  plot_1 <- umap_base(sce, color_by = i)
  plot_1l <- umap_legend(sce, color_by = i)+
    scale_color_continuous(name = i)
  legend_1 <- get_legend(plot_1l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  print(PLOT_1+ggtitle(spec_curr))
  
}

for(i in colnames(nmf_hsc_usg)){
  plot_1 <- umap_base(sce_hsc, color_by = i)
  plot_1l <- umap_legend(sce_hsc, color_by = i)+
    scale_color_continuous(name = i)
  legend_1 <- get_legend(plot_1l)

  PLOT_1 <- ggarrange(plot_1, legend_1)
  print(PLOT_1+ggtitle(paste0(spec_curr, " HSPCs only")))
}

```

## Pie charts

This is just a simple check if a program is used per cell >0. 
Does not reflect actual cell types yet.

### Stromal cells

```{r}

programs <- unique(colnames(colData(sce))[grepl("str_program", colnames(colData(sce)))])

for(i in programs){

  ggdf <- data.frame(colData(sce))
  ggdf <- ggdf[which(ggdf[,colnames(colData(sce)) == i] > 0),]
   
  col_all <- col_cts_all[
    names(col_cts_all) %in% 
      sce$Identity_ref_all[colnames(colData(sce)) == i]]

  plot1 <- ggplot(ggdf, aes(x="", y=i, fill=Identity_ref_all))+
    geom_col()+
    theme_classic()+
    coord_polar(theta = "y")+
    scale_fill_manual("Identity1", values = col_all)
  
  col_frc1 <- col_cts_fraction1[
    names(col_cts_fraction1) %in% 
      sce$Identity_ref_fraction1[colnames(colData(sce)) == i]]

  plot2 <- ggplot(ggdf, aes(x="", y=i, fill=Identity_ref_fraction1))+
    geom_col()+
    theme_classic()+
    coord_polar(theta = "y")+
    scale_fill_manual("Identity2", values = col_frc1)
  
  col_frc2 <- col_cts_fraction2[
    names(col_cts_fraction2) %in% 
      sce$Identity_ref_fraction2[colnames(colData(sce)) == i]]

  plot3 <- ggplot(ggdf, aes(x="", y=i, fill=Identity_ref_fraction2))+
    geom_col()+
    theme_classic()+
    coord_polar(theta = "y")+
    scale_fill_manual("Identity3", values = col_frc2)
  
  col_batches <- col_batch_exp_day[
    names(col_batch_exp_day) %in% 
      sce$Batch_exp_day[colnames(colData(sce)) == i]]

  plot4 <- ggplot(ggdf, aes(x="", y=i, fill=Batch_exp_day))+
    geom_col()+
    theme_classic()+
    coord_polar(theta = "y")+
    scale_fill_manual("Experimental day", values = col_batches)
  
  PLOT_A <- ggarrange(plot1, plot2)
  PLOT_B <- ggarrange(plot3, plot4)
  
  print(annotate_figure(PLOT_A, top = text_grob(i)))
  print(annotate_figure(PLOT_B, top = text_grob(i)))
}

```


