---
title: "preprocessing_report"
author: "Amy Danson, Lea Wölbert"
date: '2022-09-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report QC for all objects separately.
Some operations from the scripts must be repeated to generate relevant plots,
but any new objects cannot be saved from an .Rmd script.

#### Load libraries, source code 

```{r load,  message = FALSE}

library(SingleCellExperiment, quietly = TRUE) 
library(DropletUtils, quietly = TRUE) 
library(tidyverse, quietly = TRUE) 
library(scater, quietly = TRUE) 
library(scuttle, quietly = TRUE) 
library(scran, quietly = TRUE) 
set.seed(37)
```

```{r load_objects, message = FALSE}

sce_01 <- readRDS(file = snakemake@input[["sce_01"]])
sce_02 <- readRDS(file = snakemake@input[["sce_02"]])
sce_03 <- readRDS(file = snakemake@input[["sce_03"]])
sce_04 <- readRDS(file = snakemake@input[["sce_04"]])
sce_05 <- readRDS(file = snakemake@input[["sce_05"]])

# set name for plots
name_curr <- colData(sce_02)$Object_ID[1] # name to display in plots

```


```{r declare_cutoffs, message = FALSE}

cutoff_umis <- snakemake@params[["cutoff_umis"]]
cutoff_doublets <- snakemake@params[["cutoff_doublets"]] 
cutoff_sum <- snakemake@params[["cutoff_sum"]] 
cutoff_detected <- snakemake@params[["cutoff_detected"]]
cutoff_mitos <- snakemake@params[["cutoff_mitos"]]
color_tables <- snakemake@params[["color_tables"]]

```

```{r}
source(file = "../source/colors.R")
source(file = "../source/plotting.R")
```


## Step 1: call objects and add metadata

```{r}

print(sce_01)
print(colData(sce_01))

```

## Step 2: 

### Remove empty droplets

Check that the barcode rank and p value distributions look appropriate.
Also check if any droplets with low FDR (that are likely cells) were dropped

```{r}

# show barcoded ranks 
bcrank <- barcodeRanks(counts(sce_01)) #before droplet removal
uniq_pos <- !duplicated(bcrank$rank)

plot(bcrank$rank[uniq_pos], bcrank$total[uniq_pos], log="xy", 
     xlab="Rank", ylab="Total UMI count", cex.lab=1.2, main = name_curr)
  abline(h=metadata(bcrank)$inflection, col = marker1)
  abline(h=metadata(bcrank)$knee, col = marker2)
  legend("bottomleft", legend=c("Inflection", "Knee"), 
         col=c(marker1, marker2), cex=1.2)

# show pval distribution    
cutoff_umis <- cutoff_umis
out <- emptyDrops(counts(sce_01), lower = cutoff_umis, test.ambient=TRUE)  

hist(out$PValue[out$Total <= cutoff_umis & out$Total > 0],
     xlab="P-value",
     main = paste(name_curr, "cutoff", cutoff_umis), col="grey80")  

# show FDR of droplets that were kept or removed
check <- table(Sig=out$FDR <= 0.001, Limited=out$Limited)
print(check)

```

### Remove doublets

Check if cells with high doublet scores form conspicuous clusters

```{r}

plotUMAP(sce_05, colour_by = "doublet_score")+ # requires reduced dims
  ggtitle(name_curr)+
  theme(legend.key.size = unit(0.3, "cm"), legend.key.width = unit(0.3, "cm"))+
  theme(legend.spacing = unit(0.06, "cm"), legend.text = element_text(size = 7))

```

## Step 3: Remove outliers

Check the quality of the single cells and if cells were removed appropriately

```{r}

# get mito count just to check
mito_genes <- grep("mt-", rownames(sce_02))
qcdf_02 <- perCellQCMetrics(sce_02, subsets=list(Mito=mito_genes)) # before outlier removal

# set QC cutoffs below which cells are removed
sum_out <- qcdf_02$sum < cutoff_sum 
det_out <- qcdf_02$detected < cutoff_detected 
mito_out <- qcdf_02$subsets_Mito_percent > cutoff_mitos

# remove all cells to discard and add info to qc
remove_pos <- sum_out | det_out | mito_out
qcdf_02$removed <- remove_pos

qcdf_gg <- as.data.frame(qcdf_02[,c(1,2,5,7)])

qcdf_gg <- pivot_longer(qcdf_gg, 
                        names_to = "metric", cols = c(1, 2, 3))

ggplot(qcdf_gg[qcdf_gg$metric == "sum",], aes(y = value, x = metric))+
  theme_classic()+
  ggbeeswarm::geom_quasirandom(size = 0.1, aes(color = removed))+
  geom_boxplot(alpha = 0)+
  scale_color_manual("Removed", values = c("FALSE" = "grey80", 
                                           "TRUE" = "orange"))+
  ylab("Library size")+
  ggtitle(name_curr)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  geom_abline(intercept = cutoff_sum, color = marker1,  slope = 0) 

ggplot(qcdf_gg[qcdf_gg$metric == "detected",], aes(y = value, x = metric))+
  theme_classic()+
  ggbeeswarm::geom_quasirandom(size = 0.1, aes(color = removed))+
  geom_boxplot(alpha = 0)+
  scale_color_manual("Removed", values = c("FALSE" = "grey80", 
                                           "TRUE" = "orange"))+
  ylab("Detected genes")+
  ggtitle(name_curr)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  geom_abline(intercept = cutoff_detected, color = marker1, slope = 0) 

ggplot(qcdf_gg[qcdf_gg$metric == "subsets_Mito_percent",], 
       aes(y = value, x = metric))+
  theme_classic()+
  ggbeeswarm::geom_quasirandom(size = 0.1, aes(color = removed))+
  geom_boxplot(alpha = 0)+
  scale_color_manual("Removed", values = c("FALSE" = "grey80", 
                                           "TRUE" = "orange"))+
  ylab("Mitochondrial gene percent")+
  ggtitle(name_curr)+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())+
  geom_abline(intercept = cutoff_mitos, color = marker1, slope = 0) 

```

## Step 4: Normalization

```{r}
# calculate two different size factors for direct comparison
quick_clust <- quickCluster(sce_03) # before normalization
sizef_deconv <- calculateSumFactors(sce_03, cluster = quick_clust)
sizef_lib <- librarySizeFactors(sce_03)

plot(sizef_lib, sizef_deconv, xlab = "Library size factor",
    ylab = "Deconv. size factor", log = 'xy', pch=16, 
    main = paste("Library/Deconv", name_curr))
    abline(a=0, b=1, col="red")
    
# distribution of size factors should be close to normal dist
hist(log10(sizef_deconv), xlab="Log10[Deconv factor]", col='grey80', 
     main = name_curr)
    
# directly compare deconv factors to library size
qcdf_03 <- perCellQCMetrics(sce_03) # after outlier removal
plot(sizef_deconv, qcdf_03$sum, log = "xy", ylab = "Library size",
     xlab = "Deconv. Size factor", main = paste("Deconv/libsize", name_curr))
```

```{r}

# print final object
print(name_curr)
print(sce_04) # after normalization

print(colData(sce_04))
print(ncol(sce_04))

```

## Remaining quality

```{r}

qcdf_05 <- perCellQCMetrics(sce_05)
sce_05$sum <- qcdf_05$sum
sce_05$detected <- qcdf_05$detected

plot_1 <- umap_base(sce_05, color_by = "sum")+
  scale_color_gradientn("Library size", 
                         limits = c(0, 100000),
                         colors = c("black", "darkorange4", "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))+
  ggtitle(name_curr)
plot_1l <- umap_legend(sce_05, color_by = "sum")+
  scale_color_gradientn("Library size", 
                         limits = c(0, 1000000),
                          colors = c("black", "darkorange4", "darkorange3", "darkorange1",
                                    "orange", "orange", "gold1", "gold",
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod", 
                                    "lightgoldenrod", "lightgoldenrod"))
legend_1 <- get_legend(plot_1l)

PLOT_1 <- ggarrange(plot_1, legend_1)
PLOT_1

plot_2 <- umap_base(sce_05, color_by = "detected")+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", "orange", "lightgoldenrod"))+
  ggtitle(name_curr)
plot_2l <- umap_legend(sce_05, color_by = "detected")+
  scale_color_gradientn("Number of detected genes", 
                         limits = c(0, 10000),
                         colors = c("black", "darkorange3", "orange", "lightgoldenrod"))
legend_2 <- get_legend(plot_2l)

PLOT_2 <- ggarrange(plot_2, legend_2)
PLOT_2

```


```{r}
sessionInfo()
```
