---
title: "Tabula sapiens bone marrow custom reclustering"
author: "Lea Wölbert"
date: '2025-12-04'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r load_objects}

seu_ts_bone_marrow <- readRDS(snakemake@input[["seu_input"]])

#seu_ts_bone_marrow <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/ts_bone_marrow")

#test <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/zeb_all_hspc")

# load ensembl df to get the ENS IDs for the genes I need
ensembl_hum <- base::readRDS(snakemake@input[["ensembl_hum"]])
#ensembl_hum <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/ensembl/ensembl_hum")
head(ensembl_hum)
```

```{r output_path}
# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
#output_path <- base::paste0(base_path_data, "")
print(base_path_data)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

Just try lots of different ones.

```{r parameters}
nr_k <- 15
resolution <- 0.5
```

Use the seurat piepline roughly takin from sce_functions_reclustering.R
The seurat object should be pretty basic.

For comparability between signature-reclustered and custom-reclustered
partitions, I will not perform batch correction here, either.

Keeping it very basic here.

```{r seurat_clustering_pipeline}

# seu_ts_bone_marrow uses raw data so need to normalize first

seu_ts_bone_marrow@misc$data_use 

seu_ts_bone_marrow <- Seurat::NormalizeData( 
      seu_ts_bone_marrow,
      assay = "RNA", 
      verbose = FALSE)

seu_ts_bone_marrow <- Seurat::ScaleData(
  seu_ts_bone_marrow, 
  verbose = FALSE)

# find variable features
seu_ts_bone_marrow <- Seurat::FindVariableFeatures(
  seu_ts_bone_marrow
)

# uses scaled data
seu_ts_bone_marrow <- Seurat::RunPCA(
  seu_ts_bone_marrow,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(seu_ts_bone_marrow@reductions$pca)
print(nr_pca)

# requires PC coordinates
seu_ts_bone_marrow <- Seurat::FindNeighbors(
  seu_ts_bone_marrow, 
  k.param = nr_k,
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
seu_ts_bone_marrow <- Seurat::FindClusters(
  seu_ts_bone_marrow, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
seu_ts_bone_marrow <-  Seurat::RunUMAP(seu_ts_bone_marrow, dims = 1:10)
```

# Evaluate

With signature genes, there were 9 clusters.

```{r table}
table(seu_ts_bone_marrow@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible batch effects but I can't do much about them.

```{r umaps_base, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")
``` 

```{r umaps_dev, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$assay), ncol = 2)
``` 

There is some strong effect by sex.

## Gene expression

Now check the quality of cell type separation based on expressed genes.

```{r marker_genes}

# these must be human genes
# try straight up with the same genes as for mouse first
# but then alter as necessary/possible
  marker_genes <- c(
    "ALDH1A1", # HSC, Stem
    "PROCR",
    "CDKN1C",
    "CD34",
    "MLLT3",
    "MMRN1",
    "MPL",
    "ALCAM", 
    "HLF",
    "LMO2",
    "MECOM",
    "MEIS1",
    "PBX1",
    "MSI2",
    "ADGRL4", # MPP, progenitors, activation
    "CDK4",
    "SOX4",
    "SPI1", 
    "CEBPA", # Myeloid
    "CTSG",
    "MPO",
    "PRTN3",
    "ELANE", # Neutro
    "CEBPE", 
    "S100A8",
    "S100A9",
    "WFDC21P",
    "CCR2", # Mono
    "IRF8",
    "MS4A6A",
    "FLT3", # Lymphoid
    "SATB1",
    "DNTT",
    "CD79B", 
    "IL7R",
    "RAG1",
    "CPA3", # Mast
    "HDC", 
    "MS4A2",
    "LMO4", # Baso
    "GATA2", # also Mk, Eo Baso and HSC
    "ITGA2B", # Mk
    "PF4",
    "BLVRB", # Mk Ery
    "GATA1",
    "ZFPM1",
    "EPOR", # Ery
    "HBA1",
    "KLF1",
    "HELLS",
    "LIG1",
    "MCM6", 
    "CENPA", # CC
    "MKI67"
  )

```

```{r ensmus_ids}
# not get the ensmus IDs
head(rownames(seu_ts_bone_marrow))

marker_genes[!marker_genes %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% marker_genes]

```

#### Heatmap

To get an overview of the rough expected populations based on well-known genes.

```{r heatmap, fig.height = 10, fig.width = 5}

seu_ts_bone_marrow@meta.data$clusters <- base::paste0(
  "cluster-",
  seu_ts_bone_marrow@meta.data$seurat_clusters
)

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_bone_marrow,
  assays = "RNA",
  group.by = c("clusters")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_bone_marrow$clusters),
    names_to = "clusters",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(marker_genes))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

#### UMAPs

```{r umaps_genes, fig.width = 4.5, fig.height = 4}
plotlist <- lapply(as.list(marker_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- marker_genes

plotlist
```

## Markers

```{r markers}
# replace the original idents with the new clusters 

Seurat::Idents(seu_ts_bone_marrow) <- seu_ts_bone_marrow@meta.data$clusters
clustlist <- as.list(base::sort(base::unique(seu_ts_bone_marrow@meta.data$clusters)))

cluster_markers <- lapply(clustlist, function(x){
  
  print(x)
  markers <- Seurat::FindMarkers(
    seu_ts_bone_marrow,
    slot = "data",
    test.use = "wilcox", 
    ident.1 = x)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  # add gene names otherwise I cannot interpret
  markers$genes <- vector(length = nrow(markers))
  markers$genes <- ensembl_hum$external_gene_name[
    match(
      rownames(markers),
      ensembl_hum$ensembl_gene_id)]

  return(markers)
})
names(cluster_markers) <- base::sort(base::unique(seu_ts_bone_marrow@meta.data$clusters))
```

Gave provisional names based on the heatmap.

### HSC

```{r markers_hsc}
cluster_markers[["cluster-5"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-6"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

Sorry, throughout this text I will also use the mouse gene nomenclature 
because I'm so used to it.

"Cluster-5" is female-biased and found in all assays.
It expresses Cdkn1c, Cd34, Mllt3, Mmrn, Mpl, Mecom, Meis1, Msi2, but no
Cebpa, Ctsg, Pf4, Mcm6, or Cenpa.

The top marker genes capture some HSC genes like Mllt3, some lymphoid stuff
lke Cd37, Igkc, Sell but a lot of the rest like the RPLs et could also come
from batch or other, probably between seyes? 
Since this cluster basically represents "female HSCs".

In contrast, cluster-6 is male-exclusive. 
High CD34, Mmrn1, Alcam, Hlf, Mecom, Meis1, PBX1, Msi2 expression.
Top marker genes like CD69, MSI2, EGR1, Jun, Mmrn1, also suggest HSCs, the
others I either don't know yet or also could be batch/sex.

```{r hsc_umaps_additional, fig.width = 4.5, fig.height = 4}

hsc_additional_genes <- c(
  "HOPX", # related to HSPC biology, possible stem cell marker, stemness
  "SOCS2", # HSC stemness
  "BST2", # CD137, HSC quiescence and maintenance, division
  "DEPP1", # related to autophagy? or hypoxia
  "CD52", # related to lymphoma, lymphocytes but not stem cells
  "CD37", # more lymphoid/mature
  "PRDX1", # antioxidant, peroxidase
  "SELL", # stemness
  "RPS4X", # ribosomal really looks like batch, low qu
  "FOS", # HSC subsets, stemness but not specific
  "JUN", # HSC subsets but not specific
  "EGR1", # immature state
  "JUNB", # HSC subsets but not specific
  "ELMO1" # in AML, related to Fos, HOPX, stemness
)

plotlist <- lapply(as.list(hsc_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- hsc_additional_genes

plotlist
```

Both look HSC enough for me.

### Myeloid

```{r markers_myeloid}

cluster_markers[["cluster-1"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
  
cluster_markers[["cluster-3"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-4"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-9"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

```

Cluster 1 doesn't seem to express early markers like Lmo2.
Both sexes but barely in smart-seq2.
High Sp1, Cebpa, low Ctsg, Mpo, Prtn3 and Elane.
Highest S100a8 and S100a9 and these are also the top marker genes together 
withg lots of other S100a genes as well as Lyz, Anxa2, Lgals1, Nkg7,
Cebpd, Lst1, Capg, Retn, MS4A6A, FCER1G.
Some Ccr2 and Irf8 but not the most. 
Also cycling.

Cluster 3 is found in all batches. 
Also no early markers but very high Ctsg, Mpo, Prtn3 and Elane, so
early-mid myeloid. 
Also a bit of Gata2 and other eo/baso must stuff, like CPA3, HDC, MS4A2.
The top markers are Prtn3, Mpo, Ctsg, Igll1, Calr, Srgn, Clec11a, 
Gata2, AZU1, PRSS57.
Could even be MPP because they also express some lymnphoid stuff and even some mk ery stuff like gata2, calr

Cluster 4 top markers are Lyz, Srgn, Cxcl8, Klf6, Actr2, Il1r2, Lyst,
Klf2, Mpo, Fosb, Azu1, Vcan, Elane, Fos.
It is exclusively male. 
Seems slightly earlier neutro than cluster 1.

Cluster 9 actually looks antigen-presenting and BC lineage; with CD74,
JCHAIN, HLAs and more.
It is detected in all batches.

```{r myeloid_umaps_additional, fig.width = 4.5, fig.height = 4}

myeloid_additional_genes <- c(
  "S100A10", # S100 genes primarily myeloid
  "S100A12",
  "LYZ", # lysozyme, myeloid, neutrophil
  "TYROBP", # defense
  "ANXA2",  # neutrophil
  "RETN", # neutrophil
  "CEBPD", # neutrophil
  "LST1", # mono/activation
  "MS4A6A", # mono
  "FCER1G", # neutrophil
  "LGALS1", # several
  "NKG7", # nk cells
  "CAPG", # macrophage capping 
  "CLC", # granu/myeloid ish early progs eo baso mast
  "SNORD3A",
  "CST7", # neutro ish
  "AZU1", # neutrophil, looks earlier
  "IGLL1",  # b cells
  "CALR", # mk ery?
  "PRSS57", # neutro
  "SRGN", # immune cell infiltration
  "CLEC11A", # immune cell infiltration
  "CXCL8",
  "EREG",
  "KLF6", # myeloid
  "KLF2",
  "ACTR2",
  "LYST", # lysosomes/granules e.g. nk or tc
  "VCAN",
  "CYBB", # general immune cell actibity
  "JCHAIN", # B cell lineage. antibody related
  "CD74", # antigen presenting
  "HLA-DPA1",
  "HLA-DRA",
  "LSP1", # all leukocytes, most in myeloid
  "IRF8", # myeloid/mono/macro/Dc and b cells
  "RUNX2", # nk cells and DCs
  "CD69", # add some other important markers
  "PROM1", # hsc
  "CD2", # neg in hsc
  "CD38", # neg in hsc
  "LYL1", # mpp
  "CD33", # cmp
  "LCN2", # gmp
  "CXCR2", # proneu
  "EPX", # eo
  "FCER1A", # bm
  "CLEC9A", # APC DC
  "ITGAX", # CD11c, dc
  "GP9", # mk
  "VWF", # mk
  "FLI1" # mk
)

plotlist <- lapply(as.list(myeloid_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  #print(gene_id %in% Features(seu_ts_bone_marrow))
  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- myeloid_additional_genes

plotlist
```

### Mk and Ery

```{r markers_mkery}

cluster_markers[["cluster-0"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-2"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-7"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-8"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
```

Cluster 0 expresses PRDX2, BLVRB, CA1 and CA2, HBD, CD36,
HBB, SMIM1, UROD among top markers.
Slightly more enriched in female data.
SOME early gene expression like CD34, Mllt3, Lmo2, still
expresses Sox4, some CPA3, ITGA2b, high GATA1, KLF1.
I'm guessing middle ery-mk, transitioning from early Mk to Ery.

Cluster 2 seems male-data exclusive.
Less CD34 than cluster 0, some CPA3, ITGA2b, BLVRB, GATA1, ZFPM1, EPOR.
Top markers include PNMT, MYC, KLF1, EEF genes, APOE, CDK6.

Cluster 7 is clearly Mk. 
Also expresses early markers like CD34. 

Cluster 9 is again male-data exclusive.
Strongly Ery, with HBAs and so on.

```{r erymk_umaps_additional, fig.width = 4.5, fig.height = 4}

erymk_additional_genes <- c(
  "CA1", # ery diff
  "CA2", # ery diff
  "PRDX2", # related to anemia
  "CYTL1", # ery mk progenitor
  "BLVRB", 
  "REXO2", # ery diff
  "HBD", # hemoglobin d
  "CD36",
  "KLF1",
  "HBB", # hemoglobin
  "UROD", # heme synthesis
  "PNMT",
  "EEF2", # elongation factor protein synth
  "EEF1G", # elongation factor protein synth
  "APOE", # apoloprotein e
  "TFR2", # transferrin receptor, precursors
  "HDHD3",
  "CDK6",
  "HMGA1",
  "PLEK",
  "CXCL3", # inflammatory
  "TGFB1", # mk signalling
  "MGLL",
  "CD9", # yes, CD for mks
  "HBA2", # hemeglobins
  "HBA1",
  "HBB",
  "ALAS2", # heme production
  "HEMGN", # hemogen
  "TFRC", # transferring receptir
  "GYPA" # red blood cell surface
)

plotlist <- lapply(as.list(erymk_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  #print(gene_id %in% Features(seu_ts_bone_marrow))
  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- erymk_additional_genes

plotlist
```

# Assign

```{r assign}
seu_ts_bone_marrow$custom_label <- vector(length = ncol(seu_ts_bone_marrow))

# hscs
# they MIGHT also be MPP but I cannot tell exactly
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-5"] <- "HSC, female-biased"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-6"] <- "HSC, male-biased"

# myeloid
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-3"] <- "Early granu prog"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-1"] <- "Neutro prog"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-4"] <- "Neutro prog, male-biased"

# antigen-presenting
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-9"] <- "Antigen-presenting and lymphoid"

# mk ery
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-7"] <- "Early Mk"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-0"] <- "Erythroid, female-biased"
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-2"] <- "Erythroid, male-biased"
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-8"] <- "Late erythroid, male-biased"

seu_ts_bone_marrow$custom_label <- factor(
  seu_ts_bone_marrow$custom_label,
  levels = c(
    "HSC, female-biased",
    "HSC, male-biased",
    "Early granu prog",
    "Neutro prog",
    "Neutro prog, male-biased",
    "Antigen-presenting and lymphoid",
    "Early Mk",
    "Erythroid, female-biased",
    "Erythroid, male-biased",
    "Late erythroid, male-biased"
  ))

stopifnot(!is.na(seu_ts_bone_marrow$custom_label))
```

#### One last plot to check

```{r slightly_better_markers}

# these must be human genes
# try straight up with the same genes as for mouse first
# but then alter as necessary/possible
slightly_better_markers <- c(
    "CD34",
    "MPL",
    "MEIS1",
    "MECOM",
    "MLLT3",
    "SOCS2",
    "HOPX",
    "CDKN1C",
    "HLF",
    "MMRN1",
    "MSI2",
    "PROM1",
    "ALCAM", 
    "CD38", # MPP, progenitors, activation
    "SPI1", 
    "CEBPA", # Myeloid
    "CTSG",
    "MPO",
    "PRTN3",
    "LMO4", # Eo Baso, Mk, HSC
    "CPA3", # Mast
    "GATA2", # Baso
    "HDC", 
    "MS4A2",
    "CEBPE", # Neutro
    "ELANE", 
    "AZU1",
    "S100A8",
    "S100A9",
    "S100A10",
    "LYZ",
    "CCR2", # Mono
    "MS4A6A",
    "IRF8",
    "ITGAX", # DC
    "FLT3", # Lymphoid early-late
    "IL7R",
    "SATB1", 
    "CD79B", 
    "DNTT",
    "JCHAIN", # later 
    "CD74", # antigen-presenting
    "HLA-DPA1",
    "HLA-DRA",
    "ITGA2B", # Mk
    "PF4",
    "GP9",
    "PLEK",
    "VWF",
    "PBX1",
    "ZFPM1", # Mk -> Ery
    "GATA1",
    "BLVRB", 
    "KLF1", # Ery early
    "EPOR",
    "HBA1",
    "HBB", # ery late
    "HBA2",
    "GYPA",
    "HELLS", # cc entry
    "LIG1",
    "MCM6", 
    "CENPA", # CC
    "MKI67"
  )

slightly_better_markers[!slightly_better_markers %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% slightly_better_markers]
```

```{r last_heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_bone_marrow,
  assays = "RNA",
  group.by = c("custom_label")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_bone_marrow$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_ts_df_genes$custom_label <- factor(
  agg_ts_df_genes$custom_label, 
  levels = levels(seu_ts_bone_marrow$custom_label))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```