---
title: "Tabula sapiens bone marrow custom reclustering"
author: "Lea Wölbert"
date: '2025-12-04'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r load_objects}

# the un-clustered original object
seu_ts_bone_marrow <- readRDS(snakemake@input[["seu_input"]])

# load ensembl df to get the ENS IDs for the genes I need
ensembl_hum <- base::readRDS(snakemake@input[["ensembl_hum"]])
head(ensembl_hum)
```

```{r output_path}
# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/03_custom_clustering/ts_bone_marrow.Rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

Just try lots of different ones.

```{r parameters}
nr_k <- 15
resolution <- 0.5
```

Use the seurat piepline roughly taken from sce_functions_reclustering.R
The seurat object should be pretty basic.

For comparability between signature-reclustered and custom-reclustered
partitions, I will not perform batch correction here, either.

Keeping it very basic here.

```{r seurat_clustering_pipeline}

# seu_ts_bone_marrow uses raw data so need to normalize first

seu_ts_bone_marrow@misc$data_use 

seu_ts_bone_marrow <- Seurat::NormalizeData( 
      seu_ts_bone_marrow,
      assay = "RNA", 
      verbose = FALSE)

seu_ts_bone_marrow <- Seurat::ScaleData(
  seu_ts_bone_marrow, 
  verbose = FALSE)

# find variable features
seu_ts_bone_marrow <- Seurat::FindVariableFeatures(
  seu_ts_bone_marrow
)

# uses scaled data
seu_ts_bone_marrow <- Seurat::RunPCA(
  seu_ts_bone_marrow,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(seu_ts_bone_marrow@reductions$pca)
print(nr_pca)

# requires PC coordinates
seu_ts_bone_marrow <- Seurat::FindNeighbors(
  seu_ts_bone_marrow, 
  k.param = nr_k,
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
seu_ts_bone_marrow <- Seurat::FindClusters(
  seu_ts_bone_marrow, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
seu_ts_bone_marrow <-  Seurat::RunUMAP(seu_ts_bone_marrow, dims = 1:10)
```

# Evaluate

With signature genes, there were 9 clusters.

```{r table}
table(seu_ts_bone_marrow@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible batch effects but I won't do much about them.

```{r umaps_base, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")
``` 

```{r umaps_dev, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_bone_marrow@meta.data$assay), ncol = 2)
``` 

There is some strong effect by sex.

## Gene expression

Now check the quality of cell type separation based on expressed genes.

```{r marker_genes}

# human genes
# try with the orthologues of mouse genes I would use here
# but then altered as necessary/possible
# from supplementary table 1 of the corresponding publication
  marker_genes <- c(
    "ALDH1A1",
    "PROCR",
    "CDKN1C",
    "CD34",
    "MLLT3",
    "MMRN1",
    "MPL",
    "ALCAM", 
    "HLF",
    "LMO2",
    "MECOM",
    "MEIS1",
    "PBX1",
    "MSI2",
    "ADGRL4",
    "CDK4",
    "SOX4",
    "SPI1", 
    "CEBPA", 
    "CTSG",
    "MPO",
    "PRTN3",
    "ELANE", 
    "CEBPE", 
    "S100A8",
    "S100A9",
    "WFDC21P",
    "CCR2", 
    "IRF8",
    "MS4A6A",
    "FLT3", 
    "SATB1",
    "DNTT",
    "CD79B", 
    "IL7R",
    "RAG1",
    "CPA3", 
    "HDC", 
    "MS4A2",
    "LMO4", 
    "GATA2", 
    "ITGA2B", 
    "PF4",
    "BLVRB", 
    "GATA1",
    "ZFPM1",
    "EPOR", 
    "HBA1",
    "KLF1",
    "HELLS",
    "LIG1",
    "MCM6", 
    "CENPA", 
    "MKI67"
  )
```

```{r ensmus_ids}
# not get the ensmus IDs
head(rownames(seu_ts_bone_marrow))

marker_genes[!marker_genes %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% marker_genes]

```

#### Heatmap

To get an overview of the rough expected populations based on well-known genes.

```{r heatmap, fig.height = 10, fig.width = 5}

seu_ts_bone_marrow@meta.data$clusters <- base::paste0(
  "cluster-",
  seu_ts_bone_marrow@meta.data$seurat_clusters
)

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_bone_marrow,
  assays = "RNA",
  group.by = c("clusters")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_bone_marrow$clusters),
    names_to = "clusters",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(marker_genes))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

#### UMAPs

```{r umaps_genes, fig.width = 4.5, fig.height = 4}
plotlist <- lapply(as.list(marker_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- marker_genes

plotlist
```

## Markers

```{r markers}
# replace the original idents with the new clusters 

Seurat::Idents(seu_ts_bone_marrow) <- seu_ts_bone_marrow@meta.data$clusters
clustlist <- as.list(base::sort(base::unique(seu_ts_bone_marrow@meta.data$clusters)))

cluster_markers <- lapply(clustlist, function(x){
  
  print(x)
  markers <- Seurat::FindMarkers(
    seu_ts_bone_marrow,
    slot = "data",
    test.use = "wilcox", 
    ident.1 = x)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  # add gene names otherwise I cannot interpret
  markers$genes <- vector(length = nrow(markers))
  markers$genes <- ensembl_hum$external_gene_name[
    match(
      rownames(markers),
      ensembl_hum$ensembl_gene_id)]

  return(markers)
})
names(cluster_markers) <- base::sort(base::unique(seu_ts_bone_marrow@meta.data$clusters))
```

Gave provisional names based on the heatmap.

### HSC

```{r markers_hsc}
cluster_markers[["cluster-5"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-6"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

Using mouse and human gene name conventions interchangeably because I'm
much more used to one than the other, sorry.

"Cluster-5" is female-biased and found in all assays.
It expresses Cdkn1c, Cd34, Mllt3, Mmrn, Mpl, Mecom, Meis1, Msi2, but no
Cebpa, Ctsg, Pf4, Mcm6, or Cenpa.

The top marker genes capture some HSC genes like Mllt3, some lymphoid stuff
lke Cd37, Igkc, Sell but a lot of the rest like the RPLs et could also come
from batch or other, probably between seyes? 
Since this cluster basically represents "female HSCs".

In contrast, cluster-6 is male-exclusive. 
High Cd34, Mmrn1, Alcam, Hlf, Mecom, Meis1, Pbx1, Msi2 expression.
Top marker genes like Cd69, Msi2, Egr1, Jun also suggest HSCs, the
others I either don't know yet or also could be batch/sex.

```{r hsc_umaps_additional, fig.width = 4.5, fig.height = 4}

# checking some interesting genes from the marker gene lists
hsc_additional_genes <- c(
  "HOPX", 
  "SOCS2", 
  "BST2", 
  "DEPP1", 
  "CD52", 
  "CD37", 
  "PRDX1",
  "SELL", 
  "RPS4X", 
  "FOS", 
  "JUN", 
  "EGR1", 
  "JUNB", 
  "ELMO1"
)

plotlist <- lapply(as.list(hsc_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- hsc_additional_genes

plotlist
```

Both look HSC enough for me.

### Myeloid

```{r markers_myeloid}

cluster_markers[["cluster-1"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
  
cluster_markers[["cluster-3"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-4"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-9"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

```

Cluster 1 doesn't seem to express early markers like Lmo2.
Both sexes but barely in smart-seq2.
High Sp1, Cebpa, low Ctsg, Mpo, Prtn3 and Elane.
Highest S100a8 and S100a9 and these are also the top marker genes together 
withg lots of other S100a genes as well as Lyz, Anxa2, Lgals1, Nkg7,
Cebpd, Lst1, Capg, Retn, MS4A6A, FCER1G.
Some Ccr2 and Irf8 but not the most. 
Also cycling.

Cluster 3 is found in all batches. 
Also no early markers but very high Ctsg, Mpo, Prtn3 and Elane, so
early-mid myeloid. 
Also a bit of Gata2 and other eo/baso must stuff, like CPA3, HDC, MS4A2.
The top markers are Prtn3, Mpo, Ctsg, Igll1, Calr, Srgn, Clec11a, 
Gata2, AZU1, PRSS57.
Could even be MPP because they also express some lymnphoid stuff and
even some mk ery stuff like Gata2.

Cluster 4 top markers are Lyz, Srgn, Cxcl8, Klf6, Actr2, Il1r2, Lyst,
Klf2, Mpo, Fosb, Azu1, Vcan, Elane, Fos.
It is exclusively male. 
Seems slightly earlier neutro than cluster 1.

Cluster 9 actually looks antigen-presenting and BC lineage; with CD74,
JCHAIN, HLAs and more.
It is detected in all batches.

```{r myeloid_umaps_additional, fig.width = 4.5, fig.height = 4}

# checking some interesting genes, some of which are well known
myeloid_additional_genes <- c(
  "S100A10", # S100 genes primarily myeloid
  "S100A12",
  "LYZ", # lysozyme, myeloid, neutrophil
  "TYROBP", # defense
  "ANXA2",  # neutrophil
  "RETN", # neutrophil
  "CEBPD", # neutrophil
  "LST1", # mono/activation
  "MS4A6A", # mono
  "FCER1G", # neutrophil
  "NKG7", # nk cells
  "CAPG", # macrophage capping 
  "CLC", # granu/myeloid ish early progs eo baso mast
  "SNORD3A",
  "CST7", 
  "AZU1",
  "IGLL1", # Ig gene
  "CALR", 
  "PRSS57", # Neutro/Granu https://doi.org/10.1073/pnas.1200470109
  "SRGN", # neutros and others https://doi.org/10.3389/fonc.2013.00327
  "CLEC11A", 
  "CXCL8",
  "EREG",
  "KLF6",
  "KLF2",
  "ACTR2",
  "LYST", 
  "VCAN",
  "CYBB",
  "JCHAIN", # B cell lineage. antibody related
  "CD74", # antigen presenting
  "HLA-DPA1", # antigen presenting
  "HLA-DRA" ,# antigen presenting
  "LSP1",
  "IRF8", 
  "RUNX2", # macro TF
  "PROM1", # add some other important markers from literature https://doi.org/10.3389/fphys.2022.1009160
  "CD2", # mature lineage marker https://doi.org/10.4049/jimmunol.0802008
  "CD38" # neg in hsc https://doi.org/10.3389/fphys.2022.1009160
)

plotlist <- lapply(as.list(myeloid_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  #print(gene_id %in% Features(seu_ts_bone_marrow))
  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- myeloid_additional_genes

plotlist
```

### Mk and Ery

```{r markers_mkery}

cluster_markers[["cluster-0"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-2"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-7"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

cluster_markers[["cluster-8"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
```

Cluster 0 expresses PRDX2, BLVRB, CA1 and CA2, HBD, CD36,
HBB, SMIM1, UROD among top markers.
Slightly more enriched in female data.
SOME early gene expression like CD34, Mllt3, Lmo2, still
expresses Sox4, some CPA3, ITGA2b, high GATA1, KLF1.
I'm guessing middle ery-mk, transitioning from early Mk to Ery.

Cluster 2 seems male-data exclusive.
Less CD34 than cluster 0, some CPA3, ITGA2b, BLVRB, GATA1, ZFPM1, EPOR.
Top markers include PNMT, MYC, KLF1, EEF genes, APOE, CDK6.

Cluster 7 is clearly Mk. 
Also expresses early markers like CD34. 

Cluster 9 is again male-data exclusive.
Strongly Ery, with HBAs and so on.

```{r erymk_umaps_additional, fig.width = 4.5, fig.height = 4}

# interesting genes from the marker gene list
erymk_additional_genes <- c(
  "CA1", # ery diff
  "CA2", # ery diff
  "PRDX2", 
  "CYTL1", # ery mk progenitor
  "BLVRB", 
  "REXO2", # ery diff
  "HBD", # hemoglobin d
  "CD36",
  "KLF1",
  "HBB", # hemoglobin
  "UROD", # heme synthesis
  "PNMT",
  "EEF2", # elongation factor protein synth
  "EEF1G", # elongation factor protein synth
  "APOE", 
  "TFR2", # transferrin receptor, precursors
  "HDHD3",
  "CDK6",
  "HMGA1",
  "PLEK",
  "CXCL3", # inflammatory
  "TGFB1", # mk signalling
  "MGLL",
  "CD9", # Mk
  "HBA2", # hemeglobins
  "HBA1",
  "HBB",
  "ALAS2", # heme production
  "HEMGN", # hemogen
  "TFRC", # transferring receptir
  "GYPA" # red blood cell surface
)

plotlist <- lapply(as.list(erymk_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  #print(gene_id %in% Features(seu_ts_bone_marrow))
  plot <- FeaturePlot(seu_ts_bone_marrow, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- erymk_additional_genes

plotlist
```

# Assign

```{r assign}
seu_ts_bone_marrow$custom_label <- vector(length = ncol(seu_ts_bone_marrow))

# hscs
# they MIGHT also be MPP but I cannot tell exactly
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-5"] <- "HSC, female-biased"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-6"] <- "HSC, male-biased"

# myeloid
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-3"] <- "Early granu prog, male-biased"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-1"] <- "Neutro prog"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-4"] <- "Neutro prog, male-biased"

# antigen-presenting
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-9"] <- "Antigen-presenting and lymphoid"

# mk ery
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-7"] <- "Early Mk prog"

seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-0"] <- "Ery prog, female-biased"
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-2"] <- "Ery prog, male-biased"
seu_ts_bone_marrow$custom_label[
  seu_ts_bone_marrow$clusters == "cluster-8"] <- "Late ery prog, male-biased"

# for this dataset, the Mk and Ery lineage cells are "earlier" than the
# myeloid cells in how I determined the levels of the original cell types
seu_ts_bone_marrow$custom_label <- factor(
  seu_ts_bone_marrow$custom_label,
  levels = c(
    "HSC, female-biased",
    "HSC, male-biased",
    "Early Mk prog",
    "Ery prog, female-biased",
    "Ery prog, male-biased",
    "Late ery prog, male-biased",
    "Early granu prog, male-biased",
    "Neutro prog, male-biased",
    "Neutro prog",
     "Antigen-presenting and lymphoid"
  ))

stopifnot(!is.na(seu_ts_bone_marrow$custom_label))
```

```{r unify}

# unify some annotations by cell type in cases where cell types were split
# by high resolution clustering or by batch
# it is not always super obvious what can be unified and what cannot

seu_ts_bone_marrow$custom_label_unif <- unfactor(seu_ts_bone_marrow$custom_label)

# the two HSC clusters are both clearly HSCs
seu_ts_bone_marrow$custom_label_unif[
  seu_ts_bone_marrow$custom_label == "HSC, female-biased"] <- "HSC"
seu_ts_bone_marrow$custom_label_unif[
  seu_ts_bone_marrow$custom_label == "HSC, male-biased"] <- "HSC"

# same with the two earlier erythroid clusters
seu_ts_bone_marrow$custom_label_unif[
  seu_ts_bone_marrow$custom_label == "Ery prog, female-biased"] <- "Early ery prog"
seu_ts_bone_marrow$custom_label_unif[
  seu_ts_bone_marrow$custom_label == "Ery prog, male-biased"] <- "Early ery prog"

# same with the two neutro clusters, one is already mixed
seu_ts_bone_marrow$custom_label_unif[
  seu_ts_bone_marrow$custom_label == "Neutro prog, male-biased"] <- "Neutro prog"

seu_ts_bone_marrow$custom_label_unif <- factor(
  seu_ts_bone_marrow$custom_label_unif,
  levels = c(
    "HSC",
    "Early granu prog, male-biased",
    "Neutro prog",
    "Antigen-presenting and lymphoid",
    "Early Mk prog",
    "Early ery prog",
    "Late ery prog, male-biased"
  ))

stopifnot(!is.na(seu_ts_bone_marrow$custom_label_unif))
```

#### One last plot to check

```{r slightly_better_markers}

# human genes
# taken from supp table 1 of the corresponding publications
# for genes without reference mentioned here
slightly_better_markers <- c(
    "CD34",
    "MPL",
    "MEIS1",
    "MECOM",
    "MLLT3",
    "SOCS2",
    "HOPX", # stemness https://doi.org/10.1038/s41388-020-1340-2
    "CDKN1C",
    "HLF",
    "MMRN1",
    "MSI2",
    "PROM1", # HSC https://doi.org/10.3389/fphys.2022.1009160
    "ALCAM", 
    "CD38", # MPP, progenitors, activation https://doi.org/10.3389/fphys.2022.1009160
    "SPI1", 
    "CEBPA", 
    "CTSG",
    "MPO",
    "PRTN3",
    "LMO4", 
    "CPA3", 
    "GATA2", 
    "HDC", 
    "MS4A2",
    "CEBPE", 
    "ELANE", 
    "AZU1", # neutros https://doi.org/10.1002/1873-3468.14183
    "S100A8",
    "S100A9",
    "LYZ", # lysozyme
    "CCR2", 
    "MS4A6A",
    "IRF8",
    "ITGAX", # DCs https://doi.org/10.3389/fimmu.2025.1556992
    "FLT3", 
    "IL7R",
    "SATB1", 
    "CD79B", 
    "DNTT",
    "JCHAIN", # BC differentiation 
    "CD74", # antigen-presenting
    "HLA-DPA1", # MHCII
    "HLA-DRA", # MHCII
    "ITGA2B",
    "PF4",
    "GP9", # Mk https://doi.org/10.1038/ncomms11208
    "PLEK", # Mk https://doi.org/10.1182/blood-2006-08-038901
    "VWF", # Mk https://doi.org/10.1038/ncomms11208
    "PBX1",
    "ZFPM1", 
    "GATA1",
    "BLVRB", 
    "KLF1",
    "EPOR",
    "HBA1", # hemoglobins
    "HBB", 
    "HBA2",
    "GYPA", # Ery https://doi.org/10.1016/j.exphem.2023.05.001
    "HELLS", 
    "LIG1",
    "MCM6", 
    "CENPA", 
    "MKI67"
  )

slightly_better_markers[!slightly_better_markers %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% slightly_better_markers]
```

```{r last_heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_bone_marrow,
  assays = "RNA",
  group.by = c("custom_label")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_bone_marrow$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_ts_df_genes$custom_label <- factor(
  agg_ts_df_genes$custom_label, 
  levels = levels(seu_ts_bone_marrow$custom_label))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

```{r last_heatmap2, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_bone_marrow,
  assays = "RNA",
  group.by = c("custom_label_unif")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_bone_marrow$custom_label_unif),
    names_to = "custom_label_unif",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_ts_df_genes$custom_label_unif <- factor(
  agg_ts_df_genes$custom_label_unif, 
  levels = levels(seu_ts_bone_marrow$custom_label_unif))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = custom_label_unif,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

```{r umaps_last, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_bone_marrow, reduction = "umap", group.by = "custom_label")
``` 

Save the object now.

```{r save}
saveRDS(seu_ts_bone_marrow, output_path)
```

```{r sessioninfo}
utils::sessionInfo()
```
