---
title: "Weinreb et al mouse HSPC custom reclustering"
author: "Lea Wölbert"
date: '2025-12-04'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r load_objects}

# the un-clustered original object
seu_mus_weinreb_hspc <- readRDS(snakemake@input[["seu_input"]])
#seu_mus_weinreb_hspc <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/mus_weinreb_hspc")
```

```{r output_path}
# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/03_custom_clustering/mus_weinreb_hspc.Rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

Just try lots of different ones.

```{r parameters}
nr_k <- 30
resolution <- 0.35
```

Use the seurat pipeline roughly taken from sce_functions_reclustering.R
The seurat object should be pretty basic.

For comparability between signature-reclustered and custom-reclustered
partitions, I will not perform batch correction here, either.

Keeping it very basic here.

```{r seurat_clustering_pipeline}

# seu_mus_weinreb_hspc uses raw data so need to normalize first

seu_mus_weinreb_hspc@misc$data_use 

seu_mus_weinreb_hspc <- Seurat::NormalizeData( 
      seu_mus_weinreb_hspc,
      assay = "RNA", 
      verbose = FALSE)

seu_mus_weinreb_hspc <- Seurat::ScaleData(
  seu_mus_weinreb_hspc, 
  verbose = FALSE)

# find variable features
seu_mus_weinreb_hspc <- Seurat::FindVariableFeatures(
  seu_mus_weinreb_hspc
)

# uses scaled data
seu_mus_weinreb_hspc <- Seurat::RunPCA(
  seu_mus_weinreb_hspc,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(seu_mus_weinreb_hspc@reductions$pca)
print(nr_pca)

# requires PC coordinates
seu_mus_weinreb_hspc <- Seurat::FindNeighbors(
  seu_mus_weinreb_hspc, 
  k.param = nr_k,
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
seu_mus_weinreb_hspc <- Seurat::FindClusters(
  seu_mus_weinreb_hspc, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
seu_mus_weinreb_hspc <-  Seurat::RunUMAP(seu_mus_weinreb_hspc, dims = 1:10)
```

Shooting for about 12 clusters as this is the number of clusters when reclustering
with signature genes.

# Evaluate

```{r table}
table(seu_mus_weinreb_hspc@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible batch effects but I can't do much about them.

```{r umaps_base, fig.height = 6, fig.width = 6}
DimPlot(seu_mus_weinreb_hspc, reduction = "umap")
``` 

```{r umaps_tp, fig.height = 6, fig.width = 6}
DimPlot(seu_mus_weinreb_hspc, reduction = "umap")+
  facet_wrap(facets = vars(seu_mus_weinreb_hspc@meta.data$Time.point), ncol = 2)
``` 

There is some strong effect by time point.

## Gene expression

Now check the quality of cell type separation based on expressed genes.

```{r marker_genes}

  marker_genes <- c(
    "Procr", # HSC, Stem 
    "Cdkn1c",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Alcam", 
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Pbx1",
    "Msi2",
    "Adgrl4", # MPP, progenitors, activation
    "Cd34",
    "Cdk4",
    "Sox4",
    "Spi1", 
    "Cebpa", # Myeloid
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Elane", # Neutro
    "Cebpe", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", # Mono
    "Irf8",
    "Ms4a6c",
    "Flt3", # Lymphoid
    "Satb1",
    "Dntt",
    "Cd79b", 
    "Il7r",
    "Rag1",
    "Cpa3", # Mast
    "Hdc", 
    "Ms4a2",
    "Lmo4", # Baso
    "Gata2", # also Mk, Eo Baso and HSC
    "Itga2b", # Mk
    "Pf4",
    "Plek",
    "Cd9",
    "Blvrb", # Mk Ery
    "Gata1",
    "Zfpm1",
    "Epor", # Ery
    "Hba-a1",
    "Gypa",
    "Klf1",
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa", # CC
    "Lockd"
  )

```

#### Heatmap

To get an overview of the rough expected populations based on well-known genes.

```{r heatmap, fig.height = 10, fig.width = 5}

seu_mus_weinreb_hspc@meta.data$clusters <- base::paste0(
  "cluster-",
  seu_mus_weinreb_hspc@meta.data$seurat_clusters
)

# get average logcounts per cluster
agg_mus_weinreb_hspc <- Seurat::AverageExpression(
  seu_mus_weinreb_hspc,
  assays = "RNA",
  group.by = c("clusters")) 

# turn into df formatted for visualisation
agg_mus_weinreb_hspc_df <- agg_mus_weinreb_hspc$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(seu_mus_weinreb_hspc$clusters),
    names_to = "clusters",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_mus_weinreb_hspc_df_genes <- agg_mus_weinreb_hspc_df %>%
  dplyr::filter(genes %in% marker_genes)

print(unique(agg_mus_weinreb_hspc_df_genes$genes))

# factorise for nicer plot
agg_mus_weinreb_hspc_df_genes$genes <- factor(
  agg_mus_weinreb_hspc_df_genes$genes,
  levels = rev(marker_genes))

# scale expression per gene using z-scores
agg_mus_weinreb_hspc_df_genes$zscore_gene <- vector(
  length = nrow(agg_mus_weinreb_hspc_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_mus_weinreb_hspc_df_genes$genes)){
  
  expr_vector_g <- agg_mus_weinreb_hspc_df_genes[
    agg_mus_weinreb_hspc_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_mus_weinreb_hspc_df_genes$zscore_gene[
    agg_mus_weinreb_hspc_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_mus_weinreb_hspc_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

#### UMAPs

```{r umaps_genes, fig.width = 4.5, fig.height = 4}
plotlist <- lapply(as.list(marker_genes), function(gene){
  print(gene)

  plot <- FeaturePlot(seu_mus_weinreb_hspc, reduction = "umap", gene)
  return(plot)
})
names(plotlist) <- marker_genes

plotlist
```

## Markers

```{r markers}
# replace the original idents with the new clusters 

Seurat::Idents(seu_mus_weinreb_hspc) <- seu_mus_weinreb_hspc@meta.data$clusters
clustlist <- as.list(base::sort(base::unique(seu_mus_weinreb_hspc@meta.data$clusters)))

cluster_markers <- lapply(clustlist, function(x){
  
  print(x)
  markers <- Seurat::FindMarkers(
    seu_mus_weinreb_hspc,
    slot = "data",
    test.use = "wilcox", 
    ident.1 = x)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
names(cluster_markers) <- base::sort(base::unique(seu_mus_weinreb_hspc@meta.data$clusters))
```

Gave provisional names based on the heatmap.

### HSC and early progenitors

```{r markers_hsc_and_mpp}
# only few genes in the heatmap but let's see if it's hscs
cluster_markers[["cluster-12"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# looks early with somne mk
cluster_markers[["cluster-1"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# looks early with some lymphoid
cluster_markers[["cluster-9"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
```

All three clusters are approximately in roughly the "same UMAP space".
But each is predominantly detected in only of of the three time points.
They are also of vastly different numbers of cells.

Cluster 12 is tiny and only detected in time point 16 unless I'm mistaken.
It actually highly expresses classic Antigen-presenting markers so it may 
in fact be closer to that than HSCs. 
I suspected it may not be HSCs because this cluster only expresses Alcam 
and Procr in the heatmap.

Cluster 1 is the only cluster predodimantly found in time point 2.
It is almost not found in time point 16.
It expresses Mpl, Hlf, Lmo2, Meis1, Pbx1, Msi2, a bit of Mecom.
But also Cd34.
Ly6a, Cd34, Ldha, Ctla3a, Cdk6, Pim2, Adgrl4, Nrgn are main marker genes, 
to me it looks like a mix of HSC and MPP/progenitor genes.

Cluster 9 is predominantly detected in time point 9.
The genes atually look like very early lymphoid (Dntt, Flt3, Runx2, Satb1,
Cd34, Irf8) but with Nrgn. 
Also some cell cycle stuff (Ccnd2 for example).
Some weaker expression of some HSC genes but not in the marker list.

It seems that cells had differentiated at later time points, explaining the 
lack of HSCs at the latest time point.

```{r hsc_umaps_additional, fig.width = 4.5, fig.height = 4}

hsc_additional_genes <- c(
  "Mecom",
  "Ccl5",
  "H2-Eb1",
  "H2-Aa",
  "Cd74",
  "Cst3",
  "Id2",
  "Fscn1",
  "Ccr7",
  "Dlk1",
  "Ctla2a",
  "Ly6a",
  "Ldha",
  "Cdk6",
  "Nrgn",
  "Hmox1",
  "Runx2",
  "Ccnd2",
  "Ly6d",
  "Ccnd1",
  "Cd27"
)

plotlist <- lapply(as.list(hsc_additional_genes), function(gene){
  print(gene)

  plot <- FeaturePlot(seu_mus_weinreb_hspc, reduction = "umap", gene)
  return(plot)
})
names(plotlist) <- hsc_additional_genes

plotlist
```

### Myeloid and Lymphoid

```{r markers_myeloid}

# early myeloid
cluster_markers[["cluster-6"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# neutro
cluster_markers[["cluster-2"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# mono
cluster_markers[["cluster-3"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# baso mast? a bit early?
cluster_markers[["cluster-10"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# lymphoid a bit earlier
cluster_markers[["cluster-8"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# lymphoid a bit later
cluster_markers[["cluster-5"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

```

Cluster 6 has top markers Elane, Prtn3, Mpo, Ctsg, Nkg7 (NK), Srgn,
Calr, Prss57, Clec12 (mono), Gfi1, Cd63 (neutro) indicating a still 
mixed-lineage early myeloid progenitor.

Cluster 2 expresses clearly neutro genes like Nhp, S100a8 and S100a9, Camp,
Retnlg, Wfdc21, Chil3, Lrg1.

Cluster 3 is mono-ish/antigen-present-ish.
Irf8, Ccr1, Mgep1, Ahnak, Ly86 are mono/macro lineages.
Ctss is going into apc direction.
Some genes like Lyz2 etc also look a bit granu-ish.

Cluster 10 is very clearly Eo baso mast. 
Prss34, Cpa3, Gata2 are baso mast, Ccl4 and Ccl3 are eo. 
Fcer1a, Cd63, Itga2b, Cebpa, Hdc, Ier3 support this.

Cluster 8 is lymphoid but later than the MPP cluster 9, with Vpreb1, Igll1 etc.

Cluster 5 is only detected in time point 16, the latest time point?
It expresses more Ig genes and already some antigen-presenting genes, so I 
think it's a bit later.

```{r myeloid_umaps_additional, fig.width = 4.5, fig.height = 4}

myeloid_additional_genes <- c(
  "Nkg7",
  "Srgn",
  "Prss57",
  "Calr",
  "Clec12a",
  "Gfi1",
  "Fcnb",
  "Cd63",
  "Cst7",
  "Anxa3",
  "Ngp",
  "Lcn2",
  "Retnlg",
  "Camp",
  "Chil3",
  "F13a1",
  "Mpeg1",
  "Ahnak",
  "Lyz2",
  "Ctss",
  "Crip1",
  "S100a10",
  "Prss34",
  "Ccl4",
  "Ccl3",
  "Fcer1a",
  "Ier3",
  "Vpreb1",
  "Igll1",
  "Vpreb3",
  "Ebf1",
  "Cd79a",
  "Igkc",
  "Ighm",
  "Iglc2",
  "H2-Aa",
  "Cd19"
)

plotlist <- lapply(as.list(myeloid_additional_genes), function(gene){
  print(gene)

  plot <- FeaturePlot(seu_mus_weinreb_hspc, reduction = "umap", gene)
  return(plot)
})
names(plotlist) <- myeloid_additional_genes

plotlist
```

### Mk and Ery

```{r markers_mkery}

# ery
cluster_markers[["cluster-0"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# ery
cluster_markers[["cluster-4"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# ery
cluster_markers[["cluster-7"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)

# late ery
cluster_markers[["cluster-11"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
```

Cluster 0 is early-ish Ery? 
No Hb genes yet, but Klf1, Car1, Blvrb, Gata1, and then some Hmbs,
Prdx2, Ermap.

Cluster 4 looks similar. 
Also no Hb genes but Car1, Ermap, Klf1, Car2, Epor.

For cluster 7, things look a bit more mature with Gypa and Hbb-bt. 

Cluster 11 has four Hbs as main markers - I think it looks the latest.

```{r erymk_umaps_additional, fig.width = 4.5, fig.height = 4}

erymk_additional_genes <- c(
  "Car2",
  "Car1",
  "Gata1",
  "Ermap",
  "Hmbs",
  "Alad",
  "Hemgn",
  "Aqp1",
  "Mt2",
  "Mt1",
  "Epor",
  "Hbb-bt",
  "Tfrc",
  "Hba-a2",
  "Hbb-bs",
  "Alas2",
  "Fech"
)

plotlist <- lapply(as.list(erymk_additional_genes), function(gene){
  print(gene)
  
  plot <- FeaturePlot(seu_mus_weinreb_hspc, reduction = "umap", gene)
  return(plot)
})
names(plotlist) <- erymk_additional_genes

plotlist
```

# Assign

```{r assign}
seu_mus_weinreb_hspc$custom_label <- vector(length = ncol(seu_mus_weinreb_hspc))

seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-1"] <- "HSC and MPP"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-9"] <- "Lymphoid MPP"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-6"] <- "Early myeloid"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-2"] <- "Neutro"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-3"] <- "Mono"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-10"] <- "Eo/Baso/Mast/Mk"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-8"] <- "Lymphoid"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-5"] <- "Later lymphoid, time point 16"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-12"] <- "Antigen-presenting"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-0"] <- "Early ery 1"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-4"] <- "Early ery 2"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-7"] <- "Erythroid"
seu_mus_weinreb_hspc$custom_label[
  seu_mus_weinreb_hspc$clusters == "cluster-11"] <- "Late ery"

seu_mus_weinreb_hspc$custom_label <- factor(
  seu_mus_weinreb_hspc$custom_label,
  levels = c(
    "HSC and MPP",
    "Lymphoid MPP",
    "Early myeloid",
    "Neutro",
    "Mono",
    "Eo/Baso/Mast/Mk",
    "Lymphoid",
    "Later lymphoid, time point 16",
    "Antigen-presenting",
    "Early ery 1",
    "Early ery 2",
    "Erythroid",
    "Late ery"
  ))

stopifnot(!is.na(seu_mus_weinreb_hspc$custom_label))
```

```{r unify}

# unify some annotations by cell type in cases where cell types were split
# by high resolution clustering or by batch
# it is not always super obvious what can be unified and what cannot

seu_mus_weinreb_hspc$custom_label_unif <- unfactor(seu_mus_weinreb_hspc$custom_label)

# the two early Ery overlap quite significantly but I think the "Erthroid"
# cluster already looks more mature compared to those
seu_mus_weinreb_hspc$custom_label_unif[
  seu_mus_weinreb_hspc$custom_label == "Early ery 1"] <- "Early ery"
seu_mus_weinreb_hspc$custom_label_unif[
  seu_mus_weinreb_hspc$custom_label == "Early ery 2"] <- "Early ery"

seu_mus_weinreb_hspc$custom_label_unif <- factor(
  seu_mus_weinreb_hspc$custom_label_unif,
  levels = c(
    "HSC and MPP",
    "Lymphoid MPP",
    "Early myeloid",
    "Neutro",
    "Mono",
    "Eo/Baso/Mast/Mk",
    "Lymphoid",
    "Later lymphoid, time point 16",
    "Antigen-presenting",
    "Early ery",
    "Erythroid",
    "Late ery"
  ))

stopifnot(!is.na(seu_mus_weinreb_hspc$custom_label_unif))

```

#### One last plot to check

```{r slightly_better_markers}

# these must be human genes
# try straight up with the same genes as for mouse first
# but then alter as necessary/possible
slightly_better_markers <- c(
    "Mpl",
    "Mecom",
    "Meis1",
    "Ctla2a",
    "Ly6a",
    "Nrgn",
    "Msi2",
    "Hlf",
    "Adgrl4",
    "Cd34",
    "Cdk6",
    "Cdk4",
    "Ccnd1",
    "Runx2",
    "Flt3",
    "Sox4",
    "Satb1",
    "Tcf4",
    "Spi1",
    "Ctsg",
    "Prss57",
    "Mpo",
    "Prtn3",
    "Elane",
    "Cebpe",
    "S100a8",
    "S100a9",
    "Retnlg",
    "Ngp",
    "Irf8",
    "Ccr2",
    "Ms4a6c",
    "F13a1",
    "Cpa3",
    "Hdc",
    "Ms4a2",
    "Gata2",
    "Ccl4",
    "Ccl3",
    "Dntt1",
    "Vpreb1",
    "Igll1",
    "Vpreb3",
    "Il7r",
    "Cd79a",
    "Cd79b",
    "Igkc",
    "Ighm",
    "Iglc2",
    "Cd74",
    "Ccl5",
    "H2-Eb1",
    "H2-Aa",
    "Fscn1",
    "Ccr7",
    "Ctss",
    "Pf4",
    "Plek",
    "Itga2b",
    "Cd9",
    "Zfpm1",
    "Blvrb",
    "Klf1",
    "Ermap",
    "Epor",
    "Tspo2",
    "Hba-a1",
    "Hbb-bt",
    "Gypa",
    "Alas2"
  )

slightly_better_markers[duplicated(slightly_better_markers)]
```

```{r last_heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_mus_weinreb_hspc <- Seurat::AverageExpression(
  seu_mus_weinreb_hspc,
  assays = "RNA",
  group.by = c("custom_label")) 

# turn into df formatted for visualisation
agg_mus_weinreb_hspc_df <- agg_mus_weinreb_hspc$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(seu_mus_weinreb_hspc$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_mus_weinreb_hspc_df_genes <- agg_mus_weinreb_hspc_df %>%
  dplyr::filter(genes %in% slightly_better_markers)

print(unique(agg_mus_weinreb_hspc_df_genes$genes))

# factorise for nicer plot
agg_mus_weinreb_hspc_df_genes$genes <- factor(
  agg_mus_weinreb_hspc_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_mus_weinreb_hspc_df_genes$custom_label <- factor(
  agg_mus_weinreb_hspc_df_genes$custom_label, 
  levels = levels(seu_mus_weinreb_hspc$custom_label))

# scale expression per gene using z-scores
agg_mus_weinreb_hspc_df_genes$zscore_gene <- vector(
  length = nrow(agg_mus_weinreb_hspc_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_mus_weinreb_hspc_df_genes$genes)){
  
  expr_vector_g <- agg_mus_weinreb_hspc_df_genes[
    agg_mus_weinreb_hspc_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_mus_weinreb_hspc_df_genes$zscore_gene[
    agg_mus_weinreb_hspc_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_mus_weinreb_hspc_df_genes,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )
```

```{r last_heatmap2, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_mus_weinreb_hspc <- Seurat::AverageExpression(
  seu_mus_weinreb_hspc,
  assays = "RNA",
  group.by = c("custom_label_unif")) 

# turn into df formatted for visualisation
agg_mus_weinreb_hspc_df <- agg_mus_weinreb_hspc$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(seu_mus_weinreb_hspc$custom_label_unif),
    names_to = "custom_label_unif",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_mus_weinreb_hspc_df_genes <- agg_mus_weinreb_hspc_df %>%
  dplyr::filter(genes %in% slightly_better_markers)

print(unique(agg_mus_weinreb_hspc_df_genes$genes))

# factorise for nicer plot
agg_mus_weinreb_hspc_df_genes$genes <- factor(
  agg_mus_weinreb_hspc_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_mus_weinreb_hspc_df_genes$custom_label_unif <- factor(
  agg_mus_weinreb_hspc_df_genes$custom_label_unif, 
  levels = levels(seu_mus_weinreb_hspc$custom_label_unif))

# scale expression per gene using z-scores
agg_mus_weinreb_hspc_df_genes$zscore_gene <- vector(
  length = nrow(agg_mus_weinreb_hspc_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_mus_weinreb_hspc_df_genes$genes)){
  
  expr_vector_g <- agg_mus_weinreb_hspc_df_genes[
    agg_mus_weinreb_hspc_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_mus_weinreb_hspc_df_genes$zscore_gene[
    agg_mus_weinreb_hspc_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_mus_weinreb_hspc_df_genes,
  aes(
    x = custom_label_unif,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )
```

Save the object now.

```{r save}
saveRDS(seu_mus_weinreb_hspc, output_path)
```

```{r sessioninfo}
utils::sessionInfo()
```