---
title: "Tabula sapiens all stromal cells custom reclustering"
author: "Lea Wölbert"
date: '2025-12-09'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r load_objects}

# the un-clustered original object
seu_ts_all_stromal <- readRDS(snakemake@input[["seu_input"]])

#seu_ts_all_stromal <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/ts_all_stromal")

# load ensembl df to get the ENS IDs for the genes I need
ensembl_hum <- base::readRDS(snakemake@input[["ensembl_hum"]])
#ensembl_hum <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/ensembl/ensembl_hum")
head(ensembl_hum)
```

```{r output_path}
# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/03_custom_clustering/ts_all_stromal.Rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

Just try lots of different ones.

```{r parameters}
nr_k <- 20
resolution <- 0.5
```

Use the seurat piepline roughly takin from sce_functions_reclustering.R
The seurat object should be pretty basic.

For comparability between signature-reclustered and custom-reclustered
partitions, I will not perform batch correction here, either.

Keeping it very basic here.

```{r seurat_clustering_pipeline}

# seu_ts_all_stromal uses raw data so need to normalize first

seu_ts_all_stromal@misc$data_use 

seu_ts_all_stromal <- Seurat::NormalizeData( 
      seu_ts_all_stromal,
      assay = "RNA", 
      verbose = FALSE)

seu_ts_all_stromal <- Seurat::ScaleData(
  seu_ts_all_stromal, 
  verbose = FALSE)

# find variable features
seu_ts_all_stromal <- Seurat::FindVariableFeatures(
  seu_ts_all_stromal
)

# uses scaled data
seu_ts_all_stromal <- Seurat::RunPCA(
  seu_ts_all_stromal,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(seu_ts_all_stromal@reductions$pca)
print(nr_pca)

# requires PC coordinates
seu_ts_all_stromal <- Seurat::FindNeighbors(
  seu_ts_all_stromal, 
  k.param = nr_k,
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
seu_ts_all_stromal <- Seurat::FindClusters(
  seu_ts_all_stromal, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
seu_ts_all_stromal <-  Seurat::RunUMAP(seu_ts_all_stromal, dims = 1:10)
```

# Evaluate

With signature genes there were 13 clusters.

```{r table}
table(seu_ts_all_stromal@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible batch effects but I can't do much about them.

```{r umaps_base, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_all_stromal, reduction = "umap")
``` 

```{r umaps_dev, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_all_stromal, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_all_stromal@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
DimPlot(seu_ts_all_stromal, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_all_stromal@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay, fig.height = 6, fig.width = 6}
DimPlot(seu_ts_all_stromal, reduction = "umap")+
  facet_wrap(facets = vars(seu_ts_all_stromal@meta.data$assay), ncol = 2)
``` 

There is some clear sex and assay-dependent effects, so need to keep in mind.

Check out some single clusters.

```{r umaps_single, fig.height = 6, fig.width = 6}
Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "2"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))

Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "5"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))

Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "10"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))

Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "11"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))

Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "12"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))

Seurat::DimPlot(
  seu_ts_all_stromal[,seu_ts_all_stromal$seurat_clusters == "13"], 
  reduction = "umap")+
  ggplot2::xlim(
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,1])))+
  ggplot2::ylim(  
    limits = c(
      min(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2]),
      max(Seurat::Embeddings(seu_ts_all_stromal@reductions$umap)[,2])))
``` 

## Gene expression

Now check the quality of cell type separation based on expressed genes.

```{r marker_genes}

# these must be human genes
# try straight up with the same genes as for mouse first
# but then alter as necessary/possible
  marker_genes <- c(
    "CXCL12", 
    "KITLG", 
    "LEPR",
    "ALPL",
    "BGLAP",
    "CD200",
    "IBSP",
    "MMP13",
    "POSTN",
    "PTGIS",
    "PTH1R",
    "RUNX2",
    "SP7",
    "SPP1",
    "TNC",
    "ACAN", 
    "ADAMTS5",
    "CHAD",
    "FMOD",
    "MEF2C",
    "MGP",
    "PRELP",
    "PRG4",
    "SOX9",
    "COMP", 
    "S100A10",
    "CD34", 
    "COL3A1",
    "DCN",
    "FN1",
    "LUM",
    "PDGFRA",
    "THY1",
    "CDH5", 
    "PECAM1",
    "CA4", 
    "RGCC",
    "TIMP4",
    "EFNB2",
    "HEY1",
    "SOX17",
    "ACKR1", 
    "EMCN",
    "EPHB4",
    "ICAM1",
    "LRG1",
    "NR2F2",
    "SELP",
    "VCAM1", 
    "FLT4",
    "STAB2",
    "TFPI",
    "CCL21",
    "LYVE1", 
    "PDPN",
    "PROX1",
    "RGS4",
    "RGS5",
    "STEAP4",
    "ACTA2",
    "MYH11", 
    "TAGLN",
    "DES",
    "CD8A", # just some mature immune cell markers for contamination
    "CD4",
    "IL7R",
    "CD79A",
    "IGHM",
    "GZMB",
    "CD74",
    "S100A8",
    "S100A9",
    "CCL5",
    "CPA3", 
    "HDC",
    "GYPA", 
    "HBA1"
  )

```

```{r ensmus_ids}
# not get the ensmus IDs
head(rownames(seu_ts_all_stromal))

marker_genes[!marker_genes %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% marker_genes]
```

#### Heatmap

To get an overview of the rough expected populations based on well-known genes.

```{r heatmap, fig.height = 10, fig.width = 5}

seu_ts_all_stromal@meta.data$clusters <- base::paste0(
  "cluster-",
  seu_ts_all_stromal@meta.data$seurat_clusters
)

# get average logcounts per cluster
agg_ts_bonemarrow <- Seurat::AverageExpression(
  seu_ts_all_stromal,
  assays = "RNA",
  group.by = c("clusters")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_all_stromal$clusters),
    names_to = "clusters",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(marker_genes))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

#### UMAPs

```{r umaps_genes, fig.width = 4.5, fig.height = 4}
plotlist <- lapply(as.list(marker_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  plot <- FeaturePlot(seu_ts_all_stromal, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- marker_genes

plotlist
```

## Markers

```{r markers}
# replace the original idents with the new clusters 

Seurat::Idents(seu_ts_all_stromal) <- seu_ts_all_stromal@meta.data$clusters
clustlist <- as.list(base::sort(base::unique(seu_ts_all_stromal@meta.data$clusters)))

cluster_markers <- lapply(clustlist, function(x){
  
  print(x)
  markers <- Seurat::FindMarkers(
    seu_ts_all_stromal,
    slot = "data",
    test.use = "wilcox", 
    ident.1 = x)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  # add gene names otherwise I cannot interpret
  markers$genes <- vector(length = nrow(markers))
  markers$genes <- ensembl_hum$external_gene_name[
    match(
      rownames(markers),
      ensembl_hum$ensembl_gene_id)]

  return(markers)
})
names(cluster_markers) <- base::sort(base::unique(seu_ts_all_stromal@meta.data$clusters))
```

Gave provisional names based on the heatmap.

### Mesenchymal

```{r markers_hsc}
# adipo/CAR expression
cluster_markers[["cluster-9"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-4"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# osteo expression
cluster_markers[["cluster-0"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-3"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-7"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

cluster_markers[["cluster-2"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# chondro expression
cluster_markers[["cluster-12"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# (chondro) fibro
# contains eo baso mast contamination
cluster_markers[["cluster-10"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

Cluster 4 has very high CXCL12 expression, so by definition they are CAR cells
It is exclusively from 10X V1.
Other classic top markers are VCAM1 and LPL.

Cluster 9 also has very high CXCL12 expression = CAR cells.
It is exclusively from 10X V2.
Other top markers like VCAM1, LPL, PAPPA also make sense in an 
adipo-CAR-context.

Both also express KITLG, LEPR and a bit of ALPL.

Clusters 0, 3 and 12 are one group (with Cluster 4) and are again only 10X V1.
Cluster 0 expresses ALPL, CD200, IBSP, MMP13, SP7, but not BGLAP.
Top markers are TNC, SPP1, LIMCH1, ALPL, so these are early Osteo-CAR-ish.

Cluster 3 has high osteo markers like BGLAP but also chondro marker CHAD.
But no SOX9 or ACAN. 
Top markers are SPARC, IBSP, GGN, SPP1, I think these are osteo-early-chondro
mixed.

Cluster 12 is quite small and also belongs to the "next group" in the UMAP.
COL2A1, COL9A3, SNORC, MGP, COMP, FMOD, ACAB and PRELP in the top marker
list make it very chondro.

Another group of osteo-ish cells are clusters 2 and 7 of which 2 us only 
really found in 10X V2. 

Cluster 7 is cycling. 
The top markers look cycling exclusive but the umaps show ALPL, ISBP, MMP13,
SP7, SPP1, TNC, ACAN expression so I'm going with cycling osteo-early-chondro.
Cluster 2 is low quality or batch + neutrophil contamonation in the top markers.
Maybe it's just a mix of a lot of stuff, but the expression of osteo-markers is 
alsoseen in the UMAPs. 
It seems only one little part of cluster 2 is contaminatin that dominates the marker
gene list, but the rest do look osteo.
High IBSP, MMP13, RUNX2, SPP1.

Another group of Fibro/Chondro/PRG4+ cells is cluster 10.
It looks slightly different from the other chondro group, with high POSTN,
OGN, ASPN, S100A6, S100a5, PRG6, COL3A1, THBS4, CREB5, CRIP1, SEMA3C which
looks fibro-ish articular chondro/synovial fibro like, maybe a mix of prognenitors for
those populations.

```{r mes_umaps_additional, fig.width = 4.5, fig.height = 4}

mes_additional_genes <- c(
  "PPARG",
  "PAPPA",
  "VCAM1",
  "LPL",
  "ESM1",
  "HES1",
  "PDGFRB",
  "SNAI2",
  "CRABP1",
  "EGR1",
  "APOE",
  "CLEC11A",
  "SCARA3",
  "LIMCH1",
  "TGFBI",
  "COL1A2",
  "DPT",
  "SPARC",
  "SERPINF1",
  "BGN",
  "BAMBI",
  "TOP2A",
  "MKI67",
  "CENPF",
  "RPS17",
  "RPS10",
  "LYZ",
  "MT-ND2",
  "CPE",
  "MPO",
  "COL2A1",
  "COL9A3",
  "COL10A1",
  "SNORC",
  "DLK1",
  "COMP",
  "FMOD",
  "OGN",
  "ASPN",
  "S100A6",
  "S100A4",
  "PRG4",
  "THBS4",
  "SEMA3C",
  "CRIP1",
  "CREB5"
)

plotlist <- lapply(as.list(mes_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  print(gene_id)

  plot <- FeaturePlot(seu_ts_all_stromal, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- mes_additional_genes

plotlist
```

### Endothelial cells and Peri/SMC

```{r markers_ec}

# more venous
cluster_markers[["cluster-13"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
  
# more capillary?
cluster_markers[["cluster-5"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# a bit smeary
cluster_markers[["cluster-8"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# a bit smeary
cluster_markers[["cluster-6"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# a bit smeary
cluster_markers[["cluster-1"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# peri SMC and also kind of fibro
# contains neutro contamination
cluster_markers[["cluster-11"]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

All clusters 1, 8, 5 in one group (10X V2) and 6 (10X v1) are clearly ECs.
The mouse capillary markers CA4, TIMP4 and RGCC look kind of useless.
Sinusoidal marker STAB2 is highest in clusters 1, 5 and 6 but missing from
clusters 13 and 5 in the UMAP.

Cluster 5 has top markers PVLAP, CD34, PECAM1, ESAM, AQP1, PODXL. 
They look fairly standard to me, with RGCC and PLVAP indicating something
more capillary-ish. 
Cluster 1 looks slightly different from the other two following sinusoidal ECs
with STAB1, STAB2, ENG, KDT for some sinusoidal identity but also some fatty acid 
binding stuff, RGCC, and slightly immuno/activated looking.
Cluster 8 has top markers inclusing STAB2, CD36, FLT1, LYVE1, KDR, so I'm going
with sinusoidal.
Same with cluster 6, high NR2F1, LYVE1, FLT1, KDR. 
Also some batch/low quality-ish look with lots of R-proteins.

Cluster 13 is just a handful of cells interspersed into the V1 EC group.
The top markers are LYVE1, DNASE1L3, FLT1, STAB2, all CLEARLY related to 
sinusoidal ECs?
Even KDR and VWV and TFPI. 

Regarding Peri/SMCs, the classic genes from the heatmap are barely expressed.
In fact, the first few markers look very neutrophil-ish to me.
LAter, some NOTCH3, THY1, MYO1E, FN1, CSPG4, PDGFRB, OGN also give SOME muscle or 
mesenchymal idea.

```{r ec_umaps_additional, fig.width = 4.5, fig.height = 4}

ec_additional_genes <- c(
  "DNASE1L3",
  "FLT1",
  "KDR",
  "VWF",
  "TFPI",
  "PLVAP",
  "PODXL",
  "AQP1",
  "ESAM",
  "HSPA1A",
  "PLIN2",
  "UGCG",
  "CD36",
  "BACH1",
  "IL6ST",
  "NR2F1",
  "HBG1",
  "CTSD",
  "FABP4",
  "NEAT1",
  "STC1",
  "RPS17",
  "CTSL",
  "EGFL7",
  "RNASE1",
  "PIM3",
  "FABP5",
  "SELE",
  "APLNR",
  "MMP9",
  "CCL3",
  "MMP14",
  "GPC1",
  "CALD1",
  "DLK1",
  "NOTCH3",
  "EDNRA",
  "TGFB3",
  "MYO1E",
  "CSPG4"
)

plotlist <- lapply(as.list(ec_additional_genes), function(gene){
  print(gene)
  
  gene_id <- ensembl_hum$ensembl_gene_id[
    ensembl_hum$external_gene_name == gene]

  #print(gene_id %in% Features(seu_ts_all_stromal))
  plot <- FeaturePlot(seu_ts_all_stromal, reduction = "umap", gene_id)
  return(plot)
})
names(plotlist) <- ec_additional_genes

plotlist
```


# Assign

```{r assign}
seu_ts_all_stromal$custom_label <- vector(length = ncol(seu_ts_all_stromal))

# MES
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-4"] <- "Adipo-CAR, V1"
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-9"] <- "Adipo-CAR, V2"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-0"] <- "Osteo-CAR, V1"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-3"] <- "Osteo-Chondro, V1"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-2"] <- "Osteo-Chondro/Immune contamination, V2"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-7"] <- "Osteo Cycling"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-12"] <- "Chondrocyte"

# I think this is a mix of all fibro/chondro-like looking cells except actual chondros
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-10"] <- "Fibro/chondro/synovial/articular-like"


# EC
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-5"] <- "Arterial EC"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-8"] <- "Sinusoidal EC 1, V1"
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-6"] <- "Sinusoidal EC 2, V1"
seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-1"] <- "Sinusoidal EC 3, V2"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-13"] <- "Sinusoidal EC 4, V1"

seu_ts_all_stromal$custom_label[
  seu_ts_all_stromal$clusters == "cluster-11"] <- "Pericyte/Immune contamination"

seu_ts_all_stromal$custom_label <- factor(
  seu_ts_all_stromal$custom_label,
  levels = c(
    "Adipo-CAR, V1",
    "Adipo-CAR, V2",
    "Osteo-CAR, V1",
    "Osteo Cycling",
    "Osteo-Chondro/Immune contamination, V2",
    "Osteo-Chondro, V1",
    "Chondrocyte",
    "Fibro/chondro/synovial/articular-like",
    "Arterial EC",
    "Sinusoidal EC 1, V1",
    "Sinusoidal EC 2, V1",
    "Sinusoidal EC 3, V2",
    "Sinusoidal EC 4, V1",
    "Pericyte/Immune contamination"
  ))

stopifnot(!is.na(seu_ts_all_stromal$custom_label))
```

```{r unify}

# unify some annotations by cell type in cases where cell types were split
# by high resolution clustering or by batch
# it is not always super obvious what can be unified and what cannot

seu_ts_all_stromal$custom_label_unif <- unfactor(seu_ts_all_stromal$custom_label)

# I think the adipo CARs overlap pretty nicely but this is not the case
# for other mesenchymal cells
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Adipo-CAR, V1"] <- "Adipo-CAR"
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Adipo-CAR, V2"] <- "Adipo-CAR"

# I am merging all sinusoidal ECs
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Sinusoidal EC 1, V1"] <- "Sinusoidal EC"
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Sinusoidal EC 2, V1"] <- "Sinusoidal EC"
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Sinusoidal EC 3, V2"] <- "Sinusoidal EC"
seu_ts_all_stromal$custom_label_unif[
  seu_ts_all_stromal$custom_label == "Sinusoidal EC 4, V1"] <- "Sinusoidal EC"

seu_ts_all_stromal$custom_label_unif <- factor(
  seu_ts_all_stromal$custom_label_unif,
  levels = c(
    "Adipo-CAR",
    "Osteo-CAR, V1",
    "Osteo Cycling",
    "Osteo-Chondro/Immune contamination, V2",
    "Osteo-Chondro, V1",
    "Chondrocyte",
    "Fibro/chondro/synovial/articular-like",
    "Arterial EC",
    "Sinusoidal EC",
    "Pericyte/Immune contamination"
  ))

stopifnot(!is.na(seu_ts_all_stromal$custom_label_unif))

```

#### One last plot to check

```{r slightly_better_markers}

# these must be human genes
# after cghecking plots above
# check the 02/01/ custom clustering of our own datasets for additional sources
# also supplementary table 1 of the corresponding publication
slightly_better_markers <- c(
    "CXCL12",
    "KITLG",
    "LEPR",
    "LPL", # lipoprotein lipase adipocyte/fatty acid related
    "VCAM1", # Adipo-CAR and sinusoid
    "SNAI2", # EMT, adipose tissue mass
    "LIMCH1", # osteo-CAR
    "TNC", 
    "ALPL",
    "IBSP", 
    "SP7",
    "MMP13",
    "RUNX2",
    "SPP1", # osteo lineage, also osteochondric progenitor
    "BGLAP",
    "SPARC", # osteoblastic differentiation
    "BAMBI", # chondrogenic differentiation https://doi.org/10.1155/2016/2685147
    "CHAD", # both osteo and chondro https://doi.org/10.1038/s41598-020-67998-0  https://doi.org/10.1371/annotation/cb7d7b4c-624f-46e6-957e-13b355ca8e02
    "MGP",
    "COMP",
    "FMOD",
    "SOX9",
    "SNORC", # specific to chondrocytes
    "PRELP",
    "POSTN",
    "OGN", # osteo and cartilage
    "PRG4",
    "CREB5",
    "CRIP1",
    "ASPN",
    "S100A6", # chondro, fibro
    "S100A4", # fibro-specific 
    "THBS4", # fibro mifration
    "SEMA3C", # several fibroblast type
    "PDGFRA",
    "DCN",
    "DES",
    "CD34", # also AEC
    "PECAM1",
    "CDH5",
    "AQP1", # EC migration
    "EFNB2",
    "SOX17",
    "BCAM", # EC
    "IGFBP7", # EC recovery, attachment, .. activity
    "RGCC",
    "PLVAP",
    "APLNR", # vascular remodelling
    "STAB1",
    "CD36", # FA 
    "KDR",
    "STAB2",
    "FLT1",
    "DNASE1L3",
    "LYVE1", # sometimes also SEC not just LEC
    "VWF", # EC subsets, venous
    "MMP14", # can be osteo
    "MYO1E", # non-muscle myosin
    "MYH11", # smooth muscle
    "NOTCH3", # osteo
    "EDNRA", # peri and smc
    "CSPG4", # pericytes
    "RGS5", # SMC/Peri
    "CALD1", # pericytes
    "ACTA2", # pericytes
    "TAGLN", # pericytes
    "MYOCD", # smooth muscle
    "CPA3", # cont, baso mast
    "HDC",
    "CCL5",    
    "GZMB", # cont, granu
    "S100A8", # cont, neutro
    "S100A9",
    "LYZ",
    "HBG2", # cont, ery
    "MMP9",
    "CCL3",
    "CENPA", # cycling
    "MKI67",
    "TOP2A"
  )

slightly_better_markers[!slightly_better_markers %in% ensembl_hum$external_gene_name]
gene_ids_keep <- ensembl_hum$ensembl_gene_id[
  ensembl_hum$external_gene_name %in% slightly_better_markers]
```

```{r last_heatmap, fig.height = 15, fig.width = 5}

# get average logcounts per cluster
agg_ts_all_stromal <- Seurat::AverageExpression(
  seu_ts_all_stromal,
  assays = "RNA",
  group.by = c("custom_label")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_all_stromal$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_all_stromal$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_ts_df_genes$custom_label <- factor(
  agg_ts_df_genes$custom_label, 
  levels = levels(seu_ts_all_stromal$custom_label))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

```{r last_heatmap_unif, fig.height = 15, fig.width = 5}

# get average logcounts per cluster
agg_ts_all_stromal <- Seurat::AverageExpression(
  seu_ts_all_stromal,
  assays = "RNA",
  group.by = c("custom_label_unif")) 

# turn into df formatted for visualisation
agg_ts_df <- agg_ts_all_stromal$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "gene_ids") %>%
  tidyr::pivot_longer(
    cols = unique(seu_ts_all_stromal$custom_label_unif),
    names_to = "custom_label_unif",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_ts_df_genes <- agg_ts_df %>%
  dplyr::filter(gene_ids %in% gene_ids_keep)

agg_ts_df_genes$genes <- vector(length = nrow(agg_ts_df_genes))
agg_ts_df_genes$genes <- ensembl_hum$external_gene_name[
  match(
    agg_ts_df_genes$gene_ids,
    ensembl_hum$ensembl_gene_id
  )]

print(unique(agg_ts_df_genes$genes))

# factorise for nicer plot
agg_ts_df_genes$genes <- factor(
  agg_ts_df_genes$genes,
  levels = rev(slightly_better_markers))

agg_ts_df_genes$custom_label_unif <- factor(
  agg_ts_df_genes$custom_label_unif, 
  levels = levels(seu_ts_all_stromal$custom_label_unif))

# scale expression per gene using z-scores
agg_ts_df_genes$zscore_gene <- vector(
  length = nrow(agg_ts_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_ts_df_genes$genes)){
  
  expr_vector_g <- agg_ts_df_genes[agg_ts_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_ts_df_genes$zscore_gene[agg_ts_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_ts_df_genes,
  aes(
    x = custom_label_unif,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(angle = 90)
  )

```

Save the object now.

```{r save}
saveRDS(seu_ts_all_stromal, output_path)
```

```{r sessioninfo}
utils::sessionInfo()
```
