---
title: "Tabula muris bone marrow custom reclustering"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

General build taken from custom_mmus_hsc.Rmd etc. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r load_objects}

seu_tm_bonemarrow <- readRDS(snakemake@input[["seu_input"]])

#seu_tm_bonemarrow <- readRDS("/home/l012t/data/metadata/scRNAseq/03_sce_analysis/reclustering_bm/prepared/mus_tm_bonemarrow")

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
#output_path <- base::paste0(base_path_data, "")
print(output_path)

```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:40])
names(col_num) <- as.character(c(1:40))
```

# Clustering

### Define parameters

Just try lots of different ones.

```{r parameters}
nr_k <- 20
resolution <- 0.6
```

Use the seurat piepline roughly takin from sce_functions_reclustering.R
The seurat object should be pretty basic.

For comparability between signature-reclustered and custom-reclustered
partitions, I will not perform batch correction here, either.

Keeping it very basic here.

```{r seurat_clustering_pipeline}

# seu_tm_bonemarrow uses raw_data so need to normalize first

seu_tm_bonemarrow@misc$data_use 

seu_tm_bonemarrow <- Seurat::NormalizeData( 
      seu_tm_bonemarrow,
      assay = "RNA", 
      verbose = FALSE)

seu_tm_bonemarrow <- Seurat::ScaleData(
  seu_tm_bonemarrow, 
  verbose = FALSE)

# find variable features
seu_tm_bonemarrow <- Seurat::FindVariableFeatures(
  seu_tm_bonemarrow
)

# uses scaled data
seu_tm_bonemarrow <- Seurat::RunPCA(
  seu_tm_bonemarrow,
  npcs = 30,
  verbose = FALSE)
  
# check and possibly define nr of pcs
nr_pca <- ncol(seu_tm_bonemarrow@reductions$pca)
print(nr_pca)

# requires PC coordinates
seu_tm_bonemarrow <- Seurat::FindNeighbors(
  seu_tm_bonemarrow, 
  k.param = nr_k,
  reduction = "pca",
  dims = 1:nr_pca, 
  verbose = FALSE)
  
# requires neighbors
seu_tm_bonemarrow <- Seurat::FindClusters(
  seu_tm_bonemarrow, 
  resolution = as.numeric(resolution), 
  verbose = FALSE) 

# calculate UMAP
seu_tm_bonemarrow <-  Seurat::RunUMAP(seu_tm_bonemarrow, dims = 1:10)
```

# Evaluate

```{r table}
table(seu_tm_bonemarrow@meta.data$seurat_clusters)
```

## Clustering separation

Look at the clusters.
Check possible batch effects but I can't do much about them.

```{r umaps_base, fig.height = 6, fig.width = 6}
DimPlot(seu_tm_bonemarrow, reduction = "umap")
``` 

```{r umaps_base, fig.height = 9, fig.width = 6}
DimPlot(seu_tm_bonemarrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_tm_bonemarrow@meta.data$mouse.id), ncol = 2)
``` 

```{r umaps_base, fig.height = 9, fig.width = 6}
DimPlot(seu_tm_bonemarrow, reduction = "umap")+
  facet_wrap(facets = vars(seu_tm_bonemarrow@meta.data$mouse.sex), ncol = 2)
``` 

There is some very strong effect by sex.

## Gene expression

Now check the quality of cell type separation based on expressed genes.

```{r marker_genes}

  marker_genes <- c(
    "Aldh1a1", # HSC, Stem
    "Procr",
    "Cdkn1c",
    "Gng11",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Alcam", 
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Socs2",
    "Pbx1",
    "Slamf1",
    "Txnip",
    "Msi2",
    "Adgrl4", # MPP, progenitors, activation
    "Cd34", 
    "Cdk4",
    "Sox4",
    "Spi1", 
    "Cebpa", # Myeloid
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Elane", # Neutro
    "Cebpe", 
    "Gfi1", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", # Mono
    "Irf8",
    "Ly86",
    "Ms4a6c",
    "Flt3", # Lymphoid
    "Satb1",
    "Dntt",
    "Cd79b", 
    "Il7r",
    "Rag1",
    "Cpa3", # Mast
    "Hdc", 
    "Ms4a2",
    "Lmo4", # Baso
    "Mcpt8", 
    "Prss34",
    "Gata2", # also Mk, Eo Baso and HSC
    "Itga2b", # Mk
    "Pf4",
    "Blvrb", # Mk Ery
    "Gata1",
    "Zfpm1",
    "Epor", # Ery
    "Gfi1b",
    "Hba-a1",
    "Klf1",
    "Dctpp1", # CC entry
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa", # CC
    "Lockd", 
    "Mki67"
  )

```

#### Heatmap

To get an overview of the rough expected populations based on well-known genes.

```{r}
assays <- GetAssayData(seu_tm_bonemarrow)

any(is.infinite(assays))
any(is.nan(assays))

```

```{r heatmap, fig.height = 10, fig.width = 5}

seu_tm_bonemarrow@meta.data$clusters <- base::paste0(
  "cluster-",
  seu_tm_bonemarrow@meta.data$seurat_clusters
)

# get average logcounts per cluster
agg_tm_bonemarrow <- Seurat::AverageExpression(
  seu_tm_bonemarrow,
  assays = "RNA",
  group.by = c("clusters", "mouse.id")) # MACH HIER WEITER!!!

colnames(agg_tm_bonemarrow$RNA)

# turn into df formatted for visualisation
agg_tm_df <- agg_tm_bonemarrow$RNA %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(seu_tm_bonemarrow$clusters),
    names_to = "clusters",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_tm_df_genes <- agg_tm_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_tm_df_genes$genes <- factor(
  agg_tm_df_genes$genes,
  levels = rev(marker_genes))

# scale expression per gene using z-scores
agg_tm_df_genes$zscore_gene <- vector(
  length = nrow(agg_tm_df_genes), 
  mode = "numeric")

for(gene in base::unique(agg_tm_df_genes$genes)){
  
  expr_vector_g <- agg_tm_df_genes[agg_tm_df_genes$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_tm_df_genes$zscore_gene[agg_tm_df_genes$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_tm_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)

ggplot2::ggplot(
  agg_tm_df_genes,
  aes(
    x = clusters,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::facet_grid(cols = vars)
```
