---
title: "Custom clustering mspr hsc"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
#source("~/repositories/Interspecies_BM_phd/code/source/plotting.R")
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])
#sce_species <- base::readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/01_mnnc/comparison/sce_mspr_hsc_Object_ID-01")

fraction_curr <- "hsc"
species_curr <- "mspr"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be
# defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mspr_hsc_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

```{r parameters}
k_graph <- 8
resolution_louvain <- 0.82
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "hsc"){
  marker_genes <- c(
    "Aldh1a1", # HSC, Stem
    "Procr",
    "Cdkn1c",
    "Gng11",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Alcam", 
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Socs2",
    "Pbx1",
    "Slamf1",
    "Txnip",
    "Msi2",
    "Adgrl4", # MPP, progenitors, activation
    "Cd34", 
    "Cdk4",
    "Sox4",
    "Spi1", 
    "Cebpa", # Myeloid
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Elane", # Neutro
    "Cebpe", 
    "Gfi1", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", # Mono
    "Irf8",
    "Ly86",
    "Ms4a6c",
    "Flt3", # Lymphoid
    "Satb1",
    "Dntt",
    "Cd79b", 
    "Il7r",
    "Rag1",
    "Cpa3", # Mast
    "Hdc", 
    "Ms4a2",
    "Lmo4", # Baso
    "Mcpt8", 
    "Prss34",
    "Gata2", # also Mk, Eo Baso and HSC
    "Itga2b", # Mk
    "Pf4",
    "Blvrb", # Mk Ery
    "Gata1",
    "Zfpm1",
    "Epor", # Ery
    "Gfi1b",
    "Hba-a1",
    "Klf1",
    "Dctpp1", # CC entry
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa", # CC
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}

```

```{r genes_short}

if(fraction_curr == "hsc"){
  marker_genes_short <- c(
    "Alcam", 
    "Cdkn1c",
    "Mecom",
    "Mpl",
    "Meis1",
    "Cd34", 
    "Spi1", 
    "Mpo",
    "Elane",
    "Cebpa", 
    "Cebpe", 
    "S100a9",
    "Wfdc21",
    "Ccr2", 
    "Irf8",
    "Ly86",
    "Cd79b", 
    "Dntt",
    "Il7r",
    "Satb1",
    "Cpa3", 
    "Hdc", 
    "Lmo4", 
    "Prss34",
    "Itga2b", 
    "Pf4",
    "Zfpm1",
    "Epor",
    "Hba-a1",
    "Klf1",
    "Hells",
    "Lig1",
    "Cenpa", 
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}

```

#### Heatmap

Just to get an idea what is going on in each cluster based on the defined
marker genes.
The marker gene list is LIMITED to a small number of genes and may not be able to resolve all cell types.

```{r heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

#### Violin

Slightly downsized but depicting more than just the average per cell type.

```{r violin, fig.height = 20, fig.width = 6}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

## Markers

This is a bit challenging here because there are many small clusters that
look like mixed mature or low-quality populations and also some 
ambiguous clusters.
However, I think that even this fine-grained clustering doesn't separate some 
populations properly, especially some cycling cells from myeloid cells.
So I decided to go ahead with this fine-grained clustering. 
Most of the main clusters (HSC, MPP) stay roughly the same when using different
resolutions or ks.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### HSC

```{r markers_hsc}
cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

```{r hsc_add_plots, fig.width = 4, fig.height = 3}
# adding specific genes from the marker list to visualise in umap
hsc_markers_add <- c(
  "Ramp2",
  "Gas6",
  "Gcnt2",
  "Ldhb"
)
marker_genes_list <- as.list(hsc_markers_add)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- hsc_markers_add
plot_list_gene
```

This cluster reliably and most strongly expresses many BL6 HSC markers.

### HSC/MPP - Mk like populations

I think these clusters look like they still express markers of quiescence
that are also somewhat tied to Mk lineage, but these cells look much more HSC-
like than Mk progenitors do in BL6 or CAST.
For example, cluster 4 expresses a lot of Aldh1a1, Gng11, Pbx1, Slamf1, Meis1,
all of which are not super shocking for Mk but quite strongly compared 
to the more "canonical" HSC cluster.
Cluster 1 also expresses Slamf1 and Txnip, but then again, Slamf1 may just
not be an HSC marker in the exotic species.
Txnip makes sense in Redox state like Ery prog too.

```{r markers_hsc_mk}
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

```{r hsc_mk_add_plots, fig.width = 4, fig.height = 3}
# adding genes for umap from the marker lists for better differentiation
hsc_mk_markers_add <- c(
  "Cavin2",
  "Nupr1",
  "Plek",
  "Cd9",
  "Fli1"
)

hsc_mk_bm_markers_add <- c(
  "Trib2",
  "Ugcg",
  "Pdzph1",
  "Plxnc1",
  "Klf3"
)
marker_genes_list <- as.list(c(hsc_mk_markers_add, hsc_mk_bm_markers_add))

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- c(hsc_mk_markers_add, hsc_mk_bm_markers_add)
plot_list_gene
```

After looking at the actual markers, they really do look much more like
Mk and Mk/Ery progenitor respectively, but especially the Mk cluster
still expresses classic BL6 HSC markers a lot.
The more ery-like cluster also looks a lot like cluster 6 below.

### HSC - progenitor like population

This one seems "very dormant"? 
Cdk4 and Dctpp1 are very low.

```{r markers_hsc_prog}
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# differentiate between similar cluster colors 2 and 18
# that also seem generally low-quality both
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == 2],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == 18],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))
```

This cluster has mostly marker genes that are related to batch 
or low quality, like Tsix, mt genes and other random stuff.
Check the quality:

```{r hsc_prog_qc}
qc_df <- scuttle::perCellQCMetrics(sce_species)
sce_species$log_sum <- log(qc_df$sum)
sce_species$log_detected <- log(qc_df$detected)

umap_cont(sce_species, "log_sum")
umap_cont(sce_species, "log_detected")

qc_df$cluster_louvain_custom <- sce_species$cluster_louvain_custom
qc_df <- as.data.frame(qc_df)

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(sum)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(detected)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()
```

Both cluster 2 and 18 have very low quality compared to other clusters. 
Cluster 19 partially, too.

### MPP and progenitor populations

These seem pretty cut and dry MPP.

```{r markers_mpp}
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
```

They otherwise don't seem too different, so I'll just call them
Mpp cells 1 and 2 (not meaning the defined MPP1 and MPP2 though).

### Myeloid populations

```{r markers_myeloid}
# probably early myeloid
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# early neutro?
cluster_markers[[17]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# late neutro?
cluster_markers[[18]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# mature immune, mixed (looks neutro)
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# late mono, antigen presenting?
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
```

```{r myeloid_add_plots, fig.width = 4, fig.height = 3}
early_neutro_markers_17 <- c(
  "Alas1",
  "Anxa3",
  "Cebpe",
  "Gda"
)

more_neutro_markers_18 <- c(
  "Chil1"
)

mature_neutro_immune_markers_14 <- c(
  "Tcf7",
  "Rora",
  "Ltb",
  "Tox",
  "Bcl2"
)
mature_mono_markers_13 <- c(
  "Cd74",
  "H2-Aa",
  "Ifi30",
  "Cst3",
  "Ctss"
)

marker_genes_list <- as.list(c(
  early_neutro_markers_17,
  more_neutro_markers_18,
  mature_neutro_immune_markers_14,
  mature_mono_markers_13))

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- c(
  early_neutro_markers_17,
  more_neutro_markers_18,
  mature_neutro_immune_markers_14,
  mature_mono_markers_13)
plot_list_gene
```

The progression from early neutro to more mature neutro doesn't
appear clearly in the marker list, but from the expression patterns of
later neutro genes like Cebpe, Gfi1, S100a8, etc. 
In addition, cluster 18 has very low expression levels.
Cluster 14 looks like a variable mix of lymphoid cells (Tcf7, Il74r, Tox) and 
some mature neutros (S100a8).
Cluster 13 strongly expresses marker of antigen presenting cells.

### Lymphoid

```{r markers_lymphoid}
# earlier
cluster_markers[[16]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# later
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
```

These are also pretty cut and dry a progression frome early lymphoid cells
(Flit3, Satb1) to later, more mature B lineage cells (Cd79b)

### Rest of Bm and ery

These are populations not expressing HSC markers anymore. 
As compared to the more HSC-looking Mk/Ery populations above.

```{r markers_ery_mb}

# baso mast
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# Ery

# (earliest) pure Ery?
cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# activated Ery?
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# (early) Ery?
cluster_markers[[19]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)

# cycling Ery?
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
```

```{r ery_add_plots, fig.width = 4, fig.height = 3}
# from the marker genes, 
# these are not black-white different genes but slightly higher or lower
ery_6_markers <- c(
  "Cpa3",
  "Clec4d",
  "Tspo2",
  "Nupr1"
)

ery_9_markers <- c(
  "Blvrb",
  "Klf1",
  "Ephx2",
  "Gfi1b",
  "Ermap",
  "Emc9"
)
marker_genes_list <- as.list(c(
  ery_6_markers,
  ery_9_markers))

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- c(
  ery_6_markers,
  ery_9_markers)
plot_list_gene
```

Baso/Mast are pretty clearly defined, as usual.

The "progression" of or difference between Ery cells is not that clear. 
Cluster 9 looks more CC activated (Hells), and also expresses canonical 
Ery markers more strongly.
Cluster 6 looks a lot like cluster 1, I think these are early, undefined
progenitors of ery, mk and MB cells.
Cluster 19 looks like Ery cells (cluster 9) but in the quality plots, the
lower quality is also very visible.
Cluster 12 is clearly cycling Erys.

### Cycling 

```{r markers_rest}
# cycling mixed and some neutros/monos
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
```

This is a mixed population of cycling cells spanning from MPP to neutro
progenitor.

# Assign

```{r assign}

# hsc
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "HSC"

# mpp
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "MPP cells 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "MPP cells 2"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Low quality, early"

# myeloid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Early myeloid"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "17"] <-  "Neutro progenitor"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "18"] <-  "Low quality, neutro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Mono/antigen-presenting" 
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Late immune, mixed"

# lymphoid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "16"] <-  "Early lymphoid"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Late/Cycling lymphoid"

# baso mast
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Baso/Mast progenitor"

# ery and mk populations
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Early Mk progenitor"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Early ery/baso/mast 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Early ery/baso/mast 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Ery progenitor"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "Cycling ery"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "19"] <-  "Low quality, ery"

# cycling
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Cycling, mostly myeloid"

sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "HSC",
    "MPP cells 1",
    "MPP cells 2",
    "Early myeloid",
    "Neutro progenitor",
    "Cycling, mostly myeloid",
    "Mono/antigen-presenting",
    "Early lymphoid",
    "Late/Cycling lymphoid",
    "Baso/Mast progenitor",
    "Early Mk progenitor",
    "Early ery/baso/mast 1",
    "Early ery/baso/mast 2",
    "Ery progenitor",
    "Cycling ery",
    # add populations that will be removed due to their status
    # as mature immune cell contamination or low quality cell contamination
    "Low quality, early",
    "Low quality, neutro",
    "Low quality, ery",
    "Late immune, mixed"
    ))
stopifnot(!is.na(sce_species$custom_label))
```

```{r unif}

sce_species$custom_label_unif <- unfactor(sce_species$custom_label)

sce_species$custom_label_unif[
  sce_species$custom_label == "MPP cells 1"] <-  "MPP cells"
sce_species$custom_label_unif[
  sce_species$custom_label == "MPP cells 2"] <-  "MPP cells"

sce_species$custom_label_unif[
  sce_species$custom_label == "Early ery/baso/mast 1"] <-  "Early ery/baso/mast"
sce_species$custom_label_unif[
  sce_species$custom_label == "Early ery/baso/mast 2"] <-  "Early ery/baso/mast"

sce_species$custom_label_unif[
  sce_species$custom_label == "Low quality, early"] <-  "Low quality"
sce_species$custom_label_unif[
  sce_species$custom_label == "Low quality, neutro"] <-  "Low quality"
  sce_species$custom_label_unif[
  sce_species$custom_label == "Low quality, ery"] <-  "Low quality"

sce_species$custom_label_unif <- factor(
  sce_species$custom_label_unif,
  levels = c(
    "HSC",
    "MPP cells",
    "Early myeloid",
    "Neutro progenitor",
    "Cycling, mostly myeloid",
    "Mono/antigen-presenting",
    "Early lymphoid",
    "Late/Cycling lymphoid",
    "Baso/Mast progenitor",
    "Early Mk progenitor",
    "Early ery/baso/mast",
    "Ery progenitor",
    "Cycling ery",
    # add populations that will be removed due to their status
    # as mature immune cell contamination or low quality cell contamination
    "Low quality",
    "Late immune, mixed"
  )
)

stopifnot(!is.na(sce_species$custom_label_unif))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 5}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```
