---
title: "Custom clustering mcas str"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])

fraction_curr <- "str"
species_curr <- "mcas"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mcas_str_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:40])
names(col_num) <- as.character(c(1:40))
```

# Clustering

### Define parameters

Caroli and Cast will be tough because they contain many more contaminating
cells than cells of interest. 
So I'll do a super fine clustering to try and properly separate even the
very rare cell types of interest, like sinusoidal ECs.
I'll try at least.

```{r parameters}
k_graph <- 5
resolution_louvain <- 1
```

This was best after trying several times even though it's still a bit
less refined than the original clustering would have been I think.

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "str"){
  marker_genes <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Anxa1",
    "Bglap",
    "Bglap2",
    "Cd200",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Ptgis",
    "Pth1r",
    "Runx2",
    "Sp7",
    "Spp1",
    "Tnc",
    "Acan", # Chondro
    "Adamts5",
    "Chad",
    "Fmod",
    "Mef2c",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", # fibro chondro
    "S100a10",
    "Cd34", # fibro
    "Col3a1",
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Cdh5", # pan endo
    "Pecam1",
    "Car4", # capillary
    "Rgcc",
    "Timp4",
    "Efnb2", # Arteriolar
    "Hey1",
    "Ly6a",
    "Sox17",
    "Ackr1", # Venous/sinusoidal
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Nr2f2",
    "Selp",
    "Vcam1", # sinusoidal
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", # Lymphatic EC
    "Lyve1", 
    "Pdpn",
    "Prox1",
    "Rgs4", # pericyte
    "Rgs5",
    "Steap4",
    "Acta2", # smooth muscle
    "Myh11", 
    "Tagln",
    "Myod1", # just some skeletal muscle markers for contamination
    "Myh2",
    "Des",
    "Cd3d", # just some mature immune cell markers for contamination
    "Cd8a",
    "Cd4",
    "Il7r",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Cd74",
    "S100a8",
    "S100a9",
    "Ccl5",
    "Cd3e", # TC
    "Tcf7",
    "Cpa3", # Mast
    "Hdc",
    "Gypa", # Ery
    "Hba-a1"
  )
}

```

```{r genes_short}

if(fraction_curr == "str"){
  marker_genes_short <- c(
    "Adipoq", 
    "Cxcl12", 
    "Kitl", 
    "Lepr",
    "Alpl", 
    "Bglap",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Runx2",
    "Sp7",
    "Spp1",
    "Acan", 
    "Chad",
    "Fmod",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34", 
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Pecam1",
    "Car4",
    "Rgcc",
    "Efnb2",
    "Hey1",
    "Sox17",
    "Ackr1", 
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Rgs4", 
    "Steap4",
    "Acta2", 
    "Myh2"
  )
}

```

#### Heatmap

Just to get an overview of what is found in each cluster

```{r heatmap, fig.height = 10, fig.width = 10}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 10}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

Check markers because there are so many different cell types
including the contamination that I cannot possibly have a complete list.

The number of clusters is ridiculous as expected, so this is taking a while.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### Contamination

Get the contamination out of the way first.

I don't care that much about them as long as I can exclude that 
they are of interest, so I will not annotate them too nicely.

#### Ery

```{r markers_contamination_ery}

cluster_markers[[24]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# top markers strongly support Hb synthesis.
```

#### Neutro and other myeloid

```{r markers_contamination_neutrom}

# also looks mixed with some Lyve1, Hdc, and Cd3
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# These are mainly neutros with Camp, Wfdc21, Retnlg, S100a8 and S100a9
# check all genes

# possibly early neutro?
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# co-expression of neutro markers like S100a9 and Ngp, 
# BC markers like Igs and Baso marker like Prss34, so mixed.
# check 

# looks mixed with some Hdc and Cd3 expressing cells
cluster_markers[[27]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# the top markers are more Neutro than others I can tell.
# Retnlg, S100a8 and S100a9 all related to neutro
# check because in the heatmap there was "contamination" with Hdc and Cd3d

# baso mast
cluster_markers[[16]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# classic baso mast markers like Prss34, Gata2, Hdc
# but also some neutro like Cd63 ... might also call them mixed.
# check more
```

```{r umaps_neu, fig.height = 3, fig.width = 4}
# find the clusters
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "4"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "13"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "27"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "16"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_ery <- as.list(c(
  "S100a8",
  "S100a9",
  "Wfdc21",
  "Camp",
  "Trem1",
  "Hdc",
  "Cd3d",
  "Lyve1",
  "Lrg1",
  "Anxa1",
  "Prss34",
  "Jchain",
  "Igha",
  "Vpreb1",
  "Cd79a",
  "Mcpt8",
  "Cd63",
  "Ier3"
))

plot_list_gene <- lapply(marker_genes_list_ery, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_ery
plot_list_gene
```

Clusters 4 and 27 look classic neutro.
They do express Cd3d but no Cd8 or Cd4 so I'd stick with neutro.

Cluster 13 is a bit less clear. 
One of the few populations that express Jchain.
I think it's just mixed.

Cluster 16 looks clearly mast baso. 
Seems Cd63 is actually mast/baso and not neutro here.

#### Lymphoid

There are lots of lymphoid-looking clusters based on the heatmap.
Unfortunately I don't know many classic mature lymphoid lineage markers
by heart since they're not in the HSPC dataset.

```{r markers_contamination_lymph}

# TC lineage?
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Ccl5, Nkg7, Txk are NK
# Cd3d, Trbc2 are TC, so both

# APC?
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# H2 genes, Ccr2 and Cd74 --> antigen-presenting

# BC-ish
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Igll1, Vpreb1 are BC lineage, but the rest either suggests low quality
# or cycling, probably the latter

# TC
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# actually looks kind of DC, too with Siglech, Irf8, Runx2
# check more

# APC?
cluster_markers[[18]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# more clearly BC lineage with lots of Ig stuff

# BC-ish
cluster_markers[[21]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also quite clearly BC lineage with Vpreb3, Cd79b, Ig stuff

# BC-ish
cluster_markers[[23]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also quite clearly BC lineage with Vpreb1, Il7r, Sox4, possibly a little
# earlier, possibly low quality with Xist

# BC-ish
cluster_markers[[25]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also quite clearly BC lineage with Vpreb3, Cd79b, Sox4, Ig stuff

# BC-ish
cluster_markers[[28]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# actually also very antigen-presenting like with H2, but also Ig and classic
# BC markers, other than 10 which seemed more mono-derived?

# BC-ish
cluster_markers[[29]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# clearly BC lineage with Ig, Cd74, Cd79b, Vpreb3, Cd79a etc.

# BC-ish
cluster_markers[[30]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# clearly BC lineage with Vpreb1 and Vpreb3, Sox4, also maybe slightly early
# with Socs2?
```

```{r umaps_lym, fig.height = 3, fig.width = 4}
# find the clusters
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "2"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "12"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "14"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

# addind some known markers genes to the markers from the lists
marker_genes_list_ery <- as.list(c(
  "Cd3d",
  "Cd3e",
  "Ltb",
  "Gzmb",
  "Cd8a", # TC
  "Cd4", # TC
  "Cd27", # NK 
  "Klrg1",
  "Klrd1",
  "Nkg7",
  "Trbc1", # TC
  "Ccl5",
  "Mki67",
  "Aurkb",
  "Vpreb1",
  "Igll1",
  "Siglech",
  "Irf8",
  "Cd7",
  "Runx2",
  "Ighm",
  "Cd47"
))

plot_list_gene <- lapply(marker_genes_list_ery, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_ery
plot_list_gene
```

Cluster 2 is looking more NK than T cell.
But I still cannot exclude that TC are in there.
Cluster 12 is indeed cycling B lineage.
Cluster 14 is indeed mono/macrophage.

#### Other

```{r markers_contamination_other}

# could be skeletal muscle
cluster_markers[[26]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# yes, these are all skeletal muscle genes

# these clusters don't look like they epxress anything relevant in the heatmap
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# I'd say some sort of mature monocyte-ish cell type? Fn1, F13a1, Ccr2
# Mpeg1, Itga4 is macrophage, Lyst possibly, to. Dleu2 is TC, Zeb2 is mesenchyme?
# check more

cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks like some neutro-ery mix?
# check more

cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# could also be macrophage-ish
# check more

cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# that seems just cycling
# check more

cluster_markers[[17]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# could also be macrophage-ish
# check more
```

```{r umaps_other, fig.height = 3, fig.width = 4}
# find the clusters
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "3"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "5"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "6"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "7"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "17"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_ery <- as.list(c(
  "Fn1", # macrophages, NK cells, Tregs
  "Dleu2", # TC, BC, mono, macro
  "Mpeg1", # macrophage, phagocytes
  "Ccr2", # mono, macro, activated TC, DC, neutro
  "Zeb2", # dendritic cells, macrophages, monocyte
  "Ahnak", # CD4 TC, monos?
  "Itga4", # Tregs, CD4 TC
  "Il6ra", # TC 
  "Ms4a6c", # microglia, macrophages
  "Ass1",
  "Lgals3",
  "Lgals1",
  "F13a1", # macrophages, coagulation factor 13
  "Tspo", # actually this is the only marker gene that is ery
  "Fcer1g",
  "Klf1",
  "Epor",
  "Mpo",
  "Hopx",
  "Atf3",
  "Cd14",
  "Thbs1",
  "Ccl6",
  "Mki67",
  "Aurkb",
  "Lockd",
  "Cenpa",
  "Lig1",
  "Cr2", # Cd21
  "Cd19"
))

plot_list_gene <- lapply(marker_genes_list_ery, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_ery
plot_list_gene
```

So the blob containing clusters 3, 5, 6, 7, and 17 all look like the belong to 
the same populations.
They express lots of markers that are not super exclusive to cell types, 
like Fcer1g, Il6ra, F13a1.
Given the lack of Cd8, Cd4, Cd3 these are not activated TCs.
Given the lack of BC markers like Ig genes, Vprebs, Cd79 these are not BC lin.
Given the lack of Mpo, Elane, S100a8, Hdc, Hba-a1, they are also not these
lineages.
They do express Irf8, Ccr2, Mpeg1, so I conclude that they are all
mono/macrophage lineage of some sort.

### Expected clusters

```{r umaps_expected, fig.height = 3, fig.width = 4}
# find the cluster
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
sce_expected <- sce_species[,sce_species$cluster_louvain_custom %in% c(
  "1",
  "8",
  "9",
  "11",
  "15",
  "19",
  "20",
  "22"
)]
```

```{r markers_car}
# CAR
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Adipoq, Cxcl12, Kitl, classic CAR markers but also some osteo genes like Alpl,
Spp1, Bglap, Mmp13, Pth1r are beginning to be expressed in this cluster.
Both adipo and osteo.

```{r markers_osteo}
# osteo
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

No Cxcl12 etc, but Alpl, Bglap, Ibsp and Runx2 are expressed here, looks very 
Osteo.

```{r markers_msc_balanced}
# balanced in some way
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Portions of Alpl, Bglap, Postn, Ptgis from osteo lineage.
Portions of Chad, Fmod, Mgp, Prg4, Comp, S100a10 from Chondro lineage.
Portions of Dcn, Fn1 from fibro lineage.
No adipo-lineage.
The main markers look mostly chondro-fibro.

```{r markers_fibro}
# fibro?
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Nothing from osteo-lineage, very little from chondro lineage. 
Cd34, Dcn, Fn1, Lum.
Looks very fibro.

```{r markers_chondro}
# chondro?
cluster_markers[[22]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Acan and Sox9 are basically only expressed here; 
Chad, Mgp, Prelp, Comp are also found here.

```{r marker_genes_list_msc}
marker_genes_list_msc <- as.list(c(
  # add some more expected genes that I know even though we had them before
  "Cxcl12", # CAR 8
  "Lpl",
  "Vcam1",
  "Kitl",
  "Adipoq",
  "Lepr",
  "Spp1",
  "Alpl",
  "Bglap", # Osteo 1
  "Ibsp",
  "Car3",
  "Ptgis",
  "Cpe",
  "Runx2",
  "Dcn", # balanced 9
  "Lum",
  "Col1a1",
  "Col1a2",
  "Aspn",
  "Ogn",
  "Fmod",
  "Comp",
  "Abi3bp",
  "Pdgfra", # fibro
  "Pdgfrb",
  "Cd34",
  "Thy1",
  "Gsn",
  "Ly6a",
  "Clec3b",
  "Col2a1", # chondro
  "Snorc",
  "Mgp",
  "Prelp",
  "Chad",
  "Prg4",
  "Acan",
  "Sox9"
))

plot_list_gene <- lapply(marker_genes_list_msc, function(gene){
  plot <- umap_gene(sce_expected, gene)
})
names(plot_list_gene) <- marker_genes_list_msc
plot_list_gene
```

```{r EC_sinus}
# EC sinusoidal
cluster_markers[[19]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Tfpi, Plvap, Lrg1, Stab2, Kdr, Emcn, Ackr1 all identified as main markers,
all sinusoidal/venous markers in some way. 

```{r EC_art_cap}
# EC art cap
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Cxcl12, Cd36, Cd200 among the main genes.

```{r marker_genes_list_ec}
marker_genes_list_ec <- as.list(c(
  "Plvap", # EC sinusoidal, venular, lymphatic?
  "Fabp4",
  "Tfpi",
  "Adamts5",
  "Ackr1",
  "Selenop",
  "Stab2", 
  "Il6st",
  "Kdr",
  "Eng",
  "Vwf",
  "Emcn",
  "Lrg1",
  "Eng",
  "Ephb4",
  "Selp",
  "Icam1",
  "Vcam1",
  "Sox4", # other
  "Flt4",
  "Esam",
  "Lyve1", # lymph
  "Ccl21a",
  "Crip2", # EC arteriolar and capillary?
  "Ctla2a",
  "Ramp2",
  "Sparcl1",
  "Timp3",
  "Rgcc", # add more
  "Car4",
  "Cd36",
  "Hey1",
  "Sox17",
  "Efnb2"
))

plot_list_gene <- lapply(marker_genes_list_ec, function(gene){
  plot <- umap_gene(sce_expected, gene)
})
names(plot_list_gene) <- marker_genes_list_ec
plot_list_gene
```

```{r peri_smc}
# peri/smc
cluster_markers[[20]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Rgs5, Acta2, Tagln, Tpm2 are Peri/SMC.
Seems to contain some chondro contamination with Mgp, Aspn if I interpret 
correctly.

# Assign

In the same order as above.

```{r assign}

sce_species$custom_label <- vector(length = ncol(sce_species))

# Ery
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "24"] <-  "Erythroid"

# Neutro and myeloid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Neutro 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "27"] <-  "Neutro 2"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Mature immune, mixed 1"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "16"] <-  "Baso/Mast"

# Lymphoid
# order BC for nicer gene expression plots downstream
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Nk/TC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Antigen-presenting 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "BC lineage 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Mature immune, mixed 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "18"] <-  "BC lineage 7"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "21"] <-  "BC lineage 4"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "23"] <-  "BC lineage 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "25"] <-  "BC lineage 5"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "28"] <-  "Antigen-presenting 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "29"] <-  "BC lineage 6"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "30"] <-  "BC lineage 1"

# Other
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "26"] <-  "Skeletal muscle"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Mono/Macro lineage 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Mono/Macro lineage 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Mono/Macro lineage 4"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Mono/Macro lineage 2, cycling"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "17"] <-  "Mono/Macro lineage 5"

# Expected
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Adipo/Osteo CAR"
  # because there is significant osteo expression here, as well
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Osteo"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Balanced mesenchymal"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Fibro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "22"] <-  "Late chondro"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "19"] <-  "Sinusoidal EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Arteriolar/capillary EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "20"] <-  "Peri/SMC"

sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "Adipo/Osteo CAR",
    "Osteo",
    "Balanced mesenchymal",
    "Fibro",
    "Late chondro",
    "Arteriolar/capillary EC",
    "Sinusoidal EC",
    "Peri/SMC",
    "Erythroid",
    "Mono/Macro lineage 1", # they look a bit "earlier" than the neutros
    "Mono/Macro lineage 2, cycling",
    "Mono/Macro lineage 3",
    "Mono/Macro lineage 4",
    "Mono/Macro lineage 5",
    "Neutro 1",
    "Neutro 2",
    "Baso/Mast",
    "Nk/TC",
    "BC lineage 1",
    "BC lineage 2",
    "BC lineage 3",
    "BC lineage 4",
    "BC lineage 5",
    "BC lineage 6",
    "BC lineage 7",
    "Antigen-presenting 1",
    "Antigen-presenting 2",
    "Mature immune, mixed 1",
    "Mature immune, mixed 2",
    "Skeletal muscle"
    ))

sort(unique(sce_species$custom_label))
levels(sce_species$custom_label)
stopifnot(!is.na(sce_species$custom_label))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 7}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```