---
title: "Custom clustering mmus str"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])

fraction_curr <- "str"
species_curr <- "mmus"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mmus_str_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:30])
names(col_num) <- as.character(c(1:30))
```

# Clustering

### Define parameters

```{r parameters}
k_graph <- 13
resolution_louvain <- 0.7
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)

# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "str"){
  marker_genes <- c(
    "Adipoq", 
    "Cxcl12",
    "Kitl", 
    "Lepr",
    "Alpl",
    "Anxa1",
    "Bglap",
    "Bglap2",
    "Cd200",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Ptgis",
    "Pth1r",
    "Runx2",
    "Sp7",
    "Spp1",
    "Tnc",
    "Acan", 
    "Adamts5",
    "Chad",
    "Fmod",
    "Mef2c",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34",
    "Col3a1",
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Cdh5",
    "Pecam1",
    "Car4", 
    "Rgcc",
    "Timp4",
    "Efnb2", 
    "Hey1",
    "Ly6a",
    "Sox17",
    "Ackr1",
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Nr2f2",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Pdpn",
    "Prox1",
    "Rgs4", 
    "Rgs5",
    "Steap4",
    "Acta2",
    "Myh11", 
    "Tagln",
    "Myod1", # just some skeletal muscle markers for contamination
    "Myh2",
    "Cd3d", # just some mature immune cell markers for contamination
    "Cd8a",
    "Cd4",
    "Il7r",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Cd74",
    "S100a8",
    "S100a9",
    "Ccl5",
    "Tcf7",
    "Cpa3", 
    "Hdc",
    "Gypa", 
    "Hba-a1"
  )
}
```

```{r genes_short}

if(fraction_curr == "str"){
  marker_genes_short <- c(
    "Adipoq",
    "Cxcl12", 
    "Kitl", 
    "Lepr",
    "Alpl",
    "Bglap",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Runx2",
    "Sp7",
    "Spp1",
    "Acan", 
    "Chad",
    "Fmod",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34", 
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Pecam1",
    "Car4",
    "Rgcc",
    "Efnb2",
    "Hey1",
    "Sox17",
    "Ackr1", 
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Rgs4", 
    "Steap4",
    "Acta2", 
    "Myh2"
  )
}
```

#### Heatmap

Just to get an overview of what is found in each cluster.

```{r heatmap, fig.height = 10, fig.width = 6}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 10}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

 Check each cluster's top markers to support the already observed patterns
 and find possible additional genes.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### Mesenchymal fraction

```{r markers_mes}

# Car/Adipo
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# very strong niche, adipo, Hp, Adipoq, Cxcl14, Vcam, Lepr

cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
# looks mesenchymal/adipo but weaker, check the quality as well
# looks very similar to the corresponding low quality
# cluster in mspr, even the top marker genes which makes me a bit suspicious

# osteo
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# very strong osteo-gene expression, Spp1, Bglap, Obsp, Alpl

# osteo-chondro
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks instead more chondro-ish based on the main genes, like Acan, Chad

# balanced or mixed? 
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# the very early chondro (Prg4, Cilp, Chad) and mixed
# Fmod, Comp, Aspn expression. Could also be "early chondro"?

# fibro
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also some fibro-chondro-ish expression but sole 
# expression of Cd34, Thy1 here mean fibro
```

Check these unclear clusters; ALSO check quality.

```{r umaps_mes, fig.height = 3, fig.width = 4}

marker_genes_list_mes <- as.list(c(
  "Cebpa", # https://doi.org/10.1016/j.beem.2021.101518 MALP
  "Runx1", # osteogenic differentation, inhibits adipo  https://doi.org/10.7554/eLife.54695
  "Pappa", 
  "Pparg", # https://doi.org/10.1016/j.beem.2021.101518 MALP
  "Plin1", # lipid droplet coating https://doi.org/10.1016/j.beem.2021.101518 MALP
  "Fabp4", # fatty acid bp indicating no lipid droplet https://doi.org/10.1016/j.beem.2021.101518 
  "Lep", # white adipose tissue https://doi.org/10.1016/j.beem.2021.101518 
  "Ucp1", # brown https://doi.org/10.1016/j.beem.2021.101518 
  "Tnfrsf9", # beige https://doi.org/10.1016/j.beem.2021.101518 
  "Cidea", # brown https://doi.org/10.1016/j.beem.2021.101518 
  "Cited1", # beige https://doi.org/10.1016/j.beem.2021.101518 
  "Tbx1", # beige https://doi.org/10.1016/j.beem.2021.101518 
  "Hoxc9", # adipose depot https://doi.org/10.1016/j.beem.2021.101518 
  "Vegfc", # angiogenic factor expressed by adipo CAR/MAlP https://doi.org/10.1016/j.beem.2021.101518 
  "Vegfa", # angiogenic factor expressed by adipo CAR/MAlP https://doi.org/10.1016/j.beem.2021.101518 
  "Rspo3", # angiogenic factor expressed by adipo CAR/MAlP https://doi.org/10.1016/j.beem.2021.101518 
  "Apoe", # adipocyte marker in MALP https://doi.org/10.1016/j.beem.2021.101518 
  "Ebf3", 
  "Zeb2", # niche https://doi.org/10.1182/blood-2016-05-714659
  "Tnc",
  "Gdpd2", # https://doi.org/10.1016/j.beem.2021.101518 MALP
  "Xist", # low quality
  "Ntng1", # now from similar population in mspr, in obese adipose tissue
  "Tns3", # osteogenic vs adipocytic fate https://doi.org/10.1007/s00018-023-04930-5
  "Ptprd", # related to adipogenesis https://doi.org/10.2217/epi-2019-0267 https://doi.org/10.1016/j.cmet.2022.02.012
  "Palld", # related to adipocyte size, in stressed adipocytes https://doi.org/10.1016/j.cmet.2020.12.004
  "Cp", # related to obesity, expressed in adip tissue https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0080274
  "Rai14", # related to migration, mechanical sensing? https://doi.org/10.1038/s44319-024-00228-0
  "Atp13a3", # mechanical role in osteogenic differentiation https://doi.org/10.1016/j.advms.2024.07.003
  "Snorc", # chondro cluster markers from list
  "Col9a1",
  "Mia",
  "Matn3",
  "Thbs4",
  "Cilp2",
  "Cilp",
  "Prg4"
))

plot_list_gene <- lapply(marker_genes_list_mes, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_mes
plot_list_gene
```

```{r mes_qc}
qc_df <- scuttle::perCellQCMetrics(sce_species)
sce_species$log_sum <- log(qc_df$sum)
sce_species$log_detected <- log(qc_df$detected)

umap_cont(sce_species, "log_sum")
umap_cont(sce_species, "log_detected")

qc_df$cluster_louvain_custom <- sce_species$cluster_louvain_custom
qc_df <- as.data.frame(qc_df)

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(sum)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(detected)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()
```

Cluster 8 is also low quality but it's contamination anyways (see below).

So the quality of that additional adipo Cluster 6 is indeed very low.
The EC clusters are also generally quite low, so if I went any
lower with UMI cutoff I would lose those cells.
I'm a bit suspicious that these low-quality adipo-like cells
and the low quality chondro cells (here not as a separate cluster) 
are found in both mspr and mmus very predominantly.
I will look into the Adipo cluster but tbh, I did that for the initial
annotation and back then I couldn't really determine if these could
correspond to a specific population so I removed them.

No Plin1 would indicate no liquid droplet.
The mystery cluster doesn't express "mature" adipocytic markers more or
very differently than the adipo-CAR cluster, so it doesn't seem like 
it is in a different state of differentiation.
I really think it's just lower quality. 
The fact that Xist is higher here than in others could indicate some normali-
sation problem, which is also why other adipose genes looks slightly higher
here, they probably had dysproportionate size factors similar to Xist.

Still, check violin plots to make a bit more quantitatively sure it's really not 
more down the adipo-line.

```{r violin_mes, fig.height = 12, fig.width = 3}

# take only the relevant genes and clusters
genes_of_interest <- marker_genes_list_mes[c(
  1:4,
  6,
  15:24,
  25:31)]

sce_species_genes <- sce_species[rownames(sce_species) %in% genes_of_interest,]
sce_species_genes <- sce_species_genes[,sce_species_genes$cluster_louvain_custom %in% c("1", "6")]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = genes_of_interest)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::stat_ydensity(
      stat = "ydensity",
      draw_quantiles = c(0.25, 0.5, 0.75),
      scale = "area")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

Some are higher, a few are lower. 
In the end they are still adipo-primed cells with high Cxcl12 
expression so by definitions CARs.
So they are low quality adipo CARs after all that.

## ECs

Plus peri SMCs.

```{r markers_ecs}

# art 
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Sox17, Gja4 are arteriolar

# cap
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Cd36 is interesting, Cavs and Cavins + Timp4 + Cd36 look capillary

# art cap mix? 
cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks more strongly capillary, Rgcc, Timp4, Car4 I have used as markers

# sinusoidal 1
cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 25)
# looks sinusoidal with Lrg1,
# Emcn, Vwf and Plvap could also suggest some other H or L-type stuff
# check if they really express more sinusoidal genes

# sinusoidal 2
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also looks a bit more classic sinusoidal with Tfpi, Lrg1, Stab2, Stab1, 

# lymphatic, also seems to contain immune contamination
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# heatmap says lymphatic, markers clearly say B cell contamination 

# peri smc 
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Rgs5, Acta2, Tagln, Tpm, classic
```

```{r umaps_ec, fig.height = 3, fig.width = 4}

marker_genes_list_ec <- as.list(c(
  "Gja4", # art cluster
  "Flt1",
  "Id1",
  "Ptprb",
  "Sema3g", 
  "Podxl",
  "Slc9a3r2",
  "Stmn2",
  "Bmx", # adding some art genes from literature https://doi.org/10.1038/s41467-018-04726-3
  "Cd200",
  "Ace",
  "Fbln2",
  "Lims2",
  "Jam2",
  "Gkn3",
  "Tspan13", # cap cluster 1
  "Gpihbp1",
  "Sparcl1",
  "Cd36",
  "Hspb1", # cap cluster 2
  "Cd300lg",
  "Vwf", # sinusoidal 1
  "Plvap",
  "Ehd4", 
  "Abcg2",
  "Il6st",
  "Ctla2a",
  "Grrp1",
  "Fxyd5",
  "Mrc1", # sinusoidal 2, 
  "Mafb", 
  "Abcc9",
  "Stab1",
  "Tcn2",
  "Sele", # adding some more from literature https://doi.org/10.1038/s41467-018-04726-3
  "Dnase1l3",
  "C1qtnf1",
  "Ubd",
  "Sirpa",
  "Cd302",
  "Sox4", # in some H type ECs
  "Kdr", # some others
  "Dll4",
  "Notch1",
  "Robo4",
  "Vpreb3", # immune cont 
  "Ighm"
))

plot_list_gene <- lapply(marker_genes_list_ec, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_ec
plot_list_gene
```

The BL6 sinusoidal "classics" like Stab2 are only expressed in cluster 12. 
This also includes some literature markers Cd302, Sirpa, Ubd, etc. 
But other marker genes like Selp, Icam1, Vcam1 are expressed in both 12 and 5.
Cluster 5 expresses Cd34 and some Sox17, but also some Lrg1, Nr2f2, Selp.
It exclusively expresses Tagln (in ECs) Vwf, Ehd4, Grrp1 to some degree. 
I will go ahead and say these are not sinusoidal after all (and I will recheck
some of the other species, but it seemed less clearly different there).

Due to this mix of arterial and venous identity genes I'm going to call them
"Transitory venous".

## Clear contamination?

```{r markers_cont}

# looks like ery
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
# main markers are clearly ery

# looks like skeletal muscle
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
# main markers are clearly skeletal muscle
```

# Assign

In the same order as above.

```{r assign}

sce_species$custom_label <- vector(length = ncol(sce_species))

# Mes
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Adipo CAR"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Low qu adipo CAR"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Osteo"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Balanced mesenchymal"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Late chondro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Fibro"


# EC
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Arteriolar EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Capillary EC 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Capillary EC 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Transitory/Venous EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "Sinusoidal EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Mix lymphatic EC/immune"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Peri/SMC"

# contamination
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Erythroid"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Low qu skeletal muscle"


sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "Adipo CAR",
    "Osteo",
    "Balanced mesenchymal",
    "Late chondro",
    "Fibro",
    "Capillary EC 1",
    "Capillary EC 2",
    "Arteriolar EC",
    "Transitory/Venous EC",
    "Sinusoidal EC",
    "Peri/SMC",
    # contamination
    "Erythroid",
    "Low qu adipo CAR",
    "Low qu skeletal muscle",
    "Mix lymphatic EC/immune"
    ))

sort(unique(sce_species$custom_label))
levels(sce_species$custom_label)
stopifnot(!is.na(sce_species$custom_label))
```

```{r unif}

sce_species$custom_label_unif <- unfactor(sce_species$custom_label)

sce_species$custom_label_unif[
  sce_species$custom_label == "Capillary EC 1"] <-  "Capillary EC"
sce_species$custom_label_unif[
  sce_species$custom_label == "Capillary EC 2"] <-  "Capillary EC"

sce_species$custom_label_unif <- factor(
  sce_species$custom_label_unif,
  levels = c(
    "Adipo CAR",
    "Osteo",
    "Balanced mesenchymal",
    "Late chondro",
    "Fibro",
    "Capillary EC",
    "Arteriolar EC",
    "Transitory/Venous EC",
    "Sinusoidal EC",
    "Peri/SMC",
    # contamination
    "Erythroid",
    "Low qu adipo CAR",
    "Low qu skeletal muscle",
    "Mix lymphatic EC/immune"
  )
)

stopifnot(!is.na(sce_species$custom_label_unif))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 7}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```