---
title: "Custom clustering mcar str"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])
#sce_species <- base::readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/01_mnnc/comparison/sce_mcar_str_Object_ID-01")

fraction_curr <- "str"
species_curr <- "mcar"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be
# defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mcar_str_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:40])
names(col_num) <- as.character(c(1:40))
```

# Clustering

### Define parameters

Caroli and Cast will be tough because they contain many more contaminating
cells than cells of interest. 
So I'll do a super fine clustering to try and properly separate even the
very rare cell types of interest, like sinusoidal ECs.
I'll try at least.

```{r parameters}
k_graph <- 5
resolution_louvain <- 1.2 # 1
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "str"){
  marker_genes <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Anxa1",
    "Bglap",
    "Bglap2",
    "Cd200",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Ptgis",
    "Pth1r",
    "Runx2",
    "Sp7",
    "Spp1",
    "Tnc",
    "Acan", # Chondro
    "Adamts5",
    "Chad",
    "Fmod",
    "Mef2c",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", # fibro chondro
    "S100a10",
    "Cd34", # fibro
    "Col3a1",
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Cdh5", # pan endo
    "Pecam1",
    "Car4", # capillary
    "Rgcc",
    "Timp4",
    "Efnb2", # Arteriolar
    "Hey1",
    "Ly6a",
    "Sox17",
    "Ackr1", # Venous/sinusoidal
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Nr2f2",
    "Selp",
    "Vcam1", # sinusoidal
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", # Lymphatic EC
    "Lyve1", 
    "Pdpn",
    "Prox1",
    "Rgs4", # pericyte
    "Rgs5",
    "Steap4",
    "Acta2", # smooth muscle
    "Myh11", 
    "Tagln",
    "Myod1", # just some skeletal muscle markers for contamination
    "Myh2",
    "Des",
    "Cd3d", # just some mature immune cell markers for contamination
    "Cd8a",
    "Cd4",
    "Il7r",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Cd74",
    "S100a8",
    "S100a9",
    "Ccl5",
    "Cd3e", # TC
    "Tcf7",
    "Cpa3", # Mast
    "Hdc",
    "Gypa", # Ery
    "Hba-a1"
  )
}

```

```{r genes_short}

if(fraction_curr == "str"){
  marker_genes_short <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Bglap",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Runx2",
    "Sp7",
    "Spp1",
    "Acan", 
    "Chad",
    "Fmod",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34", 
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Pecam1",
    "Car4",
    "Rgcc",
    "Efnb2",
    "Hey1",
    "Sox17",
    "Ackr1", 
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Rgs4", 
    "Steap4",
    "Acta2", 
    "Myh2"
  )
}

```

#### Heatmap

Just to get an overview of what is found in each cluster

```{r heatmap, fig.height = 10, fig.width = 10}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

That was less challenging than I expected.

#### Violin

```{r violin, fig.height = 20, fig.width = 10}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

Check markers because there are so many different cell types
including the contamination that I cannot possibly have a complete list.

The number of clusters is ridiculous as expected, so this is taking a while.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### Contamination

Get the contamination out of the way first.

I don't care that much about them as long as I can exclude that 
they are of interest, so I will not annotate them too nicely.

#### Ery

```{r markers_contamination_ery}

cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Car1, Car2, Hba-a1, Hba-a2, Gypa, Alad

cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Alas, Hba-a2, Gypa, Hba-a1, Hemgn

cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Car, Car2, Gype, Hba-a1, Klf2, Alad

cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Ermap, Rhd, Tfrc, definitely ery-like but check out in more detail

cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Car, Car2, Rhd, Hba-a1, Ermap

cluster_markers[[17]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Alas2, Hbb-bt, Tspo2, Rhd, Hba-a1

cluster_markers[[23]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Car1, Car2, Hba-a1, Hba-a2, Gypa

cluster_markers[[33]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Hba-a2, Gypa, Hbb-bt, Rhd, Car2
```

Check out cluster 8.

```{r umaps_8, fig.height = 3, fig.width = 4}
# find the cluster
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "8"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_ery <- as.list(c(
  "Slc4a1", # in eythroid membrane
  "Kel", # in erythroid membrane, blood group antigen
  "Ermap",
  "Ctse",
  "Rhd",
  "Sox6", # erythroid maturation
  "Tfrc", # iron transporter
  # also adding some Ery classics
  "Hba-a1",
  "Gypa",
  "Klf1",
  "Epor"
))

plot_list_gene <- lapply(marker_genes_list_ery, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_ery
plot_list_gene
```

These are looking like very late RBC/Erys which is why I didn't know some
these genes from the progenitors I ususally look at.
But the top genes Slc4a1, Kel, Sox6 are definitely erythroid.

#### Neutro

```{r markers_contamination_neutro}

cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Rnase2a, S100a8, Elane, S100a9, Wfdc21, F13a1, Prtn3

cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Elane, Prtn3, Rnase2a, S100a8, Wfdc21, Ccr2, Cebpb

cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Thbs1, Trem1, Ccl6, Fn1, Ctss, Ms4a6c, Mmp8, Rgs, Cybb
# not overtly neutrophilic as far as I can tell, check out in more detail

cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Camp, Ly6a, S100a9, Trem2, Cepbe, Lrg1, Chil1, Wfdc21, Ms4a3, Cd63

cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Fn1, Ccl6, Ctss, Mmp8, Chil1, Ms4a6c, Wfdc21, F13a1, Cybb
# similar to 10 so check out but much more neutro looking

cluster_markers[[21]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Fn1, Wfdc21, S100a8, S100a9, Rnase2a, Ms4a6c

cluster_markers[[29]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Camp, Lrg1, Cd3d, Ngp, Cd52, Mmp9, Chil1, Ly6a
# looks a bit more mixed? check out

cluster_markers[[32]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Ngp, Camp, S100a8, S100a9, Trem3, Wfdc21, Cebpe, Chil1, Elane

cluster_markers[[35]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Retnlg, Prg2, Ccl6, Prg3, Cebpe, Cd63, Ear1, Ear2
# maybe eosinophilic? I don't know many eo genes. check out
```

Checking out clusters 10, 14, 29, 35

```{r umaps_neutro, fig.height = 3, fig.width = 4}
# find the clusters
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "10"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "14"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "29"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "35"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_neu <- as.list(c(
  "Thbs1", # related to  neutro serine protease function
  "Atf3", # neutrophilic responses, mature neutro
  "Trem1", # neutro migration
  "Ccl6", # one of the most abundant chemokines in mouse neutros
  "Ctss", # macrophage mabye, neutro inflammation
  "Fn1", 
  "Lrg1", # in neutro granules
  "Ltf", # lactoferrin, related to NETs?
  "Mmp9", # a subset of neutrophils
  "Asprv1", # macrophage-like  neutrophil cells
  "Cd52", # also expressed on neutrophils
  "Retnlg", # different subsets of neutrophils
  "Prg2", # eosinophil
  "Ear2",
  "Prg3", # eosinophil granules
  "Ear6"  # eosinophil granules
))

plot_list_gene <- lapply(marker_genes_list_neu, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_neu
plot_list_gene
```

So the mystery neutro clusters are actually very overtly neutrophilic
just much more mature than what I'm used to from the progenitors
in terms of expression.
Actually, cluster 29 also contains some Cd3d, so I'd rather call it mixed.
The last cluster are eosinophils which are not in the HSPC dataset.

#### Baso Mast

```{r markers_contamination_bm}

cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Prss34, Mcpt8, Gata2, Ms4a2, Fcer1a, Hdc, Cpa3, Prg3
# Mast/Baso, maybe a little eosiniphilic, too
```

#### Lymphoid

```{r markers_contamination_lymph}

cluster_markers[[20]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Siglech, Mef2c, Ramp1, Sox4, Runx2, Irgf8, Ms4a6c, Tcf4, Ighm, Rnase6
# this looks kind of mixed to me? but I don't know much about mature
# immune cells -> check out

cluster_markers[[26]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# mt genes, Cd79a, Il7r, Ighm, Igll1, Sox4, Xist
# low quality late-ish BC

cluster_markers[[28]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Igkc, Cd79a, Ighm, Cd79b, Il7r, Sox4, Igll1
# also some sort of late-ish BC

cluster_markers[[30]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Siglech, Rnase6, Iglc3, Tcf4, Irf8, Sirt2, Cd74, Ighm
# maybe something antigen presenting/DC/macro/mono/BC like?
# check out

cluster_markers[[31]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Igll1, Cd79a, Ighm, Igkc, Vpreb1, Cd79b, Il7r
```

```{r umaps_lymph, fig.height = 3, fig.width = 4}
# find the cluster
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "20"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "30"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_lymph <- as.list(c(
  "Siglech", # pDC precursors
  "Mef2c", # DC development
  "Runx2", # also specifically pDC 
  "Rnase6", # dendritic cells and macrophages, 
  "Sirt2" # several immune cells, TC, DC
))

plot_list_gene <- lapply(marker_genes_list_lymph, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_lymph
plot_list_gene
```

So clusters 20 and 30 are DCs that look BC-like = pDCs?
But clusters 20 and  30 also express  some Cd8a and Cd4, Cd3e, Ccl5, 
so I'll rather call them mixed.
Or maybe pDC and some B-lineage cells cluster together
because they look so similar, either way, let's call it DC.

#### Weird clusters 

```{r markers_contamination_weird}

cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Mpo, Elane, PRtn3, Ctsg, Ms4a3, Irf8, Clec12a
# probably actually mono progenitor, check out

cluster_markers[[24]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Mpo, Elane, PRtn3, Ctsg, Ms4a3, Irf8
# probably actually mono progenitor, too, check out

cluster_markers[[27]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Cd74, H2-A2, Ms4a6c, Ccr2 --> antigen presenting/mono

# mystery
cluster_markers[[34]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Prg4, Ccn2, Dcn, Cpe, Creb5, Sparcl1, Cfh, Cd9, Aebp1, Sdc2, Spp1
# this is just a wild mix, check out?
# kind of MSC looking but...
```

```{r umaps_weird, fig.height = 3, fig.width = 4}
# find the cluster
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "15"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "24"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "27"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)

umap_base(
  sce_species[,sce_species$cluster_louvain_custom == "34"],
  "cluster_louvain_custom")+
  ggplot2::xlim(
    c(
      min(reducedDims(sce_species)$UMAP[,1]),
      max(reducedDims(sce_species)$UMAP[,1])))+
  ggplot2::ylim(
    c(
      min(reducedDims(sce_species)$UMAP[,2]),
      max(reducedDims(sce_species)$UMAP[,2])))+
  ggplot2::scale_color_manual("cluster", values = col_num)
# it is "closer" to other MSC like clusters

marker_genes_list_weird <- as.list(c(
  "Mpo", # 15: myeloid 
  "Elane", # 15: myeloid 
  "Irf8", # 15: mono  
  "Cd34", # progenitor
  "Adgrl4", # progenitor
  "Meis1", # progenitor
  "Hlf", # progenitor
  "Flt3", # progenitor
  "Cd74", # APC
  "Prg4", # unknowm MSC-looking cluster
  "Ccn2",
  "Dcn",
  "Sparcl1",
  "Aebp1",
  "Spp1"
))

plot_list_gene <- lapply(marker_genes_list_weird, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_weird
plot_list_gene

```

They don't look early enough for "early myeloid prog".
So keep them early neutro.

### Expected clusters

```{r umaps_expected, fig.height = 3, fig.width = 4}
# find the cluster
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)
sce_expected <- sce_species[,sce_species$cluster_louvain_custom %in% c(
  "4",
  "11",
  "19",
  "16",
  "12",
  "22",
  "18",
  "25",
  "34" # adding the mystery cluster
)]

umap_base(sce_expected, "cluster_louvain_custom")+
  ggplot2::scale_color_manual("cluster", values = col_num)

marker_genes_list_weird <- as.list(c(

))
```

```{r markers_adipo}
# adipo
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Cxcl12, Apoe, Cxcl14, Gas6, Kitl, Lpl, Vcam1, Hp, Snai1 are clearly adipo CAR like.

```{r markers_osteo_fibro}
# osteo-fibro?
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Dcn, Lum, Col1a1, Col1a2, Gpx3, Fmod, Mgp, Comp, Cd9
look more fibro neutro based on main markers but these
just may be the most strongly expressed so also need to check
out the patterns of possibly more lowly expressed osteo genes.

```{r markers_osteo_chondro}
# osteo chondro?
cluster_markers[[19]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Thbs4, Fmod, Mgp, Dcn, Col1a1, Col1a2, Prelp, Aspn, Ogn, Cd9, Aebp1,
Chad, Prg4.
Very chondro-ish (early)?

```{r markers_fibro_chondro}
# chondro fibro?
cluster_markers[[16]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Dcn, Gsn, Clec3b, Col3a1, Lum, Cd9, Pcolce, Tnxb
also not overtly osteo-lineage.

```{r marker_genes_list_msc}
marker_genes_list_msc <- as.list(c(
  # add some more expected genes that I know even though we had them before
  "Cxcl12", # adipo 4
  "Cxcl14",
  "Lpl",
  "Vcam1",
  "Kitl",
  "Adipoq",
  "Lepr",
  "Dcn", # fibro-osteo 11
  "Lum",
  "Col1a1",
  "Col1a2",
  "Aspn",
  "Ogn",
  "Fmod",
  "Comp",
  "Thbs2",
  "Abi3bp",
  "Fn1",
  "Pdgfra",
  "Pdgfrb",
  "Cd34",
  "Thy1",
  "Thbs4", # osteo chondro/19
  "Mgp",
  "Prelp",
  "Chad",
  "Prg4",
  "Alpl",
  "Bglap",
  "Ibsp",
  "Mmp13",
  "Sp7",
  "Spp1",
  "Postn",
  "Runx2",
  "Gsn", # chondro fibro/16
  "Clec3b",
  "Col3a1",
  "Pcolce2",
  "Tnxb",
  "Ccl11",
  "Dpt",
  "Acan"
))

plot_list_gene <- lapply(marker_genes_list_msc, function(gene){
  plot <- umap_gene(sce_expected, gene)
})
names(plot_list_gene) <- marker_genes_list_msc
plot_list_gene
```

Cluster 4 is clearly adipocytic/CAR-like. 
The only one expressing Adipoq.
But one part also expresses osteo-lineage genes.

cluster 16 is also weakly adipocytic, with light Cxcl12 and Lpl expression.
Dcn is very high, Lum a bit lower.
SPOTTY osteo expression. 
Highest Pdgfra and Cd34, as well as Thy1.
These are mostly fibro.

Cluster 11 also has some spotty adipocytic.
Dcn is mostly high, Lum is higher.
Highest Aspn, Ogn, Fmod, Alpl, 
I think a good portion of the more "Osteo" like cells are now with the
Adipo-CAR cells.
So these are mostly osteo-chondro.
Especially since a small speckle contained in this cluster are the most
mature chondro-cells, interestingly.

Cluster 19 is the strongest Thbs4, Prelp, Chad expressing cluster, 
also expresses some fibro-stuff.
But really strong chondro-like.

Cluster 34 (mystery) doesn't express most of these markers but Prg4.

```{r EC_sinus_venular}
# EC sinusoidal, venular, lymphatic?
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Plvap, Pecam1, Vwf, Cdh5, Lrg1, Emcn, Eng.
Definitely venular-ish.

```{r EC_sinus_capillary}
# EC very sinusoidal and slightly capillary
cluster_markers[[22]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Cdh5, Flt1, Pecam1, Cavin2, Esam, Cd36, Kdr, Cav1.
It does look sinusoidal like.

```{r EC_artcap}
# EC arteriolar and capillary
cluster_markers[[18]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Cdh5, Cd36, Cav1, Cavin2, Flt1, Timp4, Cxcl12, Esam, Kdr, Pecam1.
Markers look similar but tbh the Cxcl12 makes me really believe it's 
arteriolar.

```{r marker_genes_list_ec}
marker_genes_list_ec <- as.list(c(
  "Plvap", # EC sinusoidal, venular, lymphatic?
  "Fabp4",
  "Cavin2",
  "Vwf",
  "Ramp2",
  "Clec14a",
  "Cd9", 
  "Emcn",
  "Lrg1",
  "Bcam",
  "Eng",
  "Mustn1", 
  "Stab2", # add some more well-known
  "Ackr1",
  "Ephb4",
  "Selp",
  "Icam1",
  "Vcam1",
  "Sox4",
  "Flt4",
  "Tfpi",
  "Lyve1",
  "Ccl21a",
  "Esam",
  "Tm4sf1",
  "Rbp7",
  "Cd36",
  "Slc9a3r2",
  "Adgrf5",
  "Kdr",
  "Epas1",
  "Sparcl1", # EC arteriolar and capillary?
  "Hspb1",
  "Rgcc", # add more
  "Car4",
  "Hey1",
  "Sox17",
  "Efnb2"
))

plot_list_gene <- lapply(marker_genes_list_ec, function(gene){
  plot <- umap_gene(sce_expected, gene)
})
names(plot_list_gene) <- marker_genes_list_ec
plot_list_gene
```

Cluster 12 has Plvap, Vwf, highest Emcn, highest Lrg1, Ackr1, Selp, 
but not Stab2.
A handful of Lymph EC is mingled in here, too.
Sinusoidal/venular/fenestrated/H-type coded.

Cluster 22 has spotty Plvap, Vwf, Lrg1, Ackr1 and Selp, 
but highest and exclusive Stab2, and Tfpi.
On the one hand sinusoidal, on the other mixed.
I cannot really tell which one is "more" "classic" sinusoidal
so I'll call both of them that.

Cluster 18 has highest Rgcc, Car4, some Hey1 and Sox17.
These are arteriolar/capillary.

```{r peri_smc}
# peri/smc
cluster_markers[[25]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

Try the weird cluster again.

```{r weird_MSC}
# peri/smc
cluster_markers[[34]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

```{r marker_genes_list_weird_msc}
marker_genes_list_weird_msc <- as.list(c(
  "Prg4",
  "Clic5",
  "Ccn2",
  "Dcn",
  "Cpe",
  "Fbln7",
  "Gpx3",
  "Creb5",
  "Sparcl1",
  "Pcolce",
  "Cfh",
  "Cd9",
  "Serping1",
  "Aebp1",
  # add some myofibroblast markers
  "Acta2",
  "Vim",
  "Des",
  "Vcl"
))

plot_list_gene <- lapply(marker_genes_list_weird_msc, function(gene){
  plot <- umap_gene(sce_expected, gene)
})
names(plot_list_gene) <- marker_genes_list_weird_msc
plot_list_gene

```

Prg4, Cpe, Fbln7, are co-expressed with chondro.
Gpx3, Creb5, Sparcl1, Cfh with fibro.
Clic5 is co-expressed with Aart/Cap EC.
Ccn2 is an MSC gene.

Clic5 seems pre-adipocytic?
Cpe also obesity related.

Prg4 seems to be a marker for some Chondro/fibro subset at a specific
location but I don't want to give location names (like endosteal) from
scRNAseq data alone.

I just cannot tell, these express markers from all lineages, and lack
important others, so I'll just call them "unclear". 
It's a tiny cluster anyways.

# Assign

In the same order as above

```{r assign}

sce_species$custom_label <- vector(length = ncol(sce_species))

# Ery
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Ery 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Ery 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Ery 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Mature ery"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Ery 4"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "17"] <-  "Ery 5"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "23"] <-  "Ery 6"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "33"] <-  "Ery 7"

# Neutro
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Neutro 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Neutro 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Mature neutro 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Neutro 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Mature neutro 2"
  sce_species$custom_label[
  sce_species$cluster_louvain_custom == "21"] <-  "Neutro 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "29"] <-  "Mature immune, mixed 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "32"] <-  "Neutro 4"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "35"] <-  "Eosinophil"

# Baso mast 
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Baso/Mast"

# Lymphoid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "20"] <-  "Mature immune, mixed 3"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "26"] <-  "BC lineage 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "28"] <-  "BC lineage 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "30"] <-  "Mature immune, mixed 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "31"] <-  "BC lineage 3"
  
# weird clusters 
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Early Neutro 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "24"] <-  "Early Neutro 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "27"] <-  "Antigen-presenting"

# expected clusters + the last weird cluster
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Adipo/Osteo CAR"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "16"] <-  "Fibro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Osteo-Chondro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "19"] <-  "Chondro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "34"] <-  "Unclear mesenchymal"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "Sinusoidal EC 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "22"] <-  "Sinusoidal EC 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "18"] <-  "Arteriolar/capillary EC"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "25"] <-  "Peri/SMC"

sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "Adipo/Osteo CAR",
    "Osteo-Chondro",
    "Chondro",
    "Fibro",
    "Unclear mesenchymal",
    "Arteriolar/capillary EC",
    "Sinusoidal EC 1",
    "Sinusoidal EC 2",
    "Peri/SMC",
    "Ery 1",
    "Ery 2",
    "Ery 3",
    "Ery 4",
    "Ery 5",
    "Ery 6",
    "Ery 7",
    "Mature ery",
    "Early Neutro 1",
    "Early Neutro 2",
    "Neutro 1",
    "Neutro 2",
    "Neutro 3",
    "Neutro 4",
    "Mature neutro 1",
    "Mature neutro 2",
    "Mature neutro 3",
    "Eosinophil",
    "Baso/Mast",
    "BC lineage 1",
    "BC lineage 2",
    "BC lineage 3",
    "Antigen-presenting",
    "Mature immune, mixed 1",
    "Mature immune, mixed 2",
    "Mature immune, mixed 3"

    ))
sort(unique(sce_species$custom_label))
levels(sce_species$custom_label)
stopifnot(!is.na(sce_species$custom_label))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 7}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```