---
title: "Custom clustering mcas hsc"
author: "Lea Wölbert"
date: '2025-11-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])
#sce_species <- base::readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/01_mnnc/comparison/sce_mcas_hsc_Object_ID-01")

fraction_curr <- "hsc"
species_curr <- "mcas"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be
# defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mmus_hsc_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
#colors_path <- snakemake@params[["colors_path"]]
#source(snakemake@params[["colors"]])
#source("../../source/plotting.R")

col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

```{r parameters}
k_graph <- 10
resolution_louvain <- 1.05
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "hsc"){
  marker_genes <- c(
    "Aldh1a1", # HSC, Stem
    "Procr",
    "Cdkn1c",
    "Gng11",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Alcam", 
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Socs2",
    "Pbx1",
    "Slamf1",
    "Txnip",
    "Msi2",
    "Adgrl4", # MPP, progenitors, activation
    "Cd34", 
    "Cdk4",
    "Sox4",
    "Spi1", 
    "Cebpa", # Myeloid
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Elane", # Neutro
    "Cebpe", 
    "Gfi1", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", # Mono
    "F13a1",
    "Irf8",
    "Ly86",
    "Ms4a6c",
    "Flt3", # Lymphoid
    "Satb1",
    "Dntt",
    "Cd79b", 
    "Il7r",
    "Rag1",
    "Cpa3", # Mast
    "Hdc", 
    "Ms4a2",
    "Lmo4", # Baso
    "Mcpt8", 
    "Prss34",
    "Gata2", # also Mk, Eo Baso and HSC
    "Itga2b", # Mk
    "Pf4",
    "Blvrb", # Mk Ery
    "Gata1",
    "Zfpm1",
    "Epor", # Ery
    "Gfi1b",
    "Hba-a1",
    "Klf1",
    "Dctpp1", # CC entry
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa", # CC
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}

```

```{r genes_short}

if(fraction_curr == "hsc"){
  marker_genes_short <- c(
    "Alcam", 
    "Cdkn1c",
    "Mecom",
    "Mpl",
    "Meis1",
    "Cd34", 
    "Spi1", 
    "Mpo",
    "Elane",
    "Cebpa", 
    "Cebpe", 
    "S100a9",
    "Wfdc21",
    "Ccr2", 
    "Irf8",
    "Ly86",
    "Cd79b", 
    "Dntt",
    "Il7r",
    "Satb1",
    "Cpa3", 
    "Hdc", 
    "Lmo4", 
    "Prss34",
    "Itga2b", 
    "Pf4",
    "Zfpm1",
    "Epor",
    "Hba-a1",
    "Klf1",
    "Hells",
    "Lig1",
    "Cenpa", 
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}

```

#### Heatmap

```{r heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 6}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)
```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

Check all clusters for marker genes since for the exotic species,
there may be genes we didn't anticipate.
For each probable cell type based on the heatmap.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### HSC

```{r markers_hsc}
cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

### MPP

There are many different MPP/early prog-like clusters.
Checking them out.

```{r markers_mpp}
# still kind of HSC-looking and not very activated
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# even more HSC-looking but a bit more lympho?
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# a bit more mature looking than 13
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# activated MPP like, much less HSC-like
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# just mostly activated, some MPP, not much else?
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

### Myeloid

```{r markers_myeloid}
# looks early myeloid includes mono
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# earlier myeloid +  
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# this cluster looks myeloid early to late
# also some mk ery expression
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# mono and some antigen presenting?
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

### Lymphoid

```{r markers_lymph}
# lymphoid
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

### Mk ery baso

```{r markers_mkerbm}
# roughly mk and ery
cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# purely ery
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
# mk and baso mast
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

### cycling

```{r markers_cycling}
cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 20)
```

# Assign

```{r assign}

# HSC
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "HSC"

# MPP
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "MPP, early 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "MPP, early 2"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "MPPs"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "MPP, activated 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "MPP, activated 2"  

# myeloid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Myeloid, early"  
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Myeloid, later"  
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Myeloid, mature mono"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Myeloid, mature neutro"

# lymphoid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Lymphoid"

# mk ery baso mast
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Mk/Ery"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Baso/Mast"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Erythroid"

# cycling
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Cycling"

sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "HSC",
    "MPP, early 1",
    "MPP, early 2",
    "MPPs",
    "MPP, activated 1",
    "MPP, activated 2",
    "Myeloid, early",
    "Myeloid, later",
    "Myeloid, mature neutro",
    "Myeloid, mature mono",
    "Lymphoid",
    "Baso/Mast",
    "Mk/Ery",
    "Erythroid",
    "Cycling"
  )
)

stopifnot(!is.na(sce_species$custom_label))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 5}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

