---
title: "Custom clustering comparison"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report is to compare the custom clusterings to each other based on 
the selected marker genes as well as some observed marker genes.
To make sure that a unified manner of annotation was used and very similar
populations didn't get different names due to time having passed in-between.

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

#todo input snakemake later

```{r load_objects}
sce_mmus_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds")

sce_mcas_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds")

sce_mspr_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds")

sce_mcar_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds")
```

## Control gene expression

Compare the expression of marker genes (either from the literature, see 
supplementary table of the corresponding publication) or picked from the
marker gene lists of individual species and cell types.

```{r marker_genes}
# this list is a bit longer than the one in the custom assignment .Rmd scripts
# so it can cover all species
# it's also not JUST to get an idea, but to better show cell type seperation. 
  marker_genes_list <- list(
    "CAR" = c( # classic niche 
      "Cxcl14", 
      "Ebf3", # niche cells https://doi.org/10.1101/gad.314013.118
      "Cxcl12", # also arteriolar EC
      "Kitl" # also arteriolar EC
      ),
    "Adipo" = c(
      "Adipoq",
      "Hp",
      "Lepr", 
      "Lpl",
      "Cebpa",
      "Pappa"  # adipocyte marker in MALP https://doi.org/10.7554/eLife.54695
      ),
    "Osteo" = c( # from early/CAR to more osteocyte-ish (approx)
      "Limch1", # Osteo-CAR https://doi.org/10.1016/j.jbc.2024.107158
      "Angpt4", # Osteo-CAR https://doi.org/10.1016/j.jbc.2024.107158
      "Wif1", # Osteo-CAR https://doi.org/10.1016/j.jbc.2024.107158
      "Spp1", # pre-osteoblast https://doi.org/10.1016/j.jbc.2024.107158
      "Mmp13", # pre-osteoblast https://doi.org/10.1016/j.jbc.2024.107158
      "Alpl", # Osteo-CAR https://doi.org/10.1038/s41556-019-0439-6
      "Slit2", # pre-osteoblast https://doi.org/10.1016/j.jbc.2024.107158
      "Sp7", # Osteo-CAR and osteoblast https://doi.org/10.1038/s41556-019-0439-6
      "Pth1r", # osteo ST1
      "Bglap", # osteo https://doi.org/10.1016/j.jbc.2024.107158
      "Bglap2", # osteoblasts https://doi.org/10.7554/eLife.54695
      "Bglap3", # 
      "Mepe", # osteocyte https://doi.org/10.1016/j.jbc.2024.107158
      "Ibsp", # osteoblasts https://doi.org/10.7554/eLife.54695
      "Phex" # osteocyte https://doi.org/10.1016/j.jbc.2024.107158
      ),
    "Chondro/Osteo/fibro/early mixed" = c(
      "Tnc", # osteo-X https://doi.org/10.1016/j.jbc.2024.107158
      "Ptgis", # osteo ST1
      "Postn", # osteo-X https://doi.org/10.1016/j.jbc.2024.107158
      "Ptn", # osteo-X https://doi.org/10.1016/j.jbc.2024.107158
      "Comp", # fibro-chondro ST1
      "Ogn", # in branch point cekk type to osteogenic state https://doi.org/10.1186/s13073-023-01224-0
      "Aspn", # osteo-X https://doi.org/10.1016/j.jbc.2024.107158
      "Gsn", # several undefined MSC populations 
      "Apod", # several undefined MSC populations 
      "Tnn" # osteo-X https://doi.org/10.1016/j.jbc.2024.107158
      ),
    "Early Fibro mixed" = c(
      "Dcn", # fibro ST1
      "Lum", # fibro ST1
      "Dpt", # fibro https://doi.org/10.1016/j.jbc.2024.107158
      "Clec3b", # fibro https://doi.org/10.1016/j.jbc.2024.107158
      "Cd34", # also arteriolar EC
      "Thy1", # fibro, msc, related to osteogenesis https://doi.org/10.1096/fj.201701379R
      "Pdgfra", # fibro ST1
      "Pdgfrb" # 
      ),
    "Chondro early/other/mixed" = c(
      "Fmod", # chondro ST1
      "Chad", # chondro ST1
      "Prelp", # chondro ST1
      "Mgp", # chondro ST1
      "Crispld1",  # art ch https://doi.org/10.1016/j.jbc.2024.107158
      "Cytl1" # art ch https://doi.org/10.1016/j.jbc.2024.107158
      ),
    "Articular cartilage/Synovial fibro" = c(  # these genes can be articular chondros or also early chondros or also early synovial fibros.
      "Htra4", # see mspr
      "Prg4", # art ch and syn fibro https://doi.org/10.1016/j.jbc.2024.107158
      "Creb5" # see mspr
      ),   
    "Chondro late" = c(  
      "Sox9", # chondro https://doi.org/10.1016/j.jbc.2024.107158
      "Snorc",  # chondro https://doi.org/10.1016/j.jbc.2024.107158
      "Acan"
      ),
    "Pan EC" = c(
      "Cdh5",
      "Pecam1"
      ),
    "EC other" = c( # COULD be interesting but mostly similarly expressed
      "Emcn",  # higher in H-type vessels
      "Kdr", # venous/sinusoid/unclear
      "Cd36"
      ),
   "Capillary" = c(
      "Car4",
      "Timp4",
      "Rgcc"
      ),
    "Arteriolar" = c(
      "Efnb2", 
      "Hey1", # have also seen as H type
      "Sox17" # have also seen as H type
      ),
    "Transitory/venous/fenestrated/other" = c(
      "Vwf", # venous or H type?
      "Ehd4", # angiogenesis or transitory?
      "Plvap" # fenestration
    ),
    "Sinusoidal/venous expressed" = c(
      "Selp",  # expressed in sinusoid
      "Icam1", # expressed in sinsuoid
      "Vcam1",# expressed in sinsuoid
      "Nr2f2", # venous identity
      "Ackr1", # expressed in sinusoid
      "Lrg1", # expressed in sinsuoid
      "Sele"  # expressed in sinusoid
    ),
    "Sinusoidal classic" = c(
      "Stab2", # classic sinusoid
      "Stab1", # classic sinusoid
      "Flt4", # classic sinusoid 
      "Dnase1l3", # sinusoid https://doi.org/10.1038/s41467-018-04726-3
      "Tfpi", # classic sinusoid
      "Sirpa", # sinusoid https://doi.org/10.1038/s41467-018-04726-3
      "C1qtnf1" # sinusoid https://doi.org/10.1038/s41467-018-04726-3
      ),
    "Lymphatic EC" = c(
      "Lyve1", # also sinusoid
      "Ccl21a"
      ),
    "Peri/SMC" = c(
      "Rgs5",
      "Acta2",
      "Tagln",
      "Steap4"
      ),
    "Contamination ery early/middle" = c(
      "Klf1",
      "Hmbs",
      "Hba-a1",
      "Hbb-bs",
      "Gypa"
      ),
    "Contamination ery late" = c(
      "Ermap",
      "Kel",
      "Sox6",
      "Slc4a1"
      ),
    "Contamination myeloid early" = c(
      "Mpo",
      "Elane",
      "Prtn3"
      ),
    "Contamination mono/macro" = c(
      "Irf8",
      "Ccr2",
      "F13a1",
      "Mpeg1"
      ),
    "Contamination neutro middle" = c(
      "S100a8",
      "S100a9",
      "Wfdc21",
      "Trem1",
      "Cebpe",
      "Camp",
      "Ngp",
      "Retnlg"
      ),
    "Contamination neutro late" = c(
      "Thbs1",
      "Ccl6"
      ),
    "Contamination eosinophil" = c(
      "Ear1",
      "Ear6",
      "Prg3"
    ),
    "Contamination baso mast" = c(
      "Prss34",
      "Gata2",
      "Hdc",
      "Cpa3"
      ),
    "Contamination NK lineage" = c(
      "Ccl5",
      "Nkg7"
      ), 
    "Contamination TC lineage" = c(
      "Cd8a",
      "Cd4",
      "Cd3e",
      "Cd3d",
      "Trbc1"
      ),
    "Contamination lymph early" = c(
      "Sox4",
      "Satb1",
      "Dntt",
      "Il7r"
      ),
    "Contamination BC lineage" = c(
      "Vpreb1",
      "Vpreb3",
      "Cd79a",
      "Cd79b",
      "Igha",
      "Jchain",
      "Igkc"
      ), 
    "Contamination Antigen-presenting" = c(
      "H2-Aa",
      "H2-Ab1",
      "Cd74",
      "Ctss"
      ),
    "Contamination DC" = c(
      "Siglech",
      "Cd7",
      "Runx2" # also osteo
      ), 
    "Contamination skeletal muscle" = c(
      "Acta1",
      "Tnnt3",
      "Tnnc2",
      "Tpm2"
    ),
    "CC entry/activation" = c(
      "Lig1",
      "Mcm7",
      "Hells",
      "Mcm6"
      ),
    "Cell cycle S phase" = c(
      "Aurkb",
      "Cdk1"
      ),
    "Cell cycle M phase" = c(
      "Cdc20",
      "Ccnb2",
      "Knstrn"
      ),
    "Cell cycle general" = c(
      "Mki67",
      "Lockd",
      "Cenpa" 
      )

  )
```

Make a shorter marker gene lists based on my observations
from the annotation .Rmds.

```{r best_marker_genes_list}
best_marker_genes <- c(
  "Cxcl12", # niche
  "Kitl",
  "Adipoq", # adipo
  "Pappa",
  "Limch1", # osteo-CAR
  "Angpt4",
  "Mmp13", # osteo
  "Alpl",
  "Bglap",
  "Postn", # balanced/mixed
  "Aspn", 
  "Apod", # Mixed/Fibro
  "Gsn", 
  "Clec3b", # Fibro
  "Thy1",
  "Cd34",
  "Pdgfra",
  "Mgp", # Mixed/chondro
  "Fmod", 
  "Chad", # mid chondro
  "Sox9", # late chondro
  "Snorc", 
  "Prg4", # synovial/articular
  "Creb5", 
  "Cdh5", # pan-endo 
  "Pecam1", 
  "Car4", # capillaty
  "Timp4",
  "Rgcc",
  "Efnb2", # art
  "Hey1",
  "Sox17",
  "Vwf", # transitory/venous
  "Ehd4",
  "Plvap",
  "Selp", # transitory sinusoidal/venous
  "Nr2f2",
  "Ackr1", 
  "Stab2", # pure sinusoidal
  "Stab1",
  "Flt4",
  "Lyve1", # lymphatic
  "Ccl21a",
  "Rgs5", # peri SMC
  "Acta2",
  "Tagln",
  "Klf1", # ery
  "Hba-a1",
  "Ermap",
  "Kel",
  "Elane", # early-ish myeloid
  "Prtn3",
  "Ccr2", # mono
  "Mpeg1", # macro
  "S100a8", # early-ish neutro
  "s100a9",
  "Cebpe", # mid-ish neutro
  "Ngp",
  "Rentlg", # late neutro
  "Ccl6", # late neutro and eo
  "Ear1", # eo
  "Prg3",
  "Prss34",
  "Cpa3",
  "Nkg7", # Nk
  "Ccl5",
  "Cd8a", # TC
  "Cd4",
  "Dntt", # early lymph
  "Vpreb1", # mid-ish lymph
  "Vpreb3", 
  "Cd79a", 
  "Igha", # later lymph
  "Jchain",
  "H2-Aa", # antige,-presenting
  "H2-Ab1",
  "Cd74",
  "Siglech", # other immune
  "Acta1", # skeletal muscle
  "Tnnt3",
  "Tnnc2",
  "Lig1", # CC entry
  "Hells",
  "Aurkb", # S phase
  "Cdk1",
  "Ccnb2", # M phase
  "Knstrn",
  "Mki67",
  "Lockd",
  "Cenpa" 

)
```

#### Heatmap

```{r make_heatmap_function}
make_heatmap <- function(sce_species, marker_genes){

  # get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5,
      hjust = 1))
}
```

```{r heatmap_mes, fig.height = 9, fig.width = 9}

marker_genes_mes <- unlist(marker_genes_list[1:8])

make_heatmap(sce_mmus_str, marker_genes_mes)
make_heatmap(sce_mcas_str, marker_genes_mes)
make_heatmap(sce_mspr_str, marker_genes_mes)
make_heatmap(sce_mcar_str, marker_genes_mes)
```

```{r heatmap_ec, fig.height = 6, fig.width = 9}
# and peri SMC
marker_genes_ec <- unlist(marker_genes_list[9:17])

make_heatmap(sce_mmus_str, marker_genes_ec)
make_heatmap(sce_mcas_str, marker_genes_ec)
make_heatmap(sce_mspr_str, marker_genes_ec)
make_heatmap(sce_mcar_str, marker_genes_ec)
```

```{r heatmap_ery, fig.height = 4, fig.width = 9}

marker_genes_ery <- unlist(marker_genes_list[18:19])

make_heatmap(sce_mmus_str, marker_genes_ery)
make_heatmap(sce_mcas_str, marker_genes_ery)
make_heatmap(sce_mspr_str, marker_genes_ery)
make_heatmap(sce_mcar_str, marker_genes_ery)
```

```{r heatmap_myeloid, fig.height = 6, fig.width = 9}

marker_genes_myeloid <- unlist(marker_genes_list[20:25])

make_heatmap(sce_mmus_str, marker_genes_myeloid)
make_heatmap(sce_mcas_str, marker_genes_myeloid)
make_heatmap(sce_mspr_str, marker_genes_myeloid)
make_heatmap(sce_mcar_str, marker_genes_myeloid)
```

```{r heatmap_lym, fig.height = 5.3, fig.width = 9}

marker_genes_lym <- unlist(marker_genes_list[26:31])

make_heatmap(sce_mmus_str, marker_genes_lym)
make_heatmap(sce_mcas_str, marker_genes_lym)
make_heatmap(sce_mspr_str, marker_genes_lym)
make_heatmap(sce_mcar_str, marker_genes_lym)
```

```{r heatmap_immu, fig.height = 9, fig.width = 9}

marker_genes_immu <- unlist(marker_genes_list[20:31])

make_heatmap(sce_mmus_str, marker_genes_immu)
make_heatmap(sce_mcas_str, marker_genes_immu)
make_heatmap(sce_mspr_str, marker_genes_immu)
make_heatmap(sce_mcar_str, marker_genes_immu)
```

```{r heatmap_all_marker_genes_list, fig.height = 21, fig.width = 9}

marker_genes_all <- unlist(marker_genes_list)

make_heatmap(sce_mmus_str, marker_genes_all)
make_heatmap(sce_mcas_str, marker_genes_all)
make_heatmap(sce_mspr_str, marker_genes_all)
make_heatmap(sce_mcar_str, marker_genes_all)
```

```{r best_marker_genes, fig.height = 11, fig.width = 9}
 
make_heatmap(sce_mmus_str, best_marker_genes)
make_heatmap(sce_mcas_str, best_marker_genes)
make_heatmap(sce_mspr_str, best_marker_genes)
make_heatmap(sce_mcar_str, best_marker_genes)

#best_marker_genes[duplicated(best_marker_genes)]
```

```{r session_info}
utils::sessionInfo()
```
