---
title: "Custom clustering report - gene expression"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report is to compare the custom clusterings to each other based on 
the selected marker genes as well as some observed marker genes.
Just to make sure that a unified manner of annotation was used.

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file = snakemake@params[["custom_utils"]])
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r load_objects}
sce_mmus_hsc <- readRDS(snakemake@input[["mmus_sce"]])
sce_mcas_hsc <- readRDS(snakemake@input[["mcas_sce"]])
sce_mspr_hsc <- readRDS(snakemake@input[["mspr_sce"]])
sce_mcar_hsc <- readRDS(snakemake@input[["mcar_sce"]])
```

## Control gene expression

Compare the expression of marker genes (either from the literature, see 
supplementary table of the corresponding publication) or picked from the
marker gene lists of individual species and cell types.

```{r marker_genes}
# this list is a bit longer than the one in the custom assignment .Rmd scripts
# so it can cover all species
# it's also not JUST to get an idea, but to better show cell type seperation. 

# from custom utils
marker_genes_list <- marker_genes_list_hsc
```

Make a shorter marker gene lists based on my observations
from the annotation .Rmds.

```{r best_marker_genes_list}
# from custom utils
best_marker_genes <- best_marker_genes_hsc
```

#### Heatmap

```{r make_heatmap_function}
make_heatmap <- function(sce_species, marker_genes){

  # get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
}
```

```{r heatmap_hsc, fig.height = 7, fig.width = 6}

marker_genes_hsc <- unlist(marker_genes_list[1:2])

make_heatmap(sce_mmus_hsc, marker_genes_hsc)
make_heatmap(sce_mcas_hsc, marker_genes_hsc)
make_heatmap(sce_mspr_hsc, marker_genes_hsc)
make_heatmap(sce_mcar_hsc, marker_genes_hsc)
```

```{r heatmap_myeloid, fig.height = 6, fig.width = 6}

marker_genes_myeloid <- unlist(marker_genes_list[3:6])

make_heatmap(sce_mmus_hsc, marker_genes_myeloid)
make_heatmap(sce_mcas_hsc, marker_genes_myeloid)
make_heatmap(sce_mspr_hsc, marker_genes_myeloid)
make_heatmap(sce_mcar_hsc, marker_genes_myeloid)
```

```{r heatmap_lymph, fig.height = 5, fig.width = 6}

marker_genes_lymph <- unlist(marker_genes_list[7:9])

make_heatmap(sce_mmus_hsc, marker_genes_lymph)
make_heatmap(sce_mcas_hsc, marker_genes_lymph)
make_heatmap(sce_mspr_hsc, marker_genes_lymph)
make_heatmap(sce_mcar_hsc, marker_genes_lymph)
```

```{r heatmap_mkerybm, fig.height = 8, fig.width = 6}

marker_genes_mkebm <- unlist(marker_genes_list[10:13])

make_heatmap(sce_mmus_hsc, marker_genes_mkebm)
make_heatmap(sce_mcas_hsc, marker_genes_mkebm)
make_heatmap(sce_mspr_hsc, marker_genes_mkebm)
make_heatmap(sce_mcar_hsc, marker_genes_mkebm)
```

```{r heatmap_cycling, fig.height = 4, fig.width = 6}

marker_genes_cycling <- unlist(marker_genes_list[14:17])

make_heatmap(sce_mmus_hsc, marker_genes_cycling)
make_heatmap(sce_mcas_hsc, marker_genes_cycling)
make_heatmap(sce_mspr_hsc, marker_genes_cycling)
make_heatmap(sce_mcar_hsc, marker_genes_cycling)
```

```{r heatmap_all_marker_genes_list, fig.height = 20, fig.width = 6}

marker_genes_all <- unlist(marker_genes_list)

make_heatmap(sce_mmus_hsc, marker_genes_all)
make_heatmap(sce_mcas_hsc, marker_genes_all)
make_heatmap(sce_mspr_hsc, marker_genes_all)
make_heatmap(sce_mcar_hsc, marker_genes_all)
```

```{r best_marker_genes, fig.height = 8, fig.width = 6}

make_heatmap(sce_mmus_hsc, best_marker_genes)
make_heatmap(sce_mcas_hsc, best_marker_genes)
make_heatmap(sce_mspr_hsc, best_marker_genes)
make_heatmap(sce_mcar_hsc, best_marker_genes)
```

```{r session_info}
utils::sessionInfo()
```