---
title: "Custom clustering comparison"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This report is to compare the custom clusterings to the original merged
clustering.

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r colors, message = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])
```

```{r libraries, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(ggalluvial, quietly = TRUE)
```

```{r load_objects}
sce_mmus_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mmus_str_custom.rds")
sce_mcas_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcas_str_custom.rds")
sce_mspr_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mspr_str_custom.rds")
sce_mcar_str <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/custom/sce_mcar_str_custom.rds")
```

```{r load_merged}
sce_str_all_cells <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/02_clst/sce_str-02")

# remove unnecessary assays because the objects become too large to load
assays(sce_str_all_cells) <- list(
  counts = assays(sce_str_all_cells)$counts,
  logcounts = assays(sce_str_all_cells)$logcounts
)
reducedDims(sce_str_all_cells) <- list(
  UMAP = reducedDims(sce_str_all_cells)$UMAP
)
gc()
```

```{r cluster_labels}
# add all labels to the object that contains all cells

sce_str_all_cells$custom_label <- vector(length = ncol(sce_str_all_cells))

sce_mmus_str$custom_label <- unfactor(sce_mmus_str$custom_label)
sce_mcas_str$custom_label <- unfactor(sce_mcas_str$custom_label)
sce_mspr_str$custom_label <- unfactor(sce_mspr_str$custom_label)
sce_mcar_str$custom_label <- unfactor(sce_mcar_str$custom_label)

sce_str_all_cells$custom_label[match(
  colnames(sce_mmus_str), colnames(sce_str_all_cells))] <- sce_mmus_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcas_str), colnames(sce_str_all_cells))] <- sce_mcas_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mspr_str), colnames(sce_str_all_cells))] <- sce_mspr_str$custom_label
sce_str_all_cells$custom_label[match(
  colnames(sce_mcar_str), colnames(sce_str_all_cells))] <- sce_mcar_str$custom_label

rm(
  sce_mmus_str,
  sce_mcas_str,
  sce_mspr_str,
  sce_mcar_str)
gc()
```

```{r load_anno}
sce_str_anno <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data_freeze/05_annotated_objects/sce_str.rds")

sce_str_all_cells$celltypes <- vector(length = ncol(sce_str_all_cells))
sce_str_all_cells$celltypes <- sce_str_anno$celltypes[
  match(colnames(sce_str_all_cells), colnames(sce_str_anno))]

rm(sce_str_anno)
gc()

# i have removed some cells from the finally annotated object, so those
# that are missing from that object are not called "removed"
levels(sce_str_all_cells$celltypes) <- c(
  levels(sce_str_all_cells$celltypes), 
  "removed")
sce_str_all_cells$celltypes[is.na(sce_str_all_cells$celltypes)] <- "removed"

# randomize the barcodes for nicer visualisation
sce_str_all_cells <- sce_str_all_cells[,base::sample(
  x = c(1:ncol(sce_str_all_cells)), 
  size = ncol(sce_str_all_cells),
  replace = FALSE)]

print(sort(unique(sce_str_all_cells$custom_label)))
```

## Basics

```{r factor_and_colors}

# for nicer handling factor the custom labels, too
sce_str_all_cells$custom_label <- factor(
  sce_str_all_cells$custom_label, 
  levels = c(
    "Adipo-CAR",
    "Adipo/Osteo CAR",
    "Osteo",
    
    "Balanced mesenchymal",

    "Fibro",
    "Fibro/Chondro (more Fibro)",
    "Early chondro",
    "Late chondro and Osteo",
    "Late chondro",

    "Articular/synovial progenitor-like",

    "Capillary EC",
    "Capillary EC 1",
    "Capillary EC 2",
    "Arteriolar/capillary EC",
    "Arteriolar EC",

    "Transitory/Venous EC",
    "Mixed arteriolar/sinus EC",
    "Mixed ven/sinus/lymph EC",
    "Sinusoidal EC",

    "Lymphatic EC",

    "Peri/SMC",

    "Erythroid",
    "Ery 1",
    "Ery 2",
    "Ery 3",
    "Ery 4",
    "Ery 5",
    "Ery 6",
    "Ery 7",
    "Mature ery",

    "Mono/Macro lineage 1",
    "Mono/Macro lineage 2, cycling",
    "Mono/Macro lineage 3",
    "Mono/Macro lineage 4",
    "Mono/Macro lineage 5",

    "Early Neutro 1",
    "Early Neutro 2",

    "Neutro 1",
    "Neutro 2",
    "Neutro 3",
    "Neutro 4",

    "Mature neutro 1",
    "Mature neutro 2",
    "Mature neutro 3",

    "Nk/TC",
    "Eosinophil",
    "Eo/Baso/Mast",
    "Baso/Mast",

    "BC lineage 1",
    "BC lineage 2",
    "BC lineage 3",
    "BC lineage 4",
    "BC lineage 5",
    "BC lineage 6",
    "BC lineage 7",


    "Low quality adipo-CAR",
    "Low quality osteo-chondro",

    "Mature immune, mixed 1",
    "Mature immune, mixed 2",
    "Antigen-presenting",
    "Antigen-presenting 1",
    "Antigen-presenting 2",

    "Mixed lymphatic EC/mature immune",
    "Immune/skeletal muscle, mixed",
    "Skeletal muscle",
    "Low quality skeletal muscle"
))

col_custom <- c(
    "Adipo-CAR" = "#1c6c6eff",
    "Adipo/Osteo CAR" = "#35abadff",
    "Osteo" = "#13ddc2ff",
    
    "Balanced mesenchymal" = "#35ad5bff",

    "Fibro" = "#6b9508ff",
    "Fibro/Chondro (more Fibro)" = "#bcf537ff",
    "Early chondro" = "#787a0eff",
    "Late chondro and Osteo" = "#0e7a3fff",
    "Late chondro" = "#48490cff",

    "Articular/synovial progenitor-like" = "#6ab98dff",

    "Capillary EC" = "#9eabffff",
    "Capillary EC 1" = "#5f6cc1ff",
    "Capillary EC 2" = "#3041b0ff",
    "Arteriolar/capillary EC" = "#70abfeff",
    "Arteriolar EC" = "#277dd4ff",

    "Transitory/Venous EC" = "#3e2289ff",
    "Mixed arteriolar/sinus EC" = "#6b5d92ff",
    "Mixed ven/sinus/lymph EC" = "#9572f3ff",
    "Sinusoidal EC" = "#22076eff",

    "Lymphatic EC" = "#4f7389ff",

    "Peri/SMC" = "#aa00ffff",

    "Erythroid" = "#ff005dff",
    "Ery 1" = "#e2407cff",
    "Ery 2" = "#ff92baff",
    "Ery 3" = "#a2003bff",
    "Ery 4" = "#610024ff",
    "Ery 5" = "#983056ff",
    "Ery 6" = "#ad5374ff",
    "Ery 7" = "#b87f94ff",
    "Mature ery" = "#b04747ff",

    "Mono/Macro lineage 1" = "#ff5900ff",
    "Mono/Macro lineage 2, cycling" = "#ff9257ff",
    "Mono/Macro lineage 3" = "#bd754dff",
    "Mono/Macro lineage 4" = "#ff5900ff",
    "Mono/Macro lineage 5" = "#873001ff",

    "Early Neutro 1" = "#ff8c00ff",
    "Early Neutro 2" = "#cf7f1dff",

    "Neutro 1" = "#faac4dff",
    "Neutro 2" = "#ffc986ff",
    "Neutro 3" = "#caa87eff",
    "Neutro 4" = "#9c784bff",

    "Mature neutro 1" = "#573c1bff",
    "Mature neutro 2" = "#795c39ff",
    "Mature neutro 3" = "#4b2f0dff",

    "Nk/TC" = "#b34848ff",
    "Eosinophil" = "#5f111eff",
    "Eo/Baso/Mast" = "#ff002bff",
    "Baso/Mast"  = "#9c001aff",

    "BC lineage 1" = "#ffcc00ff",
    "BC lineage 2" = "#ffe479ff",
    "BC lineage 3" = "#c3af62ff",
    "BC lineage 4" = "#958542ff",
    "BC lineage 5" = "#eeff00ff",
    "BC lineage 6" = "#d9e07bff",
    "BC lineage 7" = "#9ea534ff",


    "Low quality adipo-CAR" = "#425555ff",
    "Low quality osteo-chondro" = "#506459ff",

    "Mature immune, mixed 1" = "#afa683ff",
    "Mature immune, mixed 2" = "#7b7660ff",
    "Antigen-presenting" = "#8b7467ff",
    "Antigen-presenting 1" = "#655348ff",
    "Antigen-presenting 2" = "#4f4139ff",

    "Mixed lymphatic EC/mature immune" = "#47555eff",
    "Immune/skeletal muscle, mixed" = "#a8a8a8ff",
    "Skeletal muscle" = "#606060ff",
    "Low quality skeletal muscle" =  "#191919ff"
)

stopifnot(!is.na(sce_str_all_cells$custom_label))
```

```{r tables}
table(
  sce_str_all_cells$celltypes,
  sce_str_all_cells$custom_label)

table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcas"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcas"])

table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mspr"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mspr"])

table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcar"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcar"])
```

## Comparison

### UMAPs

```{r umap_all_custom, fig.height = 4, fig.width = 4}
umap_base(sce_str_all_cells, "custom_label")+
  ggplot2::scale_color_manual(values = col_custom)
```

```{r umap_all_orig, fig.height = 4, fig.width = 4}
umap_base(sce_str_all_cells, "celltypes")+
  ggplot2::scale_color_manual(values = col_cts_str)
```

```{r umap_spec_custom, fig.height = 8, fig.width = 8}
umap_base(sce_str_all_cells, "custom_label")+
  ggplot2::scale_color_manual(values = col_custom)+
  ggplot2::facet_wrap(facets = vars(sce_str_all_cells$Species_ID))
```

```{r umap_spec_orig, fig.height = 8, fig.width = 8}
umap_base(sce_str_all_cells, "celltypes")+
  ggplot2::scale_color_manual(values = col_cts_str)+
  ggplot2::facet_wrap(facets = vars(sce_str_all_cells$Species_ID))
```

### Tiles

```{r mmus_tiles}
mat_mmus <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

for(i in 1:ncol(mat_mmus)){
  mat_mmus[,i] <- mat_mmus[,i]/base::sum(mat_mmus[,i])
}
mmus_df <- mat_mmus %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mmus_df$Var2 <- factor(
  mmus_df$Var2, 
  levels = rev(levels(sce_str_all_cells$custom_label)[
    levels(sce_str_all_cells$custom_label) %in% unique(unfactor(mmus_df$Var2))]))

ggplot2::ggplot(
  mmus_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5, 
      hjust = 1
    )
  )
```

```{r mcas_tiles}
mat_mcas <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcas"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcas"])

for(i in 1:ncol(mat_mcas)){
  mat_mcas[,i] <- mat_mcas[,i]/base::sum(mat_mcas[,i])
}
mcas_df <- mat_mcas %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcas_df$Var2 <- factor(
  mcas_df$Var2, 
  levels = rev(levels(sce_str_all_cells$custom_label)[
    levels(sce_str_all_cells$custom_label) %in% unique(unfactor(mcas_df$Var2))]))

ggplot2::ggplot(
  mcas_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5, 
      hjust = 1
    )
  )
```

```{r mspr_tiles}
mat_mspr <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mspr"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mspr"])

for(i in 1:ncol(mat_mspr)){
  mat_mspr[,i] <- mat_mspr[,i]/base::sum(mat_mspr[,i])
}
mspr_df <- mat_mspr %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mspr_df$Var2 <- factor(
  mspr_df$Var2, 
  levels = rev(levels(sce_str_all_cells$custom_label)[
    levels(sce_str_all_cells$custom_label) %in% unique(unfactor(mspr_df$Var2))]))

ggplot2::ggplot(
  mspr_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5, 
      hjust = 1
    )
  )
```

```{r mcar_tiles}
mat_mcar <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcar"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcar"])

for(i in 1:ncol(mat_mcar)){
  mat_mcar[,i] <- mat_mcar[,i]/base::sum(mat_mcar[,i])
}
mcar_df <- mat_mcar %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcar_df$Var2 <- factor(
  mcar_df$Var2, 
  levels = rev(levels(sce_str_all_cells$custom_label)[
    levels(sce_str_all_cells$custom_label) %in% unique(unfactor(mcar_df$Var2))]))

ggplot2::ggplot(
  mcar_df, 
  aes(x = Var1, 
      y = Var2,
      fill = Freq))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::theme(
    axis.text.x = element_text(
      angle = 90,
      vjust = 0.5, 
      hjust = 1
    )
  )
```

### Alluvial

```{r mmus_alluv, fig.height = 9, fig.width = 8}
mat_mmus <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mmus"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mmus"])

mmus_df <- mat_mmus %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mmus_df$Var2_rev <- mmus_df$Var2
mmus_df$Var2_rev <- factor(
  mmus_df$Var2_rev, 
  levels = rev(levels(mmus_df$Var2)))

mmus_df2 <- mmus_df %>%
  dplyr::mutate(
    Var1_u = fct_inorder(paste0("orig ", Var1)),
    Var2_u = fct_inorder(paste0("custom ", Var2_rev))
  )

ggplot2::ggplot(data = mmus_df2,
                aes(axis1 = Var1_u,
                    axis2 = Var2_u, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1_u))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r mcas_alluv, fig.height = 9, fig.width = 8}
mat_mcas <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcas"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcas"])

mcas_df <- mat_mcas %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcas_df$Var2_rev <- mcas_df$Var2
mcas_df$Var2_rev <- factor(
  mcas_df$Var2_rev, 
  levels = rev(levels(mcas_df$Var2)))

mcas_df2 <- mcas_df %>%
  dplyr::mutate(
    Var1_u = fct_inorder(paste0("orig ", Var1)),
    Var2_u = fct_inorder(paste0("custom ", Var2_rev))
  )

ggplot2::ggplot(data = mcas_df2,
                aes(axis1 = Var1_u,
                    axis2 = Var2_u, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1_u))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r mspr_alluv, fig.height = 9, fig.width = 8}
mat_mspr <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mspr"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mspr"])

mspr_df <- mat_mspr %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mspr_df$Var2_rev <- mspr_df$Var2
mspr_df$Var2_rev <- factor(
  mspr_df$Var2_rev, 
  levels = rev(levels(mspr_df$Var2)))

mspr_df2 <- mspr_df %>%
  dplyr::mutate(
    Var1_u = fct_inorder(paste0("orig ", Var1)),
    Var2_u = fct_inorder(paste0("custom ", Var2_rev))
  )

ggplot2::ggplot(data = mspr_df2,
                aes(axis1 = Var1_u,
                    axis2 = Var2_u, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1_u))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r mcar_alluv, fig.height = 9, fig.width = 8}
mat_mcar <- table(
  sce_str_all_cells$celltypes[sce_str_all_cells$Species_ID == "mcar"],
  sce_str_all_cells$custom_label[sce_str_all_cells$Species_ID == "mcar"])

mcar_df <- mat_mcar %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

mcar_df$Var2_rev <- mcar_df$Var2
mcar_df$Var2_rev <- factor(
  mcar_df$Var2_rev, 
  levels = rev(levels(mcar_df$Var2)))

mcar_df2 <- mcar_df %>%
  dplyr::mutate(
    Var1_u = fct_inorder(paste0("orig ", Var1)),
    Var2_u = fct_inorder(paste0("custom ", Var2_rev))
  )

ggplot2::ggplot(data = mcar_df2,
                aes(axis1 = Var1_u,
                    axis2 = Var2_u, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1_u))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::scale_color_manual(values = col_cts_hsc)
```

```{r all_alluv, fig.height = 18, fig.width = 8}
mat_all <- table(
  sce_str_all_cells$celltypes,
  sce_str_all_cells$custom_label)

all_df <- mat_all %>%
  base::as.data.frame() %>%
  dplyr::filter(!is.nan(Freq))

all_df$Var2_rev <- all_df$Var2
all_df$Var2_rev <- factor(
  all_df$Var2_rev, 
  levels = rev(levels(all_df$Var2)))

all_df2 <- all_df %>%
  dplyr::mutate(
    Var1_u = fct_inorder(paste0("orig ", Var1)),
    Var2_u = fct_inorder(paste0("custom ", Var2_rev))
  )

ggplot2::ggplot(data = all_df2,
                aes(axis1 = Var1_u,
                    axis2 = Var2_u, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1_u))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(expand = c(0.15, 0.05))+
  theme_void()+
  ggplot2::theme(legend.position = "none")
knitr::knit_exit()
```

## Control gene expression

Compare the expression of marker genes (either from the literature, see 
supplementary table of the corresponding publication) or picked from the
marker gene lists of individual species and cell types.

```{r marker_genes}
# this list is a bit longer than the one in the custom assignment .Rmd scripts
# so it can cover all species
# it's also not JUST to get an idea, but to better show cell type seperation. 
  marker_genes_list <- list(
    "HSC" = c(
      "Ly6a", # HSC-specific, spotty
      "Slamf1",
      "Aldh1a1", 
      "Alcam",
      "Gng11",
      "Mllt3",
      "Socs2",

      "Cdkn1c", # HSC-specific, works for all and I known them well
      "Procr",
      "Mmrn1",
      "Mecom",
      "Mpl",

      "Pdzk1ip1", # HSC-specific, works for all but I don't know it well

      "Ltb", # mostly HSC, works for all, and I know them
      "Hlf", 
      "Ldhb",

      "Txnip", # redox functions, likely also expressed in other clusters
      "Pbx1",

      "Angpt1", # quite broad already, but highest in HSC and well-known mostly
      "Hoxa9",
      "Lmo2",
      "Meis1",
      "Msi2"
    ),
    "MPP" = c(
      "Adgrl4", 
      "Cd34", 
      "Sox4",
      "Cd48", 
      "Sell",
      "Cdk4"
    ),
    "Myeloid_general" = c(
      "Spi1", 
      "Cebpa",
      "Ctsg",
      "Mpo",
      "Prtn3",
      "F13a1"),

    "Neutro" = c(
      "Elane", 
      "Gfi1",
      "Cebpe", 
      "S100a8",
      "S100a9",
      "Wfdc21",
      "Chil1"),

    "Mono" = c(  
      "Ccr2",
      "Irf8",
      "Ly86",
      "Ms4a6c"),

    "Antigen_presenting_contamination" = c(
      "Cd74",
      "H2-Aa",
      "H2-Eb1",
      "H2-Ab1",
      "Ctss"),

    "Lymphoid_early" = c(
      "Tcf4",
      "Flt3", 
      "Satb1",
      "Dntt"),

    "Lymphoid_late" = c(
      "Ighm",
      "Il7r",
      "Rag1",
      "Vpreb1",
      "Vpreb3",
      "Cd79b", 
      "Igkc"),

    "Lymph_contamination" = c(
      "Ccl5", # NK
      "Cd3e", # TC
      "Cd8a",
      "Tcf7"),

    "Mast/Baso" = c(
      "Mcpt8", 
      "Hdc", 
      "Lmo4", 
      "Gzmb",
      "Egr1",
      "Cpa3", 
      "Ms4a2",
      "Prss34",
      "Fcer1a",
      "Gata2"#  also Mk, Eo Baso and HSC
    ),

    "Mk" = c(
      "Itga2b", 
      "Cavin2",
      "Pf4",
      "Clec4d",
      "Plek",
      "Fli1"),

    "Mk/Ery" = c(
      "Zfpm1",
      "Gata1", 
      "Trib2"),

    "Ery" = c(
      "Frat2",
      "Ass1",
      "Aqp11",
      "Klf1",
      "Blvrb",
      "Tspo2",
      "Epor", 
      "Gfi1b",
      "Gypa",
      "Hba-a1",
      "Hbb-bs"),

    "CC entry/activation" = c(
      "Lig1",
      "Mcm7",
      "Hells",
      "Mcm6"),

    "Cell cycle S phase" = c(
      "Aurkb",
      "Cdk1"
      ),

    "Cell cycle M phase" = c(
      "Cdc20",
      "Ccnb2",
      "Knstrn"),

    "Cell cycle general" = c(
      "Mki67",
      "Lockd",
      "Cenpa" 
      )
    )
```


