---
title: "Custom clustering mspr str"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])
#sce_species <- base::readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/01_mnnc/comparison/sce_mspr_str_Object_ID-01")

fraction_curr <- "str"
species_curr <- "mspr"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be
# defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mspr_str_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:30])
names(col_num) <- as.character(c(1:30))
```

# Clustering

### Define parameters

```{r parameters}
k_graph <- 15 
resolution_louvain <- 0.5
```

More finely resolved I think but it looks like it wants to 
cluster that way and the quality actually looks fine.

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "str"){
  marker_genes <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Anxa1",
    "Bglap",
    "Bglap2",
    "Cd200",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Ptgis",
    "Pth1r",
    "Runx2",
    "Sp7",
    "Spp1",
    "Tnc",
    "Acan", # Chondro
    "Adamts5",
    "Chad",
    "Fmod",
    "Mef2c",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", # fibro chondro
    "S100a10",
    "Cd34", # fibro
    "Col3a1",
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Cdh5", # pan endo
    "Pecam1",
    "Car4", # capillary
    "Rgcc",
    "Timp4",
    "Efnb2", # Arteriolar
    "Hey1",
    "Ly6a",
    "Sox17",
    "Ackr1", # Venous/sinusoidal
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Nr2f2",
    "Selp",
    "Vcam1", # sinusoidal
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", # Lymphatic EC
    "Lyve1", 
    "Pdpn",
    "Prox1",
    "Rgs4", # pericyte
    "Rgs5",
    "Steap4",
    "Acta2", # smooth muscle
    "Myh11", 
    "Tagln",
    "Myod1", # just some skeletal muscle markers for contamination
    "Myh2",
    "Des",
    "Cd3d", # just some mature immune cell markers for contamination
    "Cd8a",
    "Cd4",
    "Il7r",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Cd74",
    "S100a8",
    "S100a9",
    "Ccl5",
    "Tcf7",
    "Cpa3", # Mast
    "Hdc",
    "Gypa", # Ery
    "Hba-a1"
  )
}

```

```{r genes_short}

if(fraction_curr == "str"){
  marker_genes_short <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Bglap",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Runx2",
    "Sp7",
    "Spp1",
    "Acan", 
    "Chad",
    "Fmod",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34", 
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Pecam1",
    "Car4",
    "Rgcc",
    "Efnb2",
    "Hey1",
    "Sox17",
    "Ackr1", 
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Rgs4", 
    "Steap4",
    "Acta2", 
    "Myh2"
  )
}

```

#### Heatmap

Just to get an overview of what is found in each cluster

```{r heatmap, fig.height = 10, fig.width = 6}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 10}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

## Mes

```{r markers_mes}

# Car
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# strong adipo-CAR support

cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks mesenchymal/adipo --> Pappa, Ebf3, Teb2, Lepr
# check out some more adipo genes

# osteo
cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# very strong osteo profile in top markers

# osteo-chondro-ish?
cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks kind of balanced, fibro, too
# Lum, Ptn, Aspn, Ogn, Fmod from this mixed population
# no overt chondro osteo genes like Spp1 or Acan

# chondro-fibro?
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# actually looks more fibro
# check

# chondro?
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# check

# chondro-fibro?
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Fn1, maybe also low quality Xist
# check
```

Check these unclear clusters ALSO check quality.

```{r umaps_mes, fig.height = 3, fig.width = 4}

marker_genes_list_mes <- as.list(c(
  "Ntng1",
  "Palld",
  "Tns3", # osteogenic vs adipocytic fate https://doi.org/10.1007/s00018-023-04930-5
  "Ptprd", # related to adipogenesis in some way https://doi.org/10.2217/epi-2019-0267 https://doi.org/10.1016/j.cmet.2022.02.012
  "Cp", # related to obesity, expressed in adipose tissue https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0080274
  "Rai14", # related to migration, mechanical sensing? 
  "Grem1", # Zhong
  "Cebpa", # Zhong
  "Cebpe", # Zhong
  "Runx1", # adding some from mmus and literature. 
           # osteogenic differentation, inhibits adipo , https://doi.org/10.7554/eLife.54695
  "Pappa", # adipocyte marker in MALP, getting several genes from here: https://doi.org/10.7554/eLife.54695
  "Pparg", # adipocyte marker in MALP
  "Plin1", # # lipid droplet coating
  "Fabp4", # fatty acid bp indicating no lipid droplet
  "Lep", # white adipose tissue
  "Ucp1", # brown
  "Tnfrsf9", # beige
  "Cidea", # brown
  "Cited1", # beige
  "Tbx1", # beige
  "Hoxc9", # adipose depot
  "Vegfc", # angiogenic factor expressed by adipo CAR/MAlP
  "Vegfa", # angiogenic factor expressed by adipo CAR/MAlP
  "Rspo3", # angiogenic factor expressed by adipo CAR/MAlP
  "Apoe", # adipocyte marker in MALP
  "Ebf3", # niche cells https://doi.org/10.1101/gad.314013.118
  "Tnc", # niche https://doi.org/10.1182/blood-2011-11-393645 
  "Ltbp2", # related to adipocytic gene expession in adipocytes
  "Gdpd2", # adipocyte marker in MALP
  "Xist", # always view this as low quality
  "Mycbp2", # ?
  "Omd",
  "Ptn",
  "Gpx3",
  "Ogn",
  "Apod",
  "Clec3b",
  "Dcn",
  "Gsn",
  "Tnxb",
  "Ecm1",
  "Ccn3", 
  "Cd34", # add two known fibro genes
  "Thy1",
  "Thbs4",
  "Scx",
  "Tnmd",
  "Cilp",
  "Chodl",
  "Itgbl1",
  "Mgp",
  "Neat1",
  "Fn1"
))

plot_list_gene <- lapply(marker_genes_list_mes, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_mes
plot_list_gene
```

Cluster 2 looks more fibro-like, but also contains the transition 
from more chondro-osteo-like cells, so I think it's still mixed.
Genes like Clec3b, Pcolce2, Tnxb, Dpt are chondro and genes like
Dcn, Gsn, Thy1, Cd34 are fibro, in the umaps the transition is 
clearly visible.

Cluster 10 is starting to look VERY chondro. 
Already loosing some intermediate gene expression like Chad, Comp.
No Cd34, Thy1, Lum, Pdgfra.
Cilp, Mgp, Scx, Chodl are all chondro-related. 

```{r mes_qc}
qc_df <- scuttle::perCellQCMetrics(sce_species)
sce_species$log_sum <- log(qc_df$sum)
sce_species$log_detected <- log(qc_df$detected)

umap_cont(sce_species, "log_sum")
umap_cont(sce_species, "log_detected")

qc_df$cluster_louvain_custom <- sce_species$cluster_louvain_custom
qc_df <- as.data.frame(qc_df)

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(sum)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(detected)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()
```

Clusters 4 and 11, and a bit 1, have very low quality compared to the others.
I did remove a "low-quality" CAR-like cluster from the merged dataset, so
maybe this was mostly from Spret.
Both of these cluster express many marker genes similar to the "real" high-
quality corresponding cluster, but weakly, but lack genes that differentiate
them from them, so I view them as "low quality copies".
ECs also contain some low-quality cells but these didn't cluster separately
so I cannot get rid of them easily right now.

```{r violin_mes, fig.height = 15, fig.width = 4}

# take only the genes that are expressed anyways
genes_of_interest <- marker_genes_list_mes[c(1:31)]
# and only the relevant two clusters
sce_species_genes <- sce_species[rownames(sce_species) %in% genes_of_interest,]
sce_species_genes <- sce_species_genes[,sce_species_genes$cluster_louvain_custom %in% c("3", "4")]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = genes_of_interest)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "area")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

## ECs

Plus peri SMCs.

```{r markers_ecs}

# art 
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also looks antigen presenting? but Pecam1, Sox17 and Flt1 are EC

# cap
cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Car4, Cd36, Rgcc I have used as classic capillary

# sinusoidal 1
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# looks classic sinusoidal 

# sinusoidal 2
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# also looks classic sinusoidal 

# lymphatic 
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Ccl21a, Lyve1 and Flt4 are classic LEC markers

# peri smc 
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 15)
# Acta2, Tpm2, Tagln, Rgs4 all classic Peri SMC
```

## Clear contamination?

```{r markers_cont}

# looks like a wild mix incl. skeletal muscle
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
# Ighm, H2s, Cd79a look lymph, also Mki67 cycling
# H2-Ab1, Ctss, H2-Eb1 antigen-presenting
# check other markers from heatmap

# looks also kind of mixed?
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
# Prg4 and Clic5 are ACTUALLY the same as the mysterious MSC cluster from before
# so maybe check in the context of stromal cells
# I really don't know what this is atm.

# baso/mast/lymph mix??
cluster_markers[[16]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 30)
# looks mast, baso, eo (Prss34, Hdc, Gata2), lymph (Il7r) and also Neutro?
# (Ccl6, Cd52) ish? or are these just genes that can ALSO be expressed by neu
# check in more detail as well.
```

```{r umaps_cont, fig.height = 3, fig.width = 4}

marker_genes_list_cont <- as.list(c(
  "Ighm",
  "Cd52",
  "Cd74",
  "H2-Ab1",
  "Myh2",
  "Acta1", 
  "Tnnt3",
  "Tnnc2",
  "Tpm2",
  "Nkg7",
  "Trbc1",
  "Klf1",
  "Epor",  
  "Wfdc21",
  "Trem1",
  "Cebpe",
  "Camp",
  "Ear1",
  "Ear6",
  "Prg3",
  "Prg4",
  "Clic5",
  "Htra4",
  "F13a1", 
  "Ackr3",
  "Hbegf",
  "Calcrl",
  "Htra1",
  "Cd55",
  "Anxa8",
  "Has1",
  "Ccn2", 
  "Cpe", # add some markers of this population from mcar
  "Spp1",
  "Creb5",
  "Fbln7",
  "Ccn2",
  "Ccl3",
  "Ccl4",
  "Ccl6",
  "Gata2",
  "Il7r"
))

plot_list_gene <- lapply(marker_genes_list_cont, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes_list_cont
plot_list_gene
```

Cluster 14 really is just a mix of lots of small contaminating populations.
Including some skeletal muscle cells.

Cluster 16 are mast/eo/baso cells.

Cluster 13 does contain a small amount of contaminating neutro cells. 
Otherwise it looks like a mesenchymal population close to fibros
that similar to the mysterious cluster in mcar expresses Prg4 and Clic5.
Maybe if I work together on both I can figure out what it is now.
I will do that here and not with mcar because there are more cells here.

#### Clic5: 
- myoblast diff in skeletal muscle cells https://doi.org/10.1126/sciadv.adq6795
- fibroblast-like synoviocytes are marked by Clic5 in synovial joints https://doi.org/10.1136/ard-2021-221682
- expressed on bone-marrow derived macrophages https://doi.org/10.1038/s41467-017-00227-x

#### Prg4:
- expressed in synovial fibroblast progenitors https://doi.org/10.1136/ard-2021-221682
- related to accumulation of macrophages in synovium/synovial fibrosis https://doi.org/10.1186/s13075-021-02621-9
- reduces friction in the synovial fluid https://doi.org/10.1016/j.joca.2021.02.566
- related to wound healing and scar formation https://doi.org/10.1038/s41536-022-00228-5
- fibroadipogenic progenitors in muscle https://doi.org/10.1073/pnas.2417806122
- progenitors that repair articular cartilage wounds https://doi.org/10.1172/jci.insight.167858
- hallmark of articular cartilage chondrocytes https://doi.org/10.1038/s42003-021-01857-0
- chondro-progenitor cells in superficial zone articular cartilage https://doi.org/10.3390/ijms25115605
- lining synovial fibroblasts while Prg4 low, Thy1 high fibros are sublining https://doi.org/10.1016/j.isci.2024.109707

#### Creb5:
- TF that causes Prg4 expression in articular cartilage chondrocytes https://doi.org/10.1038/s42003-021-01857-0
- causes Prg4 expression in articular cartilage chondrocyte progenitors https://doi.org/10.3390/ijms25115605

#### Htra4:
- articular cartilage chondrocyte-specific transcriptome https://doi.org/10.1016/j.joca.2021.03.007 
- zone-specific in human cartilage superficial zone https://doi.org/10.1002/art.37760
- in Prg4-high fibroblasts related to synovium in some way https://doi.org/10.1038/s42003-025-08250-1

#### F13a1
- originating from synovial fibroblasts https://doi.org/10.1186/s41232-022-00252-4
- in synovial macrophages https://doi.org/10.1016/j.isci.2024.109707
- not in human cartilage superficial zone https://doi.org/10.1002/art.37760

#### Others
- Chadl, Chad, Acan, Comp, Fmod, Cilp2, Clec3a, Prg4, Has1, Tnxb, Pcolce2,
  Crispld1, Dpt, all in articular cartilage chondrocyte-specific transcriptome
  https://doi.org/10.1016/j.joca.2021.03.007

- Dpt, Clic2, Thbs4, Ogn, Aspn in superficial zone human https://doi.org/10.1002/art.37760

Well, so it seems it's either synovial fibro or articular cartilage, or both,
or progenitors of one or both.
They are in similar places, probably they are derived from the same cell type.
I'll call them Articular/synovial progenitor-like, because now I already did 
the work.

# Assign

In the same order as above.

```{r assign}

sce_species$custom_label <- vector(length = ncol(sce_species))

# Mes
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Adipo CAR"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Low qu adipo CAR"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Osteo"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Balanced mesenchymal"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "Fibro/Chondro (more Fibro)"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Early chondro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Low qu osteo-chondro"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Articular/synovial-like"

# EC
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Arteriolar EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Capillary EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "Transitory/Venous EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Sinusoidal EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Lymphatic EC"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "Peri/SMC"

# contamination
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Mix skeletal muscle/immune"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "16"] <-  "Eo/Baso/Mast"


sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "Adipo CAR",
    "Osteo",
    "Balanced mesenchymal",
    "Early chondro",
    "Articular/synovial-like",
    "Fibro/Chondro (more Fibro)",
    "Capillary EC",
    "Arteriolar EC",
    "Transitory/Venous EC",
    "Sinusoidal EC",
    "Lymphatic EC",
    "Peri/SMC",
    # contamination
    "Low qu adipo CAR",
    "Low qu osteo-chondro",
    "Eo/Baso/Mast",
    "Mix skeletal muscle/immune"
    ))

sort(unique(sce_species$custom_label))
levels(sce_species$custom_label)
stopifnot(!is.na(sce_species$custom_label))

#knitr::knit_exit()
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 7}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

Unfortunately, looks like removing the low-quality osteo-chondro
cells also removes chondrocytes, but I've already made the 
decision.

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```