---
title: "Custom clustering TEMPLATE"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add names.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.
I will save the selected parameters for each conditions rather than defining them 
in the config.
To make sure that the config doesn't actually get disconnected or switched between
the object and this file.
Do this individual for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
```

```{r source, message = FALSE}
source(file = snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])
#sce_species <- base::readRDS(file = "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/01_mnnc/comparison/sce_mmus_hsc_Object_ID-01")

#fraction_curr <- snakemake@wildcards[["fraction"]]
fraction_curr <- "hsc"
#species_curr <- snakemake@wildcards[["species"]]
species_curr <- "mmus"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))
```

```{r colors, meassage = FALSE, eval = FALSE, include = FALSE}
#colors_path <- snakemake@params[["colors_path"]]
#source(snakemake@params[["colors"]])
#source("../../source/plotting.R")
```

## Clustering

### Define parameters

```{r parameters}
k_graph <- 30
resolution_louvain <- 1
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(
  k_graph, 
  ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

## Evaluate

### Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps, fig.height = 3, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)
```

### Gene expression

Now check the quality of cell type separation based on expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

if(fraction_curr == "hsc"){
  marker_genes <- c(
    "Alcam", # HSC, Stem
    "Aldh1a1",
    "Cdkn1c",
    "Gata2", # also Mk
    "Gng11",
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Msi2",
    "Pbx1",
    "Procr",
    "Slamf1",
    "Socs2",
    "Txnip",
    "Adgrl4", # MPP, progenitors, activation
    "Cd34", 
    "Cdk4",
    "Flt3",
    "Sox4",
    "Spi1", 
    "Cebpa", # Myeloid
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Cd63", # Neutro
    "Cebpe", 
    "Elane",
    "Gfi1", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", # Mono
    "F13a1",
    "Fcer1g",
    "Irf8",
    "Ly86",
    "Ms4a6c",
    "Cd79b", # Lymphoid
    "Dntt",
    "Il7r",
    "Rag1",
    "Satb1",
    "Cpa3", # Mast
    "Hdc", 
    "Ms4a2",
    "Lmo4", # Baso
    "Mcpt8", 
    "Prss34",
    "Itga2b", # Mk
    "Pf4",
    "Blvrb", # Mk Ery
    "Car2",
    "Gata1",
    "Zfpm1",
    "Epor", # Ery
    "Gfi1b",
    "Hba-a1",
    "Klf1",
    "Dctpp1", # CC entry
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa", # CC
    "Lockd", 
    "Mki67"
  )
}else if(fraction_curr == "str"){
  marker_genes <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Anxa1",
    "Bglap",
    "Bglap2",
    "Cd200",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Ptgis",
    "Pth1r",
    "Runx2",
    "Sp7",
    "Spp1",
    "Tnc",
    "Acan", # Chondro
    "Adamts5",
    "Chad",
    "Fmod",
    "Mef2c",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", # fibro chondro
    "S100a10",
    "Cd34", # fibro
    "Col3a1",
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Cdh5", # pan endo
    "Pecam1",
    "Car4", # capillary
    "Rgcc",
    "Timp4",
    "Efnb2", # Arteriolar
    "Hey1",
    "Ly6a",
    "Sox17",
    "Ackr1", # Venous/sinusoidal
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Nr2f2",
    "Selp",
    "Vcam1", # sinusoidal
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", # Lymphatic EC
    "Lyve1", 
    "Pdpn",
    "Prox1",
    "Rgs4", # pericyte
    "Rgs5",
    "Steap4",
    "Acta2", # smooth muscle
    "Myh11", 
    "Tagln",
    "Myod1", # just some skeletal muscle markers for contamination
    "Myh2",
    "Des",
    "Cd3d", # just some mature immune cell markers for contamination
    "Cd8a",
    "Cd4",
    "Ilr7",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Lyz2",
    "S100a8",
    "S100a9"
  )
}

```

```{r genes_short}

if(fraction_curr == "hsc"){
  marker_genes_short <- c(
    "Alcam", 
    "Cdkn1c",
    "Gata2", 
    "Mecom",
    "Meis1",
    "Mpl",
    "Cd34", 
    "Spi1", 
    "Cebpa", 
    "Mpo",
    "Cebpe", 
    "Elane",
    "S100a9",
    "Wfdc21",
    "Ccr2", 
    "Irf8",
    "Ly86",
    "Cd79b", 
    "Dntt",
    "Il7r",
    "Satb1",
    "Cpa3", 
    "Hdc", 
    "Lmo4", 
    "Prss34",
    "Itga2b", 
    "Pf4",
    "Zfpm1",
    "Epor",
    "Hba-a1",
    "Klf1",
    "Hells",
    "Lig1",
    "Cenpa", 
    "Lockd", 
    "Mki67"
  )
}else if(fraction_curr == "str"){
  marker_genes_short <- c(
    "Adipoq", # adipo CAR
    "Cxcl12", # also arteriolar EC
    "Kitl", 
    "Lepr",
    "Alpl", # Osteo
    "Bglap",
    "Ibsp",
    "Mmp13",
    "Postn",
    "Runx2",
    "Sp7",
    "Spp1",
    "Acan", 
    "Chad",
    "Fmod",
    "Mgp",
    "Prelp",
    "Prg4",
    "Sox9",
    "Comp", 
    "S100a10",
    "Cd34", 
    "Dcn",
    "Fn1",
    "Lum",
    "Pdgfra",
    "Thy1",
    "Pecam1",
    "Car4",
    "Rgcc",
    "Efnb2",
    "Hey1",
    "Sox17",
    "Ackr1", 
    "Emcn",
    "Ephb4",
    "Icam1",
    "Lrg1",
    "Selp",
    "Vcam1", 
    "Flt4",
    "Stab2",
    "Tfpi",
    "Ccl21a", 
    "Lyve1", 
    "Rgs4", 
    "Steap4",
    "Acta2", 
    "Myh2",
    "Des",
    "Cd3d", 
    "Cd8a",
    "Cd4",
    "Ilr7",
    "Cd79a",
    "Ighm",
    "Gzmb",
    "Lyz2",
    "S100a8",
  )
}

```

#### Heatmap

```{r heatmap, fig.height = 20, fig.width = 5}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 6}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot by portion
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))

```

#### UMAPs

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r fig.height = 3, fig.width = 4}
plot_list_gene
```

```{r session_info}
utils::sessionInfo()
```
