---
title: "Custom clustering mcar hsc"
author: "Lea Wölbert"
date: '2025-11-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Find custom clusterings for each species and fraction.
Try out several clustering parameters and pick the most suitable ones. 
Also add custom labels.
This will be a rare case where I will save objects from an .Rmd file that are not
figure pdfs.

Do this individually for each of the eight conditions (4 species X 2 fractions).

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
library(scran, quietly = TRUE)
library(igraph, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(scuttle, quietly = TRUE)
library(dittoSeq, quietly = TRUE)
library(Seurat, quietly = TRUE)
library(SeuratObject, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_species <- base::readRDS(file = snakemake@input[["sce_species"]])

fraction_curr <- "hsc"
species_curr <- "mcar"

stopifnot(unique(sce_species$Fraction_ID == fraction_curr))
stopifnot(unique(sce_species$Species_ID == species_curr))

# define output path here
# since in an .Rmd file run by snakemake the only output defined
# by snakemake is the html, any other output must be defined manually
base_path_data <- snakemake@params[["OUTPUT_DAT"]]
output_path <- base::paste0(base_path_data, "/02_clst/custom/sce_mcar_hsc_custom.rds")
print(output_path)
```

```{r colors, message = FALSE}
col_num <- c(dittoSeq::dittoColors()[1:20])
names(col_num) <- as.character(c(1:20))
```

# Clustering

### Define parameters

```{r parameters}
k_graph <- 15 
resolution_louvain <- 1.3 
```

### Louvain clustering

Copied from 02_louvain_clustering.R

```{r clustering}
# get a graph of nearest neighbors
graph <- scran::buildSNNGraph(
  sce_species, 
  k = k_graph, 
  use.dimred = "PCA", # corrected
  type = "rank")

# community detection
clust <- igraph::cluster_louvain(
  graph, 
  resolution = resolution_louvain)

#-------------------------------------------------------------------------------
# add to sce object 
sce_species$cluster_louvain <- clust$membership

sce_species$cluster_louvain_custom <- base::factor(
  sce_species$cluster_louvain, 
  levels = base::sort(base::unique(sce_species$cluster_louvain)))
  
print(head(sce_species$cluster_louvain_custom))

sce_species$k_graph_custom <- base::rep(k_graph, ncol(sce_species))
sce_species$resolution_louvain_custom <- base::rep(
  resolution_louvain, 
  ncol(sce_species))

print(sce_species)
```

# Evaluate

## Clustering separation

Look at the clusters, check that there are no obvious age or batch-
dependent clusters.

```{r umaps_base, fig.height = 6, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::geom_point(size = 2, alpha = 0.5)+
  ggplot2::scale_color_manual("cluster", values = col_num)
``` 

```{r umaps_age, fig.height = 4, fig.width = 6}  
umap_base(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_grid(cols = vars(sce_species$Age_ID))+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

```{r umaps_batch, fig.height = 8, fig.width = 6}
umap_base_l(sce_species, "cluster_louvain_custom")+
  ggplot2::facet_wrap(vars(sce_species$Object_ID), ncol = 2)+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Silhouette and Purity

```{r silhouette}

# Silhouette
sil.approx <- bluster::approxSilhouette(
  reducedDim(sce_species, "PCA"), 
  clusters = sce_species$cluster_louvain_custom)

sil.data <- base::as.data.frame(sil.approx)

sil.data$closest <- factor(ifelse(sil.data$width > 0, 
                                  sce_species$cluster_louvain, 
                                  sil.data$other))
sil.data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(sil.data,
                aes(x=cluster, y=width, colour=closest))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(-1, 1)+
  ggplot2::ylab("Silhouette width")+
  ggplot2::scale_color_manual("cluster", values = col_num)


# Purity
pure_sce <- bluster::neighborPurity(
  reducedDim(sce_species, "PCA"),
  sce_species$cluster_louvain)
pure_data <- base::as.data.frame(pure_sce)
pure_data$maximum <- factor(pure_data$maximum)
pure_data$cluster <- sce_species$cluster_louvain

ggplot2::ggplot(pure_data, 
                aes(x=cluster, y=purity, colour=maximum))+
  ggbeeswarm::geom_quasirandom(size = 0.1)+
  ggplot2::theme_classic()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlab("cluster (louvain)")+
  ggplot2::ylim(0, 1)+
  ggplot2::ylab("Cluster purity")+
  ggplot2::scale_color_manual("cluster", values = col_num)
```

## Gene expression

Now check the quality of the clustering based on the expressed genes.

First, I hand-pick a small number of well-known marker genes for the 
populations that I expect in BL6, based on BL6 literature 
(Check supplementary table of the corresponding publication).

This may not be enough for the exotic species or some sub-clusters 
but let's start with this first.

```{r marker_genes}

# check corresponding supplementary table 1 for sources not mentioned here
if(fraction_curr == "hsc"){
  marker_genes <- c(
    "Aldh1a1", 
    "Procr",
    "Cdkn1c",
    "Gng11",
    "Mllt3",
    "Mmrn1",
    "Mpl",
    "Alcam", 
    "Hlf",
    "Lmo2",
    "Mecom",
    "Meis1",
    "Socs2",
    "Pbx1",
    "Slamf1",
    "Txnip",
    "Msi2",
    "Ldhb",
    "Adgrl4", 
    "Cd34", 
    "Cdk4",
    "Sox4",
    "Spi1", 
    "Cebpa", 
    "Ctsg",
    "Mpo",
    "Prtn3",
    "Elane",
    "Cebpe", 
    "Gfi1", 
    "S100a8",
    "S100a9",
    "Wfdc21",
    "Ccr2", 
    "Irf8",
    "Ly86",
    "Ms4a6c",
    "Flt3", 
    "Satb1",
    "Dntt",
    "Cd79b", 
    "Il7r",
    "Rag1", # adding some additional genes, Rag1 is lymphoid https://doi.org/10.1093/nar/27.14.2938
    "Ighm", # Ig gene
    "Vpreb1", # BC development
    "Vpreb3", # BC development
    "Igkc", # Ig gene
    "Cd3e", # TC
    "Cd8a", # TC
    "Tcf7", # lymphoid/TC https://doi.org/10.3389/fimmu.2020.00470
    "Cpa3", 
    "Hdc", 
    "Ms4a2",
    "Lmo4", 
    "Mcpt8", 
    "Prss34",
    "Gata2",
    "Itga2b", 
    "Pf4",
    "Cavin2", # Mk http://dx.doi.org/10.1097/BS9.0000000000000187
    "Plek", # Mk https://doi.org/10.1182/blood-2006-08-038901
    "Rap1b", # Mk https://doi.org/10.1182/blood-2018-03-838714
    "Cd9", # Mk https://doi.org/10.1182/blood.v97.7.1982
    "Fli1", # Mk https://doi.org/10.1182/blood-2017-02-770958
    "Blvrb",
    "Gata1",
    "Zfpm1",
    "Epor", 
    "Gfi1b",
    "Hba-a1", # hemoglobin gene
    "Klf1",
    "Dctpp1", 
    "Hells",
    "Lig1",
    "Mcm6", 
    "Cenpa",
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}
```

```{r genes_short}

# shorter version of the above list
if(fraction_curr == "hsc"){
  marker_genes_short <- c(
    "Alcam", 
    "Cdkn1c",
    "Mecom",
    "Mpl",
    "Meis1",
    "Cd34", 
    "Spi1", 
    "Mpo",
    "Elane",
    "Cebpa", 
    "Cebpe", 
    "S100a9",
    "Wfdc21",
    "Ccr2", 
    "Irf8",
    "Ly86",
    "Cd79b", 
    "Dntt",
    "Il7r",
    "Satb1",
    "Cpa3", 
    "Hdc", 
    "Lmo4", 
    "Prss34",
    "Itga2b", 
    "Pf4",
    "Zfpm1",
    "Epor",
    "Hba-a1",
    "Klf1",
    "Hells",
    "Lig1",
    "Cenpa", 
    "Lockd", 
    "Mki67"
  )
}else{
  stop()
}
```

#### Heatmap

Just to get an overview of what is found in each cluster

```{r heatmap, fig.height = 10, fig.width = 5}

# get average logcounts per cluster
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("cluster_louvain_custom")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$cluster_louvain_custom),
    names_to = "cluster_louvain_custom",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$cluster_louvain_custom <- factor(
  agg_logcounts_df$cluster_louvain_custom,
  levels = as.character(c(1:length(unique(agg_logcounts_df$cluster_louvain_custom)))))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = cluster_louvain_custom,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)
```

#### Violin

```{r violin, fig.height = 20, fig.width = 6}

sce_species_genes <- sce_species[rownames(sce_species) %in% marker_genes_short,]

# make a df for visualisation
logcounts_df <- logcounts(sce_species_genes) %>%
  as.matrix() %>%
  as.data.frame() %>% 
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = colnames(sce_species_genes),
    names_to = "cell",
    values_to = "logcounts")

# add cluster info
logcounts_df$cluster_louvain_custom <- sce_species_genes$cluster_louvain_custom[
  match(
    logcounts_df$cell,
    rownames(colData(sce_species_genes)))]

logcounts_df$genes <- factor(logcounts_df$genes, levels = marker_genes_short)

# plot 
logcounts_df %>%
  ggplot2::ggplot(
    aes(
      x = cluster_louvain_custom,
      y = logcounts,
      fill = cluster_louvain_custom))+
    ggplot2::geom_violin(
      scale = "width")+
    ggplot2::theme_classic()+
    ggplot2::theme(
      strip.text.y = element_text(angle = 0))+
    ggplot2::facet_grid(rows = vars(genes))+
    ggplot2::scale_fill_manual("cluster", values = col_num)

```

#### UMAPs

Use the previously defined marker gene list to get a better
look at the expression patterns.

```{r gene_plots}
marker_genes_list <- as.list(marker_genes)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- marker_genes
```

```{r umaps, fig.height = 3, fig.width = 4}
plot_list_gene
```

## Markers

Similarly to SPRETUS, I decided to go for a larger number of clusters.
In this case, a lot of clustering is happening in the erythroid populations,
but I also wanted to separate some other populations better.
So this clustering will result in a larger number of erythroid clusters which
can later be merged back together.
One cannot simply assume that BL6 marker genes will be individually useful or
expressed in the same pattern in CAROLI, so the consensus from many genes
must be used to assign cell type labels.

```{r get_markers, message = FALSE}
# convert to seurat
seurat <- Seurat::as.Seurat(
  sce_species,
  counts = "counts", 
  data = "logcounts") 
SeuratObject::Idents(seurat) <- sce_species$cluster_louvain_custom

# get hvgs for marker genes
hvgs <- scran::modelGeneVar(sce_species)
hvgs <- scran::getTopHVGs(hvgs, n=2000)

# find marker genes for each cluster
clustlist <- as.list(1:length(base::unique(sce_species$cluster_louvain_custom)))

cluster_markers <- lapply(clustlist, function(x){
  
  markers <- Seurat::FindMarkers(seurat,
                                 slot = "data",
                                 test.use = "wilcox", 
                                 ident.1 = x, 
                                 features = hvgs)
  
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  
  markers$which_cluster <- base::rep(x, nrow(markers))
  
  return(markers)
})
```

### HSC

These HSC contain a small population of very early lymphoid cells because
there are not enough immature lymphoid cells to form their own
cluster in CAROLI.
Also, compared to other species, many BL6 HSC genes seem more broadly
expressed (slightly useless).

```{r markers_hsc}
cluster_markers[[2]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

It actually already looks quite MPP-ish, expressing Cd34, but it is
also visible that Cd34 is not really excluded from an immature population,
so for CAROLI it's maybe simply not a good HSC-MPP differentiation marker.
Despite some more mature-looking expression, Msi2, Meis1, Hoxa9, Mecom
place it in the (most) immature stem compartment (in this species' dataset).
Or we simply don't know the right stemness genes for CAROLI.

### MPP

One cluster still expressing some HSC-ish genes.
Another cluster expressing HSC/MPP genes (slightly lower) and
slightly more Mk genes. 

```{r markers_mpp}
# MPP
cluster_markers[[1]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# MPP with slight Mk expression
cluster_markers[[7]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

Cluster 11 and 7 express similar markers to cluster 2, including S100a6, Adgrl4, Cd34.
In constrast, cluster 11 also expresses BL6 MPP and progenitor markers
(Sox4, Flt3, Sell) and much fewer HSC markers. 
Cluster 7 also expresses these genes and some cycling (Mki67).
Both start expressing Lig1 and Mcm6.
I think cluster 7 is activated MPPs in different cycling phases.

```{r hsc_mpp_genes_add, fig.width = 4, fig.height = 3}
# testing some interesting genes from the marker gene lust
hsc_mpp_genes_add <- c(
  "Hoxa9",
  "Cd302",
  "S100a6",
  "Cd9",
  "Basp1",
  "Ypel3",
  "Sell",
  "Clec4d",
  "Cd93",
  "Ccnb2",
  "Prc1"
)

marker_genes_list <- as.list(hsc_mpp_genes_add)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- hsc_mpp_genes_add
plot_list_gene
```

### Myeloid

```{r markers_myeloid}
# activated early myeloid
cluster_markers[[5]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# activated and cycling early myeloid
cluster_markers[[15]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# neutro-primed progs
cluster_markers[[10]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# neutro?
cluster_markers[[17]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# mono
cluster_markers[[18]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

```{r markers_myeloid_add, fig.width = 4, fig.height = 3}
# testing some interesting genes from the marker gene lists
markers_myeloid_add <- c(
  "Mcpt8",
  "F13a1",
  "Fabp5"
)

marker_genes_list <- as.list(markers_myeloid_add)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- markers_myeloid_add
plot_list_gene
```

Cluster 5: clear expression of early myeloid/granu markers but 
not late neutro or mono markers quite as much or rather 
neutro markers like Cebpe are just "starting".
Cluster 15: same, but also cycling.
Both also contain come mono genes like Ccr2 and F13a1 also not so
sure about the latter.
Cluster 10 is clearly neutrophilic.
Cluster 18 expresses classic antigen-presenting markers like Cd74 and H2-Aa.
Cluster 17 looks slightly neutro-like but also low quality (expressing mt genes, 
Tsix and so on as marker genes). 
Check quality.

```{r myeloid_qc, fig.width = 4, fig.height = 3}
qc_df <- scuttle::perCellQCMetrics(sce_species)
sce_species$log_sum <- log(qc_df$sum)
sce_species$log_detected <- log(qc_df$detected)

umap_cont(sce_species, "log_sum")
umap_cont(sce_species, "log_detected")

qc_df$cluster_louvain_custom <- sce_species$cluster_louvain_custom
qc_df <- as.data.frame(qc_df)

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(sum)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()

ggplot2::ggplot(
  qc_df,
  aes(x = cluster_louvain_custom,
      y = log(detected)))+
  ggplot2::geom_boxplot()+
  ggplot2::theme_classic()
```

Cluster 17 has indeed the lowest quality.

### Lymphoid

There is one cluster expressing mature lymphoid markers. 

```{r markers_lymph} 
# late lymph
cluster_markers[[16]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

Mature cells, small cluster, likely contamination.

### Baso Mast

There are actually two clusters, one probably more baso than mast 
or more early than the other.
Both express some HSC quiescence/anti-redox markers

```{r markers_bm} 
# later? more mast?
cluster_markers[[8]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# earlier? more Mk? actually this may be more Mk
cluster_markers[[14]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# earlier? weaker?
cluster_markers[[12]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)
```

```{r markers_meb_add, fig.width = 4, fig.height = 3}
# testing some interesting genes from the marker gene lists
markers_meb_add <- c(
  "Ifi204",
  "Clec2d",
  "Egr1",
  "Gzmb",
  "Fcer1a",
  "Nrgn",
  "Slpi",
  "Dscam",
  "Cdkn2d",
  "Ifitm7",
  "Ptprcap",
  "Lgals9",
  "Nudt14"
)

marker_genes_list <- as.list(markers_meb_add)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- markers_meb_add
plot_list_gene
```

Cluster 8 expresses mast/baso markers strongly, also the only
cluster expressing Gzmb, Ms4a2 in a subpopulation, making it look more
mature.

Cluster 14 looks more Mk like, with higher Cavin2, Itga2b, Meis1 expression.
Also including Fcer1a, Nrgn, Slpi, Dscam, Cdkn2d which I don't know but seem
fairly specific.
And some small remnants of Mecom, Hlf, Mpl, Mllt3, Gng11.
So I think it's quite an immature Mk/Mast/Baso-primed population that
is getting activated (Lig1, Hells).

Cluster 12 looks early but less activated. 
It's all over the place.
Parts of it express more early ery-like genes. 
Also some sort of MEB progenitor, a bit more mixed?

### Mk Ery

There is quite a bit of separation in the very Ery looking cluster.
Any cluster expressing Mk markers somewhat strongly (1, 2, 7, 8, 14,
even 15) has already been mentioned above, maybe they are interspersed.
So only ery clusters are left basically.
Not sure if we can find meaningful differences.

```{r markers_ery}
# ery 3
cluster_markers[[3]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# ery 4, maybe slightly more early
cluster_markers[[4]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# ery 6, stronger ery expression and more active cycling 
cluster_markers[[6]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# ery 9 looks similar to 6 from the heatmap
cluster_markers[[9]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# ery 11, looks like 6 and 9
cluster_markers[[11]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

# ery 13 seems even more strongly expressing ery markers
cluster_markers[[13]] %>%
  dplyr::filter(avg_log2FC > 0) %>%
  head(n = 40)

```

```{r ery_markers_add, fig.width = 4, fig.height = 3}
# testing some potentially interesting genes from the marker gene lists
ery_markers_add <- c(
  "Bex4",
  "Ermap",
  "Rgcc",
  "Ass1",
  "Trib2",
  "Frat2",
  "Aqp11",
  "Gypa"
)

marker_genes_list <- as.list(ery_markers_add)

plot_list_gene <- lapply(marker_genes_list, function(gene){
  plot <- umap_gene(sce_species, gene)
})
names(plot_list_gene) <- ery_markers_add
plot_list_gene
```

Clusters 3, 11, 9 all kind of look like normal ery progenitors.
9 seems much more cycle-y (Cdk1, Mki67). 
Cluster 4 looks still pretty early, expressing genes like Trib2 and Cpa3, 
but already clearly ery-biased.
Clusters 6 and 13 look more mature with high Hba-a1 expression.
Cluster 13 seems even more mature.

# Assign

```{r assign}

sce_species$custom_label <- vector(length = ncol(sce_species))

# HSC
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "2"] <-  "HSC"

# MPPs
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "1"] <-  "Activated MPP"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "7"] <-  "Cycling MPP"  

# myeloid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "5"] <-  "Granu/mono prog"  
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "15"] <-  "Cycling granu/mono prog" 
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "18"] <-  "Antigen-presenting"   
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "10"] <-  "Neutro prog"  
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "17"] <-  "Low quality"  

# lymphoid
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "16"] <-  "Late lymphoid prog"
  
# BM Ery Mk
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "14"] <-  "Mk/Ery/BM (more Mk/BM) prog"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "12"] <-  "Mk/Ery/BM (more ery) prog"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "8"] <-  "BM prog"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "4"] <-  "Early ery prog"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "3"] <-  "Ery prog 1"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "11"] <-  "Ery prog 2"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "9"] <-  "Ery prog 3"

sce_species$custom_label[
  sce_species$cluster_louvain_custom == "6"] <-  "Late ery prog"
sce_species$custom_label[
  sce_species$cluster_louvain_custom == "13"] <-  "Latest ery prog"

sce_species$custom_label <- factor(
  sce_species$custom_label,
  levels = c(
    "HSC",
    "Activated MPP",
    "Cycling MPP",
    "Granu/mono prog",
    "Cycling granu/mono prog",
    "Neutro prog",
    "Antigen-presenting",
    "Late lymphoid prog",
    "BM prog",
    "Mk/Ery/BM (more Mk/BM) prog",
    "Mk/Ery/BM (more ery) prog",
    "Early ery prog",
    "Ery prog 1",
    "Ery prog 2",
    "Ery prog 3",
    "Late ery prog",
    "Latest ery prog",
    # contamination
    "Low quality"
    ))
stopifnot(!is.na(sce_species$custom_label))
```

```{r unif}

sce_species$custom_label_unif <- unfactor(sce_species$custom_label)

sce_species$custom_label_unif[
  sce_species$custom_label == "Ery prog 1"] <-  "Ery prog"
sce_species$custom_label_unif[
  sce_species$custom_label == "Ery prog 2"] <-  "Ery prog"
sce_species$custom_label_unif[
  sce_species$custom_label == "Ery prog 3"] <-  "Ery prog"

sce_species$custom_label_unif <- factor(
  sce_species$custom_label_unif,
  levels = c(
    "HSC",
    "Activated MPP",
    "Cycling MPP",
    "Granu/mono prog",
    "Cycling granu/mono prog",
    "Neutro prog",
    "Antigen-presenting",
    "Late lymphoid prog",
    "BM prog",
    "Mk/Ery/BM (more Mk/BM)",
    "Mk/Ery/BM (more ery)",
    "Early ery prog",
    "Ery prog",
    "Late ery prog",
    "Latest ery prog",
    # contamination
    "Low quality"
    ))
stopifnot(!is.na(sce_species$custom_label))
```

One last plot to confirm in correct order.

```{r heatmap_confirm, fig.height = 10, fig.width = 5}

# get average logcounts per label
agg_sce_species <- scuttle::aggregateAcrossCells(
  sce_species,
  id=colData(sce_species)[,c("custom_label")], 
  statistics = "mean",
  use.assay.type = "logcounts")

# turn into df formatted for visualisation
agg_logcounts_df <- logcounts(agg_sce_species) %>%
  as.matrix() %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "genes") %>%
  tidyr::pivot_longer(
    cols = unique(agg_sce_species$custom_label),
    names_to = "custom_label",
    values_to = "avg_logcounts")

# only keep genes for visualisation
agg_logcounts_df <- agg_logcounts_df %>%
  dplyr::filter(genes %in% marker_genes)

# factorise for nicer plot
agg_logcounts_df$genes <- factor(
  agg_logcounts_df$genes,
  levels = rev(marker_genes))

agg_logcounts_df$custom_label <- factor(
  agg_logcounts_df$custom_label,
  levels = levels(sce_species$custom_label))

# scale expression per gene using z-scores
agg_logcounts_df$zscore_gene <- vector(
  length = nrow(agg_logcounts_df), 
  mode = "numeric")

for(gene in base::unique(agg_logcounts_df$genes)){
  
  expr_vector_g <- agg_logcounts_df[
    agg_logcounts_df$genes == gene,]$avg_logcounts

  # calculate z score per gene
  z_score_g <- (expr_vector_g - base::mean(expr_vector_g))/
    stats::sd(expr_vector_g)
  agg_logcounts_df$zscore_gene[agg_logcounts_df$genes == gene] <- z_score_g
}

# plot
ggplot2::ggplot(
  agg_logcounts_df,
  aes(
    x = custom_label,
    y = genes,
    fill = zscore_gene))+
  ggplot2::geom_tile()+
  ggplot2::theme_classic()+
  ggplot2::scale_fill_gradient2(
    low = "blue3",
    mid = "white", 
    high = "red3", 
    midpoint = 0)+
  ggplot2::theme(axis.text.x = element_text(angle = 90))
```

## Save 

```{r save}
print(output_path)
saveRDS(sce_species, output_path)
```

```{r session_info}
utils::sessionInfo()
```
