---
title: "Shared genes report"
author: "Lea Wölbert"
date: '2023-07-14'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on nDGEs per cell type, and possible nDGE candidates for subclustering.

#### Load objects

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(EnhancedVolcano, quietly = TRUE)
library(DESeq2, quietly = TRUE)
```

```{r source, message = FALSE}
source(file = snakemake@params[["plotting"]])
```

```{r load_objects, message = FALSE}
fc_cutoff <- snakemake@params[["fc_cutoff"]]
padj_cutoff <- snakemake@params[["padj_cutoff"]]

sce <- base::readRDS(snakemake@input[["sce_input"]])
sep <- base::readRDS(snakemake@input[["sep"]])

cluster_res_list <- base::readRDS(snakemake@input[["cluster_res"]])
shared_gene_list <- base::readRDS(snakemake@input[["cluster_sharedgenes_list"]])

cand_list <- utils::read.csv(snakemake@input[["gene_list"]], 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

fraction_curr <- snakemake@wildcards[["fraction"]]
cluster_curr <- snakemake@wildcards[["cluster"]]
```

```{r print_objects, message = FALSE}
print(sce)
print(levels(sce$annotation_cluster))
print(unique(sce$cluster_louvain))
print(cluster_curr)
print(fraction_curr)

if(!cluster_curr %in% sce$cluster_louvain){
  print("cluster irrelevant")
  knitr::knit_exit()
}
```

```{r subset}
# subset objects to required cluster and fractions
cand_list <- cand_list[cand_list$Fraction == fraction_curr,]
cand_genes <- cand_list$Gene
sce_clust <- sce[,sce$cluster_louvain == as.numeric(cluster_curr)]

ct_curr <- unfactor(sce_clust$annotation_cluster[1])
print(ct_curr)

shared_gene_list <- shared_gene_list[[ct_curr]]
cluster_res_list <- cluster_res_list[[ct_curr]]
```

# Show Pval distributions

```{r pval_distr_function}
see_pvals <- function(cluster_res, comparison){
  
  plot1 <- EnhancedVolcano::EnhancedVolcano(
    cluster_res,
    lab = "",
    x = 'log2FoldChange', 
    y = 'pvalue', 
    col = c("grey70", "grey70", "dodgerblue3", "grey70"),
    subtitle = "", 
    colAlpha = 0.9, 
    pointSize = 0.5,
    pCutoff = padj_cutoff, 
    FCcutoff = fc_cutoff)+
    ggplot2::theme_classic()+
    ggplot2::ggtitle(comparison)+
    ggplot2::theme(legend.position="none",
                   plot.subtitle = element_blank())
  
  DESeq2::plotMA(cluster_res, main = comparison)

  print(plot1)
}
```

## MMUS MCAS

```{r mmus_mcas}
print(cluster_res_list[["mmus-mcas"]])
res_df_comb_mmus_mcas <- cluster_res_list[["mmus-mcas"]]
see_pvals(cluster_res = res_df_comb_mmus_mcas, comparison = "mmus-mcas")
```

## MMUS MSPR

```{r mmus_mspr}
res_df_comb_mmus_mspr <- cluster_res_list[["mmus-mspr"]]
see_pvals(cluster_res = res_df_comb_mmus_mspr, comparison = "mmus-mspr")
```

## MMUS MCAR

```{r mmus_mcar}
res_df_comb_mmus_mcar <- cluster_res_list[["mmus-mcar"]]
see_pvals(cluster_res = res_df_comb_mmus_mcar, comparison = "mmus-mcar")
```

## MCAS MSPR

```{r mcas_mspr}
res_df_comb_mcas_mspr <- cluster_res_list[["mcas-mspr"]]
see_pvals(cluster_res = res_df_comb_mcas_mspr, comparison = "mcas-mspr")
```

## MCAS MCAR

```{r mcas_mcar}
res_df_comb_mcas_mcar <- cluster_res_list[["mcas-mcar"]]
see_pvals(cluster_res = res_df_comb_mcas_mcar, comparison = "mcas-mcar")
```

## MSPR MCAR

```{r mspr_mcar}
res_df_comb_mspr_mcar <- cluster_res_list[["mspr-mcar"]]
see_pvals(cluster_res = res_df_comb_mspr_mcar, comparison = "mspr-mcar")
```

# Subclustering genes

Candidate genes are potential subclustering genes that were picked from
literature.
These genes are here tested for their potential to separate certain cell types 
into subpopulations.
Only candidate genes that are nDGE in at least three of four species
will be considered for subclustering.

```{r print}
print(base::paste("comparison", ct_curr))

remain_cand_list <- cand_list$Gene[
  which(cand_list$Gene %in% shared_gene_list$three)]
print(remain_cand_list)
```

## UMAPS of remaining candidate subclustering genes

```{r umap_plots_function}
gene_exps <- function(gene){
  
  sce_clust$Species_ID <- factor(sce_clust$Species_ID,
                                 levels = c("mmus", "mcas", "mspr", "mcar"))

  sce$Species_ID <- factor(sce$Species_ID,
                           levels = c("mmus", "mcas", "mspr", "mcar"))
  
  # Gene expression UMAP in cluster X, own function
  plot_1 <- umap_gene(sce_clust, gene)

  # Gene expression UMAP in cluster X, separated by species
  plot_2 <- umap_gene(sce_clust, gene)+
    ggplot2::facet_grid(cols = vars(sce_clust$Species_ID))
  
  ggdf <- base::as.data.frame(colData(sce_clust))
  # normalised logcounts (multibatchnorm)
  logc <- SummarizedExperiment::assays(sce_clust)[["logcounts"]]

  ggdf$Species_ID <- factor(ggdf$Species_ID,
                            levels = c("mmus", "mcas", "mspr", "mcar"))
  
  plot_3 <- ggplot2::ggplot(ggdf,
                            aes(x = cluster_louvain,
                                y = logc[rownames(logc) == gene,]))+
    ggbeeswarm::geom_quasirandom(size = 0.2, color = "grey60")+
    ggplot2::theme_classic()+
    ggplot2::geom_boxplot(color = "black", alpha = 0)+
    ggplot2::ylab(base::paste0(gene, " expression"))+
    ggplot2::facet_grid(cols = vars(sce_clust$Species_ID))
  
  return(list(gene, plot_1, plot_2, plot_3))
}
```

```{r umap_plots}
plotlist <- lapply(as.list(remain_cand_list), gene_exps)
plotlist
```

```{r}
print(ct_curr)
print(shared_gene_list$all)
print(shared_gene_list$three)
```

```{r session_info}
utils::sessionInfo()
```