---
title: "Blind annotation of TABULA SAPIENS BONE MARROW"
author: "Lea Wölbert"
date: '2025-12-09'
output: html_document
---

Blind annotation of two human HSPC datasets and one mouse and one human
niche cell datasets after unbiased re-clustering with signature genes.
This is the script for annotating TABULA SAPIENS BONE MARROW.
The thusly obtained clusters will now be annotated based only on the expression
of signature gene orthologs within these clusters, and without knowledge of
which cells from which originally annotated cell type these clusters contain.

This is another occasion where I will save an object from an .Rmd file because
after naming it and looking at it this way, I'm already adding the annotations.

These parts of code were originally in the first iteration of Figure 4 but after
adding the custom annotations, I have decided to separate the blind annotations
from the figure, too.
I didn't change too much from the original figure script though.
Instead, I added some comments why I chose each annotation.

Check supplementary table 1 of the corresponding publication for some of the 
marker gene sources.

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load_packages, message = FALSE}
library(Seurat, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@input[["base_path"]]
metadata_path <- snakemake@input[["metadata_path"]]
```

```{r load_objects}

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  metadata_path,
  "/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")

ensembl_sign <- readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/02_endf/ensembl_sign_hsc"))
```

Other 1 is a remnant from old figure 4 iterations.

```{r load_other1}

dataset_o1 <- "ts_bone_marrow"

# list of seurat objects with reclustering vectors for each gene set+ resolution
seu_hsc_recl_list_o1 <- base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_ts_bone_marrow_list"))
```

```{r selected_seu_o1}

resolution_df_o1 <- resolution_df[resolution_df$dataset == dataset_o1,]

selected_seu_list_o1 <- lapply(seu_hsc_recl_list_o1, function(seu_list){

  conslev_curr <- seu_list[[1]]@misc$used_genes

  resdf_temp <- resolution_df_o1[
    resolution_df_o1$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution

  seu_return <- seu_list[[res_curr]]
  return(seu_return)

})

rm(seu_hsc_recl_list_o1)
gc()
```

```{r load_custom_labels}
# load the object with the custom labels to put them into one object
seu_ts_bone_marrow_cust <-  base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/03_custom_clustering/ts_bone_marrow.Rds"))
```

```{r export_paths}
output_path <- base::paste0(
  base_path, 
  "/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/ts_bone_marrow_both.Rds")
print(output_path)
```

## Blind annotation based on signature genes only

### Prep = Calculate Markers

```{r prepare_annotate_o1}

seu_sign_o1 <- selected_seu_list_o1$seu_sign
reclustered_o1 <- base::unique(seu_sign_o1$seurat_clusters)

clust_list_o1 <- as.list(reclustered_o1)

ct_markers_o1 <- lapply(clust_list_o1, function(x){
  markers <- Seurat::FindMarkers(seu_sign_o1,
                                 test.use = "wilcox",
                                 ident.1 = x)
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  # need to add only ensembl gene symbols as I will not be using others
  markers$symbol <- markers$genes
  markers$symbol <- ensembl_sign$HSAPIENS_SYMBOL[
    base::match(markers$genes, ensembl_sign$ENSG_ID)]
  return(markers)
})
names(ct_markers_o1) <- reclustered_o1

ct_markers_df_o1 <- bind_rows(ct_markers_o1)
head(ct_markers_df_o1)
```

```{r features}
print(unique(rownames(seu_sign_o1)))
print(dim(seu_sign_o1))
# the object only has 225 features, so only the signature genes
# can be used to get marker genes
```

I will not use any other genes than orthologs of signature genes44
for this annotation.
It is blind because I don't know the original identity of each cell/
cluster as I am annotating them.

### UMAPs

```{r umap_others_1}
Seurat::DimPlot(seu_sign_o1, dims = c(1,2), group.by = "seurat_clusters")
Seurat::DimPlot(seu_sign_o1, dims = c(1,2), group.by = "assay")
```

```{r umaps_dev, fig.height = 6, fig.width = 6}
Seurat::DimPlot(seu_sign_o1, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(seu_sign_o1@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
Seurat::DimPlot(seu_sign_o1, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(seu_sign_o1@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay, fig.height = 6, fig.width = 6}
Seurat::DimPlot(seu_sign_o1, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(seu_sign_o1@meta.data$assay), ncol = 2)
``` 

### Print markers

Print marker genes for each cluster for label annotation.

```{r add_annotated_reclustered}
seu_sign_o1$annotated_reclustered <- vector(length = ncol(seu_sign_o1))
# cluster 1 is smart-seq batch
```

```{r others1_cluster6}
nr <- "6"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# ITGA2B, TGFB1, ANGPT1, RAMP1, MPL, ZFPM1
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o1, markers[1:6])
# Itga2b, Tgfb1,Zfpm1, Gata2, Mpl as markers indicate Mk cells 

seu_sign_o1$annotated_reclustered[seu_sign_o1$seurat_clusters == nr] <- "Mk prog"
```

```{r others1_cluster4}
nr <- "5"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# TRIB2, ITGA2B, KLF1, GATA1, LMO2, ZFPM1, TXNIP, EPOR, GATA2, 
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# Itga2b, Klf1, Gata1, Zfpm1 indicate early Mk and Ery mixed cells

seu_sign_o1$annotated_reclustered[seu_sign_o1$seurat_clusters == nr] <- "Mk/Ery prog" 
```

```{r others1_cluster2}
nr <- "1"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# AQP1, CDC20, KLF1, BLVRB, GATA1, EPOR, CCNB2, TUBB
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# Gata1, Klf1, Blvrb, Epor are slightly more early-ish than Hb genes

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "Ery prog" 
```

```{r others1_cluster8}
nr <- "8"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# TSPO2, AQP1, HMBS, EPOR, MKI67, CENPE, CENPF, BLVRB, KLF1
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]
# Hmbs, Ermap, cycling genes, Alad make it look later

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "Late Ery prog"
```

```{r others1_cluster1}
nr <- "0"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# MS4A2, MPL, ARHGDIB, 
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "batch, Smart-Seq2" 
```

```{r others1_cluster3}
nr <- "4"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# HLF, C17orf50, CD34, MECOM, MSI2, ANGPT1, RAMP1, MEIS1, MPL
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# HLF, CD34, MECOM, MSI2 are all classic HSC genes 
# C17Orf is 1700020L24Rik btw, a neutro prog signature gene

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "HSC" 
```

```{r others1_cluster7}
nr <- "7"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# SMIM14, LSP1, LY86, CLEC12A, SPI1, MEF2C, CYBA, ANXA2, TCF4, 
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# this was not so easy, but LY86, SPI1, CLEC12A indicate something
# early myeloid, and TCF4 and LY86 a little bit lymphoid, too.
# mono and lymphoid do overlap some genes, so I went with a more mono anno

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "Mono prog"
```

```{r others1_cluster0}
nr <- "3"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# PRTN3, CTSG, MPO, MS4A3, CEBPE, ANXA2, CLEC12A, PLAC8, SPI1, LY86
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# PRTN3, CTSG and MPO, SPI1 typically appear a little earlier in mixed
# myeloid cells (not neutro or mono-restricted yet)

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "Myeloid prog" 
```

```{r others1_cluster5}
nr <- "2"
markers <- ct_markers_o1[[nr]]$genes[
  !is.na(ct_markers_o1[[nr]]$symbol) &
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05]

# CENPA, ANXA2, MKI67, AURKB, CENPE, LY86, CDK1, CLEC12A, SPI1, 
ct_markers_o1[[nr]][
  !is.na(ct_markers_o1[[nr]]$symbol) & 
    ct_markers_o1[[nr]]$avg_log2FC > 0 &
    ct_markers_o1[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

Seurat::FeaturePlot(seu_sign_o1, markers[1:6])
# these are cycling clearly, but also express myeloid markers like Spi1

seu_sign_o1$annotated_reclustered[
  seu_sign_o1$seurat_clusters == nr] <- "Cycling Myeloid prog"
```

Factorize for nicer plots.

```{r others1_levels}
seu_sign_o1$annotated_reclustered <- factor(
  seu_sign_o1$annotated_reclustered,
  levels = c(
    "HSC",
    "Mk prog",
    "Mk/Ery prog",
    "Ery prog",
    "Late Ery prog",
    "Myeloid prog",
    "Mono prog",
    "Cycling Myeloid prog",
    "batch, Smart-Seq2"))

print(sort(levels(seu_sign_o1$annotated_reclustered)))

stopifnot(!is.na(seu_sign_o1$annotated_reclustered))
```

## Add clusters from the custom clustering

```{r add_clusters}
# add the custom-labelled clusters

seu_sign_o1$signature_label <- seu_sign_o1$annotated_reclustered
seu_sign_o1$custom_label <- vector(length = ncol(seu_sign_o1))

# match to be sure but they should be identical
seu_sign_o1$custom_label <- seu_ts_bone_marrow_cust$custom_label[
  match(
    colnames(seu_sign_o1),
    colnames(seu_ts_bone_marrow_cust))]
```

```{r tables_all}
print(table(seu_sign_o1$custom_label, seu_sign_o1$signature_label))
print(table(seu_sign_o1$cell_type, seu_sign_o1$signature_label))
print(table(seu_sign_o1$custom_label, seu_sign_o1$cell_type))
```

```{r umaps_labels, fig.height = 6, fig.width = 6}
Seurat::DimPlot(seu_sign_o1, reduction = "umap", group.by = "signature_label")
```

```{r umaps_labels2, fig.height = 6, fig.width = 8}
Seurat::DimPlot(seu_sign_o1, reduction = "umap", group.by = "custom_label")
```

## Save

```{r save}
saveRDS(seu_sign_o1, output_path)
```

```{r sessioninfo}
utils::sessionInfo()
```