---
title: "Blind annotation of MUS TIKHONOVA ET AL STROMAL"
date: '2025-12-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Blind annotation of two human HSPC datasets and one mouse and one human
niche cell datasets after unbiased re-clustering with signature genes.
This is the script for annotating MUS TIKHONOVA ET AL STROMAL.
The thusly obtained clusters will now be annotated based only on the expression
of signature gene orthologs within these clusters, and without knowledge of
which cells from which originally annotated cell type these clusters contain.

This is another occasion where I will save an object from an .Rmd file because
after naming it and looking at it this way, I'm already adding the annotations.

These parts of code were originally in the first iteration of Figure 4 but after
adding the custom annotations, I have decided to separate the blind annotations
from the figure, too.
I didn't change too much from the original figure script though.
Instead, I added some comments why I chose each annotation.

Check supplementary table 1 of the corrsponding publication for some of the 
marker gene sources.

#### Load objects

```{r seed, message = FALSE}
base::RNGkind("L'Ecuyer-CMRG")
set.seed(37)
```

```{r load_packages, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(Seurat, quietly = TRUE)
```

```{r define_basepath}
base_path <- snakemake@input[["base_path"]]
metadata_path <- snakemake@input[["metadata_path"]]
```

```{r load_short}

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  metadata_path,
  "/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(
  file = resolution_df_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

ensembl_sign <- readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/02_endf/ensembl_sign_str"))

# list of gene sets for HSPCs (conserved signature, conserved markers, etc.)
# required for getting more info on the signature genes
geneset_list_str <-base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_str"))
```

Other 2 as name for this dataset is a remnant from previous figure 5 iteration.

```{r load_other2}

dataset_o2 <- "mus_tik_stromal"

# list of seurat objects with reclustering vectors for each gene set and resolution
seu_str_recl_list_o2 <- base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_mus_tik_stromal_list"))
```

```{r export_paths}
output_path <- base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/mus_tik_stromal.Rds")
print(output_path)
```

## Blind annotation

### Prep

```{r selected_seu_o2}

resolution_df_o2 <- resolution_df[resolution_df$dataset == dataset_o2,]

selected_seu_list_o2 <- lapply(seu_str_recl_list_o2, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df_o2[
    resolution_df_o2$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
  
})

rm(seu_str_recl_list_o2)
gc()
```

```{r prepare_annotate_o2}

# mus tik stromal

seu_sign_o2 <- selected_seu_list_o2$seu_sign
reclustered_o2 <- base::unique(seu_sign_o2$seurat_clusters)

clust_list_o2 <- as.list(reclustered_o2)

ct_markers_o2 <- lapply(clust_list_o2, function(x){
  markers <- Seurat::FindMarkers(seu_sign_o2,
                                 test.use = "wilcox",
                                 ident.1 = x)
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  return(markers)
})
names(ct_markers_o2) <- reclustered_o2

ct_markers_df_o2 <- bind_rows(ct_markers_o2)
head(ct_markers_df_o2)

```

```{r prep_bcd}
# make a signature gene list with cell types of origin for helping annotation

signature_df <- data.frame(
  genes = unique(ensembl_sign$MMUS_SYMBOL),
  celltype = vector(length = length(unique(ensembl_sign$MMUS_SYMBOL)), 
                    mode = "character")
)

# ugly solution but works for this purpose
for(ct in names(geneset_list_str)){
  for(g in signature_df$genes){
    if(g %in% geneset_list_str[[ct]]$conserved_signature){
      signature_df[which(signature_df$genes == g),]$celltype <- paste(
        signature_df[which(signature_df$genes == g),]$celltype, ct)
    }
  }
}
```

### UMAPs

```{r umap_others_2}
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "seurat_clusters")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "batch")
# 4 = non-pbs
# 0 = pbs
# 7 = pbs

# 2 = non-pbs
# 3 = pbs

seu_sign_o2$annotated_reclustered <- vector(length = ncol(seu_sign_o2))
```

```{r umaps_batch, fig.height = 6, fig.width = 6}
Seurat::DimPlot(seu_sign_o2, reduction = "umap")+
  ggplot2::facet_wrap(
    facets = vars(seu_sign_o2@meta.data$batch), ncol = 2)
``` 

### Find markers and assign

```{r others2_cluster0}
nr <- "0"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Mgp", "Cxcl12", "Hp", "Notch3", "Ebf3", "Dpep1", "Cxcl14",
# "Cyp1b1", "Esm1"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Cxcl12")

SOI <-  c("Mgp", "Cxcl12", "Hp", "Notch3", "Ebf3", "Dpep1", "Cxcl14", "Cyp1b1",
          "Esm1")
signature_df[signature_df$genes %in% SOI,]

# Cxcl12 is by definition a CAR cells, some other genes are also clearly niche or
# adipo-related
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Adipo/CAR1, pbs"
```

```{r others2_cluster5}
nr <- "5"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Alpl", "Limch1", "Lrp4", "Dlx5", "Olfml3", "Sp7", "Pth1r", 
# "Satb2", "Cxcl12"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Cxcl12")

SOI <-  c("Alpl", "Limch1", "Lrp4", "Dlx5", "Olfml3", "Sp7", "Pth1r", "Satb2",
          "Cxcl12")
signature_df[signature_df$genes %in% SOI,]

# Alpl, Sp7, Pth1r osteo

# still enough cxcl12 expression, Limch1 is Osteo-CAR, other genes are also
# osteo-related but not too differentiated
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Osteo/CAR"
```

```{r others2_cluster4}
nr <- "4"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Angptl1", "Kitl", "Esm1", "Igfbp5", "Mme", "C1s1", "Dpep1", 
# "Vegfc", "Cxcl12"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Cxcl12")

SOI <-  c("Angpt1", "Kitl", "Esm1", "Igfbp5", "Mme", "C1s1", "Dpep1", "Vegfc",
          "Cxcl12")
signature_df[signature_df$genes %in% SOI,]

# Kitl, Cxcl12: CAR
# Dpep1, Esm1, Mme: adipo

# standard Adipo-CAR gene expression
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Adipo/CAR3, non-pbs"
```

```{r others2_cluster7}
nr <- "7"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Mmp13", "Slit2", "Lrp4", "Ptgis", "Gpc1", "Mrc2", "S100a6", "Satb2", 
# "Pth1r", "Cfh"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Mmp13", "Slit2", "Lrp4", "Ptgis", "Gpc1", "Mrc2", "S100a6", "Satb2",
          "Pth1r", "Cfh")
signature_df[signature_df$genes %in% SOI,]

# Sp7, Runx2, Alpl, Satb2, Ptgis, S100a6  strongly osteogenic
# Mmp13, Pth1r, Tpm1 also a bit chondro

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Osteo-Chondro"
```

```{r others2_cluster10}
nr <- "9"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Aspn", "Edil3", "Fmod", "Lum", "Mmp2", "Prrx2", "Sema3c"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Aspn", "Edil3", "Fmod", "Lum", "Mmp2", "Prrx2", "Sema3c")
signature_df[signature_df$genes %in% SOI,]

# this is classic for some sort of mixed fibro/chondro possible
# transitory or progenitor population 
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Fibro-Chondro"
```

```{r others2_cluster9}
nr <- "10"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Thbs4", "Clec3b", "Crispld2", "Tnxb", "Itgbl1", "Sema3c", "Prrx2"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50,  c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Thbs4", "Clec3b", "Crispld2", "Tnxb", "Itgbl1", "Sema3c", "Prrx2")
signature_df[signature_df$genes %in% SOI,]

# still some fibro-related erxpression with Clec3b but much more clearly 
# early chondro
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Chondro"
```

```{r others2_cluster11}
nr <- "11"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Scara5", "Pdgfrb", "Snrk", "Ebf3", "Ltbp3", "Mgp", "Hp", "Foxc1"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][,c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:4])
FeaturePlot(seu_sign_o2, markers[5:8])
FeaturePlot(seu_sign_o2, markers[9:12])
FeaturePlot(seu_sign_o2, markers[13:16])

SOI <-  c("Scara5", "Pdgfrb", "Snrk", "Ebf3", "Ltbp3", "Mgp", "Hp", "Foxc1")
signature_df[signature_df$genes %in% SOI,]

# adipo- or niche related expression
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Adipo/CAR2"
```

```{r others2_cluster8}
nr <- "8"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Serpinf1", "Sgms2", "Col5a2", "Cpe", "Dcn", "Satb2", "Sp7", 
# "Serpinh1", "Lum", "Ptgis"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Serpinf1", "Sgms2", "Col5a2", "Cpe", "Dcn", "Satb2", "Sp7",
          "Serpinh1", "Lum", "Ptgis")
signature_df[signature_df$genes %in% SOI,]

# with the Lum and Dcn, still looks kind of early/mixed-ish but Sp7 
# and Satb2 are clearly osteo lineage
seu_sign_o2$annotated_reclustered[seu_sign_o2$seurat_clusters == nr] <- "Osteo1"
```

```{r others2_cluster12}
nr <- "12"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Ackr3", "Jam2", "Mustn1", "Cryab", "Sgms2", "Ptgis"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Ackr3", "Jam2", "Mustn1", "Cryab", "Sgms2", "Ptgis")
signature_df[signature_df$genes %in% SOI,]

# losing some mixed/fibro gene expression
seu_sign_o2$annotated_reclustered[seu_sign_o2$seurat_clusters == nr] <- "Osteo2"
```

```{r others2_cluster13}

nr <- "13"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Serpinf1", "Cpe", "Col5a2", "Sgms2", "Mmp2", "Thbs1", "Lum", "Col5a1", 
# "Omd", "Ptgis", "Satb2", "Sp7"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

SOI <-  c("Serpinf1", "Cpe", "Col5a2", "Sgms2", "Mmp2", "Thbs1", "Lum",
          "Col5a1", "Omd", "Ptgis", "Satb2", "Sp7")
signature_df[signature_df$genes %in% SOI,]

# high expression of Osteo-markers like Sp7
seu_sign_o2$annotated_reclustered[seu_sign_o2$seurat_clusters == nr] <- "Osteo3"
```

```{r others2_cluster6}

nr <- "6"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check for "Mcf2l", "Cd34", "Lims2", "Mecom", "Cav1", "Sparcl1" in rerun
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

# marker genes (cap)
FeaturePlot(seu_sign_o2, "Mcf2l")
FeaturePlot(seu_sign_o2, "Cd34")
FeaturePlot(seu_sign_o2, "Lims2")
FeaturePlot(seu_sign_o2, "Mecom")
FeaturePlot(seu_sign_o2, "Cav1")
FeaturePlot(seu_sign_o2, "Sparcl1")

# arteriolar genes
FeaturePlot(seu_sign_o2, "Sox18")
FeaturePlot(seu_sign_o2, "Efnb2")

# cap gene
FeaturePlot(seu_sign_o2, "Rgcc") 

SOI <-  c("Mcf2l", "Cd34", "Lims2", "Mecom", "Cav1", "Sparcl1")
signature_df[signature_df$genes %in% SOI,]

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Art/Cap-EC"
```

```{r others2_cluster1}

nr <- "1"

markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check for "Axl", "Lpar6", "Clec14a", "Ephb4" in rerun
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Axl")
FeaturePlot(seu_sign_o2, "Lpar6")
FeaturePlot(seu_sign_o2, "Aebp1")
FeaturePlot(seu_sign_o2, "Tgfbi")
FeaturePlot(seu_sign_o2, "Clec14a")
FeaturePlot(seu_sign_o2, "Prex2")
FeaturePlot(seu_sign_o2, "Lrg1")

# AEC
FeaturePlot(seu_sign_o2, "Sox18")
FeaturePlot(seu_sign_o2, "Clec14a")

# VEC
FeaturePlot(seu_sign_o2, "Nr2f2")
FeaturePlot(seu_sign_o2, "Ephb4")
FeaturePlot(seu_sign_o2, "Vcam1")

# capillary
FeaturePlot(seu_sign_o2, "Rgcc")
FeaturePlot(seu_sign_o2, "Cd34")

# this cluster is mostly marked by NEGATIVE markers

SOI <-  c("Axl", "Lpar6", "Clec14a", "Ephb4")
signature_df[signature_df$genes %in% SOI,]

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Ven EC1, non-pbs"
```

```{r others2_cluster2}

nr <- "2"
markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Tmcc3", "Dll4", "Klf4", "Itga6", "Nr2f2", "Vcam1", "Nfkbia", "Spry2"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,][1:50, c(1, 3, 6)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Tmcc3")
FeaturePlot(seu_sign_o2, "Spry2")
FeaturePlot(seu_sign_o2, "Sdc4")
FeaturePlot(seu_sign_o2, "Nfkbia")

FeaturePlot(seu_sign_o2, "Klf4") # angiogenesis
FeaturePlot(seu_sign_o2, "Rgcc") # capillary
 
# VEC 
FeaturePlot(seu_sign_o2, "Nr2f2") 
FeaturePlot(seu_sign_o2, "Ephb4") 
FeaturePlot(seu_sign_o2, "Vcam1") 

SOI <-  c("Tmcc3", "Dll4", "Klf4", "Itga6", "Nr2f2", "Vcam1", "Nfkbia", "Spry2")
signature_df[signature_df$genes %in% SOI,]

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Ven EC3, non-pbs"
```

```{r others2_cluster3}

nr <- "3"

markers <- ct_markers_o2[[nr]]$genes[
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL]

# check "Fabp4", "Ier3", "Cd36", "Lrg1", "Srgn", "Cldn5", "Fam167b", "Vcam1"
ct_markers_o2[[nr]][
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05 &
    ct_markers_o2[[nr]]$genes %in% ensembl_sign$MMUS_SYMBOL,]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

FeaturePlot(seu_sign_o2, "Vcam1") # activation
FeaturePlot(seu_sign_o2, "Kdr") # activation
FeaturePlot(seu_sign_o2, "Cd36") # activation all kind of unspecific

FeaturePlot(seu_sign_o2, "Ier3") 
FeaturePlot(seu_sign_o2, "Fabp4")
FeaturePlot(seu_sign_o2, "Fam167b") 

FeaturePlot(seu_sign_o2, "Il6st")
FeaturePlot(seu_sign_o2, "Fabp4")
FeaturePlot(seu_sign_o2, "Lrg1")

SOI <-  c("Fabp4", "Ier3", "Cd36", "Lrg1", "Srgn", "Cldn5", "Fam167b", "Vcam1")
signature_df[signature_df$genes %in% SOI,]

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Ven EC2, pbs"
```

```{r factor}
seu_sign_o2$annotated_reclustered <- factor(
  seu_sign_o2$annotated_reclustered,
  levels = c(
    "Adipo/CAR1, pbs",
    "Adipo/CAR2",
    "Adipo/CAR3, non-pbs",
    "Osteo/CAR",
    "Osteo-Chondro",
    "Fibro-Chondro",
    "Chondro",
    "Osteo1",
    "Osteo2",
    "Osteo3",
    "Art/Cap-EC",
    "Ven EC1, non-pbs",
    "Ven EC2, pbs",
    "Ven EC3, non-pbs"
))
stopifnot(!is.na(seu_sign_o2$annotated_reclustered))
```

## Add clusters from the signature-based re-clustering

For this dataset, there was no custom labelling so there is nothing to add.

## Save

```{r save}
saveRDS(seu_sign_o2, output_path)
```

```{r session_info}
utils::sessionInfo()
```
