---
title: "Blind annotation of LI ET AL STROMAL"
date: '2025-12-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Blind annotation of two human HSPC datasets and one mouse and one human
niche cell datasets after unbiased re-clustering with signature genes.
This is the script for annotating LI ET AL STROMAL.
The obtained clusters will now be annotated based only on the expression
of signature gene orthologs within these clusters, and without knowledge of
which cells from which originally annotated cell type these clusters contain.

This is another occasion where I will save an object from an .Rmd file because
after naming it and looking at it this way, I'm already adding the annotations.

These parts of code were originally in the first iteration of Figure 4 but after
adding the custom annotations, I have decided to separate the blind annotations
from the figure, too.
I didn't change too much from the original figure script though.
Instead, I added some comments why I chose each annotation.

Check supplementary table 1 of the corresponding publication for some of the 
marker gene sources.

#### Load objects

```{r seed, message = FALSE}
base::RNGkind("L'Ecuyer-CMRG")
set.seed(37)
```

```{r load_packages, message = FALSE}
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(Seurat, quietly = TRUE)
```

```{r define_basepath}
base_path <- snakemake@input[["base_path"]]
metadata_path <- snakemake@input[["metadata_path"]]
```

```{r load_short}

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  metadata_path,
  "/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(
  file = resolution_df_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

ensembl_sign <- readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/02_endf/ensembl_sign_str"))

# list of gene sets for HSPCs (conserved signature, conserved markers, etc.)
# required for getting more info on the signature genes
geneset_list_str <-base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/01_gens/geneset_list_str"))
```

The name other 3 remains from previous figure 5 iterations.

```{r load_other3}

dataset_o3 <- "li_all_stromal"

# list of seurat objects with reclustering vectors for each gene set and resolution
seu_str_recl_list_o3 <- base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_li_all_stromal_list"))
```

```{r export_paths}
output_path <- base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/li_all_stromal.Rds")
print(output_path)
```

## Blind Re-annotation 

### Prepare

Calculate markers from the selected object (selected resolution).
Because the object was subset only to signature gene orthologs, 
all resulting marker genes are by definition also signature gene orhtologs.

```{r selected_seu_o3}

# get the selected object based on the chosen resolution from the metadata df
resolution_df_o3 <- resolution_df[resolution_df$dataset == dataset_o3,]

selected_seu_list_o3 <- lapply(seu_str_recl_list_o3, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df_o3[
    resolution_df_o3$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
  
})

rm(seu_str_recl_list_o3)
gc()
```

```{r prepare_annotate_o3}

seu_sign_o3 <- selected_seu_list_o3$seu_sign
reclustered_o3 <- base::unique(seu_sign_o3$seurat_clusters)

clust_list_o3 <- as.list(reclustered_o3)

ct_markers_o3 <- lapply(clust_list_o3, function(x){
  markers <- Seurat::FindMarkers(seu_sign_o3,
                                 test.use = "wilcox",
                                 ident.1 = x)
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  # need to add only ensembl gene symbols as I will not be using others
  markers$symbol <- markers$genes
  markers$symbol <- ensembl_sign$HSAPIENS_SYMBOL[
    base::match(markers$genes, ensembl_sign$ENSG_ID)]
  return(markers)
})
names(ct_markers_o3) <- reclustered_o3

ct_markers_df_o3 <- bind_rows(ct_markers_o3)
head(ct_markers_df_o3)

```

```{r prep_bcd}
# make a signature gene list with cell types of origin for helping annotation

signature_df <- data.frame(
  genes = unique(ensembl_sign$MMUS_SYMBOL),
  celltype = vector(length = length(unique(ensembl_sign$MMUS_SYMBOL)), 
                    mode = "character")
)

# ugly solution but works for this purpose
for(ct in names(geneset_list_str)){
  for(g in signature_df$genes){
    if(g %in% geneset_list_str[[ct]]$conserved_signature){
      signature_df[which(signature_df$genes == g),]$celltype <- paste(
        signature_df[which(signature_df$genes == g),]$celltype, ct)
    }
  }
}
```

```{r annotated_reclustered}
# make empty vector to fill
seu_sign_o3$annotated_reclustered <- vector(length = ncol(seu_sign_o3))
```

### UMAPs

```{r umap_others_3}
Seurat::DimPlot(seu_sign_o3, dims = c(1,2), group.by = "seurat_clusters")
Seurat::DimPlot(seu_sign_o3, dims = c(1,2), group.by = "orig.ident")
# no coldata with possible batch?
```

### Print markers and annotate

order: 6, 1, 2, 0, 7, 4, 5, 3

```{r others3_cluster6}
nr <- "6"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CXCL12", "CXCL14", "EDIL3", "ABI3BP", "SOD3", "COL6A3", "ABCA8" rerun
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000107562") #CXCL12
FeaturePlot(seu_sign_o3, "ENSG00000145824") #CXCL14
FeaturePlot(seu_sign_o3, "ENSG00000164176") #EDIL3
FeaturePlot(seu_sign_o3, "ENSG00000154175") #ABI3BP
FeaturePlot(seu_sign_o3, "ENSG00000109610") #SOD3
FeaturePlot(seu_sign_o3, "ENSG00000163359") #COL6A3
FeaturePlot(seu_sign_o3, "ENSG00000141338") #ABCA8

# genes of interest
GOI <- c("CXCL12", "CXCL14", "EDIL3", "ABI3BP", "SOD3", "COL6A3", "ABCA8")
# mouse symbols of interest
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
# cell types of origin for signature genes
signature_df[signature_df$genes %in% SOI,]

# CXCL12, CXCL14
# ABCA8, EDIL3, ABI3BP, SOD3, COL6A3 fibro-like from list of signature genes

# by definition, CXCL12 expressing cells are CARs
# these also express some fibro genes
seu_sign_o3$annotated_reclustered[
  seu_sign_o3$seurat_clusters == nr] <- "Fibro-like CAR2"
```

```{r others3_cluster1}
nr <- "1"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CXCL12", "FSTL1", "MME", "HP", "TJP1", "SPARCL1", "FBN1", "RSPO3",
# "FBLN5" 
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6]) # a lot of these are lowly expressed
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000107562") #CXCL12
FeaturePlot(seu_sign_o3, "ENSG00000163430") #FSTL1
FeaturePlot(seu_sign_o3, "ENSG00000196549") #MME
FeaturePlot(seu_sign_o3, "ENSG00000257017") #HP
FeaturePlot(seu_sign_o3, "ENSG00000104067") #TJP1
FeaturePlot(seu_sign_o3, "ENSG00000152583") #SPARCL1
FeaturePlot(seu_sign_o3, "ENSG00000166147") #FBN1
FeaturePlot(seu_sign_o3, "ENSG00000146374") #RSPO3
FeaturePlot(seu_sign_o3, "ENSG00000140092") #FBLN5

GOI <- c("CXCL12", "FSTL1", "MME", "HP", "TJP1", "SPARCL1", "FBN1",
         "RSPO3", "FBLN5")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

# CXCL12: CAR cell
# FSTL1, HP, MME, FBN1, RSPO3 all from "Adipo/CAR"

# CAR cells without much other lineage expression 
seu_sign_o3$annotated_reclustered[seu_sign_o3$seurat_clusters == nr] <- "CAR"
```

```{r others3_cluster2}
nr <- "2"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CXCL12", "KITL", "PDGFRB", "FBN1", "COL3A1", "C1S", "SPARCL1", 
# "ABCA8" 
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][,c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000107562") #CXCL12
FeaturePlot(seu_sign_o3, "ENSG00000049130") #KITL
FeaturePlot(seu_sign_o3, "ENSG00000113721") #PDGFRB
FeaturePlot(seu_sign_o3, "ENSG00000166147") #FBN1
FeaturePlot(seu_sign_o3, "ENSG00000168542") #COL3A1
FeaturePlot(seu_sign_o3, "ENSG00000182326") #C1S
FeaturePlot(seu_sign_o3, "ENSG00000152583") #SPARCL1
FeaturePlot(seu_sign_o3, "ENSG00000141338") #ABCA8

GOI <- c("CXCL12", "KITL", "PDGFRB", "FBN1", "COL3A1", "C1S", "SPARCL1",
         "ABCA8")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

# CXCL14, CXCL12
# PDGFRB, FBN1, COL3A1, C1S, ABCA8 = fibro-like
# SPARCL1

# again some fibro-like expression in addition to CXCL12
seu_sign_o3$annotated_reclustered[
  seu_sign_o3$seurat_clusters == nr] <- "Fibro-like CAR1"
```

```{r others3_cluster0}
nr <- "0"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CXCL12", "EMP1", "ID1", "DEPP1", "LMNA", "FABP4", "OSR2", "PPP1R15A" 
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][,c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000107562") #CXCL12
FeaturePlot(seu_sign_o3, "ENSG00000134531") #EMP1
FeaturePlot(seu_sign_o3, "ENSG00000166147") #ID1
FeaturePlot(seu_sign_o3, "ENSG00000160789") #LMNA  
FeaturePlot(seu_sign_o3, "ENSG00000170323") #FABP4
FeaturePlot(seu_sign_o3, "ENSG00000164920") #OSR2
FeaturePlot(seu_sign_o3, "ENSG00000087074") #PPP1R15A

GOI <- c("CXCL12", "EMP1", "ID1", "LMNA", "FABP4", "OSR2", "PPP1R15A")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

# CXCL12: CAR cell
# ID1, LMNA, OSR2, 
# FABP4, PPP1R15A -> adipocytic lineage

seu_sign_o3$annotated_reclustered[
  seu_sign_o3$seurat_clusters == nr] <- "Adipo-like CAR"
```

```{r}
DimPlot(seu_sign_o3, dims = c(1,2), group.by = "seurat_clusters")
```

```{r others3_cluster7}
nr <- "7"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CPE", "SP7", "PTPRD", "RUNX2", "PALS2", "OMD", "MMP13", "PTH1R", "CD9" 
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000107562") #CXCL12 actually only weakly expr.

GOI <- c("CPE", "SP7", "PTPRD", "RUNX2", "PALS2", "OMD", "MMP13",
         "PTH1R", "CD9")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

# CPE, SP7, PTPRD, RUNX2, PALS2, OMD, MMP13, PTH1R, CD9

# some of these genes are clearly osteo (SP7), some more clearly chondro (CD9)
seu_sign_o3$annotated_reclustered[
  seu_sign_o3$seurat_clusters == nr] <- "Osteo-Chondro"
```

```{r}
DimPlot(seu_sign_o3, dims = c(1,2), group.by = "seurat_clusters")
```

I know these are ECs but I couldn't properly discern which subtypes.

```{r others3_cluster5}
nr <- "5"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CD34", "RGCC", "TSPAN13", "CAPG", "MGST1", "GPX1",
# "PPP1R14B", "TXN", "CD36", "NOTCH1"
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000106537") # TSPAN13
FeaturePlot(seu_sign_o3, "ENSG00000149564") # ESAM # BARELY EXPR
FeaturePlot(seu_sign_o3, "ENSG00000042493") # CAPG
FeaturePlot(seu_sign_o3, "ENSG00000174059") # CD34
FeaturePlot(seu_sign_o3, "ENSG00000102760") # RGCC
FeaturePlot(seu_sign_o3, "ENSG00000008394") # MGST1
FeaturePlot(seu_sign_o3, "ENSG00000135218") # CD36
FeaturePlot(seu_sign_o3, "ENSG00000148400") # NOTCH1

GOI <- c("CD34", "RGCC", "TSPAN13", "CAPG", "MGST1", "GPX1",
         "PPP1R14B", "TXN", "CD36", "NOTCH1")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

seu_sign_o3$annotated_reclustered[seu_sign_o3$seurat_clusters == nr] <- "EC1"
```

```{r others3_cluster3}
nr <- "3"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "PECAM1", "FLI1", "ICAM2", "CRIP1", "FCER1G", "SH2D3C"
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000261371") #PECAM1 
FeaturePlot(seu_sign_o3, "ENSG00000151702") #FLI1

FeaturePlot(seu_sign_o3, "ENSG00000108622") #ICAM2
FeaturePlot(seu_sign_o3, "ENSG00000213145") #CRIP1
FeaturePlot(seu_sign_o3, "ENSG00000158869") #FCER1G
FeaturePlot(seu_sign_o3, "ENSG00000108622") #ICAM2
FeaturePlot(seu_sign_o3, "ENSG00000095370") #SH2D3C

GOI <- c("PECAM1", "FLI1", "ICAM2", "CRIP1", "FCER1G", "SH2D3C")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

seu_sign_o3$annotated_reclustered[seu_sign_o3$seurat_clusters == nr] <- "EC3"
```

```{r others3_cluster4}
nr <- "4"
markers <- ct_markers_o3[[nr]]$gene[
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05]

# check "CELF2", "LYN", "PECAM1", "SRGN", "NFKBIA"
ct_markers_o3[[nr]][
  !is.na(ct_markers_o3[[nr]]$symbol) & 
    ct_markers_o3[[nr]]$avg_log2FC > 0 &
    ct_markers_o3[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o3, markers[1:6])
FeaturePlot(seu_sign_o3, markers[7:12])
FeaturePlot(seu_sign_o3, markers[13:18])

FeaturePlot(seu_sign_o3, "ENSG00000048740") #CELF2
FeaturePlot(seu_sign_o3, "ENSG00000254087") #LYN
FeaturePlot(seu_sign_o3, "ENSG00000261371") #PECAM1
FeaturePlot(seu_sign_o3, "ENSG00000122862") #SRGN
FeaturePlot(seu_sign_o3, "ENSG00000100906") #NFKBIA

GOI <- c("CELF2", "LYN", "PECAM1", "SRGN", "NFKBIA")
SOI <- ensembl_sign$MMUS_SYMBOL[ensembl_sign$HSAPIENS_SYMBOL %in% GOI]
signature_df[signature_df$genes %in% SOI,]

seu_sign_o3$annotated_reclustered[seu_sign_o3$seurat_clusters == nr] <- "EC2"
```

```{r factor}
seu_sign_o3$annotated_reclustered <- factor(
  seu_sign_o3$annotated_reclustered,
  levels = c(
    "Fibro-like CAR1",
    "Fibro-like CAR2",
    "CAR",
    "Adipo-like CAR",
    "Osteo-Chondro",
    "EC1",
    "EC2",
    "EC3"
    ))
stopifnot(!is.na(seu_sign_o3$annotated_reclustered))
```

## Add clusters from the signature-based re-clustering

For this dataset, there was no custom labelling so there is nothing to add.

## Save

```{r save}
saveRDS(seu_sign_o3, output_path)
```

```{r session_info}
utils::sessionInfo()
```
