---
title: "Blind annotation of TABULA SAPIENS HSPC PROGENITORS"
author: "Lea Wölbert"
date: '2025-12-09'
output: html_document
---

Blind annotation of two human HSPC datasets and one mouse and one human
niche cell datasets after unbiased re-clustering with signature genes.
This is the script for annotating TABULA SAPIENS HSPC/PROGENITORS from fetal data.
The thusly obtained clusters will now be annotated based only on the expression
of signature gene orthologs within these clusters, and without knowledge of
which cells from which originally annotated cell type these clusters contain.

This is another occasion where I will save an object from an .Rmd file because
after naming it and looking at it this way, I'm already adding the annotations.

These parts of code were originally in the first iteration of Figure 4 but after
adding the custom annotations, I have decided to separate the blind annotations
from the figure, too.
I didn't change too much from the original figure script though.
Instead, I added some comments why I chose each annotation.

#### Load objects

```{r seed, message = FALSE}
RNGkind("L'Ecuyer-CMRG") 
set.seed(37)
```

```{r load_packages, message = FALSE}
library(Seurat, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(ggpubr, quietly = TRUE)
library(tidyverse, quietly = TRUE)
```

```{r base_path}
base_path <- snakemake@input[["base_path"]]
metadata_path <- snakemake@input[["metadata_path"]]
#base_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/"
```

```{r load_objects}

# data frame containing the correct resolutions for other datasets as visually determined
resolution_df_path <- base::paste0(
  metadata_path,
  "/scRNAseq/03_sce_analysis/reclustering_bm/reclustering_other_resolution.txt")

resolution_df <- utils::read.csv(file = resolution_df_path, 
                               header = TRUE, 
                               sep = ";", 
                               check.names=FALSE, 
                               stringsAsFactors=FALSE, 
                               as.is=TRUE, 
                               colClasses = "character")

ensembl_sign <- readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/01_reclustering_own/02_endf/ensembl_sign_hsc"))
```

Name "other 2" is a remnant from previous figure iterations

```{r load_other2}
dataset_o2 <- "ts_hscs_progenitors"

# list of seurat objects with reclustering vectors for each gene set+ resolution
seu_hsc_recl_list_o2 <- base::readRDS(base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/02_reclustering_other/03_recl/reclustered_ts_hscs_progenitors_list"))
```

```{r export_paths}
output_path <- base::paste0(
  base_path,
  "/sce_objects/03_sce_analysis/04_signatures/04_blind_annotation/ts_hscs_progenitors.Rds")
print(output_path)
```

## Blindly annotate

### Prep

```{r selected_seu_o2}

resolution_df_o2 <- resolution_df[resolution_df$dataset == dataset_o2,]

selected_seu_list_o2 <- lapply(seu_hsc_recl_list_o2, function(seu_list){
  
  conslev_curr <- seu_list[[1]]@misc$used_genes
  
  resdf_temp <- resolution_df_o2[
    resolution_df_o2$conservation_level == conslev_curr,]
  res_curr <- resdf_temp$resolution
  
  seu_return <- seu_list[[res_curr]]
  return(seu_return)
  
})

rm(seu_hsc_recl_list_o2)
gc()
```

```{r prepare_annotate_o2}

seu_sign_o2 <- selected_seu_list_o2$seu_sign
reclustered_o2 <- base::unique(seu_sign_o2$seurat_clusters)

clust_list_o2 <- as.list(reclustered_o2)

ct_markers_o2 <- lapply(clust_list_o2, function(x){
  markers <- Seurat::FindMarkers(seu_sign_o2,
                                 test.use = "wilcox",
                                 ident.1 = x)
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  # need to add only ensembl gene symbols as I will not be using others
  markers$symbol <- markers$genes
  markers$symbol <- ensembl_sign$HSAPIENS_SYMBOL[
    base::match(markers$genes, ensembl_sign$ENSG_ID)]
  return(markers)
})
names(ct_markers_o2) <- reclustered_o2

ct_markers_df_o2 <- bind_rows(ct_markers_o2)
head(ct_markers_df_o2)
```

Get more markers between specific sub-clusters.

```{r prepare_annotate_o2_hsc_only}

seu_sign_o2_hsc <- seu_sign_o2[,seu_sign_o2$seurat_clusters %in%
                                 c("4", "10", "1")]
reclustered_o2_hsc <- base::unique(seu_sign_o2_hsc$seurat_clusters)

clust_list_o2_hsc <- as.list(unfactor(reclustered_o2_hsc))

ct_markers_o2_hsc <- lapply(clust_list_o2_hsc, function(x){
  markers <- Seurat::FindMarkers(seu_sign_o2_hsc,
                                 test.use = "wilcox",
                                 ident.1 = x)
  markers <- markers[base::order(base::abs(markers$avg_log2FC), 
                                 decreasing=TRUE),]
  markers$which_cluster <- base::rep(x, nrow(markers))
  markers <- rownames_to_column(markers, var = "genes")
  
  # need to add only ensembl gene symbols as I will not be using others
  markers$symbol <- markers$genes
  markers$symbol <- ensembl_sign$HSAPIENS_SYMBOL[
    base::match(markers$genes, ensembl_sign$ENSG_ID)]
  return(markers)
})
names(ct_markers_o2_hsc) <- reclustered_o2_hsc

ct_markers_df_o2_hsc <- bind_rows(ct_markers_o2_hsc)
head(ct_markers_df_o2_hsc)
```

### UMAPs

```{r umap_others_2}
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "seurat_clusters")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "assay")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "donor_id")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "sex")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "development_stage")
DimPlot(seu_sign_o2, dims = c(1,2), group.by = "tissue")

seu_sign_o2$annotated_reclustered <- vector(length = ncol(seu_sign_o2))
# clusters 3, 9 and 7 are batch V1
# clusters 1, 8 and 6 are batch V2
```

```{r umaps_dev, fig.height = 10, fig.width = 6}
DimPlot(seu_sign_o2, reduction = "umap")+
  facet_wrap(facets = vars(seu_sign_o2@meta.data$development_stage), ncol = 2)
``` 

```{r umaps_sex, fig.height = 3, fig.width = 6}
DimPlot(seu_sign_o2, reduction = "umap")+
  facet_wrap(facets = vars(seu_sign_o2@meta.data$sex), ncol = 2)
``` 

```{r umaps_assay, fig.height = 3, fig.width = 6}
DimPlot(seu_sign_o2, reduction = "umap")+
  facet_wrap(facets = vars(seu_sign_o2@meta.data$assay), ncol = 2)
``` 

### Show markers

```{r others2_cluster0}
nr <- "0"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# ANXA2, CTSZ, CLEC12A, H1-10, LY86, BLVRB, CEBPE, MPO
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
# Early myeloid (MPO, CLEC12a, SPI1) granu (ANXA2, CEBPE,CTSZ)
# mono (LY86, CLEC12A, LAMP1 I think, too) 
# I had previously called this "cycling", too but cannot explain why atm

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Early Granu/Mono"
```

```{r others2_cluster1}
nr <- "2"
markers <- ct_markers_o2[[nr]]$gene[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# PRTN3, TIFAB, MPO, MS4A3, CTSG, CEBPE, MS4A2, MS4A2, CALR, CLEC12A
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# CALR, SRGN, CEBPE, PRSS57, HSP90B1, CALR, ANXA1 go more into
# granu/neutro direction 

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Granu"
```

```{r others2_cluster6}
nr <- "4"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# CTSG, PRTN3, MPO, SRGN, CEPBE, MS4A2, CLEC12A
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# strong CTSG, PRTN3, MPO, without much downstream neutro genes
# like S100A genes always look (earlier) myeloid to me, not clearly
# neutro, granu or mono

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Myeloid"
```

```{r others2_cluster10}
nr <- "10"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# TMEM176B, TMEM176A, ANXA2, CLEC12A, TSPO, LY86, CTSZ, CEBPE, 
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
FeaturePlot(seu_sign_o2, markers[13:18])

# TMEM176B and TMEM176A are mono dendritic genes, The others are 
# mostly generally myeloid except CEBPE, LAMP1 is found in monos
seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Mono/Dendritic"
```

```{r others2_cluster4}
nr <- "5"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# LTB, TRIB2, TCF4, MEF2C, SMIM14, LY86, RAMP1, LSP1, FUT8, CD34, LMO4, CD37
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# LTB, TCF4, LY86, RAMP1, CD37 are lymphoid
# more "classic" lymphoid markers are not in the signature

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Lymphoid"
```

```{r others2_cluster3}
nr <- "3"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# MPL, HLF, MECOM, CD34, ANGPT1, RAMP1, MEIS1
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# MPL, HLF, CD34, MECOM, MEIS1 all classic HSC genes in human

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "HSC1 10xV1"
```

```{r others2_cluster5}
nr <- "7"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# CDCA3, CENPA, CENPE, CDC20, MKI67, AURKB, CDK1, H1-4, MPL, NKX2-3, CD34, 
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
# clear cycling genes but lack of lineage markers and expression of
# MPL, NKX2-3, CD34 etc a little bit down the line instead of lineage markers

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Cycling Early 10xV1"
```

```{r others2_cluster2}
nr <- "1"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# HLF, NFIA, VAMP5, MECOM, ANGPT1, MEIS1, RAMP1, GCNT2, CD34, TXNIP
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# again HLF, MSI2, MECOM are classic HSC, maybe a slight Mk hint with VAMP5 
# but nonetheless mostly HSC

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "HSC2 10xV2"
```

```{r others2_cluster9}
nr <- "9"
markers <- ct_markers_o2[[nr]]$genes[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# GATA1, ITGA2B, KLF1, GATA2, MPL, CENPV, EPOR, LMO4, ZFPM1, AQP1, MEIS1
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
FeaturePlot(seu_sign_o2, markers[7:12])
# GATA1, ITGA2B, GATA2, KLF1, MPL, EPOR indicate some early Mk ery prog

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Mk/Ery1 10xV1"
```

```{r others2_cluster8}
nr <- "8"
markers <- ct_markers_o2[[nr]]$gene[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# CENPA, AURKB, CDCA3, CDK1, CDC20, CENPE, MKI67
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# cycling genes, lack of lineage markers and also lack of HSC markers 
# I can recognize, making it "only" cycling

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Cycling 10xV2"
```

```{r others2_cluster7}
nr <- "6"
markers <- ct_markers_o2[[nr]]$gene[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# KLF1, ITGA2B, GATA1, EPOR, AQP1, GATA2, LMO2, ZFPM1, CDK1, CENPV, AURKB
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# KLF1, ITGA2B, GATA1, EOIR, GATA2 together, also TRIB2, so mk/ery again

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "Mk/Ery2 10xV2"
```

```{r others2_cluster11}
nr <- "11"
markers <- ct_markers_o2[[nr]]$gene[
  !is.na(ct_markers_o2[[nr]]$symbol) &
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05]

# C17orf50, RAMP1, HLF, CD34, MPL, ANPGT1, MECOM
ct_markers_o2[[nr]][
  !is.na(ct_markers_o2[[nr]]$symbol) & 
    ct_markers_o2[[nr]]$avg_log2FC > 0 &
    ct_markers_o2[[nr]]$p_val_adj < 0.05,][1:50, c(1, 3, 6, 8)]

FeaturePlot(seu_sign_o2, markers[1:6])
# interesting that C17orf50 appears here again like in ts_bonemarrow
# NKF2-3, HLF, MPL, MECOM are all HSC signature genes

seu_sign_o2$annotated_reclustered[
  seu_sign_o2$seurat_clusters == nr] <- "HSC3"
unique(seu_sign_o2$annotated_reclustered)
```

Factorize for nicer plots.

```{r others2_level}
seu_sign_o2$annotated_reclustered <- factor(
  seu_sign_o2$annotated_reclustered,
  levels = c(  
    "HSC1 10xV1",
    "HSC2 10xV2",
    "HSC3",
    "Cycling Early 10xV1",
    "Cycling 10xV2",
    "Granu",
    "Myeloid",
    "Mono/Dendritic",
    "Early Granu/Mono",
    "Lymphoid",
    "Mk/Ery1 10xV1",
    "Mk/Ery2 10xV2"))

stopifnot(!is.na(seu_sign_o2$annotated_reclustered))
```

## Add clusters from the signature-based re-clustering

For this dataset, there was custom labelling so there is nothing to add.

## Save

```{r save}
saveRDS(seu_sign_o2, output_path)
```

```{r sessioninfo}
utils::sessionInfo()
```