---
title: "Subclustering results report - Dotplots"
author: "Lea Wölbert"
date: '2023-07-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on the results of the final annotation.
This report is used to make many plots for presentations, etc., so it is 
very chaotic and frequently adjusted.
Only for dotplots using ggplot (for now).

Since I use a lot of plots from this report, there are many different
more or less good-looking versions of the same plot in this script.

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r testing, eval = FALSE, include = FALSE}
source("../../source/plotting.R")
colors_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/colors/colors.txt"
source("../../source/colors.R")

dtplt_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/scRNAseq/02_sce_anno/gene_list_dotplot.txt"
dtplt_short_path<- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/scRNAseq/02_sce_anno/gene_list_dotplot_short.txt"

source("../../source/colors.R")

#genelist_shared <- base::readRDS(file = snakemake@input[["genes_list_shared"]])
sce <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10")
umap_base_l(sce_anno, color_by = "celltypes")+
  scale_color_manual(values = col_cts_str)

```

```{r load, message = FALSE}
library(scuttle, quietly = TRUE) 
```

```{r sources, message = FALSE}
#source(file = snakemake@params[["plotting"]])
#source(file = snakemake@params[["functions"]])
```

```{r load_objects, message = FALSE}
#load SCE objects
sce <- base::readRDS(file = snakemake@input[["sce_input"]])

nr_hvgs <- snakemake@params[["nr_hvgs"]]
fraction_curr <- snakemake@wildcards[["fraction"]]

# load gene list for dotplots
dtplt_path <- snakemake@input[["gene_list_dtplt"]]
# load shorter gene list for prettier dotplots 
dtplt_short_path <- snakemake@input[["gene_list_dtplt_short"]]
```

```{r load_objects_csv, message = FALSE}
gene_list_dtplt <- utils::read.csv(dtplt_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")
gene_list_dtplt <- gene_list_dtplt[gene_list_dtplt$fraction == fraction_curr,]

gene_list_dtplt_short <- utils::read.csv(dtplt_short_path, 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")

# subset to get list to annotate ECs
gene_list_dtplt_ECs <- gene_list_dtplt_short[
  gene_list_dtplt_short$use == "ECs",] 
gene_list_dtplt_short <- gene_list_dtplt_short[
  gene_list_dtplt_short$use == "short",] 
```

```{r colors, message = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

# for factorizing cell types later
if(fraction_curr == "hsc"){
  col_cts <- col_cts_hsc
}else if(fraction_curr == "str"){
  col_cts <- col_cts_str
}

# factorize as required
sce$Species_ID <- factor(sce$Species_ID, levels = names(col_spc))
```

```{r print}
print(col_cts)
```

# Gene expression

## Prepare

```{r aggregate}
agg <- scuttle::aggregateAcrossCells(
  sce, 
  id=colData(sce)[,c("celltypes")], 
  statistics = "mean",
  use.assay.type = "logcounts") # normalized (MultiBatchNorm)
```

```{r vis_df_all}
# generate a dataframe containing only genes that will be visualised
all_genes <- base::unique(c(gene_list_dtplt$gene,
                            gene_list_dtplt_short$gene, 
                            gene_list_dtplt_ECs$gene))

vis_df_all <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg)$logcounts[
    rownames(SummarizedExperiment::assays(agg)$logcounts) %in% all_genes,]))
```

```{r vis_df_all_prep}
# further prepare the dataframe for visualisation
vis_df_all <- tibble::rownames_to_column(vis_df_all, "celltypes")

vis_df_all <- tidyr::pivot_longer(
  vis_df_all, 
  cols = c(2:ncol(vis_df_all)), 
  values_to = "expression", 
  names_to = "gene")
```

```{r perc_expressed}
# subset and factor

vis_df_all$celltypes <- factor(
  vis_df_all$celltypes, 
  levels = base::rev(names(col_cts)))

gene_list_dtplt$gene[which(duplicated(gene_list_dtplt$gene))]
gene_list_dtplt_short$gene[which(duplicated(gene_list_dtplt_short$gene))]
gene_list_dtplt_ECs$gene[which(duplicated(gene_list_dtplt_ECs$gene))]

vis_df_main <- vis_df_all[vis_df_all$gene %in% gene_list_dtplt$gene,]
vis_df_short <- vis_df_all[vis_df_all$gene %in% gene_list_dtplt_short$gene,]
vis_df_ECs <- vis_df_all[vis_df_all$gene %in% gene_list_dtplt_ECs$gene,]

vis_df_main$gene <- factor(
  vis_df_main$gene, 
  levels = gene_list_dtplt$gene)

vis_df_short$gene <- factor(
  vis_df_short$gene, 
  levels = gene_list_dtplt_short$gene)

vis_df_ECs$gene <- factor(
  vis_df_ECs$gene, 
  levels = gene_list_dtplt_ECs$gene)
```

```{r perc_expressed_func}

get_perc_expressed <- function(vis_df){
  
  vis_df$percent_expressed <- vector(length = nrow(vis_df))
  
  cts <- base::unique(vis_df$celltypes)
  gs <- base::unique(vis_df$gene)
  
  for(ct in cts){
    
    lc_temp <- SummarizedExperiment::assays(sce)$logcounts[,sce$celltypes == ct]
    
    for(g in gs){
      
      perc_expr <- length(which(lc_temp[
        rownames(lc_temp) == g,] > 0))/ncol(lc_temp)*100
      
      vis_df$percent_expressed[
        which(vis_df$celltypes == ct & vis_df$gene == g)] <- perc_expr
      
    }
  }
  
  return(vis_df)
}

vis_df_main <- get_perc_expressed(vis_df_main)

# only needs to be calculated for str
if(fraction_curr == "str"){
  vis_df_short <- get_perc_expressed(vis_df_short)
  vis_df_ECs <- get_perc_expressed(vis_df_ECs)
  
  vis_df_ECs <- vis_df_ECs[vis_df_ECs$celltypes %in% c(
    "Sinusoidal ECs",
    "Arteriolar/Capillary ECs",
    "Lymphatic ECs",
    "Mixed ECs"),]
}
```

```{r scale, eval = FALSE, include = FALSE}
# scale
get_scale <- function(vis_df){
  
  vis_df$scaled <- vector(length = nrow(vis_df))

  cts <- base::unique(vis_df$celltypes)
  gs <- base::unique(vis_df$gene)
  
  for(ct in cts){
    
    vis_df$scaled[which(vis_df$celltypes == ct)] <- base::scale(
      vis_df$expression[which(vis_df$celltypes == ct)])
  }
  return(vis_df)
}

vis_df <- get_scale(vis_df)
vis_df_short <- get_scale(vis_df_short)
vis_df_ECs <- get_scale(vis_df_ECs)

```

```{r rowmax, eval = FALSE, include = FALSE}
# obtain the maximum per row (cell type) and put into an extra dataframe
# for comparison of normalized counts per cell type vs without normalisation
rowmax <- function(dataframe){
  max_vector <- vector()
  
  for(i in 1:nrow(dataframe)){
    max <- base::max(dataframe[i,])
    max_vector <- c(max_vector, max)
  }
  return(max_vector)
}

vis_df_normalized_per_celltype <- vis_df

vis_df_normalized_per_celltype <- vis_df_normalized_per_celltype/
  rowmax(vis_df_normalized_per_celltype)

vis_df_normalized_per_celltype <- tibble::rownames_to_column(
  vis_df_normalized_per_celltype, 
  "celltypes")

vis_df_normalized_per_celltype <- tidyr::pivot_longer(
  vis_df_normalized_per_celltype,
  cols = c(2:ncol(vis_df_normalized_per_celltype)), 
  values_to = "expression",
  names_to = "gene")

```

```{r plot_function}

standard_dotplot <- function(vis_df){
  
  plot_return <- ggplot(vis_df,
         aes(x = gene,
             y = celltypes,
             color = percent_expressed,
             size = expression))+
    geom_point()+
    theme_classic()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
          #legend.position = "none",
          axis.text.y = element_text(hjust = 0),
          axis.title = element_blank())+
    scale_size("logcounts", 
               limits=c(0, max(vis_df$expression)),
               range = c(0, 5))+
    scale_color_continuous(limits = c(0, 100),
                           low = "white",
                           high = "blue")
  
  return(plot_return)
  
}
```


## HSCs

```{r hsc_ggdotplot1, fig.width = 7, fig.height = 3.5}
if(fraction_curr == "hsc"){standard_dotplot(vis_df_main)}
```

```{r hsc_ggdotplot1_rev, fig.width = 7, fig.height = 3.5}
if(fraction_curr == "hsc"){
  
  ggplot(vis_df_main,
           aes(x = gene,
               y = celltypes,
               color = expression,
               size = percent_expressed))+
      geom_point()+
      theme_classic()+
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
            axis.text.y = element_text(hjust = 0),
            axis.title = element_blank())+
      scale_size("expression", 
                 limits=c(0, max(vis_df_main$percent_expressed)),
                 range = c(0, 5))+
      scale_color_continuous("logcounts",
                             limits = c(0, max(vis_df_main$expression)),
                             low = "white",
                             high = "blue")
  
}
```

## Stromal cells

### Main genes

```{r str_ggdotplot1, fig.width = 15, fig.height = 4}
# for stromal all/main
if(fraction_curr == "str"){standard_dotplot(vis_df_main)}
```

```{r str_ggdotplot1_rev, fig.width = 15, fig.height = 4}
# for stromal all/main
if(fraction_curr == "str"){
  ggplot(vis_df_main,
           aes(x = gene,
               y = celltypes,
               color = expression,
               size = percent_expressed))+
      geom_point()+
      theme_classic()+
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
            axis.text.y = element_text(hjust = 0),
            axis.title = element_blank())+
      scale_size("expression", 
                 limits=c(0, max(vis_df_main$percent_expressed)),
                 range = c(0, 5))+
      scale_color_continuous("logcounts",
                             limits = c(0, max(vis_df_main$expression)),
                             low = "white",
                             high = "blue")
}
```

### Short genes

```{r str_ggdotplot_short2, fig.width = 10, fig.height = 4}
# for shorter gene list
if(fraction_curr == "str"){standard_dotplot(vis_df_short)}
```

### ECs

```{r vectors}
c_arteries <- c(
  "Gja4",
  "Gja5",
  "Efnb2",
  "Cd34",
  "Sox17",
  "Dkk2",
  "Dll4",
  "Bmx",
  "C1qtnf9",
  "Tspan13",
  "Slc9a3r2",
  "Slc6a6",
  "Fbln2",
  "Ly6a",
  "Tm4sf1",
  "Sema3g",
  "Ass1",
  "Rbp7",
  "S100a6",
  "Nes"
)

c_arteriolar_short <- c(
 "Ly6a",     
 "Slc9a3r2",
 "Tspan13",
 "Efnb2"
)

c_capillary <- c(
  "Car4",
  "Rgcc",
  "Fibin",
  "Jag1",
  "Timp4",
  "Cd300lg",
  "Aplnr",
  "Lrg1",
  "Cldn5",
  "Slc6a6",
  "Fibin",
  "Jag1",
  "Plaur",
  "Gja1",
  "Klhl21"
)

c_cap_short <- c(
  "Car4",
  "Rgcc",
  "Timp4"
)

c_venous <- c(
  "Ackr1",
  "Vcam1",
  "Icam1",
  "Icam2",
  "Icam3",
  "Sele",
  "Selp",
  "Plvap",
  "Ephb4",
  "Emcn",
  "Nr2f2",
  "Vwf",
  "Lrg1",
  "Flt4",
  "Cotl",
  "Sox4"
  
)

c_ven_short <- c(
  "Vwf",
  "Nr2f2",
  "Selp",
  "Icam1",
  "Ackr1"
)

c_sinus <- c(
  "Stab2",
  "Tfpi",
  "Ubd",
  "Ctsl",
  "Il6st",
  "Sepp1",
  "Adamts5", 
  "Flt4", 
  "Abcc9",
  "Ackr1" 
 )

c_sin_short <- c(
  "Stab2",
  "Tfpi",
  "Ubd",
  "Flt4"
 )

c_lymph <- c(
  "Flt4",
  "Lyve1",
  "Ccl21a",
  "Pdpn",
  "Prox1"
)

c_lym_short <- c(
  "Lyve1",
  "Ccl21a",
  "Pdpn",
  "Prox1"
)

```

```{r str_ggdotplot_ECs, fig.width = 15, fig.height = 3}
# for endothelial cells
if(fraction_curr == "str"){standard_dotplot(vis_df_ECs)}
```

#### Arteriolar

```{r str_ggdotplot_art, fig.width = 7, fig.height = 3}
if(fraction_curr == "str"){
  vis_df_ECs_art <- vis_df_ECs[vis_df_ECs$gene %in% c_arteries,]
  standard_dotplot(vis_df_ECs_art)
}
```

#### Capillary

```{r str_ggdotplot_cap, fig.width = 7, fig.height = 3}
if(fraction_curr == "str"){
  vis_df_ECs_cap <- vis_df_ECs[vis_df_ECs$gene %in% c_capillary,]
  standard_dotplot(vis_df_ECs_cap)
}
```

#### Sinusoidal/Venous

```{r str_ggdotplot_ven, fig.width = 7, fig.height = 3}
if(fraction_curr == "str"){
  vis_df_ECs_ven <- vis_df_ECs[vis_df_ECs$gene %in% c_venous,]
  standard_dotplot(vis_df_ECs_ven)
}
```

```{r str_ggdotplot_sin, fig.width = 6, fig.height = 3}
if(fraction_curr == "str"){
  vis_df_ECs_ven <- vis_df_ECs[vis_df_ECs$gene %in% c_sinus,]
  standard_dotplot(vis_df_ECs_ven)
}
```

#### Lymphatic

```{r str_ggdotplot_lym, fig.width = 3, fig.height = 3}
if(fraction_curr == "str"){
  vis_df_ECs_lym <- vis_df_ECs[vis_df_ECs$gene %in% c_lymph,]
  standard_dotplot(vis_df_ECs_lym)
}
```


```{r knitexit}
knitr::knit_exit()
```

## Specific cell types for visualisation

```{r agg_again}
if(fraction_curr == "str"){
  cell_type <- "Adipo/CARs"
}else if(fraction_curr == "hsc"){
  cell_type <- "HSCs"
}

sce_temp <- sce[,sce$celltypes == cell_type]
agg <- scuttle::aggregateAcrossCells(sce_temp,
                                     id=colData(sce_temp)[,c("Species_ID")], 
                                     statistics = "mean",
                                     use.assay.type = "logcounts")
```

Separate given cell types by species.
Do this only for the short list

```{r generate_vis_df_spec}
vis_df_spec <- base::as.data.frame(
  t(SummarizedExperiment::assays(agg)$logcounts[
    rownames(SummarizedExperiment::assays(agg)$logcounts) %in% 
      gene_list_dtplt_short$gene,]))

vis_df_spec <- tibble::rownames_to_column(vis_df_spec, "species")
vis_df_spec <- tidyr::pivot_longer(vis_df_spec,
                                   cols = c(2:ncol(vis_df_spec)), 
                                   values_to = "expression", 
                                   names_to = "gene")
```

```{r perc_expressed_spec}
vis_df_spec <- get_perc_expressed(vis_df_spec)

vis_df_spec$species <- factor(vis_df_spec$species, 
                              levels = base::rev(names(col_spc)))
vis_df_spec$gene <- factor(vis_df_spec$gene,
                           levels = gene_list_dtplt_short$gene)
```

```{r hsc_ggdotplot_spec, fig.width = 5.5, fig.height = 1.5}
if(fraction_curr == "hsc"){
  standard_dotplot(vis_df_spec)+
    ggplot2::ggtitle(cell_type)
}
```

```{r hsc_ggdotplot_spec2, fig.width = 12, fig.height = 1.5}
if(fraction_curr == "hsc"){
  standard_dotplot(vis_df_spec)+
    ggplot2::ggtitle(cell_type)
}
knitr::knit_exit()
```

```{r str_ggdotplot4, fig.width = 5.5, fig.height = 1.5}
if(fraction_curr == "str"){
 
  ggplot2::ggplot(vis_df_short, 
                  aes(x = gene, 
                      y = species, 
                      color = percent_expressed, 
                      size = expression))+
    ggplot2::geom_point()+
    ggplot2::theme_classic()+
    ggplot2::theme(legend.position = "none",
                   plot.title = element_text(size = 11),
                   axis.text.x = element_blank(),
                   axis.title.x = element_blank())+
    ggplot2::scale_color_continuous(limits=c(0, 100), 
                                    low = ("white"), 
                                    high = "blue")+
    ggplot2::scale_size("logcounts", limits=c(0, 9), range = c(0,4))+
    ggplot2::ggtitle(cell_type)
}
```

```{r session_info}
utils::sessionInfo()
```