---
title: "Subclustering mclust report"
author: "Lea Wölbert"
date: '2023-07-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
#
```

```{r sources, message = FALSE}
source(file = snakemake@params[["plotting"]])
```

```{r wildcards, message = FALSE}
fraction_curr <- snakemake@wildcards[["fraction"]]
cluster_curr <- snakemake@wildcards[["cluster"]]
if(cluster_curr == 2){cluster_curr <- "2_4"}
```

```{r sce, message = FALSE}
#load SCE objects
sce_clust <- readRDS(file = snakemake@input[["sce_input"]])
celltypes <- unique(sce_clust$annotation_cluster)
```

Subset SCE

```{r print_celltypes}
print(celltypes)
```

List of genes used for subclustering or evaluation

```{r gene_list}
gene_list_sub <- read.csv(snakemake@input[["gene_list_subclustering"]], 
  header = TRUE, 
  sep = ";", 
  check.names=FALSE, 
  stringsAsFactors=FALSE, 
  as.is=TRUE, 
  colClasses = "character")
```

Subset and get only genes used for subclustering and to the correct fraction
and cluster

```{r subset_genes}

gene_list_sub <- gene_list_sub[gene_list_sub$fraction == fraction_curr,]
gene_list_sub <- gene_list_sub[gene_list_sub$cluster == cluster_curr,]

subcl <- unique(gene_list_sub$gene[gene_list_sub$purpose == "subclustering"])
print(subcl)
```

List of genes that are shared by at least three of four species in this cluster
 
```{r shared_genes}
genes_list_shared <- readRDS(file = snakemake@input[["genes_list_shared"]])
```

Subset shared gene list to the current cluster/cell type.

```{r subset_shared_genes}
if(length(celltypes) == 1){
  genes_shared <- genes_list_shared[[celltypes]]$at_least_three
}else if(length(celltypes) == 2){
  genes_shared <- intersect(genes_list_shared[[celltypes[1]]]$at_least_three, 
                            genes_list_shared[[celltypes[2]]]$at_least_three)
}
```

Print which genes used for subclustering are shared, and which were
excluded for not being shared.
Technically, there should be none of the latter if gene_list_subclustering was
configured correctly.

Also print which genes are shared/could be used for subclustering in addition.

```{r print}
print(paste("subcl genes used:", subcl[which(subcl %in% genes_shared)]))
print(paste("subcl genes not shared, not used:", 
            subcl[which(!subcl %in% genes_shared)]))

genes_all <- unique(gene_list_sub$gene)
poss_subcl <- genes_all[which(genes_all %in% genes_shared)]

print(paste("subcl genes possible:", poss_subcl))
```

# Subclustering QC

```{r qc_plots}

coord_x <- c(min(reducedDims(sce_clust)$UMAP[,1]), 
             max(reducedDims(sce_clust)$UMAP[,1]))
coord_y <- c(min(reducedDims(sce_clust)$UMAP[,2]), 
             max(reducedDims(sce_clust)$UMAP[,2]))

plot_1 <- umap_base_l(sce_clust[,sce_clust$Species_ID == "mmus"], 
                      color_by = "annotation_subcluster")+
  ggtitle("mmus")+
  xlim(coord_x)+
  ylim(coord_y)
  
plot_2 <- umap_base_l(sce_clust[,sce_clust$Species_ID == "mcas"], 
                      color_by = "annotation_subcluster")+
  ggtitle("mcas")+
  xlim(coord_x)+
  ylim(coord_y)
  
plot_3 <- umap_base_l(sce_clust[,sce_clust$Species_ID == "mspr"], 
                      color_by = "annotation_subcluster")+
  ggtitle("mspr")+
  xlim(coord_x)+
  ylim(coord_y)
    
plot_4 <- umap_base_l(sce_clust[,sce_clust$Species_ID == "mcar"], 
                      color_by = "annotation_subcluster")+
  ggtitle("mcar")+
  xlim(coord_x)+
  ylim(coord_y)

plot_5 <- umap_base_l(sce_clust, color_by = "annotation_subcluster")+
      facet_grid(cols = vars(sce_clust$annotation_subcluster),
                 rows = vars(sce_clust$Species_ID))

plot_1
plot_2
plot_3
plot_4
plot_5
knitr::knit_exit() # TODO: Find error in chunks below
```

```{r plot_function}
make_plot_list <- function(gene){
  
  print(gene)
  
  ggdf <- as.data.frame(colData(sce_clust))
  logc <- assays(sce_clust)[["logcounts"]]

  ggdf$Species_ID <- factor(ggdf$Species_ID,
                            levels = c("mmus", "mcas", "mspr", "mcar"))
  
  plot1 <- umap_gene(sce_clust, gene)

  plot2 <- umap_gene(sce_clust, gene)+
    facet_grid(cols = vars(sce_clust$Species_ID))

  plot3 <- ggplot(ggdf, aes(x = annotation_subcluster,
                             y = logc[rownames(logc) == gene,]))+
    geom_violin(fill = "grey60")+
    theme_classic()+
    geom_boxplot(color = "black", alpha = 0)+
    ylab(paste0(gene, " expression"))+
    facet_grid(cols = vars(sce_clust$Species_ID))+
    theme(axis.text.x = element_text(angle = 90))+
    ggtitle(gene)
  
  print(list(plot1, plot2, plot3))
}
```

```{r plotlist}
genes_all <- genes_all[genes_all %in% rownames(sce)]
lapply(as.list(genes_all), make_plot_list)
```

```{r session_info}
sessionInfo()
```