---
title: "Subclustering results report - UMAPs"
author: "Lea Wölbert"
date: '2023-07-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report on the results of the final annotation.
This report is used to make many plots for presentations, etc., so it is 
very chaotic and frequently adjusted.
Only for generating pretty umaps of.

Since I use a lot of plots from this report, there are many different
more or less good-looking versions of the same plot in this script.

#### Load libraries, source code 

```{r seed, message = FALSE}
set.seed(37)
```

```{r eval = FALSE, include = FALSE}
source("../../source/plotting.R")
colors_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/colors/colors.txt"
source("../../source/colors.R")

dtplt_path <- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/scRNAseq/02_sce_anno/gene_list_dotplot.txt"
dtplt_short_path<- "/omics/odcf/analysis/OE0538_projects/DO-0008/data/metadata/scRNAseq/02_sce_anno/gene_list_dotplot_short.txt"

source("../../source/colors.R")

#genelist_shared <- base::readRDS(file = snakemake@input[["genes_list_shared"]])
sce <- readRDS("/omics/odcf/analysis/OE0538_projects/DO-0008/data/scRNAseq/main_analysis/sce_objects/02_sce_anno/10_anns/sce_str-10")
umap_base_l(sce_anno, color_by = "celltypes")+
  scale_color_manual(values = col_cts_str)

```

```{r load, message = FALSE}
#library(scuttle, quietly = TRUE) 
```

```{r sources, message = FALSE}
source(file = snakemake@params[["plotting"]])
#source(file = snakemake@params[["functions"]])
```

```{r load_objects, message = FALSE}
#load SCE objects
sce <- base::readRDS(file = snakemake@input[["sce_input"]])
fraction_curr <- snakemake@wildcards[["fraction"]]
```

```{r colors, message = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

if(fraction_curr == "hsc"){
  col_cat <- col_cat_hsc
  col_cts <- col_cts_hsc
}else if(fraction_curr == "str"){
  col_cat <- col_cat_str
  col_cts <- col_cts_str
}

# factorize as required
sce$Species_ID <- factor(sce$Species_ID, levels = names(col_spc))

# mix for nicer plots
sce <- sce[,base::sample(1:ncol(sce), ncol(sce), replace = F)]
```

```{r legends}
# make legends for later use
plot_cat <- umap_legend(sce, color_by = "category")+
  ggplot2::scale_color_manual("Category", values = col_cat)
legend_cat <- ggpubr::get_legend(plot_cat)

plot_cts <- umap_legend(sce, color_by = "celltypes")+
  ggplot2::scale_color_manual("Cell types", values = col_cts)
legend_cts <- ggpubr::get_legend(plot_cts)

plot_species <- umap_legend(sce, color_by = "Species_ID")+
  ggplot2:: scale_color_manual("species", 
                              values = col_spc)+
  ggplot2::theme(legend.text= element_text(size = 11),
                 legend.title = element_text(size = 14, face = "bold"))
legend_species <- ggpubr::get_legend(plot_species)
```

# Subclustering 

## Overview

### Category

```{r overview_cat}
# overview_cat
plot <- umap_base(sce, color_by = "category")+ 
  ggplot2::ggtitle(fraction_curr)+
  ggplot2::scale_color_manual(values = col_cat)
PLOT <- ggpubr::ggarrange(plot, legend_cat, widths = c(2, 1))
PLOT
```

### Cell types

```{r overview_cts}
# overview_cts
plot <- umap_base(sce, color_by = "celltypes")+ 
  ggplot2::ggtitle(fraction_curr)+
  ggplot2::scale_color_manual(values = col_cts)  
PLOT <- ggpubr::ggarrange(plot, legend_cts, widths = c(2, 1))
PLOT
```

```{r overview_cts2}
# overview_cts2
plot <- umap_base(sce, color_by = "celltypes")+ 
  ggplot2::ggtitle(fraction_curr)+
  ggplot2::scale_color_manual(values = col_cts)+
  ggplot2::theme(axis.text = element_blank(),
                 axis.title = element_blank(),
                 axis.line = element_blank(),
                 plot.title = element_blank())
PLOT <- ggpubr::ggarrange(plot, legend_cts, widths = c(2, 1))
PLOT
```

```{r overview_cts2_size, fig.width = 4, fig.height = 4}
plot
```

### Antibodies

```{r hsc_antibodies}

if(fraction_curr == "hsc"){
  
  plot_ab <- umap_legend(sce, color_by = "Antibody_combination")+
    ggplot2::scale_color_manual("Antibody combination", 
                                values = col_ab_hsc)+
    ggplot2::theme(legend.text= element_text(size = 11),
                   legend.title = element_text(size = 14, face = "bold"))
  legend_ab <- ggpubr::get_legend(plot_ab)
  
  plot_2 <- umap_base(sce, color_by = "Antibody_combination")+ 
    ggplot2::ggtitle(fraction_curr)+
    ggplot2::scale_color_manual(values = col_ab_hsc)+
    ggplot2::theme(axis.text = element_blank(),
                   axis.title = element_blank(),
                   axis.line = element_blank(),
                   plot.title = element_blank())
  
  PLOT_2 <- ggpubr::ggarrange(plot_2, legend_ab, widths = c(2, 1))
  PLOT_2
}
```

```{r str_antibodies}
if(fraction_curr == "str"){
  plot_ab <- umap_legend(sce, color_by = "Antibody_combination")+
    scale_color_manual("Antibody combination", 
                       values = col_ab_str)+
    theme(legend.text= element_text(size = 11),
          legend.title = element_text(size = 14, face = "bold"))
  legend_ab <- get_legend(plot_ab)

  
  plot_2 <- umap_base(sce, color_by = "Antibody_combination")+ 
    ggtitle(paste(fraction_curr))+
    scale_color_manual("Antibody combination", 
                       values = col_ab_str)+
    theme(axis.text = element_blank(),
          axis.title = element_blank(),
          axis.line = element_blank(),
          plot.title = element_blank())

    PLOT_2 <- ggarrange(plot_2, legend_ab, widths = c(2, 1))
    PLOT_2
}
```

```{r antibody_plot_size, fig.width = 2, fig.height = 2}
plot_2
```

### Cell types by species 

```{r species_plots}
# species_plots
plot_2 <- umap_base(sce, color_by = "Species_ID")+ 
  ggplot2::ggtitle(paste(fraction_curr))+
  ggplot2::scale_color_manual(values = col_spc)+
  ggplot2::theme(axis.text = element_blank(),
                 axis.title = element_blank(),
                 axis.line = element_blank(),
                 plot.title = element_blank())

PLOT_2 <- ggpubr::ggarrange(plot_2, legend_species, widths = c(2, 1))
PLOT_2
```

```{r species_plots_size, fig.width = 2, fig.height = 2}
# species_plots_size
plot_2
```

```{r species_plots2}
# species_plots_size
coord_x <- c(base::min(SingleCellExperiment::reducedDims(sce)$UMAP[,1]), 
             base::max(SingleCellExperiment::reducedDims(sce)$UMAP[,1]))
coord_y <- c(base::min(SingleCellExperiment::reducedDims(sce)$UMAP[,2]), 
             base::max(SingleCellExperiment::reducedDims(sce)$UMAP[,2]))

plot_1 <- umap_base_l(sce[,sce$Species_ID == "mmus"], 
                      color_by = "celltypes")+
  ggplot2::ggtitle("mmus")+
  ggplot2::scale_color_manual("Cell types", values = col_cts)+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlim(coord_x)+
  ggplot2::ylim(coord_y)
plot_2 <- umap_base_l(sce[,sce$Species_ID == "mcas"], 
                      color_by = "celltypes")+
  ggplot2::ggtitle("mcas")+
  ggplot2::scale_color_manual("Cell types", values = col_cts)+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlim(coord_x)+
  ggplot2::ylim(coord_y)
plot_3 <- umap_base_l(sce[,sce$Species_ID == "mspr"], 
                      color_by = "celltypes")+
  ggplot2::ggtitle("mspr")+
  ggplot2::scale_color_manual("Cell types", values = col_cts)+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlim(coord_x)+
  ggplot2::ylim(coord_y)
plot_4 <- umap_base_l(sce[,sce$Species_ID == "mcar"], 
                      color_by = "celltypes")+
  ggplot2::ggtitle("mcar")+
  ggplot2::scale_color_manual("Cell types", values = col_cts)+
  ggplot2::theme(legend.position = "none")+
  ggplot2::xlim(coord_x)+
  ggplot2::ylim(coord_y)

PLOT_1 <- ggpubr::ggarrange(plot_1, legend_cts, widths = c(2, 1))
PLOT_2 <- ggpubr::ggarrange(plot_2, legend_cts, widths = c(2, 1))
PLOT_3 <- ggpubr::ggarrange(plot_3, legend_cts, widths = c(2, 1))
PLOT_4 <- ggpubr::ggarrange(plot_4, legend_cts, widths = c(2, 1))

PLOT_1
PLOT_2
PLOT_3
PLOT_4
```

```{r species_separation2}
#species_separation2
print(base::table(sce$Species_ID, sce$celltypes))
```

```{r species_separation3}
#species_separation3
print(base::table(sce$Species_ID,
                  sce$celltypes)/
        BiocGenerics::rowSums(base::table(sce$Species_ID, 
                                          sce$celltypes))*100)
```

```{r species_separation4, fig.width = 7, fig.height = 5}
#species_separation4
plot_a <- umap_base(sce, color_by = "celltypes")+ 
  ggplot2::ggtitle(fraction_curr)+
  ggplot2::scale_color_manual(values = col_cts)+
  ggplot2::theme(axis.text = element_blank(),
                 axis.title = element_blank(),
                 axis.line = element_blank(),
                 plot.title = element_blank())+
  ggplot2::facet_wrap( ~ sce$Species_ID, ncol = 2)+
  ggplot2::theme(strip.text = element_text(face = "bold", size = 14),
                 strip.background = element_rect(linetype = 0))

plot_a

PLOT_all <- ggpubr::ggarrange(plot_a, legend_cts, widths = c(2.5, 1))
PLOT_all
```

```{r species_separation5, fig.width = 8, fig.height = 5}
#species_separation5
PLOT_all <- ggpubr::ggarrange(plot_a, legend_cts, widths = c(2.5, 1.5))
PLOT_all
```

```{r sessioninfo}
utils::sessionInfo()
```
