---
title: "Batch correction report"
author: "Lea Wölbert"
date: '2022-12-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report for batch correction at fraction level.
This is a very long report with lots of plots.

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r load, message = FALSE}
library(scuttle, quietly = TRUE) 
```

```{r source, message = FALSE}
source(file =  snakemake@params[["functions"]])
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}

batch_use <- snakemake@params[["batch_use"]]
nr_hvgs <- snakemake@params[["nr_hvgs"]]

sce_input_raw <- base::readRDS(file = snakemake@input[["sce_input_raw"]])
sce_mnn <- base::readRDS(file = snakemake@input[["sce_input_corrected"]])

sce_list <- list(list(sce_input_raw, " after merging"),
                 list(sce_mnn, " MNN batch correction"))

sce_list <- lapply(sce_list, function(x){
  x[[1]] <- factor_reference_cts(x[[1]])
  print(x[[1]])
  return(x)
})

# randomize counts for nicer plots
sce_list <- lapply(sce_list, function(x){
  x[[1]] <- x[[1]][,base::sample(c(1:ncol(x[[1]])), size = ncol(x[[1]]))]
  return(x)
})
```

```{r colors}
colors_ref_path <- snakemake@params[["colors_ref_path"]]
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

fraction_curr <- sce_list[[1]][[1]]$Fraction_ID[1]

if(fraction_curr == "hsc"){
  col_cts_ref["unassigned"] <- "honeydew2"
}else if(fraction_curr == "str"){
  col_cts_ref["unassigned"] <- "mistyrose3"
}

col_cts_baccin <- col_cts_ref[base::match(
  levels(sce_list[[1]][[1]]$baccin_celltype_scmapclust), names(col_cts_ref))]
col_cts_dahlin <- col_cts_ref[base::match(
  levels(sce_list[[1]][[1]]$dahlin_celltype_scmapclust), names(col_cts_ref))]
col_cts_dolgalev <- col_cts_ref[base::match(
  levels(sce_list[[1]][[1]]$dolgalev_celltype_scmapclust), names(col_cts_ref))]

col_batch_exp_day <- col_alp[names(col_alp) %in%
                               sce_list[[1]][[1]]$Batch_exp_day]
col_batch_seq <- col_num[names(col_num) %in% 
                           sce_list[[1]][[1]]$Batch_sequencing]

col_age <- col_age[names(col_age) %in% sce_list[[1]][[1]]$Age_ID]
col_frc <- col_frc[names(col_frc) %in% sce_list[[1]][[1]]$Fraction_ID]
col_spc <- col_spc[names(col_spc) %in% sce_list[[1]][[1]]$Species_ID]
```

# Reference cell types 

Cells were annotated with scmapclust and three reference datasets.
Own functions.

```{r plot_function}

identity_plots <- function(sce, title){
  
  title_full <- base::paste0(fraction_curr, title)
  
  plot_1 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_baccin)
  plot_1l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual("baccin_celltype_scmapclust",
                                values = col_cts_baccin)
  legend_1 <- ggpubr::get_legend(plot_1l)

  plot_2 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_dahlin)
  plot_2l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual("dahlin_celltype_scmapclust",
                                values = col_cts_dahlin)
  legend_2 <- ggpubr::get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)
  plot_3l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual("dolgalev_celltype_scmapclust",
                                values = col_cts_dolgalev)
  legend_3 <- ggpubr::get_legend(plot_3l)
  
  PLOT_1 <- ggpubr::ggarrange(plot_1, legend_1)
  PLOT_2 <- ggpubr::ggarrange(plot_2, legend_2)
  PLOT_3 <- ggpubr::ggarrange(plot_3, legend_3)

  return(list(PLOT_1, PLOT_2, PLOT_3))
}
```

```{r print_basic}
print(fraction_curr)
print(base::table(sce_input_raw$baccin_celltype_scmapclust))
print(base::table(sce_input_raw$dahlin_celltype_scmapclust))
print(base::table(sce_input_raw$dolgalev_celltype_scmapclust))
```

### After merging

```{r after_merging_plots1}
identity_plotlist <- lapply(sce_list, function(x){
  plotlist <- identity_plots(x[[1]], title = x[[2]])
  return(plotlist)
})
```

```{r after_merging_plots2, fig.width=8, fig.height=4}
identity_plotlist
```

# All QC metrics

```{r qc_plots_function}

qc_plots <- function(sce, title){
  
  title_full <- base::paste0(fraction_curr, title)

  plot_1 <- umap_base(sce, color_by = "Object_ID")+
    ggplot2::ggtitle(title_full)
  plot_1l <- umap_legend(sce, color_by = "Object_ID")+
    ggplot2::scale_color_discrete("Object_ID")
  legend_1 <- ggpubr::get_legend(plot_1l)
  
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual(values = col_batch_seq)+
    ggplot2::ggtitle(title_full)
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual("Batch_sequencing", values = col_batch_seq)
  legend_2 <- ggpubr::get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+ 
    ggplot2::scale_color_manual(values =  col_batch_exp_day)+
    ggplot2::ggtitle(title_full)
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual("Batch_experimental_day", 
                                values = col_batch_exp_day)
  legend_3 <- ggpubr::get_legend(plot_3l)
  
  plot_4 <- umap_base(sce, color_by = "Date_collected")+ 
    ggplot2::ggtitle(title_full)
  plot_4l <- umap_legend(sce, color_by = "Date_collected")+
    ggplot2::scale_color_discrete("Date of collection")
  legend_4 <- ggpubr::get_legend(plot_4l)
  
  # check everything else
  plot_5 <- umap_base(sce, color_by = "Species_ID")+ 
    ggplot2::scale_color_manual(values = col_spc)+
    ggplot2::ggtitle(title_full)
  plot_5l <- umap_legend(sce, color_by = "Species_ID")+
    ggplot2::scale_color_manual("Species", values = col_spc)
  legend_5 <- ggpubr::get_legend(plot_5l)
  
  plot_6 <- umap_base(sce, color_by = "Age_ID")+ 
    ggplot2::scale_color_manual(values = col_age)+
    ggplot2::ggtitle(title_full)
  plot_6l <- umap_legend(sce, color_by = "Age_ID")+
    ggplot2::scale_color_manual("Age", values = col_age)
  legend_6 <- ggpubr::get_legend(plot_6l) 
  
  plot_7 <- umap_base(sce, color_by = "Fraction_ID")+ 
    ggplot2::scale_color_manual(values = col_frc)+
    ggplot2::ggtitle(title_full)
  plot_7l <- umap_legend(sce, color_by = "Fraction_ID")+
    ggplot2::scale_color_manual("Fraction", values = col_frc)
  legend_7 <- ggpubr::get_legend(plot_7l) 
 
  plot_8 <- umap_base(sce, color_by = "Antibody_combination")+ 
    ggplot2::ggtitle(title_full)
  plot_8l <- umap_legend(sce, color_by = "Antibody_combination")+
    ggplot2::scale_color_discrete("Antibody combination")
  legend_8 <- ggpubr::get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "doublet_score")+ 
    ggplot2::ggtitle(title_full)
  plot_9l <- umap_legend(sce, color_by = "doublet_score")+
    ggplot2::scale_color_continuous("Doublet score")
  legend_9 <- ggpubr::get_legend(plot_9l)
  
  plot_10 <- umap_base(sce, color_by = "Keep_sample")+ 
    ggplot2::ggtitle(title_full)
  plot_10l <- umap_legend(sce, color_by = "Keep_sample")+
    ggplot2::scale_color_discrete("Keep_sample")
  legend_10 <- ggpubr::get_legend(plot_10l)
  
  qcdf <- scuttle::perCellQCMetrics(sce)
  sce$sum <- qcdf$sum
  sce$detected <- qcdf$detected

  plot_11 <- umap_base(sce, color_by = "sum")+
    ggplot2::scale_color_gradientn("Library size", 
                                   limits = c(0, 100000),
                                   colors = color_vector_sum)+
    ggplot2::ggtitle(title_full)
  plot_11l <- umap_legend(sce, color_by = "sum")+
    ggplot2::scale_color_gradientn("Library size", 
                                   limits = c(0, 1000000),
                                   colors = color_vector_sum)
  legend_11 <- ggpubr::get_legend(plot_11l)

  plot_12 <- umap_base(sce, color_by = "detected")+
    ggplot2::scale_color_gradientn("Number of detected genes", 
                                   limits = c(0, 10000),
                                   colors = color_vector_det)+
    ggplot2::ggtitle(base::paste0(fraction_curr, title))
  plot_12l <- umap_legend(sce, color_by = "detected")+
    ggplot2::scale_color_gradientn("Number of detected genes", 
                                   limits = c(0, 10000),
                                   colors = color_vector_det)
  legend_12 <- ggpubr::get_legend(plot_12l)
  
  PLOT_12 <- ggpubr::ggarrange(plot_12, legend_12)

  PLOT_1 <- ggpubr::ggarrange(plot_1, legend_1)
  PLOT_2 <- ggpubr::ggarrange(plot_2, legend_2)
  PLOT_3 <- ggpubr::ggarrange(plot_3, legend_3)
  PLOT_4 <- ggpubr::ggarrange(plot_4, legend_4)
  PLOT_5 <- ggpubr::ggarrange(plot_5, legend_5)
  PLOT_6 <- ggpubr::ggarrange(plot_6, legend_6)
  PLOT_7 <- ggpubr::ggarrange(plot_7, legend_7)
  PLOT_8 <- ggpubr::ggarrange(plot_8, legend_8)
  PLOT_9 <- ggpubr::ggarrange(plot_9, legend_9)
  PLOT_10 <- ggpubr::ggarrange(plot_10, legend_10)
  PLOT_11 <- ggpubr::ggarrange(plot_11, legend_11)
  PLOT_12 <- ggpubr::ggarrange(plot_12, legend_12)

  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10, PLOT_11, PLOT_12)
  return(plot_list)
  
}

qc_plotlist <- lapply(sce_list, function(x){
  plotlist <- qc_plots(x[[1]], title = x[[2]])
  return(plotlist)
})
```

```{r qc_plots, fig.width=10, fig.height=5}
qc_plotlist
```

## Facet grid UMAPs

To merge across ages and fractions using MNNcorrect, every age and fraction 
should have at least one overlapping BATCH.

Also, batches should have overlapping cell types for use of MNNCorrect or
Seurat.

```{r facet_plot_function}

facet_plots <- function(sce, title){
  
  title_full <- base::paste0(fraction_curr, title)
  
  # VARS SPECIES
  plot_1 <- umap_base(sce, color_by = "Object_ID")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_1l <- umap_legend(sce, color_by = "Object_ID")+
    ggplot2::scale_color_discrete("Object_ID")
  legend_1 <- ggpubr::get_legend(plot_1l)
  
  plot_2 <- umap_base(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual(values = col_batch_seq)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_2l <- umap_legend(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual("Batch_sequencing", values = col_batch_seq)
  legend_2 <- ggpubr::get_legend(plot_2l)
  
  plot_3 <- umap_base(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual(values = col_batch_exp_day)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_3l <- umap_legend(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual("Batch_experimental_day", 
                                values = col_batch_exp_day)
  legend_3 <- ggpubr::get_legend(plot_3l)
  
  plot_19 <- umap_base(sce, color_by = "Age_ID")+
    ggplot2::scale_color_manual(values = col_age)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_19l <- umap_legend(sce, color_by = "Age_ID")+
    ggplot2::scale_color_manual("Age", values = col_age)
  legend_19 <- ggpubr::get_legend(plot_19l)
  
  plot_21 <- umap_base(sce, color_by = "Antibody_combination")+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_21l <- umap_legend(sce, color_by = "Antibody_combination")+
    ggplot2::scale_color_discrete(name = "Antibody combination")
  legend_21 <- ggpubr::get_legend(plot_21l)

  # AGE
  plot_4 <- umap_base(sce, color_by = "Object_ID")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  plot_4l <- umap_legend(sce, color_by = "Object_ID")+
    ggplot2::scale_color_discrete("Object_ID")
  legend_4 <- ggpubr::get_legend(plot_4l)
  
  plot_5 <- umap_base(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual(values = col_batch_seq)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  plot_5l <- umap_legend(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual("Batch_sequencing", values = col_batch_seq)
  legend_5 <- ggpubr::get_legend(plot_5l)
  
  plot_6 <- umap_base(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual(values = col_batch_exp_day)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  plot_6l <- umap_legend(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual("Batch_experimental_day",
                                values = col_batch_exp_day)
  legend_6 <- ggpubr::get_legend(plot_6l)
  
  plot_20 <- umap_base(sce, color_by = "Species_ID")+
    ggplot2::scale_color_manual(values = col_spc)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  plot_20l <- umap_legend(sce, color_by = "Species_ID")+
    ggplot2::scale_color_manual("Species_ID", values = col_spc)
  legend_20 <- ggpubr::get_legend(plot_20l)
  
  # Identities
  plot_7 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_baccin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_7l <- umap_legend(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual("baccin_celltype_scmapclust", 
                                values = col_cts_baccin)
  legend_baccin <- ggpubr::get_legend(plot_7l)
  
  plot_8 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dahlin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_8l <- umap_legend(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual("dahlin_celltype_scmapclust",
                                values = col_cts_dahlin)
  legend_dahlin <- ggpubr::get_legend(plot_8l)
  
  plot_9 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Species_ID))
  plot_9l <- umap_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual("dolgalev_celltype_scmapclust",
                       values = col_cts_dolgalev)
  legend_dolgalev <- ggpubr::get_legend(plot_9l)
  
  
  plot_10 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_baccin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  
  plot_11 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dahlin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  
  plot_12 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Age))
  
  
  plot_13 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_baccin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_exp_day))
  
  plot_14 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dahlin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_exp_day))
  
  plot_15 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_exp_day))
  
  
  plot_16 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_baccin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_17 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dahlin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_sequencing))
  
  plot_18 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Batch_sequencing)) 
  
  plot_22 <- umap_base(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_baccin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Antibody_combination))
  
  plot_23 <- umap_base(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dahlin)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Antibody_combination))
  
  plot_24 <- umap_base(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)+
    ggplot2::ggtitle(title_full)+
    ggplot2::facet_grid(cols = vars(sce$Antibody_combination)) 

  PLOT_1 <- ggpubr::ggarrange(plot_1, legend_1, widths = c(3, 1))
  PLOT_2 <- ggpubr::ggarrange(plot_2, legend_2, widths = c(3, 1))
  PLOT_3 <- ggpubr::ggarrange(plot_3, legend_3, widths = c(3, 1))
  PLOT_4 <- ggpubr::ggarrange(plot_4, legend_4, widths = c(3, 1))
  PLOT_5 <- ggpubr::ggarrange(plot_5, legend_5, widths = c(3, 1))
  PLOT_6 <- ggpubr::ggarrange(plot_6, legend_6, widths = c(3, 1))
  PLOT_7 <- ggpubr::ggarrange(plot_7, legend_baccin, widths = c(3, 1))
  PLOT_8 <- ggpubr::ggarrange(plot_8, legend_dahlin, widths = c(3, 1))
  PLOT_9 <- ggpubr::ggarrange(plot_9, legend_dolgalev, widths = c(3, 1))
  PLOT_10 <- ggpubr::ggarrange(plot_10, legend_baccin, widths = c(3, 1))
  PLOT_11 <- ggpubr::ggarrange(plot_11, legend_dahlin, widths = c(3, 1))
  PLOT_12 <- ggpubr::ggarrange(plot_12, legend_dolgalev, widths = c(3, 1))
  PLOT_13 <- ggpubr::ggarrange(plot_13, legend_baccin, widths = c(3, 1))
  PLOT_14 <- ggpubr::ggarrange(plot_14, legend_dahlin, widths = c(3, 1))
  PLOT_15 <- ggpubr::ggarrange(plot_13, legend_dolgalev, widths = c(3, 1))
  PLOT_16 <- ggpubr::ggarrange(plot_14, legend_baccin, widths = c(3, 1))
  PLOT_17 <- ggpubr::ggarrange(plot_13, legend_dahlin, widths = c(3, 1))
  PLOT_18 <- ggpubr::ggarrange(plot_14, legend_dolgalev, widths = c(3, 1))  
  PLOT_19 <- ggpubr::ggarrange(plot_19, legend_19, widths = c(3, 1))  
  PLOT_20 <- ggpubr::ggarrange(plot_20, legend_20, widths = c(3, 1))  
  PLOT_21 <- ggpubr::ggarrange(plot_21, legend_21, widths = c(3, 1))  
  PLOT_22 <- ggpubr::ggarrange(plot_22, legend_baccin, widths = c(3, 1))  
  PLOT_23 <- ggpubr::ggarrange(plot_23, legend_dahlin, widths = c(3, 1))  
  PLOT_24 <- ggpubr::ggarrange(plot_24, legend_dolgalev, widths = c(3, 1))  
  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_19, # Species
                    PLOT_7, PLOT_8, PLOT_9, PLOT_21, # Species
                    
                    PLOT_4, PLOT_5, PLOT_6, PLOT_20, # Age
                    PLOT_10, PLOT_11, PLOT_12, # Age
                    
                    PLOT_13, PLOT_14, PLOT_15, # batchexpday
                    PLOT_16, PLOT_17, PLOT_18, # batchseq
                    PLOT_22, PLOT_23, PLOT_24) # antibody combination
  
  return(plot_list)
}

facet_plotlist <- lapply(sce_list, function(x){
  plotlist <- facet_plots(x[[1]], title = x[[2]])
  return(plotlist)
})
```

```{r facet_plots, fig.width=12, fig.height=4}
facet_plotlist
```

## PCAs

```{r pca_function_pc2}

make_qc_pcas2 <- function(sce, title){
  
  title_full <- base::paste0(fraction_curr, title)

  plot_a <- pca_base2(sce, color_by = "baccin_celltype_scmapclust")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_baccin)
  plot_al <- pca_legend(sce, color_by = "baccin_celltype_scmapclust")+
    ggplot2::scale_color_manual("baccin_celltype_scmapclust", 
                                values = col_cts_baccin)
  legend_a <- ggpubr::get_legend(plot_al)

  plot_b <- pca_base2(sce, color_by = "dahlin_celltype_scmapclust")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_dahlin)
  plot_bl <- pca_legend(sce, color_by = "dahlin_celltype_scmapclust")+
    ggplot2::scale_color_manual("dahlin_celltype_scmapclust", 
                                values = col_cts_dahlin)
  legend_b <- ggpubr::get_legend(plot_bl)
  
  plot_c <- pca_base2(sce, color_by = "dolgalev_celltype_scmapclust")+ 
    ggplot2::ggtitle(title_full)+
    ggplot2::scale_color_manual(values = col_cts_dolgalev)
  plot_cl <- pca_legend(sce, color_by = "dolgalev_celltype_scmapclust")+
    ggplot2::scale_color_manual("dolgalev_celltype_scmapclust", 
                                values = col_cts_dolgalev)
  legend_c <- ggpubr::get_legend(plot_cl)
  
  
  plot_1 <- pca_base2(sce, color_by = "Object_ID")+ 
    ggplot2::ggtitle(title_full)
  plot_1l <- pca_legend(sce, color_by = "Object_ID")+
    ggplot2::scale_color_discrete("Object_ID")
  legend_1 <- ggpubr::get_legend(plot_1l)
  
  plot_2 <- pca_base2(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual(values = col_batch_seq)+
    ggplot2::ggtitle(title_full)
  plot_2l <- pca_legend(sce, color_by = "Batch_sequencing")+
    ggplot2::scale_color_manual("Batch_sequencing", values = col_batch_seq)
  legend_2 <- ggpubr::get_legend(plot_2l)
  
  plot_3 <- pca_base2(sce, color_by = "Batch_exp_day")+ 
    ggplot2::scale_color_manual(values = col_batch_exp_day)+
    ggplot2::ggtitle(title_full)
  plot_3l <- pca_legend(sce, color_by = "Batch_exp_day")+
    ggplot2::scale_color_manual("Batch_experimental_day", 
                                values = col_batch_exp_day)
  legend_3 <- ggpubr::get_legend(plot_3l)
  
  plot_4 <- pca_base2(sce, color_by = "Date_collected")+ 
    ggplot2::ggtitle(title_full)
  plot_4l <- pca_legend(sce, color_by = "Date_collected")+
    ggplot2::scale_color_discrete("Date of collection")
  legend_4 <- ggpubr::get_legend(plot_4l)
  
  # check everything else
  plot_5 <- pca_base2(sce, color_by = "Species_ID")+ 
    ggplot2::scale_color_manual(values = col_spc)+
    ggplot2::ggtitle(title_full)
  plot_5l <- pca_legend(sce, color_by = "Species_ID")+
    ggplot2::scale_color_manual("Species", values = col_spc)
  legend_5 <- ggpubr::get_legend(plot_5l)
  
  plot_6 <- pca_base2(sce, color_by = "Age_ID")+ 
    ggplot2::scale_color_manual(values = col_age)+
    ggplot2::ggtitle(title_full)
  plot_6l <- pca_legend(sce, color_by = "Age_ID")+
    ggplot2::scale_color_manual("Age", values = col_age)
  legend_6 <- ggpubr::get_legend(plot_6l) 
  
  plot_7 <- pca_base2(sce, color_by = "Fraction_ID")+ 
    ggplot2::scale_color_manual(values = col_frc)+
    ggplot2::ggtitle(title_full)
  plot_7l <- pca_legend(sce, color_by = "Fraction_ID")+
    ggplot2::scale_color_manual("Fraction", values = col_frc)
  legend_7 <- ggpubr::get_legend(plot_7l) 
 
  plot_8 <- pca_base2(sce, color_by = "Antibody_combination")+ 
    ggplot2::ggtitle(title_full)
  plot_8l <- pca_legend(sce, color_by = "Antibody_combination")+
    ggplot2::scale_color_discrete("Antibody combination")
  legend_8 <- ggpubr::get_legend(plot_8l)
  
  plot_9 <- pca_base2(sce, color_by = "doublet_score")+ 
    ggplot2::ggtitle(title_full)
  plot_9l <- pca_legend(sce, color_by = "doublet_score")+
    ggplot2::scale_color_continuous("Doublet score")
  legend_9 <- ggpubr::get_legend(plot_9l)
  
  plot_10 <- pca_base2(sce, color_by = "Keep_sample")+ 
    ggplot2::ggtitle(title_full)
  plot_10l <- pca_legend(sce, color_by = "Keep_sample")+
    ggplot2::scale_color_discrete("Keep_sample")
  legend_10 <- ggpubr::get_legend(plot_10l)
  
  PLOT_A <- ggpubr::ggarrange(plot_a, legend_a)
  PLOT_B <- ggpubr::ggarrange(plot_b, legend_b)
  PLOT_C <- ggpubr::ggarrange(plot_c, legend_c)
  PLOT_1 <- ggpubr::ggarrange(plot_1, legend_1)
  PLOT_2 <- ggpubr::ggarrange(plot_2, legend_2)
  PLOT_3 <- ggpubr::ggarrange(plot_3, legend_3)
  PLOT_4 <- ggpubr::ggarrange(plot_4, legend_4)
  PLOT_5 <- ggpubr::ggarrange(plot_5, legend_5)
  PLOT_6 <- ggpubr::ggarrange(plot_6, legend_6)
  PLOT_7 <- ggpubr::ggarrange(plot_7, legend_7)
  PLOT_8 <- ggpubr::ggarrange(plot_8, legend_8)
  PLOT_9 <- ggpubr::ggarrange(plot_9, legend_9)
  PLOT_10 <- ggpubr::ggarrange(plot_10, legend_10)

  plot_list <- list(PLOT_1, PLOT_2, PLOT_3, PLOT_4, PLOT_5, PLOT_6, PLOT_7,
                    PLOT_8, PLOT_9, PLOT_10, PLOT_A, PLOT_B, PLOT_C)
  return(plot_list)
  
}

pca2_plotlist <- lapply(sce_list, function(x){
  plotlist <- make_qc_pcas2(x[[1]], title = x[[2]])
  return(plotlist)
})
```

```{r pcas_pc2, fig.width=10, fig.height=5}
pca2_plotlist
```

# Cell type proportions

To get an idea of cell numbers (according to refernce dataset).
Also helps understand balance of cell types in batches.

Using only one object is enough for this.

```{r subset_sce}
sce <- sce_list[[1]][[1]]
```

## Per Sample

```{r ct_props, fig.width=10, fig.height=12}

ggdf <- base::data.frame(colData(sce))

# BACCIN
plot1 <- ggplot2::ggplot(ggdf,
                         aes(x = Object_ID,
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot_baccin_leg <- ggplot2::ggplot(ggdf, 
                                   aes(x = Object_ID, 
                                       y=baccin_celltype_scmapclust,
                                       fill=baccin_celltype_scmapclust))+
  ggplot2::geom_col()+
  ggplot2::scale_fill_manual("baccin_celltype_scmapclust", 
                             values = col_cts_baccin)
legend_baccin <- ggpubr::get_legend(plot_baccin_leg)

plot2 <- ggplot2::ggplot(ggdf, 
                         aes(x = Object_ID, 
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DAHLIN
plot3 <- ggplot2::ggplot(ggdf, 
                         aes(x = Object_ID, 
                             fill=dahlin_celltype_scmapclust,
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot_dahlin_leg <- ggplot2::ggplot(ggdf, 
                                   aes(x = Object_ID, 
                                       y=dahlin_celltype_scmapclust,
                                       fill=dahlin_celltype_scmapclust))+
  ggplot2::geom_col()+
  ggplot2::scale_fill_manual("dahlin_celltype_scmapclust", 
                             values = col_cts_dahlin)
legend_dahlin <- ggpubr::get_legend(plot_dahlin_leg)
                    
plot4 <- ggplot2::ggplot(ggdf, 
                         aes(x = Object_ID, 
                             fill=dahlin_celltype_scmapclust, 
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DOLGALEV
plot5 <- ggplot2::ggplot(ggdf, 
                         aes(x = Object_ID, 
                             fill=dolgalev_celltype_scmapclust,
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot_dolgalev_leg <- ggplot2::ggplot(ggdf, 
                                     aes(x = Object_ID, 
                                         y=dolgalev_celltype_scmapclust,
                                         fill=dolgalev_celltype_scmapclust))+
  ggplot2::geom_col()+
  ggplot2::scale_fill_manual("dolgalev_celltype_scmapclust", 
                             values = col_cts_dolgalev)
legend_dolgalev <- ggpubr::get_legend(plot_dolgalev_leg)
                    
plot6 <- ggplot2::ggplot(ggdf, 
                         aes(x = Object_ID, 
                             fill=dolgalev_celltype_scmapclust, 
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

PLOT1 <- ggpubr::ggarrange(plot1, legend_baccin, ncol = 1)
PLOT2 <- ggpubr::ggarrange(plot2, legend_baccin, ncol = 1)
PLOT3 <- ggpubr::ggarrange(plot3, legend_dahlin, ncol = 1)
PLOT4 <- ggpubr::ggarrange(plot4, legend_dahlin, ncol = 1)
PLOT5 <- ggpubr::ggarrange(plot5, legend_dolgalev, ncol = 1)
PLOT6 <- ggpubr::ggarrange(plot6, legend_dolgalev, ncol = 1)

PLOT1
PLOT2
PLOT3
PLOT4
PLOT5
PLOT6
```

```{r plotx}

plot_x <- ggplot2::ggplot(ggdf, 
                          aes(x = Object_ID, 
                              fill=baccin_celltype_scmapclust, 
                              color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual("baccin_celltype_scmapclust", 
                             values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)+
  ggplot2::facet_grid(cols = vars(ggdf$Antibody_combination))

plot_x
```

## Per Experimental Day

```{r per_experimental_day}

# BACCIN
plot1 <- ggplot2::ggplot(ggdf, 
                         aes(x = Batch_exp_day,
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot2 <- ggplot2::ggplot(ggdf, 
                         aes(x = Batch_exp_day, 
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DAHLIN
plot3 <- ggplot2::ggplot(ggdf, 
                         aes(x = Batch_exp_day, 
                             fill=dahlin_celltype_scmapclust,
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot4 <- ggplot2::ggplot(ggdf, 
                         aes(x = Batch_exp_day, 
                             fill=dahlin_celltype_scmapclust, 
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DOLGALEV
plot5 <- ggplot2::ggplot(ggdf, 
                         aes(x = Batch_exp_day, 
                             fill=dolgalev_celltype_scmapclust,
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot6 <- ggplot2::ggplot(ggdf,
                         aes(x = Batch_exp_day, 
                             fill=dolgalev_celltype_scmapclust, 
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

PLOT1 <- ggpubr::ggarrange(plot1, legend_baccin)
PLOT2 <- ggpubr::ggarrange(plot2, legend_baccin)
PLOT3 <- ggpubr::ggarrange(plot3, legend_dahlin)
PLOT4 <- ggpubr::ggarrange(plot4, legend_dahlin)
PLOT5 <- ggpubr::ggarrange(plot5, legend_dolgalev)
PLOT6 <- ggpubr::ggarrange(plot6, legend_dolgalev)

PLOT1
PLOT2
PLOT3
PLOT4
PLOT5
PLOT6
```

## Per species

```{r per_species}

# BACCIN
plot1 <- ggplot2::ggplot(ggdf, 
                         aes(x = Species_ID,
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot2 <- ggplot2::ggplot(ggdf,
                         aes(x = Species_ID, 
                             fill=baccin_celltype_scmapclust, 
                             color = baccin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_baccin)+
  ggplot2::scale_color_manual(values = col_cts_baccin)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DAHLIN
plot3 <- ggplot2::ggplot(ggdf,
                         aes(x = Species_ID, 
                             fill=dahlin_celltype_scmapclust,
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack")+ 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


plot4 <- ggplot2::ggplot(ggdf, 
                         aes(x = Species_ID, 
                             fill=dahlin_celltype_scmapclust, 
                             color = dahlin_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dahlin)+
  ggplot2::scale_color_manual(values = col_cts_dahlin)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)


# DOLGALEV
plot5 <- ggplot2::ggplot(ggdf,
                         aes(x = Species_ID, 
                             fill=dolgalev_celltype_scmapclust,
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "stack") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none",
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

plot6 <- ggplot2::ggplot(ggdf, 
                         aes(x = Species_ID, 
                             fill=dolgalev_celltype_scmapclust, 
                             color = dolgalev_celltype_scmapclust))+
  ggplot2::geom_bar(position = "fill") + 
  ggplot2::theme_classic()+
  ggplot2::scale_fill_manual(values = col_cts_dolgalev)+
  ggplot2::scale_color_manual(values = col_cts_dolgalev)+
  ggplot2::theme(legend.position = "none", 
                 axis.text.x = element_text(angle = 90))+
  ggplot2::ggtitle(fraction_curr)

PLOT1 <- ggpubr::ggarrange(plot1, legend_baccin)
PLOT2 <- ggpubr::ggarrange(plot2, legend_baccin)
PLOT3 <- ggpubr::ggarrange(plot3, legend_dahlin)
PLOT4 <- ggpubr::ggarrange(plot4, legend_dahlin)
PLOT5 <- ggpubr::ggarrange(plot5, legend_dolgalev)
PLOT6 <- ggpubr::ggarrange(plot6, legend_dolgalev)

PLOT1
PLOT2
PLOT3
PLOT4
PLOT5
PLOT6

print(length(which(sce_mnn$Species_ID == "mmus")))
print(length(which(sce_mnn$Species_ID == "mcar")))
print(length(which(sce_mnn$Species_ID == "mcas")))
print(length(which(sce_mnn$Species_ID == "mspr")))
```

```{r session_info}
utils::sessionInfo()
```