---
title: "Clustering comparison report"
author: "Lea Wölbert"
date: '2024-07-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report for clustering of all species separately, to compare to the original 
clustering obtained from all species merged.
To make sure that the merged clustering is not too different from species-
specific clustering.

#### Load libraries, source code 

```{r seed,  message = FALSE}
set.seed(37)
```

```{r, message = FALSE}
library(ggalluvial, quietly = TRUE)
library(SummarizedExperiment, quietly = TRUE)
library(SingleCellExperiment, quietly = TRUE)
library(BiocSingular, quietly = TRUE)
```

```{r source, message = FALSE}
source(file =  snakemake@params[["plotting"]])
```

```{r load_objects}
sce_l <- base::readRDS(file = snakemake@input[["sce_l"]])
sce_ls <- base::readRDS(file = snakemake@input[["sce_l_species"]])
sce_ct <- base::readRDS(file = snakemake@input[["sce_l_celltypes"]])

fraction_curr <- snakemake@wildcards[["fraction"]]
species_curr <- snakemake@wildcards[["species"]]

individual_curr <- base::paste0(species_curr, "_", fraction_curr)
```

```{r colors, meassage = FALSE}
colors_path <- snakemake@params[["colors_path"]]
source(snakemake@params[["colors"]])

# randomize counts for nicer plots
sce_l <- sce_l[,base::sample(c(1:ncol(sce_l)), size = ncol(sce_l))]
sce_ls <- sce_ls[,base::sample(c(1:ncol(sce_ls)), size = ncol(sce_ls))]
sce_ct <- sce_ct[,base::sample(c(1:ncol(sce_ct)), size = ncol(sce_ct))]

# get nicer looking cluster color vectors
col_clust_l <- col_num[names(col_num) %in% sce_l$cluster_louvain]
```

# Comparison

## All clusters

For all clusters (before removal and annotation).

```{r prepare_ggdf_cluster}
# subset the merged object to the species-specific cells only
# this will keep original UMAP coordinates 
sce_gg_cl <- sce_l[,which(colnames(sce_l) %in% colnames(sce_ls))]

sce_gg_cl$species_cluster_louvain <- vector(length = ncol(sce_gg_cl))
# add info on species-specific clustering in species_cluster_louvain
sce_gg_cl$species_cluster_louvain <- sce_ls$cluster_louvain[base::match(
  colnames(sce_gg_cl),
  colnames(sce_ls))]

ggdf_cl <- base::as.data.frame(SummarizedExperiment::colData(sce_gg_cl))
ggdf_cl$species_cluster_louvain <- base::paste0(ggdf_cl$species_cluster_louvain, 
                                                "_species")
```

```{r umap_clus_legend}
# make legend
umap_l <- umap_legend(sce_l, color_by = "cluster_louvain")+
  ggplot2::scale_color_manual("original cluster", values = col_clust_l)
legend <- ggpubr::get_legend(umap_l)
umap_l_species <- umap_legend(sce_gg_cl, color_by = "species_cluster_louvain")+
  ggplot2::scale_color_discrete("cluster_species")
legend_species <- ggpubr::get_legend(umap_l_species)
umap_l_species_full <- umap_legend(sce_ls, 
                                   color_by = "cluster_louvain")+
  ggplot2::scale_color_discrete("cluster_species")
legend_species_full <- ggpubr::get_legend(umap_l_species_full)
```

```{r umap_clust1}

base::table(sce_l$cluster_louvain)

umap1 <- umap_base(sce_l, color_by = "cluster_louvain")+
  ggplot2::scale_color_manual(values = col_clust_l)+
  ggplot2::ggtitle(base::paste(individual_curr, 
                               "fraction merge, original clusters"))
UMAP1 <- ggpubr::ggarrange(umap1, legend)
UMAP1
```

```{r umap_clust2}

base::table(sce_gg_cl$species_cluster_louvain)

umap2 <- umap_base(sce_gg_cl, color_by = "cluster_louvain")+
  ggplot2::scale_color_manual(values = col_clust_l)+
  ggplot2::ggtitle(base::paste(individual_curr,
                               "species-specific cells, original clusters"))
UMAP2 <- ggpubr::ggarrange(umap2, legend)
UMAP2

umap3 <- umap_base(sce_gg_cl, color_by = "species_cluster_louvain")+
  ggplot2::ggtitle(base::paste(individual_curr,
                               "species-specific cells, species clusters"))
UMAP3 <- ggpubr::ggarrange(umap3, legend_species)
UMAP3
```

```{r umap_clust3}
umap4 <- umap_base(sce_ls, color_by = "cluster_louvain")+
  ggplot2::ggtitle(base::paste(individual_curr, 
                               "species-specific UMAP, species clusters"))
UMAP4 <- ggpubr::ggarrange(umap4, legend_species_full)
UMAP4

# add original clusters to object with species-specific UMAP coordinates
sce_ls$cluster_orig <- vector(length = ncol(sce_ls))
sce_ls$cluster_orig[base::match(
  colnames(sce_gg_cl),
  colnames(sce_ls))] <- as.character(sce_gg_cl$cluster_louvain)


umap5 <- umap_base(sce_ls, color_by = "cluster_orig")+
  ggplot2::ggtitle(base::paste(
    individual_curr, "species-specific UMAP, original clusters"))+
  ggplot2::scale_color_manual(
    "cluster", 
    values = col_clust_l)
UMAP5 <- ggpubr::ggarrange(umap5, legend)
UMAP5
```

```{r prepare_ggtable_cluster}
ggtable_cl <- base::as.data.frame(table(ggdf_cl$cluster_louvain,
                                     ggdf_cl$species_cluster_louvain))

ggtable_cl$Var1 <- factor(ggtable_cl$Var1,
                          levels = levels(ggdf_cl$cluster_louvain))

ggtable_cl$Var2 <- factor(ggtable_cl$Var2, 
                          levels = base::paste0(levels(ggdf_cl$cluster_louvain),
                                             "_species"))
```

```{r plot1_clust, fig.height = 5, fig.width = 4}
ggplot2::ggplot(data = ggtable_cl,
                aes(axis1 = Var2,
                    axis2 = Var1, 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::ggtitle(base::paste("species-specific cells", individual_curr))+
  ggplot2::scale_fill_manual(values = col_clust_l)
```

```{r plot1_clust_rev, fig.height = 5, fig.width = 4}
ggplot2::ggplot(data = ggtable_cl,
                aes(axis1 = reorder(Var2, -Freq),
                    axis2 = reorder(Var1, -Freq), 
                    y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                            expand = c(0.15, 0.05))+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::ggtitle(base::paste("species-specific cells", individual_curr))+
  ggplot2::scale_fill_manual(values = col_clust_l)
```

Add annotation if possible.

```{r prepare_clust_anno}
ggtable_cl$ct <- vector(length = nrow(ggtable_cl))

ggtable_cl$ct <- unfactor(sce_ct$annotation_cluster[base::match(
  ggtable_cl$Var1,
  sce_ct$cluster_louvain)])

ggtable_cl$ct[is.na(ggtable_cl$ct)] <- "removed"

ggtable_cl$ct <- factor(ggtable_cl$ct,
                        levels = c(levels(sce_ct$annotation_cluster),
                                   "removed"))
```

```{r plot2_clust, fig.height = 5, fig.width = 4}
ggplot2::ggplot(data = ggtable_cl, aes(axis1 = Var2, axis2 = ct, y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                             expand = c(0.15, 0.05))+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::ggtitle(base::paste("species-specific", individual_curr))+
  ggplot2::scale_fill_manual(values = col_clust_l)
```

```{r plot2_clust_rev, fig.height = 5, fig.width = 4}
ggplot2::ggplot(data = ggtable_cl, aes(axis1 = reorder(Var2, -Freq), 
                                       axis2 = reorder(ct, -Freq),
                                       y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                             expand = c(0.15, 0.05))+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::ggtitle(base::paste("species-specific", individual_curr))+
  ggplot2::scale_fill_manual(values = col_clust_l)
```

## Only remaining cell types

```{r prepare_ggdf_ct}
# because some cell types were removed, the number of cells is smaller
sce_gg_ct <- sce_ct[,which(colnames(sce_ct) %in% colnames(sce_ls))]

sce_gg_ct$species_cluster_louvain <- vector(length = ncol(sce_gg_ct))
sce_gg_ct$species_cluster_louvain <- sce_ls$cluster_louvain[base::match(
  colnames(sce_gg_ct),
  colnames(sce_ls))]

ggdf_ct <- base::as.data.frame(SummarizedExperiment::colData(sce_gg_ct))

ggdf_ct$species_cluster_louvain <- base::paste0(ggdf_ct$species_cluster_louvain, 
                                                "_species")
```

```{r umap_ct_legend}

intermed_col_ct <- c(col_num[1:length(base::unique(sce_ct$annotation_cluster))],
                     "grey86")
intermed_col_ct
levels(sce_ct$annotation_cluster)

names(intermed_col_ct) <- c(levels(sce_ct$annotation_cluster), "removed")

# make legend
umap_l_ct <- umap_legend(sce_gg_ct, color_by = "annotation_cluster")+
  ggplot2::scale_color_manual("annotation_cluster", values = intermed_col_ct)
legend_ct <- ggpubr::get_legend(umap_l_ct)
```

```{r umap_ct1}

base::table(sce_ct$annotation_cluster)

umap1 <- umap_base(sce_ct, color_by = "annotation_cluster")+
  ggplot2::ggtitle(base::paste(individual_curr, 
                               "fraction merge, annotated cell types"))+
  ggplot2::scale_color_manual(values = intermed_col_ct)
UMAP1 <- ggpubr::ggarrange(umap1, legend_ct)
UMAP1
```

```{r umap_ct2}
base::table(sce_gg_ct$species_cluster_louvain)

umap2 <- umap_base(sce_gg_ct, color_by = "annotation_cluster")+
  ggplot2::ggtitle(base::paste(individual_curr,
                               "species-specific cells, annotated cell types"))+
  ggplot2::scale_color_manual(values = intermed_col_ct)
UMAP2 <- ggpubr::ggarrange(umap2, legend_ct)
UMAP2

umap3 <- umap_base(sce_gg_ct, color_by = "species_cluster_louvain")+
  ggplot2::ggtitle(base::paste(individual_curr,
                               "species-specific cells, species clusters"))
UMAP3 <- ggpubr::ggarrange(umap3, legend_species)
UMAP3
```

```{r umap_ct3}
umap4 <- umap_base(sce_ls, color_by = "cluster_louvain")+
  ggplot2::ggtitle(base::paste(individual_curr, 
                               "species-specific UMAP, species clusters"))
UMAP4 <- ggpubr::ggarrange(umap4, legend_species_full)
UMAP4

# add annotated cell types to object with species-specific UMAP coordinates
sce_ls$annotation_orig <- vector(length = ncol(sce_ls))

base::table(sce_ls$cluster_louvain[base::match(
  colnames(sce_gg_ct),
  colnames(sce_ls))])

base::table(sce_ls$annotation_orig[base::match(
  colnames(sce_gg_ct),
  colnames(sce_ls))])

sce_ls$annotation_orig[base::match(
  colnames(sce_gg_ct),
  colnames(sce_ls))] <- unfactor(sce_gg_ct$annotation_cluster)

sce_ls$annotation_orig[sce_ls$annotation_orig == FALSE] <- "removed"

base::table(sce_ls$annotation_orig)

umap5 <- umap_base(sce_ls, color_by = "annotation_orig")+
  ggplot2::ggtitle(base::paste(
    individual_curr, "species-specific UMAP, annotated cell types"))+
  ggplot2::scale_color_manual(values = intermed_col_ct)
UMAP5 <- ggpubr::ggarrange(umap5, legend_ct)
UMAP5

table(sce_ls$annotation_orig)
```

```{r prepare_ggtable_ct}
ggtable_ct <- base::as.data.frame(base::table(ggdf_ct$annotation_cluster,
                                           ggdf_ct$species_cluster_louvain))

ggtable_ct$Var1 <- factor(ggtable_ct$Var1,
                          levels = levels(ggdf_ct$annotation_cluster))
ggtable_ct$Var2 <- factor(ggtable_ct$Var2, 
                          levels = base::paste0(levels(ggdf_ct$cluster_louvain),
                                                "_species"))
```

```{r plot1_ct, fig.height = 5, fig.width = 4}
ggplot2::ggplot(data = ggtable_ct, aes(axis1 = Var2, axis2 = Var1, y = Freq))+
  ggalluvial::geom_alluvium(aes(fill = Var1))+
  ggalluvial::geom_stratum()+
  ggplot2::geom_text(stat = "stratum",
                     aes(label = after_stat(stratum)))+
  ggplot2::scale_x_discrete(limits = c("cluster_louvain", 
                                       "species_cluster_louvain"),
                             expand = c(0.15, 0.05))+
  ggplot2::theme_void()+
  ggplot2::theme(legend.position = "none")+
  ggplot2::ggtitle(base::paste("species-specific", individual_curr))+
  ggplot2::scale_fill_manual("annotated", values = intermed_col_ct)
```

```{r session_info}
utils::sessionInfo()
```
